### 2022-08-05 appendix/appendix_a.Rmd
Add files via upload

# Participant background information

```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
library(magrittr)
library(tidyverse)
library(readr)
library(kableExtra)
library("papaja")
library(stringr)
h1 = c("No", "Status", "Sex", "AOA", "Age", "LOE", "Deaf schools", "Home language", "Parents' hearing")

df_table_a <- read.csv("appendix_a_table.csv", sep = ";")
names(df_table_a) <- h1
apa_table(df_table_a,
          caption ="Participant demographic information based on self-report",
          note = "AOA = age of acquisition, LOE = level of education, HS = high school, AD = associate degree, MS = middle school, BA/BS = Bachelor of Arts/Science, TR = (Spoken) Turkish.",
          escape = FALSE,
          placement = "h",
          font_size = "footnotesize",
          landscape = FALSE
)
```

### 2022-08-05 appendix/appendix_a.tex
Add files via upload

\clearpage

\begin{appendix}
\hypertarget{participant-background-information}{%
\section{Participant background
information}\label{participant-background-information}}

\begin{table}[h]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:unnamed-chunk-12}Participant demographic information based on self-report}

\footnotesize{

\begin{tabular}{lllllllll}
\toprule
No & \multicolumn{1}{c}{Status} & \multicolumn{1}{c}{Sex} & \multicolumn{1}{c}{AOA} & \multicolumn{1}{c}{Age} & \multicolumn{1}{c}{LOE} & \multicolumn{1}{c}{Deaf schools} & \multicolumn{1}{c}{Home language} & \multicolumn{1}{c}{Parents' hearing}\\
\midrule
1 & Native & M & 0-3 & 33 & 12 & PS, MS, HS & TİD, home sign & Both deaf\\
2 & Native & F & 0-3 & 27 & 12 & PS, MS, HS & TİD, home sign & Both deaf\\
3 & Native & F & 0-3 & 27 & 12 & PS, MS, HS & Home sign & Both deaf\\
4 & Native & M & 0-3 & 29 & 12 & PS, MS, HS & TİD, home sign & Both deaf\\
5 & Native & M & 0-3 & 28 & 12 & PS, MS & TİD & Both deaf\\
6 & Native & M & 0-3 & 32 & 12 & PS, MS & TİD & F deaf, M hearing\\
7 & Native & M & 0-3 & 35 & 12 & PS, MS, HS & TİD & Both deaf\\
8 & Native & F & 0-3 & 23 & 12 & PS, MS, HS & TİD & Both deaf\\
9 & Native & F & 0-3 & 30 & 12 & PS, MS, HS & TİD & Both deaf\\
10 & Native & F & 0-3 & 25 & 12 & PS, MS & TİD & Both deaf\\
11 & Native & F & 0-3 & 24 & 14 & PS, MS, HS & TİD & Both deaf\\
12 & Native & M & 0-3 & 18 & 12 & PS, MS & TİD & Both deaf\\
13 & Native & F & 0-3 & 21 & 16 & PS, MS, HS & TİD & Both deaf\\
14 & Native & M & 0-3 & 19 & 16 & PS, MS, HS & TİD & Both deaf\\
15 & Native & F & 0-3 & 25 & 12 & PS, MS & TİD & Both deaf\\
16 & Late & F & 4-7 & 35 & 14 & PS, MS & TİD, home sign, TR & Both hearing\\
17 & Late & M & 4-7 & 36 & 12 & MS, HS & TR, home sign & Both hearing\\
18 & Late & M & 4-7 & 33 & 12 & PS, HS & TR & Both hearing\\
19 & Late & F & 4-7 & 33 & 12 & PS & TR, home sign & Both hearing\\
20 & Late & F & 4-7 & 31 & 14 & PS, MS & Home sign & Both hearing\\
21 & Late & M & 4-7 & 35 & 12 & PS, MS, HS & Home sign & Both hearing\\
22 & Late & F & 8-12 & 49 & 8 & PS, MS & Home sign, TR & Both hearing\\
23 & Late & M & 8-12 & 43 & 16 & PS, MS & TR & Both hearing\\
24 & Late & M & 8-12 & 28 & 16 & MS & TR & Both hearing\\
25 & Late & F & 8-12 & 24 & 12 & PS, MS, HS & TR, TİD & Both hearing\\
26 & Late & F & 8-12 & 24 & 8 & PS, MS & Home sign & Both hearing\\
27 & Late & M & 8-12 & 33 & 12 & PS, MS & Home sign & Both hearing\\
28 & Late & M & 13-17 & 50 & 8 & MS & TR, home sign & Both hearing\\
29 & Late & M & 13-17 & 31 & 16 & PS, MS, HS & TR & Both hearing\\
\bottomrule
\addlinespace
\end{tabular}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} AOA = age of acquisition, LOE = level of education, HS = high school, AD = associate degree, MS = middle school, BA/BS = Bachelor of Arts/Science, TR = (Spoken) Turkish.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}
\end{appendix}

### 2022-08-05 appendix/appendix_b.Rmd
Add files via upload

\setlength{\columnsep}{12pt}
\twocolumn
\linespread{1}

# Episodes in the stimuli.

\justifying
\small

**Narration 1** \vspace{-0.3cm}

1\. Tom chases Jerry, wearing a wooden plank \vspace{-0.3cm}

2\. Tom and Jerry enter a tube \vspace{-0.3cm}

3\. Tom falls down and Jerry runs away \vspace{-0.3cm}

4\. The tube gets thinner and Tom keeps chasing Jerry \vspace{-0.3cm}

5\. Jerry gets out and escapes \vspace{-0.3cm}

6\. Tom follows Jerry in the shape of a stick

**Narration 2** \vspace{-0.3cm}

1\. Tom walks past a female cat sitting on a couch \vspace{-0.3cm}

2\. Tom turns to and approaches the female cat \vspace{-0.3cm}

3\. The female cat blushes \vspace{-0.3cm}

4\. Tom offers a fish in aquarium to the female cat \vspace{-0.3cm}

5\. The female cat rejects the offer \vspace{-0.3cm}

6\. Tom offers a bird in cage to the female cat \vspace{-0.3cm}

7\. Episode 5 is repeated \vspace{-0.3cm}

8\. Tom leaves by walking in a flirtatious manner

**Narration 3** \vspace{-0.3cm}

1\. Jerry falls down and tries to walk up the stairs \vspace{-0.3cm}

2\. Tom pulls the carpet back \vspace{-0.3cm}

3\. Jerry struggles \vspace{-0.3cm}

4\. A piano falls down the stairs \vspace{-0.3cm}

5\. Jerry manages to escape \vspace{-0.3cm}

6\. Tom sees the piano and tries to run away \vspace{-0.3cm}

7\. The piano smashes into Tom and flattens him \vspace{-0.3cm}

8\. Tom falls on the ground

**Narration 4** \vspace{-0.3cm}

1\. Tom hides behind a cardboard house \vspace{-0.3cm}

2\. Jerry walks toward the house with a toy girl mouse \vspace{-0.3cm}

3\. The toy leaves Jerry who stops at the entrance \vspace{-0.3cm}

4\. The toy enters the house \vspace{-0.3cm}

5\. The toy gets smashed into pieces \vspace{-0.3cm}

6\. Jerry is surprised \vspace{-0.3cm}

7\. Tom hurts his stomach after swallowing the toy \vspace{-0.3cm}

8\. Tom looks at his broken teeth on a mirror \vspace{-0.3cm}

9\. Tom gets angry and throws down the mirror

**Narration 5** \vspace{-0.3cm}

1\. Tom fakes sleeping and suddenly gets up \vspace{-0.3cm}

2\. Tom pulls a rope inside a couch \vspace{-0.3cm}

3\. Jerry comes out of the couch holding a bar of soap \vspace{-0.3cm}

4\. Jerry washes Tom's mouth with the bar and flees \vspace{-0.3cm}

5\. Tom spits the soap out \vspace{-0.3cm}

6\. Jerry escapes to the nest with the rope \vspace{-0.3cm}

7\. Tom catches the rope and stars pulling it \vspace{-0.3cm}

8\. A mouse trap comes out and catches Tom's nose \vspace{-0.3cm}

9\. Tom yells and removes the trap

**Narration 6** \vspace{-0.3cm}

1\. Tom points a sword at Jerry and catches him \vspace{-0.3cm}

2\. Tom looks at the ill mouse \vspace{-0.3cm}

3\. Red spots appear on Jerry's face \vspace{-0.3cm}

4\. Tom gets scared and drops Jerry \vspace{-0.3cm}

5\. Tom rushes into bathroom \vspace{-0.3cm}

6\. Tom washes his hands \vspace{-0.3cm}

7\. Tom opens a drawer \vspace{-0.3cm}

8\. Tom swallows pills one after another \vspace{-0.3cm}

9\. Feeling ill, Tom looks out of the window \vspace{-0.3cm}

10\. Tom and Jerry look at each other

**Narration 7** \vspace{-0.3cm}

1\. Two male cats chase one another \vspace{-0.3cm}

2\. Tom approaches a female cat and kisses her cheek \vspace{-0.3cm}

3\. Tom hides away when the black cat returns \vspace{-0.3cm}

4\. Tom pokes his head behind the couch again \vspace{-0.3cm}

5\. Episode 3 is repeated \vspace{-0.3cm}

6\. The black cat sees the female cat and hugs her \vspace{-0.3cm}

7\. Tom appears and the black cat kisses Tom instead \vspace{-0.3cm}

8\. The two male cats move away from one another

**Narration 8** \vspace{-0.3cm}

1\. Jerry hits Tom's knee with a hammer \vspace{-0.3cm}

2\. Tom suddenly jumps up \vspace{-0.3cm}

3\. Tom yells and holds his knee \vspace{-0.3cm}

4\. Jerry runs towards Tom holding a thermometer \vspace{-0.3cm}

5\. Jerry sticks the thermometer into Tom's mouth \vspace{-0.3cm}

6\. Jerry lights a fire under the thermometer \vspace{-0.3cm}

7\. The red liquid starts to rise up \vspace{-0.3cm}

8\. The thermometer explodes

**Narration 9** \vspace{-0.3cm}

1\. Tom sits on Jerry who tries to get loose \vspace{-0.3cm}

2\. Jerry looks around and sees a hat ribbon\vspace{-0.3cm}

3\. Jerry finds a needle attached to the hat \vspace{-0.3cm}

4\. Jerry takes the needle and sticks Tom with it \vspace{-0.3cm}

5\. Tom turns red and jumps \vspace{-0.3cm}

6\. Jerry places the needle between two cushions \vspace{-0.3cm}

7\. Tom dodges the needle

**Narration 10** \vspace{-0.3cm}

1\. Wearing a stethoscope, Tom catches Jerry \vspace{-0.3cm}

2\. Jerry yells into the stethoscope \vspace{-0.3cm}

3\. Tom's head pops up \vspace{-0.3cm}

4\. Jerry runs and enter a mouse nest \vspace{-0.3cm}

5\. Tom inserts a gun into the nest \vspace{-0.3cm}

6\. The gun bends over toward Tom \vspace{-0.3cm}

7\. Tom fires the gun and his head explodes

\hspace{0.5cm}

### 2022-08-05 appendix/tables/edited_table.tex
Add files via upload

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  english,
  doc,mask]{apa6}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{ifxetex,ifluatex}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Age of Acquisition Effects on Reference Tracking in Turkish Sign Language (TİD)},
  pdflang={en-EN},
  pdfkeywords={Reference tracking, Turkish Sign Language (TİD), Age of acqusition, Signed narrative, Pragmatic competence},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
% Manuscript styling
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

% Table formatting
\usepackage{longtable}
\usepackage{lscape}
% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
\usepackage{multirow}		% Table styling
\usepackage{tabularx}		% Control Column width
\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
\usepackage{threeparttablex}            % Lets threeparttable work with longtable

% Create new environments so endfloat can handle them
% \newenvironment{ltable}
%   {\begin{landscape}\begin{center}\begin{threeparttable}}
%   {\end{threeparttable}\end{center}\end{landscape}}
\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

% Enables adjusting longtable caption width to table width
% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

% \setlength{\parindent}{0.5in}
% \setlength{\parskip}{0pt plus 0pt minus 0pt}

% Overwrite redefinition of paragraph and subparagraph by the default LaTeX template
% See https://github.com/crsh/papaja/issues/292
\makeatletter
\renewcommand{\paragraph}{\@startsection{paragraph}{4}{\parindent}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-1em}%
  {\normalfont\normalsize\bfseries\itshape\typesectitle}}

\renewcommand{\subparagraph}[1]{\@startsection{subparagraph}{5}{1em}%
  {0\baselineskip \@plus 0.2ex \@minus 0.2ex}%
  {-\z@\relax}%
  {\normalfont\normalsize\itshape\hspace{\parindent}{#1}\textit{\addperi}}{\relax}}
\makeatother

% \usepackage{etoolbox}
\makeatletter
\patchcmd{\HyOrg@maketitle}
  {\section{\normalfont\normalsize\abstractname}}
  {\section*{\normalfont\normalsize\abstractname}}
  {}{\typeout{Failed to patch abstract.}}
\patchcmd{\HyOrg@maketitle}
  {\section{\protect\normalfont{\@title}}}
  {\section*{\protect\normalfont{\@title}}}
  {}{\typeout{Failed to patch title.}}
\makeatother
\shorttitle{AOA Effects on Reference Tracking in TİD}
\keywords{Reference tracking, Turkish Sign Language (TİD), Age of acqusition, Signed narrative, Pragmatic competence\newline\indent Word count: X}
\usepackage{csquotes}
\usepackage{floatrow}
\floatsetup[table]{capposition=top}
\floatsetup[figure]{capposition=top}
\usepackage{times}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{langsci-gb4e}
\usepackage{multicol}
\setlength{\columnsep}{-3cm}
\usepackage{setspace}
\AtBeginEnvironment{tabular}{\singlespacing}
\AtBeginEnvironment{lltable}{\singlespacing}
\AtBeginEnvironment{tablenotes}{\doublespacing}
\captionsetup[table]{font={stretch=1.5}}
\captionsetup[figure]{font={stretch=1.5}}
\ifxetex
  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\else
  \usepackage[main=english]{babel}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\fi
\ifluatex
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

\title{Age of Acquisition Effects on Reference Tracking in Turkish Sign Language (TİD)}
\author{Onur Keleş\textsuperscript{1}, Furkan Atmaca\textsuperscript{2}, \& Kadir Gökgöz\textsuperscript{1}}
\date{}

\authornote{

This work was supported by Boğaziçi Üniversitesi Bilimsel Araştırma Projeleri {[}Boğaziçi University Research Fund{]}, {[}14458{]}. We thank all of our 31 participants.

The authors made the following contributions. Onur Keleş: Conceptualization, Data Curation, Investigation, Methodology, Writing - Original Draft Preparation, Writing - Review \& Editing; Furkan Atmaca: Formal Analysis, Methodology, Software, Validation, Visualization, Writing - Review \& Editing; Kadir Gökgöz: Funding Acquisition, Project Administration, Supervision, Writing - Review \& Editing.

Correspondence concerning this article should be addressed to Onur Keleş, Bebek, 34342 Beşiktaş/İstanbul. E-mail: \href{mailto:keles.onur@hotmail.com}{\nolinkurl{keles.onur@hotmail.com}}

}

\affiliation{\vspace{0.5cm}\textsuperscript{1} Boğaziçi University\\\textsuperscript{2} Leipzig University}

\abstract{
We investigated the reference tracking strategies among native and late adult deaf signers of Turkish Sign Language (TİD). Given the previous research that suggests over-redundancy in reference tracking and poorer morphosyntactic competence for late signers, we expected that late signers use fewer instances of null marking and mixing of perspectives in their tracking than the native group. Using a story-telling paradigm, we tested 15 native and 16 late signers. Participants were shown 10 short wordless video clips from a cartoon, and asked to retell them. The overall results indicated that the observed distribution of referring expressions was overall consistent with the suggested models of reference control within the topicality and saliency paradigm for both spoken and signed languages. However, native acquisition increased the use of null marking, with no significant effect in nominal production. We also found that native signers used mixed perspectives in narrative discourse more than late signers. Given these findings, we suggest that language deprivation might negatively affect late signers' pragmatic competence (e.g., sensitivity to economy of form) and their use of complex morphosyntactic structures in reference tracking, which are known to be age-sensitive.
}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Narratives are extensive discourses that construct an important part of human communication. In order to create a systematic and coherent organization of experiences, narrators must choose appropriate linguistic forms to refer to entities. Certain forms such as nominal expressions are commonly used for the first mentions of a referent in a discourse, and other forms such as pronominals and null marking may track previously mentioned entities which could either be maintained across multiple clauses or re-introduced back into the discourse. The selection of these referring expressions and anaphoric tools (henceforth REATs) is contingent on cognitive abilities, such that the encoder must rely on the activation of the working memory to keep in mind previously referred expressions and long strings of information (Bamberg, 1997; Morgan, 2005). Pragmatic abilities too play a pivotal role in the tracking of referents: The form with the least amount of redundant information must be selected based on the accessibility levels of the referent (Ahn, 2019; Ariel, 1990). This denotes that acquiring adult-like narrative competence is not a trivial job for children. Speakers and signers of a native language start to refer to entities and use complex anaphoric or cataphoric expressions appropriately following age 10 (Aksu-Koç \& Nicolopoulou, 2015; Hickmann, Hendriks, Roland, \& Liang, 1996; Rathmann, Mann, \& Morgan, 2007). Similarly, second language learners find it difficult to produce a coherent and structured narrative where it is important to use REATs appropriately in line with the cognitive statuses of the referents. This finding comes both from research examining L2 acquisition of spoken narratives (Gundel, Hedberg, \& Zacharski, 1993; Hickmann, Hendriks, Roland, \& Liang, 1996; Williams, 1988) and signed narratives (Bel, Ortells, \& Morgan, 2015; Frederiksen \& Mayberry, 2019) which are conveyed in different communication modalities. Yet, sign languages and deaf populations have remained mostly understudied with respect to the acquisition of narrative and pragmatic competence.

Sign languages are the natural languages of the visual-spatial modality prominently used among deaf populations. The cognitive and linguistic development of native signers (i.e., those who start acquiring a signed language from one or more deaf caregivers following birth) parallels that of native speakers (Chamberlain, Morford, \& Mayberry, 1999; Petitto, 1997). However, reports indicate that at least over 90\% of all deaf individuals are born into hearing families who do not know how to sign (Weaver \& Starner, 2011; Woll, 2013). This means that the auditory modality used in the household is not accessible to many deaf children. That is why the age of acquisition of a first functional language varies greatly among deaf children, with many of whom being deprived of their first languages typically until after age 7 when they start formal education in a deaf school. Although the lack of a systematic and comprehensible form of communication in the early years of life is compensated with an internally consistent primitive gesture-based system called home sign (Goldin-Meadow, 2005), this neither fully resembles natural languages nor reinstates first language acquisition. Consequently, language deprivation has been found to have negative effects on certain cognitive and linguistic measures among late signing children. To name a few, delayed exposure to a first functional language is reported to negatively affect deaf children's morphosyntactic abilities (Boudreault \& Mayberry, 2006; Karadöller, Sümer, \& Özyürek, 2017; Newport, 1990), phonological processing and phonemic awareness (Corina, Hafer, \& Welch, 2014; MacSweeney, Waters, Brammer, Woll, \& Goswami, 2008), vocabulary knowledge (Anderson, 2005; Lu, Jones, \& Morgan, 2016), and narrative abilities (Becker, 2009).

Given that there is a paucity of studies that examine deaf adult signers with delayed exposure to a sign language, it is not very clear whether the narrative abilities of late signers differ from those of native signers. To expand on the previous literature, the present study aims to investigate the reference tracking strategies of native and late adult signers in Turkish Sign Language (TİD) narratives through a story-telling paradigm. We explore the nature of REATs selection and perspective taking as a possible indicator of pragmatic competence between our acquisition groups. The paper is organized as follows: Section 1 (i) gives an overview of REATs selection and perspective taking in signed languages, (ii) describes hitherto attested REATs in TİD, and (iii) reviews previous findings on the effect of delayed first language acquisition on signers' narrative competence. Section 2 and 3 present the methodology and results of this paper, respectively. Finally, Section 4 provides a general discussion of the reference tracking strategies among native and late signers in relation to the effects of delayed first language acquisition on pragmatic competence.

\hypertarget{reference-tracking-strategies-in-natural-languages}{%
\subsection{Reference Tracking Strategies in Natural Languages}\label{reference-tracking-strategies-in-natural-languages}}

\hypertarget{reats-selection}{%
\subsubsection{REATs Selection}\label{reats-selection}}

Natural languages can be systematically used to keep track of previous referents in the discourse and mark the changing cognitive statuses of the REATs to create a more coherent narrative. Among the theoretical frameworks that aim to explain that different linguistic forms are selected in accordance with the addressee's mental representations of the referents in the discourse is the Givenness Hierarchy (Gundel, Hedberg, \& Zacharski, 1993). According to this model, given or familiar information is expressed in more attenuated forms such as pronominal use and null marking since the mental status of the expressed referent is already at the center of attention, as indicated by the present contextual information (Chafe, 1976). The Givenness Hierarchy is akin to other previously proposed models of reference control within the topicality and saliency paradigm such as the Accessibility Hierarchy (Ariel, 1990), Topic Continuity (Givon, 1983), and Familiarity Scale (Prince, 1981). Gundel, Hedberg, and Zacharski (1993) argued that the distribution of forms is not random and abides by the Gricean Maxim of Quantity (Grice, 1975) in that the encoder must make the most informative and succinct choice possible. On this ground, for highly accessible referents, linguistic forms with the least phonetic content are selected for the least amount of processing effort based on the economy principle.

It is also possible to categorize these referring expressions into three according to their discourse statuses (Gullberg, 2006). A referent can be introduced, maintained, that is, continued across two or more clauses, or reintroduced into the discourse following a topic shift. Based on the principles of reference accessibility and relevance, the introduction and reintroduction of a referent are often accomplished by full noun phrases which can be commonly modified by an adjective or demonstrative in spoken languages and also by optional markers such as constructed action, classifier constructions, or index points in sign languages. Maintenance or topic continuity, on the other hand, is typically marked with null marking, pronominal forms, and classifiers across signed languages since the continued referent is still highly accessible by the addressee.

\hypertarget{perspective-taking-in-signed-narratives}{%
\subsubsection{Perspective taking in signed narratives}\label{perspective-taking-in-signed-narratives}}

Sign language research has identified a split, albeit using different terminology, between what is commonly referred to as the character perspective and the observer perspective to explain some of the attested structures in signed narratives. The character perspective (also referred to as the large-scale use of the space) depends on the signer's mental conceptualization of the narrated event. In other words, the signers engage with the space in front of themselves as if they are interacting with the objects and characters in the story. This is often achieved with the signer taking on the role of a referent in the narrative and performing the actions by employing a linguistic device called constructed action (Metzger, 1995) or referential shift (Emmorey, Corina, \& Bellugi, 1995) that are marked by certain non-manual cues such as the facial expression and the breaking of the eye contact. This is to say, the narrator's head, face, torso, arms, and as an optional marker, hands become one with those of the associated referent(s) (Smith \& Cormier, 2014). Constructed action does not necessarily entail a body shift to a locus in the signing space. Instead, the signer mentally rotates the space and faces the addressee in most cases.In the observer perspective or the small-scale use of the signing space, referents are fully or partially represented not by the signer's body but by the manually articulated classifier, verbal or nominal constructions signed in front of the body.

Signers typically switch between these perspectives multiple times while narrating. While the observer perspective is commonly used to describe an intransitive event and to show change of location (e.g., path movement), the character perspective is used to inform the addressee about transitive events and to depict what the constructed character is doing. Switching between perspectives can be done sequentially: Signers can talk about different characters in a narrative by providing locative information and describe intransitive events from the observer perspective through different verbal and classifier constructions. The character perspective can immediately follow an observer perspective representation and describe the transitive or manual activities carried out by the character(s).

Capitalizing on the visual-spatial modality, signers can also engage in simultaneous mixing of perspectives. Morgan (2005) noted that these two perspectives, shifted referential space versus fixed referential space in his terminology, can be mixed for the simultaneous articulation of multiple referents. In this representation, while the signer constructs the actions of a referent via constructed action, the hands or part of the signer's body can be partitioned (see Dudis, 2004) and leave the mentally rotated space. The partitioned limbs (e.g., selected fingers) can then function as a whole entity classifier for another referent. The signer's construction or enactment of the animate referent can interact with the objects situated in the observer perspective through body part classifiers or verbal constructions. This way, the signer provides spatial information from both the observer and character perspective. This is to ensure the achievement of spatial coherence and efficiency of information. Simultaneous encoding of spatial information is then motivated by the economy principle of language.

\hypertarget{tracking-referents-in-turkish-sign-language-tid}{%
\subsection{Tracking Referents in Turkish Sign Language (TİD)}\label{tracking-referents-in-turkish-sign-language-tid}}

The linguistic inventory for discourse referents in TİD includes nominal constructions (bare nouns, fingerspelled nouns, and modified nouns) for low accessibility referents. Null marking (e.g., plain verb, agreement verb, and constructed action), classifier predicates (e.g., whole entity classifiers, body part classifiers, handling classifiers, and extension classifiers), and overt pronouns (pronominal index or IX) are selected for high accessibility referents. This interim section describes the REAT types attested in TİD under four main headings.

\begin{multicols}{2}
\begin{exe}
\ex
\begin{xlist}
\ex Full noun \\\glll
\includegraphics[width=2.5cm]{pictures/1a_1.png} \includegraphics[width=2.5cm]{pictures/1a_2.png} \\ \textsc{mouse} \textsc{cat} \\ \\ 
\ex Noun modified by adjective \\\glll
\includegraphics[width=2.5cm]{pictures/1c_1.png} \includegraphics[width=2.5cm]{pictures/1c_2.png} \\ \textsc{mouse} \textsc{sick} \\ \\ 
\ex Fingerspelled noun\\\glll 
\includegraphics[width=2.5cm]{pictures/1b_1.png} \includegraphics[width=2.5cm]{pictures/1b_2.png} \includegraphics[width=2.5cm]{pictures/1b_3.png}\\ \textsc{t-} \textsc{-o-} \textsc{-m} \\ \\
\ex Noun modified by index\\\glll 
\includegraphics[width=2.5cm]{pictures/1d_1.png} \includegraphics[width=2.5cm]{pictures/1d_2.png}\\ \textsc{ix} \textsc{cat} \\ \\
\end{xlist}
\end{exe}
\end{multicols}
\vspace{-1.25cm}
\newpage
\ea 
\ea Pronominal \\\glll
\includegraphics[width=2.5cm]{pictures/2_1.png} \includegraphics[width=2.5cm]{pictures/2_2.png} \\ \textsc{ix} \textsc{sick} \\ \\
\z \z
\vspace{-0.75cm}
\begin{multicols}{4}
\begin{exe}
\ex
\begin{xlist}
\ex Whole entity \\\glll 
\includegraphics[width=2.25cm]{pictures/3a.png} \\ {\textsc{dh: mouse}} \\ {\textsc{ndh: tunnel}}\\ \\
\ex Body part \\\glll
\includegraphics[width=2.25cm]{pictures/3b.png} \\ \textsc{walk} \\ \\ 
\ex Handling \\\glll 
\includegraphics[width=2.25cm]{pictures/3c.png} \\ \textsc{hold} \\ \\
\ex Extension \\\glll 
\includegraphics[width=2.25cm]{pictures/3d_1.png} \includegraphics[width=2.25cm]{pictures/3d_2.png}\\ \textsc{wood} \textsc{plank}\\ \\
\end{xlist}
\end{exe}
\end{multicols}

\begin{multicols}{3}
\begin{exe}
\ex
\begin{xlist}
\ex Constructed action \\\glll 
\includegraphics[width=2.5cm]{pictures/4c.png} \\ \textsc{struggle} \\ \\

\ex Agreement verb  \\\glll 
\includegraphics[width=2.5cm]{pictures/4b.png} \\ {\textsc{$_{3a}$kiss$_{3b}$}} \\ \\

\ex Plain verb  \\\glll
\includegraphics[width=2.5cm]{pictures/4a_1.png} \includegraphics[width=2.5cm]{pictures/4a_2.png}  \\ \textsc{restroom} \textsc{enter}\\ \\
\end{xlist}
\end{exe}
\end{multicols}

\hypertarget{nominal}{%
\subsubsection{\texorpdfstring{Nominal (\textsc{nom})}{Nominal ()}}\label{nominal}}

In TİD, it is possible to use bare noun phrases as in (1a) and fingerspelled nouns as in (1b) to introduce or reintroduce referents into the discourse. The first mention of a referent can also be tracked with noun phrases modified by adjectives as in (1c) which either follow or precede a nominal (Özsoy \& Nuhbalaoğlu, 2014). Alternatively, the noun phrase can retain its basic meaning and the modifier can help determine its position on the accessibility hierarchy by expressing the definiteness of the nominal. Definiteness in TİD can be expressed with a prenominal or postnominal \textsc{ix} functioning as a determiner or a demonstrative as in (1d).

\hypertarget{pronominal}{%
\subsubsection{\texorpdfstring{Pronominal \textsc{ix} (\textsc{pro})}{Pronominal  ()}}\label{pronominal}}

As discussed previously, the \textsc{ix} sign (extended index finger) has various functions in signed languages, one of which is the \textsc{pro} use which refers to established nominals by pointing to a previously assigned abstract or physical locus in the signing space when reintroducing a topic in discourse (Emmorey, 1996). That is to say, pronominal \textsc{ix} is licensed only if the referent is physically present in the space or the non-present antecedent is already associated with a locus. Once such a spatial relation is set up by the signer, it is possible to bring those referents back into the discourse by pointing to their location. The same \textsc{ix} function has also been identified for TİD (Sevinç, 2006). The example in (2) illustrates a spatially-anchored pronominal reference for topic reintroduction. These signs are thus rarely used for the first mentions of an entity (Perniss \& Özyürek, 2015). For the maintenance condition where the referent is highly accessible, some studies reported a relatively few instances of pronominal \textsc{ix} as signers favored other grammatical tools for reference such as null marking or classifier constructions (Cormier, Smith, \& Zwets, 2013; Frederiksen \& Mayberry, 2016, 2019). However, recent work pointed out what is referred to as the referentially underspecified pronominal \textsc{ix} which does not require the non-present antecedent to be spatially anchored to a previously established locus. Although other interpretations consider pointing signs without an noun phrases as examples of demonstratives (Koulidobrova \& Lillo-Martin, 2016), for the purposes of the present study, we will refer to those constructions as pronominal \textsc{ix} and particularly examine their non-deictic or personal pronoun realizations as a possible discourse strategy to mainly mark topic shift among native and late signers of TİD.

\hypertarget{classifier}{%
\subsubsection{\texorpdfstring{Classifier (\textsc{cl})}{Classifier ()}}\label{classifier}}

\textsc{cl} predicates are highly complex morphological constructions which are very common in sign languages (Supalla, 1986). These are manually signed iconic representations of animate or inanimate objects, and they embody several properties pertaining to the nominal referent including its location in the signing space, movement, path, size, and shape (Talmy, 2003). As in most sign languages hitherto studied, \textsc{cl} constructions are also common and found in TİD (see Dikyuva, Makaroğlu, \& Arık, 2015 for a review). Following the classification of Engberg-Pedersen (1993), we distinguish between four types of \textsc{ix} constructions. Whole entity classifiers (\textsc{wecl}) represent the entirety of any referent whereas body part classifiers (\textsc{bpcl}) only reflect a specific part of an animate entity (e.g., limbs, head, tongue, teeth, and so forth). Handling classifiers (\textsc{hcl}) occur with transitive verbs and simply depict how an object is handled or manipulated by the agent. On this ground, \textsc{hcl} depicts the manner of the referring action while the former two constructions encode the location of the entity as well as its motion in space. Extension classifiers (\textsc{extcl}), on the other hand, denote the size and shape of an entity by following a trajectory to trace its shape. We exemplify these constructions in (3). As discussed before, \textsc{hcl} is transitive in nature and depicts the handling action performed by an animate agent. This is why most research tend to consider \textsc{hcl} as an extension to constructed action and group them with the character perspective where the signer's body aligns with the referent (Cormier, Smith, \& Zwets, 2013; Dudis, 2004; Özyürek \& Perniss, 2010). However, the other \textsc{ix} constructions co-occur with the observer perspective. That is to say, the signer does not mentally rotate the space nor employ constructed action but instead narrates the event from the observer perspective by fully or partly stating the referent via different linguistic forms. As expected, these constructions are argued to be higher on the accessibility hierarchy than \textsc{nom} (\textsc{wecl} more so than \textsc{bpcl} due to its less transparent and iconic nature). This is because they take on certain semantic properties of the referent, but they are reported to rank lower than \textsc{pro} pointing signs which are less explicit in form (Morgan, 2005). This contrasts with the finding that \textsc{pro} may be dispreferred in favor of \textsc{cl} constructions (Cormier, Smith, \& Zwets, 2013; Frederiksen \& Mayberry, 2016) contra Bel, Ortells, and Morgan (2015); Morgan (1999) who observed an abundance of \textsc{pro} within their participants' signed discourse. Considering the Givenness Hierarchy (Givon, 1983) that suggests more attenuated content for highly accessible referents, we expect \textsc{pro} to be higher on the hierarchy than \textsc{cl} in the present article. Nevertheless, \textsc{cl} constructions are commonly used by signers to maintain previous referring expressions although some studies reported a sporadic use of \textsc{cl} for referent introduction and reintroduction (e.g., Morgan, 2005).

\hypertarget{null-marking}{%
\subsubsection{Null Marking}\label{null-marking}}

Parallel to spoken languages, many sign languages have been reported to allow null subjects and/or objects and make use of null marking (\textsc{NULL}) when referring to highly accessible referents which are still active in the mental representation of the addressee. In signed discourse, this is achieved by employing verbal constructions when the event is narrated via the observer perspective. To identify the predicate types in ASL, Padden (1986) suggested a triadic categorization of verbs in sign languages: plain verbs, agreement verbs, and spatial verbs. According to this classification, plain verbs (\textsc{plain}) (4a) do not agree with the subject or the object. This means that they are not inflected for number or person. This contrasts with agreeing verbs (\textsc{agr}) (4b) the starting or the ending point of which can be modified to specific loci for person and number agreement while spatial verbs (as in \textsc{ix book put3a} `I placed the book there') take locational affixes but like plain verbs, they are not inflected for number or person. TİD too licenses \textsc{null} arguments. Following this classification, Sevinç (2006) argued that although it is common in TİD to drop arguments, there were certain limitations on the \textsc{null} marking of the arguments for \textsc{agr} and \textsc{plain} verbs. It is also possible and common for signers to use constructed action (\textsc{ca}) as in (4c) and take on the role of a character in the narrative while omitting arguments. This means that they do not assign locations to previously established referents but instead designate their own body as the specified locus. This change in the spatial scale is often marked with non-manual markers, with a break in eye gaze and the facial expression being the most informative cues (Engberg-Pedersen, 1993).

\hypertarget{mixing-of-perspectives}{%
\subsubsection{\texorpdfstring{\textbf{Mixing of perspectives}}{Mixing of perspectives}}\label{mixing-of-perspectives}}

As described in previous sections, it is also possible to mix both perspectives to bring two or more referents simultaneously into the discourse (Dudis, 2004; Janzen, 2004; Liddell, 2003). The fusion of character perspective and observer perspective is attested in TİD (Perniss \& Özyürek, 2015). In this mix, the signer's face and torso are mapped onto the narrated character with \textsc{ca} and the signer's hands or arms are partitioned off from the character perspective to sign verbal or \textsc{cl} predicates. These manual constructions are viewed externally by the signer from the observer perspective. In the example (4a), the signer takes on the role of Tom, an animate character in the narrative, holding a gun. The signer's right-hand leaves the mentally rotated space and starts to depict the gun being bent over toward Tom (the signer's head) by constructing an agreeing whole entity classifier. Two referents with varying degrees of cognitive statuses are then tracked simultaneously via two different REATs. Likewise, the signer in (4b) first introduces the female cat in the narrative by lexically identifying it as the syntactic object. The agentive subject (Tom), however, is highly accessible to the addressee. For this reason, the signer omits the subject argument and only signs the predicate \textsc{see} which agrees with unspecified loci. The second predicate \textsc{KISS} is represented with \textsc{bpcl} the endpoint of which this time aligns with the location of the signer who takes on the role of the previously established referent by using the character perspective. In the following clause, the event shifts from the perspective of the female cat to Tom. Here, the signer depicts Tom's hiding behind a couch and his left hand is partitioned off to represent a third referent (the black cat) with a \textsc{WECL}. In the second mix, Tom secretly watches the black cat running around. By employing these morphologically complex constructions that result in the mixture or blend of the two perspectives, signers can make the most of the visual modality for the simultaneous tracking of multiple referents with varying accessibility levels. Prioritizing informativeness and efficiency of information, this observation is in line with the presuppositions of the Gricean Maxim of Quantity (Grice, 1975) and the Relevance Theory (Sperber \& Wilson, 1986).

\newpage
\begin{exe}
\ex 
\begin{xlist}
\ex Perspective mixing with CA and WECL \\\glll
\includegraphics[width=2.5cm]{pictures/5a_1.png} \includegraphics[width=2.5cm]{pictures/5a_2.png} \includegraphics[width=2.5cm]{pictures/5a_3.png}\\ {\textsc{cp: hold.gun}} \\ \textsc{op:} {\textsc{ }} {\textsc{$_{3a}$gun.bend$_1$}} \\
\glt `The gun is bending toward Tom.'
\ex Perspective mixing across two sentences \\\glll 
\includegraphics[width=2.5cm]{pictures/5b_1.png} \includegraphics[width=2.5cm]{pictures/5b_3.png} \includegraphics[width=2.5cm]{pictures/5b_2.png} \includegraphics[width=2.5cm]{pictures/5b_4.png}\\ \textsc{cp:} {\textsc{}} {\textsc{being.kissed}} {\textsc{hide}} \\ {\textsc{op: girl}} {\textsc{$_{3a}$see$_{3b}$}} {\textsc{$_{3b}$kiss$_1$}} {\textsc{$_{3a}$run$_{3b}$}} \\
\glt `Tom sees and kisses the female cat. Then the gray cat comes running while Tom hides behind the couch.'
\end{xlist}
\end{exe}

\hypertarget{learning-how-to-track-referents-in-signed-languages}{%
\subsection{Learning How to Track Referents in Signed Languages}\label{learning-how-to-track-referents-in-signed-languages}}

Considering the work on sign language discourse that indicates native sign language development of deaf children parallels that of hearing children, all else being equal, it is not surprising that the pragmatic competence is also achieved similarly, albeit with different milestones (Morgan, 2000). Deaf children of deaf parents have been observed to lag behind hearing age-matched peers at 4 years of age in terms of the organization of a narrative, but the general understanding is that they catch up within two years (Vercaingne-Ménard, Godard, \& Labelle, 2001). In both communication modalities, children start to produce more complex and coherent stories with appropriate and informative linguistic forms to indicate anaphoric or cataphoric references after age 10 (Aksu-Koç \& Nicolopoulou, 2015; Hickmann, Hendriks, Roland, \& Liang, 1996; Rathmann, Mann, \& Morgan, 2007). Prior to this age, both signing and speaking children fail to refer to entities consistently since their pragmatic abilities are yet to develop. The manipulation of linguistic forms for discourse functions to signal topic continuance or shift is difficult to master until the beginning of formal education. These findings are also comparable across different sign languages including British Sign Language (BSL) (Morgan, 2005), American Sign Language (ASL) (Loew, 1984), German Sign Language (Deutsche Gebärdensprache - DGS) (Becker, 2009), and Canadian Sign Language (LSQ) (Vercaingne-Ménard, Godard, \& Labelle, 2001), pointing out to the fact that the mastery of reference tracking in narratives is one of the most protracted achievements in the acquisition of signed discourse.

Given the preliminary evidence that topic continuity and salience marking is also difficult for second language learners regardless of modality Hickmann \& Hendriks (1999), it has been suggested under the Interface Hypothesis that the syntax-discourse interface is vulnerable and especially challenging to acquire (Sorace, 2011). For this reason, it is thus one of the latest components of language to be learned. Similarly, Sorace and Serratrice (2009) discuss that this interface entails much more than morphosyntactic knowledge but also includes cognitive and pragmatic resources that help provide discourse coherence. They propose that the second language discourse is over-redundant. This means that second language learners produce more overt pronouns or fuller forms of language for highly accessible referents possibly as a compensatory strategy in an attempt to reduce the increasing load on cognitive processing.

Previously cited studies indicated that acquisition of reference tracking and thereby the mastery of perspectives in discourse production is challenging for deaf children of deaf parents (i.e., native signers) and even more so for second language learners. Still, it is unclear whether linguistically deprived signers (i.e., deaf signers with delayed exposure to a sign language) also display redundancy in their reference tracking. In response to this question, Becker (2009) examined DGS signing deaf native and late children's narrative abilities. She reported that late signing children generated more explicit forms such as \textsc{nom} when referring to the characters in the story compared to native learners. Her findings also revealed that late learners failed to establish unambiguous references and hence were less successful in the use of the signing space. Becker (2009) concluded with the claim that first language deprivation results in an inadequate representation of the addressee and less efficient use of morphosyntax. A similar line of research came from Smith and Cormier (2014). Likewise, they explored the use of the two perspectives employed with or without \textsc{ca} among BSL signing deaf native and late children. The results indicated that only native signers employed a non-prototypical alignment of perspectives to increase efficiency. In the study, the native signers used the character perspective without \textsc{ca}, with the opposite pattern observed for late learners. The authors reasoned that the former had been a more complex instance of reference tracking and the difference might stem from the fact that late signers have hearing parents and thus engage in less joint attention which translates into fewer opportunities to learn appropriate eye gaze patterns. Comparably, Cormier, Smith, and Zwets (2013) investigated the use of constructed action in BSL among deaf native and late adult signers and found that late signers were redundant in topic continuity. In other words, they produced more explicit forms like \textsc{nom}, and fewer instances of \textsc{null} when maintaining previous referents while Ferrara and Johnston (2012) failed to find such a preference for less phonetic content for highly accessible referents by native signers. Most relevantly to the present study, Gür (2018) replicated some of these findings for TİD in her work on the age of sign language acquisition effects on narrative abilities. She found that signers with early acquisition engaged in more correct instances of role shift while narrating. Given the limited amount of research involving late adult signers, the effects of delayed exposure to language on later discourse competence have been understudied. As discussed in the first section, the morphosyntactic competence of late signers falls behind that of native learners (e.g., Boudreault \& Mayberry, 2006; Newport, 1990) and these differences persist into adulthood. As a result, this discrepancy may continue to have an inhibiting effect on the syntax-discourse interface late in adulthood.

\hypertarget{present-study}{%
\subsection{Present Study}\label{present-study}}

In an attempt at answering this question, the present study aims to examine the reference tracking strategies among native and late deaf adult signers of TİD in relation to REATs selection and perspective taking. Based on previous research, we hypothesize that late signers will become more over-redundant in their reference tracking compared to native signers. This would translate to a preference by late signers for \textsc{nom} over \textsc{null} especially in the maintenance and re-introduction settings. In contrast, we would expect that native signers use more \textsc{null} markers in their tracking based on the economy principle. As for the perspective taking, we hypothesize that late learners will employ fewer instances of perspective mixing to refer to multiple references because of less developed morphosyntactic and pragmatic abilities in comparison to native signers. In contrast, we predict that native signers will use the signing space more economically and use more zero anaphora along with better use of the spatial scale.

\hypertarget{method}{%
\section{Method}\label{method}}

\hypertarget{participants}{%
\subsection{Participants}\label{participants}}

We tested 29 Deaf adult signers, 15 with exposure to TİD from birth (i.e., native signers), 14 with exposure between ages 3 and 17 (i.e., late signers) (MAOA = 7.7 years, SD = 3.3 years). 26 participants were pre-lingually deaf, that is they were either born deaf or they had acquired deafness before the age of 3. The remaining 3 participants became deaf between the ages of 4-7. Participants' level of education (LOE) (MLOE = 12.3 years, SD = 2.0 years) and the average year of language use following the start of exposure to TİD did not differ much between the acquisition groups (see Appendix for more detailed demographic information). All participating native signers were born into deaf families and started acquiring TİD from at least one of their caregivers after birth. The parents of 14 participating native signers were deaf. One native signer had a deaf father and a hearing mother. Comparably, late learners were all born into hearing families and started learning TİD later in life after age 3, usually with the start of formal education. The age of acquisition for late signers ranged between years 3-17 (MAOA = 8.70 years, SD = 3.52 years). The parents of all participating late signers were hearing, and the reported home languages spoken in childhood consisted mainly of Turkish and Home Sign. All participants indicated that their preferred language of communication was TİD which they used for most contexts. On a 1 (poor) to 5 (proficient) point scale, participants' self-reported proficiency in TİD was on average 4.97 points.

\hypertarget{task-procedure}{%
\subsection{\texorpdfstring{\textbf{Task Procedure}}{Task Procedure}}\label{task-procedure}}

We used a story-telling paradigm to examine how native and late signers of TİD track references in signed discourse. Participants were presented with 10 short extracts from a cartoon and asked to retell them. The stimulus stories were extracted from the cartoon Tom and Jerry. The duration of clips ranged from 17 seconds to 28 seconds. Each clip had six to ten events which consisted of one or multiple (up to three) recurring animate characters for which the present context required the introduction, maintenance, and re-introduction of referents.

Participants sat in front of a laptop computer which had the stimulus stories. The instructions were all given by a deaf research assistant who was present in all sessions and ensured that the task requirements were clear. The informed consent form was translated into TİD and explained to the participants by the same assistant before data recording. Participants were recorded by a digital camera, and they narrated the content of each clip to an additional experimenter who sat across the participants. The order of clips was randomized for each participant.

\hypertarget{coding}{%
\subsection{Coding}\label{coding}}

We coded the discourse status (Introduction, Maintenance, Re-introduction), REAT type (Nominal, Pronominal, Classifier, Null), and perspectives (Character, Observer, Mixed) (see Table 2 for the observed instances), identifying 4523 reference markings in total. We used ELAN (Sloetjes \& Wittenburg, 2008) for data annotation. First Author annotated all the narratives. We first identified the clause boundaries in each sentence indicated by the presence of predicates and informative non-manual markers (e.g., head nod). On one tier, we had the grammatical tools for a single reference with the subcategories Nominal, Pronominal, Zero Anaphora, and Classifier. Following the classification of Frederiksen and Mayberry (2019), we considered all referents mentioned for the first time as introduced regardless of their syntactic function in the sentence. Only subjects were coded for maintenance and re-introduction. We coded subject referents as maintained only if they had been signed in the previous clause. To this end, the referents which had been discontinued across one or more sentences were all regarded reintroduced into the discourse. We classified all handling classifiers which depicted the action of a referent under constructed action in line with previous interpretations (Cormier, Smith, \& Zwets, 2013).

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{reats-selection-1}{%
\subsection{REATs Selection}\label{reats-selection-1}}

We collected 4168 data points as a result of the experiment. Each data point is a referent tracked by the participant. Each of the referents is classified in terms of its discourse function (Introduction, Maintenance, Re-introduction), its grammatical function (Classifier, Nominal, Pronominal, Null), and the group of the participant (Native, Late). Figure \ref{fig:count-plot} shows the numerical distribution of the responses according to the aforementioned variables. The use of nominal or null is more numerous than the classifier and pronominal reference tracking. There is a noticeable difference between participant groups in the use of null marking for maintaining a referent. Native signers of TİD have used it more compared to the late signers.

\begin{figure}
\centering
\includegraphics{narration_paper_files/figure-latex/count-plot-1.pdf}
\caption{\label{fig:count-plot}Referent counts by Discourse function (Introduction, Maintanence, Re-introduction), Grammatical function (Classifier, Nominal, Pronominal, Zero anaphora), and Participant group (Native, Late)}
\end{figure}

For more inference on the data and have pairwise comparisons, we fit a regression model to the grammatical function of the responses using Discourse Status and Group as predictors. We used the brms package (Bürkner, 2018) in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late). We did not define priors ourselves but used the preset uninformed priors. Figure \ref{fig:model-gram} shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50\% credible intervals, and the thin line is the 95\% credible intervals.

\begin{figure}
\centering
\includegraphics{narration_paper_files/figure-latex/model-gram-1.pdf}
\caption{\label{fig:model-gram}Regression model plot for the Grammatical Function (Nominal, Pronominal, Zero) with the predictors Discourse function (Introduction, Maintenance, Re-introduction) and Participant group (Native, Late)}
\end{figure}

In a categorical distribution, the model takes one of the responses as a reference and compares all other responses to it. For example the first coefficient, nominal intercept, in Figure \ref{fig:model-gram} is the comparison of Nominal referent use versus a classifier referent use. It means that participants produced more nominal referents than classifier referents in general regardless of participant group and discourse function. The second coefficient shows that zero anaphora referents are also used more than classifier referents by the participants in general, but not as much as nominal referents.

In the comparison of nominal versus classifier referents, being a native acquirer of TiD did not have an effect. Nominal referents were used more in the introduction setting of a discourse compared to a re-introduction setting, again the whole comparison is being made against a classifier referent. In contrast to the introduction setting, nominal referents were used less in the maintain discourse setting compared to the re-introduction setting, within the greater comparison to classifier referents. There does not seem to be an interaction between introduction setting of discourse and being a native acquirer of TiD, however there looks like an interaction between the Maintain setting of the discourse with being a native acquirer of TiD. In that, native signers produced less nominal referents in the maintain setting of the discourse compared to the other combinations.

In zero anaphora referents compared to classifier referents, being a native acquirer of TiD increased the odds of using zero anaphora in general. Zero anaphora were used less as introduction setting of discourse compared to the re-introduction setting. This contrasts with the nominal referent (4th coefficient) as it was used more in the introduction setting versus re-introduction. In contrast to the introduction setting of the discourse, zero anaphora referents were used more as maintain setting of discourse compared to the re-introduction setting, yet again within the greater comparison to the classifier referents. There seems to be an interaction between the introduction setting of discourse and being a native acquirer of TiD. In that, the odds of using zero anaphora referent further increases if the participant is a native acquirer of TiD and the referent is used in the introduction setting. There is another interaction with regards to the maintain discourse setting, contrasting the first interaction. In that, the odds of using zero anaphora decrease if the participant is a native acquirer of TiD and the referent is used in the maintain setting of discourse.

We infer the following from the results of our first model. Participants used nominal referents and zero anaphora most of the time compared to classifier referents. Being a native acquirer of TiD increases the overall chances of using zero anaphora referents. However, this effect interacts with the discourse setting. The chances of a native acquirer of TiD using zero anaphora referent increases if they are to use it as a maintain function of the discourse. This interaction is reversed in the case of nominal referents. The chances of a native acquirer of TiD using nominal referents decrease if they are to use it as a maintaion function of the discourse. Furthermore, the chances of using zero anaphora referents decreases if the native participants are to use it as the introduction function of the discourse.

In addition to the discourse and grammatical function of a referent, we also coded the perspective of the participant in their response. In the responses where the participants simulated the story in their summary, they could have: (i) acted as the character, (ii) mixed character and observer perspectives, or (iii) been a complete observer in their reference tracking. Figure \ref{fig:plot-perspective} shows the distribution of the responses by the Discourse function, Perspective (Character, Mixed, Observer), and the participant group.

\begin{figure}
\centering
\includegraphics{narration_paper_files/figure-latex/plot-perspective-1.pdf}
\caption{\label{fig:plot-perspective}Distribution of the simulated reference tracking by Discourse function (Introduction, Maintenance, Re-introduction), Perspective (Character, Mixed, Observer), and the Participant group (Native, Late).}
\end{figure}

For more inference on the data and have pairwise comparisons, we fit a regression model to the perspective of the responses using Discourse function and the participant group as predictors. We used the brms package in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse (Intro, Maintenance, Re-intro) and Participant group (Native, Late). We did not define priors and used the preset uninformed priors. Figure \ref{fig:model-gram} shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50\% credible intervals, and the thin line is the 95\% credible intervals.

\begin{figure}
\centering
\includegraphics{narration_paper_files/figure-latex/model-pers-1.pdf}
\caption{\label{fig:model-pers}Regression model plot for the Perspective of the reference with the predictors Participant group (Native, Late) and Discourse function (Introduction, Maintain, Re-introduction)}
\end{figure}

The results of this model too are interpreted with a reference level of the response in mind, here the Observer perspective is the reference level with regards to which we make the comparisons. Participants in general used less mixed perspective in their referents compared to the observer perspective (1st coefficient), they also used less character perspective compared to the observer perspective. The magnitude of the decrease for Mixed versus Observer is greater than that of Character versus Observer. This means that participants used more Observer and Character perspectives than Mixed perspectives. The posterior probability distributions for both intercepts are left skewed, this shows that there is higher variance towards the lower end of the estimate. This means that there are extreme cases where Mixed and Character perspectives were used even less than the Observer perspective.

Being a native acquirer of TiD increased the chances of using mixed perspective compared to using the observer perspective. Mixed perspective was used less in the introduction setting of the discourse compared to the re-introduction setting. In contrast, mixed perspective was used more in the maintain setting of the discourse compared to the re-introduction setting. This means that mixed perspective was used more as introduction and re-introduction settings than the maintain setting. There does not seem to be an interaction between being a native acquirer of TiD and the discourse function that effects the chances of using mixed perspective compared to the observer perspective.

Being a native acuqirer of TiD increased the chances of using character perspective compared to suing the observer perspective. In a similar spirit to the effect of introduction setting in mixed versus observer (4th coefficient), character observer was used less as the introduction setting of the discourse compared to the re-introduction setting. In contrast, character perspective was used more in the maintain setting of the discourse compared to the re-introduction setting. This means that character perspective was used more as introduction and re-introduction settings than the maintain setting. There seems to be interactions of being a native acquirer of TiD and discourse settings this time. The chances of character perspective further increase if the participant is a native acquirer of TiD and they are to use the referent in the introduction setting of the discourse. The chances of character perspective decrease if the participant is a native acquirer of TiD and they are to use the referent in the maintain setting of the discourse.

We infer the following from the results of our second model. Participants used Observer and Character perspective more than Mixed perspective in general. Being a native acquirer of TiD increases the chances of using Mixed or Character perspective compared to the observer perspective. The discourse setting introduction decreases the chances of using Character or Mixed perspective, but this is partly undone for the character perspective by this setting's interaction with being a native acquirer of TiD.

In general, the posterior probability distributions of our second model's coefficients are left or right skewed. This means that there was high variance towards the extreme in a way that our data or experiment setting was not enough to prune them out. We are noting this difference of the first and second models to show that, although we treat posterior probability distributions as categorical statements of an effect, not all interpretations of the estimates are the same.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

\newpage

\hypertarget{references}{%
\section{References}\label{references}}

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-ahn2019}{}%
Ahn, D. (2019). \emph{The determinacy scale: A competition mechanism for anaphoric expressions} (PhD thesis).

\leavevmode\hypertarget{ref-aksu-kouxe72015}{}%
Aksu-Koç, A., \& Nicolopoulou, A. (2015). Character reference in young children's narratives: A crosslinguistic comparison of English, Greek, and Turkish. \emph{Lingua}, \emph{155}, 62--84. \url{https://doi.org/10.1016/j.lingua.2014.04.006}

\leavevmode\hypertarget{ref-anderson2005}{}%
Anderson, D. (2005). \emph{Lexical Development of Deaf Children Acquiring Signed Languages}. Oxford University Press. Retrieved from \url{https://www.oxfordscholarship.com/view/10.1093/acprof:oso/9780195180947.001.0001/acprof-9780195180947-chapter-6}

\leavevmode\hypertarget{ref-ariel1990}{}%
Ariel, M. (1990). \emph{Accessing Noun-Phrase Antecedents}. Routledge.

\leavevmode\hypertarget{ref-bamberg1997}{}%
Bamberg, M. G. (1997). Positioning between structure and performance. \emph{Journal of Narrative and Life History}, \emph{7}(1-4), 335342.

\leavevmode\hypertarget{ref-becker2009}{}%
Becker, C. (2009). Narrative competences of Deaf children in German Sign Language. \emph{Sign Language \& Linguistics}, \emph{12}(2), 113--160. \url{https://doi.org/10.1075/sll.12.2.02bec}

\leavevmode\hypertarget{ref-bel2015}{}%
Bel, A., Ortells, M., \& Morgan, G. (2015). Reference control in the narratives of adult sign language learners. \emph{International Journal of Bilingualism}, \emph{19}(5), 608--624. \url{https://doi.org/10.1177/1367006914527186}

\leavevmode\hypertarget{ref-boudreault2006}{}%
Boudreault, P., \& Mayberry, R. I. (2006). Grammatical processing in american sign language: Age of first-language acquisition effects in relation to syntactic structure. \emph{Language and Cognitive Processes}, \emph{21}(5), 608--635. \url{https://doi.org/10.1080/01690960500139363}

\leavevmode\hypertarget{ref-buxfcrkner2018}{}%
Bürkner, P.-C. (2018). Advanced Bayesian Multilevel Modeling with the R Package brms. \emph{The R Journal}, \emph{10}(1), 395. \url{https://doi.org/10.32614/RJ-2018-017}

\leavevmode\hypertarget{ref-chafe1976}{}%
Chafe, W. (1976). \emph{Givenness, contrastiveness, definiteness, subjects, topics, and point of view}. Retrieved from \href{https:///paper/Givenness\%2C-contrastiveness\%2C-definiteness\%2C-subjects\%2C-Chafe/805a12a9bee00562f6c2663722e9f7f3f5991848}{/paper/Givenness\%2C-contrastiveness\%2C-definiteness\%2C-subjects\%2C-Chafe/805a12a9bee00562f6c2663722e9f7f3f5991848}

\leavevmode\hypertarget{ref-chamberlain_language_1999}{}%
Chamberlain, C., Morford, J. P., \& Mayberry, R. I. (Eds.). (1999). \emph{Language acquisition by eye}. Mahwah, N.J: Psychology Press.

\leavevmode\hypertarget{ref-corina2014}{}%
Corina, D. P., Hafer, S., \& Welch, K. (2014). Phonological Awareness for American Sign Language. \emph{The Journal of Deaf Studies and Deaf Education}, \emph{19}(4), 530--545. \url{https://doi.org/10.1093/deafed/enu023}

\leavevmode\hypertarget{ref-cormier2013}{}%
Cormier, K., Smith, S., \& Zwets, M. (2013). Framing constructed action in British Sign Language narratives. \emph{Journal of Pragmatics}, \emph{55}, 119--139. \url{https://doi.org/10.1016/j.pragma.2013.06.002}

\leavevmode\hypertarget{ref-dikyuva2015}{}%
Dikyuva, H., Makaroğlu, B., \& Arık, E. (2015). Türk i{ş}aret dili dilbilgisi kitab{ı}. \emph{Aile Ve Sosyal Politikalar Bakanl{ı}{ğ}{ı} Yay{ı}nlar{ı}: Ankara}.

\leavevmode\hypertarget{ref-dudis2004}{}%
Dudis, P. G. (2004). Body partitioning and real-space blends. \emph{Cognitive Linguistics}, \emph{15}(2). \url{https://doi.org/10.1515/cogl.2004.009}

\leavevmode\hypertarget{ref-emmorey1996}{}%
Emmorey, K. (1996). \emph{The confluence of space and language in signed languages}. In \emph{Language, Speech, and Communication} (pp. 171--209). Cambridge, MA, US: The MIT Press.

\leavevmode\hypertarget{ref-emmorey1995}{}%
Emmorey, K., Corina, D., \& Bellugi, U. (1995). \emph{Differential processing of topographic and referential functions of space.}

\leavevmode\hypertarget{ref-engberg-pedersen1993}{}%
Engberg-Pedersen, E. (1993). \emph{Space in danish sign language: The semantics and morphosyntax of the use of space in a visual language}.

\leavevmode\hypertarget{ref-ferrara2012}{}%
Ferrara, L., \& Johnston, T. (2012). Enactment in discourse: Conceptualization through language and gesture. \emph{Language, Culture and Mind V: Integrating Semiotic Resources in Communication and Creativity. Catholic University of Portugal, Lisbon}.

\leavevmode\hypertarget{ref-frederiksen2016}{}%
Frederiksen, A. T., \& Mayberry, R. I. (2016). Who's on First? Investigating the referential hierarchy in simple native ASL narratives. \emph{Lingua}, \emph{180}, 49--68. \url{https://doi.org/10.1016/j.lingua.2016.03.007}

\leavevmode\hypertarget{ref-frederiksen2019}{}%
Frederiksen, A. T., \& Mayberry, R. I. (2019). Reference tracking in early stages of different modality L2 acquisition: Limited over-explicitness in novice ASL signers{'} referring expressions. \emph{Second Language Research}, \emph{35}(2), 253--283. \url{https://doi.org/10.1177/0267658317750220}

\leavevmode\hypertarget{ref-givon1983}{}%
Givon, T. (1983). \emph{Topic continuity in discourse}. John Benjamins Publishing Company. Retrieved from \url{https://benjamins.com/catalog/tsl.3}

\leavevmode\hypertarget{ref-goldin-meadow2005}{}%
Goldin-Meadow, S. (2005). \emph{The Resilience of Language: What Gesture Creation in Deaf Children Can Tell Us about how All Children Learn Language}. Psychology Press.

\leavevmode\hypertarget{ref-grice1975}{}%
Grice, H. P. (1975). \emph{Logic and conversation}. Brill.

\leavevmode\hypertarget{ref-gullberg2006}{}%
Gullberg, M. (2006). Handling Discourse: Gestures, Reference Tracking, and Communication Strategies in Early L2: Language Learning Vol. 56, No. 1. \emph{Language Learning}, \emph{56}(1), 155--196. \url{https://doi.org/10.1111/j.0023-8333.2006.00344.x}

\leavevmode\hypertarget{ref-gundel1993}{}%
Gundel, J. K., Hedberg, N., \& Zacharski, R. (1993). Cognitive status and the form of referring expressions in discourse. \emph{Language}, \emph{69}(2), 274--307. \url{https://doi.org/10.2307/416535}

\leavevmode\hypertarget{ref-guxfcr2018}{}%
Gür, C. (2018). \emph{The Effects of Late Language Acquisition on the Development of Narrative Skills: Insights from Turkish Sign Language} (PhD thesis). Erzurum.

\leavevmode\hypertarget{ref-hickmann1999}{}%
Hickmann, M., \& Hendriks, H. (1999). Cohesion and anaphora in children's narratives: a comparison of English, French, German, and Mandarin Chinese. \emph{Journal of Child Language}, \emph{26}(2), 419--452. \url{https://doi.org/10.1017/S0305000999003785}

\leavevmode\hypertarget{ref-hickmann1996}{}%
Hickmann, M., Hendriks, H., Roland, F., \& Liang, J. (1996). The marking of new information in children's narratives: a comparison of English, French, German and Mandarin Chinese*. \emph{Journal of Child Language}, \emph{23}(3), 591--619. \url{https://doi.org/10.1017/S0305000900008965}

\leavevmode\hypertarget{ref-janzen2004}{}%
Janzen, T. (2004). Space rotation, perspective shift, and verb morphology in ASL. \emph{Cognitive Linguistics}, \emph{15}(2). \url{https://doi.org/10.1515/cogl.2004.006}

\leavevmode\hypertarget{ref-karaduxf6ller2017}{}%
Karadöller, D. Z., Sümer, B., \& Özyürek, A. (2017). \emph{Effects of Delayed Language Exposure on Spatial Language Acquisition by Signing Children and Adults}. 5.

\leavevmode\hypertarget{ref-koulidobrova2016}{}%
Koulidobrova, E., \& Lillo-Martin, D. (2016). A {`}point{'}of inquiry: The case of the (non-) pronominal IX in ASL. \emph{The Impact of Pronominal Form on Interpretation}, \emph{221}, 250.

\leavevmode\hypertarget{ref-liddell2003}{}%
Liddell, S. K. (2003). \emph{Grammar, Gesture, and Meaning in American Sign Language}. Cambridge University Press.

\leavevmode\hypertarget{ref-loew1984}{}%
Loew, R. C. (1984). \emph{Roles and reference in American sign language : a developmental perspective}. Retrieved from \href{https:///paper/Roles-and-reference-in-American-sign-language-\%3A-a-Loew/315c2c539d3bd88fbc13564775141ba4a60ac5aa}{/paper/Roles-and-reference-in-American-sign-language-\%3A-a-Loew/315c2c539d3bd88fbc13564775141ba4a60ac5aa}

\leavevmode\hypertarget{ref-lu2016}{}%
Lu, J., Jones, A., \& Morgan, G. (2016). The impact of input quality on early sign development in native and non-native language learners*. \emph{Journal of Child Language}, \emph{43}(3), 537--552. \url{https://doi.org/10.1017/S0305000915000835}

\leavevmode\hypertarget{ref-macsweeney2008}{}%
MacSweeney, M., Waters, D., Brammer, M. J., Woll, B., \& Goswami, U. (2008). Phonological processing in deaf signers and the impact of age of first language acquisition. \emph{NeuroImage}, \emph{40}(3), 1369--1379. \url{https://doi.org/10.1016/j.neuroimage.2007.12.047}

\leavevmode\hypertarget{ref-metzger1995}{}%
Metzger, M. (1995). Constructed action and constructed dialogue in american sign language. \emph{Sociolinguistics in Deaf Communities}, 255271.

\leavevmode\hypertarget{ref-morgan1999}{}%
Morgan, G. (1999). Event packaging in british. \emph{Storytelling and Conversation: Discourse in Deaf Communities}, \emph{5}, 27.

\leavevmode\hypertarget{ref-morgan2000}{}%
Morgan, G. (2000). \emph{Discourse cohesion in sign and speech}. \url{https://doi.org/10.1177/13670069000040030101}

\leavevmode\hypertarget{ref-morgan2005}{}%
Morgan, G. (2005). \emph{The Development of Narrative Skills in British Sign Language} (B. Schick, M. Marschark, \& P. E. Spencer, Eds.). Oxford University Press. \url{https://doi.org/10.1093/acprof:oso/9780195180947.001.0001}

\leavevmode\hypertarget{ref-newport1990}{}%
Newport, E. L. (1990). Maturational Constraints on Language Learning. \emph{Cognitive Science}, \emph{14}(1), 11--28. \url{https://doi.org/10.1207/s15516709cog1401_2}

\leavevmode\hypertarget{ref-uxf6zsoy2014}{}%
Özsoy, S., \& Nuhbalaoğlu, D. (2014). \emph{Formal and experimental approaches in sign languages}. 9--11. University of Venice.

\leavevmode\hypertarget{ref-uxf6zyuxfcrek2010}{}%
Özyürek, A., \& Perniss, P. (2010). \emph{Event representations in signed languages} (E. Pederson \& J. Bohnemeyer, Eds.). In (pp. 84--107). Cambridge: Cambridge University Press. \url{https://doi.org/10.1017/CBO9780511782039.005}

\leavevmode\hypertarget{ref-padden1986}{}%
Padden, C. (1986). \emph{Verbs and role-shifting in american sign language}. \emph{44}, 57. National Association of the Deaf Silver Spring, MD.

\leavevmode\hypertarget{ref-perniss2015}{}%
Perniss, P., \& Özyürek, A. (2015). Visible Cohesion: A Comparison of Reference Tracking in Sign, Speech, and Co-Speech Gesture. \emph{Topics in Cognitive Science}, \emph{7}(1), 36--60. \url{https://doi.org/10.1111/tops.12122}

\leavevmode\hypertarget{ref-petitto1997}{}%
Petitto, L. A. (1997). \emph{In the Beginning: On the Genetic and Environmental Factors that Make Early Language Acquisition Possible} (M. Gopnik, Ed.). England: Oxford University Press. Retrieved from \url{https://www.academia.edu/8722491/In_the_Beginning_On_the_Genetic_and_Environmental_Factors_that_Make_Early_Language_Acquisition_Possible}

\leavevmode\hypertarget{ref-prince1981}{}%
Prince, E. (1981). \emph{Toward a taxonomy of given - new information} (P. Cole, Ed.). New York: Academic Press.

\leavevmode\hypertarget{ref-rathmann2007}{}%
Rathmann, C., Mann, W., \& Morgan, G. (2007). Narrative Structure and Narrative Development in Deaf Children. \emph{Deafness \& Education International}, \emph{9}(4), 187--196. \url{https://doi.org/10.1179/146431507790559932}

\leavevmode\hypertarget{ref-sevinuxe72006}{}%
Sevinç, A. M. (2006). Grammatical relations and word order in turkish sign language (t{İ}d). \emph{Unpublished Master's Thesis, Middle East Technical University, Turkey}.

\leavevmode\hypertarget{ref-sloetjes2008}{}%
Sloetjes, H., \& Wittenburg, P. (2008, May). \emph{Annotation by category: ELAN and ISO DCR}. Marrakech, Morocco: European Language Resources Association (ELRA). Retrieved from \url{http://www.lrec-conf.org/proceedings/lrec2008/pdf/208_paper.pdf}

\leavevmode\hypertarget{ref-smith2014}{}%
Smith, S., \& Cormier, K. (2014). In or Out?: Spatial Scale and Enactment in Narratives of Native and Nonnative Signing Deaf Children Acquiring British Sign Language. \emph{Sign Language Studies}, \emph{14}(3), 275--301. \url{https://doi.org/10.1353/sls.2014.0008}

\leavevmode\hypertarget{ref-sorace2011}{}%
Sorace, A. (2011). Pinning down the concept of {``}interface{''} in bilingualism. \emph{Linguistic Approaches to Bilingualism}, \emph{1}(1), 133.

\leavevmode\hypertarget{ref-sorace2009}{}%
Sorace, A., \& Serratrice, L. (2009). Internal and external interfaces in bilingual language development: Beyond structural overlap. \emph{International Journal of Bilingualism}, \emph{13}(2), 195--210. \url{https://doi.org/10.1177/1367006909339810}

\leavevmode\hypertarget{ref-sperber1986}{}%
Sperber, D., \& Wilson, D. (1986). \emph{Relevance: Communication and cognition} (Vol. 142). Harvard University Press Cambridge, MA.

\leavevmode\hypertarget{ref-supalla1986}{}%
Supalla, T. (1986). The classifier system in american sign language. \emph{Noun Classes and Categorization}, \emph{7}, 181214.

\leavevmode\hypertarget{ref-talmy2003}{}%
Talmy, L. (2003). The representation of spatial structure in spoken and signed language. \emph{Perspectives on Classifier Constructions in Sign Language}, 169195.

\leavevmode\hypertarget{ref-vercaingne-muxe9nard2001}{}%
Vercaingne-Ménard, A., Godard, L., \& Labelle, M. (2001). The emergence of narrative discourse in two young deaf children. \emph{Signed Languages: Discoveries from International Research}, 120134.

\leavevmode\hypertarget{ref-weaver2011}{}%
Weaver, K. A., \& Starner, T. (2011). \emph{The proceedings of the 13th international ACM SIGACCESS conference}. 91. Dundee, Scotland, UK: ACM Press. \url{https://doi.org/10.1145/2049536.2049554}

\leavevmode\hypertarget{ref-williams1988}{}%
Williams, J. (1988). Zero Anaphora in Second Language Acquisition: A Comparison among Three Varieties of English. \emph{Studies in Second Language Acquisition}, \emph{10}(3), 339--370. \url{https://doi.org/10.1017/S0272263100007488}

\leavevmode\hypertarget{ref-woll2013}{}%
Woll, B. (2013). \emph{Sign language and spoken language development in young children: Measuring vocabulary by means of the CDI} (L. Meurant, A. Sinte, M. Van Herreweghe, \& M. Vermeerbergen, Eds.). Berlin, Boston: DE GRUYTER. \url{https://doi.org/10.1515/9781614511472.15}

\end{CSLReferences}

\newpage

\begin{table}[tbp]

\begin{center}
\begin{threeparttable}

\caption{\label{tab:table-a}Participant demographic information based on self-report}

\small{

\begin{tabular*}{\textwidth}{c @{\extracolsep{\fill}}lllllllll}
\toprule
No & \multicolumn{1}{c}{Status} & \multicolumn{1}{c}{Sex} & \multicolumn{1}{c}{AOA} & \multicolumn{1}{c}{Age} & \multicolumn{1}{c}{\shortstack{Hearing \\loss}} & \multicolumn{1}{c}{LOE} & \multicolumn{1}{c}{\shortstack{Home \\language}} & \multicolumn{1}{c}{\shortstack{Parents' \\hearing}}\\
\midrule
1 & Native & M & 0-3 & 33 & Prelingual & HS & TİD, HS & Both deaf\\
2 & Native & F & 0-3 & 27 & Prelingual & HS & TİD, HS & Both deaf\\
3 & Native & F & 0-3 & 27 & Prelingual & HS & HS & Both deaf\\
4 & Native & M & 0-3 & 29 & Prelingual & HS & TİD, HS & Both deaf\\
5 & Native & M & 0-3 & 28 & 4-7 years & HS & TİD & Both deaf\\
6 & Native & M & 0-3 & 32 & Prelingual & HS & TİD & F deaf, M hearing\\
7 & Native & M & 0-3 & 35 & Prelingual & HS & TİD & Both deaf\\
8 & Native & F & 0-3 & 23 & Prelingual & HS & TİD & Both deaf\\
9 & Native & F & 0-3 & 30 & Prelingual & HS & TİD & Both deaf\\
10 & Native & F & 0-3 & 25 & Prelingual & HS & TİD & Both deaf\\
11 & Native & F & 0-3 & 24 & Prelingual & HS & TİD & Both deaf\\
12 & Native & M & 0-3 & 26 & Prelingual & AD & TİD & Both deaf\\
13 & Native & M & 0-3 & 18 & Prelingual & HS & TİD & Both deaf\\
14 & Native & F & 0-3 & 21 & Prelingual & BA & TİD & Both deaf\\
15 & Native & M & 0-3 & 19 & Prelingual & HS & TİD & Both deaf\\
16 & Native & F & 0-3 & 25 & Prelingual & HS & TİD & Both deaf\\
17 & Late & F & 4-7 & 35 & Prelingual & AD & TİD, HS, TR & Both hearing\\
18 & Late & M & 13-17 & 50 & 4-7 years & MS & TR, HS & Both hearing\\
19 & Late & F & 8-12 & 49 & Prelingual & MS & HS, TR & Both hearing\\
20 & Late & M & 8-12 & 43 & Prelingual & BA & TR & Both hearing\\
21 & Late & M & 4-7 & 36 & 4-7 years & HS & TR, HS & Both hearing\\
22 & Late & M & 8-12 & 28 & Prelingual & BA & TR & Both hearing\\
23 & Late & M & 4-7 & 33 & Prelingual & HS & TR & Both hearing\\
24 & Late & M & 13-17 & 31 & Prelingual & BA & TR & Both hearing\\
25 & Late & F & 4-7 & 33 & Prelingual & HS & TR, HS & Both hearing\\
26 & Late & F & 8-12 & 24 & Prelingual & HS & HS & Both hearing\\
27 & Late & M & 8-12 & 33 & Prelingual & MS & HS & Both hearing\\
28 & Late & F & 4-7 & 31 & Prelingual & AD & HS & Both hearing\\
29 & Late & M & 4-7 & 35 & Prelingual & HS & HS & Both hearing\\
\bottomrule
\addlinespace
\end{tabular*}

}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} AOA = age of acquisition, LOE = level of education, HS = high school, AD = associate degree, MS = middle school, BA/BS = Bachelor of Arts/Science, TR = (Spoken) Turkish.}
\end{tablenotes}

\end{threeparttable}
\end{center}

\end{table}

\endgroup

\end{document}

### 2022-08-05 manuscript_v2.Rmd
Add files via upload

---
title: Reference Tracking Strategies of Native and Late Signers in Turkish Sign Language
abstract: |
  We investigated the reference tracking strategies among native and late adult deaf signers of Turkish Sign Language. Consistent with the theories of saliency and referential accessibility, regardless of their acquisiton groups, signers mainly used nominals and extension classifiers for introducing referents. To maintain a referent in the previous clause, signers chose null marking (e.g., constructed action, agreement and plain verbs). For topic shifts or re-introduced contexts, nominals and pronouns were chiefly favored although we observed very little pronominal use. As for age of acquisition effects, we only report limited over-redundancy for late signers who used null marking less compared to native signers especially for introduced and maintained contexts. Given our limited support for redundancy by late signers, we conclude that native signers might utilize the spatial affordances of the visual-spatial modality better than late signers only to a certain extent.
keywords: "Referential accessibility, signed narratives, pragmatic competence, delayed language acquisition, Turkish Sign Language"
journal: "Journal of Pragmatics"
date: "`r Sys.Date()`"
bibliography: ./bibliography/references.bib
times: true
numbersections: true
classoption: "review"
csl: ./bibliography/elsevier-harvard.csl
output: rticles::elsevier_article
header-includes:
   - \usepackage{floatrow}
   - \usepackage{hyperref}
   - \usepackage{enumitem}  
   - \usepackage{float}
   - \floatplacement{figure}{H}
   - \floatplacement{table}{H}
   - \floatsetup[figure]{capposition=top}
   - \floatsetup[table]{capposition=top}
   - \usepackage{times}
   - \usepackage{smartdiagram}
   - \usepackage{ragged2e}
   - \usepackage{tikz}
   - \usepackage{xcolor}
   - \usepackage[utf8]{inputenc}
   - \usepackage[T1]{fontenc}
   - \usepackage{graphicx}
   - \usepackage{./sty/langsci-gb4e}
   - \usepackage{caption}
   - \usepackage{multicol}
   - \setlength{\columnsep}{-3cm}
   - \usepackage{setspace}
---

# Introduction

Narratives are extensive discourses that construct an important part of human communication. In order to create a systematic and coherent organization of experiences, narrators must choose appropriate linguistic forms to refer to entities [@ariel1990; @givon1983; @gundel1993]: Certain forms such as nominal expressions are commonly used for the first mention of a referent in a discourse, and other forms such as pronouns and null marking may track previously mentioned entities which could either be maintained across multiple clauses or re-introduced back into the discourse. The selection of these referring expressions (henceforth REs) is contingent on cognitive abilities, such that individuals must rely on the activation of the working memory to keep in mind previously referred expressions and long strings of information [@morgan2005; @bamberg1997]. Pragmatic abilities too play a pivotal role in the tracking of referents. The form with the least amount of redundant information must be selected based on the accessibility levels of the referent (i.e., how active a referent is in the addressee's mind given a point in discourse) [@ahn2019; @ariel1990]. Because of these cognitive and pragmatic necessities, it is not surprising to observe that producing a coherent narrative is one of the latest milestones in language acquisition among children [@hickmann1996]. As can be expected, there is a plethora of research examining how narrative competence is achieved in both monolingual and bilingual acquisition contexts among children and adults. Although narrative competence and referent selection have been studied extensively for typically developing speakers [@aksu-koç2015; @hickmann1996; @williams1988], deaf individuals who acquired a sign language in atypical learning contexts (i.e., deaf signers with linguistic deprivation) have remained mostly understudied with an exception of handful of studies [@cormier2013; @gür2018; @becker2009]. Nevertheless, it is paramount that we investigate and describe the effects of language deprivation among deaf adults on narrative competence more in depth. This will help us understand whether the loss of systematic and frequent language input in the early years of life affect how deaf signers produce a narrative in their adulthood. Given this, the present study will focus on deaf adult signers with language deprivation and examine their reference tracking strategies in signed narratives.

Sign Languages are natural languages of the Deaf communities all around the world. They are independent languages, which do not derive from majority spoken language(s) but emerge naturally when deaf people form a community. Turkish Sign Language (Türk İşaret Dili -- TİD) is and has been the officially recognized language of the Deaf community in Turkey since 2005 [see @nuhbalaoglu2018 for a brief review of the historical development of the language]. Deaf or hearing children acquiring a SL within a family where there is one or more deaf caregivers display the same developmental milestones as hearing children acquiring a spoken language in a hearing family [@chamberlain_language_1999; @meier1990]. However, reports indicate that over 90% of all deaf individuals are born into hearing families who do not know how to sign (i.e., those with language deprivation) [@woll2013]. This means that the auditory modality used in the household is not accessible to many deaf children. That is why the age of acquisition of a first functional language varies greatly among deaf children, with many of whom being deprived of their first languages typically until after age 7 when they start formal education in a deaf school. First language deprivation or delay has been found to have negative effects on certain cognitive and linguistic measures among late signing children. To name a few, delayed exposure to a first functional language is reported to negatively affect deaf children's morphosyntactic production and processing [@kayaba2022; @boudreault2006; @newport1990; @cheng2021; @karadöller2017; @cormier2012] and narrative abilities [e.g., @becker2009]. In our overall research, we are concerned with the effects of this late exposure in the ultimate adult forms of TİD users.

## Referring Expression (RE) Selection

Natural languages can be systematically used to keep track of previous referents in the discourse and mark the changing cognitive statuses of the REs (i.e., from being active to less active in the addressee's mind) to create a more coherent narrative. Among the theoretical frameworks that aim to explain that different linguistic forms are selected in accordance with the addressee's mental representations of the referents in the discourse is the Givenness Hierarchy [@gundel1993]. According to this model, given or familiar information is expressed in more attenuated forms such as pronominal use and null marking since the mental status of the expressed referent is already at the center of attention, as indicated by the present contextual information [@chafe1976]. The Givenness Hierarchy is akin to other previously proposed models of reference control within the topicality and saliency paradigm such as the Accessibility Hierarchy [@ariel1990], Topic Continuity [@givon1983], and Familiarity Scale [@prince1981]. @gundel1993 argued that the distribution of forms is not random but abides by the Gricean Maxim of Quantity [@grice1975]. That is to say, individuals must make the most informative and succinct choice possible to retain economy of form. When multiple expressions are available for the same referent as in an utterance like (1), phonetically simplest form is selected with the least redundant information are selected based on an economy principle. In (1), REs for "Eddie" are given in bold. All three REs (nominal, pronominal, and null marking) refer to the same entity; however, the quantity of marking decreases as the referent's accessibility increases. On this ground, for highly accessible referents (i.e., those that are highly active in the mind), linguistic forms with the least phonetic content are selected for the least amount of processing effort.

```{=tex}
\begin{exe}
    \ex
    So \textbf{Eddie} turned around. \hfill Williams (1988; p. 343, ex. 7) \\
\textbf{He} said, "youse got a problem?" \\
"Yeah, we want you," they say.
So-\textbf{$\emptyset$} walked right up to them \\
\end{exe}
```
It is also possible to categorize REs into three according to their discourse status [@gullberg2006]. A referent can be introduced or mentioned for the first time, maintained, that is, continued across two or more clauses, or re-introduced into the discourse following a topic shift. As a result, interlocutors mark some REs differently than others based on the principles of reference accessibility. For spoken languages, this observation is well established [@ariel1990; @givon1983; @gundel1993; @debreslioska2013; @ahrenholz2005; @chafe1976; @hickmann1996; @azar2015]. These research indicate that the introduction and re-introduction of a referent are often accomplished by full noun phrases which can be commonly modified by an adjective or demonstrative. Maintenance, on the other hand, is typically marked with pronominals and null markers since the continued referent is still highly accessible to the addressee.

However, this particular observation is not peculiar to spoken languages. There is an emerging body of literature that reports the distribution of REs in signed narratives. Building on @gundel1993's work, @swabey2002 demonstrated that, like other spoken languages such as Russian, Chinese, and Japanese, American Sign Language (ASL) signers use bare nouns for introduced and reintroduced discourse contexts whereas null marking is used for referents which are "in-focus" or very accessible to the addressee. Overt REs like pronouns or nouns are disprefered for such referents. @swabey2002 concluded that "native" signers in her study abided by the Gricean maxim of quantity and were thus economic in their reference tracking. Similar findings have been reported for German Sign Language (DGS) [@perniss2015] and British Sign Language (BSL) [@morgan2005]. It is important to note, however, that these research either collapsed native and late signers into a single group and labeled them as "native-like" or explicitly compared native deaf signers with late deaf signers (with language deprivation) in terms of narrative competence.

### Signed vs. Spoken REs

Commonly in both spoken and sign languages nouns and noun phrases are used as referring expressions. Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions [@lillo-martin1990]. In addition, the pointing sign is used to referentially anchor a referent to a location in signing space. For instance, when the signer introduces \textsc{elvan} for the first time, s/he can assign her to the right. This referential locus can later be used by agreement verbs and classifier predicates (more below). For instance, in (2a) there are two referents, \textsc{elvan}, who is referentially-assigned to the right side of the signer, represented with the "a" subscript and \textsc{demet}, who is referentially-assigned to the left side of the signer, represented with the "b" subscript. When the signer makes a contrast in (2b), as a pronominal use, she points to the right side to refer to ELVAN and to the left to refer to \textsc{demet}. We schematically represents the referential locus assigned to the referents in (3).

```{=tex}
\begin{exe}
    \ex 
\begin{xlist}
    \ex
    \textsc{lab place elvan ix\textsubscript{a} ten years work. demet ix\textsubscript{b} last year start.} \\
    ‘In the lab, Elvan has been working for 10 years. Demet started last year.’
    \\
    \ex
    \textsc{ix\textsubscript{a} e-l-a-n know. ix\textsubscript{b} learn.}\\
    ‘She (Elvan) knows ELAN (the annotation program). She (Demet) is learning.’
    \end{xlist}
    \end{exe}

    \begin{exe}
    \ex Signing space, signer, example referential loci for the referents \textsc{elvan} (a) and \textsc{demet} (b) \\

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300

%Curve Lines [id:da7006577339157659] 
\draw    (100.25,77.81) .. controls (136.19,41.85) and (183.62,42.17) .. (222.5,78.13) ;
%Straight Lines [id:da8428827309782937] 
\draw    (112.9,128.8) -- (210.21,128.8) ;
%Curve Lines [id:da39290455666055124] 
\draw    (222.19,100.13) .. controls (221.89,103.9) and (217.55,123.77) .. (210.21,128.8) ;
%Curve Lines [id:da9909903731270634] 
\draw    (101.17,100.45) .. controls (100.86,104.22) and (108.2,124.02) .. (112.9,128.8) ;

% Text Node
\draw (88.75,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-b }};
% Text Node
\draw (91.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Demet)}};
% Text Node
\draw (208.25,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-a }};
% Text Node
\draw (212.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Elvan)}};
% Text Node
\draw (144,115.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Signer}};
\end{tikzpicture}
    \end{exe}
```
Grammatical gender as in Romance languages or noun classes as in Bantu languages are some reference tracking tools that are familiar from spoken languages. A somehow less known reference-tracking tool in spoken languages is *verbal classifiers*, which morphologically attach to the verb and often cross-reference a core argument in a clause [@aikhenvald2017]. Classifiers denote sets of objects based on semantic criteria such as size, shape, body-part, texture and sometimes function. Usually, the verbal classifier and the cross-referenced noun co-occur in the same clause when the argument is introduced into discourse. The classifier can track the argument on its own on subsequent occurrences. In sign languages, a verbal classifier morpheme is encoded by a *handshape* that is articulated phonologically co-temporally with a movement, which encodes the verbal meaning [@supalla1990; @benedicto2004]. These two infused morphemes form a complex predicate. Some classifier handshapes and the set of objects that they denote are below. Some classifier handshapes and the set of objects that they denote are presented below in (4).

```{=tex}
\begin{exe}
\ex 
\includegraphics[width=25pt]{pictures/flathand.png}: flat entities, surfaces, vehicles.\\
\includegraphics[width=25pt]{pictures/flat.png}: human agent holding a flat entity.\\
\includegraphics[width=25pt]{pictures/small_legged.png}: small animal legs.\\
\includegraphics[width=25pt]{pictures/index.png}: thin entities (like a pencil); upright beings.\\
\includegraphics[width=25pt]{pictures/hold.png}: human agent holding a thin entity (like a pencil or a branch of flower). 
\end{exe}
```
In (5), in the first clause, a referential locus, i.e. locus-a, is assigned to the locative source argument \textsc{cage}. In the second clause, the core argument \textsc{bird} is cross-referenced with the classifier handshape, i.e. small-legged animal. It flies from the referential locus of the \textsc{cage} (locus-a) to the referential locus of the goal \textsc{tree} (locus-b).

```{=tex}
\begin{exe}
    \ex 
\begin{xlist}
    \ex \textsc{cage} assigned to the referential locus-a. \\ \\
    \textsc{garden bird cage-a open} \\
    ‘In the garden, there is a cage and it is open.’
    \\

    \ex \textsc{tree} assigned to the referential locus-b. The source and goal tracked on the predicates \textsc{fly} and \textsc{land}. \\ \\
    \textsc{bird fly[manner]-a tree-b land. CL-b}\\
    ‘The bird flies (from the cage assigned to the left) to a tree (assigned to the right).’
    \end{xlist}
    \end{exe}
    \hfill (Adapted from Gökgöz and Sevgi 2020:318)
```
## Linguistic Tools for Tracking Referents in Signed Narratives

The linguistic inventory for discourse referents can include nominal constructions (bare nouns, fingerspelled nouns, and modified nouns), null markers (plain verb, agreement verb, and constructed action), classifier predicates (whole entity classifiers, body part classifiers, handling classifiers, and extension classifiers), and overt pronouns (pronominal index or IX) [@frederiksen2016]. This interim section describes such constructions attested for sign languages under four main headings (nominals, pronouns, classifiers, and null marking).

\clearpage

```{=tex}
\begin{multicols}{2}
\begin{exe}
\ex
\begin{xlist}
\ex Full noun \\\glll
\includegraphics[width=50pt]{pictures/1a_1.png}\\ \textsc{mouse\footnotemark} \\ \\ 
\ex Noun modified by classifier \\\glll
\includegraphics[width=50pt]{pictures/1c_2.png} \includegraphics[width=50pt]{pictures/1c_1.png} \\ \textsc{soap} \textsc{soap.cl} \\ \\ 
\ex Fingerspelled noun\\\glll 
\includegraphics[width=50pt]{pictures/1b_1.png} \includegraphics[width=50pt]{pictures/1b_2.png} \includegraphics[width=50pt]{pictures/1b_3.png}\\ \textsc{t-} \textsc{o-} \textsc{m} \\ \\
\ex Noun modified by index (IX)\\\glll 
\includegraphics[width=50pt]{pictures/1d_1.png} \includegraphics[width=50pt]{pictures/1d_2.png}\\ IX \textsc{cat} \\ \\
\end{xlist}
\end{exe}
\end{multicols}
\ea 
\ea Pronominal \\\glll
\includegraphics[width=50pt]{pictures/2_1.png} \includegraphics[width=50pt]{pictures/2_2.png} \\ IX \textsc{sick} \\ \\
\z \z
\begin{multicols}{4}
\begin{exe}
\ex
\begin{xlist}
\ex Whole entity \\\glll 
\includegraphics[width=50pt]{pictures/3e.png} \\ \textsc{fish.fall}\\ \\
\ex Body part \\\glll
\includegraphics[width=50pt]{pictures/3b.png} \\ \textsc{walk} \\ \\ 
\ex Handling \\\glll 
\includegraphics[width=50pt]{pictures/3c.png} \\ \textsc{hold} \\ \\
\ex Extension \\\glll 
\includegraphics[width=50pt]{pictures/3d_1.png} \includegraphics[width=50pt]{pictures/3a.png}\\ \textsc{plank} {\textsc{dh: mouse}} \\ {} {\textsc{ndh: tunnel}}\\ \\
\end{xlist}
\end{exe}
\end{multicols}
\begin{multicols}{3}
\begin{exe}
\ex
\begin{xlist}
\ex Constructed action \\\glll 
\includegraphics[width=50pt]{pictures/4c.png} \\ \textsc{struggle} \\ \\

\ex Plain verb  \\\glll
\includegraphics[width=50pt]{pictures/4a_2.png}  \\ \textsc{enter}\\ \\

\ex Agreement verb  \\\glll 
\includegraphics[width=50pt]{pictures/4b_1.png} 
\includegraphics[width=50pt]{pictures/4b_2.png} \\ {\textsc{$_{1}$hit$_{3}$}} {} \\ \\ 
\end{xlist}
\end{exe}
\end{multicols}
\footnotetext{We use small capitals to indicate sign glosses.}
```
### Nominals

In TİD and many other sign languages, it is possible to use bare noun phrases (\textsc{bare nom}) as in (1a) as well as fingerspelled nouns (\textsc{fs nom}) as in (1c). Fingerspelling refers to the use of the manual alphabet which is derived or borrowed from a surrounding written language [@kubu2008]. Nouns can also be modified (\textsc{mod nom}) by classifiers as in (1b) [@frederiksen2016] or by a prenominal or postnominal IX functioning as a determiner or a demonstrative as in (1d) [@maclaughlin1997].

### Pronouns

The IX sign (extended index finger) has various functions in sign languages, one of which is the pronominal use which refers to established nominals by pointing to a previously assigned abstract or physical locus in the signing space when re-introducing a topic in discourse [@emmorey1996]. The same IX function has also been identified for TİD [@sevinç2006]. The example in (2) illustrates a spatially-anchored pronominal reference for topic re-introduction. These signs are thus rarely used for the first mention of an entity [@perniss2015]. Other interpretations consider pointing signs without noun phrases as examples of demonstratives and not personal pronouns [@koulidobrova2016]. It is important to mention here that in some spoken languages (e.g., German and Dutch), demonstratives can also refer to persons [@bosch2007] and behave similarly to personal pronouns. However, for the purposes of the present study, we will refer to IX constructions as pronominal IX and particularly examine their non-deictic or personal pronoun realizations as a possible discourse strategy to mainly mark topic shift among native and late signers of TİD.

### Classifiers

Classifier predicates are highly complex morphological constructions which are very common in sign languages [@supalla1986]. These are manually signed iconic representations of animate or inanimate objects, and they embody several properties pertaining to the nominal referent including its location in the signing space, movement, path, size, and shape [@talmy2003]. As in most sign languages hitherto studied, classifier constructions are also common and found in TİD [see @dikyuva2015 for a review]. Following the classification of @engberg-pedersen1993, we distinguish between four types of classifier constructions. Whole entity classifiers (WECL) represent the entirety of any referent whereas body part classifiers (BPCL) only reflect a specific part of an animate entity (e.g., limbs, head, tongue, teeth, and so forth). Handling classifiers (HCL) occur with transitive verbs and depict how an object is handled or manipulated by agent. On this ground, an HCL[^1] depicts the manner of the referring action while the former two constructions encode the location of the entity as well as its motion in space. Extension classifiers (ExtCL), on the other hand, denote the size and shape of an entity by following a trajectory to trace its shape. We exemplify these constructions in (3).

[^1]: As discussed before, an HCL is transitive in nature and depicts the handling action performed by ananimate agent. This is why most research tend to consider HCLs as an extension to constructed action [@dudis2004; @cormier2013].

### Null marking

Parallel to spoken languages, many sign languages have been reported to allow null subjects and/or objects and make use of null arguments when referring to highly accessible referents which are still active in the mental representation of the addressee. In signed discourse, this can be achieved by employing certain verbal constructions. To identify the predicate types in ASL, @padden1986 suggested a triadic categorization of verbs in sign languages: plain verbs, agreement verbs, and spatial verbs. According to this classification, plain verbs (plain) (4b) do not agree with the subject or the object. This means that they are not inflected for number or person. This contrasts with agreeing verbs (agreement) (4c) the starting or the ending point of which can be modified to specific loci for person and number agreement while spatial verbs (as in \textsc{ix book put} 'I placed the book there') agree with locative arguments but like plain verbs, they are not inflected for number or person. TİD too licenses null arguments. Following this classification, @sevinç2006 argued that although it is common in TİD to drop arguments, there were certain limitations on the null marking of the arguments for agreement and plain verbs.

## Learning How to Track Referents in Sign Languages

In both communication modalities (visual and auditory), children start to produce more complex and coherent stories with appropriate and informative linguistic forms to indicate anaphoric or cataphoric references after age 10 [@hickmann1996; @rathmann2007]. Prior to this age, both signing and speaking children fail to refer to entities consistently since their pragmatic abilities are yet to develop. The manipulation of linguistic forms for discourse functions to signal topic continuance or shift is difficult to master until the beginning of formal education. These findings are also comparable across different sign languages including British Sign Language (BSL) [@morgan2005], ASL [@loew1984], and German Sign Language (Deutsche Gebärdensprache - DGS) [@becker2009], pointing out to the fact that the mastery of reference tracking in narratives is one of the most protracted achievements in the acquisition of signed discourse.

Like in first language development, narrative skills have been extensively studied among second language learners [see @bel2015; @frederiksen2019]. Comparing 11 deaf adults signers and 13 hearing advanced learners of Catalan Sign Language (Llengua de Signes Catalana - LSC), @bel2015 found that hearing signers resort to more overt forms (nominals and pronouns) than native signers to avoid potential ambiguities among referents. Nevertheless, their results did not reach statistical significant for the interaction of discourse status and acquisition group. A similar line of research came from @frederiksen2019 who found a statistically significant difference between both groups, that is native signers and hearing L2 learners of ASL, only for classifier use especially in re-introduced contexts. L2 learners in the study used more nominals and fewer instances of null marking in number, indicative of a limited effect of overexplicitness among second language learners. However, these studies treated late and native deaf signers as a single group. That is, the control group consisted of both native signers who had one or more deaf parents and also native-like [@cormier2012] late deaf signers who had hearing parents and their age of acquisition varied. Given the preliminary evidence that topic continuity and salience marking is also difficult for second language learners regardless of modality, it has been suggested under the Interface Hypothesis that the syntax-discourse interface is vulnerable and especially challenging to acquire [@sorace2011] for second language learners. Similarly, @sorace2009 discuss that this interface entails much more than morphosyntactic knowledge but also includes cognitive and pragmatic resources that help provide discourse coherence. They propose that the second language discourse is over-redundant. This means that second language learners produce more overt pronouns or fuller forms of language for highly accessible referents possibly as a compensatory strategy in an attempt to reduce the increasing load on cognitive processing.

Still, only very few research examined whether linguistically deprived signers (i.e., deaf signers with delayed exposure to a sign language) also display redundancy in their reference tracking. @becker2009 examined 6 DGS signing deaf children (3 native and 3 late)'s narrative abilities. She reported that late signing children generated more explicit forms such as nominals when referring to the characters in the story compared to native learners, concluding that first language deprivation results in an inadequate representation of the addressee and less efficient use of morphosyntax. Comparably, @cormier2013 investigated the use of constructed action in 15 BSL among deaf native and late adult signers and found that late signers were redundant in topic continuity. In other words, the late signers produced more explicit forms like nominals, and fewer instances of null marking when maintaining previous referents. Most relevantly to the present study, @gür2018 studied the development of narrative skills among 10 native and 10 late deaf children and deaf adult signers of TİD. She found that native-signing children and adults had more instances of explicit referent introduction, set the scene more frequently, and employed constructed action in an appropriate manner more than corresponding late-signing children and adults. Although this study presented preliminary findings on age of acquisition effects on narrative competence, it does not investigate how native and late signers employ different linguistic tools based on the accessibility of the referent. Therefore, to our knowledge, there is a lack of research directly examining the effects of first language deprivation on deaf adult signers' selection of REs in signed narratives.

Considering this, it is not very clear whether the negative effects on narrative abilities observed among late signing deaf children hold when they become adults. Understanding the impact of language deprivation or delayed first language learning on narrative competence is crucial for three reasons. First, studying this population will shed light on the possible effect of lack of early language experience on future cognitive and pragmatic abilities since the narrative competence is dependent on such faculties. Second, examining reference tracking strategies in particular will also inform about the linguistic inventory that native and late deaf signers use to refer to entities. Finally, any possible difference between the two acquisition groups in terms of narrative competence will have further practical implications for the importance of early access to a first functional sign language to retain future linguistic and cognitive abilities.

## Present Study

In an attempt at addressing this gap in the literature, the present study aims to investigate the reference tracking strategies of native and late adult signers in Turkish Sign Language (Türk İşaret Dili - TİD) narratives with a story-telling paradigm. We explore the forms available for referent tracking in TİD and examine the nature of RE selection as a possible indicator of pragmatic competence between our acquisition groups.

Overall, considering what has been suggested for the distribution of REs in signed narratives, we expect more explicit forms like nominals for referent introduction and and less explicit forms such as null marking, pronominal and/or classifier use for referent maintainance. For the re-introduction contexts, where referents which are discontinued across one or more clauses are brought back into the discourse, we expect a less clear picture. Some re-introduced referents will have higher accessibility and will thus be tracked with more implicit REs, while others that have lower accessibility will be tracked with more explicit REs Also based on previous research, we hypothesize that late signers will become more over-redundant in their reference tracking compared to native signers. This would translate to a preference by late signers for nominals over null marking especially in maintenance and re-introduction contexts. In contrast, we would expect that native signers use more null marking to track referents.

# Method

## Participants

We tested 29 deaf adult signers, 15 with exposure to TİD from birth (i.e., native signers), 14 with exposure between ages 3 and 17 (i.e., late signers) (*M*\textsubscript{Age of acquisition} = 7.7 years, *SD* = 3.3 years). The criterion to be classified as a late signer was to have hearing and non-signing caregivers. All the signers satisfying this criterion were regarded as late signers \footnote{Note that we are essentially comparing deaf of deaf and deaf of hearing adult signers. Although we did not set a cut-off point for nativeness, all the deaf of hearing participants indicated that they learned TİD after age 3. In order to capture in-group variation among late signers, we run an additional supplementary analysis}. All participants indicated that their preferred language of communication was TİD which they used for most contexts. Detailed information about the linguistic and educational background of the participants can be found in \ref{section:a}.

All participating native signers (7 males, 8 females) were born into deaf families and started acquiring TİD from at least one of their caregivers after birth. The parents of 14 participating native signers were deaf. One native signer had a deaf father and a hearing mother. In addition to TİD, the home language for some native signers (n = 4) in early childhood included the use of home sign\footnote{Home signs are gesture-based communication systems that emerge between the deaf children and their primary caregivers. Although home sign communication may have internal consistency, they are not fully-flegged linguistic systems like natural languages}. The age of the native participants ranged between 18 and 35 (*M* = 26.4 years, *SD* = 5.0 years). This also corresponded to the total average years of TİD use for native signers. All native signers were pre-lingually deaf, that is they were either born deaf or they had become deaf before the age of 3, with the exception of one signer who became deaf between the ages of 4 and 7. Native signers reported to complete at least high school education and their level of education (LOE) was between 12 and 16 years (*M* = 12.7 years, *SD* = 1.5 years). In addition, all native participants attended at least two deaf schools, one being primary and the other middle school. 10 native participants also attended a deaf school for their high school education. Based on a self-reported 5-point scale of proficiency (1: poor, 5: proficient), native participants declared that they have high receptive and productive skills in TİD (*M* = 5 points, *SD* = 0 points). As for their proficiency in (spoken) Turkish, their self-reported rankings were on average 3.7 (*SD* = 0.7), 3.7 (*SD* = 0.7), and 2.6 (*SD* = 1.2) for reading, writing, and speaking abilities, respectively.

Comparably, late learners (8 males, 6 females) were all born into hearing families and started learning TİD later in life after age 3, usually with the start of formal education. The age of sign language acquisition for late signers ranged between years 3-17 (*M* = 8.7 years, *SD* = 3.5 years). The parents of all participating late signers were hearing, and the reported home languages used in childhood mainly consisted of Turkish and home sign. The average age of late signers was 34.6 years (*SD* = 8 years) and ranged between 24 and 50. The derived mean number of years for TİD use following its acquisition was comparable to native signers with 25.9 years (*SD* = 7.9 years). Except for two signers who became deaf between the ages of 4 and 7, all the other late signers were pre-lingually deaf. All late signers completed middle school education and the average year of schooling was 12.3 years (*SD* = 2.8 years). All late participants attended at least one deaf school. While 8 participants attended two deaf schools (either primary and middle school, or middle and high school), 3 participants reported attend three deaf schools. Using the same scale, late signers self-rated their abilities in TİD and (spoken) Turkish. Both of their receptive and productive TİD abilities rated on average 4.9 points (*SD* = 0.3 points). Late signers' mean rating for their reading abilities in Turkish was 3.5 points (*SD* = 0.7 points) while their writing and speaking abilities rated on average 3.4 (*SD* = 0.5 points) and 3.1 points (*SD* = 0.7 points), respectively.

Native and late participants' sex, average years of schooling, TİD use following its acquisition, and their self-reported receptive and productive TİD ratings, and their reading, writing, and speaking ratings for Turkish did not statistically differ; Bayesian *p*'s \> 0.1. However, participants' ages statistically differed between the two acquisition groups; *t*(27) = 3.4, Bayesian *p* \< 0.01. Overall, the native group consisted of younger participants. This was to ensure that the average years of TİD use among native signers approximated those of late signers.

## Stimuli

We used a story-telling paradigm to examine how native and late signers of TİD track reference in signed discourse. Participants were presented with 10 short video clips extracted from the cartoon *Tom and Jerry* and asked to retell them. Each clip had six to ten episodes (*M* = `r mean(c(6, 8, 8, 9, 9, 10, 8, 8, 7, 7))` episodes). The duration of the clips ranged from 17 seconds to 28 seconds (*M* = `r mean(c(19, 23, 28, 24, 22, 26, 17, 22, 23, 23))` seconds). The narratives included on average `r round(mean(c(2, 4, 2, 3, 2, 2, 3, 2, 2, 2)))` recurring animate characters for which the present context required the introduction, maintenance, and re-introduction of referents. In addition, in each narrative, there were on average `r round(mean(c(1, 1, 3, 2, 4, 4, 1, 3, 3, 3)))` inanimate objects with which the animate referents interacted. The detailed description of the episodes in each narrative can be found in \ref{section:b}.

```{r tab-2, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warning=FALSE}
library(magrittr)
library(tidyverse)
library(readr)
library(kableExtra)
library("papaja")
library(stringr)
options(scipen=1, digits=2)

df_table_2 <- read.csv("./appendix/tables/table_tier.csv", sep = ";")
apa_table(df_table_2,
          caption ="Annotation Tiers with their Tags and Descriptions",
          note = "NOM = nominal, PRO = pronominal, CL = classifier, NULL = null anaphora.",
          escape = FALSE,
          align = c("p{3.2cm}", rep("p{3.8cm}", ncol(df_table_2))),
          placement = "p",
          font_size = "footnotesize",
          landscape = FALSE
)
```

## Task Procedure

Participants sat in front of a laptop computer which had the stimulus stories. The instructions were all given by a deaf research assistant who was present in all sessions and ensured that the task requirements were clear. The informed consent form was translated into TİD and explained to the participants by the same assistant before data recording. Participants filled in the consent form and a background questionnaire form before the onset of the trials. Following this step, signers were seated across a hearing researcher who is an advanced late learner of TİD. We instructed the participants to narrate as they would to a deaf friend. All trials were recorded by a digital camera placed in front of the signers. The participants were instructed to watch the narrative clip as many times as they needed, and narrate the content of each clip either to the experimenter or the recording digital camera\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining e did not make it obligatory for participants to either narrate to the digital camera or to the experimenter for few reasons. First, a native or native-like deaf experimenter was not always available, and a hearing late signer of TİD attended some sessions as the experimenter. Therefore, most signers preferred to narrate to the camera. Furhermore, some signers (4 in number) did not feel comfortable narrating to the camera and, as a result, narrated to the deaf research assistant, if available. We ran an additional model excluding the 4 signers who narrated to the experimenter and compared it with the original model including these signers in Appendix. The results imply that there is not much of a difference.}.

```{r echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='asis'}
# generic libraries
library(readxl) #read excel
library(gdata) # reorder.factor
library(magrittr) # pipe
library(tidyverse) # tidy
library(janitor)
library(MASS)
theme_set(theme_bw()) #set ggplot theme
library(extrafont)
loadfonts()

# statistics libraries
library(brms) # bayesian modeling
library(bayesplot) # plot options for bayesian models
bayesplot_theme_set(new = theme_bw()) # set plot theme
library(tidybayes)
library(bayestestR)
theme_apa <- function(base_size = 10, base_family = "serif", box = FALSE) {
  adapted_theme <- ggplot2::theme_bw(base_size, base_family) +
    ggplot2::theme(legend.position = "bottom",
                   plot.title = ggplot2::element_text(size = ggplot2::rel(1.1), margin = ggplot2::margin(0, 0, ggplot2::rel(14), 0))
                   # , axis.title = ggplot2::element_text(size = ggplot2::rel(1.1))
                   , axis.title.x = ggplot2::element_text(size = ggplot2::rel(1), lineheight = ggplot2::rel(1.1), margin = ggplot2::margin(ggplot2::rel(12), 0, 0, 0))
                   , axis.title.y = ggplot2::element_text(size = ggplot2::rel(1), lineheight = ggplot2::rel(1.1), margin = ggplot2::margin(0, ggplot2::rel(12), 0, 0))
                   , axis.title.y.right = ggplot2::element_text(size = ggplot2::rel(1), lineheight = ggplot2::rel(1.1), margin = ggplot2::margin(0, 0, 0, ggplot2::rel(12)))
                   , axis.ticks.length = ggplot2::unit(ggplot2::rel(6), "points")
                   , axis.text = ggplot2::element_text(size = ggplot2::rel(0.9))
                   , axis.text.x = ggplot2::element_text(size = ggplot2::rel(1), margin = ggplot2::margin(ggplot2::rel(6), 0, 0, 0))
                   , axis.text.y = ggplot2::element_text(size = ggplot2::rel(1), margin = ggplot2::margin(0, ggplot2::rel(6), 0, 0))
                   , axis.text.y.right = ggplot2::element_text(size = ggplot2::rel(1), margin = ggplot2::margin(0, 0, 0, ggplot2::rel(6)))
                   , axis.line = ggplot2::element_line()
                   # , axis.line.x = ggplot2::element_line()
                   # , axis.line.y = ggplot2::element_line()
                   , legend.title = ggplot2::element_text()
                   , legend.key = ggplot2::element_rect(fill = NA, color = NA)
                   , legend.key.width = ggplot2::unit(ggplot2::rel(20), "points")
                   , legend.key.height = ggplot2::unit(ggplot2::rel(20), "points")
                   , legend.margin = ggplot2::margin(
                     t = ggplot2::rel(16)
                     , r = ggplot2::rel(16)
                     , b = ggplot2::rel(16)
                     , l = ggplot2::rel(16)
                     , unit = "points"
                   )
                   , panel.spacing = ggplot2::unit(ggplot2::rel(14), "points")
                   , panel.grid.major.x = ggplot2::element_blank()
                   , panel.grid.minor.x = ggplot2::element_blank()
                   , panel.grid.major.y = ggplot2::element_blank()
                   , panel.grid.minor.y = ggplot2::element_blank()
                   , strip.background = ggplot2::element_rect(fill = NA, color = NA)
                   , strip.text.x = ggplot2::element_text(size = ggplot2::rel(1.2), margin = ggplot2::margin(0, 0, ggplot2::rel(10), 0))
                   , strip.text.y = ggplot2::element_text(size = ggplot2::rel(1.2), margin = ggplot2::margin(0, 0, 0, ggplot2::rel(10)))
    )
  if(box) {
    adapted_theme <- adapted_theme + ggplot2::theme(panel.border = ggplot2::element_rect(color = "black"))
  } else {
    adapted_theme <- adapted_theme + ggplot2::theme(panel.border = ggplot2::element_blank())
  }
  adapted_theme
}

# Read df and drop NAs
df_narration <- read.csv("data/narration_data.csv", sep = ";") %>% tidyr::drop_na()

df_misc <- read.csv("./appendix/appendix_a_table.csv", sep=";") %>% clean_names() %>% 
  dplyr::mutate(Subject = as.integer(no),
                Group = case_when(aoa == "0-3" ~ "native",
                                  aoa == "4-7" ~ "late1",
                                  T ~ "late2") %>% as.factor()) %>% dplyr::select(Subject, Group)
df_narration %<>% left_join(df_misc)

# Coding nominal, Zero, Classifier, and Pronominal categories
df_narration$ref_type <- with(df_narration, case_when(
  GTR1 %in% c("Bare noun", "Point noun", "Noun Point",
              "Fingerspelled noun", "CL noun", "Noun CL") ~ "Nominal",
  GTR1 %in% c("Constructed action", "Constructed Action", "Plain verb", "Person agreement verb") ~ "Null",
  GTR1 %in%  c("Whole entity CL", "Body part CL",
               "SASS - Tracing", "SASS - Static") ~ "Classifier",
  T ~ "Pronominal"))

# Coding nominal, Zero, Classifier, and Pronominal categories
df_narration$GTR1_plot <- with(df_narration, case_when(
  GTR1 %in% c("Bare noun") ~ "BARE NOM",
  GTR1 %in% c("Point noun", "Noun Point","CL noun", "Noun CL") ~ "M NOM",
  GTR1 %in% c("Point noun", "Noun Point","CL noun", "Noun CL") ~ "FS NOM",
  GTR1 %in% c("Constructed action", "Constructed Action") ~ "CA",
  GTR1 %in% c("Plain verb") ~ "PLAIN",
  GTR1 %in% c("Person agreement verb") ~ "AGR",
  GTR1 %in% c("Whole entity CL") ~ "WECL",
  GTR1 %in%  c("Body part CL")  ~ "BPCL",
  GTR1 %in%  c("SASS - Tracing", "SASS - Static")  ~ "SASS",
  T ~ "Pronominal"))

# Encode vector types
df_narration$Narrative %<>% as.integer() 
df_narration$Subject %<>% as.integer()
df_narration$GTR1 %<>% as.factor() 
df_narration$Nativeness %<>% dplyr::recode(`0` = "Late", `1` = "Native") %>% as.factor() %>% reorder.factor(new.order = c("Native","Late"))
df_narration$Discourse %<>% as.factor()
df_narration$ref_type %<>% as.factor()

# Summary of the data
summary_df <- df_narration %>% 
  group_by(Nativeness, ref_type) %>% 
  summarise(Count = n()) %>% 
  mutate(Percentage = formattable::percent(Count/sum(Count))) %>% 
  arrange(desc(Percentage))

# Descriptive plots
df_narration$ref_type2 <- df_narration$ref_type %>% recode('Classifier' = 'CL', 'Null' = 'NULL',
                                                           'Pronominal' = 'PRO', 'Nominal' = 'NOM')
# dropping Pronominal references because they are too few.
df_model1 <- df_narration %>% subset(ref_type !="Pronominal") %>% drop.levels()

# Contrast coding
df_model1$Nativeness %<>% reorder.factor(new.order = c("Native","Late"))
contrasts(df_model1$Nativeness) <- contr.sum(2)/2
contrasts(df_model1$Discourse) <- contr.sum(3)/2
contrasts(df_model1$Group) <- contr.sdif(3)

# counts of ref types
ref_count <- df_narration %>% group_by(ref_type) %>% summarise(count=n())

# Model for reference type
model_1 <- brm(ref_type ~ Nativeness*Discourse + (1|Subject), 
               data = df_model1, family = categorical(link = "logit"), 
               iter = 3000, warmup = 2000, chains = 4, cores = 4, file ="./rds/model_1.rds")

# Posterior plot for reference type
df_model_1 <- mcmc_intervals_data(model_1, pars = vars(starts_with("b_")))
df_model_1$parameter %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
                                        `b_muNull_Intercept` = "Null_Intercept",
                                        `b_muNominal_Nativeness1` = "Nom_Native",
                                        `b_muNull_Nativeness1` = "Null_Native",
                                        `b_muNominal_Discourse1` = "Nom_Intro",
                                        `b_muNominal_Discourse2` = "Nom_Maintain",
                                        'b_muNominal_Nativeness1:Discourse1' = 'Nom_Native*Intro',
                                        'b_muNominal_Nativeness1:Discourse2' = 'Nom_Native*Maintain',
                                        'b_muNull_Nativeness1:Discourse1' = 'Null_Native*Intro',
                                        'b_muNull_Nativeness1:Discourse2' = 'Null_Native*Maintain',
                                        `b_muNull_Discourse1` = "Null_Intro",
                                        `b_muNull_Discourse2` = "Null_Maintain")
saveRDS(df_model_1, "./rds/df_model_1.rds")

# Supplementary model with 3 groups of acquisition
model_3 <- brm(ref_type ~ Group*Discourse + (1|Subject), 
               data = df_model1, family = categorical(link = "logit"), 
               iter = 3000, warmup = 2000, chains = 4, cores = 4, file ="./rds/model_3.rds")

# Posterior plot for reference type
df_model_3 <- mcmc_intervals_data(model_3, pars = vars(starts_with("b_")))
df_model_3$parameter %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
                                        `b_muNull_Intercept` = "Null_Intercept",
                                        `b_muNominal_Group2M1` = "Nom_Late1vsNative",
                                        `b_muNominal_Group3M2` = "Nom_Late2vsLate1",
                                        `b_muNominal_Discourse1` = "Nom_Intro",
                                        `b_muNominal_Discourse2` = "Nom_Maintain",
                                        `b_muNominal_Group2M1:Discourse1` = 'Nom_L1vsN*Intro',
                                        `b_muNominal_Group3M2:Discourse1` = 'Nom_L2vsL1*Intro',
                                        `b_muNominal_Group2M1:Discourse2` = 'Nom_L1vsN*Maintain',
                                        `b_muNominal_Group3M2:Discourse2` = 'Nom_L2vsL1*Maintain',
                                        `b_muNull_Group2M1` = "Null_Late1vsNative",
                                        `b_muNull_Group3M2` = "Null_Late2vsLate1",
                                        `b_muNull_Discourse1` = "Null_Intro",
                                        `b_muNull_Discourse2` = "Null_Maintain",
                                        `b_muNull_Group2M1:Discourse1` = 'Null_L1vsN*Intro',
                                        `b_muNull_Group3M2:Discourse1` = 'Null_L2vsL1*Intro',
                                        `b_muNull_Group2M1:Discourse2` = 'Null_L1vsN*Maintain',
                                        `b_muNull_Group3M2:Discourse2` = 'Null_L2vsL1*Maintain')
saveRDS(df_model_3, "./rds/df_model_3.rds")
#df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "NULL"))

```

## Data Annotation

We coded the discourse status (Introduction, Maintenance, Re-introduction) and RE type (Nominal, Pronominal, Classifier, Null). We used ELAN Linguistic Annotation Software [@crasborn2008] for data annotation. First Author who is a hearing signer of TİD annotated all the data and translated the narratives into Turkish together with a native deaf signer of TİD.

Following @berman2013relating, we first identified the clause boundaries in each sentence indicated by the presence of predicates and informative non-manual markers (e.g., head nod). On the second tier, we coded the discourse status for each referent. Having adapted the classification of @gullberg2006, we considered all referents mentioned for the first time as introduced regardless of their syntactic function in the sentence. Only subjects were coded for maintenance and re-introduction. We coded subject referents as maintained only if they had been referred to in the previous clause. To this end, the referents which had been discontinued across one or more sentences were all regarded re-introduced into the discourse. On a third tier, we had the RE type for a single reference with the main categories nominal, pronominal, null marking, and classifier. Under nominal, we identified bare nouns, modified nouns, and fingerspelled nouns. For classifiers, we distinguished among three types of constructions: WECL, BPCL, and ExtCL. Finally, we had CA, agreement, plain tags for null constructions. We were unable to observe any instance of HCLs occurring in isolation (i.e., not part of constructed action). The annotation tiers are presented in Table \ref{tab:tab-2}.

## Data Analysis

In our analysis, we use Bayesian regression models with the help of the brms package [@bürkner2018] in R. In a Bayesian regression the prior and the likelihood is used to calculate the posterior. Instead of a single filter of significance, we get a posterior probability distribution. We do not define our own priors and use the brms package defaults. We are using Bayesian regression to avoid some common pitfalls of other approaches [@wagenmakers2007; @kruschke2011].

For this particular type of experimental data, we fit the responses to a multinomial distribution. Most of the time, experimental results have yes/no-true/false responses which are fit to binomial distributions, reading/response times which are fit to log-normal distributions, and ratings which are fit to cumulative distributions. In these cases, the researcher is not particularly interested in the \`intercept' coefficient which shows the mean value of the response. The researcher is usually interested in the remaining coefficients which give information about predictor level comparisons.

In a multinomial distribution, the responses are in a categorical relationship where the levels can be A,B,C, and more. These levels are not ordered in time, they are not cumulative, and they are not additive. If we were to make an analogy, they are like countries on Earth which can have differences and similarities, but they don't have to. In a multinomial distribution, one of the responses becomes the level of comparison and other levels are compared against it. In a way, there are multiple binomial distributions. However, the intercepts in a multinomial distribution end up having more of an interest for the researcher. This is because the intercept is a comparison of two levels of the response to the exclusion of the remaining response levels. There is another consideration that should go into interpreting the results of a multinomial distribution. Different from the intercepts, the remaining coefficients are predictor level comparisons within the grand comparison of the response levels. To indicate which predictor level comparison belongs to which response comparison, we name our coefficients with prefixes followed by an underscore.

In our model plots, we report median estimates as points, 50% credible intervals as a thick line and 95% credible intervals as a thin line. These numbers are relatively arbitrary, and they are intended to show the posterior probability distribution of the coefficients. We use interval plots to save space and have a more compact plot. If a considerable amount (\>90%) of the posterior probability distribution is towards a sign (-/+) we interpret it as a categorial effect, if a good amount (\>50%) of the distribution is towards a sign (-/+) we interpret it as a tendency, and we see the remaining distributions as indifference. In addition to these, we look at how the distribution skewed in the extremes or how \`wide' it is from the median estimate to inform us mainly about uncertainty and variation.

# Results

We collected `r nrow(df_narration)` data points as a result of the experiment. Each data point is a referent tracked by the participant. Each of the REs is classified in terms of its Discourse Status (Introduction, Maintenance, Re-introduction), its Type (Nominal, Pronominal, Classifer, Null)\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer Null). However, we allude to the numerical distribution of the subtypes in this and following sections}, and Participant Group (Native, Late)\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17). Treating age of acquisition as a continous variable would not be informative due to the small sample size in each interval.}.

Across our acquisition groups, we observed `r sum(df_narration$Discourse == "Introduction")` (`r sum(df_narration$Discourse == "Introduction")/nrow(df_narration)`) instances of referent introduction, `r sum(df_narration$Discourse == "Maintenance")` (`r sum(df_narration$Discourse == "Maintenance")/nrow(df_narration)`) instances of referent maintenance, and `r sum(df_narration$Discourse == "Re-introduction")` (`r sum(df_narration$Discourse == "Re-introduction")/nrow(df_narration)`) instances of referent re-introduction. For the introduction of a referent, the majority of the REs were nominal constructions (Table \ref{tab:my-table}). However, we observed a sporadic use of classifiers and null marking respectively. When the discourse status was maintained, signers extensively used null markers to refer to the entities in the immediately previous clause. Classifier constructions followed null markers in frequency in maintained contexts. We observed only a few instances of nominal use for referent maintenance. In re-introduction contexts, signers used a greater proportion of nominal markings which were followed by null, classifier, and pronominal use.

```{=tex}
\begin{table}
\small
\caption{Distribution of all discourse status by referring expression types}
\label{tab:my-table}
\begin{tabular}{llll}
\hline
                     & Introduced    & Maintained     & Re-introduced \\ \hline
\textit{Nominal:}    & 92.78\% (989) & 5.63\% (93)    & 63.65\% (907) \\
\textit{Pronominal:} & 0.09\% (1)    & 0.24\% (4)     & 2.46\% (35)   \\
\textit{Classifier:} & 5.35\% (57)   & 18.29\% (302)  & 8\% (114)     \\
\textit{Null:}       & 1.78\% (19)   & 75.83\% (1252) & 25.89\% (369) \\ \hline
\end{tabular}%
\end{table}
```
Table \ref{tab:my-table2} shows that for all discourse statuses the most numerous nominal type was \textsc{bare nom} which was followed by nouns modified by index or classifiers. In number, signers used \textsc{mod nom} more in introduction and re-introduction compared to maintenance settings. Signers produced a total number of `r sum(df_narration$ref_type == "Pronominal")` pronouns (Table \ref{tab:my-table3}). A great majority of the pronominal constructions were used to re-introduce a referent into the discourse. We observed only a very small amount of pronoun use for referent introduction and maintenance.

```{=tex}
\begin{table}
\small
\caption{Distribution of nominal referring expressions by discourse status}
\label{tab:my-table2}
\begin{tabular}{lllll}
\hline
                        & Bare Noun  & Modified Noun & FS Noun \\ \hline
\textit{Introduced:}    & 48.93\% (842) & 53.56\% (139)     & 100\% (8) \\
\textit{Maintained:}    & 4.82\% (83)  & 3.85\% (10)      & 0\% (0) \\
\textit{Re-introduced:} & 46.25\% (796) & 42.69\% (111)     & 0\% (0) \\ \hline
\end{tabular}
\end{table}

\begin{table}
\small
\caption{Distribution of pronominal referring expressions by discourse status}
\label{tab:my-table3}
\begin{tabular}{ll}
\hline
                        & Pronominal  \\ \hline
\textit{Introduced:}    & 2.5\% (1)   \\
\textit{Maintained:}    & 10\% (4)    \\
\textit{Re-introduced:} & 87.5\% (35) \\ \hline
\end{tabular}
\end{table}
```
As for the classifier constructions (Table \ref{tab:my-table4}), WECL was the most common type, followed by BPCL and ExtCL. For the first mention of a referent, signers mostly preferred to use ExtCL. However, for referent maintenance and re-introduction WECL was used more compared to BPCL and ExtCL.

Among the null markers, CA was most frequently employed, and it was followed by plain verbs and agreement verbs (Table \ref{tab:my-table5}). Overall, signers used very few null markings for the introduction of a referent and the majority of the existing ones were CA. Additionally, signers used CA more for referent re-introduction than maintenance. We observed a similar instances of agreement verbs for maintained and re-introduced contexts. In contrast, plain verbs were favored more in the maintenance setting than re-introduction.

```{=tex}
\begin{table}
\small
\caption{Distribution of classifier referring expressions by discourse status}
\label{tab:my-table4}
\begin{tabular}{llll}
\hline
                        & WECL          & BPCL        & ExtCL         \\ \hline
\textit{Introduced:}    & 2.14\% (6)    & 2.4\% (3)   & 71.64\% (48) \\
\textit{Maintained:}    & 67.26\% (189) & 77.6\% (97) & 23.88\% (16) \\
\textit{Re-introduced:} & 30.6\% (86)   & 20\% (25)   & 4.48\% (3)   \\ \hline
\end{tabular}
\end{table}
```
Across the discourse statuses, there is also a difference between participant groups in the use of null marking for reference tracking. Native signers of TİD used it more (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "Native") / nrow(df_narration), digits=2)`) compared to the late signers (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "Late") / nrow(df_narration), digits=2)`). Comparably, late signers used nominals more frequently (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "Late") / nrow(df_narration), digits=2)`) than native signers (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "Native") / nrow(df_narration), digits=2)`). Figure 1 shows the numerical distribution of the responses by RE type and Group.
```{=tex}
\begin{table}
\small
\caption{Distribution of null referring expressions by discourse status}
\label{tab:my-table5}
\begin{tabular}{llll}
\hline
                        & Constructed Action & Plain Verb    & Agreement Verb \\ \hline
\textit{Introduced:}    & 1.58\% (17)        & 0.28\% (1)    & 0.5\% (1)      \\
\textit{Maintained:}    & 71.64\% (773)      & 89.97\% (323) & 77.23\% (156)  \\
\textit{Re-introduced:} & 26.78\% (289)      & 9.75\% (35)   & 22.28\% (45)   \\ \hline
\end{tabular}
\end{table}
```

```{r fig:count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), RE Type (Nominal, Pronominal, Classifier, Null), and Participant Group (Native, Late)"}
df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "NULL"))

p1 <- df_narration %>%
  dplyr::select(Nativeness, ref_type2, Discourse) %>%
  group_by(Nativeness, Discourse) %>%
  ggplot(aes(ref_type2, fill = Nativeness)) + geom_bar(position = "dodge") +
  stat_count(geom = "text", aes(color = Nativeness, label = ..count..), vjust = 0,
             position = position_dodge(width = 0.85), size = 2) +
  facet_grid(~Discourse) + ylab("Count") + xlab("REATs") +
  scale_color_grey() +
  scale_fill_grey() +
  theme_apa() 
p1
```

For more inference on the data and having pairwise comparisons, we fit a regression model to RE Type using Discourse Status and Group as predictors. We used the brms package [@bürkner2018] in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late). As the overall number of produced pronouns was very little (40 data points), we excluded pronominal use from the following analyses. For the current analysis, we did not define priors ourselves but used the preset uninformed priors. Figure 2[^2] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals. There are 3 responses we are considering in this model, and they are Classifier, Nominative, and Null. Classifier response is the level that other responses Nominative and Null are compared against.

[^2]: We ran another model where we used the Nativeness predictor with 3 levels 0-3:Native, 4-7:Late1, and 7-above:Late2. However, our model did not converg properly. We provide the model results as a plot and table in the Appendix.

```{r fig:model-gram, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
p3 <- df_model_1 %>% 
  ggplot(aes(m, y=factor(parameter,levels = rev(levels(factor(parameter)))))) +
  geom_point(size = 2.5) +
  geom_errorbarh(aes(xmin = l, xmax = h), alpha = 1, height = 0, size = 1.5) +
  geom_errorbarh(aes(xmin = ll, xmax = hh, height = 0)) + vline_0(linetype = "dashed") +
  xlab("Estimate (log odds)") + ylab("Coefficients") + theme_bw() + theme_apa()
p3
```

In a multinomial distribution, the model takes one of the responses as a reference and compares all other responses to it. For example, the first coefficient, Nom_Intercept (nominal intercept), in Figure 2 is the comparison of a nominal versus a classifier. It means that participants produced more nominals than classifiers in general regardless of participant group and discourse status. The second coefficient shows that null markings were also used more than classifiers by the participants in general, but not as much as nominals.

In the comparison of nominals and classifiers, being a native acquirer of TİD did not have an effect. Nominals were used more when the discourse status was introduction compared to re-introduction. Again, the whole comparison is made against a classifier. In contrast to the introduction status, nominals were used less in maintenance contexts compared to the re-introduction contexts within the greater comparison to classifier referents. There does not seem to be an interaction between the introduction status of discourse and being a native acquirer of TİD. However, there seems to be an interaction between maintaining a referent and being a native acquirer of TİD. In that, there were fewer nominal referents when it was a native acquirer of TİD and the discourse context was maintenance.

Compared to classifiers, being a native acquirer of TİD increased the odds of using null marking in general. Null constructions were used less in introduction contexts than in re-introduction. This contrasts with the nominals (4th coefficient) as they were used more in referent introduction versus re-introduction. In contrast to introduction contexts, null markers were used more in referent maintenance compared to re-introduction, yet again within the greater comparison to the classifiers. There seems to be an interaction between introducing a referent and being a native acquirer of TİD. In that, the odds of using null marking increases if the participant is a native acquirer of TİD and the referent is used in an introduction context. There is another interaction with regards to referent maintenance, contrasting the first interaction. In that, the odds of using null markers decrease if the participant is a native acquirer of TİD and the referent is maintained. This does not mean that late signers used more null markers in maintenance contexts. It means that the use of null markers did not linearly increase as the addition of using a null marker instead of a classifier as a native plus using null markers instead of a classifier for maintaining a referent versus re-introducing it.

# General Discussion

In our study, we aimed to (i) describe how signers track referents in signed discourse in TİD and (ii) examine whether language deprivation results in more over-redundant referent selection. To this end, we investigated the distribution of referring expressions (REs) that the native and late signers select when employing referents in different discourse statuses (introduced, maintained, and re-introduced).

In our first set of hypotheses regarding referent tracking strategies, we had the general expectation that signers would use more explicit forms (mainly nominals) for referent introduction and less implicit forms such as null marking, pronominal use, and classifiers. In re-introduced contexts, we predicted that signers would use both explicit and implicit markers depending on the accessibility of the tracked referent.

For age of acquisition effects, based on what is proposed for child and adult deaf signers with language deprivation in terms of their morphosyntactic [@boudreault2006; @cormier2012; @kayaba2022] and narrative development [@becker2009; @cormier2013], we expected that late signers would be more over-redundant in their referent tracking, akin to second language learners [@frederiksen2019; @bel2015]. That is, we had the expectation that late signers would use more explicit markers like nouns and fewer implicit markers like null marking even for highly accessible referents (when the discourse status is maintained or even re-introduced in same cases) than the control group (i.e., native signers).

## Tracking referents in TİD

The results of the present study partly support our hypotheses. In line with our expectations, regardless of their acquisition groups (native or late) (i) signers mainly used nominals to introduce a referent into the discourse, (ii) used null marking and classifiers but not pronouns for maintained contexts, and (iii) employed both nominals and less explicit markers (classifiers, null markers and pronouns) to re-introduce a referent.

However, the proportion of the observed discourse statuses (26% for introduction, 40% for maintenance, 34% for re-introduction across acquisition groups) in this study differed from that of some other studies. For a comparison, while the percentage of introduced referents were similar in both research (26% versus 24%), the participants in the present study brought back referents into the discourse more frequently compared to the participants in @frederiksen2019 and @frederiksen2016 which reported that re-introduction contexts accounted for only 7% of the referents in ASL. On the other hand, maintenance percentages were lower in the present study compared to @frederiksen2016 (40% versus 69%). Investigating the distribution of REs across discourse statuses in ASL, a similar study by @czubek2017 also reported contrasting numbers (17% for introduction, 64% for maintenance, and 19% for re-introduction). We suggest that the discrepancy might be due to the nature of the stimuli used in these research. In @frederiksen2016, the authors used 4-6-episode-long picture story with 2 animate characters and 1 inanimate object. In contrast, the stimuli in the present study consisted of 8-episode-long video clips on average with a mean number of 2 animate characters and 3 inanimate objects. Although the number of animate characters and inanimate characters in @czubek2017's stimuli matched that of the present study, their narrative structure (a 6-episode-long picture story) was still shorter compared to our narrative stimuli. We believe fewer possible referents and/or a simpler narrative structure in these studies may have contributed to the lower number of re-introduced contexts compared to the present study. Due to the story length and composition in this study, the participants had to track more recurring animate or inanimate referents for a longer period of time. As a result, our signers engaged in referent re-introduction more than what was reported in previous studies [@czubek2017; @frederiksen2016].

As for the distribution of REs across different discourses statuses, we observed that signers overwhelmingly preferred to use nominals over other REs for the first mention of a referent and used null markers such as CA, and agreement and plain verb for referents with high accessibility (i.e., when the discourse status was maintained). This finding parallels other accounts of referential accessibility suggested for other sign languages [see @frederiksen2016; @czubek2017; and @swabey2002; for ASL, @morgan2005; for BSL @bel2015; and @barberà2009 for LSC. Our results also indicate that modified nouns (e.g., nouns that are preceded or followed by a classifier or IX) appeared more in introduced contexts compared to re-introduced contexts. This brings along the assumption that modified nouns are used more to track referents with lower accessibility, probably due to their more explicit structure.

Interestingly and to our surprise, the signers in this study used pronominal IX very sporadically with only 40 observed instances that were mainly used to re-introduce referents rather than maintain them. This finding is comparable with what has been reported for ASL [@swabey2002; @frederiksen2016; @czubek2017] but not with LSC [@bel2015], BSL [@morgan2005] and DGS [@perniss2015], in which the signers extensively made overt pronominal references (up to 30%) to maintain entities. Comparable to our findings, It has been reported that pronouns do not frequently occur in the local context and instead follows a topic shift and thus rank higher than classifiers and null markers different genre [see @barberà2009 for LSC; and @nuhbalaoglu2018 for TİD]. Considering this, the claim that the IX sign functions like demonstratives that are also used to track referents in re-introduced contexts has not been borne out [@koulidobrova2016]. Similar to these research, our findings indicate that the pronominal IX in TİD was used more concurrently with nominal constructions and pronominal IX was used very infrequently in isolation. The observation that overt pronouns are rare in ASL narratives comes from studies that include a simple picture retelling task with one protagonist [e.g., @frederiksen2016] as well as a more complex narrative design as in [@czubek2017]. However, @czubek2017's study also incorporated different genres and reported more instances of overt pronouns in personal narratives (compared to picture or video retelling). In fact, this finding suggests that the use of pronouns might be genre-dependent. Moreover, another production study on referent tracking in ASL reported object preference for pronominals [@frederiksen2017]. Nonetheless, we only coded subjects for maintained and re-introduced referents. We argue that all of these factors (genre-dependency as well as object preference) might contribute to the explanation of why overt pronouns in this data are dispreferred in favor of other markers for highly accessible referents such as null marking and classifiers.

A deeper look into the allocation of classifier constructions revealed that WECLs were the most common type of classifiers for the maintenance of a referent followed by BPCLs and ExtCLs. We observed a similar alignment for re-introduced contexts. In contrast, a great number of the overall ExtCLs were used to introduce a new referent into the discourse. That is to say, whereas WECLs were the preferred classifier types to signal highly accessible referents, ExtCLs constructions were mainly used for referents which were not easily accessible. We suggest that the less iconic and transparent nature of WECLs allowed signers to continue a previous referent without being redundant. As pointed out by @czubek2017, signers prefer to rely on ExtCLs for the first mention of an entity because these constructions are informative and explicit by nature, and they provide visual and descriptive cues about an entity and its location in space. Nevertheless, we only observed `r sum(df_narration$GTR1_plot == "SASS" & df_narration$Discourse == "Re-introduction")` instances of ExtCLs for re-introduced contexts, replicating @frederiksen2019's findings for ASL.

As for the distribution of the null markers, approximately 90% of the plain verbs were used to maintain a previous referent, making these constructions least implicit REs used for TİD[^3]. Agreement verbs and CA were the following implicit markers of reference, `r sum(df_narration$GTR1 == "Person agreement verb" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1 == "Person agreement verb")` and `r sum(df_narration$GTR1_plot == "CA" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1_plot == "CA")` of which were used for maintained contexts, respectively. This contrasts with some of the research [@czubek2017] in which agreement verbs ranked lowest on scale, indicating that these constructions were more implicit compared to plain verbs. However, there seems to be a common ground for the position of CA markers which have been claimed to rank higher than other null markers as in the present study [@frederiksen2016; @czubek2017; @barberà2009]. In Figure \ref{fig:scale}, we report the distribution of REs in TİD as a function of referent accessibility and compare our accessibility hierarchy with that of [@frederiksen2016] for ASL. In the hierarchies, the left most REs (e.g., modified and bare nouns for TİD) are used when the accessibility is low. If the referent is already active in the addressee's mind and its accessibility is high, then the right most markers (e.g., agreement and plain verbs for TİD) are selected.

[^3]: As in other sign languages, in TİD, the lexical meaning of signs has an effect on the morphological grouping. For instance, usually (di)transitive agreement verbs denote some sort of physical or metaphorical transfer of an object or abstract notion (Meir 1999). As for plain verbs, both phonology of a sign and its semantics have an effect on morphology. For instance, \textsc{eat} in TİD is signed nearby the mouth and since it needs to remain in that phonological location, it is not suitable for agreement and thus it is regarded as a plain verb. Likewise the experiencer verb \textsc{love} is signed on the torso and as such it does not take part in agreement. These generalizations noted we should also be cautious in labelling a verb as categorially one type or the other. For instance, some intransitive plain verbs like \textsc{fall} or \textsc{jump} in TİD may display spatial person agreement with the single argument in the clause (Gökgöz and Sevgi, 2020). See also for a similar discussion in Brazilian Sign Language.}.

```{=tex}
\begin{figure}
\caption{Comparing the present study and Frederikse and Mayberry (2016) in terms of proposed accessibility hierarchies}
\label{fig:scale}
\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,270); %set diagram left start at 0, and has height of 270

%Straight Lines [id:da19719799699706264] 
\draw [fill={rgb, 255:red, 175; green, 40; blue, 40 }  ,fill opacity=1 ][line width=1.5]    (114.9,108.91) -- (135.9,108.91) -- (253.78,108.53) -- (495.74,108.99) (164.88,102.31) -- (164.92,115.31)(214.88,102.16) -- (214.92,115.16)(264.91,102.06) -- (264.89,115.06)(314.91,102.15) -- (314.89,115.15)(364.91,102.24) -- (364.89,115.24)(414.91,102.34) -- (414.89,115.34)(464.91,102.43) -- (464.89,115.43) ;
\draw [shift={(499.74,109)}, rotate = 180.11] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;
\draw [shift={(110.9,108.91)}, rotate = 0] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;

% Text Node
\draw (498,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{High}};
% Text Node
\draw (81,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{Low}};
% Text Node
\draw (150,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize M NOM}};
% Text Node
\draw (191,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize BARE NOM}};
% Text Node
\draw (250,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize ExtCL}};
% Text Node
\draw (285,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize WECL \& BPCL}};
% Text Node
\draw (359,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize CA}};
% Text Node
\draw (405,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize AGR}};
% Text Node
\draw (452,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize PLAIN}};
\end{tikzpicture}

\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt        
\begin{minipage}{\textwidth}
    \footnotesize
    \centering
    Accessibility hierarchy in the present study for TİD
    \end{minipage}

\vspace{1 cm}

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,270); %set diagram left start at 0, and has height of 270

%Straight Lines [id:da19719799699706264] 
\draw [fill={rgb, 255:red, 175; green, 40; blue, 40 }  ,fill opacity=1 ][line width=1.5]    (114.9,108.91) -- (135.9,108.91) -- (253.78,108.53) -- (495.74,108.99) (164.88,102.31) -- (164.92,115.31)(214.88,102.16)(264.91,102.06) -- (264.89,115.06)(314.91,102.15)(364.91,102.24) -- (364.89,115.24)(414.91,102.34)(464.91,102.43) -- (464.89,115.43) ;
\draw [shift={(499.74,109)}, rotate = 180.11] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;
\draw [shift={(110.9,108.91)}, rotate = 0] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;

% Text Node
\draw (498,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{High}};
% Text Node
\draw (81,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{Low}};
% Text Node
\draw (150,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize NOM}};
% Text Node
\draw (250,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize ExtCL}};
% Text Node% Text Node
\draw (333,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize AGR \& CA}};
% Text Node
\draw (445,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize PLAIN}};
\end{tikzpicture}
\begin{minipage}{\textwidth}
    \footnotesize
    \centering
    Accessibility hierarchy in Frederiksen and Mayberry (2016) for ASL
    \end{minipage}
\end{figure}
```
Both hierarchies illustrated in Figure \ref{fig:scale} consider nominals and ExtCLs as markers of lowly accessibly referents. Likewise, in both hierarchies, as markers of highly accessible referents, different types of null markers are further distinguished. Plain verbs rank the lowest on the scales, preceded by agreement verbs and constructed action. On a similar ground, both scales lack pronouns, which occurred very infrequently in the data. Although @frederiksen2016 did not include other classifier types in the hierarchy, we proposed that WECLs and BPCLs must ranker higher than null markers but lower than ExtCLs that were mainly used by interlocutors to introduce a referent into the discourse context.

## Age of acquisition effects on referent tracking

The results in this study only somewhat support our second set of hypotheses that we set for the effects of age of acquisition on participants' RE selection. That is, while being a native acquirer of TİD increased the odds of using null marking in reference tracking across different discourse statuses, we failed to find such a difference for overall nominal production. Taking a closer look at the interaction terms, we observed that native signers, compared to late signers, were less likely to use nominal constructions when the discourse status was maintained. Native acquisition further increased the odds of null markers when the discourse status was introduced. However, we found the opposite alignment for the interaction of maintenance and native acquisition. That is, the coefficient for this interaction term slightly decreased the combined main effects of having a maintained context and being a native signer. This means that these main effects were greater than the interaction or added effects, pointing out to the implication that the type of discourse status (introduced, maintained, re-introduced) accounts for the distribution of REs in TİD narratives better than age of acquisition (native, late) which, in comparison, had a less salient effect.

While late signers were only partially over-redundant in maintained contexts, they were more so in introduced contexts. In line with our study, a tendency for limited over-redundancy has been suggested for late signing children in DGS [@becker2009] and late signing adults in BSL specifically for the use of constructed action [@cormier2013]. As for TİD, to our knowledge, there is only one study [@gür2018] that directly aimed at examining age of acquisition effects on adult signers' narrative skills. @gür2018 observed that late signing deaf adults and children set the scene less frequently and used constructed action less accurately compared to their native counterparts. However, this particular investigation did not present any data on the distribution of REs between two acquisition groups. In the present study, the finding that native signers preferred to use null arguments even for the first mention of a referent (`r sum(df_narration$ref_type == "Null" & df_narration$Discourse == "Introduction")` in total) is surprising given some of the previous literature in the topicality and saliency framework [@cormier2013; @swabey2002]. These research reported very little or no use of null marking by native signers for introduced contexts.

Given the paucity of research directed at the late signing group, we now turn to the referent tracking strategies observed for second language learners of a sign language. Our study replicated some of the findings of the line of research that investigated deaf or hearing bimodal bilingual signers [@bel2015; @frederiksen2019].\footnote{For similar results with heritage learners of sign language, see Reynolds (2018).} The authors reported that more instances of null marking in the context of referent introduction was attested for native signers than deaf or hearing bimodal bilingual signers. Although using null marking to introduce a referent into the discourse would normally result in referential ambiguity, it is important to note that our stimuli consisted of 10 interconnected narratives with 2 or more recurring characters. It may be possible that native signers, having already introduced the referent with an explicit nominal marker in a different discourse context, found it felicitous to refer to the same entity with more implicit markers in the following narratives even for introduced contexts. Then, our results imply a difference in pragmatic sensitivity to referential accessibility between native and late signers, which is more robustly reflected in introduced contexts.

Since we only observed a small difference between the two acquisition groups for the use of certain REs, the aforementioned findings allude to the implication that the effect on language deprivation on pragmatic factors should also be considered in different contexts. For one, both groups of signers can be tested with different genres of narratives and with a more complex story design. In order to capture possible variation in the accessibility levels of re-introduced referents, a more fine-tuned protocol [as in @toole1996] can be used for measuring referent distance and the number of other possible competitors. Taken altogether, our results accentuates the necessity of engaging in signed narrative production early in life as well as the importance of early intervention for late signing deaf children.

# Conclusion

The present study aimed to investigate the reference tracking strategies of native and late deaf adult signers in TİD narratives. In line with our expectations, we found that the distribution of REs (nominal, pronominal, classifier, and null marking) followed from previous accounts of saliency and referential accessibility. For introduced contexts, signers produced more nominals and ExtCLs, and to maintain a previous referent, hey mainly used null arguments (plain verbs, agreement verbs and constructed action) and certain classifiers (e.g., BPCL and WECL) to track highly accessible referents. We observed only very few instances of pronouns (mostly produced by late signers), unlike our expectations. The ones we observed were mostly reserved for re-introduced contexts along with nominals. Based on our findings across the two groups, we proposed a hierarchy of referent accessibility for TİD. As for the age of acquisition effects, we observed limited over-redundancy for signers with delayed language acquisition (i.e., late signers) who employed fewer implicit forms of reference tracking compared to native signers. Native signers were less likely to use a nominal for maintained contexts and more likely to use null marking for introduced contexts than late signers. We conclude that native signers might only partially adhere to pragmatic norms and leverage accessibility better than late signers since they use less form when the referent is more active in the mind.

\clearpage

\appendix

# Participant background information

\label{section:a}

```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
library(magrittr)
library(tidyverse)
library(readr)
library(kableExtra)
library("papaja")
library(stringr)
h1 = c("No", "Status", "Sex", "AOA", "Age", "LOE", "Deaf schools", "Home language", "Parents")

df_table_a <- read.csv("./appendix/appendix_a_table.csv", sep = ";")
names(df_table_a) <- h1
df_table_a %<>% dplyr::select(-Parents) 
apa_table(df_table_a,
          caption ="Participant demographic information based on self-report",
          note = "AOA = age of acquisition, LOE = level of education, PS = primary school, MS = middle school, HS = high school, TR = (Spoken) Turkish.",
          escape = FALSE,
          placement = "hbt!",
          font_size = "small",
          landscape = FALSE
)
```

\clearpage

# Effect of narrating to the camera vs. experimenter
\label{section:b}
```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}

# df for appendix (those who narrated to the experimenter are removed)
df_only_camera <- df_narration %>% filter(!Subject %in% c(16, 20, 21, 24))

# Coding nominal, Zero, Classifier, and Pronominal categories
df_only_camera$ref_type <- with(df_only_camera, case_when(
  GTR1 %in% c("Bare noun", "Point noun", "Noun Point",
              "Fingerspelled noun", "CL noun", "Noun CL") ~ "Nominal",
  GTR1 %in% c("Constructed action", "Constructed Action", "Plain verb", "Person agreement verb") ~ "Null",
  GTR1 %in%  c("Whole entity CL", "Body part CL",
               "SASS - Tracing", "SASS - Static") ~ "Classifier",
  T ~ "Pronominal"))

# Encode vector types
df_only_camera$Narrative %<>% as.integer() 
df_only_camera$Subject %<>% as.integer()
df_only_camera$GTR1 %<>% as.factor() 
df_only_camera$Nativeness %<>% dplyr::recode(`0` = "Late", `1` = "Native") %>% as.factor() %>% reorder.factor(new.order = c("Native","Late"))
df_only_camera$Discourse %<>% as.factor()
df_only_camera$ref_type %<>% as.factor()

# dropping Pronominal references because they are too few.
df_model_camera <- df_only_camera %>% subset(ref_type !="Pronominal") %>% drop.levels()

# Contrast coding
df_model_camera$Nativeness %<>% reorder.factor(new.order = c("Native","Late"))
contrasts(df_model_camera$Nativeness) <- contr.sum(2)/2
contrasts(df_model_camera$Discourse) <- contr.sum(3)/2
contrasts(df_model_camera$Group) <- contr.sdif(3)

# counts of ref types
ref_count <- df_only_camera %>% group_by(ref_type) %>% summarise(count=n())

# Model for reference type
model_camera <- brm(ref_type ~ Nativeness*Discourse + (1|Subject), 
               data = df_model_camera, family = categorical(link = "logit"), 
               iter = 3000, warmup = 2000, chains = 4, cores = 4, file ="./rds/model_camera.rds")

# Posterior plot for reference type
df_model_camera <- mcmc_intervals_data(model_1, pars = vars(starts_with("b_")))
df_model_camera$parameter %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
                                        `b_muNull_Intercept` = "Null_Intercept",
                                        `b_muNominal_Nativeness1` = "Nom_Native",
                                        `b_muNull_Nativeness1` = "Null_Native",
                                        `b_muNominal_Discourse1` = "Nom_Intro",
                                        `b_muNominal_Discourse2` = "Nom_Maintain",
                                        'b_muNominal_Nativeness1:Discourse1' = 'Nom_Native*Intro',
                                        'b_muNominal_Nativeness1:Discourse2' = 'Nom_Native*Maintain',
                                        'b_muNull_Nativeness1:Discourse1' = 'Null_Native*Intro',
                                        'b_muNull_Nativeness1:Discourse2' = 'Null_Native*Maintain',
                                        `b_muNull_Discourse1` = "Null_Intro",
                                        `b_muNull_Discourse2` = "Null_Maintain")
saveRDS(df_model_camera, "./rds/df_model_camera.rds")
```

```{r fig:model-camera, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
p4 <- df_model_camera %>% 
  ggplot(aes(m, y=factor(parameter,levels = rev(levels(factor(parameter)))))) +
  geom_point(size = 2.5) +
  geom_errorbarh(aes(xmin = l, xmax = h), alpha = 1, height = 0, size = 1.5) +
  geom_errorbarh(aes(xmin = ll, xmax = hh, height = 0)) + vline_0(linetype = "dashed") +
  xlab("Estimate (log odds)") + ylab("Coefficients") + theme_bw() + theme_apa()
p4
```

\clearpage

# Additional model with 3 levels for Nativeness

\label{section:c}

```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warning=FALSE}
library(magrittr)
library(tidyverse)
library(readr)
library(kableExtra)
library("papaja")
library(janitor)
df <- summary(model_3) %>% .$fixed
df <- as.data.frame(df)
df_table_extra <- df %>%  clean_names() %>% rownames_to_column("Coefficients") %>% dplyr::select(-7,-8)

names(df_table_extra) <- c("Coefficients","Estimate", "EstError","lower95","upper95","Rhat")
df_table_extra$Coefficients %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
                                        `muNull_Intercept` = "Null_Intercept",
                                        `muNominal_Group2M1` = "Nom_Late1vsNative",
                                        `muNominal_Group3M2` = "Nom_Late2vsLate1",
                                        `muNominal_Discourse1` = "Nom_Intro",
                                        `muNominal_Discourse2` = "Nom_Maintain",
                                        `muNominal_Group2M1:Discourse1` = 'Nom_L1vsN*Intro',
                                        `muNominal_Group3M2:Discourse1` = 'Nom_L2vsL1*Intro',
                                        `muNominal_Group2M1:Discourse2` = 'Nom_L1vsN*Maintain',
                                        `muNominal_Group3M2:Discourse2` = 'Nom_L2vsL1*Maintain',
                                        `muNull_Group2M1` = "Null_Late1vsNative",
                                        `muNull_Group3M2` = "Null_Late2vsLate1",
                                        `muNull_Discourse1` = "Null_Intro",
                                        `muNull_Discourse2` = "Null_Maintain",
                                        `muNull_Group2M1:Discourse1` = 'Null_L1vsN*Intro',
                                        `muNull_Group3M2:Discourse1` = 'Null_L2vsL1*Intro',
                                        `muNull_Group2M1:Discourse2` = 'Null_L1vsN*Maintain',
                                        `muNull_Group3M2:Discourse2` = 'Null_L2vsL1*Maintain')
df_table_extra %<>% mutate( across(.cols = "Coefficients",~str_replace( ., "_","\\\\_")))

apa_table(df_table_extra,
          format = "latex",
          booktabs = T,
          escape = FALSE,
          landscape = FALSE,
          font_size = "small",
          placement = "hbt!",
          caption ="Model results for Nativeness with 3 levels",
          note = "Rhat values above 1 indicate that the regression model chains did not converge properly. There were 4 chains in our regression model."
)
```

\clearpage

```{=tex}
\setlength{\columnsep}{10pt}
\twocolumn
\linespread{1}
\justifying
\footnotesize
```
# Episodes in the stimuli

\label{section:b}

**Narration 1**

1\. Tom chases Jerry, wearing a wooden plank

2\. Tom and Jerry enter a tube

3\. Tom falls down and Jerry runs away

4\. Tube gets thinner and Tom chases Jerry

5\. Jerry gets out and escapes

6\. Tom follows Jerry in shape of a stick

**Narration 2**

1\. Tom walks past a female cat sitting on a couch

2\. Tom turns to and approaches female cat

3\. Female cat blushes

4\. Tom offers a fish in aquarium to female cat

5\. Female cat rejects offer

6\. Tom offers a bird in cage to female cat

7\. Episode 5 is repeated

8\. Tom leaves by walking in a flirtatious manner

**Narration 3**

1\. Jerry falls down and tries to walk up stairs

2\. Tom pulls carpet back

3\. Jerry struggles

4\. A piano falls down stairs

5\. Jerry manages to escape

6\. Tom sees piano and tries to run away

7\. Piano smashes into Tom and flattens him

8\. Tom falls on ground

**Narration 4**

1\. Tom hides behind a cardboard house

2\. Jerry walks to house with a toy mouse

3\. Toy leaves Jerry who stops at entrance

4\. Toy enters house

5\. Toy gets smashed into pieces

6\. Jerry is surprised

7\. Tom hurts his stomach after swallowing toy

8\. Tom looks at his broken teeth on a mirror

9\. Tom gets angry and throws down mirror

**Narration 5**

1\. Tom fakes sleeping and suddenly gets up

2\. Tom pulls a rope inside a couch

3\. Jerry appears holding a bar of soap

4\. Jerry washes Tom's mouth with bar and flees

5\. Tom spits soap out

6\. Jerry escapes to nest with rope

7\. Tom catches rope and stars pulling it

8\. A mouse trap appears and catches Tom's nose

9\. Tom yells and removes trap

**Narration 6**

1\. Tom points a sword at Jerry and catches him

2\. Tom looks at Jerry

3\. Red spots appear on Jerry's face

4\. Tom gets scared and drops Jerry

5\. Tom rushes into bathroom

6\. Tom washes his hands

7\. Tom opens a drawer

8\. Tom swallows pills one after another

9\. Feeling ill, Tom looks out of window

10\. Tom and Jerry look at each other

**Narration 7**

1\. Two male cats chase one another

2\. Tom sees a female cat and kisses her

3\. Tom hides away when black cat returns

4\. Tom pokes his head behind couch again

5\. Episode 3 is repeated

6\. black cat sees female cat and hugs her

7\. Tom appears and black cat kisses Tom

8\. Two male cats move away from one another

**Narration 8**

1\. Jerry hits Tom's knee with a hammer

2\. Tom suddenly jumps up

3\. Tom yells and holds his knee

4\. Jerry appraches Tom holding a thermometer

5\. Jerry sticks thermometer into Tom's mouth

6\. Jerry lights a fire under thermometer

7\. Red liquid starts to rise up

8\. Thermometer explodes

**Narration 9**

1\. Tom sits on Jerry who tries to get loose

2\. Jerry looks around and sees a hat ribbon

3\. Jerry finds a needle attached to hat

4\. Jerry takes needle and sticks Tom with it

5\. Tom turns red and jumps

6\. Jerry placesn eedle between two cushions

7\. Tom dodges needle

**Narration 10**

1\. Wearing a stethoscope, Tom catches Jerry

2\. Jerry yells into stethoscope

3\. Tom's head pops up

4\. Jerry runs and enter a mouse nest

5\. Tom inserts a gun into nest

6\. Gun bends over toward Tom

7\. Tom fires gun and his head explodes \clearpage

\onecolumn

# References

### 2022-08-05 manuscript_v2.tex
Add files via upload

--- 2022-08-05 manuscript_v2.Rmd
@@
----
-title: Reference Tracking Strategies of Native and Late Signers in Turkish Sign Language
-abstract: |
-  We investigated the reference tracking strategies among native and late adult deaf signers of Turkish Sign Language.
-Consistent with the theories of saliency and referential accessibility, regardless of their acquisiton groups, signers mainly used nominals and extension classifiers for introducing referents.
-To maintain a referent in the previous clause, signers chose null marking (e.g., constructed action, agreement and plain verbs).
-For topic shifts or re-introduced contexts, nominals and pronouns were chiefly favored although we observed very little pronominal use.
-As for age of acquisition effects, we only report limited over-redundancy for late signers who used null marking less compared to native signers especially for introduced and maintained contexts.
-Given our limited support for redundancy by late signers, we conclude that native signers might utilize the spatial affordances of the visual-spatial modality better than late signers only to a certain extent.
-keywords: "Referential accessibility, signed narratives, pragmatic competence, delayed language acquisition, Turkish Sign Language"
-journal: "Journal of Pragmatics"
-date: "`r Sys.Date()`"
-bibliography: ./bibliography/references.bib
-times: true
-numbersections: true
-classoption: "review"
-csl: ./bibliography/elsevier-harvard.csl
-output: rticles::elsevier_article
-header-includes:
-   - \usepackage{floatrow}
-   - \usepackage{hyperref}
-   - \usepackage{enumitem}  
-   - \usepackage{float}
-   - \floatplacement{figure}{H}
-   - \floatplacement{table}{H}
-   - \floatsetup[figure]{capposition=top}
-   - \floatsetup[table]{capposition=top}
-   - \usepackage{times}
-   - \usepackage{smartdiagram}
-   - \usepackage{ragged2e}
-   - \usepackage{tikz}
-   - \usepackage{xcolor}
-   - \usepackage[utf8]{inputenc}
-   - \usepackage[T1]{fontenc}
-   - \usepackage{graphicx}
-   - \usepackage{./sty/langsci-gb4e}
-   - \usepackage{caption}
-   - \usepackage{multicol}
-   - \setlength{\columnsep}{-3cm}
-   - \usepackage{setspace}
----
-
-# Introduction
-
-Narratives are extensive discourses that construct an important part of human communication.
-In order to create a systematic and coherent organization of experiences, narrators must choose appropriate linguistic forms to refer to entities [@ariel1990; @givon1983; @gundel1993]: Certain forms such as nominal expressions are commonly used for the first mention of a referent in a discourse, and other forms such as pronouns and null marking may track previously mentioned entities which could either be maintained across multiple clauses or re-introduced back into the discourse.
-The selection of these referring expressions (henceforth REs) is contingent on cognitive abilities, such that individuals must rely on the activation of the working memory to keep in mind previously referred expressions and long strings of information [@morgan2005; @bamberg1997].
-Pragmatic abilities too play a pivotal role in the tracking of referents.
-The form with the least amount of redundant information must be selected based on the accessibility levels of the referent (i.e., how active a referent is in the addressee's mind given a point in discourse) [@ahn2019; @ariel1990].
-Because of these cognitive and pragmatic necessities, it is not surprising to observe that producing a coherent narrative is one of the latest milestones in language acquisition among children [@hickmann1996].
-As can be expected, there is a plethora of research examining how narrative competence is achieved in both monolingual and bilingual acquisition contexts among children and adults.
-Although narrative competence and referent selection have been studied extensively for typically developing speakers [@aksu-koç2015; @hickmann1996; @williams1988], deaf individuals who acquired a sign language in atypical learning contexts (i.e., deaf signers with linguistic deprivation) have remained mostly understudied with an exception of handful of studies [@cormier2013; @gür2018; @becker2009].
-Nevertheless, it is paramount that we investigate and describe the effects of language deprivation among deaf adults on narrative competence more in depth.
-This will help us understand whether the loss of systematic and frequent language input in the early years of life affect how deaf signers produce a narrative in their adulthood.
-Given this, the present study will focus on deaf adult signers with language deprivation and examine their reference tracking strategies in signed narratives.
-
-Sign Languages are natural languages of the Deaf communities all around the world.
-They are independent languages, which do not derive from majority spoken language(s) but emerge naturally when deaf people form a community.
-Turkish Sign Language (Türk İşaret Dili -- TİD) is and has been the officially recognized language of the Deaf community in Turkey since 2005 [see @nuhbalaoglu2018 for a brief review of the historical development of the language].
-Deaf or hearing children acquiring a SL within a family where there is one or more deaf caregivers display the same developmental milestones as hearing children acquiring a spoken language in a hearing family [@chamberlain_language_1999; @meier1990].
-However, reports indicate that over 90% of all deaf individuals are born into hearing families who do not know how to sign (i.e., those with language deprivation) [@woll2013].
-This means that the auditory modality used in the household is not accessible to many deaf children.
-That is why the age of acquisition of a first functional language varies greatly among deaf children, with many of whom being deprived of their first languages typically until after age 7 when they start formal education in a deaf school.
-First language deprivation or delay has been found to have negative effects on certain cognitive and linguistic measures among late signing children.
-To name a few, delayed exposure to a first functional language is reported to negatively affect deaf children's morphosyntactic production and processing [@kayaba2022; @boudreault2006; @newport1990; @cheng2021; @karadöller2017; @cormier2012] and narrative abilities [e.g., @becker2009].
-In our overall research, we are concerned with the effects of this late exposure in the ultimate adult forms of TİD users.
-
-## Referring Expression (RE) Selection
-
-Natural languages can be systematically used to keep track of previous referents in the discourse and mark the changing cognitive statuses of the REs (i.e., from being active to less active in the addressee's mind) to create a more coherent narrative.
-Among the theoretical frameworks that aim to explain that different linguistic forms are selected in accordance with the addressee's mental representations of the referents in the discourse is the Givenness Hierarchy [@gundel1993].
-According to this model, given or familiar information is expressed in more attenuated forms such as pronominal use and null marking since the mental status of the expressed referent is already at the center of attention, as indicated by the present contextual information [@chafe1976].
-The Givenness Hierarchy is akin to other previously proposed models of reference control within the topicality and saliency paradigm such as the Accessibility Hierarchy [@ariel1990], Topic Continuity [@givon1983], and Familiarity Scale [@prince1981].
-@gundel1993 argued that the distribution of forms is not random but abides by the Gricean Maxim of Quantity [@grice1975].
-That is to say, individuals must make the most informative and succinct choice possible to retain economy of form.
-When multiple expressions are available for the same referent as in an utterance like (1), phonetically simplest form is selected with the least redundant information are selected based on an economy principle.
-In (1), REs for "Eddie" are given in bold.
-All three REs (nominal, pronominal, and null marking) refer to the same entity; however, the quantity of marking decreases as the referent's accessibility increases.
-On this ground, for highly accessible referents (i.e., those that are highly active in the mind), linguistic forms with the least phonetic content are selected for the least amount of processing effort.
-
-```{=tex}
+\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
+%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
+\usepackage[hyphens]{url}
+
+  \journal{Journal of Pragmatics} % Sets Journal name
+
+\usepackage{lineno} % add
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
+\usepackage{graphicx}
+%%%%%%%%%%%%%%%% end my additions to header
+
+\usepackage[T1]{fontenc}
+\usepackage{lmodern}
+\usepackage{amssymb,amsmath}
+\usepackage{ifxetex,ifluatex}
+\usepackage{fixltx2e} % provides \textsubscript
+% use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[utf8]{inputenc}
+\else % if luatex or xelatex
+  \usepackage{fontspec}
+  \ifxetex
+    \usepackage{xltxtra,xunicode}
+  \fi
+  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
+  \newcommand{\euro}{€}
+\fi
+% use microtype if available
+\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
+\bibliographystyle{elsarticle-harv}
+\usepackage{graphicx}
+\ifxetex
+  \usepackage[setpagesize=false, % page size defined by xetex
+              unicode=false, % unicode breaks when used with xetex
+              xetex]{hyperref}
+\else
+  \usepackage[unicode=true]{hyperref}
+\fi
+\hypersetup{breaklinks=true,
+            bookmarks=true,
+            pdfauthor={},
+            pdftitle={Reference Tracking Strategies of Native and Late Signers in Turkish Sign Language},
+            colorlinks=false,
+            urlcolor=blue,
+            linkcolor=magenta,
+            pdfborder={0 0 0}}
+\urlstyle{same}  % don't use monospace font for urls
+
+\setcounter{secnumdepth}{5}
+% Pandoc toggle for numbering sections (defaults to be off)
+
+% Pandoc citation processing
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+% for Pandoc 2.8 to 2.10.1
+\newenvironment{cslreferences}%
+  {}%
+  {\par}
+% For Pandoc 2.11+
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+
+% Pandoc header
+\usepackage{floatrow}
+\usepackage{hyperref}
+\usepackage{enumitem}
+\usepackage{float}
+\floatplacement{figure}{H}
+\floatplacement{table}{H}
+\floatsetup[figure]{capposition=top}
+\floatsetup[table]{capposition=top}
+\usepackage{times}
+\usepackage{smartdiagram}
+\usepackage{ragged2e}
+\usepackage{tikz}
+\usepackage{xcolor}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage{graphicx}
+\usepackage{./sty/langsci-gb4e}
+\usepackage{caption}
+\usepackage{multicol}
+\setlength{\columnsep}{-3cm}
+\usepackage{setspace}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+
+\begin{document}
+\begin{frontmatter}
+
+  \title{Reference Tracking Strategies of Native and Late Signers in
+Turkish Sign Language}
+
+  \begin{abstract}
+  We investigated the reference tracking strategies among native and
+  late adult deaf signers of Turkish Sign Language. Consistent with the
+  theories of saliency and referential accessibility, regardless of
+  their acquisiton groups, signers mainly used nominals and extension
+  classifiers for introducing referents. To maintain a referent in the
+  previous clause, signers chose null marking (e.g., constructed action,
+  agreement and plain verbs). For topic shifts or re-introduced
+  contexts, nominals and pronouns were chiefly favored although we
+  observed very little pronominal use. As for age of acquisition
+  effects, we only report limited over-redundancy for late signers who
+  used null marking less compared to native signers especially for
+  introduced and maintained contexts. Given our limited support for
+  redundancy by late signers, we conclude that native signers might
+  utilize the spatial affordances of the visual-spatial modality better
+  than late signers only to a certain extent.
+  \end{abstract}
+   \begin{keyword} Referential accessibility, signed narratives,
+pragmatic competence, delayed language acquisition, Turkish Sign
+Language\end{keyword}
+ \end{frontmatter}
+
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
+
+Narratives are extensive discourses that construct an important part of
+human communication. In order to create a systematic and coherent
+organization of experiences, narrators must choose appropriate
+linguistic forms to refer to entities (Ariel, 1990; Givon, 1983; Gundel
+et al., 1993): Certain forms such as nominal expressions are commonly
+used for the first mention of a referent in a discourse, and other forms
+such as pronouns and null marking may track previously mentioned
+entities which could either be maintained across multiple clauses or
+re-introduced back into the discourse. The selection of these referring
+expressions (henceforth REs) is contingent on cognitive abilities, such
+that individuals must rely on the activation of the working memory to
+keep in mind previously referred expressions and long strings of
+information (Bamberg, 1997; Morgan, 2005). Pragmatic abilities too play
+a pivotal role in the tracking of referents. The form with the least
+amount of redundant information must be selected based on the
+accessibility levels of the referent (i.e., how active a referent is in
+the addressee's mind given a point in discourse) (Ahn, 2019; Ariel,
+1990). Because of these cognitive and pragmatic necessities, it is not
+surprising to observe that producing a coherent narrative is one of the
+latest milestones in language acquisition among children (Hickmann et
+al., 1996). As can be expected, there is a plethora of research
+examining how narrative competence is achieved in both monolingual and
+bilingual acquisition contexts among children and adults. Although
+narrative competence and referent selection have been studied
+extensively for typically developing speakers (Aksu-Koç and
+Nicolopoulou, 2015; Hickmann et al., 1996; Williams, 1988), deaf
+individuals who acquired a sign language in atypical learning contexts
+(i.e., deaf signers with linguistic deprivation) have remained mostly
+understudied with an exception of handful of studies (Becker, 2009;
+Cormier et al., 2013; Gür, 2018). Nevertheless, it is paramount that we
+investigate and describe the effects of language deprivation among deaf
+adults on narrative competence more in depth. This will help us
+understand whether the loss of systematic and frequent language input in
+the early years of life affect how deaf signers produce a narrative in
+their adulthood. Given this, the present study will focus on deaf adult
+signers with language deprivation and examine their reference tracking
+strategies in signed narratives.
+
+Sign Languages are natural languages of the Deaf communities all around
+the world. They are independent languages, which do not derive from
+majority spoken language(s) but emerge naturally when deaf people form a
+community. Turkish Sign Language (Türk İşaret Dili -- TİD) is and has
+been the officially recognized language of the Deaf community in Turkey
+since 2005 (see Nuhbalaoglu, 2018 for a brief review of the historical
+development of the language). Deaf or hearing children acquiring a SL
+within a family where there is one or more deaf caregivers display the
+same developmental milestones as hearing children acquiring a spoken
+language in a hearing family (Chamberlain et al., 1999; Meier and
+Newport, 1990). However, reports indicate that over 90\% of all deaf
+individuals are born into hearing families who do not know how to sign
+(i.e., those with language deprivation) (Woll, 2013). This means that
+the auditory modality used in the household is not accessible to many
+deaf children. That is why the age of acquisition of a first functional
+language varies greatly among deaf children, with many of whom being
+deprived of their first languages typically until after age 7 when they
+start formal education in a deaf school. First language deprivation or
+delay has been found to have negative effects on certain cognitive and
+linguistic measures among late signing children. To name a few, delayed
+exposure to a first functional language is reported to negatively affect
+deaf children's morphosyntactic production and processing (Boudreault
+and Mayberry, 2006; Cheng and Mayberry, 2021; Cormier et al., 2012;
+Karadöller et al., 2017; Kayabaşı and Gökgöz, 2022; Newport, 1990) and
+narrative abilities (e.g., Becker, 2009). In our overall research, we
+are concerned with the effects of this late exposure in the ultimate
+adult forms of TİD users.
+
+\hypertarget{referring-expression-re-selection}{%
+\subsection{Referring Expression (RE)
+Selection}\label{referring-expression-re-selection}}
+
+Natural languages can be systematically used to keep track of previous
+referents in the discourse and mark the changing cognitive statuses of
+the REs (i.e., from being active to less active in the addressee's mind)
+to create a more coherent narrative. Among the theoretical frameworks
+that aim to explain that different linguistic forms are selected in
+accordance with the addressee's mental representations of the referents
+in the discourse is the Givenness Hierarchy (Gundel et al., 1993).
+According to this model, given or familiar information is expressed in
+more attenuated forms such as pronominal use and null marking since the
+mental status of the expressed referent is already at the center of
+attention, as indicated by the present contextual information (Chafe,
+1976). The Givenness Hierarchy is akin to other previously proposed
+models of reference control within the topicality and saliency paradigm
+such as the Accessibility Hierarchy (Ariel, 1990), Topic Continuity
+(Givon, 1983), and Familiarity Scale (Prince, 1981). Gundel et al.
+(1993) argued that the distribution of forms is not random but abides by
+the Gricean Maxim of Quantity (Grice, 1975). That is to say, individuals
+must make the most informative and succinct choice possible to retain
+economy of form. When multiple expressions are available for the same
+referent as in an utterance like (1), phonetically simplest form is
+selected with the least redundant information are selected based on an
+economy principle. In (1), REs for ``Eddie'' are given in bold. All
+three REs (nominal, pronominal, and null marking) refer to the same
+entity; however, the quantity of marking decreases as the referent's
+accessibility increases. On this ground, for highly accessible referents
+(i.e., those that are highly active in the mind), linguistic forms with
+the least phonetic content are selected for the least amount of
+processing effort.
+
 \begin{exe}
@@
 \end{exe}
-```
-It is also possible to categorize REs into three according to their discourse status [@gullberg2006].
-A referent can be introduced or mentioned for the first time, maintained, that is, continued across two or more clauses, or re-introduced into the discourse following a topic shift.
-As a result, interlocutors mark some REs differently than others based on the principles of reference accessibility.
-For spoken languages, this observation is well established [@ariel1990; @givon1983; @gundel1993; @debreslioska2013; @ahrenholz2005; @chafe1976; @hickmann1996; @azar2015].
-These research indicate that the introduction and re-introduction of a referent are often accomplished by full noun phrases which can be commonly modified by an adjective or demonstrative.
-Maintenance, on the other hand, is typically marked with pronominals and null markers since the continued referent is still highly accessible to the addressee.
-
-However, this particular observation is not peculiar to spoken languages.
-There is an emerging body of literature that reports the distribution of REs in signed narratives.
-Building on @gundel1993's work, @swabey2002 demonstrated that, like other spoken languages such as Russian, Chinese, and Japanese, American Sign Language (ASL) signers use bare nouns for introduced and reintroduced discourse contexts whereas null marking is used for referents which are "in-focus" or very accessible to the addressee.
-Overt REs like pronouns or nouns are disprefered for such referents.
-@swabey2002 concluded that "native" signers in her study abided by the Gricean maxim of quantity and were thus economic in their reference tracking.
-Similar findings have been reported for German Sign Language (DGS) [@perniss2015] and British Sign Language (BSL) [@morgan2005].
-It is important to note, however, that these research either collapsed native and late signers into a single group and labeled them as "native-like" or explicitly compared native deaf signers with late deaf signers (with language deprivation) in terms of narrative competence.
-
-### Signed vs. Spoken REs
-
-Commonly in both spoken and sign languages nouns and noun phrases are used as referring expressions.
-Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions [@lillo-martin1990].
-In addition, the pointing sign is used to referentially anchor a referent to a location in signing space.
-For instance, when the signer introduces \textsc{elvan} for the first time, s/he can assign her to the right.
-This referential locus can later be used by agreement verbs and classifier predicates (more below).
-For instance, in (2a) there are two referents, \textsc{elvan}, who is referentially-assigned to the right side of the signer, represented with the "a" subscript and \textsc{demet}, who is referentially-assigned to the left side of the signer, represented with the "b" subscript.
-When the signer makes a contrast in (2b), as a pronominal use, she points to the right side to refer to ELVAN and to the left to refer to \textsc{demet}.
-We schematically represents the referential locus assigned to the referents in (3).
-
-```{=tex}
+
+It is also possible to categorize REs into three according to their
+discourse status (Gullberg, 2006). A referent can be introduced or
+mentioned for the first time, maintained, that is, continued across two
+or more clauses, or re-introduced into the discourse following a topic
+shift. As a result, interlocutors mark some REs differently than others
+based on the principles of reference accessibility. For spoken
+languages, this observation is well established (Ahrenholz, 2005; Ariel,
+1990; Azar and Özyürek, 2015; Chafe, 1976; Debreslioska et al., 2013;
+Givon, 1983; Gundel et al., 1993; Hickmann et al., 1996). These research
+indicate that the introduction and re-introduction of a referent are
+often accomplished by full noun phrases which can be commonly modified
+by an adjective or demonstrative. Maintenance, on the other hand, is
+typically marked with pronominals and null markers since the continued
+referent is still highly accessible to the addressee.
+
+However, this particular observation is not peculiar to spoken
+languages. There is an emerging body of literature that reports the
+distribution of REs in signed narratives. Building on Gundel et al.
+(1993)'s work, Swabey (2002) demonstrated that, like other spoken
+languages such as Russian, Chinese, and Japanese, American Sign Language
+(ASL) signers use bare nouns for introduced and reintroduced discourse
+contexts whereas null marking is used for referents which are
+``in-focus'' or very accessible to the addressee. Overt REs like
+pronouns or nouns are disprefered for such referents. Swabey (2002)
+concluded that ``native'' signers in her study abided by the Gricean
+maxim of quantity and were thus economic in their reference tracking.
+Similar findings have been reported for German Sign Language (DGS)
+(Perniss and Özyürek, 2015) and British Sign Language (BSL) (Morgan,
+2005). It is important to note, however, that these research either
+collapsed native and late signers into a single group and labeled them
+as ``native-like'' or explicitly compared native deaf signers with late
+deaf signers (with language deprivation) in terms of narrative
+competence.
+
+\hypertarget{signed-vs.-spoken-res}{%
+\subsubsection{Signed vs.~Spoken REs}\label{signed-vs.-spoken-res}}
+
+Commonly in both spoken and sign languages nouns and noun phrases are
+used as referring expressions. Spoken languages use pronouns, determines
+and demonstratives where sign languages use a pointing sign, usually
+glossed as IX, which assumes similar functions (Lillo-Martin and Klima,
+1990). In addition, the pointing sign is used to referentially anchor a
+referent to a location in signing space. For instance, when the signer
+introduces \textsc{elvan} for the first time, s/he can assign her to the
+right. This referential locus can later be used by agreement verbs and
+classifier predicates (more below). For instance, in (2a) there are two
+referents, \textsc{elvan}, who is referentially-assigned to the right
+side of the signer, represented with the ``a'' subscript and
+\textsc{demet}, who is referentially-assigned to the left side of the
+signer, represented with the ``b'' subscript. When the signer makes a
+contrast in (2b), as a pronominal use, she points to the right side to
+refer to ELVAN and to the left to refer to \textsc{demet}. We
+schematically represents the referential locus assigned to the referents
+in (3).
+
 \begin{exe}
@@
     \end{exe}
-```
-Grammatical gender as in Romance languages or noun classes as in Bantu languages are some reference tracking tools that are familiar from spoken languages.
-A somehow less known reference-tracking tool in spoken languages is *verbal classifiers*, which morphologically attach to the verb and often cross-reference a core argument in a clause [@aikhenvald2017].
-Classifiers denote sets of objects based on semantic criteria such as size, shape, body-part, texture and sometimes function.
-Usually, the verbal classifier and the cross-referenced noun co-occur in the same clause when the argument is introduced into discourse.
-The classifier can track the argument on its own on subsequent occurrences.
-In sign languages, a verbal classifier morpheme is encoded by a *handshape* that is articulated phonologically co-temporally with a movement, which encodes the verbal meaning [@supalla1990; @benedicto2004].
-These two infused morphemes form a complex predicate.
-Some classifier handshapes and the set of objects that they denote are below.
-Some classifier handshapes and the set of objects that they denote are presented below in (4).
-
-```{=tex}
+
+Grammatical gender as in Romance languages or noun classes as in Bantu
+languages are some reference tracking tools that are familiar from
+spoken languages. A somehow less known reference-tracking tool in spoken
+languages is \emph{verbal classifiers}, which morphologically attach to
+the verb and often cross-reference a core argument in a clause
+(Aikhenvald, 2017). Classifiers denote sets of objects based on semantic
+criteria such as size, shape, body-part, texture and sometimes function.
+Usually, the verbal classifier and the cross-referenced noun co-occur in
+the same clause when the argument is introduced into discourse. The
+classifier can track the argument on its own on subsequent occurrences.
+In sign languages, a verbal classifier morpheme is encoded by a
+\emph{handshape} that is articulated phonologically co-temporally with a
+movement, which encodes the verbal meaning (Benedicto and Brentari,
+2004; Supalla, 1990). These two infused morphemes form a complex
+predicate. Some classifier handshapes and the set of objects that they
+denote are below. Some classifier handshapes and the set of objects that
+they denote are presented below in (4).
+
 \begin{exe}
@@
 \end{exe}
-```
-In (5), in the first clause, a referential locus, i.e.
-locus-a, is assigned to the locative source argument \textsc{cage}.
-In the second clause, the core argument \textsc{bird} is cross-referenced with the classifier handshape, i.e.
-small-legged animal.
-It flies from the referential locus of the \textsc{cage} (locus-a) to the referential locus of the goal \textsc{tree} (locus-b).
-
-```{=tex}
+
+In (5), in the first clause, a referential locus, i.e.~locus-a, is
+assigned to the locative source argument \textsc{cage}. In the second
+clause, the core argument \textsc{bird} is cross-referenced with the
+classifier handshape, i.e.~small-legged animal. It flies from the
+referential locus of the \textsc{cage} (locus-a) to the referential
+locus of the goal \textsc{tree} (locus-b).
+
 \begin{exe}
@@
     \hfill (Adapted from Gökgöz and Sevgi 2020:318)
-```
-## Linguistic Tools for Tracking Referents in Signed Narratives
-
-The linguistic inventory for discourse referents can include nominal constructions (bare nouns, fingerspelled nouns, and modified nouns), null markers (plain verb, agreement verb, and constructed action), classifier predicates (whole entity classifiers, body part classifiers, handling classifiers, and extension classifiers), and overt pronouns (pronominal index or IX) [@frederiksen2016].
-This interim section describes such constructions attested for sign languages under four main headings (nominals, pronouns, classifiers, and null marking).
+
+\hypertarget{linguistic-tools-for-tracking-referents-in-signed-narratives}{%
+\subsection{Linguistic Tools for Tracking Referents in Signed
+Narratives}\label{linguistic-tools-for-tracking-referents-in-signed-narratives}}
+
+The linguistic inventory for discourse referents can include nominal
+constructions (bare nouns, fingerspelled nouns, and modified nouns),
+null markers (plain verb, agreement verb, and constructed action),
+classifier predicates (whole entity classifiers, body part classifiers,
+handling classifiers, and extension classifiers), and overt pronouns
+(pronominal index or IX) (Frederiksen and Mayberry, 2016). This interim
+section describes such constructions attested for sign languages under
+four main headings (nominals, pronouns, classifiers, and null marking).
 
@@
 
-```{=tex}
 \begin{multicols}{2}
@@
 \footnotetext{We use small capitals to indicate sign glosses.}
-```
-### Nominals
-
-In TİD and many other sign languages, it is possible to use bare noun phrases (\textsc{bare nom}) as in (1a) as well as fingerspelled nouns (\textsc{fs nom}) as in (1c).
-Fingerspelling refers to the use of the manual alphabet which is derived or borrowed from a surrounding written language [@kubu2008].
-Nouns can also be modified (\textsc{mod nom}) by classifiers as in (1b) [@frederiksen2016] or by a prenominal or postnominal IX functioning as a determiner or a demonstrative as in (1d) [@maclaughlin1997].
-
-### Pronouns
-
-The IX sign (extended index finger) has various functions in sign languages, one of which is the pronominal use which refers to established nominals by pointing to a previously assigned abstract or physical locus in the signing space when re-introducing a topic in discourse [@emmorey1996].
-The same IX function has also been identified for TİD [@sevinç2006].
-The example in (2) illustrates a spatially-anchored pronominal reference for topic re-introduction.
-These signs are thus rarely used for the first mention of an entity [@perniss2015].
-Other interpretations consider pointing signs without noun phrases as examples of demonstratives and not personal pronouns [@koulidobrova2016].
-It is important to mention here that in some spoken languages (e.g., German and Dutch), demonstratives can also refer to persons [@bosch2007] and behave similarly to personal pronouns.
-However, for the purposes of the present study, we will refer to IX constructions as pronominal IX and particularly examine their non-deictic or personal pronoun realizations as a possible discourse strategy to mainly mark topic shift among native and late signers of TİD.
-
-### Classifiers
-
-Classifier predicates are highly complex morphological constructions which are very common in sign languages [@supalla1986].
-These are manually signed iconic representations of animate or inanimate objects, and they embody several properties pertaining to the nominal referent including its location in the signing space, movement, path, size, and shape [@talmy2003].
-As in most sign languages hitherto studied, classifier constructions are also common and found in TİD [see @dikyuva2015 for a review].
-Following the classification of @engberg-pedersen1993, we distinguish between four types of classifier constructions.
-Whole entity classifiers (WECL) represent the entirety of any referent whereas body part classifiers (BPCL) only reflect a specific part of an animate entity (e.g., limbs, head, tongue, teeth, and so forth).
-Handling classifiers (HCL) occur with transitive verbs and depict how an object is handled or manipulated by agent.
-On this ground, an HCL[^1] depicts the manner of the referring action while the former two constructions encode the location of the entity as well as its motion in space.
-Extension classifiers (ExtCL), on the other hand, denote the size and shape of an entity by following a trajectory to trace its shape.
-We exemplify these constructions in (3).
-
-[^1]: As discussed before, an HCL is transitive in nature and depicts the handling action performed by ananimate agent.
-This is why most research tend to consider HCLs as an extension to constructed action [@dudis2004; @cormier2013].
-
-### Null marking
-
-Parallel to spoken languages, many sign languages have been reported to allow null subjects and/or objects and make use of null arguments when referring to highly accessible referents which are still active in the mental representation of the addressee.
-In signed discourse, this can be achieved by employing certain verbal constructions.
-To identify the predicate types in ASL, @padden1986 suggested a triadic categorization of verbs in sign languages: plain verbs, agreement verbs, and spatial verbs.
-According to this classification, plain verbs (plain) (4b) do not agree with the subject or the object.
-This means that they are not inflected for number or person.
-This contrasts with agreeing verbs (agreement) (4c) the starting or the ending point of which can be modified to specific loci for person and number agreement while spatial verbs (as in \textsc{ix book put} 'I placed the book there') agree with locative arguments but like plain verbs, they are not inflected for number or person.
-TİD too licenses null arguments.
-Following this classification, @sevinç2006 argued that although it is common in TİD to drop arguments, there were certain limitations on the null marking of the arguments for agreement and plain verbs.
-
-## Learning How to Track Referents in Sign Languages
-
-In both communication modalities (visual and auditory), children start to produce more complex and coherent stories with appropriate and informative linguistic forms to indicate anaphoric or cataphoric references after age 10 [@hickmann1996; @rathmann2007].
-Prior to this age, both signing and speaking children fail to refer to entities consistently since their pragmatic abilities are yet to develop.
-The manipulation of linguistic forms for discourse functions to signal topic continuance or shift is difficult to master until the beginning of formal education.
-These findings are also comparable across different sign languages including British Sign Language (BSL) [@morgan2005], ASL [@loew1984], and German Sign Language (Deutsche Gebärdensprache - DGS) [@becker2009], pointing out to the fact that the mastery of reference tracking in narratives is one of the most protracted achievements in the acquisition of signed discourse.
-
-Like in first language development, narrative skills have been extensively studied among second language learners [see @bel2015; @frederiksen2019].
-Comparing 11 deaf adults signers and 13 hearing advanced learners of Catalan Sign Language (Llengua de Signes Catalana - LSC), @bel2015 found that hearing signers resort to more overt forms (nominals and pronouns) than native signers to avoid potential ambiguities among referents.
-Nevertheless, their results did not reach statistical significant for the interaction of discourse status and acquisition group.
-A similar line of research came from @frederiksen2019 who found a statistically significant difference between both groups, that is native signers and hearing L2 learners of ASL, only for classifier use especially in re-introduced contexts.
-L2 learners in the study used more nominals and fewer instances of null marking in number, indicative of a limited effect of overexplicitness among second language learners.
-However, these studies treated late and native deaf signers as a single group.
-That is, the control group consisted of both native signers who had one or more deaf parents and also native-like [@cormier2012] late deaf signers who had hearing parents and their age of acquisition varied.
-Given the preliminary evidence that topic continuity and salience marking is also difficult for second language learners regardless of modality, it has been suggested under the Interface Hypothesis that the syntax-discourse interface is vulnerable and especially challenging to acquire [@sorace2011] for second language learners.
-Similarly, @sorace2009 discuss that this interface entails much more than morphosyntactic knowledge but also includes cognitive and pragmatic resources that help provide discourse coherence.
-They propose that the second language discourse is over-redundant.
-This means that second language learners produce more overt pronouns or fuller forms of language for highly accessible referents possibly as a compensatory strategy in an attempt to reduce the increasing load on cognitive processing.
-
-Still, only very few research examined whether linguistically deprived signers (i.e., deaf signers with delayed exposure to a sign language) also display redundancy in their reference tracking.
-@becker2009 examined 6 DGS signing deaf children (3 native and 3 late)'s narrative abilities.
-She reported that late signing children generated more explicit forms such as nominals when referring to the characters in the story compared to native learners, concluding that first language deprivation results in an inadequate representation of the addressee and less efficient use of morphosyntax.
-Comparably, @cormier2013 investigated the use of constructed action in 15 BSL among deaf native and late adult signers and found that late signers were redundant in topic continuity.
-In other words, the late signers produced more explicit forms like nominals, and fewer instances of null marking when maintaining previous referents.
-Most relevantly to the present study, @gür2018 studied the development of narrative skills among 10 native and 10 late deaf children and deaf adult signers of TİD.
-She found that native-signing children and adults had more instances of explicit referent introduction, set the scene more frequently, and employed constructed action in an appropriate manner more than corresponding late-signing children and adults.
-Although this study presented preliminary findings on age of acquisition effects on narrative competence, it does not investigate how native and late signers employ different linguistic tools based on the accessibility of the referent.
-Therefore, to our knowledge, there is a lack of research directly examining the effects of first language deprivation on deaf adult signers' selection of REs in signed narratives.
-
-Considering this, it is not very clear whether the negative effects on narrative abilities observed among late signing deaf children hold when they become adults.
-Understanding the impact of language deprivation or delayed first language learning on narrative competence is crucial for three reasons.
-First, studying this population will shed light on the possible effect of lack of early language experience on future cognitive and pragmatic abilities since the narrative competence is dependent on such faculties.
-Second, examining reference tracking strategies in particular will also inform about the linguistic inventory that native and late deaf signers use to refer to entities.
-Finally, any possible difference between the two acquisition groups in terms of narrative competence will have further practical implications for the importance of early access to a first functional sign language to retain future linguistic and cognitive abilities.
-
-## Present Study
-
-In an attempt at addressing this gap in the literature, the present study aims to investigate the reference tracking strategies of native and late adult signers in Turkish Sign Language (Türk İşaret Dili - TİD) narratives with a story-telling paradigm.
-We explore the forms available for referent tracking in TİD and examine the nature of RE selection as a possible indicator of pragmatic competence between our acquisition groups.
-
-Overall, considering what has been suggested for the distribution of REs in signed narratives, we expect more explicit forms like nominals for referent introduction and and less explicit forms such as null marking, pronominal and/or classifier use for referent maintainance.
-For the re-introduction contexts, where referents which are discontinued across one or more clauses are brought back into the discourse, we expect a less clear picture.
-Some re-introduced referents will have higher accessibility and will thus be tracked with more implicit REs, while others that have lower accessibility will be tracked with more explicit REs Also based on previous research, we hypothesize that late signers will become more over-redundant in their reference tracking compared to native signers.
-This would translate to a preference by late signers for nominals over null marking especially in maintenance and re-introduction contexts.
-In contrast, we would expect that native signers use more null marking to track referents.
-
-# Method
-
-## Participants
-
-We tested 29 deaf adult signers, 15 with exposure to TİD from birth (i.e., native signers), 14 with exposure between ages 3 and 17 (i.e., late signers) (*M*\textsubscript{Age of acquisition} = 7.7 years, *SD* = 3.3 years).
-The criterion to be classified as a late signer was to have hearing and non-signing caregivers.
-All the signers satisfying this criterion were regarded as late signers \footnote{Note that we are essentially comparing deaf of deaf and deaf of hearing adult signers.
+
+\hypertarget{nominals}{%
+\subsubsection{Nominals}\label{nominals}}
+
+In TİD and many other sign languages, it is possible to use bare noun
+phrases (\textsc{bare nom}) as in (1a) as well as fingerspelled nouns
+(\textsc{fs nom}) as in (1c). Fingerspelling refers to the use of the
+manual alphabet which is derived or borrowed from a surrounding written
+language (Kubuş, 2008). Nouns can also be modified (\textsc{mod nom}) by
+classifiers as in (1b) (Frederiksen and Mayberry, 2016) or by a
+prenominal or postnominal IX functioning as a determiner or a
+demonstrative as in (1d) (Maclaughlin, 1997).
+
+\hypertarget{pronouns}{%
+\subsubsection{Pronouns}\label{pronouns}}
+
+The IX sign (extended index finger) has various functions in sign
+languages, one of which is the pronominal use which refers to
+established nominals by pointing to a previously assigned abstract or
+physical locus in the signing space when re-introducing a topic in
+discourse (Emmorey, 1996). The same IX function has also been identified
+for TİD (Sevinç, 2006). The example in (2) illustrates a
+spatially-anchored pronominal reference for topic re-introduction. These
+signs are thus rarely used for the first mention of an entity (Perniss
+and Özyürek, 2015). Other interpretations consider pointing signs
+without noun phrases as examples of demonstratives and not personal
+pronouns (Koulidobrova and Lillo-Martin, 2016). It is important to
+mention here that in some spoken languages (e.g., German and Dutch),
+demonstratives can also refer to persons (Bosch and Umbach, 2007) and
+behave similarly to personal pronouns. However, for the purposes of the
+present study, we will refer to IX constructions as pronominal IX and
+particularly examine their non-deictic or personal pronoun realizations
+as a possible discourse strategy to mainly mark topic shift among native
+and late signers of TİD.
+
+\hypertarget{classifiers}{%
+\subsubsection{Classifiers}\label{classifiers}}
+
+Classifier predicates are highly complex morphological constructions
+which are very common in sign languages (Supalla, 1986). These are
+manually signed iconic representations of animate or inanimate objects,
+and they embody several properties pertaining to the nominal referent
+including its location in the signing space, movement, path, size, and
+shape (Talmy, 2003). As in most sign languages hitherto studied,
+classifier constructions are also common and found in TİD (see Dikyuva
+et al., 2015 for a review). Following the classification of
+Engberg-Pedersen (1993), we distinguish between four types of classifier
+constructions. Whole entity classifiers (WECL) represent the entirety of
+any referent whereas body part classifiers (BPCL) only reflect a
+specific part of an animate entity (e.g., limbs, head, tongue, teeth,
+and so forth). Handling classifiers (HCL) occur with transitive verbs
+and depict how an object is handled or manipulated by agent. On this
+ground, an HCL\footnote{As discussed before, an HCL is transitive in
+  nature and depicts the handling action performed by ananimate agent.
+  This is why most research tend to consider HCLs as an extension to
+  constructed action (Cormier et al., 2013; Dudis, 2004).} depicts the
+manner of the referring action while the former two constructions encode
+the location of the entity as well as its motion in space. Extension
+classifiers (ExtCL), on the other hand, denote the size and shape of an
+entity by following a trajectory to trace its shape. We exemplify these
+constructions in (3).
+
+\hypertarget{null-marking}{%
+\subsubsection{Null marking}\label{null-marking}}
+
+Parallel to spoken languages, many sign languages have been reported to
+allow null subjects and/or objects and make use of null arguments when
+referring to highly accessible referents which are still active in the
+mental representation of the addressee. In signed discourse, this can be
+achieved by employing certain verbal constructions. To identify the
+predicate types in ASL, Padden (1986) suggested a triadic categorization
+of verbs in sign languages: plain verbs, agreement verbs, and spatial
+verbs. According to this classification, plain verbs (plain) (4b) do not
+agree with the subject or the object. This means that they are not
+inflected for number or person. This contrasts with agreeing verbs
+(agreement) (4c) the starting or the ending point of which can be
+modified to specific loci for person and number agreement while spatial
+verbs (as in \textsc{ix book put} `I placed the book there') agree with
+locative arguments but like plain verbs, they are not inflected for
+number or person. TİD too licenses null arguments. Following this
+classification, Sevinç (2006) argued that although it is common in TİD
+to drop arguments, there were certain limitations on the null marking of
+the arguments for agreement and plain verbs.
+
+\hypertarget{learning-how-to-track-referents-in-sign-languages}{%
+\subsection{Learning How to Track Referents in Sign
+Languages}\label{learning-how-to-track-referents-in-sign-languages}}
+
+In both communication modalities (visual and auditory), children start
+to produce more complex and coherent stories with appropriate and
+informative linguistic forms to indicate anaphoric or cataphoric
+references after age 10 (Hickmann et al., 1996; Rathmann et al., 2007).
+Prior to this age, both signing and speaking children fail to refer to
+entities consistently since their pragmatic abilities are yet to
+develop. The manipulation of linguistic forms for discourse functions to
+signal topic continuance or shift is difficult to master until the
+beginning of formal education. These findings are also comparable across
+different sign languages including British Sign Language (BSL) (Morgan,
+2005), ASL (Loew, 1984), and German Sign Language (Deutsche
+Gebärdensprache - DGS) (Becker, 2009), pointing out to the fact that the
+mastery of reference tracking in narratives is one of the most
+protracted achievements in the acquisition of signed discourse.
+
+Like in first language development, narrative skills have been
+extensively studied among second language learners (see Bel et al.,
+2015; Frederiksen and Mayberry, 2019). Comparing 11 deaf adults signers
+and 13 hearing advanced learners of Catalan Sign Language (Llengua de
+Signes Catalana - LSC), Bel et al. (2015) found that hearing signers
+resort to more overt forms (nominals and pronouns) than native signers
+to avoid potential ambiguities among referents. Nevertheless, their
+results did not reach statistical significant for the interaction of
+discourse status and acquisition group. A similar line of research came
+from Frederiksen and Mayberry (2019) who found a statistically
+significant difference between both groups, that is native signers and
+hearing L2 learners of ASL, only for classifier use especially in
+re-introduced contexts. L2 learners in the study used more nominals and
+fewer instances of null marking in number, indicative of a limited
+effect of overexplicitness among second language learners. However,
+these studies treated late and native deaf signers as a single group.
+That is, the control group consisted of both native signers who had one
+or more deaf parents and also native-like (Cormier et al., 2012) late
+deaf signers who had hearing parents and their age of acquisition
+varied. Given the preliminary evidence that topic continuity and
+salience marking is also difficult for second language learners
+regardless of modality, it has been suggested under the Interface
+Hypothesis that the syntax-discourse interface is vulnerable and
+especially challenging to acquire (Sorace, 2011) for second language
+learners. Similarly, Sorace and Serratrice (2009) discuss that this
+interface entails much more than morphosyntactic knowledge but also
+includes cognitive and pragmatic resources that help provide discourse
+coherence. They propose that the second language discourse is
+over-redundant. This means that second language learners produce more
+overt pronouns or fuller forms of language for highly accessible
+referents possibly as a compensatory strategy in an attempt to reduce
+the increasing load on cognitive processing.
+
+Still, only very few research examined whether linguistically deprived
+signers (i.e., deaf signers with delayed exposure to a sign language)
+also display redundancy in their reference tracking. Becker (2009)
+examined 6 DGS signing deaf children (3 native and 3 late)`s narrative
+abilities. She reported that late signing children generated more
+explicit forms such as nominals when referring to the characters in the
+story compared to native learners, concluding that first language
+deprivation results in an inadequate representation of the addressee and
+less efficient use of morphosyntax. Comparably, Cormier et al. (2013)
+investigated the use of constructed action in 15 BSL among deaf native
+and late adult signers and found that late signers were redundant in
+topic continuity. In other words, the late signers produced more
+explicit forms like nominals, and fewer instances of null marking when
+maintaining previous referents. Most relevantly to the present study,
+Gür (2018) studied the development of narrative skills among 10 native
+and 10 late deaf children and deaf adult signers of TİD. She found that
+native-signing children and adults had more instances of explicit
+referent introduction, set the scene more frequently, and employed
+constructed action in an appropriate manner more than corresponding
+late-signing children and adults. Although this study presented
+preliminary findings on age of acquisition effects on narrative
+competence, it does not investigate how native and late signers employ
+different linguistic tools based on the accessibility of the referent.
+Therefore, to our knowledge, there is a lack of research directly
+examining the effects of first language deprivation on deaf adult
+signers' selection of REs in signed narratives.
+
+Considering this, it is not very clear whether the negative effects on
+narrative abilities observed among late signing deaf children hold when
+they become adults. Understanding the impact of language deprivation or
+delayed first language learning on narrative competence is crucial for
+three reasons. First, studying this population will shed light on the
+possible effect of lack of early language experience on future cognitive
+and pragmatic abilities since the narrative competence is dependent on
+such faculties. Second, examining reference tracking strategies in
+particular will also inform about the linguistic inventory that native
+and late deaf signers use to refer to entities. Finally, any possible
+difference between the two acquisition groups in terms of narrative
+competence will have further practical implications for the importance
+of early access to a first functional sign language to retain future
+linguistic and cognitive abilities.
+
+\hypertarget{present-study}{%
+\subsection{Present Study}\label{present-study}}
+
+In an attempt at addressing this gap in the literature, the present
+study aims to investigate the reference tracking strategies of native
+and late adult signers in Turkish Sign Language (Türk İşaret Dili - TİD)
+narratives with a story-telling paradigm. We explore the forms available
+for referent tracking in TİD and examine the nature of RE selection as a
+possible indicator of pragmatic competence between our acquisition
+groups.
+
+Overall, considering what has been suggested for the distribution of REs
+in signed narratives, we expect more explicit forms like nominals for
+referent introduction and and less explicit forms such as null marking,
+pronominal and/or classifier use for referent maintainance. For the
+re-introduction contexts, where referents which are discontinued across
+one or more clauses are brought back into the discourse, we expect a
+less clear picture. Some re-introduced referents will have higher
+accessibility and will thus be tracked with more implicit REs, while
+others that have lower accessibility will be tracked with more explicit
+REs Also based on previous research, we hypothesize that late signers
+will become more over-redundant in their reference tracking compared to
+native signers. This would translate to a preference by late signers for
+nominals over null marking especially in maintenance and re-introduction
+contexts. In contrast, we would expect that native signers use more null
+marking to track referents.
+
+\hypertarget{method}{%
+\section{Method}\label{method}}
+
+\hypertarget{participants}{%
+\subsection{Participants}\label{participants}}
+
+We tested 29 deaf adult signers, 15 with exposure to TİD from birth
+(i.e., native signers), 14 with exposure between ages 3 and 17 (i.e.,
+late signers) (\emph{M}\textsubscript{Age of acquisition} = 7.7 years,
+\emph{SD} = 3.3 years). The criterion to be classified as a late signer
+was to have hearing and non-signing caregivers. All the signers
+satisfying this criterion were regarded as late signers
+\footnote{Note that we are essentially comparing deaf of deaf and deaf of hearing adult signers.
 Although we did not set a cut-off point for nativeness, all the deaf of hearing participants indicated that they learned TİD after age 3.
 In order to capture in-group variation among late signers, we run an additional supplementary analysis}.
-All participants indicated that their preferred language of communication was TİD which they used for most contexts.
-Detailed information about the linguistic and educational background of the participants can be found in \ref{section:a}.
-
-All participating native signers (7 males, 8 females) were born into deaf families and started acquiring TİD from at least one of their caregivers after birth.
-The parents of 14 participating native signers were deaf.
-One native signer had a deaf father and a hearing mother.
-In addition to TİD, the home language for some native signers (n = 4) in early childhood included the use of home sign\footnote{Home signs are gesture-based communication systems that emerge between the deaf children and their primary caregivers.
+All participants indicated that their preferred language of
+communication was TİD which they used for most contexts. Detailed
+information about the linguistic and educational background of the
+participants can be found in \ref{section:a}.
+
+All participating native signers (7 males, 8 females) were born into
+deaf families and started acquiring TİD from at least one of their
+caregivers after birth. The parents of 14 participating native signers
+were deaf. One native signer had a deaf father and a hearing mother. In
+addition to TİD, the home language for some native signers (n = 4) in
+early childhood included the use of home
+sign\footnote{Home signs are gesture-based communication systems that emerge between the deaf children and their primary caregivers.
 Although home sign communication may have internal consistency, they are not fully-flegged linguistic systems like natural languages}.
-The age of the native participants ranged between 18 and 35 (*M* = 26.4 years, *SD* = 5.0 years).
-This also corresponded to the total average years of TİD use for native signers.
-All native signers were pre-lingually deaf, that is they were either born deaf or they had become deaf before the age of 3, with the exception of one signer who became deaf between the ages of 4 and 7.
-Native signers reported to complete at least high school education and their level of education (LOE) was between 12 and 16 years (*M* = 12.7 years, *SD* = 1.5 years).
-In addition, all native participants attended at least two deaf schools, one being primary and the other middle school.
-10 native participants also attended a deaf school for their high school education.
-Based on a self-reported 5-point scale of proficiency (1: poor, 5: proficient), native participants declared that they have high receptive and productive skills in TİD (*M* = 5 points, *SD* = 0 points).
-As for their proficiency in (spoken) Turkish, their self-reported rankings were on average 3.7 (*SD* = 0.7), 3.7 (*SD* = 0.7), and 2.6 (*SD* = 1.2) for reading, writing, and speaking abilities, respectively.
-
-Comparably, late learners (8 males, 6 females) were all born into hearing families and started learning TİD later in life after age 3, usually with the start of formal education.
-The age of sign language acquisition for late signers ranged between years 3-17 (*M* = 8.7 years, *SD* = 3.5 years).
-The parents of all participating late signers were hearing, and the reported home languages used in childhood mainly consisted of Turkish and home sign.
-The average age of late signers was 34.6 years (*SD* = 8 years) and ranged between 24 and 50.
-The derived mean number of years for TİD use following its acquisition was comparable to native signers with 25.9 years (*SD* = 7.9 years).
-Except for two signers who became deaf between the ages of 4 and 7, all the other late signers were pre-lingually deaf.
-All late signers completed middle school education and the average year of schooling was 12.3 years (*SD* = 2.8 years).
-All late participants attended at least one deaf school.
-While 8 participants attended two deaf schools (either primary and middle school, or middle and high school), 3 participants reported attend three deaf schools.
-Using the same scale, late signers self-rated their abilities in TİD and (spoken) Turkish.
-Both of their receptive and productive TİD abilities rated on average 4.9 points (*SD* = 0.3 points).
-Late signers' mean rating for their reading abilities in Turkish was 3.5 points (*SD* = 0.7 points) while their writing and speaking abilities rated on average 3.4 (*SD* = 0.5 points) and 3.1 points (*SD* = 0.7 points), respectively.
-
-Native and late participants' sex, average years of schooling, TİD use following its acquisition, and their self-reported receptive and productive TİD ratings, and their reading, writing, and speaking ratings for Turkish did not statistically differ; Bayesian *p*'s \> 0.1.
-However, participants' ages statistically differed between the two acquisition groups; *t*(27) = 3.4, Bayesian *p* \< 0.01.
-Overall, the native group consisted of younger participants.
-This was to ensure that the average years of TİD use among native signers approximated those of late signers.
-
-## Stimuli
-
-We used a story-telling paradigm to examine how native and late signers of TİD track reference in signed discourse.
-Participants were presented with 10 short video clips extracted from the cartoon *Tom and Jerry* and asked to retell them.
-Each clip had six to ten episodes (*M* = `r mean(c(6, 8, 8, 9, 9, 10, 8, 8, 7, 7))` episodes).
-The duration of the clips ranged from 17 seconds to 28 seconds (*M* = `r mean(c(19, 23, 28, 24, 22, 26, 17, 22, 23, 23))` seconds).
-The narratives included on average `r round(mean(c(2, 4, 2, 3, 2, 2, 3, 2, 2, 2)))` recurring animate characters for which the present context required the introduction, maintenance, and re-introduction of referents.
-In addition, in each narrative, there were on average `r round(mean(c(1, 1, 3, 2, 4, 4, 1, 3, 3, 3)))` inanimate objects with which the animate referents interacted.
-The detailed description of the episodes in each narrative can be found in \ref{section:b}.
-
-```{r tab-2, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warning=FALSE}
-library(magrittr)
-library(tidyverse)
-library(readr)
-library(kableExtra)
-library("papaja")
-library(stringr)
-options(scipen=1, digits=2)
-
-df_table_2 <- read.csv("./appendix/tables/table_tier.csv", sep = ";")
-apa_table(df_table_2,
-          caption ="Annotation Tiers with their Tags and Descriptions",
-          note = "NOM = nominal, PRO = pronominal, CL = classifier, NULL = null anaphora.",
-          escape = FALSE,
-          align = c("p{3.2cm}", rep("p{3.8cm}", ncol(df_table_2))),
-          placement = "p",
-          font_size = "footnotesize",
-          landscape = FALSE
-)
-```
-
-## Task Procedure
-
-Participants sat in front of a laptop computer which had the stimulus stories.
-The instructions were all given by a deaf research assistant who was present in all sessions and ensured that the task requirements were clear.
-The informed consent form was translated into TİD and explained to the participants by the same assistant before data recording.
-Participants filled in the consent form and a background questionnaire form before the onset of the trials.
-Following this step, signers were seated across a hearing researcher who is an advanced late learner of TİD.
-We instructed the participants to narrate as they would to a deaf friend.
-All trials were recorded by a digital camera placed in front of the signers.
-The participants were instructed to watch the narrative clip as many times as they needed, and narrate the content of each clip either to the experimenter or the recording digital camera\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining e did not make it obligatory for participants to either narrate to the digital camera or to the experimenter for few reasons.
+The age of the native participants ranged between 18 and 35 (\emph{M} =
+26.4 years, \emph{SD} = 5.0 years). This also corresponded to the total
+average years of TİD use for native signers. All native signers were
+pre-lingually deaf, that is they were either born deaf or they had
+become deaf before the age of 3, with the exception of one signer who
+became deaf between the ages of 4 and 7. Native signers reported to
+complete at least high school education and their level of education
+(LOE) was between 12 and 16 years (\emph{M} = 12.7 years, \emph{SD} =
+1.5 years). In addition, all native participants attended at least two
+deaf schools, one being primary and the other middle school. 10 native
+participants also attended a deaf school for their high school
+education. Based on a self-reported 5-point scale of proficiency (1:
+poor, 5: proficient), native participants declared that they have high
+receptive and productive skills in TİD (\emph{M} = 5 points, \emph{SD} =
+0 points). As for their proficiency in (spoken) Turkish, their
+self-reported rankings were on average 3.7 (\emph{SD} = 0.7), 3.7
+(\emph{SD} = 0.7), and 2.6 (\emph{SD} = 1.2) for reading, writing, and
+speaking abilities, respectively.
+
+Comparably, late learners (8 males, 6 females) were all born into
+hearing families and started learning TİD later in life after age 3,
+usually with the start of formal education. The age of sign language
+acquisition for late signers ranged between years 3-17 (\emph{M} = 8.7
+years, \emph{SD} = 3.5 years). The parents of all participating late
+signers were hearing, and the reported home languages used in childhood
+mainly consisted of Turkish and home sign. The average age of late
+signers was 34.6 years (\emph{SD} = 8 years) and ranged between 24 and
+50. The derived mean number of years for TİD use following its
+acquisition was comparable to native signers with 25.9 years (\emph{SD}
+= 7.9 years). Except for two signers who became deaf between the ages of
+4 and 7, all the other late signers were pre-lingually deaf. All late
+signers completed middle school education and the average year of
+schooling was 12.3 years (\emph{SD} = 2.8 years). All late participants
+attended at least one deaf school. While 8 participants attended two
+deaf schools (either primary and middle school, or middle and high
+school), 3 participants reported attend three deaf schools. Using the
+same scale, late signers self-rated their abilities in TİD and (spoken)
+Turkish. Both of their receptive and productive TİD abilities rated on
+average 4.9 points (\emph{SD} = 0.3 points). Late signers' mean rating
+for their reading abilities in Turkish was 3.5 points (\emph{SD} = 0.7
+points) while their writing and speaking abilities rated on average 3.4
+(\emph{SD} = 0.5 points) and 3.1 points (\emph{SD} = 0.7 points),
+respectively.
+
+Native and late participants' sex, average years of schooling, TİD use
+following its acquisition, and their self-reported receptive and
+productive TİD ratings, and their reading, writing, and speaking ratings
+for Turkish did not statistically differ; Bayesian \emph{p}'s
+\textgreater{} 0.1. However, participants' ages statistically differed
+between the two acquisition groups; \emph{t}(27) = 3.4, Bayesian
+\emph{p} \textless{} 0.01. Overall, the native group consisted of
+younger participants. This was to ensure that the average years of TİD
+use among native signers approximated those of late signers.
+
+\hypertarget{stimuli}{%
+\subsection{Stimuli}\label{stimuli}}
+
+We used a story-telling paradigm to examine how native and late signers
+of TİD track reference in signed discourse. Participants were presented
+with 10 short video clips extracted from the cartoon \emph{Tom and
+Jerry} and asked to retell them. Each clip had six to ten episodes
+(\emph{M} = 8 episodes). The duration of the clips ranged from 17
+seconds to 28 seconds (\emph{M} = 22.7 seconds). The narratives included
+on average 2 recurring animate characters for which the present context
+required the introduction, maintenance, and re-introduction of
+referents. In addition, in each narrative, there were on average 2
+inanimate objects with which the animate referents interacted. The
+detailed description of the episodes in each narrative can be found in
+\ref{section:b}.
+
+\begin{table}[p]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:tab-2}Annotation Tiers with their Tags and Descriptions}
+
+\footnotesize{
+
+\begin{tabular}{p{3.2cm}p{3.8cm}p{3.8cm}}
+\toprule
+Tier & \multicolumn{1}{c}{Tag} & \multicolumn{1}{c}{Description}\\
+\midrule
+Clause Boundary &  & The boundary between clauses indicated by the presence of predicates and/or certain prosodic cues (e.g., head nod)\\
+Discourse Status & Introduced & A first mention of an entity in a given discourse regardless of any syntactic role\\
+ & Maintained & A subject referent continued across two or more immediate clauses\\
+ & Reintroduced & A subject referent discontinued across at least one clause\\
+REAT Type & NOM: Bare Noun & A non-modified noun phrase (e.g., CAT, MOUSE)\\
+ & NOM: Modified Noun & A noun phrase modified by an index sign or an adjective\\
+ & NOM: Fingerspelled Noun & Borrowed nouns from Turkish spelled by the manual alphabet (e.g., K-E-D-İ ‘cat’)\\
+ & PRO: Pronominal IX & An index sign used in isolation to refer to an entity in discourse\\
+ & CL: Whole Entity Classifier & A manual construction that refers to the entirety of an entity\\
+ & CL: Body Part Classifier & A manual construction that refers to a part of an entity\\
+ & CL: Handling Classifier & A manual construction that refers to the handling of an item\\
+ & CL: Extension Classifier & A manual construction that denotes the size and shape of an entity\\
+ & NULL: Constructed Action & A multifunctional construction where the signers take on the role of a referent and perform their actions\\
+ & NULL: Plain Verb & A non-agreeing verb (e.g., LOVE)\\
+ & NULL: Agreement Verb & A verbal construction the beginning and ending of which agree with person and number\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} NOM = nominal, PRO = pronominal, CL = classifier, NULL = null anaphora.}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\hypertarget{task-procedure}{%
+\subsection{Task Procedure}\label{task-procedure}}
+
+Participants sat in front of a laptop computer which had the stimulus
+stories. The instructions were all given by a deaf research assistant
+who was present in all sessions and ensured that the task requirements
+were clear. The informed consent form was translated into TİD and
+explained to the participants by the same assistant before data
+recording. Participants filled in the consent form and a background
+questionnaire form before the onset of the trials. Following this step,
+signers were seated across a hearing researcher who is an advanced late
+learner of TİD. We instructed the participants to narrate as they would
+to a deaf friend. All trials were recorded by a digital camera placed in
+front of the signers. The participants were instructed to watch the
+narrative clip as many times as they needed, and narrate the content of
+each clip either to the experimenter or the recording digital
+camera\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining e did not make it obligatory for participants to either narrate to the digital camera or to the experimenter for few reasons.
 First, a native or native-like deaf experimenter was not always available, and a hearing late signer of TİD attended some sessions as the experimenter.
@@
 
-```{r echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='asis'}
-# generic libraries
-library(readxl) #read excel
-library(gdata) # reorder.factor
-library(magrittr) # pipe
-library(tidyverse) # tidy
-library(janitor)
-library(MASS)
-theme_set(theme_bw()) #set ggplot theme
-library(extrafont)
-loadfonts()
-
-# statistics libraries
-library(brms) # bayesian modeling
-library(bayesplot) # plot options for bayesian models
-bayesplot_theme_set(new = theme_bw()) # set plot theme
-library(tidybayes)
-library(bayestestR)
-theme_apa <- function(base_size = 10, base_family = "serif", box = FALSE) {
-  adapted_theme <- ggplot2::theme_bw(base_size, base_family) +
-    ggplot2::theme(legend.position = "bottom",
-                   plot.title = ggplot2::element_text(size = ggplot2::rel(1.1), margin = ggplot2::margin(0, 0, ggplot2::rel(14), 0))
-                   # , axis.title = ggplot2::element_text(size = ggplot2::rel(1.1))
-                   , axis.title.x = ggplot2::element_text(size = ggplot2::rel(1), lineheight = ggplot2::rel(1.1), margin = ggplot2::margin(ggplot2::rel(12), 0, 0, 0))
-                   , axis.title.y = ggplot2::element_text(size = ggplot2::rel(1), lineheight = ggplot2::rel(1.1), margin = ggplot2::margin(0, ggplot2::rel(12), 0, 0))
-                   , axis.title.y.right = ggplot2::element_text(size = ggplot2::rel(1), lineheight = ggplot2::rel(1.1), margin = ggplot2::margin(0, 0, 0, ggplot2::rel(12)))
-                   , axis.ticks.length = ggplot2::unit(ggplot2::rel(6), "points")
-                   , axis.text = ggplot2::element_text(size = ggplot2::rel(0.9))
-                   , axis.text.x = ggplot2::element_text(size = ggplot2::rel(1), margin = ggplot2::margin(ggplot2::rel(6), 0, 0, 0))
-                   , axis.text.y = ggplot2::element_text(size = ggplot2::rel(1), margin = ggplot2::margin(0, ggplot2::rel(6), 0, 0))
-                   , axis.text.y.right = ggplot2::element_text(size = ggplot2::rel(1), margin = ggplot2::margin(0, 0, 0, ggplot2::rel(6)))
-                   , axis.line = ggplot2::element_line()
-                   # , axis.line.x = ggplot2::element_line()
-                   # , axis.line.y = ggplot2::element_line()
-                   , legend.title = ggplot2::element_text()
-                   , legend.key = ggplot2::element_rect(fill = NA, color = NA)
-                   , legend.key.width = ggplot2::unit(ggplot2::rel(20), "points")
-                   , legend.key.height = ggplot2::unit(ggplot2::rel(20), "points")
-                   , legend.margin = ggplot2::margin(
-                     t = ggplot2::rel(16)
-                     , r = ggplot2::rel(16)
-                     , b = ggplot2::rel(16)
-                     , l = ggplot2::rel(16)
-                     , unit = "points"
-                   )
-                   , panel.spacing = ggplot2::unit(ggplot2::rel(14), "points")
-                   , panel.grid.major.x = ggplot2::element_blank()
-                   , panel.grid.minor.x = ggplot2::element_blank()
-                   , panel.grid.major.y = ggplot2::element_blank()
-                   , panel.grid.minor.y = ggplot2::element_blank()
-                   , strip.background = ggplot2::element_rect(fill = NA, color = NA)
-                   , strip.text.x = ggplot2::element_text(size = ggplot2::rel(1.2), margin = ggplot2::margin(0, 0, ggplot2::rel(10), 0))
-                   , strip.text.y = ggplot2::element_text(size = ggplot2::rel(1.2), margin = ggplot2::margin(0, 0, 0, ggplot2::rel(10)))
-    )
-  if(box) {
-    adapted_theme <- adapted_theme + ggplot2::theme(panel.border = ggplot2::element_rect(color = "black"))
-  } else {
-    adapted_theme <- adapted_theme + ggplot2::theme(panel.border = ggplot2::element_blank())
-  }
-  adapted_theme
-}
-
-# Read df and drop NAs
-df_narration <- read.csv("data/narration_data.csv", sep = ";") %>% tidyr::drop_na()
-
-df_misc <- read.csv("./appendix/appendix_a_table.csv", sep=";") %>% clean_names() %>% 
-  dplyr::mutate(Subject = as.integer(no),
-                Group = case_when(aoa == "0-3" ~ "native",
-                                  aoa == "4-7" ~ "late1",
-                                  T ~ "late2") %>% as.factor()) %>% dplyr::select(Subject, Group)
-df_narration %<>% left_join(df_misc)
-
-# Coding nominal, Zero, Classifier, and Pronominal categories
-df_narration$ref_type <- with(df_narration, case_when(
-  GTR1 %in% c("Bare noun", "Point noun", "Noun Point",
-              "Fingerspelled noun", "CL noun", "Noun CL") ~ "Nominal",
-  GTR1 %in% c("Constructed action", "Constructed Action", "Plain verb", "Person agreement verb") ~ "Null",
-  GTR1 %in%  c("Whole entity CL", "Body part CL",
-               "SASS - Tracing", "SASS - Static") ~ "Classifier",
-  T ~ "Pronominal"))
-
-# Coding nominal, Zero, Classifier, and Pronominal categories
-df_narration$GTR1_plot <- with(df_narration, case_when(
-  GTR1 %in% c("Bare noun") ~ "BARE NOM",
-  GTR1 %in% c("Point noun", "Noun Point","CL noun", "Noun CL") ~ "M NOM",
-  GTR1 %in% c("Point noun", "Noun Point","CL noun", "Noun CL") ~ "FS NOM",
-  GTR1 %in% c("Constructed action", "Constructed Action") ~ "CA",
-  GTR1 %in% c("Plain verb") ~ "PLAIN",
-  GTR1 %in% c("Person agreement verb") ~ "AGR",
-  GTR1 %in% c("Whole entity CL") ~ "WECL",
-  GTR1 %in%  c("Body part CL")  ~ "BPCL",
-  GTR1 %in%  c("SASS - Tracing", "SASS - Static")  ~ "SASS",
-  T ~ "Pronominal"))
-
-# Encode vector types
-df_narration$Narrative %<>% as.integer() 
-df_narration$Subject %<>% as.integer()
-df_narration$GTR1 %<>% as.factor() 
-df_narration$Nativeness %<>% dplyr::recode(`0` = "Late", `1` = "Native") %>% as.factor() %>% reorder.factor(new.order = c("Native","Late"))
-df_narration$Discourse %<>% as.factor()
-df_narration$ref_type %<>% as.factor()
-
-# Summary of the data
-summary_df <- df_narration %>% 
-  group_by(Nativeness, ref_type) %>% 
-  summarise(Count = n()) %>% 
-  mutate(Percentage = formattable::percent(Count/sum(Count))) %>% 
-  arrange(desc(Percentage))
-
-# Descriptive plots
-df_narration$ref_type2 <- df_narration$ref_type %>% recode('Classifier' = 'CL', 'Null' = 'NULL',
-                                                           'Pronominal' = 'PRO', 'Nominal' = 'NOM')
-# dropping Pronominal references because they are too few.
-df_model1 <- df_narration %>% subset(ref_type !="Pronominal") %>% drop.levels()
-
-# Contrast coding
-df_model1$Nativeness %<>% reorder.factor(new.order = c("Native","Late"))
-contrasts(df_model1$Nativeness) <- contr.sum(2)/2
-contrasts(df_model1$Discourse) <- contr.sum(3)/2
-contrasts(df_model1$Group) <- contr.sdif(3)
-
-# counts of ref types
-ref_count <- df_narration %>% group_by(ref_type) %>% summarise(count=n())
-
-# Model for reference type
-model_1 <- brm(ref_type ~ Nativeness*Discourse + (1|Subject), 
-               data = df_model1, family = categorical(link = "logit"), 
-               iter = 3000, warmup = 2000, chains = 4, cores = 4, file ="./rds/model_1.rds")
-
-# Posterior plot for reference type
-df_model_1 <- mcmc_intervals_data(model_1, pars = vars(starts_with("b_")))
-df_model_1$parameter %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
-                                        `b_muNull_Intercept` = "Null_Intercept",
-                                        `b_muNominal_Nativeness1` = "Nom_Native",
-                                        `b_muNull_Nativeness1` = "Null_Native",
-                                        `b_muNominal_Discourse1` = "Nom_Intro",
-                                        `b_muNominal_Discourse2` = "Nom_Maintain",
-                                        'b_muNominal_Nativeness1:Discourse1' = 'Nom_Native*Intro',
-                                        'b_muNominal_Nativeness1:Discourse2' = 'Nom_Native*Maintain',
-                                        'b_muNull_Nativeness1:Discourse1' = 'Null_Native*Intro',
-                                        'b_muNull_Nativeness1:Discourse2' = 'Null_Native*Maintain',
-                                        `b_muNull_Discourse1` = "Null_Intro",
-                                        `b_muNull_Discourse2` = "Null_Maintain")
-saveRDS(df_model_1, "./rds/df_model_1.rds")
-
-# Supplementary model with 3 groups of acquisition
-model_3 <- brm(ref_type ~ Group*Discourse + (1|Subject), 
-               data = df_model1, family = categorical(link = "logit"), 
-               iter = 3000, warmup = 2000, chains = 4, cores = 4, file ="./rds/model_3.rds")
-
-# Posterior plot for reference type
-df_model_3 <- mcmc_intervals_data(model_3, pars = vars(starts_with("b_")))
-df_model_3$parameter %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
-                                        `b_muNull_Intercept` = "Null_Intercept",
-                                        `b_muNominal_Group2M1` = "Nom_Late1vsNative",
-                                        `b_muNominal_Group3M2` = "Nom_Late2vsLate1",
-                                        `b_muNominal_Discourse1` = "Nom_Intro",
-                                        `b_muNominal_Discourse2` = "Nom_Maintain",
-                                        `b_muNominal_Group2M1:Discourse1` = 'Nom_L1vsN*Intro',
-                                        `b_muNominal_Group3M2:Discourse1` = 'Nom_L2vsL1*Intro',
-                                        `b_muNominal_Group2M1:Discourse2` = 'Nom_L1vsN*Maintain',
-                                        `b_muNominal_Group3M2:Discourse2` = 'Nom_L2vsL1*Maintain',
-                                        `b_muNull_Group2M1` = "Null_Late1vsNative",
-                                        `b_muNull_Group3M2` = "Null_Late2vsLate1",
-                                        `b_muNull_Discourse1` = "Null_Intro",
-                                        `b_muNull_Discourse2` = "Null_Maintain",
-                                        `b_muNull_Group2M1:Discourse1` = 'Null_L1vsN*Intro',
-                                        `b_muNull_Group3M2:Discourse1` = 'Null_L2vsL1*Intro',
-                                        `b_muNull_Group2M1:Discourse2` = 'Null_L1vsN*Maintain',
-                                        `b_muNull_Group3M2:Discourse2` = 'Null_L2vsL1*Maintain')
-saveRDS(df_model_3, "./rds/df_model_3.rds")
-#df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "NULL"))
-
-```
-
-## Data Annotation
-
-We coded the discourse status (Introduction, Maintenance, Re-introduction) and RE type (Nominal, Pronominal, Classifier, Null).
-We used ELAN Linguistic Annotation Software [@crasborn2008] for data annotation.
-First Author who is a hearing signer of TİD annotated all the data and translated the narratives into Turkish together with a native deaf signer of TİD.
-
-Following @berman2013relating, we first identified the clause boundaries in each sentence indicated by the presence of predicates and informative non-manual markers (e.g., head nod).
-On the second tier, we coded the discourse status for each referent.
-Having adapted the classification of @gullberg2006, we considered all referents mentioned for the first time as introduced regardless of their syntactic function in the sentence.
-Only subjects were coded for maintenance and re-introduction.
-We coded subject referents as maintained only if they had been referred to in the previous clause.
-To this end, the referents which had been discontinued across one or more sentences were all regarded re-introduced into the discourse.
-On a third tier, we had the RE type for a single reference with the main categories nominal, pronominal, null marking, and classifier.
-Under nominal, we identified bare nouns, modified nouns, and fingerspelled nouns.
-For classifiers, we distinguished among three types of constructions: WECL, BPCL, and ExtCL.
-Finally, we had CA, agreement, plain tags for null constructions.
-We were unable to observe any instance of HCLs occurring in isolation (i.e., not part of constructed action).
-The annotation tiers are presented in Table \ref{tab:tab-2}.
-
-## Data Analysis
-
-In our analysis, we use Bayesian regression models with the help of the brms package [@bürkner2018] in R.
-In a Bayesian regression the prior and the likelihood is used to calculate the posterior.
-Instead of a single filter of significance, we get a posterior probability distribution.
-We do not define our own priors and use the brms package defaults.
-We are using Bayesian regression to avoid some common pitfalls of other approaches [@wagenmakers2007; @kruschke2011].
-
-For this particular type of experimental data, we fit the responses to a multinomial distribution.
-Most of the time, experimental results have yes/no-true/false responses which are fit to binomial distributions, reading/response times which are fit to log-normal distributions, and ratings which are fit to cumulative distributions.
-In these cases, the researcher is not particularly interested in the \`intercept' coefficient which shows the mean value of the response.
-The researcher is usually interested in the remaining coefficients which give information about predictor level comparisons.
-
-In a multinomial distribution, the responses are in a categorical relationship where the levels can be A,B,C, and more.
-These levels are not ordered in time, they are not cumulative, and they are not additive.
-If we were to make an analogy, they are like countries on Earth which can have differences and similarities, but they don't have to.
-In a multinomial distribution, one of the responses becomes the level of comparison and other levels are compared against it.
-In a way, there are multiple binomial distributions.
-However, the intercepts in a multinomial distribution end up having more of an interest for the researcher.
-This is because the intercept is a comparison of two levels of the response to the exclusion of the remaining response levels.
-There is another consideration that should go into interpreting the results of a multinomial distribution.
-Different from the intercepts, the remaining coefficients are predictor level comparisons within the grand comparison of the response levels.
-To indicate which predictor level comparison belongs to which response comparison, we name our coefficients with prefixes followed by an underscore.
-
-In our model plots, we report median estimates as points, 50% credible intervals as a thick line and 95% credible intervals as a thin line.
-These numbers are relatively arbitrary, and they are intended to show the posterior probability distribution of the coefficients.
-We use interval plots to save space and have a more compact plot.
-If a considerable amount (\>90%) of the posterior probability distribution is towards a sign (-/+) we interpret it as a categorial effect, if a good amount (\>50%) of the distribution is towards a sign (-/+) we interpret it as a tendency, and we see the remaining distributions as indifference.
-In addition to these, we look at how the distribution skewed in the extremes or how \`wide' it is from the median estimate to inform us mainly about uncertainty and variation.
-
-# Results
-
-We collected `r nrow(df_narration)` data points as a result of the experiment.
-Each data point is a referent tracked by the participant.
-Each of the REs is classified in terms of its Discourse Status (Introduction, Maintenance, Re-introduction), its Type (Nominal, Pronominal, Classifer, Null)\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer Null).
-However, we allude to the numerical distribution of the subtypes in this and following sections}, and Participant Group (Native, Late)\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17).
+\hypertarget{data-annotation}{%
+\subsection{Data Annotation}\label{data-annotation}}
+
+We coded the discourse status (Introduction, Maintenance,
+Re-introduction) and RE type (Nominal, Pronominal, Classifier, Null). We
+used ELAN Linguistic Annotation Software (Crasborn and Sloetjes, 2008)
+for data annotation. First Author who is a hearing signer of TİD
+annotated all the data and translated the narratives into Turkish
+together with a native deaf signer of TİD.
+
+Following Berman and Slobin (2013), we first identified the clause
+boundaries in each sentence indicated by the presence of predicates and
+informative non-manual markers (e.g., head nod). On the second tier, we
+coded the discourse status for each referent. Having adapted the
+classification of Gullberg (2006), we considered all referents mentioned
+for the first time as introduced regardless of their syntactic function
+in the sentence. Only subjects were coded for maintenance and
+re-introduction. We coded subject referents as maintained only if they
+had been referred to in the previous clause. To this end, the referents
+which had been discontinued across one or more sentences were all
+regarded re-introduced into the discourse. On a third tier, we had the
+RE type for a single reference with the main categories nominal,
+pronominal, null marking, and classifier. Under nominal, we identified
+bare nouns, modified nouns, and fingerspelled nouns. For classifiers, we
+distinguished among three types of constructions: WECL, BPCL, and ExtCL.
+Finally, we had CA, agreement, plain tags for null constructions. We
+were unable to observe any instance of HCLs occurring in isolation
+(i.e., not part of constructed action). The annotation tiers are
+presented in Table \ref{tab:tab-2}.
+
+\hypertarget{data-analysis}{%
+\subsection{Data Analysis}\label{data-analysis}}
+
+In our analysis, we use Bayesian regression models with the help of the
+brms package (Bürkner, 2018) in R. In a Bayesian regression the prior
+and the likelihood is used to calculate the posterior. Instead of a
+single filter of significance, we get a posterior probability
+distribution. We do not define our own priors and use the brms package
+defaults. We are using Bayesian regression to avoid some common pitfalls
+of other approaches (Kruschke, 2011; Wagenmakers, 2007).
+
+For this particular type of experimental data, we fit the responses to a
+multinomial distribution. Most of the time, experimental results have
+yes/no-true/false responses which are fit to binomial distributions,
+reading/response times which are fit to log-normal distributions, and
+ratings which are fit to cumulative distributions. In these cases, the
+researcher is not particularly interested in the `intercept' coefficient
+which shows the mean value of the response. The researcher is usually
+interested in the remaining coefficients which give information about
+predictor level comparisons.
+
+In a multinomial distribution, the responses are in a categorical
+relationship where the levels can be A,B,C, and more. These levels are
+not ordered in time, they are not cumulative, and they are not additive.
+If we were to make an analogy, they are like countries on Earth which
+can have differences and similarities, but they don't have to. In a
+multinomial distribution, one of the responses becomes the level of
+comparison and other levels are compared against it. In a way, there are
+multiple binomial distributions. However, the intercepts in a
+multinomial distribution end up having more of an interest for the
+researcher. This is because the intercept is a comparison of two levels
+of the response to the exclusion of the remaining response levels. There
+is another consideration that should go into interpreting the results of
+a multinomial distribution. Different from the intercepts, the remaining
+coefficients are predictor level comparisons within the grand comparison
+of the response levels. To indicate which predictor level comparison
+belongs to which response comparison, we name our coefficients with
+prefixes followed by an underscore.
+
+In our model plots, we report median estimates as points, 50\% credible
+intervals as a thick line and 95\% credible intervals as a thin line.
+These numbers are relatively arbitrary, and they are intended to show
+the posterior probability distribution of the coefficients. We use
+interval plots to save space and have a more compact plot. If a
+considerable amount (\textgreater90\%) of the posterior probability
+distribution is towards a sign (-/+) we interpret it as a categorial
+effect, if a good amount (\textgreater50\%) of the distribution is
+towards a sign (-/+) we interpret it as a tendency, and we see the
+remaining distributions as indifference. In addition to these, we look
+at how the distribution skewed in the extremes or how `wide' it is from
+the median estimate to inform us mainly about uncertainty and variation.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+We collected 4142 data points as a result of the experiment. Each data
+point is a referent tracked by the participant. Each of the REs is
+classified in terms of its Discourse Status (Introduction, Maintenance,
+Re-introduction), its Type (Nominal, Pronominal, Classifer,
+Null)\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer Null).
+However, we allude to the numerical distribution of the subtypes in this and following sections},
+and Participant Group (Native,
+Late)\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17).
 Treating age of acquisition as a continous variable would not be informative due to the small sample size in each interval.}.
 
-Across our acquisition groups, we observed `r sum(df_narration$Discourse == "Introduction")` (`r sum(df_narration$Discourse == "Introduction")/nrow(df_narration)`) instances of referent introduction, `r sum(df_narration$Discourse == "Maintenance")` (`r sum(df_narration$Discourse == "Maintenance")/nrow(df_narration)`) instances of referent maintenance, and `r sum(df_narration$Discourse == "Re-introduction")` (`r sum(df_narration$Discourse == "Re-introduction")/nrow(df_narration)`) instances of referent re-introduction.
-For the introduction of a referent, the majority of the REs were nominal constructions (Table \ref{tab:my-table}).
-However, we observed a sporadic use of classifiers and null marking respectively.
-When the discourse status was maintained, signers extensively used null markers to refer to the entities in the immediately previous clause.
-Classifier constructions followed null markers in frequency in maintained contexts.
-We observed only a few instances of nominal use for referent maintenance.
-In re-introduction contexts, signers used a greater proportion of nominal markings which were followed by null, classifier, and pronominal use.
-
-```{=tex}
+Across our acquisition groups, we observed 1066 (0.26) instances of
+referent introduction, 1651 (0.4) instances of referent maintenance, and
+1425 (0.34) instances of referent re-introduction. For the introduction
+of a referent, the majority of the REs were nominal constructions (Table
+\ref{tab:my-table}). However, we observed a sporadic use of classifiers
+and null marking respectively. When the discourse status was maintained,
+signers extensively used null markers to refer to the entities in the
+immediately previous clause. Classifier constructions followed null
+markers in frequency in maintained contexts. We observed only a few
+instances of nominal use for referent maintenance. In re-introduction
+contexts, signers used a greater proportion of nominal markings which
+were followed by null, classifier, and pronominal use.
+
 \begin{table}
@@
 \end{table}
-```
-Table \ref{tab:my-table2} shows that for all discourse statuses the most numerous nominal type was \textsc{bare nom} which was followed by nouns modified by index or classifiers.
-In number, signers used \textsc{mod nom} more in introduction and re-introduction compared to maintenance settings.
-Signers produced a total number of `r sum(df_narration$ref_type == "Pronominal")` pronouns (Table \ref{tab:my-table3}).
-A great majority of the pronominal constructions were used to re-introduce a referent into the discourse.
-We observed only a very small amount of pronoun use for referent introduction and maintenance.
-
-```{=tex}
+
+Table \ref{tab:my-table2} shows that for all discourse statuses the most
+numerous nominal type was \textsc{bare nom} which was followed by nouns
+modified by index or classifiers. In number, signers used
+\textsc{mod nom} more in introduction and re-introduction compared to
+maintenance settings. Signers produced a total number of 40 pronouns
+(Table \ref{tab:my-table3}). A great majority of the pronominal
+constructions were used to re-introduce a referent into the discourse.
+We observed only a very small amount of pronoun use for referent
+introduction and maintenance.
+
 \begin{table}
@@
 \end{table}
-```
-As for the classifier constructions (Table \ref{tab:my-table4}), WECL was the most common type, followed by BPCL and ExtCL.
-For the first mention of a referent, signers mostly preferred to use ExtCL.
-However, for referent maintenance and re-introduction WECL was used more compared to BPCL and ExtCL.
-
-Among the null markers, CA was most frequently employed, and it was followed by plain verbs and agreement verbs (Table \ref{tab:my-table5}).
-Overall, signers used very few null markings for the introduction of a referent and the majority of the existing ones were CA.
-Additionally, signers used CA more for referent re-introduction than maintenance.
-We observed a similar instances of agreement verbs for maintained and re-introduced contexts.
-In contrast, plain verbs were favored more in the maintenance setting than re-introduction.
-
-```{=tex}
+
+As for the classifier constructions (Table \ref{tab:my-table4}), WECL
+was the most common type, followed by BPCL and ExtCL. For the first
+mention of a referent, signers mostly preferred to use ExtCL. However,
+for referent maintenance and re-introduction WECL was used more compared
+to BPCL and ExtCL.
+
+Among the null markers, CA was most frequently employed, and it was
+followed by plain verbs and agreement verbs (Table \ref{tab:my-table5}).
+Overall, signers used very few null markings for the introduction of a
+referent and the majority of the existing ones were CA. Additionally,
+signers used CA more for referent re-introduction than maintenance. We
+observed a similar instances of agreement verbs for maintained and
+re-introduced contexts. In contrast, plain verbs were favored more in
+the maintenance setting than re-introduction.
+
 \begin{table}
@@
 \end{table}
-```
-Across the discourse statuses, there is also a difference between participant groups in the use of null marking for reference tracking.
-Native signers of TİD used it more (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "Native") / nrow(df_narration), digits=2)`) compared to the late signers (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "Late") / nrow(df_narration), digits=2)`).
-Comparably, late signers used nominals more frequently (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "Late") / nrow(df_narration), digits=2)`) than native signers (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "Native") / nrow(df_narration), digits=2)`).
-Figure 1 shows the numerical distribution of the responses by RE type and Group.
-```{=tex}
+
+Across the discourse statuses, there is also a difference between
+participant groups in the use of null marking for reference tracking.
+Native signers of TİD used it more (0.23) compared to the late signers
+(0.16). Comparably, late signers used nominals more frequently (0.25)
+than native signers (0.23). Figure 1 shows the numerical distribution of
+the responses by RE type and Group.
+
 \begin{table}
@@
 \end{table}
-```
-
-```{r fig:count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), RE Type (Nominal, Pronominal, Classifier, Null), and Participant Group (Native, Late)"}
-df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "NULL"))
-
-p1 <- df_narration %>%
-  dplyr::select(Nativeness, ref_type2, Discourse) %>%
-  group_by(Nativeness, Discourse) %>%
-  ggplot(aes(ref_type2, fill = Nativeness)) + geom_bar(position = "dodge") +
-  stat_count(geom = "text", aes(color = Nativeness, label = ..count..), vjust = 0,
-             position = position_dodge(width = 0.85), size = 2) +
-  facet_grid(~Discourse) + ylab("Count") + xlab("REATs") +
-  scale_color_grey() +
-  scale_fill_grey() +
-  theme_apa() 
-p1
-```
-
-For more inference on the data and having pairwise comparisons, we fit a regression model to RE Type using Discourse Status and Group as predictors.
-We used the brms package [@bürkner2018] in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late).
-As the overall number of produced pronouns was very little (40 data points), we excluded pronominal use from the following analyses.
-For the current analysis, we did not define priors ourselves but used the preset uninformed priors.
-Figure 2[^2] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals.
-There are 3 responses we are considering in this model, and they are Classifier, Nominative, and Null.
-Classifier response is the level that other responses Nominative and Null are compared against.
-
-[^2]: We ran another model where we used the Nativeness predictor with 3 levels 0-3:Native, 4-7:Late1, and 7-above:Late2.
-However, our model did not converg properly.
-We provide the model results as a plot and table in the Appendix.
-
-```{r fig:model-gram, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
-p3 <- df_model_1 %>% 
-  ggplot(aes(m, y=factor(parameter,levels = rev(levels(factor(parameter)))))) +
-  geom_point(size = 2.5) +
-  geom_errorbarh(aes(xmin = l, xmax = h), alpha = 1, height = 0, size = 1.5) +
-  geom_errorbarh(aes(xmin = ll, xmax = hh, height = 0)) + vline_0(linetype = "dashed") +
-  xlab("Estimate (log odds)") + ylab("Coefficients") + theme_bw() + theme_apa()
-p3
-```
-
-In a multinomial distribution, the model takes one of the responses as a reference and compares all other responses to it.
-For example, the first coefficient, Nom_Intercept (nominal intercept), in Figure 2 is the comparison of a nominal versus a classifier.
-It means that participants produced more nominals than classifiers in general regardless of participant group and discourse status.
-The second coefficient shows that null markings were also used more than classifiers by the participants in general, but not as much as nominals.
-
-In the comparison of nominals and classifiers, being a native acquirer of TİD did not have an effect.
-Nominals were used more when the discourse status was introduction compared to re-introduction.
-Again, the whole comparison is made against a classifier.
-In contrast to the introduction status, nominals were used less in maintenance contexts compared to the re-introduction contexts within the greater comparison to classifier referents.
-There does not seem to be an interaction between the introduction status of discourse and being a native acquirer of TİD.
-However, there seems to be an interaction between maintaining a referent and being a native acquirer of TİD.
-In that, there were fewer nominal referents when it was a native acquirer of TİD and the discourse context was maintenance.
-
-Compared to classifiers, being a native acquirer of TİD increased the odds of using null marking in general.
-Null constructions were used less in introduction contexts than in re-introduction.
-This contrasts with the nominals (4th coefficient) as they were used more in referent introduction versus re-introduction.
-In contrast to introduction contexts, null markers were used more in referent maintenance compared to re-introduction, yet again within the greater comparison to the classifiers.
-There seems to be an interaction between introducing a referent and being a native acquirer of TİD.
-In that, the odds of using null marking increases if the participant is a native acquirer of TİD and the referent is used in an introduction context.
-There is another interaction with regards to referent maintenance, contrasting the first interaction.
-In that, the odds of using null markers decrease if the participant is a native acquirer of TİD and the referent is maintained.
-This does not mean that late signers used more null markers in maintenance contexts.
-It means that the use of null markers did not linearly increase as the addition of using a null marker instead of a classifier as a native plus using null markers instead of a classifier for maintaining a referent versus re-introducing it.
-
-# General Discussion
-
-In our study, we aimed to (i) describe how signers track referents in signed discourse in TİD and (ii) examine whether language deprivation results in more over-redundant referent selection.
-To this end, we investigated the distribution of referring expressions (REs) that the native and late signers select when employing referents in different discourse statuses (introduced, maintained, and re-introduced).
-
-In our first set of hypotheses regarding referent tracking strategies, we had the general expectation that signers would use more explicit forms (mainly nominals) for referent introduction and less implicit forms such as null marking, pronominal use, and classifiers.
-In re-introduced contexts, we predicted that signers would use both explicit and implicit markers depending on the accessibility of the tracked referent.
-
-For age of acquisition effects, based on what is proposed for child and adult deaf signers with language deprivation in terms of their morphosyntactic [@boudreault2006; @cormier2012; @kayaba2022] and narrative development [@becker2009; @cormier2013], we expected that late signers would be more over-redundant in their referent tracking, akin to second language learners [@frederiksen2019; @bel2015].
-That is, we had the expectation that late signers would use more explicit markers like nouns and fewer implicit markers like null marking even for highly accessible referents (when the discourse status is maintained or even re-introduced in same cases) than the control group (i.e., native signers).
-
-## Tracking referents in TİD
-
-The results of the present study partly support our hypotheses.
-In line with our expectations, regardless of their acquisition groups (native or late) (i) signers mainly used nominals to introduce a referent into the discourse, (ii) used null marking and classifiers but not pronouns for maintained contexts, and (iii) employed both nominals and less explicit markers (classifiers, null markers and pronouns) to re-introduce a referent.
-
-However, the proportion of the observed discourse statuses (26% for introduction, 40% for maintenance, 34% for re-introduction across acquisition groups) in this study differed from that of some other studies.
-For a comparison, while the percentage of introduced referents were similar in both research (26% versus 24%), the participants in the present study brought back referents into the discourse more frequently compared to the participants in @frederiksen2019 and @frederiksen2016 which reported that re-introduction contexts accounted for only 7% of the referents in ASL.
-On the other hand, maintenance percentages were lower in the present study compared to @frederiksen2016 (40% versus 69%).
-Investigating the distribution of REs across discourse statuses in ASL, a similar study by @czubek2017 also reported contrasting numbers (17% for introduction, 64% for maintenance, and 19% for re-introduction).
-We suggest that the discrepancy might be due to the nature of the stimuli used in these research.
-In @frederiksen2016, the authors used 4-6-episode-long picture story with 2 animate characters and 1 inanimate object.
-In contrast, the stimuli in the present study consisted of 8-episode-long video clips on average with a mean number of 2 animate characters and 3 inanimate objects.
-Although the number of animate characters and inanimate characters in @czubek2017's stimuli matched that of the present study, their narrative structure (a 6-episode-long picture story) was still shorter compared to our narrative stimuli.
-We believe fewer possible referents and/or a simpler narrative structure in these studies may have contributed to the lower number of re-introduced contexts compared to the present study.
-Due to the story length and composition in this study, the participants had to track more recurring animate or inanimate referents for a longer period of time.
-As a result, our signers engaged in referent re-introduction more than what was reported in previous studies [@czubek2017; @frederiksen2016].
-
-As for the distribution of REs across different discourses statuses, we observed that signers overwhelmingly preferred to use nominals over other REs for the first mention of a referent and used null markers such as CA, and agreement and plain verb for referents with high accessibility (i.e., when the discourse status was maintained).
-This finding parallels other accounts of referential accessibility suggested for other sign languages [see @frederiksen2016; @czubek2017; and @swabey2002; for ASL, @morgan2005; for BSL @bel2015; and @barberà2009 for LSC.
-Our results also indicate that modified nouns (e.g., nouns that are preceded or followed by a classifier or IX) appeared more in introduced contexts compared to re-introduced contexts.
-This brings along the assumption that modified nouns are used more to track referents with lower accessibility, probably due to their more explicit structure.
-
-Interestingly and to our surprise, the signers in this study used pronominal IX very sporadically with only 40 observed instances that were mainly used to re-introduce referents rather than maintain them.
-This finding is comparable with what has been reported for ASL [@swabey2002; @frederiksen2016; @czubek2017] but not with LSC [@bel2015], BSL [@morgan2005] and DGS [@perniss2015], in which the signers extensively made overt pronominal references (up to 30%) to maintain entities.
-Comparable to our findings, It has been reported that pronouns do not frequently occur in the local context and instead follows a topic shift and thus rank higher than classifiers and null markers different genre [see @barberà2009 for LSC; and @nuhbalaoglu2018 for TİD].
-Considering this, the claim that the IX sign functions like demonstratives that are also used to track referents in re-introduced contexts has not been borne out [@koulidobrova2016].
-Similar to these research, our findings indicate that the pronominal IX in TİD was used more concurrently with nominal constructions and pronominal IX was used very infrequently in isolation.
-The observation that overt pronouns are rare in ASL narratives comes from studies that include a simple picture retelling task with one protagonist [e.g., @frederiksen2016] as well as a more complex narrative design as in [@czubek2017].
-However, @czubek2017's study also incorporated different genres and reported more instances of overt pronouns in personal narratives (compared to picture or video retelling).
-In fact, this finding suggests that the use of pronouns might be genre-dependent.
-Moreover, another production study on referent tracking in ASL reported object preference for pronominals [@frederiksen2017].
-Nonetheless, we only coded subjects for maintained and re-introduced referents.
-We argue that all of these factors (genre-dependency as well as object preference) might contribute to the explanation of why overt pronouns in this data are dispreferred in favor of other markers for highly accessible referents such as null marking and classifiers.
-
-A deeper look into the allocation of classifier constructions revealed that WECLs were the most common type of classifiers for the maintenance of a referent followed by BPCLs and ExtCLs.
-We observed a similar alignment for re-introduced contexts.
-In contrast, a great number of the overall ExtCLs were used to introduce a new referent into the discourse.
-That is to say, whereas WECLs were the preferred classifier types to signal highly accessible referents, ExtCLs constructions were mainly used for referents which were not easily accessible.
-We suggest that the less iconic and transparent nature of WECLs allowed signers to continue a previous referent without being redundant.
-As pointed out by @czubek2017, signers prefer to rely on ExtCLs for the first mention of an entity because these constructions are informative and explicit by nature, and they provide visual and descriptive cues about an entity and its location in space.
-Nevertheless, we only observed `r sum(df_narration$GTR1_plot == "SASS" & df_narration$Discourse == "Re-introduction")` instances of ExtCLs for re-introduced contexts, replicating @frederiksen2019's findings for ASL.
-
-As for the distribution of the null markers, approximately 90% of the plain verbs were used to maintain a previous referent, making these constructions least implicit REs used for TİD[^3].
-Agreement verbs and CA were the following implicit markers of reference, `r sum(df_narration$GTR1 == "Person agreement verb" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1 == "Person agreement verb")` and `r sum(df_narration$GTR1_plot == "CA" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1_plot == "CA")` of which were used for maintained contexts, respectively.
-This contrasts with some of the research [@czubek2017] in which agreement verbs ranked lowest on scale, indicating that these constructions were more implicit compared to plain verbs.
-However, there seems to be a common ground for the position of CA markers which have been claimed to rank higher than other null markers as in the present study [@frederiksen2016; @czubek2017; @barberà2009].
-In Figure \ref{fig:scale}, we report the distribution of REs in TİD as a function of referent accessibility and compare our accessibility hierarchy with that of [@frederiksen2016] for ASL.
-In the hierarchies, the left most REs (e.g., modified and bare nouns for TİD) are used when the accessibility is low.
-If the referent is already active in the addressee's mind and its accessibility is high, then the right most markers (e.g., agreement and plain verbs for TİD) are selected.
-
-[^3]: As in other sign languages, in TİD, the lexical meaning of signs has an effect on the morphological grouping.
-For instance, usually (di)transitive agreement verbs denote some sort of physical or metaphorical transfer of an object or abstract notion (Meir 1999).
-As for plain verbs, both phonology of a sign and its semantics have an effect on morphology.
-For instance, \textsc{eat} in TİD is signed nearby the mouth and since it needs to remain in that phonological location, it is not suitable for agreement and thus it is regarded as a plain verb.
-Likewise the experiencer verb \textsc{love} is signed on the torso and as such it does not take part in agreement.
-These generalizations noted we should also be cautious in labelling a verb as categorially one type or the other.
-For instance, some intransitive plain verbs like \textsc{fall} or \textsc{jump} in TİD may display spatial person agreement with the single argument in the clause (Gökgöz and Sevgi, 2020).
-See also for a similar discussion in Brazilian Sign Language.}.
-
-```{=tex}
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_v2_files/figure-latex/fig:count-plot-1.pdf}
+\caption{Referent counts by Discourse Status (Introduction, Maintanence,
+Re-introduction), RE Type (Nominal, Pronominal, Classifier, Null), and
+Participant Group (Native, Late)}
+\end{figure}
+
+For more inference on the data and having pairwise comparisons, we fit a
+regression model to RE Type using Discourse Status and Group as
+predictors. We used the brms package (Bürkner, 2018) in R, with a
+categorical distribution on the responses and sum contrasts for the
+predictors Discourse Status (Introduction, Maintenance, Re-introduction)
+and Participant Group (Native, Late). As the overall number of produced
+pronouns was very little (40 data points), we excluded pronominal use
+from the following analyses. For the current analysis, we did not define
+priors ourselves but used the preset uninformed priors. Figure
+2\footnote{We ran another model where we used the Nativeness predictor
+  with 3 levels 0-3:Native, 4-7:Late1, and 7-above:Late2. However, our
+  model did not converg properly. We provide the model results as a plot
+  and table in the Appendix.} shows the posterior probability
+distributions where the dot is the median estimate, the thick line is
+the 50\% credible intervals, and the thin line is the 95\% credible
+intervals. There are 3 responses we are considering in this model, and
+they are Classifier, Nominative, and Null. Classifier response is the
+level that other responses Nominative and Null are compared against.
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_v2_files/figure-latex/fig:model-gram-1.pdf}
+\caption{Regression model plot for the RE Type (Nominal, Null) with the
+predictors Discourse Status (Introduction, Maintenance, Re-introduction)
+and Participant Group (Native, Late)}
+\end{figure}
+
+In a multinomial distribution, the model takes one of the responses as a
+reference and compares all other responses to it. For example, the first
+coefficient, Nom\_Intercept (nominal intercept), in Figure 2 is the
+comparison of a nominal versus a classifier. It means that participants
+produced more nominals than classifiers in general regardless of
+participant group and discourse status. The second coefficient shows
+that null markings were also used more than classifiers by the
+participants in general, but not as much as nominals.
+
+In the comparison of nominals and classifiers, being a native acquirer
+of TİD did not have an effect. Nominals were used more when the
+discourse status was introduction compared to re-introduction. Again,
+the whole comparison is made against a classifier. In contrast to the
+introduction status, nominals were used less in maintenance contexts
+compared to the re-introduction contexts within the greater comparison
+to classifier referents. There does not seem to be an interaction
+between the introduction status of discourse and being a native acquirer
+of TİD. However, there seems to be an interaction between maintaining a
+referent and being a native acquirer of TİD. In that, there were fewer
+nominal referents when it was a native acquirer of TİD and the discourse
+context was maintenance.
+
+Compared to classifiers, being a native acquirer of TİD increased the
+odds of using null marking in general. Null constructions were used less
+in introduction contexts than in re-introduction. This contrasts with
+the nominals (4th coefficient) as they were used more in referent
+introduction versus re-introduction. In contrast to introduction
+contexts, null markers were used more in referent maintenance compared
+to re-introduction, yet again within the greater comparison to the
+classifiers. There seems to be an interaction between introducing a
+referent and being a native acquirer of TİD. In that, the odds of using
+null marking increases if the participant is a native acquirer of TİD
+and the referent is used in an introduction context. There is another
+interaction with regards to referent maintenance, contrasting the first
+interaction. In that, the odds of using null markers decrease if the
+participant is a native acquirer of TİD and the referent is maintained.
+This does not mean that late signers used more null markers in
+maintenance contexts. It means that the use of null markers did not
+linearly increase as the addition of using a null marker instead of a
+classifier as a native plus using null markers instead of a classifier
+for maintaining a referent versus re-introducing it.
+
+\hypertarget{general-discussion}{%
+\section{General Discussion}\label{general-discussion}}
+
+In our study, we aimed to (i) describe how signers track referents in
+signed discourse in TİD and (ii) examine whether language deprivation
+results in more over-redundant referent selection. To this end, we
+investigated the distribution of referring expressions (REs) that the
+native and late signers select when employing referents in different
+discourse statuses (introduced, maintained, and re-introduced).
+
+In our first set of hypotheses regarding referent tracking strategies,
+we had the general expectation that signers would use more explicit
+forms (mainly nominals) for referent introduction and less implicit
+forms such as null marking, pronominal use, and classifiers. In
+re-introduced contexts, we predicted that signers would use both
+explicit and implicit markers depending on the accessibility of the
+tracked referent.
+
+For age of acquisition effects, based on what is proposed for child and
+adult deaf signers with language deprivation in terms of their
+morphosyntactic (Boudreault and Mayberry, 2006; Cormier et al., 2012;
+Kayabaşı and Gökgöz, 2022) and narrative development (Becker, 2009;
+Cormier et al., 2013), we expected that late signers would be more
+over-redundant in their referent tracking, akin to second language
+learners (Bel et al., 2015; Frederiksen and Mayberry, 2019). That is, we
+had the expectation that late signers would use more explicit markers
+like nouns and fewer implicit markers like null marking even for highly
+accessible referents (when the discourse status is maintained or even
+re-introduced in same cases) than the control group (i.e., native
+signers).
+
+\hypertarget{tracking-referents-in-tid}{%
+\subsection{Tracking referents in TİD}\label{tracking-referents-in-tid}}
+
+The results of the present study partly support our hypotheses. In line
+with our expectations, regardless of their acquisition groups (native or
+late) (i) signers mainly used nominals to introduce a referent into the
+discourse, (ii) used null marking and classifiers but not pronouns for
+maintained contexts, and (iii) employed both nominals and less explicit
+markers (classifiers, null markers and pronouns) to re-introduce a
+referent.
+
+However, the proportion of the observed discourse statuses (26\% for
+introduction, 40\% for maintenance, 34\% for re-introduction across
+acquisition groups) in this study differed from that of some other
+studies. For a comparison, while the percentage of introduced referents
+were similar in both research (26\% versus 24\%), the participants in
+the present study brought back referents into the discourse more
+frequently compared to the participants in Frederiksen and Mayberry
+(2019) and Frederiksen and Mayberry (2016) which reported that
+re-introduction contexts accounted for only 7\% of the referents in ASL.
+On the other hand, maintenance percentages were lower in the present
+study compared to Frederiksen and Mayberry (2016) (40\% versus 69\%).
+Investigating the distribution of REs across discourse statuses in ASL,
+a similar study by Czubek (2017) also reported contrasting numbers (17\%
+for introduction, 64\% for maintenance, and 19\% for re-introduction).
+We suggest that the discrepancy might be due to the nature of the
+stimuli used in these research. In Frederiksen and Mayberry (2016), the
+authors used 4-6-episode-long picture story with 2 animate characters
+and 1 inanimate object. In contrast, the stimuli in the present study
+consisted of 8-episode-long video clips on average with a mean number of
+2 animate characters and 3 inanimate objects. Although the number of
+animate characters and inanimate characters in Czubek (2017)'s stimuli
+matched that of the present study, their narrative structure (a
+6-episode-long picture story) was still shorter compared to our
+narrative stimuli. We believe fewer possible referents and/or a simpler
+narrative structure in these studies may have contributed to the lower
+number of re-introduced contexts compared to the present study. Due to
+the story length and composition in this study, the participants had to
+track more recurring animate or inanimate referents for a longer period
+of time. As a result, our signers engaged in referent re-introduction
+more than what was reported in previous studies (Czubek, 2017;
+Frederiksen and Mayberry, 2016).
+
+As for the distribution of REs across different discourses statuses, we
+observed that signers overwhelmingly preferred to use nominals over
+other REs for the first mention of a referent and used null markers such
+as CA, and agreement and plain verb for referents with high
+accessibility (i.e., when the discourse status was maintained). This
+finding parallels other accounts of referential accessibility suggested
+for other sign languages {[}see Frederiksen and Mayberry (2016); Czubek
+(2017); and Swabey (2002); for ASL, Morgan (2005); for BSL Bel et al.
+(2015); and Barberà and Massó (2009) for LSC. Our results also indicate
+that modified nouns (e.g., nouns that are preceded or followed by a
+classifier or IX) appeared more in introduced contexts compared to
+re-introduced contexts. This brings along the assumption that modified
+nouns are used more to track referents with lower accessibility,
+probably due to their more explicit structure.
+
+Interestingly and to our surprise, the signers in this study used
+pronominal IX very sporadically with only 40 observed instances that
+were mainly used to re-introduce referents rather than maintain them.
+This finding is comparable with what has been reported for ASL (Czubek,
+2017; Frederiksen and Mayberry, 2016; Swabey, 2002) but not with LSC
+(Bel et al., 2015), BSL (Morgan, 2005) and DGS (Perniss and Özyürek,
+2015), in which the signers extensively made overt pronominal references
+(up to 30\%) to maintain entities. Comparable to our findings, It has
+been reported that pronouns do not frequently occur in the local context
+and instead follows a topic shift and thus rank higher than classifiers
+and null markers different genre (see Barberà and Massó, 2009 for LSC;
+and Nuhbalaoglu, 2018 for TİD). Considering this, the claim that the IX
+sign functions like demonstratives that are also used to track referents
+in re-introduced contexts has not been borne out (Koulidobrova and
+Lillo-Martin, 2016). Similar to these research, our findings indicate
+that the pronominal IX in TİD was used more concurrently with nominal
+constructions and pronominal IX was used very infrequently in isolation.
+The observation that overt pronouns are rare in ASL narratives comes
+from studies that include a simple picture retelling task with one
+protagonist (e.g., Frederiksen and Mayberry, 2016) as well as a more
+complex narrative design as in (Czubek, 2017). However, Czubek (2017)'s
+study also incorporated different genres and reported more instances of
+overt pronouns in personal narratives (compared to picture or video
+retelling). In fact, this finding suggests that the use of pronouns
+might be genre-dependent. Moreover, another production study on referent
+tracking in ASL reported object preference for pronominals (Frederiksen
+and Mayberry, 2017). Nonetheless, we only coded subjects for maintained
+and re-introduced referents. We argue that all of these factors
+(genre-dependency as well as object preference) might contribute to the
+explanation of why overt pronouns in this data are dispreferred in favor
+of other markers for highly accessible referents such as null marking
+and classifiers.
+
+A deeper look into the allocation of classifier constructions revealed
+that WECLs were the most common type of classifiers for the maintenance
+of a referent followed by BPCLs and ExtCLs. We observed a similar
+alignment for re-introduced contexts. In contrast, a great number of the
+overall ExtCLs were used to introduce a new referent into the discourse.
+That is to say, whereas WECLs were the preferred classifier types to
+signal highly accessible referents, ExtCLs constructions were mainly
+used for referents which were not easily accessible. We suggest that the
+less iconic and transparent nature of WECLs allowed signers to continue
+a previous referent without being redundant. As pointed out by Czubek
+(2017), signers prefer to rely on ExtCLs for the first mention of an
+entity because these constructions are informative and explicit by
+nature, and they provide visual and descriptive cues about an entity and
+its location in space. Nevertheless, we only observed 3 instances of
+ExtCLs for re-introduced contexts, replicating Frederiksen and Mayberry
+(2019)'s findings for ASL.
+
+As for the distribution of the null markers, approximately 90\% of the
+plain verbs were used to maintain a previous referent, making these
+constructions least implicit REs used for TİD\footnote{As in other sign
+  languages, in TİD, the lexical meaning of signs has an effect on the
+  morphological grouping. For instance, usually (di)transitive agreement
+  verbs denote some sort of physical or metaphorical transfer of an
+  object or abstract notion (Meir 1999). As for plain verbs, both
+  phonology of a sign and its semantics have an effect on morphology.
+  For instance, \textsc{eat} in TİD is signed nearby the mouth and since
+  it needs to remain in that phonological location, it is not suitable
+  for agreement and thus it is regarded as a plain verb. Likewise the
+  experiencer verb \textsc{love} is signed on the torso and as such it
+  does not take part in agreement. These generalizations noted we should
+  also be cautious in labelling a verb as categorially one type or the
+  other. For instance, some intransitive plain verbs like \textsc{fall}
+  or \textsc{jump} in TİD may display spatial person agreement with the
+  single argument in the clause (Gökgöz and Sevgi, 2020). See also for a
+  similar discussion in Brazilian Sign Language.\}.}. Agreement verbs
+and CA were the following implicit markers of reference, 0.77 and 0.72
+of which were used for maintained contexts, respectively. This contrasts
+with some of the research (Czubek, 2017) in which agreement verbs ranked
+lowest on scale, indicating that these constructions were more implicit
+compared to plain verbs. However, there seems to be a common ground for
+the position of CA markers which have been claimed to rank higher than
+other null markers as in the present study (Barberà and Massó, 2009;
+Czubek, 2017; Frederiksen and Mayberry, 2016). In Figure
+\ref{fig:scale}, we report the distribution of REs in TİD as a function
+of referent accessibility and compare our accessibility hierarchy with
+that of (Frederiksen and Mayberry, 2016) for ASL. In the hierarchies,
+the left most REs (e.g., modified and bare nouns for TİD) are used when
+the accessibility is low. If the referent is already active in the
+addressee's mind and its accessibility is high, then the right most
+markers (e.g., agreement and plain verbs for TİD) are selected.
+
 \begin{figure}
@@
 \end{figure}
-```
-Both hierarchies illustrated in Figure \ref{fig:scale} consider nominals and ExtCLs as markers of lowly accessibly referents.
-Likewise, in both hierarchies, as markers of highly accessible referents, different types of null markers are further distinguished.
-Plain verbs rank the lowest on the scales, preceded by agreement verbs and constructed action.
-On a similar ground, both scales lack pronouns, which occurred very infrequently in the data.
-Although @frederiksen2016 did not include other classifier types in the hierarchy, we proposed that WECLs and BPCLs must ranker higher than null markers but lower than ExtCLs that were mainly used by interlocutors to introduce a referent into the discourse context.
-
-## Age of acquisition effects on referent tracking
-
-The results in this study only somewhat support our second set of hypotheses that we set for the effects of age of acquisition on participants' RE selection.
-That is, while being a native acquirer of TİD increased the odds of using null marking in reference tracking across different discourse statuses, we failed to find such a difference for overall nominal production.
-Taking a closer look at the interaction terms, we observed that native signers, compared to late signers, were less likely to use nominal constructions when the discourse status was maintained.
-Native acquisition further increased the odds of null markers when the discourse status was introduced.
-However, we found the opposite alignment for the interaction of maintenance and native acquisition.
-That is, the coefficient for this interaction term slightly decreased the combined main effects of having a maintained context and being a native signer.
-This means that these main effects were greater than the interaction or added effects, pointing out to the implication that the type of discourse status (introduced, maintained, re-introduced) accounts for the distribution of REs in TİD narratives better than age of acquisition (native, late) which, in comparison, had a less salient effect.
-
-While late signers were only partially over-redundant in maintained contexts, they were more so in introduced contexts.
-In line with our study, a tendency for limited over-redundancy has been suggested for late signing children in DGS [@becker2009] and late signing adults in BSL specifically for the use of constructed action [@cormier2013].
-As for TİD, to our knowledge, there is only one study [@gür2018] that directly aimed at examining age of acquisition effects on adult signers' narrative skills.
-@gür2018 observed that late signing deaf adults and children set the scene less frequently and used constructed action less accurately compared to their native counterparts.
-However, this particular investigation did not present any data on the distribution of REs between two acquisition groups.
-In the present study, the finding that native signers preferred to use null arguments even for the first mention of a referent (`r sum(df_narration$ref_type == "Null" & df_narration$Discourse == "Introduction")` in total) is surprising given some of the previous literature in the topicality and saliency framework [@cormier2013; @swabey2002].
-These research reported very little or no use of null marking by native signers for introduced contexts.
-
-Given the paucity of research directed at the late signing group, we now turn to the referent tracking strategies observed for second language learners of a sign language.
-Our study replicated some of the findings of the line of research that investigated deaf or hearing bimodal bilingual signers [@bel2015; @frederiksen2019].\footnote{For similar results with heritage learners of sign language, see Reynolds (2018).} The authors reported that more instances of null marking in the context of referent introduction was attested for native signers than deaf or hearing bimodal bilingual signers.
-Although using null marking to introduce a referent into the discourse would normally result in referential ambiguity, it is important to note that our stimuli consisted of 10 interconnected narratives with 2 or more recurring characters.
-It may be possible that native signers, having already introduced the referent with an explicit nominal marker in a different discourse context, found it felicitous to refer to the same entity with more implicit markers in the following narratives even for introduced contexts.
-Then, our results imply a difference in pragmatic sensitivity to referential accessibility between native and late signers, which is more robustly reflected in introduced contexts.
-
-Since we only observed a small difference between the two acquisition groups for the use of certain REs, the aforementioned findings allude to the implication that the effect on language deprivation on pragmatic factors should also be considered in different contexts.
-For one, both groups of signers can be tested with different genres of narratives and with a more complex story design.
-In order to capture possible variation in the accessibility levels of re-introduced referents, a more fine-tuned protocol [as in @toole1996] can be used for measuring referent distance and the number of other possible competitors.
-Taken altogether, our results accentuates the necessity of engaging in signed narrative production early in life as well as the importance of early intervention for late signing deaf children.
-
-# Conclusion
-
-The present study aimed to investigate the reference tracking strategies of native and late deaf adult signers in TİD narratives.
-In line with our expectations, we found that the distribution of REs (nominal, pronominal, classifier, and null marking) followed from previous accounts of saliency and referential accessibility.
-For introduced contexts, signers produced more nominals and ExtCLs, and to maintain a previous referent, hey mainly used null arguments (plain verbs, agreement verbs and constructed action) and certain classifiers (e.g., BPCL and WECL) to track highly accessible referents.
-We observed only very few instances of pronouns (mostly produced by late signers), unlike our expectations.
-The ones we observed were mostly reserved for re-introduced contexts along with nominals.
-Based on our findings across the two groups, we proposed a hierarchy of referent accessibility for TİD.
-As for the age of acquisition effects, we observed limited over-redundancy for signers with delayed language acquisition (i.e., late signers) who employed fewer implicit forms of reference tracking compared to native signers.
-Native signers were less likely to use a nominal for maintained contexts and more likely to use null marking for introduced contexts than late signers.
-We conclude that native signers might only partially adhere to pragmatic norms and leverage accessibility better than late signers since they use less form when the referent is more active in the mind.
+
+Both hierarchies illustrated in Figure \ref{fig:scale} consider nominals
+and ExtCLs as markers of lowly accessibly referents. Likewise, in both
+hierarchies, as markers of highly accessible referents, different types
+of null markers are further distinguished. Plain verbs rank the lowest
+on the scales, preceded by agreement verbs and constructed action. On a
+similar ground, both scales lack pronouns, which occurred very
+infrequently in the data. Although Frederiksen and Mayberry (2016) did
+not include other classifier types in the hierarchy, we proposed that
+WECLs and BPCLs must ranker higher than null markers but lower than
+ExtCLs that were mainly used by interlocutors to introduce a referent
+into the discourse context.
+
+\hypertarget{age-of-acquisition-effects-on-referent-tracking}{%
+\subsection{Age of acquisition effects on referent
+tracking}\label{age-of-acquisition-effects-on-referent-tracking}}
+
+The results in this study only somewhat support our second set of
+hypotheses that we set for the effects of age of acquisition on
+participants' RE selection. That is, while being a native acquirer of
+TİD increased the odds of using null marking in reference tracking
+across different discourse statuses, we failed to find such a difference
+for overall nominal production. Taking a closer look at the interaction
+terms, we observed that native signers, compared to late signers, were
+less likely to use nominal constructions when the discourse status was
+maintained. Native acquisition further increased the odds of null
+markers when the discourse status was introduced. However, we found the
+opposite alignment for the interaction of maintenance and native
+acquisition. That is, the coefficient for this interaction term slightly
+decreased the combined main effects of having a maintained context and
+being a native signer. This means that these main effects were greater
+than the interaction or added effects, pointing out to the implication
+that the type of discourse status (introduced, maintained,
+re-introduced) accounts for the distribution of REs in TİD narratives
+better than age of acquisition (native, late) which, in comparison, had
+a less salient effect.
+
+While late signers were only partially over-redundant in maintained
+contexts, they were more so in introduced contexts. In line with our
+study, a tendency for limited over-redundancy has been suggested for
+late signing children in DGS (Becker, 2009) and late signing adults in
+BSL specifically for the use of constructed action (Cormier et al.,
+2013). As for TİD, to our knowledge, there is only one study (Gür, 2018)
+that directly aimed at examining age of acquisition effects on adult
+signers' narrative skills. Gür (2018) observed that late signing deaf
+adults and children set the scene less frequently and used constructed
+action less accurately compared to their native counterparts. However,
+this particular investigation did not present any data on the
+distribution of REs between two acquisition groups. In the present
+study, the finding that native signers preferred to use null arguments
+even for the first mention of a referent (19 in total) is surprising
+given some of the previous literature in the topicality and saliency
+framework (Cormier et al., 2013; Swabey, 2002). These research reported
+very little or no use of null marking by native signers for introduced
+contexts.
+
+Given the paucity of research directed at the late signing group, we now
+turn to the referent tracking strategies observed for second language
+learners of a sign language. Our study replicated some of the findings
+of the line of research that investigated deaf or hearing bimodal
+bilingual signers (Bel et al., 2015; Frederiksen and Mayberry,
+2019).\footnote{For similar results with heritage learners of sign language, see Reynolds (2018).}
+The authors reported that more instances of null marking in the context
+of referent introduction was attested for native signers than deaf or
+hearing bimodal bilingual signers. Although using null marking to
+introduce a referent into the discourse would normally result in
+referential ambiguity, it is important to note that our stimuli
+consisted of 10 interconnected narratives with 2 or more recurring
+characters. It may be possible that native signers, having already
+introduced the referent with an explicit nominal marker in a different
+discourse context, found it felicitous to refer to the same entity with
+more implicit markers in the following narratives even for introduced
+contexts. Then, our results imply a difference in pragmatic sensitivity
+to referential accessibility between native and late signers, which is
+more robustly reflected in introduced contexts.
+
+Since we only observed a small difference between the two acquisition
+groups for the use of certain REs, the aforementioned findings allude to
+the implication that the effect on language deprivation on pragmatic
+factors should also be considered in different contexts. For one, both
+groups of signers can be tested with different genres of narratives and
+with a more complex story design. In order to capture possible variation
+in the accessibility levels of re-introduced referents, a more
+fine-tuned protocol (as in Toole, 1996) can be used for measuring
+referent distance and the number of other possible competitors. Taken
+altogether, our results accentuates the necessity of engaging in signed
+narrative production early in life as well as the importance of early
+intervention for late signing deaf children.
+
+\hypertarget{conclusion}{%
+\section{Conclusion}\label{conclusion}}
+
+The present study aimed to investigate the reference tracking strategies
+of native and late deaf adult signers in TİD narratives. In line with
+our expectations, we found that the distribution of REs (nominal,
+pronominal, classifier, and null marking) followed from previous
+accounts of saliency and referential accessibility. For introduced
+contexts, signers produced more nominals and ExtCLs, and to maintain a
+previous referent, hey mainly used null arguments (plain verbs,
+agreement verbs and constructed action) and certain classifiers (e.g.,
+BPCL and WECL) to track highly accessible referents. We observed only
+very few instances of pronouns (mostly produced by late signers), unlike
+our expectations. The ones we observed were mostly reserved for
+re-introduced contexts along with nominals. Based on our findings across
+the two groups, we proposed a hierarchy of referent accessibility for
+TİD. As for the age of acquisition effects, we observed limited
+over-redundancy for signers with delayed language acquisition (i.e.,
+late signers) who employed fewer implicit forms of reference tracking
+compared to native signers. Native signers were less likely to use a
+nominal for maintained contexts and more likely to use null marking for
+introduced contexts than late signers. We conclude that native signers
+might only partially adhere to pragmatic norms and leverage
+accessibility better than late signers since they use less form when the
+referent is more active in the mind.
 
@@
 
-# Participant background information
+\hypertarget{participant-background-information}{%
+\section{Participant background
+information}\label{participant-background-information}}
 
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
-library(magrittr)
-library(tidyverse)
-library(readr)
-library(kableExtra)
-library("papaja")
-library(stringr)
-h1 = c("No", "Status", "Sex", "AOA", "Age", "LOE", "Deaf schools", "Home language", "Parents")
-
-df_table_a <- read.csv("./appendix/appendix_a_table.csv", sep = ";")
-names(df_table_a) <- h1
-df_table_a %<>% dplyr::select(-Parents) 
-apa_table(df_table_a,
-          caption ="Participant demographic information based on self-report",
-          note = "AOA = age of acquisition, LOE = level of education, PS = primary school, MS = middle school, HS = high school, TR = (Spoken) Turkish.",
-          escape = FALSE,
-          placement = "hbt!",
-          font_size = "small",
-          landscape = FALSE
-)
-```
+\begin{table}[hbt!]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:unnamed-chunk-2}Participant demographic information based on self-report}
+
+\small{
+
+\begin{tabular}{llllllll}
+\toprule
+No & \multicolumn{1}{c}{Status} & \multicolumn{1}{c}{Sex} & \multicolumn{1}{c}{AOA} & \multicolumn{1}{c}{Age} & \multicolumn{1}{c}{LOE} & \multicolumn{1}{c}{Deaf schools} & \multicolumn{1}{c}{Home language}\\
+\midrule
+2 & Native & M & 0-3 & 33 & 12 & PS, MS, HS & TİD, home sign\\
+3 & Native & F & 0-3 & 27 & 12 & PS, MS, HS & TİD, home sign\\
+13 & Native & F & 0-3 & 27 & 12 & PS, MS, HS & Home sign\\
+14 & Native & M & 0-3 & 29 & 12 & PS, MS, HS & TİD, home sign\\
+18 & Native & M & 0-3 & 28 & 12 & PS, MS & TİD\\
+20 & Native & M & 0-3 & 32 & 12 & PS, MS & TİD\\
+22 & Native & M & 0-3 & 35 & 12 & PS, MS, HS & TİD\\
+24 & Native & F & 0-3 & 23 & 12 & PS, MS, HS & TİD\\
+26 & Native & F & 0-3 & 30 & 12 & PS, MS, HS & TİD\\
+27 & Native & F & 0-3 & 25 & 12 & PS, MS & TİD\\
+29 & Native & F & 0-3 & 24 & 14 & PS, MS, HS & TİD\\
+30 & Native & M & 0-3 & 18 & 12 & PS, MS & TİD\\
+33 & Native & F & 0-3 & 21 & 16 & PS, MS, HS & TİD\\
+34 & Native & M & 0-3 & 19 & 16 & PS, MS, HS & TİD\\
+38 & Native & F & 0-3 & 25 & 12 & PS, MS & TİD\\
+10 & Late & F & 4-7 & 35 & 14 & PS, MS & TİD, home sign, TR\\
+23 & Late & M & 4-7 & 36 & 12 & MS, HS & TR, home sign\\
+4 & Late & M & 4-7 & 33 & 12 & PS, HS & TR\\
+16 & Late & F & 4-7 & 33 & 12 & PS & TR, home sign\\
+21 & Late & F & 4-7 & 31 & 14 & PS, MS & Home sign\\
+25 & Late & M & 4-7 & 35 & 12 & PS, MS, HS & Home sign\\
+35 & Late & F & 8-12 & 49 & 8 & PS, MS & Home sign, TR\\
+36 & Late & M & 8-12 & 43 & 16 & PS, MS & TR\\
+12 & Late & M & 8-12 & 28 & 16 & MS & TR\\
+15 & Late & F & 8-12 & 24 & 12 & PS, MS, HS & TR, TİD\\
+19 & Late & F & 8-12 & 24 & 8 & PS, MS & Home sign\\
+28 & Late & M & 8-12 & 33 & 12 & PS, MS & Home sign\\
+31 & Late & M & 13-17 & 50 & 8 & MS & TR, home sign\\
+32 & Late & M & 13-17 & 31 & 16 & PS, MS, HS & TR\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} AOA = age of acquisition, LOE = level of education, PS = primary school, MS = middle school, HS = high school, TR = (Spoken) Turkish.}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
 
@@
 
-# Effect of narrating to the camera vs. experimenter
+\hypertarget{effect-of-narrating-to-the-camera-vs.-experimenter}{%
+\section{Effect of narrating to the camera
+vs.~experimenter}\label{effect-of-narrating-to-the-camera-vs.-experimenter}}
+
 \label{section:b}
-```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
-
-# df for appendix (those who narrated to the experimenter are removed)
-df_only_camera <- df_narration %>% filter(!Subject %in% c(16, 20, 21, 24))
-
-# Coding nominal, Zero, Classifier, and Pronominal categories
-df_only_camera$ref_type <- with(df_only_camera, case_when(
-  GTR1 %in% c("Bare noun", "Point noun", "Noun Point",
-              "Fingerspelled noun", "CL noun", "Noun CL") ~ "Nominal",
-  GTR1 %in% c("Constructed action", "Constructed Action", "Plain verb", "Person agreement verb") ~ "Null",
-  GTR1 %in%  c("Whole entity CL", "Body part CL",
-               "SASS - Tracing", "SASS - Static") ~ "Classifier",
-  T ~ "Pronominal"))
-
-# Encode vector types
-df_only_camera$Narrative %<>% as.integer() 
-df_only_camera$Subject %<>% as.integer()
-df_only_camera$GTR1 %<>% as.factor() 
-df_only_camera$Nativeness %<>% dplyr::recode(`0` = "Late", `1` = "Native") %>% as.factor() %>% reorder.factor(new.order = c("Native","Late"))
-df_only_camera$Discourse %<>% as.factor()
-df_only_camera$ref_type %<>% as.factor()
-
-# dropping Pronominal references because they are too few.
-df_model_camera <- df_only_camera %>% subset(ref_type !="Pronominal") %>% drop.levels()
-
-# Contrast coding
-df_model_camera$Nativeness %<>% reorder.factor(new.order = c("Native","Late"))
-contrasts(df_model_camera$Nativeness) <- contr.sum(2)/2
-contrasts(df_model_camera$Discourse) <- contr.sum(3)/2
-contrasts(df_model_camera$Group) <- contr.sdif(3)
-
-# counts of ref types
-ref_count <- df_only_camera %>% group_by(ref_type) %>% summarise(count=n())
-
-# Model for reference type
-model_camera <- brm(ref_type ~ Nativeness*Discourse + (1|Subject), 
-               data = df_model_camera, family = categorical(link = "logit"), 
-               iter = 3000, warmup = 2000, chains = 4, cores = 4, file ="./rds/model_camera.rds")
-
-# Posterior plot for reference type
-df_model_camera <- mcmc_intervals_data(model_1, pars = vars(starts_with("b_")))
-df_model_camera$parameter %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
-                                        `b_muNull_Intercept` = "Null_Intercept",
-                                        `b_muNominal_Nativeness1` = "Nom_Native",
-                                        `b_muNull_Nativeness1` = "Null_Native",
-                                        `b_muNominal_Discourse1` = "Nom_Intro",
-                                        `b_muNominal_Discourse2` = "Nom_Maintain",
-                                        'b_muNominal_Nativeness1:Discourse1' = 'Nom_Native*Intro',
-                                        'b_muNominal_Nativeness1:Discourse2' = 'Nom_Native*Maintain',
-                                        'b_muNull_Nativeness1:Discourse1' = 'Null_Native*Intro',
-                                        'b_muNull_Nativeness1:Discourse2' = 'Null_Native*Maintain',
-                                        `b_muNull_Discourse1` = "Null_Intro",
-                                        `b_muNull_Discourse2` = "Null_Maintain")
-saveRDS(df_model_camera, "./rds/df_model_camera.rds")
-```
-
-```{r fig:model-camera, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
-p4 <- df_model_camera %>% 
-  ggplot(aes(m, y=factor(parameter,levels = rev(levels(factor(parameter)))))) +
-  geom_point(size = 2.5) +
-  geom_errorbarh(aes(xmin = l, xmax = h), alpha = 1, height = 0, size = 1.5) +
-  geom_errorbarh(aes(xmin = ll, xmax = hh, height = 0)) + vline_0(linetype = "dashed") +
-  xlab("Estimate (log odds)") + ylab("Coefficients") + theme_bw() + theme_apa()
-p4
-```
+
+\begin{figure}
+\centering
+\includegraphics{manuscript_v2_files/figure-latex/fig:model-camera-1.pdf}
+\caption{Regression model plot for the RE Type (Nominal, Null) with the
+predictors Discourse Status (Introduction, Maintenance, Re-introduction)
+and Participant Group (Native, Late)}
+\end{figure}
 
@@
 
-# Additional model with 3 levels for Nativeness
+\hypertarget{additional-model-with-3-levels-for-nativeness}{%
+\section{Additional model with 3 levels for
+Nativeness}\label{additional-model-with-3-levels-for-nativeness}}
 
@@
 
-```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warning=FALSE}
-library(magrittr)
-library(tidyverse)
-library(readr)
-library(kableExtra)
-library("papaja")
-library(janitor)
-df <- summary(model_3) %>% .$fixed
-df <- as.data.frame(df)
-df_table_extra <- df %>%  clean_names() %>% rownames_to_column("Coefficients") %>% dplyr::select(-7,-8)
-
-names(df_table_extra) <- c("Coefficients","Estimate", "EstError","lower95","upper95","Rhat")
-df_table_extra$Coefficients %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
-                                        `muNull_Intercept` = "Null_Intercept",
-                                        `muNominal_Group2M1` = "Nom_Late1vsNative",
-                                        `muNominal_Group3M2` = "Nom_Late2vsLate1",
-                                        `muNominal_Discourse1` = "Nom_Intro",
-                                        `muNominal_Discourse2` = "Nom_Maintain",
-                                        `muNominal_Group2M1:Discourse1` = 'Nom_L1vsN*Intro',
-                                        `muNominal_Group3M2:Discourse1` = 'Nom_L2vsL1*Intro',
-                                        `muNominal_Group2M1:Discourse2` = 'Nom_L1vsN*Maintain',
-                                        `muNominal_Group3M2:Discourse2` = 'Nom_L2vsL1*Maintain',
-                                        `muNull_Group2M1` = "Null_Late1vsNative",
-                                        `muNull_Group3M2` = "Null_Late2vsLate1",
-                                        `muNull_Discourse1` = "Null_Intro",
-                                        `muNull_Discourse2` = "Null_Maintain",
-                                        `muNull_Group2M1:Discourse1` = 'Null_L1vsN*Intro',
-                                        `muNull_Group3M2:Discourse1` = 'Null_L2vsL1*Intro',
-                                        `muNull_Group2M1:Discourse2` = 'Null_L1vsN*Maintain',
-                                        `muNull_Group3M2:Discourse2` = 'Null_L2vsL1*Maintain')
-df_table_extra %<>% mutate( across(.cols = "Coefficients",~str_replace( ., "_","\\\\_")))
-
-apa_table(df_table_extra,
-          format = "latex",
-          booktabs = T,
-          escape = FALSE,
-          landscape = FALSE,
-          font_size = "small",
-          placement = "hbt!",
-          caption ="Model results for Nativeness with 3 levels",
-          note = "Rhat values above 1 indicate that the regression model chains did not converge properly. There were 4 chains in our regression model."
-)
-```
+\begin{table}[hbt!]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:unnamed-chunk-4}Model results for Nativeness with 3 levels}
+
+\small{
+
+\begin{tabular}{llllll}
+\toprule
+Coefficients & \multicolumn{1}{c}{Estimate} & \multicolumn{1}{c}{EstError} & \multicolumn{1}{c}{lower95} & \multicolumn{1}{c}{upper95} & \multicolumn{1}{c}{Rhat}\\
+\midrule
+muNominal\_Intercept & 1.44 & 0.17 & 1.12 & 1.81 & 1.02\\
+Null\_Intercept & -18.01 & 19.44 & -80.15 & -2.32 & 1.08\\
+Nom\_Late1vsNative & 0.25 & 0.43 & -0.52 & 1.16 & 1.02\\
+Nom\_Late2vsLate1 & -0.28 & 0.45 & -1.24 & 0.55 & 1.02\\
+Nom\_Intro & 4.24 & 0.45 & 3.50 & 5.24 & 1.06\\
+Nom\_Maintain & -5.33 & 0.30 & -5.97 & -4.79 & 1.05\\
+Nom\_L1vsN*Intro & 2.56 & 1.29 & 0.49 & 5.68 & 1.07\\
+Nom\_L2vsL1*Intro & -4.00 & 1.25 & -6.96 & -2.01 & 1.07\\
+Nom\_L1vsN*Maintain & -0.33 & 0.83 & -2.17 & 1.13 & 1.06\\
+Nom\_L2vsL1*Maintain & 0.47 & 0.83 & -1.01 & 2.29 & 1.07\\
+Null\_Late1vsNative & -6.19 & 28.69 & -75.09 & 57.19 & 1.06\\
+Null\_Late2vsLate1 & 10.31 & 26.94 & -34.28 & 88.61 & 1.06\\
+Null\_Intro & -77.04 & 77.76 & -326.00 & -14.23 & 1.08\\
+Null\_Maintain & 38.72 & 38.88 & 7.28 & 162.93 & 1.08\\
+Null\_L1vsN*Intro & -25.29 & 114.78 & -300.34 & 227.46 & 1.06\\
+Null\_L2vsL1*Intro & 40.13 & 107.78 & -137.50 & 352.79 & 1.06\\
+Null\_L1vsN*Maintain & 12.85 & 57.38 & -114.06 & 150.55 & 1.06\\
+Null\_L2vsL1*Maintain & -20.72 & 53.89 & -177.66 & 68.26 & 1.06\\
+\bottomrule
+\addlinespace
+\end{tabular}
+
+}
+
+\begin{tablenotes}[para]
+\normalsize{\textit{Note.} Rhat values above 1 indicate that the regression model chains did not converge properly. There were 4 chains in our regression model.}
+\end{tablenotes}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
 
@@
 
-```{=tex}
 \setlength{\columnsep}{10pt}
@@
 \footnotesize
-```
-# Episodes in the stimuli
+
+\hypertarget{episodes-in-the-stimuli}{%
+\section{Episodes in the stimuli}\label{episodes-in-the-stimuli}}
 
@@
 
-**Narration 1**
-
-1\. Tom chases Jerry, wearing a wooden plank
-
-2\. Tom and Jerry enter a tube
-
-3\. Tom falls down and Jerry runs away
-
-4\. Tube gets thinner and Tom chases Jerry
-
-5\. Jerry gets out and escapes
-
-6\. Tom follows Jerry in shape of a stick
-
-**Narration 2**
-
-1\. Tom walks past a female cat sitting on a couch
-
-2\. Tom turns to and approaches female cat
-
-3\. Female cat blushes
-
-4\. Tom offers a fish in aquarium to female cat
-
-5\. Female cat rejects offer
-
-6\. Tom offers a bird in cage to female cat
-
-7\. Episode 5 is repeated
-
-8\. Tom leaves by walking in a flirtatious manner
-
-**Narration 3**
-
-1\. Jerry falls down and tries to walk up stairs
-
-2\. Tom pulls carpet back
-
-3\. Jerry struggles
-
-4\. A piano falls down stairs
-
-5\. Jerry manages to escape
-
-6\. Tom sees piano and tries to run away
-
-7\. Piano smashes into Tom and flattens him
-
-8\. Tom falls on ground
-
-**Narration 4**
-
-1\. Tom hides behind a cardboard house
-
-2\. Jerry walks to house with a toy mouse
-
-3\. Toy leaves Jerry who stops at entrance
-
-4\. Toy enters house
-
-5\. Toy gets smashed into pieces
-
-6\. Jerry is surprised
-
-7\. Tom hurts his stomach after swallowing toy
-
-8\. Tom looks at his broken teeth on a mirror
-
-9\. Tom gets angry and throws down mirror
-
-**Narration 5**
-
-1\. Tom fakes sleeping and suddenly gets up
-
-2\. Tom pulls a rope inside a couch
-
-3\. Jerry appears holding a bar of soap
-
-4\. Jerry washes Tom's mouth with bar and flees
-
-5\. Tom spits soap out
-
-6\. Jerry escapes to nest with rope
-
-7\. Tom catches rope and stars pulling it
-
-8\. A mouse trap appears and catches Tom's nose
-
-9\. Tom yells and removes trap
-
-**Narration 6**
-
-1\. Tom points a sword at Jerry and catches him
-
-2\. Tom looks at Jerry
-
-3\. Red spots appear on Jerry's face
-
-4\. Tom gets scared and drops Jerry
-
-5\. Tom rushes into bathroom
-
-6\. Tom washes his hands
-
-7\. Tom opens a drawer
-
-8\. Tom swallows pills one after another
-
-9\. Feeling ill, Tom looks out of window
-
-10\. Tom and Jerry look at each other
-
-**Narration 7**
-
-1\. Two male cats chase one another
-
-2\. Tom sees a female cat and kisses her
-
-3\. Tom hides away when black cat returns
-
-4\. Tom pokes his head behind couch again
-
-5\. Episode 3 is repeated
-
-6\. black cat sees female cat and hugs her
-
-7\. Tom appears and black cat kisses Tom
-
-8\. Two male cats move away from one another
-
-**Narration 8**
-
-1\. Jerry hits Tom's knee with a hammer
-
-2\. Tom suddenly jumps up
-
-3\. Tom yells and holds his knee
-
-4\. Jerry appraches Tom holding a thermometer
-
-5\. Jerry sticks thermometer into Tom's mouth
-
-6\. Jerry lights a fire under thermometer
-
-7\. Red liquid starts to rise up
-
-8\. Thermometer explodes
-
-**Narration 9**
-
-1\. Tom sits on Jerry who tries to get loose
-
-2\. Jerry looks around and sees a hat ribbon
-
-3\. Jerry finds a needle attached to hat
-
-4\. Jerry takes needle and sticks Tom with it
-
-5\. Tom turns red and jumps
-
-6\. Jerry placesn eedle between two cushions
-
-7\. Tom dodges needle
-
-**Narration 10**
-
-1\. Wearing a stethoscope, Tom catches Jerry
-
-2\. Jerry yells into stethoscope
-
-3\. Tom's head pops up
-
-4\. Jerry runs and enter a mouse nest
-
-5\. Tom inserts a gun into nest
-
-6\. Gun bends over toward Tom
-
-7\. Tom fires gun and his head explodes \clearpage
+\textbf{Narration 1}
+
+1. Tom chases Jerry, wearing a wooden plank
+
+2. Tom and Jerry enter a tube
+
+3. Tom falls down and Jerry runs away
+
+4. Tube gets thinner and Tom chases Jerry
+
+5. Jerry gets out and escapes
+
+6. Tom follows Jerry in shape of a stick
+
+\textbf{Narration 2}
+
+1. Tom walks past a female cat sitting on a couch
+
+2. Tom turns to and approaches female cat
+
+3. Female cat blushes
+
+4. Tom offers a fish in aquarium to female cat
+
+5. Female cat rejects offer
+
+6. Tom offers a bird in cage to female cat
+
+7. Episode 5 is repeated
+
+8. Tom leaves by walking in a flirtatious manner
+
+\textbf{Narration 3}
+
+1. Jerry falls down and tries to walk up stairs
+
+2. Tom pulls carpet back
+
+3. Jerry struggles
+
+4. A piano falls down stairs
+
+5. Jerry manages to escape
+
+6. Tom sees piano and tries to run away
+
+7. Piano smashes into Tom and flattens him
+
+8. Tom falls on ground
+
+\textbf{Narration 4}
+
+1. Tom hides behind a cardboard house
+
+2. Jerry walks to house with a toy mouse
+
+3. Toy leaves Jerry who stops at entrance
+
+4. Toy enters house
+
+5. Toy gets smashed into pieces
+
+6. Jerry is surprised
+
+7. Tom hurts his stomach after swallowing toy
+
+8. Tom looks at his broken teeth on a mirror
+
+9. Tom gets angry and throws down mirror
+
+\textbf{Narration 5}
+
+1. Tom fakes sleeping and suddenly gets up
+
+2. Tom pulls a rope inside a couch
+
+3. Jerry appears holding a bar of soap
+
+4. Jerry washes Tom's mouth with bar and flees
+
+5. Tom spits soap out
+
+6. Jerry escapes to nest with rope
+
+7. Tom catches rope and stars pulling it
+
+8. A mouse trap appears and catches Tom's nose
+
+9. Tom yells and removes trap
+
+\textbf{Narration 6}
+
+1. Tom points a sword at Jerry and catches him
+
+2. Tom looks at Jerry
+
+3. Red spots appear on Jerry's face
+
+4. Tom gets scared and drops Jerry
+
+5. Tom rushes into bathroom
+
+6. Tom washes his hands
+
+7. Tom opens a drawer
+
+8. Tom swallows pills one after another
+
+9. Feeling ill, Tom looks out of window
+
+10. Tom and Jerry look at each other
+
+\textbf{Narration 7}
+
+1. Two male cats chase one another
+
+2. Tom sees a female cat and kisses her
+
+3. Tom hides away when black cat returns
+
+4. Tom pokes his head behind couch again
+
+5. Episode 3 is repeated
+
+6. black cat sees female cat and hugs her
+
+7. Tom appears and black cat kisses Tom
+
+8. Two male cats move away from one another
+
+\textbf{Narration 8}
+
+1. Jerry hits Tom's knee with a hammer
+
+2. Tom suddenly jumps up
+
+3. Tom yells and holds his knee
+
+4. Jerry appraches Tom holding a thermometer
+
+5. Jerry sticks thermometer into Tom's mouth
+
+6. Jerry lights a fire under thermometer
+
+7. Red liquid starts to rise up
+
+8. Thermometer explodes
+
+\textbf{Narration 9}
+
+1. Tom sits on Jerry who tries to get loose
+
+2. Jerry looks around and sees a hat ribbon
+
+3. Jerry finds a needle attached to hat
+
+4. Jerry takes needle and sticks Tom with it
+
+5. Tom turns red and jumps
+
+6. Jerry placesn eedle between two cushions
+
+7. Tom dodges needle
+
+\textbf{Narration 10}
+
+1. Wearing a stethoscope, Tom catches Jerry
+
+2. Jerry yells into stethoscope
+
+3. Tom's head pops up
+
+4. Jerry runs and enter a mouse nest
+
+5. Tom inserts a gun into nest
+
+6. Gun bends over toward Tom
+
+7. Tom fires gun and his head explodes \clearpage
 
@@
 
-# References
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-ahn2019}{}}%
+Ahn, Dorothy, 2019. The determinacy scale: A competition mechanism for
+anaphoric expressions (PhD thesis).
+
+\leavevmode\vadjust pre{\hypertarget{ref-ahrenholz2005}{}}%
+Ahrenholz, Bernt, 2005. Reference to persons and objects in the function
+of subject in learner varieties. The structure of learner varieties
+1964.
+
+\leavevmode\vadjust pre{\hypertarget{ref-aikhenvald2017}{}}%
+Aikhenvald, Alexandra, 2017. A typology of noun categorization devices.
+
+\leavevmode\vadjust pre{\hypertarget{ref-aksu-kouxe72015}{}}%
+Aksu-Koç, Ayhan, Nicolopoulou, Ageliki, 2015. Character reference in
+young children's narratives: A crosslinguistic comparison of English,
+Greek, and Turkish. Lingua, Reference and referentiality in native and
+learner grammars 155, 62--84.
+doi:\href{https://doi.org/10.1016/j.lingua.2014.04.006}{10.1016/j.lingua.2014.04.006}
+
+\leavevmode\vadjust pre{\hypertarget{ref-ariel1990}{}}%
+Ariel, Mira, 1990. Accessing Noun-Phrase Antecedents. Routledge.
+
+\leavevmode\vadjust pre{\hypertarget{ref-azar2015}{}}%
+Azar, Zeynep, Özyürek, Aslı, 2015. Discourse management: Reference
+tracking in speech and gesture in Turkish narratives. Dutch Journal of
+Applied Linguistics 4, 222--240.
+doi:\href{https://doi.org/10.1075/dujal.4.2.06aza}{10.1075/dujal.4.2.06aza}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bamberg1997}{}}%
+Bamberg, Michael GW, 1997. Positioning between structure and
+performance. Journal of narrative and life history 7, 335342.
+
+\leavevmode\vadjust pre{\hypertarget{ref-barberuxe02009}{}}%
+Barberà, Gemma, Massó, Guillem, 2009. Keeping track in sign language:
+Coreference annotation using centering theory, in: Talk Presented at the
+Sign Language Corpora: Linguistic Issues Workshop, University College
+London.
+
+\leavevmode\vadjust pre{\hypertarget{ref-becker2009}{}}%
+Becker, Claudia, 2009. Narrative competences of Deaf children in German
+Sign Language. Sign Language \& Linguistics 12, 113--160.
+doi:\href{https://doi.org/10.1075/sll.12.2.02bec}{10.1075/sll.12.2.02bec}
+
+\leavevmode\vadjust pre{\hypertarget{ref-bel2015}{}}%
+Bel, Aurora, Ortells, Marta, Morgan, Gary, 2015. Reference control in
+the narratives of adult sign language learners. International Journal of
+Bilingualism 19, 608--624.
+doi:\href{https://doi.org/10.1177/1367006914527186}{10.1177/1367006914527186}
+
+\leavevmode\vadjust pre{\hypertarget{ref-benedicto2004}{}}%
+Benedicto, Elena, Brentari, Diane, 2004. Where did all the arguments
+go?: Argument-changing properties of classifiers in ASL. Natural
+Language \& Linguistic Theory 22, 743810.
+
+\leavevmode\vadjust pre{\hypertarget{ref-berman2013relating}{}}%
+Berman, Ruth A, Slobin, Dan Isaac, 2013. Relating events in narrative: A
+crosslinguistic developmental study. Psychology Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-bosch2007}{}}%
+Bosch, Peter, Umbach, Carla, 2007. Reference determination for
+demonstrative pronouns. ZAS Papers in Linguistics 48, 39--51.
+doi:\href{https://doi.org/10.21248/zaspil.48.2007.353}{10.21248/zaspil.48.2007.353}
+
+\leavevmode\vadjust pre{\hypertarget{ref-boudreault2006}{}}%
+Boudreault, Patrick, Mayberry, Rachel I., 2006. Grammatical processing
+in american sign language: Age of first-language acquisition effects in
+relation to syntactic structure. Language and Cognitive Processes 21,
+608--635.
+doi:\href{https://doi.org/10.1080/01690960500139363}{10.1080/01690960500139363}
+
+\leavevmode\vadjust pre{\hypertarget{ref-buxfcrkner2018}{}}%
+Bürkner, Paul-Christian, 2018. Advanced bayesian multilevel modeling
+with the R package brms. The R Journal 10, 395.
+doi:\href{https://doi.org/10.32614/RJ-2018-017}{10.32614/RJ-2018-017}
+
+\leavevmode\vadjust pre{\hypertarget{ref-chafe1976}{}}%
+Chafe, W., 1976.
+\href{https:///paper/Givenness\%2C-contrastiveness\%2C-definiteness\%2C-subjects\%2C-Chafe/805a12a9bee00562f6c2663722e9f7f3f5991848}{Givenness,
+contrastiveness, definiteness, subjects, topics, and point of view}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-chamberlain_language_1999}{}}%
+Chamberlain, Charlene, Morford, Jill P., Mayberry, Rachel I. (Eds.),
+1999. Language acquisition by eye. Psychology Press, Mahwah, N.J.
+
+\leavevmode\vadjust pre{\hypertarget{ref-cheng2021}{}}%
+Cheng, Qi, Mayberry, Rachel I., 2021. When event knowledge overrides
+word order in sentence comprehension: Learning a first language after
+childhood. Developmental Science.
+doi:\href{https://doi.org/10.1111/desc.13073}{10.1111/desc.13073}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cormier2012}{}}%
+Cormier, Kearsy, Schembri, Adam, Vinson, David, Orfanidou, Eleni, 2012.
+First language acquisition differs from second language acquisition in
+prelingually deaf signers: Evidence from sensitivity to grammaticality
+judgement in British Sign Language. Cognition 124, 50--65.
+doi:\href{https://doi.org/10.1016/j.cognition.2012.04.003}{10.1016/j.cognition.2012.04.003}
+
+\leavevmode\vadjust pre{\hypertarget{ref-cormier2013}{}}%
+Cormier, Kearsy, Smith, Sandra, Zwets, Martine, 2013. Framing
+constructed action in British Sign Language narratives. Journal of
+Pragmatics 55, 119--139.
+doi:\href{https://doi.org/10.1016/j.pragma.2013.06.002}{10.1016/j.pragma.2013.06.002}
+
+\leavevmode\vadjust pre{\hypertarget{ref-crasborn2008}{}}%
+Crasborn, O., Sloetjes, H., 2008.
+\href{https://www.semanticscholar.org/paper/Enhanced-ELAN-functionality-for-sign-language-Crasborn-Sloetjes/e1623ec6b2bc3532129a3682eb263cc2e7d5389c}{Enhanced
+ELAN functionality for sign language corpora}. undefined.
+
+\leavevmode\vadjust pre{\hypertarget{ref-czubek2017}{}}%
+Czubek, Todd Alan, 2017. A comprehensive study of referring expressions
+in ASL 244.
+
+\leavevmode\vadjust pre{\hypertarget{ref-debreslioska2013}{}}%
+Debreslioska, Sandra, Özyürek, Asli, Gullberg, Marianne, Perniss,
+Pamela, 2013. Gestural viewpoint signals referent accessibility.
+Discourse Processes 50, 431--456.
+doi:\href{https://doi.org/10.1080/0163853X.2013.824286}{10.1080/0163853X.2013.824286}
+
+\leavevmode\vadjust pre{\hypertarget{ref-dikyuva2015}{}}%
+Dikyuva, Hasan, Makaroğlu, Bahtiyar, Arık, Engin, 2015. Türk i{ş}aret
+dili dilbilgisi kitab{ı}. Aile ve Sosyal Politikalar Bakanl{ı}{ğ}{ı}
+Yay{ı}nlar{ı}: Ankara.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dudis2004}{}}%
+Dudis, Paul G., 2004. Body partitioning and real-space blends. Cognitive
+Linguistics 15.
+doi:\href{https://doi.org/10.1515/cogl.2004.009}{10.1515/cogl.2004.009}
+
+\leavevmode\vadjust pre{\hypertarget{ref-emmorey1996}{}}%
+Emmorey, Karen, 1996. The confluence of space and language in signed
+languages, in: Language, Speech, and Communication. The MIT Press,
+Cambridge, MA, US, pp. 171--209.
+
+\leavevmode\vadjust pre{\hypertarget{ref-engberg-pedersen1993}{}}%
+Engberg-Pedersen, Elisabeth, 1993. Space in danish sign language: The
+semantics and morphosyntax of the use of space in a visual language.
+
+\leavevmode\vadjust pre{\hypertarget{ref-frederiksen2016}{}}%
+Frederiksen, Anne Therese, Mayberry, Rachel I., 2016. Who's on First?
+Investigating the referential hierarchy in simple native ASL narratives.
+Lingua 180, 49--68.
+doi:\href{https://doi.org/10.1016/j.lingua.2016.03.007}{10.1016/j.lingua.2016.03.007}
+
+\leavevmode\vadjust pre{\hypertarget{ref-frederiksen2017}{}}%
+Frederiksen, Anne Therese, Mayberry, Rachel I., 2017. Pronominalizing
+subjects: Referentially empty pronouns in ASL.
+
+\leavevmode\vadjust pre{\hypertarget{ref-frederiksen2019}{}}%
+Frederiksen, Anne Therese, Mayberry, Rachel I., 2019. Reference tracking
+in early stages of different modality L2 acquisition: Limited
+over-explicitness in novice ASL signers{'} referring expressions. Second
+Language Research 35, 253--283.
+doi:\href{https://doi.org/10.1177/0267658317750220}{10.1177/0267658317750220}
+
+\leavevmode\vadjust pre{\hypertarget{ref-givon1983}{}}%
+Givon, T., 1983. \href{https://benjamins.com/catalog/tsl.3}{Topic
+continuity in discourse}. John Benjamins Publishing Company.
+
+\leavevmode\vadjust pre{\hypertarget{ref-grice1975}{}}%
+Grice, Herbert P., 1975. Logic and conversation. Brill, p. 4158.
+
+\leavevmode\vadjust pre{\hypertarget{ref-gullberg2006}{}}%
+Gullberg, Marianne, 2006. Handling Discourse: Gestures, Reference
+Tracking, and Communication Strategies in Early L2: Language Learning
+Vol. 56, No. 1. Language Learning 56, 155--196.
+doi:\href{https://doi.org/10.1111/j.0023-8333.2006.00344.x}{10.1111/j.0023-8333.2006.00344.x}
+
+\leavevmode\vadjust pre{\hypertarget{ref-gundel1993}{}}%
+Gundel, Jeanette K., Hedberg, Nancy, Zacharski, Ron, 1993. Cognitive
+status and the form of referring expressions in discourse. Language 69,
+274--307. doi:\href{https://doi.org/10.2307/416535}{10.2307/416535}
+
+\leavevmode\vadjust pre{\hypertarget{ref-guxfcr2018}{}}%
+Gür, Cansu, 2018. The Effects of Late Language Acquisition on the
+Development of Narrative Skills: Insights from Turkish Sign Language
+(PhD thesis). Erzurum.
+
+\leavevmode\vadjust pre{\hypertarget{ref-hickmann1996}{}}%
+Hickmann, Maya, Hendriks, Henriëtte, Roland, Françoise, Liang, James,
+1996. The marking of new information in children's narratives: a
+comparison of English, French, German and Mandarin Chinese*. Journal of
+Child Language 23, 591--619.
+doi:\href{https://doi.org/10.1017/S0305000900008965}{10.1017/S0305000900008965}
+
+\leavevmode\vadjust pre{\hypertarget{ref-karaduxf6ller2017}{}}%
+Karadöller, Dilay Z, Sümer, Beyza, Özyürek, Aslı, 2017. Effects of
+Delayed Language Exposure on Spatial Language Acquisition by Signing
+Children and Adults 5.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kayaba2022}{}}%
+Kayabaşı, Demet, Gökgöz, Kadir, 2022. The causative-inchoative
+alternation and age-of-acquisition effects on multi-predicate
+constructions in turkish sign language. Language Learning and
+Development 0, 1--26.
+doi:\href{https://doi.org/10.1080/15475441.2022.2047690}{10.1080/15475441.2022.2047690}
+
+\leavevmode\vadjust pre{\hypertarget{ref-koulidobrova2016}{}}%
+Koulidobrova, Elena, Lillo-Martin, Diane, 2016. A {`}point{'}of inquiry:
+The case of the (non-) pronominal IX in ASL. The impact of pronominal
+form on interpretation 221, 250.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kruschke2011}{}}%
+Kruschke, John K., 2011. Bayesian assessment of null values via
+parameter estimation and model comparison. Perspectives on Psychological
+Science 6, 299312.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kubu2008}{}}%
+Kubuş, Okan, 2008. AN ANALYSIS OF TURKISH SIGN LANGUAGE (T{İ}D)
+PHONOLOGY AND MORPHOLOGY (PhD thesis). Ankara.
+
+\leavevmode\vadjust pre{\hypertarget{ref-lillo-martin1990}{}}%
+Lillo-Martin, Diane, Klima, Edward S., 1990. Pointing out differences:
+ASL pronouns in syntactic theory. Theoretical issues in sign language
+research, Volume 1: Linguistics 191.
+
+\leavevmode\vadjust pre{\hypertarget{ref-loew1984}{}}%
+Loew, Ruth Carolyn, 1984.
+\href{https:///paper/Roles-and-reference-in-American-sign-language-\%3A-a-Loew/315c2c539d3bd88fbc13564775141ba4a60ac5aa}{Roles
+and reference in American sign language : a developmental perspective}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-maclaughlin1997}{}}%
+Maclaughlin, Dawn, 1997.
+\href{http://lear.unive.it/jspui/handle/11707/7016}{The Structure of
+Determiner Phrases: Evidence From American Sign Language.} (PhD thesis).
+
+\leavevmode\vadjust pre{\hypertarget{ref-meier1990}{}}%
+Meier, Richard P., Newport, Elissa L., 1990. Out of the hands of babes:
+On a possible sign advantage in language acquisition. Language 66, 1.
+doi:\href{https://doi.org/10.2307/415277}{10.2307/415277}
+
+\leavevmode\vadjust pre{\hypertarget{ref-morgan2005}{}}%
+Morgan, Gary, 2005.
+\href{http://www.oxfordscholarship.com/view/10.1093/acprof:oso/9780195180947.001.0001/acprof-9780195180947}{The
+Development of Narrative Skills in British Sign Language}, in: Schick,
+Brenda, Marschark, Marc, Spencer, Patricia Elizabeth (Eds.),. Oxford
+University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-newport1990}{}}%
+Newport, Elissa L., 1990. Maturational Constraints on Language Learning.
+Cognitive Science 14, 11--28.
+doi:\href{https://doi.org/10.1207/s15516709cog1401_2}{10.1207/s15516709cog1401\_2}
+
+\leavevmode\vadjust pre{\hypertarget{ref-nuhbalaoglu2018}{}}%
+Nuhbalaoglu, Derya, 2018. Comprehension and production of referential
+expressions in German Sign Language and Turkish Sign Language: An
+empirical approach (PhD thesis). Germany.
+
+\leavevmode\vadjust pre{\hypertarget{ref-padden1986}{}}%
+Padden, Carol, 1986. Verbs and role-shifting in american sign language.
+National Association of the Deaf Silver Spring, MD, p. 57.
+
+\leavevmode\vadjust pre{\hypertarget{ref-perniss2015}{}}%
+Perniss, Pamela, Özyürek, Asli, 2015. Visible Cohesion: A Comparison of
+Reference Tracking in Sign, Speech, and Co-Speech Gesture. Topics in
+Cognitive Science 7, 36--60.
+doi:\href{https://doi.org/10.1111/tops.12122}{10.1111/tops.12122}
+
+\leavevmode\vadjust pre{\hypertarget{ref-prince1981}{}}%
+Prince, EF, 1981. Toward a taxonomy of given - new information, in:
+Cole, Peter (Ed.),. Academic Press, New York, pp. 223--255.
+
+\leavevmode\vadjust pre{\hypertarget{ref-rathmann2007}{}}%
+Rathmann, Christian, Mann, Wolfgang, Morgan, Gary, 2007. Narrative
+Structure and Narrative Development in Deaf Children. Deafness \&
+Education International 9, 187--196.
+doi:\href{https://doi.org/10.1179/146431507790559932}{10.1179/146431507790559932}
+
+\leavevmode\vadjust pre{\hypertarget{ref-sevinuxe72006}{}}%
+Sevinç, Ayça Müge, 2006. Grammatical relations and word order in turkish
+sign language (t{İ}d). Unpublished master's thesis, Middle East
+Technical University, Turkey.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sorace2011}{}}%
+Sorace, Antonella, 2011. Pinning down the concept of {``}interface{''}
+in bilingualism. Linguistic approaches to bilingualism 1, 133.
+
+\leavevmode\vadjust pre{\hypertarget{ref-sorace2009}{}}%
+Sorace, Antonella, Serratrice, Ludovica, 2009.
+\href{https://doi.org/10.1177/1367006909339810}{Internal and external
+interfaces in bilingual language development: Beyond structural
+overlap}. International Journal of Bilingualism 13, 195--210.
+
+\leavevmode\vadjust pre{\hypertarget{ref-supalla1986}{}}%
+Supalla, Ted, 1986. The classifier system in american sign language.
+Noun classes and categorization 7, 181214.
+
+\leavevmode\vadjust pre{\hypertarget{ref-supalla1990}{}}%
+Supalla, Ted, 1990. Serial verbs of motion in ASL. Theoretical issues in
+sign language research 1, 127152.
+
+\leavevmode\vadjust pre{\hypertarget{ref-swabey2002}{}}%
+Swabey, Laurie Anne, 2002. The Cognitive Status, Form and Distribution
+of Referring Expressions in ASL and English Narratives (PhD thesis).
+Minnepolis, USA.
+
+\leavevmode\vadjust pre{\hypertarget{ref-talmy2003}{}}%
+Talmy, Leonard, 2003. The representation of spatial structure in spoken
+and signed language. Perspectives on classifier constructions in sign
+language 169195.
+
+\leavevmode\vadjust pre{\hypertarget{ref-toole1996}{}}%
+Toole, Janine, 1996. The effect of genre on referential choice. p. 263.
+doi:\href{https://doi.org/10.1075/pbns.38.16too}{10.1075/pbns.38.16too}
+
+\leavevmode\vadjust pre{\hypertarget{ref-wagenmakers2007}{}}%
+Wagenmakers, Eric-Jan, 2007. A practical solution to the pervasive
+problems of p values. Psychonomic bulletin \& review 14, 779804.
+doi:\href{https://doi.org/10.3758/bf03194105}{10.3758/bf03194105}
+
+\leavevmode\vadjust pre{\hypertarget{ref-williams1988}{}}%
+Williams, Jessica, 1988. Zero Anaphora in Second Language Acquisition: A
+Comparison among Three Varieties of English. Studies in Second Language
+Acquisition 10, 339--370.
+doi:\href{https://doi.org/10.1017/S0272263100007488}{10.1017/S0272263100007488}
+
+\leavevmode\vadjust pre{\hypertarget{ref-woll2013}{}}%
+Woll, Bencie, 2013. Sign language and spoken language development in
+young children: Measuring vocabulary by means of the CDI, in: Meurant,
+Laurence, Sinte, Aurélie, Van Herreweghe, Mieke, Vermeerbergen, Myriam
+(Eds.),. DE GRUYTER, Berlin, Boston, pp. 15--34.
+doi:\href{https://doi.org/10.1515/9781614511472.15}{10.1515/9781614511472.15}
+
+\end{CSLReferences}
+
+\end{document}

### 2022-08-05 workspace.Rmd
Add files via upload

---
title: Reference Tracking and Perspective Taking Strategies of Native and Late Signers in Turkish Sign Language
abstract: |
  We investigated the reference tracking and perspective taking strategies among native and late adult deaf signers of Turkish Sign Language. Consistent with the theories of referential accessibility, signers used more explicit forms for introducing and re-introducing referents and less implicit forms for maintaining referents. As for age of acquisition effects, we report limited over-redundancy for late signers who used null marking less compared to native signers. Furthermore, native signers mixed observer and character perspectives more frequently for the concurrent tracking of multiple referents. We conclude that native signers utilize the spatial affordances of the visual-spatial modality better than late signers.
keywords: "Referential accessibility, signed narratives, pragmatic competence, delayed language acquisition, Turkish Sign Language"
journal: "Journal of Pragmatics"
date: "`r Sys.Date()`"
bibliography: ./bibliography/references.bib
times: true
numbersections: true
classoption: "review"
csl: ./bibliography/elsevier-harvard.csl
output: rticles::elsevier_article
header-includes:
   - \usepackage{floatrow}
   - \usepackage{hyperref}
   - \usepackage{enumitem}  
   - \usepackage{float}
   - \floatplacement{table}{H}
   - \floatsetup[table]{capposition=bottom}
   - \floatplacement{figure}{H}
   - \floatsetup[figure]{capposition=bottom}
   - \usepackage{times}
   - \usepackage{smartdiagram}
   - \usepackage{ragged2e}
   - \usepackage{tikz}
   - \usepackage{xcolor}
   - \usepackage[utf8]{inputenc}
   - \usepackage[T1]{fontenc}
   - \usepackage{graphicx}
   - \usepackage{./sty/langsci-gb4e}
   - \usepackage{caption}
   - \usepackage{multicol}
   - \setlength{\columnsep}{-3cm}
   - \usepackage{setspace}

---

# Introduction

Narratives are extensive discourses that construct an important part of human communication. In order to create a systematic and coherent organization of experiences, narrators must choose appropriate linguistic forms to refer to entities. Certain forms such as nominal expressions are commonly used for the first mention of a referent in a discourse, and other forms such as pronouns and null marking may track previously mentioned entities which could either be maintained across multiple clauses or re-introduced back into the discourse. The selection of these referring expressions and anaphoric tools (henceforth REATs) is contingent on cognitive abilities, such that the encoder must rely on the activation of the working memory to keep in mind previously referred expressions and long strings of information [@morgan2005; @bamberg1997]. Pragmatic abilities too play a pivotal role in the tracking of referents: The form with the least amount of redundant information must be selected based on the accessibility levels of the referent [@ahn2019; @ariel1990]. In particular, second language learners find it difficult to produce a coherent and structured narrative where it is important to use REATs appropriately in line with the cognitive statuses of the referents. This finding comes both from research examining L2 acquisition of spoken narratives [@hickmann1996; @williams1988] and signed narratives [@bel2015; @frederiksen2019] which are conveyed in different communication modalities. Yet, sign languages and deaf populations have remained mostly understudied with respect to the acquisition of narrative and pragmatic competence.

Sign languages are the natural languages of the visual-spatial modality prominently used among deaf populations. The cognitive and linguistic development of native signers (i.e., those who start acquiring a sign language from one or more deaf caregivers following birth) parallels that of native speakers [@chamberlain_language_1999]. However, reports indicate that over 90% of all deaf individuals are born into hearing families who do not know how to sign [@woll2013]. This means that the auditory modality used in the household is not accessible to many deaf children. That is why the age of acquisition of a first functional language varies greatly among deaf children, with many of whom being deprived of their first languages typically until after age 7 when they start formal education in a deaf school. Consequently, language deprivation has been found to have negative effects on certain cognitive and linguistic measures among late signing children. To name a few, delayed exposure to a first functional language is reported to negatively affect deaf children's morphosyntactic abilities [@boudreault2006; @newport1990] and narrative abilities [@becker2009].

Since there is a paucity of studies that examine deaf adult signers with delayed exposure to a sign language, it is not very clear whether the narrative abilities of late signers differ from those of native signers. To expand on the previous literature, the present study aims to investigate the reference tracking strategies of native and late adult signers in Turkish Sign Language (Türk İşaret Dili - TİD) narratives through a story-telling paradigm. We explore the nature of REAT selection as a possible indicator of pragmatic competence between our acquisition groups. The paper is organized as follows: Section 1 (i) gives an overview of REAT selection in sign languages, (ii) describes attested REATs in TİD, and (iii) reviews previous findings on the effect of delayed first language acquisition on signers' narrative competence. Section 2 and 3 present the methodology and results of this paper, respectively. Finally, Section 4 provides a general discussion of the reference tracking strategies of native and late signers in relation to the effects of delayed first language acquisition on pragmatic competence.

## REAT Selection

Natural languages can be systematically used to keep track of previous referents in the discourse and mark the changing cognitive statuses of the REATs to create a more coherent narrative. Among the theoretical frameworks that aim to explain that different linguistic forms are selected in accordance with the addressee's mental representations of the referents in the discourse is the Givenness Hierarchy [@gundel1993]. According to this model, given or familiar information is expressed in more attenuated forms such as pronominal use and null marking since the mental status of the expressed referent is already at the center of attention, as indicated by the present contextual information [@chafe1976]. The Givenness Hierarchy is akin to other previously proposed models of reference control within the topicality and saliency paradigm such as the Accessibility Hierarchy [@ariel1990], Topic Continuity [@givon1983], and Familiarity Scale [@prince1981]. @gundel1993 argued that the distribution of forms is not random and abides by the Gricean Maxim of Quantity [@grice1975] in that the individual must make the most informative and succinct choice possible. On this ground, for highly accessible referents, linguistic forms with the least phonetic content are selected for the least amount of processing effort.

It is also possible to categorize REATs into three according to their discourse statuses [@gullberg2006]. A referent can be introduced, maintained, that is, continued across two or more clauses, or re-introduced into the discourse following a topic shift. Based on the principles of reference accessibility, the introduction and re-introduction of a referent are often accomplished by full noun phrases which can be commonly modified by an adjective or demonstrative in spoken languages and also by optional markers such as classifier constructions or index points in sign languages. Maintenance, on the other hand, is typically marked with pronominals, classifiers, and null markers in many sign languages since the continued referent is still highly accessible to the addressee.

## Tracking Referents in Turkish Sign Language (TİD)

The linguistic inventory for discourse referents in TİD includes nominal constructions (bare nouns, fingerspelled nouns, and modified nouns), null markers (plain verb, agreement verb, and constructed action), classifier predicates (whole entity classifiers, body part classifiers, handling classifiers, and extension classifiers), and overt pronouns (pronominal index or IX). This interim section describes the REAT types attested in TİD under four main headings.

```{=tex}
\begin{multicols}{2}
\begin{exe}
\ex
\begin{xlist}
\ex Full noun \\\glll
\includegraphics[width=50pt]{pictures/1a_1.png}\\ \textsc{mouse\footnotemark} \\ \\ 
\ex Noun modified by classifier \\\glll
\includegraphics[width=50pt]{pictures/1c_2.png} \includegraphics[width=50pt]{pictures/1c_1.png} \\ \textsc{soap} \textsc{soap.cl} \\ \\ 
\ex Fingerspelled noun\\\glll 
\includegraphics[width=50pt]{pictures/1b_1.png} \includegraphics[width=50pt]{pictures/1b_2.png} \includegraphics[width=50pt]{pictures/1b_3.png}\\ \textsc{t-} \textsc{o-} \textsc{m} \\ \\
\ex Noun modified by index (\textsc{IX})\\\glll 
\includegraphics[width=50pt]{pictures/1d_1.png} \includegraphics[width=50pt]{pictures/1d_2.png}\\ \textsc{ix} \textsc{cat} \\ \\
\end{xlist}
\end{exe}
\end{multicols}
\ea 
\ea Pronominal \\\glll
\includegraphics[width=50pt]{pictures/2_1.png} \includegraphics[width=50pt]{pictures/2_2.png} \\ \textsc{ix} \textsc{sick} \\ \\
\z \z
\begin{multicols}{4}
\begin{exe}
\ex
\begin{xlist}
\ex Whole entity \\\glll 
\includegraphics[width=50pt]{pictures/3e.png} \\ \textsc{fish.fall}\\ \\
\ex Body part \\\glll
\includegraphics[width=50pt]{pictures/3b.png} \\ \textsc{walk} \\ \\ 
\ex Handling \\\glll 
\includegraphics[width=50pt]{pictures/3c.png} \\ \textsc{hold} \\ \\
\ex Extension \\\glll 
\includegraphics[width=50pt]{pictures/3d_1.png} \includegraphics[width=50pt]{pictures/3a.png}\\ \textsc{plank} {\textsc{dh: mouse}} \\ {} {\textsc{ndh: tunnel}}\\ \\
\end{xlist}
\end{exe}
\end{multicols}
\begin{multicols}{3}
\begin{exe}
\ex
\begin{xlist}
\ex Constructed action \\\glll 
\includegraphics[width=50pt]{pictures/4c.png} \\ \textsc{struggle} \\ \\

\ex Plain verb  \\\glll
\includegraphics[width=50pt]{pictures/4a_2.png}  \\ \textsc{enter}\\ \\

\ex Agreement verb  \\\glll 
\includegraphics[width=50pt]{pictures/4b_1.png} 
\includegraphics[width=50pt]{pictures/4b_2.png} \\ {\textsc{$_{1}$hit$_{3}$}} {} \\ \\ 
\end{xlist}
\end{exe}
\end{multicols}
\footnotetext{We use small capitals to indicate sign glosses.}
```
### Nominals

In TİD, it is possible to use bare noun phrases (\textsc{bare nom}) as in (1a) and fingerspelled nouns (\textsc{fs nom}) as in (1c) to introduce or re-introduce referents into the discourse. The first mention of a referent can also be tracked with modified nouns (\textsc{mod nom}). Nouns can be modified by classifiers as in (1b) which either follow or precede a nominal. Alternatively, the noun phrase can retain its basic meaning and the modifier can help determine its position on the accessibility hierarchy by expressing the definiteness of the nominal. Definiteness in TİD can be expressed with a prenominal or postnominal \textsc{ix} functioning as a determiner or a demonstrative as in (1d).

### Pronominals \textsc{ix}

The \textsc{ix} sign (extended index finger) has various functions in sign languages, one of which is the pronominal use which refers to established nominals by pointing to a previously assigned abstract or physical locus in the signing space when re-introducing a topic in discourse [@emmorey1996]. The same \textsc{ix} function has also been identified for TİD [@sevinç2006]. The example in (2) illustrates a spatially-anchored pronominal reference for topic re-introduction. These signs are thus rarely used for the first mention of an entity [@perniss2015]. Although other interpretations consider pointing signs without an noun phrases as examples of demonstratives [@koulidobrova2016], for the purposes of the present study, we will refer to those constructions as pronominal \textsc{ix} and particularly examine their non-deictic or personal pronoun realizations as a possible discourse strategy to mainly mark topic shift among native and late signers of TİD.

### Classifiers

Classifier predicates are highly complex morphological constructions which are very common in sign languages [@supalla1986]. These are manually signed iconic representations of animate or inanimate objects, and they embody several properties pertaining to the nominal referent including its location in the signing space, movement, path, size, and shape [@talmy2003]. As in most sign languages hitherto studied, classifier constructions are also common and found in TİD [see @dikyuva2015 for a review]. Following the classification of @engberg-pedersen1993, we distinguish between four types of classifier constructions. Whole entity classifiers (\textsc{wecl}) represent the entirety of any referent whereas body part classifiers (\textsc{bpcl}) only reflect a specific part of an animate entity (e.g., limbs, head, tongue, teeth, and so forth). Handling classifiers (\textsc{hcl}) occur with transitive verbs and depict how an object is handled or manipulated by agent. On this ground, an \textsc{hcl}\footnote{As discussed before, an \textsc{hcl} is transitive in nature and depicts the handling action performed by an animate agent. This is why most research tend to consider \textsc{hcl}s as an extension to constructed action and group them with the character perspective where the signer's body aligns with the referent (Cormier et al., 2013; Dudis, 2004).} depicts the manner of the referring action while the former two constructions encode the location of the entity as well as its motion in space. Extension classifiers (\textsc{extcl}), on the other hand, denote the size and shape of an entity by following a trajectory to trace its shape. We exemplify these constructions in (3). As expected, these constructions are argued to be higher on the accessibility hierarchy than nominals.

### Null Marking

Parallel to spoken languages, many sign languages have been reported to allow null subjects and/or objects and make use of null marking when referring to highly accessible referents which are still active in the mental representation of the addressee. In signed discourse, this is achieved by employing verbal constructions when the event is narrated via the observer perspective. To identify the predicate types in ASL, @padden1986 suggested a triadic categorization of verbs in sign languages: plain verbs, agreement verbs, and spatial verbs. According to this classification, plain verbs (\textsc{plain}) (4b) do not agree with the subject or the object. This means that they are not inflected for number or person. This contrasts with agreeing verbs (\textsc{agr}) (4c) the starting or the ending point of which can be modified to specific loci for person and number agreement while spatial verbs (as in \textsc{ix book put} 'I placed the book there') agree with locative arguments but like plain verbs, they are not inflected for number or person. TİD too licenses null arguments. Following this classification, @sevinç2006 argued that although it is common in TİD to drop arguments, there were certain limitations on the null marking of the arguments for \textsc{agr} and \textsc{plain} verbs.

## Learning How to Track Referents in sign languages

Deaf children of deaf parents have been observed to lag behind hearing age-matched peers at 4 years of age in terms of the organization of a narrative, but the general understanding is that they catch up within two years [@vercaingne-ménard2001]. In both communication modalities, children start to produce more complex and coherent stories with appropriate and informative linguistic forms to indicate anaphoric or cataphoric references after age 10 [@hickmann1996; @rathmann2007]. Prior to this age, both signing and speaking children fail to refer to entities consistently since their pragmatic abilities are yet to develop. The manipulation of linguistic forms for discourse functions to signal topic continuance or shift is difficult to master until the beginning of formal education. These findings are also comparable across different sign languages including British Sign Language (BSL) [@morgan2005], American Sign Language (ASL) [@loew1984], and German Sign Language (Deutsche Gebärdensprache - DGS) [@becker2009], pointing out to the fact that the mastery of reference tracking in narratives is one of the most protracted achievements in the acquisition of signed discourse.

Given the preliminary evidence that topic continuity and salience marking is also difficult for second language learners regardless of modality [see @bel2015; @frederiksen2019 for sign languages; and @hickmann1999; @williams1988 for spoken languages], it has been suggested under the Interface Hypothesis that the syntax-discourse interface is vulnerable and especially challenging to acquire [@sorace2011]. Similarly, @sorace2009 discuss that this interface entails much more than morphosyntactic knowledge but also includes cognitive and pragmatic resources that help provide discourse coherence. They propose that the second language discourse is over-redundant. This means that second language learners produce more overt pronouns or fuller forms of language for highly accessible referents possibly as a compensatory strategy in an attempt to reduce the increasing load on cognitive processing.

Still, only very few research examined whether linguistically deprived signers (i.e., deaf signers with delayed exposure to a sign language) also display redundancy in their reference tracking. @becker2009 examined DGS signing deaf native and late children's narrative abilities. She reported that late signing children generated more explicit forms such as nominals when referring to the characters in the story compared to native learners, concluding that first language deprivation results in an inadequate representation of the addressee and less efficient use of morphosyntax. Comparably, @cormier2013 investigated the use of constructed action in BSL among deaf native and late adult signers and found that late signers were redundant in topic continuity. In other words, the late signers produced more explicit forms like nominals, and fewer instances of null marking when maintaining previous referents.

## Present Study

In an attempt at addressing this gap in the literature, the present study aims to examine the reference tracking strategies among native and late deaf adult signers of TİD. Overall, considering what has been suggested for the distribution of REATs in signed narratives, we expect more explicit forms like nominals for referent introduction and and less explicit forms such as null marking, pronominal and/or classifier use for referent maintainance. For the re-introduction contexts, where referents which are discontinued across one or more clauses are brought back into the discourse, we expect a less clear picture. Some re-introduced referents will have higher accessibility and will thus be tracked with more implicit REATs, while others that have lower accessibility will be tracked with more explicit REATs. Also based on previous research, we hypothesize that late signers will become more over-redundant in their reference tracking compared to native signers. This would translate to a preference by late signers for nominals over null marking especially in maintenance and re-introduction contexts. In contrast, we would expect that native signers use more null marking to track referents.

# Method

## Participants

We tested 29 deaf adult signers, 15 with exposure to TİD from birth (i.e., native signers), 14 with exposure between ages 3 and 17 (i.e., late signers) (*M*\textsubscript{Age of acquisition} = 7.7 years, *SD* = 3.3 years). All participants indicated that their preferred language of communication was TİD which they used for most contexts. Detailed information about the linguistic and educational background of the participants can be found in Appendix \ref{section:a}.

All participating native signers (7 males, 8 females) were born into deaf families and started acquiring TİD from at least one of their caregivers after birth. The parents of 14 participating native signers were deaf. One native signer had a deaf father and a hearing mother. In addition to TİD, the home language for some native signers (n = 4) in early childhood included the use of home sign\footnote{Home signs are gesture-based communication systems that emerge between the deaf children and their primary caregivers. Although home sign communication may have internal consistency, they are not fully-flegged linguistic systems like natural languages}. The age of the native participants ranged between 18 and 35 (*M* = 26.4 years, *SD* = 5.0 years). This also corresponded to the total average years of TİD use for native signers. All native signers were pre-lingually deaf, that is they were either born deaf or they had become deaf before the age of 3, with the exception of one signer who became deaf between the ages of 4 and 7. Native signers reported to complete at least high school education and their level of education (LOE) was between 12 and 16 years (*M* = 12.7 years, *SD* = 1.5 years). In addition, all native participants attended at least two deaf schools, one being primary and the other middle school. 10 native participants also attended a deaf school for their high school education. Based on a self-reported 5-point scale of proficiency (1: poor, 5: proficient), native participants declared that they have high receptive and productive skills in TİD (*M* = 5 points, *SD* = 0 points). As for their proficiency in (spoken) Turkish, their self-reported rankings were on average 3.7 (*SD* = 0.7), 3.7 (*SD* = 0.7), and 2.6 (*SD* = 1.2) for reading, writing, and speaking abilities, respectively.

Comparably, late learners (8 males, 6 females) were all born into hearing families and started learning TİD later in life after age 3, usually with the start of formal education. The age of sign language acquisition for late signers ranged between years 3-17 (*M* = 8.7 years, *SD* = 3.5 years). The parents of all participating late signers were hearing, and the reported home languages used in childhood mainly consisted of Turkish and home sign. The average age of late signers was 34.6 years (*SD* = 8 years) and ranged between 24 and 50. The derived mean number of years for TİD use following its acquisition was comparable to native signers with 25.9 years (*SD* = 7.9 years). Except for two signers who became deaf between the ages of 4 and 7, all the other late signers were pre-lingually deaf. All late signers completed middle school education and the average year of schooling was 12.3 years (*SD* = 2.8 years). All late participants attended at least one deaf school. While 8 participants attended two deaf schools (either primary and middle school, or middle and high school), 3 participants reported attend three deaf schools. Using the same scale, late signers self-rated their abilities in TİD and (spoken) Turkish. Both of their receptive and productive TİD abilities rated on average 4.9 points (*SD* = 0.3 points). Late signers' mean rating for their reading abilities in Turkish was 3.5 points (*SD* = 0.7 points) while their writing and speaking abilities rated on average 3.4 (*SD* = 0.5 points) and 3.1 points (*SD* = 0.7 points), respectively.

Native and late participants' sex, average years of schooling, TİD use following its acquisition, and their self-reported receptive and productive TİD ratings, and their reading, writing, and speaking ratings for Turkish did not statistically differ; Bayesian *p*'s \> 0.1. However, participants' ages statistically differed between the two acquisition groups; *t*(27) = 3.4, Bayesian *p* \< 0.01. Overall, the native group consisted of younger participants. This was to ensure that the average years of TİD use among native signers approximated those of late signers.

## Stimuli

We used a story-telling paradigm to examine how native and late signers of TİD track reference in signed discourse. Participants were presented with 10 short video clips extracted from the cartoon *Tom and Jerry* and asked to retell them. Each clip had six to ten episodes (*M* = `r mean(c(6, 8, 8, 9, 9, 10, 8, 8, 7, 7))` episodes). The duration of the clips ranged from 17 seconds to 28 seconds (*M* = `r mean(c(19, 23, 28, 24, 22, 26, 17, 22, 23, 23))` seconds). The narratives included on average `r round(mean(c(2, 4, 2, 3, 2, 2, 3, 2, 2, 2)))` recurring animate characters for which the present context required the introduction, maintenance, and re-introduction of referents. In addition, in each narrative, there were on average `r round(mean(c(1, 1, 3, 2, 4, 4, 1, 3, 3, 3)))` inanimate objects with which the animate referents interacted. This allowed signers to switch from one perspective to another while narrating and also engage in multiple perspectives at the same time. The detailed description of the episodes in each narrative can be found in \ref{section:b}.

```{r tab-2, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warning=FALSE}
library(magrittr)
library(tidyverse)
library(readr)
library(kableExtra)
library("papaja")
library(stringr)
options(scipen=1, digits=2)

df_table_2 <- read.csv("./appendix/tables/table_tier.csv", sep = ";")
apa_table(df_table_2,
          caption ="Annotation Tiers with their Tags and Descriptions",
          note = "NOM = nominal, PRO = pronominal, CL = classifier, NULL = null anaphora.",
          escape = FALSE,
          align = c("p{3.2cm}", rep("p{3.8cm}", ncol(df_table_2))),
          placement = "p",
          font_size = "footnotesize",
          landscape = FALSE
)
```

## Task Procedure

Participants sat in front of a laptop computer which had the stimulus stories. The instructions were all given by a deaf research assistant who was present in all sessions and ensured that the task requirements were clear. The informed consent form was translated into TİD and explained to the participants by the same assistant before data recording. Participants filled in the consent form and a background questionnaire form before the onset of the trials. Following this step, signers were seated across a hearing experimenter who is a late learner of TİD. All trials were recorded by a digital camera placed in front of the signers. The participants were instructed to watch the narrative clip as many times as they needed, and narrate the content of each clip either to the experimenter or the recording digital camera.

## Data Annotation

We coded the discourse status (Introduction, Maintenance, Re-introduction), REAT type (Nominal, Pronominal, Classifier, Null), and perspectives (Character, Observer, Mixed), identifying 4523 reference markings in total. We used ELAN Linguistic Annotation Software [@crasborn2008] for data annotation. First Author who is a hearing signer of TİD annotated all the data.

Following @berman2013relating, we first identified the clause boundaries in each sentence indicated by the presence of predicates and informative non-manual markers (e.g., head nod). On the second tier, we coded the discourse status for each referent. Having adapted the classification of @gullberg2006, we considered all referents mentioned for the first time as introduced regardless of their syntactic function in the sentence. Only subjects were coded for maintenance and re-introduction. We coded subject referents as maintained only if they had been signed in the previous clause. To this end, the referents which had been discontinued across one or more sentences were all regarded re-introduced into the discourse. On a third tier, we had the REAT type for a single reference with the main categories nominal, pronominal, null marking, and classifier. Under nominal, we identified bare nouns \textsc{bare nom}, modified nouns \textsc{mod nom}, and fingerspelled nouns \textsc{fs nom}. For classifiers, we distinguished among three types of constructions: \textsc{wecl}, \textsc{bpcl}, and \textsc{extcl}. Finally, we had \textsc{ca}, \textsc{agr}, \textsc{plain} tags for null constructions. We were unable to observe any instance of \textsc{hcl}s occurring in isolation (i.e., not part of constructed action). The annotation tiers are presented in Table \ref{tab:tab-2}.

# Results

```{r echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='asis'}
# generic libraries
library(readxl) #read excel
library(gdata) # reorder.factor
library(magrittr) # pipe
library(tidyverse) # tidy
theme_set(theme_bw()) #set ggplot theme
library(extrafont)
loadfonts()

# statistics libraries
library(brms) # bayesian modeling
library(bayesplot) # plot options for bayesian models
bayesplot_theme_set(new = theme_bw()) # set plot theme
library(tidybayes)
library(bayestestR)

theme_apa <- function(base_size = 10, base_family = "serif", box = FALSE) {
  adapted_theme <- ggplot2::theme_bw(base_size, base_family) +
    ggplot2::theme(legend.position = "bottom",
      plot.title = ggplot2::element_text(size = ggplot2::rel(1.1), margin = ggplot2::margin(0, 0, ggplot2::rel(14), 0))
      # , axis.title = ggplot2::element_text(size = ggplot2::rel(1.1))
      , axis.title.x = ggplot2::element_text(size = ggplot2::rel(1), lineheight = ggplot2::rel(1.1), margin = ggplot2::margin(ggplot2::rel(12), 0, 0, 0))
      , axis.title.y = ggplot2::element_text(size = ggplot2::rel(1), lineheight = ggplot2::rel(1.1), margin = ggplot2::margin(0, ggplot2::rel(12), 0, 0))
      , axis.title.y.right = ggplot2::element_text(size = ggplot2::rel(1), lineheight = ggplot2::rel(1.1), margin = ggplot2::margin(0, 0, 0, ggplot2::rel(12)))
      , axis.ticks.length = ggplot2::unit(ggplot2::rel(6), "points")
      , axis.text = ggplot2::element_text(size = ggplot2::rel(0.9))
      , axis.text.x = ggplot2::element_text(size = ggplot2::rel(1), margin = ggplot2::margin(ggplot2::rel(6), 0, 0, 0))
      , axis.text.y = ggplot2::element_text(size = ggplot2::rel(1), margin = ggplot2::margin(0, ggplot2::rel(6), 0, 0))
      , axis.text.y.right = ggplot2::element_text(size = ggplot2::rel(1), margin = ggplot2::margin(0, 0, 0, ggplot2::rel(6)))
      , axis.line = ggplot2::element_line()
      # , axis.line.x = ggplot2::element_line()
      # , axis.line.y = ggplot2::element_line()
      , legend.title = ggplot2::element_text()
      , legend.key = ggplot2::element_rect(fill = NA, color = NA)
      , legend.key.width = ggplot2::unit(ggplot2::rel(20), "points")
      , legend.key.height = ggplot2::unit(ggplot2::rel(20), "points")
      , legend.margin = ggplot2::margin(
        t = ggplot2::rel(16)
        , r = ggplot2::rel(16)
        , b = ggplot2::rel(16)
        , l = ggplot2::rel(16)
        , unit = "points"
      )
      , panel.spacing = ggplot2::unit(ggplot2::rel(14), "points")
      , panel.grid.major.x = ggplot2::element_blank()
      , panel.grid.minor.x = ggplot2::element_blank()
      , panel.grid.major.y = ggplot2::element_blank()
      , panel.grid.minor.y = ggplot2::element_blank()
      , strip.background = ggplot2::element_rect(fill = NA, color = NA)
      , strip.text.x = ggplot2::element_text(size = ggplot2::rel(1.2), margin = ggplot2::margin(0, 0, ggplot2::rel(10), 0))
      , strip.text.y = ggplot2::element_text(size = ggplot2::rel(1.2), margin = ggplot2::margin(0, 0, 0, ggplot2::rel(10)))
    )
  if(box) {
    adapted_theme <- adapted_theme + ggplot2::theme(panel.border = ggplot2::element_rect(color = "black"))
  } else {
    adapted_theme <- adapted_theme + ggplot2::theme(panel.border = ggplot2::element_blank())
  }
  adapted_theme
}

# Read df and drop NAs
df_narration <- read.csv("data/narration_data.csv", sep = ";") %>% drop_na()

# Coding nominal, Zero, Classifier, and Pronominal categories
df_narration$ref_type <- with(df_narration, case_when(
  GTR1 %in% c("Bare noun", "Point noun", "Noun Point",
              "Fingerspelled noun", "CL noun", "Noun CL") ~ "Nominal",
  GTR1 %in% c("Constructed action", "Constructed Action", "Plain verb", "Person agreement verb") ~ "Null",
  GTR1 %in%  c("Whole entity CL", "Body part CL",
               "SASS - Tracing", "SASS - Static") ~ "Classifier",
  T ~ "Pronominal"))

# Coding nominal, Zero, Classifier, and Pronominal categories
df_narration$GTR1_plot <- with(df_narration, case_when(
  GTR1 %in% c("Bare noun") ~ "BARE NOM",
  GTR1 %in% c("Point noun", "Noun Point","CL noun", "Noun CL") ~ "M NOM",
  GTR1 %in% c("Point noun", "Noun Point","CL noun", "Noun CL") ~ "FS NOM",
  GTR1 %in% c("Constructed action", "Constructed Action") ~ "CA",
  GTR1 %in% c("Plain verb") ~ "PLAIN",
  GTR1 %in% c("Person agreement verb") ~ "AGR",
  GTR1 %in% c("Whole entity CL") ~ "WECL",
  GTR1 %in%  c("Body part CL")  ~ "BPCL",
  GTR1 %in%  c("SASS - Tracing", "SASS - Static")  ~ "SASS",
  T ~ "Pronominal"))

# Encode vector types
df_narration$Narrative %<>% as.integer() 
df_narration$Subject %<>% as.integer()
df_narration$GTR1 %<>% as.factor() 
df_narration$Nativeness %<>% dplyr::recode(`0` = "Late", `1` = "Native") %>% as.factor() %>% reorder.factor(new.order = c("Native","Late"))
df_narration$Discourse %<>% as.factor()
df_narration$ref_type %<>% as.factor()

# Summary of the data
summary_df <- df_narration %>% 
  group_by(ref_type, GTR1) %>% 
  summarise(Count = n()) %>% 
  mutate(Percentage = formattable::percent(Count/sum(Count))) %>% 
  arrange(desc(Percentage))

# Descriptive plots
df_narration$ref_type2 <- df_narration$ref_type %>% recode('Classifier' = 'CL', 'Null' = 'NULL',
                                                           'Pronominal' = 'PRO', 'Nominal' = 'NOM')

# Contrast coding
contrasts(df_narration$Nativeness) <- contr.sum(2)/2
contrasts(df_narration$Discourse) <- contr.sum(3)/2

# dropping Pronominal references because they are too few.
df_model1 <- df_narration %>% subset(ref_type !="Pronominal") %>% drop.levels()

# Contrast coding
df_model1$Nativeness %<>% reorder.factor(new.order = c("Native","Late"))
contrasts(df_model1$Nativeness) <- contr.sum(2)/2
contrasts(df_model1$Discourse) <- contr.sum(3)/2

# counts of ref types
ref_count <- df_narration %>% group_by(ref_type) %>% summarise(count=n())

# Model for reference type
model_1 <- brm(ref_type ~ Nativeness*Discourse + (1|Subject), 
               data = df_model1, family = categorical(link = "logit"), 
               iter = 3000, warmup = 2000, chains = 4, cores = 4, file ="./rds/model_1.rds")

# Posterior plot for reference type
df_model_1 <- mcmc_intervals_data(model_1, pars = vars(starts_with("b_")))
df_model_1$parameter %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
                                        `b_muNull_Intercept` = "Null_Intercept",
                                        `b_muNominal_Nativeness1` = "Nom_Native",
                                        `b_muNull_Nativeness1` = "Null_Native",
                                        `b_muNominal_Discourse1` = "Nom_Intro",
                                        `b_muNominal_Discourse2` = "Nom_Maintain",
                                        'b_muNominal_Nativeness1:Discourse1' = 'Nom_Native*Intro',
                                        'b_muNominal_Nativeness1:Discourse2' = 'Nom_Native*Maintain',
                                        'b_muNull_Nativeness1:Discourse1' = 'Null_Native*Intro',
                                        'b_muNull_Nativeness1:Discourse2' = 'Null_Native*Maintain',
                                        `b_muNull_Discourse1` = "Null_Intro",
                                        `b_muNull_Discourse2` = "Null_Maintain")
saveRDS(df_model_1, "./rds/df_model_1.rds")
```

## REAT Type

We collected `r nrow(df_narration)` data points as a result of the experiment. Each data point is a referent tracked by the participant. Each of the REATs is classified in terms of its Discourse Status (Introduction, Maintenance, Re-introduction), its Type (Nominal, Pronominal, Classifer Null)\footnote{In our statistical analyses, we will not consider the subtypes of REATs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer Null). However, we allude to the numerical distribution of the subtypes in this and following sections}, and Participant Group (Native, Late)\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17). Treating age of acquisition as a continous variable would not be informative due to the small sample size in each interval.}. Figure 1 shows the numerical distribution of the responses according to the aforementioned variables.

Across our acquisition groups, we observed `r sum(df_narration$Discourse == "Introduction")` (`r sum(df_narration$Discourse == "Introduction")/nrow(df_narration)`) instances of referent introduction, `r sum(df_narration$Discourse == "Maintenance")` (`r sum(df_narration$Discourse == "Maintenance")/nrow(df_narration)`) instances of referent maintenance, and `r sum(df_narration$Discourse == "Re-introduction")` (`r sum(df_narration$Discourse == "Re-introduction")/nrow(df_narration)`) instances of referent re-introduction. For the introduction of a referent, the majority of the REATs were nominal constructions (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Discourse == "Introduction") / sum(df_narration$Discourse == "Introduction"),digits=3)`). However, we observed a sporadic use of classifiers (`r round(sum(df_narration$GTR1_CAT == "Classifier" & df_narration$Discourse == "Introduction") / sum(df_narration$Discourse == "Introduction"),digits=3)`), and null marking (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Discourse == "Introduction") / sum(df_narration$Discourse == "Introduction"),digits=3)`) respectively. When the discourse status was maintained, signers extensively used null markers (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Discourse == "Maintenance") / sum(df_narration$Discourse == "Maintenance"),digits=3)`). to refer to the entities in the immediately previous clause. Classifier constructions (`r round(sum(df_narration$GTR1_CAT == "Classifier" & df_narration$Discourse == "Maintenance") / sum(df_narration$Discourse == "Maintenance"),digits=3)`) followed null markers in maintained contexts. We observed only a few instances of nominal use (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Discourse == "Maintenance") / sum(df_narration$Discourse == "Maintenance"),digits=3)`) for referent maintenance. In re-introduction contexts, signers used a greater proportion of nominal markings (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Discourse == "Re-introduction") / sum(df_narration$Discourse == "Re-introduction"),digits=3)`) which were followed by null (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Discourse == "Re-introduction") / sum(df_narration$Discourse == "Re-introduction"),digits=3)`), classifier (`r round(sum(df_narration$GTR1_CAT == "Classifier" & df_narration$Discourse == "Re-introduction") / sum(df_narration$Discourse == "Re-introduction"),digits=3)`), and pronominal use (`r round(sum(df_narration$GTR1_CAT == "Pronominal" & df_narration$Discourse == "Re-introduction") / sum(df_narration$Discourse == "Re-introduction"),digits=3)`).

For all discourse statuses, the most numerous nominal type was \textsc{bare nom} (`r sum(df_narration$GTR1 == "Bare noun") / sum(df_narration$ref_type == "Nominal")`) which was followed by nouns modified by \textsc{ix} (`r (sum(df_narration$GTR1 == "Point noun") + sum(df_narration$GTR1 == "Noun Point")) / sum(df_narration$ref_type == "Nominal")`), and nouns modified by classifiers (`r (sum(df_narration$GTR1 == "CL noun") + sum(df_narration$GTR1 == "Noun CL")) / sum(df_narration$ref_type == "Nominal")`). In number, signers used \textsc{mod nom} more in introduction (`r ((sum(df_narration$GTR1 == "Point noun" & df_narration$Discourse == "Introduction") + sum(df_narration$GTR1 == "Noun Point" & df_narration$Discourse == "Introduction") + sum(df_narration$GTR1 == "CL noun" & df_narration$Discourse == "Introduction") + sum(df_narration$GTR1 == "Noun CL" & df_narration$Discourse == "Introduction")) ) / sum(df_narration$ref_type == "Nominal" & df_narration$Discourse == "Introduction")`) and re-introduction (`r ((sum(df_narration$GTR1 == "Point noun" & df_narration$Discourse == "Re-introduction") + sum(df_narration$GTR1 == "Noun Point" & df_narration$Discourse == "Re-introduction") + sum(df_narration$GTR1 == "CL noun" & df_narration$Discourse == "Re-introduction") + sum(df_narration$GTR1 == "Noun CL" & df_narration$Discourse == "Re-introduction")) ) / sum(df_narration$ref_type == "Nominal" & df_narration$Discourse == "Re-introduction")`) compared to maintenance settings (`r ((sum(df_narration$GTR1 == "Point noun" & df_narration$Discourse == "Maintenance") + sum(df_narration$GTR1 == "Noun Point" & df_narration$Discourse == "Maintenance") + sum(df_narration$GTR1 == "CL noun" & df_narration$Discourse == "Maintenance") + sum(df_narration$GTR1 == "Noun CL" & df_narration$Discourse == "Maintenance")) ) / sum(df_narration$ref_type == "Nominal" & df_narration$Discourse == "Maintenance")`). Among the null markers, \textsc{ca} was most frequently employed with a proportion of (`r sum(df_narration$GTR1_plot == "CA") / sum(df_narration$ref_type == "Null")`), and it was followed by \textsc{plain} verbs (`r sum(df_narration$GTR1_plot == "PLAIN") / sum(df_narration$ref_type == "Null")`) and \textsc{agr} verbs (`r sum(df_narration$GTR1_plot == "AGR") / sum(df_narration$ref_type == "Null")`). Overall, signers used very few null markings for the introduction of a referent and the majority of the existing ones were \textsc{ca} (`r sum(df_narration$GTR1_plot == "CA" & df_narration$Discourse == "Introduction") / sum(df_narration$ref_type == "Null" & df_narration$Discourse == "Introduction")`). Additionally, signers used \textsc{ca} more for referent re-introduction (`r sum(df_narration$GTR1_plot == "CA" & df_narration$Discourse == "Re-introduction") / sum(df_narration$ref_type == "Null" & df_narration$Discourse == "Re-introduction")`) than maintenance (`r sum(df_narration$GTR1_plot == "CA" & df_narration$Discourse == "Maintenance") / sum(df_narration$ref_type == "Null" & df_narration$Discourse == "Maintenance")`). We observed a similar instances of \textsc{agr} verbs for maintained (`r sum(df_narration$GTR1_plot == "AGR" & df_narration$Discourse == "Maintenance") / sum(df_narration$ref_type == "Null" & df_narration$Discourse == "Maintenance")`) and re-introduced contexts (`r sum(df_narration$GTR1_plot == "AGR" & df_narration$Discourse == "Re-introduction") / sum(df_narration$ref_type == "Null" & df_narration$Discourse == "Re-introduction")`). In contrast, \textsc{plain} verbs were favored more in the maintenance setting (`r sum(df_narration$GTR1_plot == "PLAIN" & df_narration$Discourse == "Maintenance") / sum(df_narration$ref_type == "Null" & df_narration$Discourse == "Maintenance")`) than re-introduction (`r sum(df_narration$GTR1_plot == "PLAIN" & df_narration$Discourse == "Re-introduction") / sum(df_narration$ref_type == "Null" & df_narration$Discourse == "Re-introduction")`). As for the classifier constructions, \textsc{wecl} was the most common type (`r sum(df_narration$GTR1_plot == "WECL") / sum(df_narration$ref_type == "Classifier")`), followed by \textsc{bpcl} (`r sum(df_narration$GTR1_plot == "BPCL") / sum(df_narration$ref_type == "Classifier")`) and \textsc{extcl} (`r sum(df_narration$GTR1_plot == "SASS") / sum(df_narration$ref_type == "Classifier")`). For the first mention of a referent, signers mostly preferred to use \textsc{extcl} (`r sum(df_narration$GTR1_plot == "SASS" & df_narration$Discourse == "Introduction") / sum(df_narration$ref_type == "Classifier" & df_narration$Discourse == "Introduction")`). However, for referent maintenance and re-introduction \textsc{wecl} (`r sum(df_narration$GTR1_plot == "WECL" & df_narration$Discourse == "Maintenance") / sum(df_narration$ref_type == "Classifier" & df_narration$Discourse == "Maintenance")` and `r sum(df_narration$GTR1_plot == "WECL" & df_narration$Discourse == "Re-introduction") / sum(df_narration$ref_type == "Classifier" & df_narration$Discourse == "Re-introduction")` respectively) was used more compared to \textsc{bpcl} (`r sum(df_narration$GTR1_plot == "BPCL" & df_narration$Discourse == "Maintenance") / sum(df_narration$ref_type == "Classifier" & df_narration$Discourse == "Maintenance")` and `r sum(df_narration$GTR1_plot == "BPCL" & df_narration$Discourse == "Re-introduction") / sum(df_narration$ref_type == "Classifier" & df_narration$Discourse == "Re-introduction")` respectively) and \textsc{extcl} (`r sum(df_narration$GTR1_plot == "SASS" & df_narration$Discourse == "Maintenance") / sum(df_narration$ref_type == "Classifier" & df_narration$Discourse == "Maintenance")` and `r sum(df_narration$GTR1_plot == "SASS" & df_narration$Discourse == "Re-introduction") / sum(df_narration$ref_type == "Classifier" & df_narration$Discourse == "Re-introduction")` respectively).

Across the discourse statuses, there is also a noticeable difference between participant groups in the use of null marking for reference tracking. Native signers of TİD used it more (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "Native") / nrow(df_narration), digits=2)`) compared to the late signers (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "Late") / nrow(df_narration), digits=2)`). Comparably, late signers used nominals more frequently (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "Late") / nrow(df_narration), digits=2)`) than native signers (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "Native") / nrow(df_narration), digits=2)`).

```{r fig:count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), REAT Type (Nominal, Pronominal, Classifier, Null), and Participant Group (Native, Late)"}
df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "NULL"))

p1 <- df_narration %>%
  select(Nativeness, ref_type2, Discourse) %>%
  group_by(Nativeness, Discourse) %>%
  ggplot(aes(ref_type2, fill = Nativeness)) + geom_bar(position = "dodge") +
  stat_count(geom = "text", aes(color = Nativeness, label = ..count..), vjust = 0,
             position = position_dodge(width = 0.85), size = 2) +
  facet_grid(~Discourse) + ylab("Count") + xlab("REATs") +
  scale_color_grey() +
  scale_fill_grey() +
  theme_apa() 
p1
```

For more inference on the data and having pairwise comparisons, we fit a regression model to REAT Type using Discourse Status and Group as predictors. We used the brms package [@bürkner2018] in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late). As the overall number of produced pronouns was very little (40 data points), we excluded pronominal use from the following analyses. For the current analysis, we did not define priors ourselves but used the preset uninformed priors. Figure 2 shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals.

```{r fig:model-gram, fig.cap="Regression model plot for the REAT Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
p3 <- df_model_1 %>% 
  ggplot(aes(m, y=factor(parameter,levels = rev(levels(factor(parameter)))))) +
  geom_point(size = 2.5) +
  geom_errorbarh(aes(xmin = l, xmax = h), alpha = 1, height = 0, size = 1.5) +
  geom_errorbarh(aes(xmin = ll, xmax = hh, height = 0)) + vline_0(linetype = "dashed") +
  xlab("Estimate (log odds)") + ylab("Coefficients") + theme_bw() + theme_apa()
p3
```

In a categorical distribution, the model takes one of the responses as a reference and compares all other responses to it. For example, the first coefficient, nominal intercept, in Figure 2 is the comparison of a nominal versus a classifier. It means that participants produced more nominals than classifiers in general regardless of participant group and discourse status. The second coefficient shows that null markings were also used more than classifiers by the participants in general, but not as much as nominals.

In the comparison of nominals and classifiers, being a native acquirer of TİD did not have an effect. Nominals were used more when the discourse status was introduction compared to re-introduction. Again, the whole comparison is made against a classifier. In contrast to the introduction status, nominals were used less in maintenance contexts compared to the re-introduction contexts, within the greater comparison to classifier referents. There does not seem to be an interaction between the introduction status of discourse and being a native acquirer of TİD; however, there seems like an interaction between maintaining a referent and being a native acquirer of TİD. In that, native signers produced fewer nominal referents than late signers while maintaining referents compared to the other combinations.

Compared to classifiers, being a native acquirer of TİD increased the odds of using null marking in general. Null constructions were used less in introduction contexts than in re-introduction. This contrasts with the nominals (4th coefficient) as they were used more in referent introduction versus re-introduction. In contrast to introduction contexts, null markers were used more in referent maintenance compared to re-introduction, yet again within the greater comparison to the classifiers. There seems to be an interaction between introducing a referent and being a native acquirer of TİD. In that, the odds of using null marking further increases if the participant is a native acquirer of TİD and the referent is used in introduction contexts. There is another interaction with regards to referent maintenance, contrasting the first interaction. In that, the odds of using null markers decrease if the participant is a native acquirer of TİD and the referent is maintained.

# General Discussion

In our study, we aimed to examine the reference tracking strategies of native and late deaf adult signers in TİD narratives. To this end, we investigated the distribution of REATs that the signers select when employing referents in different discourse statuses. The results of the present study partly support our hypotheses. In line with our expectations, (i) signers mainly used explicit markers (e.g., nominals) to introduce a referent into the discourse, (ii) used non-explicit markers (e.g., null marking and classifiers) for maintained contexts, and (iii) employed both types of markers (explicit and implicit) to re-introduce a referent. However, the proportion of the observed discourse statuses (26% for introduction, 40% for maintenance, 34% for re-introduction across acquisition groups) in this study differed from that of some other studies. For a comparison, while the percentage of introduced referents were similar in both research (26% versus 24%), the participants in the present study brought back referents into the discourse more frequently compared to the participants in @frederiksen2019 and @frederiksen2016 which reported that re-introduction contexts accounted for only 7% of the referents in ASL. On the other hand, maintenance percentages were lower in the present study compared to @frederiksen2016 (40% versus 69%). Investigating the distribution of REATs across discourse statuses in ASL, a similar study by @czubek2017 also reported contrasting numbers (17% for introduction, 64% for maintenance, and 19% for re-introduction). We suggest that the discrepancy might be due to the nature of the stimuli used in these research. In @frederiksen2016, the authors used one 6-episode-long picture story with 2 animate characters and 1 inanimate object. In contrast, the stimuli in the present study consisted of 8-episode-long video clips on average with a mean number of 2 animate characters and 3 inanimate objects. Although the number of animate characters and inanimate characters in @czubek2017's stimuli matched that of the present study, their narrative structure (a 6-episode-long picture story) was still shorter compared to our narrative stimuli. We believe fewer possible referents and/or a simpler narrative structure in these studies may have contributed to the lower number of re-introduced contexts compared to the present study.

As for the distribution of REATs across different discourses statuses, we observed that signers overwhelmingly preferred to use nominals over other REATs for the first mention of a referent and used null markers such as \textsc{ca}, and \textsc{agr} and \textsc{plain} verb for referents with high accessibility (i.e., when the discourse status was maintained). This finding parallels other accounts of referential accessibility suggested for other sign languages [see @frederiksen2019; @czubek2017; and @swabey2002; for ASL, @morgan2005; for BSL @bel2015; and @barberà2009 for Catalan Sign Language (Llengua de Signes Catalana - LSC)]. Our results also indicate that modified nouns (e.g., nouns that are preceded or followed by a classifier or IX) appeared more in introduced contexts (`r sum(df_narration$GTR1_plot == "M NOM" & df_narration$Discourse == "Introduction") / sum(df_narration$Discourse == "Introduction")`) compared to re-introduced contexts (`r sum(df_narration$GTR1_plot == "M NOM" & df_narration$Discourse == "Re-introduction") / sum(df_narration$Discourse == "Re-introduction")`). This brings along the assumption that modified nouns are used more to track referents with lower accessibility, probably due to their more explicit structure. Interestingly, the signers in this study used pronominal IX very sporadically with only 40 observed instances which were mainly used to re-introduce referents rather than maintain them. This finding is comparable with what has been reported for ASL [@swabey2002; @frederiksen2019] but not with LSC [@bel2015], BSL[@morgan2005] and DGS [@perniss2015], in which the signers extensively made overt pronominal references (up to 30%) to maintain entities. Like in @frederiksen2019, our findings indicate that the pronominal \textsc{ix} was used more concurrently with nominal constructions (`r (sum(df_narration$GTR1 == "Point noun") + sum(df_narration$GTR1 == "Noun Point"))`). In contrast, pronominal \textsc{ix} was used very infrequently in isolation. We then suggest that overt pronouns in this data are dispreferred in favor of other markers for highly accessible referents such as null marking and classifiers [@wulf2002].

A deeper look into the allocation of classifier constructions revealed that \textsc{wecl}s were the most common type of classifiers for the maintenance of a referent followed by \textsc{bpcl}s and \textsc{extcl}s. We observed a similar alignment for re-introduced contexts. In contrast, a great number of the overall \textsc{extcl}s were used to introduce a new referent into the discourse. That is to say, whereas \textsc{wecl}s were the preferred classifier types to signal highly accessible referents, \textsc{extcl}s constructions were mainly used for referents which were not easily accessible. We suggest that the less iconic and transparent nature of \textsc{wecl}s allowed signers to continue a previous referent without being redundant. As pointed out by @czubek2017, signers prefer to rely on \textsc{extcl}s for the first mention of an entity because these constructions are informative and explicit by nature, and they provide visual and descriptive cues about an entity and its location in space. Nevertheless, we only observed `r sum(df_narration$GTR1_plot == "SASS" & df_narration$Discourse == "Re-introduction")` instances of \textsc{extcl}s for re-introduced contexts, replicating @frederiksen2019's findings for ASL. The sporadic use of these markers for re-introduction is counter-intuitive given that an \textsc{extcl}s is argued to be akin to other nominal constructions [@mckee2011]. This finding has encouraged us to examine their distribution more closely. Revisiting the data, we have been able to make a distinction between static \textsc{extcl}s and tracing \textsc{extcl}s. @zwitserlood2012 has argued that while the former type of classifiers have referential power (i.e., used to refer to entities), the latter lacks this property and has instead an adjectival function (i.e. used for modification). Following from this, we have identified `r sum(df_narration$GTR1 == "SASS - Static" & df_narration$Discourse == "Introduction")` instances of static \textsc{extcl}s and `r sum(df_narration$GTR1 == "SASS - Tracing" & df_narration$Discourse == "Introduction")` instances of tracing \textsc{extcl}s for referent introduction. Furthermore, signers only used static \textsc{extcl}s for maintained contexts (`r sum(df_narration$GTR1 == "SASS - Static" & df_narration$Discourse == "Introduction")` instances). On a similar ground, when the discourse status was re-introduced, only observed tokens were instances of static \textsc{extcl}s. This brings along the implication that tracing \textsc{extcl}s function like a nominal in that it is only used for introduction contexts. In contrast, static \textsc{extcl}s act similarly to other types of classifiers with respect to referential accessibility (i.e., mainly used for referent maintenance). As for the distribution of the null markers, approximately 90% of the \textsc{plain} verbs were used to maintain a previous referent, making these constructions least implicit REATs used for TİD. \textsc{agr} verbs and \textsc{ca} were the following implicit markers of reference, `r sum(df_narration$GTR1 == "Person agreement verb" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1 == "Person agreement verb")` and `r sum(df_narration$GTR1_plot == "CA" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1_plot == "CA")` of which were used for maintained contexts respectively. This contrasts with some of the research [@czubek2017] in which \textsc{agr} verbs ranked lowest on scale, indicating that these constructions were more implicit compared to \textsc{plain} verbs. However, there seems to be a common ground for the position of \textsc{ca} markers which have been claimed to rank higher than other null markers as in the present study [@frederiksen2016; @czubek2017; @barberà2009]. In Figure \ref{fig:scale}, we report the distribution of REATs in TİD as a function of accessibility.

```{=tex}
\begin{figure}
\caption{Distribution of REATs as a function of accessibility suggested for TİD}
\label{fig:scale}
\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt

\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
%uncomment if require: \path (0,270); %set diagram left start at 0, and has height of 270

%Straight Lines [id:da19719799699706264] 
\draw [fill={rgb, 255:red, 175; green, 40; blue, 40 }  ,fill opacity=1 ][line width=1.5]    (114.9,108.91) -- (135.9,108.91) -- (253.78,108.53) -- (495.74,108.99) (164.88,102.31) -- (164.92,115.31)(214.88,102.16) -- (214.92,115.16)(264.91,102.06) -- (264.89,115.06)(314.91,102.15) -- (314.89,115.15)(364.91,102.24) -- (364.89,115.24)(414.91,102.34) -- (414.89,115.34)(464.91,102.43) -- (464.89,115.43) ;
\draw [shift={(499.74,109)}, rotate = 180.11] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;
\draw [shift={(110.9,108.91)}, rotate = 0] [fill={rgb, 255:red, 0; green, 0; blue, 0 }  ][line width=0.08]  [draw opacity=0] (12.77,-6.13) -- (0,0) -- (12.77,6.13) -- cycle    ;

% Text Node
\draw (498,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{High}};
% Text Node
\draw (81,100) node [anchor=north west][inner sep=0.75pt]   [align=left] {\textbf{Low}};
% Text Node
\draw (150,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize M NOM}};
% Text Node
\draw (191,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize BARE NOM}};
% Text Node
\draw (250,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize EXTCL}};
% Text Node
\draw (285,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize WECL \& BPCL}};
% Text Node
\draw (359,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize CA}};
% Text Node
\draw (405,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize AGR}};
% Text Node
\draw (452,88) node [anchor=north west][inner sep=0.75pt]  [font=\fontsize{0.83em}{1em}\selectfont] [align=left] {{\scriptsize PLAIN}};
\end{tikzpicture}
\end{figure}
```
The results in this study only partially support our second set of hypotheses that we set for the effects of age of acquisition on participants' REAT selection. That is, while being a native acquirer of TİD increased the odds of using null marking in reference tracking, across different discourse statuses, we failed to find such a difference for overall nominal production. Taking a closer look at the interaction terms, we observed that native signers, compared to late signers, were less likely to use nominal constructions when the discourse status was maintained. Native acquisition further increased the odds of null REATs when the discourse status was introduced. However, we found the opposite alignment for the interaction of maintenance and native acquisition. That is, the coefficient for this interaction term slightly decreased the combined main effects of having a maintained context and being a native signer. This means that these main effects were greater than the interaction or added effects, pointing out to the implication that the type of discourse status (introduced, maintained, re-introduced) accounts for the distribution of REATs in TİD narratives better than age of acquisition (native, late) which, in comparison, had a less salient effect.

While late signers were only partially over-redundant in maintained contexts, they were more so in introduced contexts. In line with our study, a tendency for limited over-redundancy has been suggested for late signers of DGS [@becker2009] and second language learners of ASL [@frederiksen2019] and LSC [@bel2015]. However, the finding that native signers preferred to use implicit markers even for the first mention of a referent (`r sum(df_narration$ref_type == "Null" & df_narration$Discourse == "Introduction")` in total) is surprising given some of the previous literature in the topicality and saliency framework [@cormier2013; @swabey2002]. These research reported very little or no use of null marking by native signers for introduced contexts. Yet, or study replicated the findings of another line of research that investigated deaf or hearing bimodal bilingual signers [@bel2015; @reynolds2018]. The authors reported that more instances of null marking in the context of referent introduction was attested for native signers than deaf or hearing bimodal bilingual signers. Although using null marking to introduce a referent into the discourse would normally result in referential ambiguity, it is important to note that our stimuli consisted of 10 interconnected narratives with 2 or more recurring characters. It may be possible that native signers, having already introduced the referent with an explicit nominal marker in a different discourse context, found it felicitous to refer to the same entity with more implicit markers in the following narratives even for introduced contexts. Then, our results imply a difference in pragmatic sensitivity to referential accessibility between native and late signers, which is more robustly reflected in introduced contexts.

# Conclusion

The present study aimed to investigate the reference tracking strategies of native and late deaf adult signers in TİD narratives. In line with our expectations, we found that the distribution of REATs (nominal, pronominal, classifier, and null marking) and perspectives (character, observer, and mixed) followed from previous accounts of referential accessibility. That is, signers preferred more explicit forms for referents with low accessibility, and they used more implicit forms for highly accessible referents. As for the age of acquisition effects, we observed limited over-redundancy for signers with delayed language acquisition (i.e., late signers) who employed fewer implicit forms of reference tracking compared to native signers. We conclude that native signers might adhere to pragmatic norms and leverage accessibility better than late signers. In addition to the negative and long-lasting effects of delayed first language acquisition on late signers' morphosyntactic abilities [e.g., @boudreault2006; @newport1990], the aforementioned findings allude to the implication that pragmatic and cognitive factors should also be considered. Our results accentuates the necessity of engaging in signed narrative production early in life as well as the importance of early intervention for late signing deaf children.

\clearpage

\appendix

# Participant background information

\label{section:a}

```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
library(magrittr)
library(tidyverse)
library(readr)
library(kableExtra)
library("papaja")
library(stringr)
h1 = c("No", "Status", "Sex", "AOA", "Age", "LOE", "Deaf schools", "Home language", "Parents")

df_table_a <- read.csv("./appendix/appendix_a_table.csv", sep = ";")
names(df_table_a) <- h1
df_table_a %<>% select(-Parents) 
apa_table(df_table_a,
          caption ="Participant demographic information based on self-report",
          note = "AOA = age of acquisition, LOE = level of education, PM = primary school, MS = middle school, HS = high school, TR = (Spoken) Turkish.",
          escape = FALSE,
          placement = "hbt!",
          font_size = "small",
          landscape = FALSE
)
```

\clearpage

```{=tex}
\setlength{\columnsep}{10pt}
\twocolumn
\linespread{1}
\justifying
\footnotesize
```
# Episodes in the stimuli

\label{section:b}

**Narration 1**

1\. Tom chases Jerry, wearing a wooden plank

2\. Tom and Jerry enter a tube

3\. Tom falls down and Jerry runs away

4\. Tube gets thinner and Tom chases Jerry

5\. Jerry gets out and escapes

6\. Tom follows Jerry in shape of a stick

**Narration 2**

1\. Tom walks past a female cat sitting on a couch

2\. Tom turns to and approaches female cat

3\. Female cat blushes

4\. Tom offers a fish in aquarium to female cat

5\. Female cat rejects offer

6\. Tom offers a bird in cage to female cat

7\. Episode 5 is repeated

8\. Tom leaves by walking in a flirtatious manner

**Narration 3**

1\. Jerry falls down and tries to walk up stairs

2\. Tom pulls carpet back

3\. Jerry struggles

4\. A piano falls down stairs

5\. Jerry manages to escape

6\. Tom sees piano and tries to run away

7\. Piano smashes into Tom and flattens him

8\. Tom falls on ground

**Narration 4**

1\. Tom hides behind a cardboard house

2\. Jerry walks to house with a toy mouse

3\. Toy leaves Jerry who stops at entrance

4\. Toy enters house

5\. Toy gets smashed into pieces

6\. Jerry is surprised

7\. Tom hurts his stomach after swallowing toy

8\. Tom looks at his broken teeth on a mirror

9\. Tom gets angry and throws down mirror

**Narration 5**

1\. Tom fakes sleeping and suddenly gets up

2\. Tom pulls a rope inside a couch

3\. Jerry appears holding a bar of soap

4\. Jerry washes Tom's mouth with bar and flees

5\. Tom spits soap out

6\. Jerry escapes to nest with rope

7\. Tom catches rope and stars pulling it

8\. A mouse trap appears and catches Tom's nose

9\. Tom yells and removes trap

**Narration 6**

1\. Tom points a sword at Jerry and catches him

2\. Tom looks at Jerry

3\. Red spots appear on Jerry's face

4\. Tom gets scared and drops Jerry

5\. Tom rushes into bathroom

6\. Tom washes his hands

7\. Tom opens a drawer

8\. Tom swallows pills one after another

9\. Feeling ill, Tom looks out of window

10\. Tom and Jerry look at each other

**Narration 7**

1\. Two male cats chase one another

2\. Tom sees a female cat and kisses her

3\. Tom hides away when black cat returns

4\. Tom pokes his head behind couch again

5\. Episode 3 is repeated

6\. black cat sees female cat and hugs her

7\. Tom appears and black cat kisses Tom

8\. Two male cats move away from one another

**Narration 8**

1\. Jerry hits Tom's knee with a hammer

2\. Tom suddenly jumps up

3\. Tom yells and holds his knee

4\. Jerry appraches Tom holding a thermometer

5\. Jerry sticks thermometer into Tom's mouth

6\. Jerry lights a fire under thermometer

7\. Red liquid starts to rise up

8\. Thermometer explodes

**Narration 9**

1\. Tom sits on Jerry who tries to get loose

2\. Jerry looks around and sees a hat ribbon

3\. Jerry finds a needle attached to hat

4\. Jerry takes needle and sticks Tom with it

5\. Tom turns red and jumps

6\. Jerry placesn eedle between two cushions

7\. Tom dodges needle

**Narration 10**

1\. Wearing a stethoscope, Tom catches Jerry

2\. Jerry yells into stethoscope

3\. Tom's head pops up

4\. Jerry runs and enter a mouse nest

5\. Tom inserts a gun into nest

6\. Gun bends over toward Tom

7\. Tom fires gun and his head explodes \clearpage

\onecolumn

# References

### 2022-08-05 appendix/demographics.R

@@ -1,0 +1,41 @@
+library(dplyr) # magic
+library(magrittr) # piped magic
+library(tidyr) # dark magic
+library(ggplot2) # plots things
+library(readxl) 
+library(irr) 
+library(stats)
+library("Bolstad")
+df <- read.csv("demographics.csv", sep = ";")
+df$Sex %<>% recode('F' = 0, 'M' = 1)
+
+summary <- df %>% 
+  group_by(Status) %>% 
+  summarise(M_Age = mean(Age.),
+            SD_Age = sd(Age.),
+            SD_TID_R = sd(TID_Receptive),
+            M_TID_R = mean(TID_Receptive),
+            SD_TID_P = sd(TID_Productive),
+            M_TID_P = mean(TID_Productive),
+            SD_TR_R = sd(Turkish_Reading),
+            M_TR_R = mean(Turkish_Reading),
+            SD_TR_W = sd(Turkish_Writing),
+            M_TR_W = mean(Turkish_Writing),
+            SD_TR_S = sd(Turkish_Speaking),
+            M_TR_S = mean(Turkish_Speaking),
+            M_Use = mean(Use),
+            SD_Use = sd(Use),
+            M_LOE = mean(LOE),
+            SD_LOE = sd(LOE))
+
+bayes.t.test(formula = Sex ~ Status, data = df)
+bayes.t.test(Use ~ Status, data = df)
+bayes.t.test(LOE ~ Status, data = df)
+bayes.t.test(Age.  ~ Status, data = df)
+bayes.t.test(TID_Receptive ~ Status, data = df)
+bayes.t.test(TID_Productive ~ Status, data = df)
+bayes.t.test(Turkish_Writing ~ Status, data = df)
+bayes.t.test(Turkish_Speaking ~ Status, data = df)
+bayes.t.test(Turkish_Reading ~ Status, data = df)
+
+

### 2022-08-05 manuscript_v2.Rmd
updated files

@@
 That is to say, individuals must make the most informative and succinct choice possible to retain economy of form.
-When multiple expressions are available for the same referent as in an utterance like (1), phonetically simplest form is selected with the least redundant information are selected based on an economy principle.
-In (1), REs for "Eddie" are given in bold.
+When multiple expressions are available for the same referent as in an utterance like (\ref{ex:1}), phonetically simplest form is selected with the least redundant information are selected based on an economy principle.
+In the utterance, REs for "Eddie" are given in bold.
 All three REs (nominal, pronominal, and null marking) refer to the same entity; however, the quantity of marking decreases as the referent's accessibility increases.
@@
 \begin{exe}
-    \ex
+    \ex \label{ex:1}
     So \textbf{Eddie} turned around. \hfill Williams (1988; p. 343, ex. 7) \\
@@
 "Yeah, we want you," they say.
-So-\textbf{$\emptyset$} walked right up to them \\
+So-$\pmb{\emptyset}$ walked right up to them \\
 \end{exe}
@@
 There is an emerging body of literature that reports the distribution of REs in signed narratives.
-Building on @gundel1993's work, @swabey2002 demonstrated that, like other spoken languages such as Russian, Chinese, and Japanese, American Sign Language (ASL) signers use bare nouns for introduced and reintroduced discourse contexts whereas null marking is used for referents which are "in-focus" or very accessible to the addressee.
+Building on @gundel1993's work, @swabey2002 demonstrated that, like other spoken languages such as Russian, Chinese, and Japanese, American Sign Language (ASL) signers use bare nouns for introduced and re-introduced discourse contexts whereas null marking is used for referents which are "in-focus" or very accessible to the addressee.
 Overt REs like pronouns or nouns are disprefered for such referents.
 @swabey2002 concluded that "native" signers in her study abided by the Gricean maxim of quantity and were thus economic in their reference tracking.
-Similar findings have been reported for German Sign Language (DGS) [@perniss2015] and British Sign Language (BSL) [@morgan2005].
+Similar findings have been reported for German Sign Language (Deutsche Gebärdensprache - DGS) [@perniss2015] and British Sign Language (BSL) [@morgan2005].
 It is important to note, however, that these research either collapsed native and late signers into a single group and labeled them as "native-like" or explicitly compared native deaf signers with late deaf signers (with language deprivation) in terms of narrative competence.
@@
 
-Commonly in both spoken and sign languages nouns and noun phrases are used as referring expressions.
+Commonly in both spoken and sign languages nouns and noun phrases are used as REs.
 Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions [@lillo-martin1990].
@@
 This referential locus can later be used by agreement verbs and classifier predicates (more below).
-For instance, in (2a) there are two referents, \textsc{elvan}, who is referentially-assigned to the right side of the signer, represented with the "a" subscript and \textsc{demet}, who is referentially-assigned to the left side of the signer, represented with the "b" subscript.
-When the signer makes a contrast in (2b), as a pronominal use, she points to the right side to refer to ELVAN and to the left to refer to \textsc{demet}.
-We schematically represents the referential locus assigned to the referents in (3).
+For instance, in (\ref{ex:2a}) there are two referents, \textsc{elvan}, who is referentially-assigned to the right side of the signer, represented with the "a" subscript and \textsc{demet}, who is referentially-assigned to the left side of the signer, represented with the "b" subscript.
+When the signer makes a contrast in (\ref{ex:2b}), as a pronominal use, she points to the right side to refer to ELVAN and to the left to refer to \textsc{demet}.
+We schematically represents the referential locus assigned to the referents in (\ref{ex:3}).
 
@@
 \begin{xlist}
-    \ex
+    \ex \label{ex:2a}
     \textsc{lab place elvan ix\textsubscript{a} ten years work. demet ix\textsubscript{b} last year start.} \\
@@
     \\
-    \ex
+    \ex \label{ex:2b}
     \textsc{ix\textsubscript{a} e-l-a-n know. ix\textsubscript{b} learn.}\\
@@
     \begin{exe}
-    \ex Signing space, signer, example referential loci for the referents \textsc{elvan} (a) and \textsc{demet} (b) \\
-
+    \ex \label{ex:3} Signing space, signer, example referential loci for the referents \textsc{elvan} (a) and \textsc{demet} (b) \\
 \tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
@@
 Some classifier handshapes and the set of objects that they denote are below.
-Some classifier handshapes and the set of objects that they denote are presented below in (4).
+Some classifier handshapes and the set of objects that they denote are presented below in (\ref{ex:4}).
 
@@
 \begin{exe}
-\ex 
+\ex \label{ex:4}
 \includegraphics[width=25pt]{pictures/flathand.png}: flat entities, surfaces, vehicles.\\
@@
 ```
-In (5), in the first clause, a referential locus, i.e.
+In (\ref{ex:5}), in the first clause, a referential locus, i.e.
 locus-a, is assigned to the locative source argument \textsc{cage}.
@@
 \begin{exe}
-    \ex 
+    \ex \label{ex:5}
 \begin{xlist}
@@
 \begin{xlist}
-\ex Full noun \\\glll
+\ex \label{ex:6a} Full noun \\\glll
 \includegraphics[width=50pt]{pictures/1a_1.png}\\ \textsc{mouse\footnotemark} \\ \\ 
-\ex Noun modified by classifier \\\glll
+\ex \label{ex:6b} Noun modified by classifier \\\glll
 \includegraphics[width=50pt]{pictures/1c_2.png} \includegraphics[width=50pt]{pictures/1c_1.png} \\ \textsc{soap} \textsc{soap.cl} \\ \\ 
-\ex Fingerspelled noun\\\glll 
+\ex \label{ex:6c} Fingerspelled noun\\\glll 
 \includegraphics[width=50pt]{pictures/1b_1.png} \includegraphics[width=50pt]{pictures/1b_2.png} \includegraphics[width=50pt]{pictures/1b_3.png}\\ \textsc{t-} \textsc{o-} \textsc{m} \\ \\
-\ex Noun modified by index (IX)\\\glll 
-\includegraphics[width=50pt]{pictures/1d_1.png} \includegraphics[width=50pt]{pictures/1d_2.png}\\ IX \textsc{cat} \\ \\
+\ex \label{ex:6d} Noun modified by index (\textsc{IX})\\\glll 
+\includegraphics[width=50pt]{pictures/1d_1.png} \includegraphics[width=50pt]{pictures/1d_2.png}\\ \textsc{ix} \textsc{cat} \\ \\
 \end{xlist}
@@
 \ea 
-\ea Pronominal \\\glll
-\includegraphics[width=50pt]{pictures/2_1.png} \includegraphics[width=50pt]{pictures/2_2.png} \\ IX \textsc{sick} \\ \\
+\ea \label{ex:7} Pronominal \\\glll
+\includegraphics[width=50pt]{pictures/2_1.png} \includegraphics[width=50pt]{pictures/2_2.png} \\ \textsc{ix} \textsc{sick} \\ \\
 \z \z
@@
 \begin{exe}
-\ex
+\ex \label{ex:8}
 \begin{xlist}
@@
 \begin{xlist}
-\ex Constructed action \\\glll 
+\ex \label{ex:9a} Constructed action \\\glll 
 \includegraphics[width=50pt]{pictures/4c.png} \\ \textsc{struggle} \\ \\
 
-\ex Plain verb  \\\glll
+\ex \label{ex:9b} Plain verb  \\\glll
 \includegraphics[width=50pt]{pictures/4a_2.png}  \\ \textsc{enter}\\ \\
 
-\ex Agreement verb  \\\glll 
+\ex \label{ex:9c} Agreement verb  \\\glll 
 \includegraphics[width=50pt]{pictures/4b_1.png} 
@@
 
-In TİD and many other sign languages, it is possible to use bare noun phrases (\textsc{bare nom}) as in (1a) as well as fingerspelled nouns (\textsc{fs nom}) as in (1c).
+In TİD and many other sign languages, it is possible to use bare noun phrases as in (\ref{ex:6a}) as well as fingerspelled nouns as in (\ref{ex:6c}).
 Fingerspelling refers to the use of the manual alphabet which is derived or borrowed from a surrounding written language [@kubu2008].
-Nouns can also be modified (\textsc{mod nom}) by classifiers as in (1b) [@frederiksen2016] or by a prenominal or postnominal IX functioning as a determiner or a demonstrative as in (1d) [@maclaughlin1997].
+Nouns can also be modified by classifiers as in (\ref{ex:6b}) [@frederiksen2016] or by a prenominal or postnominal IX functioning as a determiner or a demonstrative as in (\ref{ex:6d}) [@maclaughlin1997].
 
@@
 The same IX function has also been identified for TİD [@sevinç2006].
-The example in (2) illustrates a spatially-anchored pronominal reference for topic re-introduction.
+The example in (\ref{ex:7}) illustrates a spatially-anchored pronominal reference for topic re-introduction.
 These signs are thus rarely used for the first mention of an entity [@perniss2015].
@@
 Extension classifiers (ExtCL), on the other hand, denote the size and shape of an entity by following a trajectory to trace its shape.
-We exemplify these constructions in (3).
+We exemplify these constructions in (\ref{ex:8}).
 
@@
 Parallel to spoken languages, many sign languages have been reported to allow null subjects and/or objects and make use of null arguments when referring to highly accessible referents which are still active in the mental representation of the addressee.
-In signed discourse, this can be achieved by employing certain verbal constructions.
-To identify the predicate types in ASL, @padden1986 suggested a triadic categorization of verbs in sign languages: plain verbs, agreement verbs, and spatial verbs.
-According to this classification, plain verbs (plain) (4b) do not agree with the subject or the object.
+In signed discourse, this can be achieved by employing certain verbal constructions or constructed action (CA).
+The latter construction exemplified in (\ref{ex:9a}) refers to cases in which the signers engage with the space in front of themselves as if they are interacting with the objects and characters in the story.
+This is often achieved with the signer taking on the role of a referent in the narrative and performing the actions [@metzger1995].
+CA is often marked by certain non-manual cues such as the facial expression and the breaking of the eye contact.
+This is to say, the narrator's head, face, torso, arms, and as an optional marker, hands represent those of the associated referent(s) [@smith2014].
+As for the verbal constructions, to identify the predicate types in ASL, @padden1986 suggested a triadic categorization of verbs in sign languages: plain verbs, agreement verbs, and spatial verbs.
+According to this classification, plain verbs (\ref{ex:9b}) do not agree with the subject or the object.
 This means that they are not inflected for number or person.
-This contrasts with agreeing verbs (agreement) (4c) the starting or the ending point of which can be modified to specific loci for person and number agreement while spatial verbs (as in \textsc{ix book put} 'I placed the book there') agree with locative arguments but like plain verbs, they are not inflected for number or person.
+This contrasts with agreeing verbs (\ref{ex:9c}) the starting or the ending point of which can be modified to specific loci for person and number agreement while spatial verbs (as in \textsc{ix book put} 'I placed the book there') agree with locative arguments but like plain verbs, they are not inflected for number or person.
 TİD too licenses null arguments.
@@
 The manipulation of linguistic forms for discourse functions to signal topic continuance or shift is difficult to master until the beginning of formal education.
-These findings are also comparable across different sign languages including British Sign Language (BSL) [@morgan2005], ASL [@loew1984], and German Sign Language (Deutsche Gebärdensprache - DGS) [@becker2009], pointing out to the fact that the mastery of reference tracking in narratives is one of the most protracted achievements in the acquisition of signed discourse.
+These findings are also comparable across different sign languages including BSL [@morgan2005], ASL [@loew1984], and DGS [@becker2009], pointing out to the fact that the mastery of reference tracking in narratives is one of the most protracted achievements in the acquisition of signed discourse.
 
@@
 She reported that late signing children generated more explicit forms such as nominals when referring to the characters in the story compared to native learners, concluding that first language deprivation results in an inadequate representation of the addressee and less efficient use of morphosyntax.
-Comparably, @cormier2013 investigated the use of constructed action in 15 BSL among deaf native and late adult signers and found that late signers were redundant in topic continuity.
+Comparably, @cormier2013 investigated the use of CA in 15 BSL among deaf native and late adult signers and found that late signers were redundant in topic continuity.
 In other words, the late signers produced more explicit forms like nominals, and fewer instances of null marking when maintaining previous referents.
 Most relevantly to the present study, @gür2018 studied the development of narrative skills among 10 native and 10 late deaf children and deaf adult signers of TİD.
-She found that native-signing children and adults had more instances of explicit referent introduction, set the scene more frequently, and employed constructed action in an appropriate manner more than corresponding late-signing children and adults.
+She found that native-signing children and adults had more instances of explicit referent introduction, set the scene more frequently, and employed CA in an appropriate manner more than corresponding late-signing children and adults.
 Although this study presented preliminary findings on age of acquisition effects on narrative competence, it does not investigate how native and late signers employ different linguistic tools based on the accessibility of the referent.
@@
 
-In an attempt at addressing this gap in the literature, the present study aims to investigate the reference tracking strategies of native and late adult signers in Turkish Sign Language (Türk İşaret Dili - TİD) narratives with a story-telling paradigm.
+In an attempt at addressing this gap in the literature, the present study aims to investigate the reference tracking strategies of native and late adult signers in TİD narratives with a story-telling paradigm.
 We explore the forms available for referent tracking in TİD and examine the nature of RE selection as a possible indicator of pragmatic competence between our acquisition groups.
@@
 Therefore, most signers preferred to narrate to the camera.
-Furhermore, some signers (4 in number) did not feel comfortable narrating to the camera and, as a result, narrated to the deaf research assistant, if available.
-We ran an additional model excluding the 4 signers who narrated to the experimenter and compared it with the original model including these signers in Appendix.
-The results imply that there is not much of a difference.}.
+Furhermore, some signers did not feel comfortable narrating to the camera and, as a result, narrated to the deaf research assistant, if available.
+We ran an additional model excluding the 4 signers who narrated to the experimenter and compared it with the original model including these signers (see \ref{section:c} The results imply that there is not much of a difference.}.
 
@@
 Finally, we had CA, agreement, plain tags for null constructions.
-We were unable to observe any instance of HCLs occurring in isolation (i.e., not part of constructed action).
+We were unable to observe any instance of HCLs occurring in isolation (i.e., not part of CA).
 The annotation tiers are presented in Table \ref{tab:tab-2}.
@@
 \small
-\caption{Distribution of all discourse status by referring expression types}
+\caption{Distribution of all discourse status by RE types}
 \label{tab:my-table}
@@
 \small
-\caption{Distribution of nominal referring expressions by discourse status}
+\caption{Distribution of nominals by discourse status}
 \label{tab:my-table2}
@@
 \small
-\caption{Distribution of pronominal referring expressions by discourse status}
+\caption{Distribution of pronominals by discourse status}
 \label{tab:my-table3}
@@
 \small
-\caption{Distribution of classifier referring expressions by discourse status}
+\caption{Distribution of classifiers by discourse status}
 \label{tab:my-table4}
@@
 Figure 1 shows the numerical distribution of the responses by RE type and Group.
+
 ```{=tex}
@@
 \small
-\caption{Distribution of null referring expressions by discourse status}
+\caption{Distribution of null markers by discourse status}
 \label{tab:my-table5}
@@
 ```
-
 ```{r fig:count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), RE Type (Nominal, Pronominal, Classifier, Null), and Participant Group (Native, Late)"}
@@
 However, our model did not converg properly.
-We provide the model results as a plot and table in the Appendix.
+We provide the model results as a plot and table in the \ref{section:d}.
 
@@
 Likewise, in both hierarchies, as markers of highly accessible referents, different types of null markers are further distinguished.
-Plain verbs rank the lowest on the scales, preceded by agreement verbs and constructed action.
+Plain verbs rank the lowest on the scales, preceded by agreement verbs and CA.
 On a similar ground, both scales lack pronouns, which occurred very infrequently in the data.
@@
 While late signers were only partially over-redundant in maintained contexts, they were more so in introduced contexts.
-In line with our study, a tendency for limited over-redundancy has been suggested for late signing children in DGS [@becker2009] and late signing adults in BSL specifically for the use of constructed action [@cormier2013].
+In line with our study, a tendency for limited over-redundancy has been suggested for late signing children in DGS [@becker2009] and late signing adults in BSL specifically for the use of CA[@cormier2013].
 As for TİD, to our knowledge, there is only one study [@gür2018] that directly aimed at examining age of acquisition effects on adult signers' narrative skills.
-@gür2018 observed that late signing deaf adults and children set the scene less frequently and used constructed action less accurately compared to their native counterparts.
+@gür2018 observed that late signing deaf adults and children set the scene less frequently and used CA less accurately compared to their native counterparts.
 However, this particular investigation did not present any data on the distribution of REs between two acquisition groups.
@@
 In line with our expectations, we found that the distribution of REs (nominal, pronominal, classifier, and null marking) followed from previous accounts of saliency and referential accessibility.
-For introduced contexts, signers produced more nominals and ExtCLs, and to maintain a previous referent, hey mainly used null arguments (plain verbs, agreement verbs and constructed action) and certain classifiers (e.g., BPCL and WECL) to track highly accessible referents.
+For introduced contexts, signers produced more nominals and ExtCLs, and to maintain a previous referent, hey mainly used null arguments (CA, and plain and agreement verbs) and certain classifiers (e.g., BPCL and WECL) to track highly accessible referents.
 We observed only very few instances of pronouns (mostly produced by late signers), unlike our expectations.
@@
 
+```{=tex}
+\setlength{\columnsep}{10pt}
+\twocolumn
+\linespread{1}
+\justifying
+\footnotesize
+```
+
+# Episodes in the stimuli
+
+\label{section:b}
+
+**Narration 1**
+
+1\. Tom chases Jerry, wearing a wooden plank
+
+2\. Tom and Jerry enter a tube
+
+3\. Tom falls down and Jerry runs away
+
+4\. Tube gets thinner and Tom chases Jerry
+
+5\. Jerry gets out and escapes
+
+6\. Tom follows Jerry in shape of a stick
+
+**Narration 2**
+
+1\. Tom walks past a female cat sitting on a couch
+
+2\. Tom turns to and approaches female cat
+
+3\. Female cat blushes
+
+4\. Tom offers a fish in aquarium to female cat
+
+5\. Female cat rejects offer
+
+6\. Tom offers a bird in cage to female cat
+
+7\. Episode 5 is repeated
+
+8\. Tom leaves by walking in a flirtatious manner
+
+**Narration 3**
+
+1\. Jerry falls down and tries to walk up stairs
+
+2\. Tom pulls carpet back
+
+3\. Jerry struggles
+
+4\. A piano falls down stairs
+
+5\. Jerry manages to escape
+
+6\. Tom sees piano and tries to run away
+
+7\. Piano smashes into Tom and flattens him
+
+8\. Tom falls on ground
+
+**Narration 4**
+
+1\. Tom hides behind a cardboard house
+
+2\. Jerry walks to house with a toy mouse
+
+3\. Toy leaves Jerry who stops at entrance
+
+4\. Toy enters house
+
+5\. Toy gets smashed into pieces
+
+6\. Jerry is surprised
+
+7\. Tom hurts his stomach after swallowing toy
+
+8\. Tom looks at his broken teeth on a mirror
+
+9\. Tom gets angry and throws down mirror
+
+**Narration 5**
+
+1\. Tom fakes sleeping and suddenly gets up
+
+2\. Tom pulls a rope inside a couch
+
+3\. Jerry appears holding a bar of soap
+
+4\. Jerry washes Tom's mouth with bar and flees
+
+5\. Tom spits soap out
+
+6\. Jerry escapes to nest with rope
+
+7\. Tom catches rope and stars pulling it
+
+8\. A mouse trap appears and catches Tom's nose
+
+9\. Tom yells and removes trap
+
+**Narration 6**
+
+1\. Tom points a sword at Jerry and catches him
+
+2\. Tom looks at Jerry
+
+3\. Red spots appear on Jerry's face
+
+4\. Tom gets scared and drops Jerry
+
+5\. Tom rushes into bathroom
+
+6\. Tom washes his hands
+
+7\. Tom opens a drawer
+
+8\. Tom swallows pills one after another
+
+9\. Feeling ill, Tom looks out of window
+
+10\. Tom and Jerry look at each other
+
+**Narration 7**
+
+1\. Two male cats chase one another
+
+2\. Tom sees a female cat and kisses her
+
+3\. Tom hides away when black cat returns
+
+4\. Tom pokes his head behind couch again
+
+5\. Episode 3 is repeated
+
+6\. black cat sees female cat and hugs her
+
+7\. Tom appears and black cat kisses Tom
+
+8\. Two male cats move away from one another
+
+**Narration 8**
+
+1\. Jerry hits Tom's knee with a hammer
+
+2\. Tom suddenly jumps up
+
+3\. Tom yells and holds his knee
+
+4\. Jerry appraches Tom holding a thermometer
+
+5\. Jerry sticks thermometer into Tom's mouth
+
+6\. Jerry lights a fire under thermometer
+
+7\. Red liquid starts to rise up
+
+8\. Thermometer explodes
+
+**Narration 9**
+
+1\. Tom sits on Jerry who tries to get loose
+
+2\. Jerry looks around and sees a hat ribbon
+
+3\. Jerry finds a needle attached to hat
+
+4\. Jerry takes needle and sticks Tom with it
+
+5\. Tom turns red and jumps
+
+6\. Jerry placesn eedle between two cushions
+
+7\. Tom dodges needle
+
+**Narration 10**
+
+1\. Wearing a stethoscope, Tom catches Jerry
+
+2\. Jerry yells into stethoscope
+
+3\. Tom's head pops up
+
+4\. Jerry runs and enter a mouse nest
+
+5\. Tom inserts a gun into nest
+
+6\. Gun bends over toward Tom
+
+7\. Tom fires gun and his head explodes
+
+\clearpage
+
+\onecolumn
+
 # Effect of narrating to the camera vs. experimenter
-\label{section:b}
+
+\label{section:c}
+
 ```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
@@
 
-\label{section:c}
+\label{section:d}
 
@@
 
-```{=tex}
-\setlength{\columnsep}{10pt}
-\twocolumn
-\linespread{1}
-\justifying
-\footnotesize
-```
-# Episodes in the stimuli
-
-\label{section:b}
-
-**Narration 1**
-
-1\. Tom chases Jerry, wearing a wooden plank
-
-2\. Tom and Jerry enter a tube
-
-3\. Tom falls down and Jerry runs away
-
-4\. Tube gets thinner and Tom chases Jerry
-
-5\. Jerry gets out and escapes
-
-6\. Tom follows Jerry in shape of a stick
-
-**Narration 2**
-
-1\. Tom walks past a female cat sitting on a couch
-
-2\. Tom turns to and approaches female cat
-
-3\. Female cat blushes
-
-4\. Tom offers a fish in aquarium to female cat
-
-5\. Female cat rejects offer
-
-6\. Tom offers a bird in cage to female cat
-
-7\. Episode 5 is repeated
-
-8\. Tom leaves by walking in a flirtatious manner
-
-**Narration 3**
-
-1\. Jerry falls down and tries to walk up stairs
-
-2\. Tom pulls carpet back
-
-3\. Jerry struggles
-
-4\. A piano falls down stairs
-
-5\. Jerry manages to escape
-
-6\. Tom sees piano and tries to run away
-
-7\. Piano smashes into Tom and flattens him
-
-8\. Tom falls on ground
-
-**Narration 4**
-
-1\. Tom hides behind a cardboard house
-
-2\. Jerry walks to house with a toy mouse
-
-3\. Toy leaves Jerry who stops at entrance
-
-4\. Toy enters house
-
-5\. Toy gets smashed into pieces
-
-6\. Jerry is surprised
-
-7\. Tom hurts his stomach after swallowing toy
-
-8\. Tom looks at his broken teeth on a mirror
-
-9\. Tom gets angry and throws down mirror
-
-**Narration 5**
-
-1\. Tom fakes sleeping and suddenly gets up
-
-2\. Tom pulls a rope inside a couch
-
-3\. Jerry appears holding a bar of soap
-
-4\. Jerry washes Tom's mouth with bar and flees
-
-5\. Tom spits soap out
-
-6\. Jerry escapes to nest with rope
-
-7\. Tom catches rope and stars pulling it
-
-8\. A mouse trap appears and catches Tom's nose
-
-9\. Tom yells and removes trap
-
-**Narration 6**
-
-1\. Tom points a sword at Jerry and catches him
-
-2\. Tom looks at Jerry
-
-3\. Red spots appear on Jerry's face
-
-4\. Tom gets scared and drops Jerry
-
-5\. Tom rushes into bathroom
-
-6\. Tom washes his hands
-
-7\. Tom opens a drawer
-
-8\. Tom swallows pills one after another
-
-9\. Feeling ill, Tom looks out of window
-
-10\. Tom and Jerry look at each other
-
-**Narration 7**
-
-1\. Two male cats chase one another
-
-2\. Tom sees a female cat and kisses her
-
-3\. Tom hides away when black cat returns
-
-4\. Tom pokes his head behind couch again
-
-5\. Episode 3 is repeated
-
-6\. black cat sees female cat and hugs her
-
-7\. Tom appears and black cat kisses Tom
-
-8\. Two male cats move away from one another
-
-**Narration 8**
-
-1\. Jerry hits Tom's knee with a hammer
-
-2\. Tom suddenly jumps up
-
-3\. Tom yells and holds his knee
-
-4\. Jerry appraches Tom holding a thermometer
-
-5\. Jerry sticks thermometer into Tom's mouth
-
-6\. Jerry lights a fire under thermometer
-
-7\. Red liquid starts to rise up
-
-8\. Thermometer explodes
-
-**Narration 9**
-
-1\. Tom sits on Jerry who tries to get loose
-
-2\. Jerry looks around and sees a hat ribbon
-
-3\. Jerry finds a needle attached to hat
-
-4\. Jerry takes needle and sticks Tom with it
-
-5\. Tom turns red and jumps
-
-6\. Jerry placesn eedle between two cushions
-
-7\. Tom dodges needle
-
-**Narration 10**
-
-1\. Wearing a stethoscope, Tom catches Jerry
-
-2\. Jerry yells into stethoscope
-
-3\. Tom's head pops up
-
-4\. Jerry runs and enter a mouse nest
-
-5\. Tom inserts a gun into nest
-
-6\. Gun bends over toward Tom
-
-7\. Tom fires gun and his head explodes \clearpage
-
-\onecolumn
-
 # References

### 2022-08-05 manuscript_v2.tex
updated files

@@
 economy of form. When multiple expressions are available for the same
-referent as in an utterance like (1), phonetically simplest form is
-selected with the least redundant information are selected based on an
-economy principle. In (1), REs for ``Eddie'' are given in bold. All
-three REs (nominal, pronominal, and null marking) refer to the same
-entity; however, the quantity of marking decreases as the referent's
-accessibility increases. On this ground, for highly accessible referents
-(i.e., those that are highly active in the mind), linguistic forms with
-the least phonetic content are selected for the least amount of
-processing effort.
+referent as in an utterance like (\ref{ex:1}), phonetically simplest
+form is selected with the least redundant information are selected based
+on an economy principle. In the utterance, REs for ``Eddie'' are given
+in bold. All three REs (nominal, pronominal, and null marking) refer to
+the same entity; however, the quantity of marking decreases as the
+referent's accessibility increases. On this ground, for highly
+accessible referents (i.e., those that are highly active in the mind),
+linguistic forms with the least phonetic content are selected for the
+least amount of processing effort.
 
 \begin{exe}
-    \ex
+    \ex \label{ex:1}
     So \textbf{Eddie} turned around. \hfill Williams (1988; p. 343, ex. 7) \\
@@
 languages such as Russian, Chinese, and Japanese, American Sign Language
-(ASL) signers use bare nouns for introduced and reintroduced discourse
+(ASL) signers use bare nouns for introduced and re-introduced discourse
 contexts whereas null marking is used for referents which are
@@
 maxim of quantity and were thus economic in their reference tracking.
-Similar findings have been reported for German Sign Language (DGS)
-(Perniss and Özyürek, 2015) and British Sign Language (BSL) (Morgan,
-2005). It is important to note, however, that these research either
-collapsed native and late signers into a single group and labeled them
-as ``native-like'' or explicitly compared native deaf signers with late
-deaf signers (with language deprivation) in terms of narrative
-competence.
+Similar findings have been reported for German Sign Language (Deutsche
+Gebärdensprache - DGS) (Perniss and Özyürek, 2015) and British Sign
+Language (BSL) (Morgan, 2005). It is important to note, however, that
+these research either collapsed native and late signers into a single
+group and labeled them as ``native-like'' or explicitly compared native
+deaf signers with late deaf signers (with language deprivation) in terms
+of narrative competence.
 
@@
 Commonly in both spoken and sign languages nouns and noun phrases are
-used as referring expressions. Spoken languages use pronouns, determines
-and demonstratives where sign languages use a pointing sign, usually
-glossed as IX, which assumes similar functions (Lillo-Martin and Klima,
-1990). In addition, the pointing sign is used to referentially anchor a
+used as REs. Spoken languages use pronouns, determines and
+demonstratives where sign languages use a pointing sign, usually glossed
+as IX, which assumes similar functions (Lillo-Martin and Klima, 1990).
+In addition, the pointing sign is used to referentially anchor a
 referent to a location in signing space. For instance, when the signer
@@
 right. This referential locus can later be used by agreement verbs and
-classifier predicates (more below). For instance, in (2a) there are two
-referents, \textsc{elvan}, who is referentially-assigned to the right
-side of the signer, represented with the ``a'' subscript and
+classifier predicates (more below). For instance, in (\ref{ex:2a}) there
+are two referents, \textsc{elvan}, who is referentially-assigned to the
+right side of the signer, represented with the ``a'' subscript and
 \textsc{demet}, who is referentially-assigned to the left side of the
 signer, represented with the ``b'' subscript. When the signer makes a
-contrast in (2b), as a pronominal use, she points to the right side to
-refer to ELVAN and to the left to refer to \textsc{demet}. We
+contrast in (\ref{ex:2b}), as a pronominal use, she points to the right
+side to refer to ELVAN and to the left to refer to \textsc{demet}. We
 schematically represents the referential locus assigned to the referents
-in (3).
+in (\ref{ex:3}).
 
@@
 denote are below. Some classifier handshapes and the set of objects that
-they denote are presented below in (4).
+they denote are presented below in (\ref{ex:4}).
 
 \begin{exe}
-\ex 
+\ex \label{ex:4}
 \includegraphics[width=25pt]{pictures/flathand.png}: flat entities, surfaces, vehicles.\\
@@
 
-In (5), in the first clause, a referential locus, i.e.~locus-a, is
-assigned to the locative source argument \textsc{cage}. In the second
+In (\ref{ex:5}), in the first clause, a referential locus, i.e.~locus-a,
+is assigned to the locative source argument \textsc{cage}. In the second
 clause, the core argument \textsc{bird} is cross-referenced with the
@@
 In TİD and many other sign languages, it is possible to use bare noun
-phrases (\textsc{bare nom}) as in (1a) as well as fingerspelled nouns
-(\textsc{fs nom}) as in (1c). Fingerspelling refers to the use of the
-manual alphabet which is derived or borrowed from a surrounding written
-language (Kubuş, 2008). Nouns can also be modified (\textsc{mod nom}) by
-classifiers as in (1b) (Frederiksen and Mayberry, 2016) or by a
-prenominal or postnominal IX functioning as a determiner or a
-demonstrative as in (1d) (Maclaughlin, 1997).
+phrases as in (\ref{ex:6a}) as well as fingerspelled nouns as in
+(\ref{ex:6c}). Fingerspelling refers to the use of the manual alphabet
+which is derived or borrowed from a surrounding written language (Kubuş,
+2008). Nouns can also be modified by classifiers as in (\ref{ex:6b})
+(Frederiksen and Mayberry, 2016) or by a prenominal or postnominal IX
+functioning as a determiner or a demonstrative as in (\ref{ex:6d})
+(Maclaughlin, 1997).
 
@@
 discourse (Emmorey, 1996). The same IX function has also been identified
-for TİD (Sevinç, 2006). The example in (2) illustrates a
+for TİD (Sevinç, 2006). The example in (\ref{ex:7}) illustrates a
 spatially-anchored pronominal reference for topic re-introduction. These
@@
 entity by following a trajectory to trace its shape. We exemplify these
-constructions in (3).
+constructions in (\ref{ex:8}).
 
@@
 mental representation of the addressee. In signed discourse, this can be
-achieved by employing certain verbal constructions. To identify the
-predicate types in ASL, Padden (1986) suggested a triadic categorization
-of verbs in sign languages: plain verbs, agreement verbs, and spatial
-verbs. According to this classification, plain verbs (plain) (4b) do not
-agree with the subject or the object. This means that they are not
-inflected for number or person. This contrasts with agreeing verbs
-(agreement) (4c) the starting or the ending point of which can be
-modified to specific loci for person and number agreement while spatial
-verbs (as in \textsc{ix book put} `I placed the book there') agree with
-locative arguments but like plain verbs, they are not inflected for
-number or person. TİD too licenses null arguments. Following this
-classification, Sevinç (2006) argued that although it is common in TİD
-to drop arguments, there were certain limitations on the null marking of
-the arguments for agreement and plain verbs.
+achieved by employing certain verbal constructions or constructed action
+(CA). The latter construction exemplified in (\ref{ex:9a}) refers to
+cases in which the signers engage with the space in front of themselves
+as if they are interacting with the objects and characters in the story.
+This is often achieved with the signer taking on the role of a referent
+in the narrative and performing the actions (Metzger, 1995). CA is often
+marked by certain non-manual cues such as the facial expression and the
+breaking of the eye contact. This is to say, the narrator's head, face,
+torso, arms, and as an optional marker, hands represent those of the
+associated referent(s) (Smith and Cormier, 2014). As for the verbal
+constructions, to identify the predicate types in ASL, Padden (1986)
+suggested a triadic categorization of verbs in sign languages: plain
+verbs, agreement verbs, and spatial verbs. According to this
+classification, plain verbs (\ref{ex:9b}) do not agree with the subject
+or the object. This means that they are not inflected for number or
+person. This contrasts with agreeing verbs (\ref{ex:9c}) the starting or
+the ending point of which can be modified to specific loci for person
+and number agreement while spatial verbs (as in \textsc{ix book put} `I
+placed the book there') agree with locative arguments but like plain
+verbs, they are not inflected for number or person. TİD too licenses
+null arguments. Following this classification, Sevinç (2006) argued that
+although it is common in TİD to drop arguments, there were certain
+limitations on the null marking of the arguments for agreement and plain
+verbs.
 
@@
 beginning of formal education. These findings are also comparable across
-different sign languages including British Sign Language (BSL) (Morgan,
-2005), ASL (Loew, 1984), and German Sign Language (Deutsche
-Gebärdensprache - DGS) (Becker, 2009), pointing out to the fact that the
-mastery of reference tracking in narratives is one of the most
-protracted achievements in the acquisition of signed discourse.
+different sign languages including BSL (Morgan, 2005), ASL (Loew, 1984),
+and DGS (Becker, 2009), pointing out to the fact that the mastery of
+reference tracking in narratives is one of the most protracted
+achievements in the acquisition of signed discourse.
 
@@
 less efficient use of morphosyntax. Comparably, Cormier et al. (2013)
-investigated the use of constructed action in 15 BSL among deaf native
-and late adult signers and found that late signers were redundant in
-topic continuity. In other words, the late signers produced more
-explicit forms like nominals, and fewer instances of null marking when
-maintaining previous referents. Most relevantly to the present study,
-Gür (2018) studied the development of narrative skills among 10 native
-and 10 late deaf children and deaf adult signers of TİD. She found that
-native-signing children and adults had more instances of explicit
-referent introduction, set the scene more frequently, and employed
-constructed action in an appropriate manner more than corresponding
-late-signing children and adults. Although this study presented
-preliminary findings on age of acquisition effects on narrative
-competence, it does not investigate how native and late signers employ
-different linguistic tools based on the accessibility of the referent.
-Therefore, to our knowledge, there is a lack of research directly
-examining the effects of first language deprivation on deaf adult
-signers' selection of REs in signed narratives.
+investigated the use of CA in 15 BSL among deaf native and late adult
+signers and found that late signers were redundant in topic continuity.
+In other words, the late signers produced more explicit forms like
+nominals, and fewer instances of null marking when maintaining previous
+referents. Most relevantly to the present study, Gür (2018) studied the
+development of narrative skills among 10 native and 10 late deaf
+children and deaf adult signers of TİD. She found that native-signing
+children and adults had more instances of explicit referent
+introduction, set the scene more frequently, and employed CA in an
+appropriate manner more than corresponding late-signing children and
+adults. Although this study presented preliminary findings on age of
+acquisition effects on narrative competence, it does not investigate how
+native and late signers employ different linguistic tools based on the
+accessibility of the referent. Therefore, to our knowledge, there is a
+lack of research directly examining the effects of first language
+deprivation on deaf adult signers' selection of REs in signed
+narratives.
 
@@
 study aims to investigate the reference tracking strategies of native
-and late adult signers in Turkish Sign Language (Türk İşaret Dili - TİD)
-narratives with a story-telling paradigm. We explore the forms available
-for referent tracking in TİD and examine the nature of RE selection as a
-possible indicator of pragmatic competence between our acquisition
-groups.
+and late adult signers in TİD narratives with a story-telling paradigm.
+We explore the forms available for referent tracking in TİD and examine
+the nature of RE selection as a possible indicator of pragmatic
+competence between our acquisition groups.
 
@@
 were unable to observe any instance of HCLs occurring in isolation
-(i.e., not part of constructed action). The annotation tiers are
-presented in Table \ref{tab:tab-2}.
+(i.e., not part of CA). The annotation tiers are presented in Table
+\ref{tab:tab-2}.
 
@@
   model did not converg properly. We provide the model results as a plot
-  and table in the Appendix.} shows the posterior probability
+  and table in the \ref{section:d}.} shows the posterior probability
 distributions where the dot is the median estimate, the thick line is
@@
 of null markers are further distinguished. Plain verbs rank the lowest
-on the scales, preceded by agreement verbs and constructed action. On a
-similar ground, both scales lack pronouns, which occurred very
-infrequently in the data. Although Frederiksen and Mayberry (2016) did
-not include other classifier types in the hierarchy, we proposed that
-WECLs and BPCLs must ranker higher than null markers but lower than
-ExtCLs that were mainly used by interlocutors to introduce a referent
-into the discourse context.
+on the scales, preceded by agreement verbs and CA. On a similar ground,
+both scales lack pronouns, which occurred very infrequently in the data.
+Although Frederiksen and Mayberry (2016) did not include other
+classifier types in the hierarchy, we proposed that WECLs and BPCLs must
+ranker higher than null markers but lower than ExtCLs that were mainly
+used by interlocutors to introduce a referent into the discourse
+context.
 
@@
 late signing children in DGS (Becker, 2009) and late signing adults in
-BSL specifically for the use of constructed action (Cormier et al.,
-2013). As for TİD, to our knowledge, there is only one study (Gür, 2018)
-that directly aimed at examining age of acquisition effects on adult
-signers' narrative skills. Gür (2018) observed that late signing deaf
-adults and children set the scene less frequently and used constructed
-action less accurately compared to their native counterparts. However,
-this particular investigation did not present any data on the
-distribution of REs between two acquisition groups. In the present
-study, the finding that native signers preferred to use null arguments
-even for the first mention of a referent (19 in total) is surprising
-given some of the previous literature in the topicality and saliency
-framework (Cormier et al., 2013; Swabey, 2002). These research reported
-very little or no use of null marking by native signers for introduced
-contexts.
+BSL specifically for the use of CA(Cormier et al., 2013). As for TİD, to
+our knowledge, there is only one study (Gür, 2018) that directly aimed
+at examining age of acquisition effects on adult signers' narrative
+skills. Gür (2018) observed that late signing deaf adults and children
+set the scene less frequently and used CA less accurately compared to
+their native counterparts. However, this particular investigation did
+not present any data on the distribution of REs between two acquisition
+groups. In the present study, the finding that native signers preferred
+to use null arguments even for the first mention of a referent (19 in
+total) is surprising given some of the previous literature in the
+topicality and saliency framework (Cormier et al., 2013; Swabey, 2002).
+These research reported very little or no use of null marking by native
+signers for introduced contexts.
 
@@
 contexts, signers produced more nominals and ExtCLs, and to maintain a
-previous referent, hey mainly used null arguments (plain verbs,
-agreement verbs and constructed action) and certain classifiers (e.g.,
-BPCL and WECL) to track highly accessible referents. We observed only
-very few instances of pronouns (mostly produced by late signers), unlike
-our expectations. The ones we observed were mostly reserved for
-re-introduced contexts along with nominals. Based on our findings across
-the two groups, we proposed a hierarchy of referent accessibility for
-TİD. As for the age of acquisition effects, we observed limited
-over-redundancy for signers with delayed language acquisition (i.e.,
-late signers) who employed fewer implicit forms of reference tracking
-compared to native signers. Native signers were less likely to use a
-nominal for maintained contexts and more likely to use null marking for
-introduced contexts than late signers. We conclude that native signers
-might only partially adhere to pragmatic norms and leverage
-accessibility better than late signers since they use less form when the
-referent is more active in the mind.
+previous referent, hey mainly used null arguments (CA, and plain and
+agreement verbs) and certain classifiers (e.g., BPCL and WECL) to track
+highly accessible referents. We observed only very few instances of
+pronouns (mostly produced by late signers), unlike our expectations. The
+ones we observed were mostly reserved for re-introduced contexts along
+with nominals. Based on our findings across the two groups, we proposed
+a hierarchy of referent accessibility for TİD. As for the age of
+acquisition effects, we observed limited over-redundancy for signers
+with delayed language acquisition (i.e., late signers) who employed
+fewer implicit forms of reference tracking compared to native signers.
+Native signers were less likely to use a nominal for maintained contexts
+and more likely to use null marking for introduced contexts than late
+signers. We conclude that native signers might only partially adhere to
+pragmatic norms and leverage accessibility better than late signers
+since they use less form when the referent is more active in the mind.
 
@@
 
+\setlength{\columnsep}{10pt}
+\twocolumn
+\linespread{1}
+\justifying
+\footnotesize
+
+\hypertarget{episodes-in-the-stimuli}{%
+\section{Episodes in the stimuli}\label{episodes-in-the-stimuli}}
+
+\label{section:b}
+
+\textbf{Narration 1}
+
+1. Tom chases Jerry, wearing a wooden plank
+
+2. Tom and Jerry enter a tube
+
+3. Tom falls down and Jerry runs away
+
+4. Tube gets thinner and Tom chases Jerry
+
+5. Jerry gets out and escapes
+
+6. Tom follows Jerry in shape of a stick
+
+\textbf{Narration 2}
+
+1. Tom walks past a female cat sitting on a couch
+
+2. Tom turns to and approaches female cat
+
+3. Female cat blushes
+
+4. Tom offers a fish in aquarium to female cat
+
+5. Female cat rejects offer
+
+6. Tom offers a bird in cage to female cat
+
+7. Episode 5 is repeated
+
+8. Tom leaves by walking in a flirtatious manner
+
+\textbf{Narration 3}
+
+1. Jerry falls down and tries to walk up stairs
+
+2. Tom pulls carpet back
+
+3. Jerry struggles
+
+4. A piano falls down stairs
+
+5. Jerry manages to escape
+
+6. Tom sees piano and tries to run away
+
+7. Piano smashes into Tom and flattens him
+
+8. Tom falls on ground
+
+\textbf{Narration 4}
+
+1. Tom hides behind a cardboard house
+
+2. Jerry walks to house with a toy mouse
+
+3. Toy leaves Jerry who stops at entrance
+
+4. Toy enters house
+
+5. Toy gets smashed into pieces
+
+6. Jerry is surprised
+
+7. Tom hurts his stomach after swallowing toy
+
+8. Tom looks at his broken teeth on a mirror
+
+9. Tom gets angry and throws down mirror
+
+\textbf{Narration 5}
+
+1. Tom fakes sleeping and suddenly gets up
+
+2. Tom pulls a rope inside a couch
+
+3. Jerry appears holding a bar of soap
+
+4. Jerry washes Tom's mouth with bar and flees
+
+5. Tom spits soap out
+
+6. Jerry escapes to nest with rope
+
+7. Tom catches rope and stars pulling it
+
+8. A mouse trap appears and catches Tom's nose
+
+9. Tom yells and removes trap
+
+\textbf{Narration 6}
+
+1. Tom points a sword at Jerry and catches him
+
+2. Tom looks at Jerry
+
+3. Red spots appear on Jerry's face
+
+4. Tom gets scared and drops Jerry
+
+5. Tom rushes into bathroom
+
+6. Tom washes his hands
+
+7. Tom opens a drawer
+
+8. Tom swallows pills one after another
+
+9. Feeling ill, Tom looks out of window
+
+10. Tom and Jerry look at each other
+
+\textbf{Narration 7}
+
+1. Two male cats chase one another
+
+2. Tom sees a female cat and kisses her
+
+3. Tom hides away when black cat returns
+
+4. Tom pokes his head behind couch again
+
+5. Episode 3 is repeated
+
+6. black cat sees female cat and hugs her
+
+7. Tom appears and black cat kisses Tom
+
+8. Two male cats move away from one another
+
+\textbf{Narration 8}
+
+1. Jerry hits Tom's knee with a hammer
+
+2. Tom suddenly jumps up
+
+3. Tom yells and holds his knee
+
+4. Jerry appraches Tom holding a thermometer
+
+5. Jerry sticks thermometer into Tom's mouth
+
+6. Jerry lights a fire under thermometer
+
+7. Red liquid starts to rise up
+
+8. Thermometer explodes
+
+\textbf{Narration 9}
+
+1. Tom sits on Jerry who tries to get loose
+
+2. Jerry looks around and sees a hat ribbon
+
+3. Jerry finds a needle attached to hat
+
+4. Jerry takes needle and sticks Tom with it
+
+5. Tom turns red and jumps
+
+6. Jerry placesn eedle between two cushions
+
+7. Tom dodges needle
+
+\textbf{Narration 10}
+
+1. Wearing a stethoscope, Tom catches Jerry
+
+2. Jerry yells into stethoscope
+
+3. Tom's head pops up
+
+4. Jerry runs and enter a mouse nest
+
+5. Tom inserts a gun into nest
+
+6. Gun bends over toward Tom
+
+7. Tom fires gun and his head explodes
+
+\clearpage
+
+\onecolumn
+
 \hypertarget{effect-of-narrating-to-the-camera-vs.-experimenter}{%
@@
 
-\label{section:b}
+\label{section:c}
 
@@
 
-\setlength{\columnsep}{10pt}
-\twocolumn
-\linespread{1}
-\justifying
-\footnotesize
-
-\hypertarget{episodes-in-the-stimuli}{%
-\section{Episodes in the stimuli}\label{episodes-in-the-stimuli}}
-
-\label{section:b}
-
-\textbf{Narration 1}
-
-1. Tom chases Jerry, wearing a wooden plank
-
-2. Tom and Jerry enter a tube
-
-3. Tom falls down and Jerry runs away
-
-4. Tube gets thinner and Tom chases Jerry
-
-5. Jerry gets out and escapes
-
-6. Tom follows Jerry in shape of a stick
-
-\textbf{Narration 2}
-
-1. Tom walks past a female cat sitting on a couch
-
-2. Tom turns to and approaches female cat
-
-3. Female cat blushes
-
-4. Tom offers a fish in aquarium to female cat
-
-5. Female cat rejects offer
-
-6. Tom offers a bird in cage to female cat
-
-7. Episode 5 is repeated
-
-8. Tom leaves by walking in a flirtatious manner
-
-\textbf{Narration 3}
-
-1. Jerry falls down and tries to walk up stairs
-
-2. Tom pulls carpet back
-
-3. Jerry struggles
-
-4. A piano falls down stairs
-
-5. Jerry manages to escape
-
-6. Tom sees piano and tries to run away
-
-7. Piano smashes into Tom and flattens him
-
-8. Tom falls on ground
-
-\textbf{Narration 4}
-
-1. Tom hides behind a cardboard house
-
-2. Jerry walks to house with a toy mouse
-
-3. Toy leaves Jerry who stops at entrance
-
-4. Toy enters house
-
-5. Toy gets smashed into pieces
-
-6. Jerry is surprised
-
-7. Tom hurts his stomach after swallowing toy
-
-8. Tom looks at his broken teeth on a mirror
-
-9. Tom gets angry and throws down mirror
-
-\textbf{Narration 5}
-
-1. Tom fakes sleeping and suddenly gets up
-
-2. Tom pulls a rope inside a couch
-
-3. Jerry appears holding a bar of soap
-
-4. Jerry washes Tom's mouth with bar and flees
-
-5. Tom spits soap out
-
-6. Jerry escapes to nest with rope
-
-7. Tom catches rope and stars pulling it
-
-8. A mouse trap appears and catches Tom's nose
-
-9. Tom yells and removes trap
-
-\textbf{Narration 6}
-
-1. Tom points a sword at Jerry and catches him
-
-2. Tom looks at Jerry
-
-3. Red spots appear on Jerry's face
-
-4. Tom gets scared and drops Jerry
-
-5. Tom rushes into bathroom
-
-6. Tom washes his hands
-
-7. Tom opens a drawer
-
-8. Tom swallows pills one after another
-
-9. Feeling ill, Tom looks out of window
-
-10. Tom and Jerry look at each other
-
-\textbf{Narration 7}
-
-1. Two male cats chase one another
-
-2. Tom sees a female cat and kisses her
-
-3. Tom hides away when black cat returns
-
-4. Tom pokes his head behind couch again
-
-5. Episode 3 is repeated
-
-6. black cat sees female cat and hugs her
-
-7. Tom appears and black cat kisses Tom
-
-8. Two male cats move away from one another
-
-\textbf{Narration 8}
-
-1. Jerry hits Tom's knee with a hammer
-
-2. Tom suddenly jumps up
-
-3. Tom yells and holds his knee
-
-4. Jerry appraches Tom holding a thermometer
-
-5. Jerry sticks thermometer into Tom's mouth
-
-6. Jerry lights a fire under thermometer
-
-7. Red liquid starts to rise up
-
-8. Thermometer explodes
-
-\textbf{Narration 9}
-
-1. Tom sits on Jerry who tries to get loose
-
-2. Jerry looks around and sees a hat ribbon
-
-3. Jerry finds a needle attached to hat
-
-4. Jerry takes needle and sticks Tom with it
-
-5. Tom turns red and jumps
-
-6. Jerry placesn eedle between two cushions
-
-7. Tom dodges needle
-
-\textbf{Narration 10}
-
-1. Wearing a stethoscope, Tom catches Jerry
-
-2. Jerry yells into stethoscope
-
-3. Tom's head pops up
-
-4. Jerry runs and enter a mouse nest
-
-5. Tom inserts a gun into nest
-
-6. Gun bends over toward Tom
-
-7. Tom fires gun and his head explodes \clearpage
-
-\onecolumn
-
 \hypertarget{references}{%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-ariel1990}{}}%
-Ariel, Mira, 1990. Accessing Noun-Phrase Antecedents. Routledge.
+Ariel, Mira, 1990. Accessing noun-phrase antecedents. Routledge.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-dudis2004}{}}%
-Dudis, Paul G., 2004. Body partitioning and real-space blends. Cognitive
-Linguistics 15.
-doi:\href{https://doi.org/10.1515/cogl.2004.009}{10.1515/cogl.2004.009}
+Dudis, Paul G., 2004.
+\href{https://www.degruyter.com/view/j/cogl.2004.15.issue-2/cogl.2004.009/cogl.2004.009.xml}{Body
+partitioning and real-space blends}. Cognitive Linguistics 15.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-frederiksen2016}{}}%
-Frederiksen, Anne Therese, Mayberry, Rachel I., 2016. Who's on First?
+Frederiksen, Anne Therese, Mayberry, Rachel I., 2016. Who's on first?
 Investigating the referential hierarchy in simple native ASL narratives.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-gullberg2006}{}}%
-Gullberg, Marianne, 2006. Handling Discourse: Gestures, Reference
-Tracking, and Communication Strategies in Early L2: Language Learning
+Gullberg, Marianne, 2006. Handling discourse: Gestures, reference
+tracking, and communication strategies in early L2: Language Learning
 Vol. 56, No. 1. Language Learning 56, 155--196.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-guxfcr2018}{}}%
-Gür, Cansu, 2018. The Effects of Late Language Acquisition on the
-Development of Narrative Skills: Insights from Turkish Sign Language
+Gür, Cansu, 2018. The effects of late language acquisition on the
+development of narrative skills: Insights from Turkish Sign Language
 (PhD thesis). Erzurum.
@@
 Karadöller, Dilay Z, Sümer, Beyza, Özyürek, Aslı, 2017. Effects of
-Delayed Language Exposure on Spatial Language Acquisition by Signing
-Children and Adults 5.
+delayed language exposure on spatial language acquisition by signing
+children and adults 5.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-kubu2008}{}}%
-Kubuş, Okan, 2008. AN ANALYSIS OF TURKISH SIGN LANGUAGE (T{İ}D)
-PHONOLOGY AND MORPHOLOGY (PhD thesis). Ankara.
+Kubuş, Okan, 2008. An analysis of Turkish Sign Language (T{İ}D)
+phonology and morphology (PhD thesis). Ankara.
 
@@
 Maclaughlin, Dawn, 1997.
-\href{http://lear.unive.it/jspui/handle/11707/7016}{The Structure of
-Determiner Phrases: Evidence From American Sign Language.} (PhD thesis).
+\href{http://lear.unive.it/jspui/handle/11707/7016}{The structure of
+determiner phrases: Evidence from American Sign Language.} (PhD thesis).
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-metzger1995}{}}%
+Metzger, Melanie, 1995. Constructed action and constructed dialogue in
+american sign language. Sociolinguistics in Deaf communities 255271.
+
 \leavevmode\vadjust pre{\hypertarget{ref-morgan2005}{}}%
@@
 \href{http://www.oxfordscholarship.com/view/10.1093/acprof:oso/9780195180947.001.0001/acprof-9780195180947}{The
-Development of Narrative Skills in British Sign Language}, in: Schick,
+development of narrative skills in British Sign Language}, in: Schick,
 Brenda, Marschark, Marc, Spencer, Patricia Elizabeth (Eds.),. Oxford
@@
 \leavevmode\vadjust pre{\hypertarget{ref-newport1990}{}}%
-Newport, Elissa L., 1990. Maturational Constraints on Language Learning.
+Newport, Elissa L., 1990. Maturational constraints on language learning.
 Cognitive Science 14, 11--28.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-perniss2015}{}}%
-Perniss, Pamela, Özyürek, Asli, 2015. Visible Cohesion: A Comparison of
-Reference Tracking in Sign, Speech, and Co-Speech Gesture. Topics in
+Perniss, Pamela, Özyürek, Asli, 2015. Visible Cohesion: A comparison of
+reference tracking in sign, speech, and co-speech gesture. Topics in
 Cognitive Science 7, 36--60.
@@
 Rathmann, Christian, Mann, Wolfgang, Morgan, Gary, 2007. Narrative
-Structure and Narrative Development in Deaf Children. Deafness \&
+structure and narrative development in deaf children. Deafness \&
 Education International 9, 187--196.
@@
 Technical University, Turkey.
+
+\leavevmode\vadjust pre{\hypertarget{ref-smith2014}{}}%
+Smith, Sandra, Cormier, Kearsy, 2014. In or out?: Spatial scale and
+enactment in narratives of native and nonnative signing deaf children
+acquiring British Sign Language. Sign Language Studies 14, 275--301.
+doi:\href{https://doi.org/10.1353/sls.2014.0008}{10.1353/sls.2014.0008}
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-swabey2002}{}}%
-Swabey, Laurie Anne, 2002. The Cognitive Status, Form and Distribution
-of Referring Expressions in ASL and English Narratives (PhD thesis).
+Swabey, Laurie Anne, 2002. The cognitive status, form and distribution
+of referring expressions in ASL and English narratives (PhD thesis).
 Minnepolis, USA.
@@
 \leavevmode\vadjust pre{\hypertarget{ref-williams1988}{}}%
-Williams, Jessica, 1988. Zero Anaphora in Second Language Acquisition: A
-Comparison among Three Varieties of English. Studies in Second Language
+Williams, Jessica, 1988. Zero anaphora in second language acquisition: A
+comparison among three varieties of English. Studies in Second Language
 Acquisition 10, 339--370.
@@
 Laurence, Sinte, Aurélie, Van Herreweghe, Mieke, Vermeerbergen, Myriam
-(Eds.),. DE GRUYTER, Berlin, Boston, pp. 15--34.
+(Eds.),. De Gruyter, Berlin, Boston, pp. 15--34.
 doi:\href{https://doi.org/10.1515/9781614511472.15}{10.1515/9781614511472.15}

### 2022-08-05 manuscript_v2.Rmd
updated

@@
 ---
+editor_options:
+  chunk_output_type: console
 title: Reference Tracking Strategies of Native and Late Signers in Turkish Sign Language
@@
 As can be expected, there is a plethora of research examining how narrative competence is achieved in both monolingual and bilingual acquisition contexts among children and adults.
-Although narrative competence and referent selection have been studied extensively for typically developing speakers [@aksu-koç2015; @hickmann1996; @williams1988], deaf individuals who acquired a sign language in atypical learning contexts (i.e., deaf signers with linguistic deprivation) have remained mostly understudied with an exception of handful of studies [@cormier2013; @gür2018; @becker2009].
+Although narrative competence and referent selection have been studied extensively for typically developing speakers [@aksu-koc2015; @hickmann1996; @williams1988], deaf individuals who acquired a sign language in atypical learning contexts (i.e., deaf signers with linguistic deprivation) have remained mostly understudied with an exception of handful of studies [@cormier2013; @gur2018; @becker2009].
 Nevertheless, it is paramount that we investigate and describe the effects of language deprivation among deaf adults on narrative competence more in depth.
@@
 First language deprivation or delay has been found to have negative effects on certain cognitive and linguistic measures among late signing children.
-To name a few, delayed exposure to a first functional language is reported to negatively affect deaf children's morphosyntactic production and processing [@kayaba2022; @boudreault2006; @newport1990; @cheng2021; @karadöller2017; @cormier2012] and narrative abilities [e.g., @becker2009].
+To name a few, delayed exposure to a first functional language is reported to negatively affect deaf children's morphosyntactic production and processing [@kayabasi2020; @boudreault2006; @newport1990; @cheng2021; @karadoller2017; @cormier2012] and narrative abilities [e.g., @becker2009].
 In our overall research, we are concerned with the effects of this late exposure in the ultimate adult forms of TİD users.
@@
 In TİD and many other sign languages, it is possible to use bare noun phrases as in (\ref{ex:6a}) as well as fingerspelled nouns as in (\ref{ex:6c}).
-Fingerspelling refers to the use of the manual alphabet which is derived or borrowed from a surrounding written language [@kubu2008].
+Fingerspelling refers to the use of the manual alphabet which is derived or borrowed from a surrounding written language [@kubus2008].
 Nouns can also be modified by classifiers as in (\ref{ex:6b}) [@frederiksen2016] or by a prenominal or postnominal IX functioning as a determiner or a demonstrative as in (\ref{ex:6d}) [@maclaughlin1997].
@@
 The IX sign (extended index finger) has various functions in sign languages, one of which is the pronominal use which refers to established nominals by pointing to a previously assigned abstract or physical locus in the signing space when re-introducing a topic in discourse [@emmorey1996].
-The same IX function has also been identified for TİD [@sevinç2006].
+The same IX function has also been identified for TİD [@sevinc2006].
 The example in (\ref{ex:7}) illustrates a spatially-anchored pronominal reference for topic re-introduction.
@@
 TİD too licenses null arguments.
-Following this classification, @sevinç2006 argued that although it is common in TİD to drop arguments, there were certain limitations on the null marking of the arguments for agreement and plain verbs.
+Following this classification, @sevinc2006 argued that although it is common in TİD to drop arguments, there were certain limitations on the null marking of the arguments for agreement and plain verbs.
 
@@
 In other words, the late signers produced more explicit forms like nominals, and fewer instances of null marking when maintaining previous referents.
-Most relevantly to the present study, @gür2018 studied the development of narrative skills among 10 native and 10 late deaf children and deaf adult signers of TİD.
+Most relevantly to the present study, @gur2018 studied the development of narrative skills among 10 native and 10 late deaf children and deaf adult signers of TİD.
 She found that native-signing children and adults had more instances of explicit referent introduction, set the scene more frequently, and employed CA in an appropriate manner more than corresponding late-signing children and adults.
@@
 library(tidyverse) # tidy
+library(ggstance)
 library(janitor)
@@
   dplyr::mutate(Subject = as.integer(no),
-                Group = case_when(aoa == "0-3" ~ "native",
-                                  aoa == "4-7" ~ "late1",
-                                  T ~ "late2") %>% as.factor()) %>% dplyr::select(Subject, Group)
+                Group = case_when(aoa == "0-3" ~ "Native",
+                                  aoa == "4-7" ~ "Late1",
+                                  T ~ "Late2") %>% as.factor()) %>% dplyr::select(Subject, Group)
 df_narration %<>% left_join(df_misc)
@@
 contrasts(df_model1$Discourse) <- contr.sum(3)/2
+df_model1$Group %<>% reorder.factor(new.order = c("Late2","Late1","Native"))
 contrasts(df_model1$Group) <- contr.sdif(3)
@@
                                         `b_muNull_Intercept` = "Null_Intercept",
-                                        `b_muNominal_Group2M1` = "Nom_Late1vsNative",
-                                        `b_muNominal_Group3M2` = "Nom_Late2vsLate1",
+                                        `b_muNominal_Group2M1` = "Nom_Late1-Late2",
+                                        `b_muNominal_Group3M2` = "Nom_Native-Late1",
                                         `b_muNominal_Discourse1` = "Nom_Intro",
                                         `b_muNominal_Discourse2` = "Nom_Maintain",
-                                        `b_muNominal_Group2M1:Discourse1` = 'Nom_L1vsN*Intro',
-                                        `b_muNominal_Group3M2:Discourse1` = 'Nom_L2vsL1*Intro',
-                                        `b_muNominal_Group2M1:Discourse2` = 'Nom_L1vsN*Maintain',
-                                        `b_muNominal_Group3M2:Discourse2` = 'Nom_L2vsL1*Maintain',
-                                        `b_muNull_Group2M1` = "Null_Late1vsNative",
-                                        `b_muNull_Group3M2` = "Null_Late2vsLate1",
+                                        `b_muNominal_Group2M1:Discourse1` = 'Nom_L1-L2*Intro',
+                                        `b_muNominal_Group3M2:Discourse1` = 'Nom_N-L1*Intro',
+                                        `b_muNominal_Group2M1:Discourse2` = 'Nom_L1-L2*Maintain',
+                                        `b_muNominal_Group3M2:Discourse2` = 'Nom_N-L1*Maintain',
+                                        `b_muNull_Group2M1` = "Null_Late1-Late2",
+                                        `b_muNull_Group3M2` = "Null_Native-Late1",
                                         `b_muNull_Discourse1` = "Null_Intro",
                                         `b_muNull_Discourse2` = "Null_Maintain",
-                                        `b_muNull_Group2M1:Discourse1` = 'Null_L1vsN*Intro',
-                                        `b_muNull_Group3M2:Discourse1` = 'Null_L2vsL1*Intro',
-                                        `b_muNull_Group2M1:Discourse2` = 'Null_L1vsN*Maintain',
-                                        `b_muNull_Group3M2:Discourse2` = 'Null_L2vsL1*Maintain')
+                                        `b_muNull_Group2M1:Discourse1` = 'Null_L1-L2*Intro',
+                                        `b_muNull_Group3M2:Discourse1` = 'Null_N-L1*Intro',
+                                        `b_muNull_Group2M1:Discourse2` = 'Null_L1-L2*Maintain',
+                                        `b_muNull_Group3M2:Discourse2` = 'Null_N-L1*Maintain')
 saveRDS(df_model_3, "./rds/df_model_3.rds")
 #df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "NULL"))
-
 ```
@@
 
-In our analysis, we use Bayesian regression models with the help of the brms package [@bürkner2018] in R.
+In our analysis, we use Bayesian regression models with the help of the brms package [@burkner2018] in R.
 In a Bayesian regression the prior and the likelihood is used to calculate the posterior.
@@
 ```
-```{r fig:count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), RE Type (Nominal, Pronominal, Classifier, Null), and Participant Group (Native, Late)"}
+
+```{r count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), RE Type (Nominal, Pronominal, Classifier, Null), and Participant Group (Native, Late)"}
 df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "NULL"))
@@
 
-For more inference on the data and having pairwise comparisons, we fit a regression model to RE Type using Discourse Status and Group as predictors.
-We used the brms package [@bürkner2018] in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late).
+For more inference on the data and having pairwise comparisons, we fit a regression model to RE Type using discourse status and acquisition group as predictors.
+We used the brms package [@burkner2018] in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late).
 As the overall number of produced pronouns was very little (40 data points), we excluded pronominal use from the following analyses.
-For the current analysis, we did not define priors ourselves but used the preset uninformed priors.
-Figure 2[^2] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals.
-There are 3 responses we are considering in this model, and they are Classifier, Nominative, and Null.
+For the current analysis, we did not define priors ourselves and used the preset uninformed priors.
+\@ref{fig:model-gram}[^2] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals.
+There are 3 responses we are considering in this model.
+They are; Classifier, Nominative, and Null.
 Classifier response is the level that other responses Nominative and Null are compared against.
@@
 [^2]: We ran another model where we used the Nativeness predictor with 3 levels 0-3:Native, 4-7:Late1, and 7-above:Late2.
-However, our model did not converg properly.
+However, our model did not converge properly.
 We provide the model results as a plot and table in the \ref{section:d}.
 
-```{r fig:model-gram, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
+```{r model-gram, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
 p3 <- df_model_1 %>% 
@@
 In contrast to the introduction status, nominals were used less in maintenance contexts compared to the re-introduction contexts within the greater comparison to classifier referents.
-There does not seem to be an interaction between the introduction status of discourse and being a native acquirer of TİD.
-However, there seems to be an interaction between maintaining a referent and being a native acquirer of TİD.
-In that, there were fewer nominal referents when it was a native acquirer of TİD and the discourse context was maintenance.
+There seems to be an interaction between the introduction status of discourse and being a native acquirer of TİD.
+When the acquirer is a native speaker and the discourse function is introduction, the chances of observing a nominal referent gets even higher.
+Also, there seems to be an interaction between maintaining a referent and being a native acquirer of TİD.
+When it was a native acquirer of TİD and the discourse context was maintenance, the chances of observing a nominal referent gets even lower.
 
@@
 
-For age of acquisition effects, based on what is proposed for child and adult deaf signers with language deprivation in terms of their morphosyntactic [@boudreault2006; @cormier2012; @kayaba2022] and narrative development [@becker2009; @cormier2013], we expected that late signers would be more over-redundant in their referent tracking, akin to second language learners [@frederiksen2019; @bel2015].
+For age of acquisition effects, based on what is proposed for child and adult deaf signers with language deprivation in terms of their morphosyntactic [@boudreault2006; @cormier2012; @kayabasi2020] and narrative development [@becker2009; @cormier2013], we expected that late signers would be more over-redundant in their referent tracking, akin to second language learners [@frederiksen2019; @bel2015].
 That is, we had the expectation that late signers would use more explicit markers like nouns and fewer implicit markers like null marking even for highly accessible referents (when the discourse status is maintained or even re-introduced in same cases) than the control group (i.e., native signers).
@@
 As for the distribution of REs across different discourses statuses, we observed that signers overwhelmingly preferred to use nominals over other REs for the first mention of a referent and used null markers such as CA, and agreement and plain verb for referents with high accessibility (i.e., when the discourse status was maintained).
-This finding parallels other accounts of referential accessibility suggested for other sign languages [see @frederiksen2016; @czubek2017; and @swabey2002; for ASL, @morgan2005; for BSL @bel2015; and @barberà2009 for LSC.
+This finding parallels other accounts of referential accessibility suggested for other sign languages [see @frederiksen2016; @czubek2017; and @swabey2002; for ASL, @morgan2005; for BSL @bel2015; and @barbera2009 for LSC.
 Our results also indicate that modified nouns (e.g., nouns that are preceded or followed by a classifier or IX) appeared more in introduced contexts compared to re-introduced contexts.
@@
 This finding is comparable with what has been reported for ASL [@swabey2002; @frederiksen2016; @czubek2017] but not with LSC [@bel2015], BSL [@morgan2005] and DGS [@perniss2015], in which the signers extensively made overt pronominal references (up to 30%) to maintain entities.
-Comparable to our findings, It has been reported that pronouns do not frequently occur in the local context and instead follows a topic shift and thus rank higher than classifiers and null markers different genre [see @barberà2009 for LSC; and @nuhbalaoglu2018 for TİD].
+Comparable to our findings, It has been reported that pronouns do not frequently occur in the local context and instead follows a topic shift and thus rank higher than classifiers and null markers different genre [see @barbera2009 for LSC; and @nuhbalaoglu2018 for TİD].
 Considering this, the claim that the IX sign functions like demonstratives that are also used to track referents in re-introduced contexts has not been borne out [@koulidobrova2016].
@@
 This contrasts with some of the research [@czubek2017] in which agreement verbs ranked lowest on scale, indicating that these constructions were more implicit compared to plain verbs.
-However, there seems to be a common ground for the position of CA markers which have been claimed to rank higher than other null markers as in the present study [@frederiksen2016; @czubek2017; @barberà2009].
+However, there seems to be a common ground for the position of CA markers which have been claimed to rank higher than other null markers as in the present study [@frederiksen2016; @czubek2017; @barbera2009].
 In Figure \ref{fig:scale}, we report the distribution of REs in TİD as a function of referent accessibility and compare our accessibility hierarchy with that of [@frederiksen2016] for ASL.
@@
 In line with our study, a tendency for limited over-redundancy has been suggested for late signing children in DGS [@becker2009] and late signing adults in BSL specifically for the use of CA[@cormier2013].
-As for TİD, to our knowledge, there is only one study [@gür2018] that directly aimed at examining age of acquisition effects on adult signers' narrative skills.
-@gür2018 observed that late signing deaf adults and children set the scene less frequently and used CA less accurately compared to their native counterparts.
+As for TİD, to our knowledge, there is only one study [@gur2018] that directly aimed at examining age of acquisition effects on adult signers' narrative skills.
+@gur2018 observed that late signing deaf adults and children set the scene less frequently and used CA less accurately compared to their native counterparts.
 However, this particular investigation did not present any data on the distribution of REs between two acquisition groups.
@@
 contrasts(df_model_camera$Discourse) <- contr.sum(3)/2
-contrasts(df_model_camera$Group) <- contr.sdif(3)
 
@@
 # Posterior plot for reference type
-df_model_camera <- mcmc_intervals_data(model_1, pars = vars(starts_with("b_")))
+df_model_camera <- mcmc_intervals_data(model_camera, pars = vars(starts_with("b_")))
 df_model_camera$parameter %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
@@
 
-```{r fig:model-camera, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
-p4 <- df_model_camera %>% 
-  ggplot(aes(m, y=factor(parameter,levels = rev(levels(factor(parameter)))))) +
-  geom_point(size = 2.5) +
-  geom_errorbarh(aes(xmin = l, xmax = h), alpha = 1, height = 0, size = 1.5) +
-  geom_errorbarh(aes(xmin = ll, xmax = hh, height = 0)) + vline_0(linetype = "dashed") +
-  xlab("Estimate (log odds)") + ylab("Coefficients") + theme_bw() + theme_apa()
+```{r model-camera, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
+df_model_1 %<>% mutate(model = "All")
+df_model_camera %<>% mutate(model = "Camera")
+model_comparison <- rbind(df_model_1, df_model_camera)
+
+p4 <- model_comparison %>% 
+  ggplot(aes(m, y=factor(parameter, levels = rev(levels(factor(parameter)))), group = model, color = model)) +
+  geom_point(size = 2.5, position = position_dodge(width = 0.6)) +
+  geom_errorbarh(aes(xmin = l, xmax = h), alpha = 1, height = 0, size = 1.5,position = position_dodge(width = 0.6)) +
+  geom_errorbarh(aes(xmin = ll, xmax = hh, height = 0), position = position_dodge(width = 0.6)) + vline_0(linetype = "dashed") +
+  xlab("Estimate (log odds)") + ylab("Coefficients") + theme_apa()
 p4
@@
 library(janitor)
-df <- summary(model_3) %>% .$fixed
-df <- as.data.frame(df)
+df <- summary(model_3) %>% .$fixed %>% as_tibble()
 df_table_extra <- df %>% clean_names() %>% rownames_to_column("Coefficients") %>% dplyr::select(-7,-8)
 
-names(df_table_extra) <- c("Coefficients","Estimate", "EstError","lower95","upper95","Rhat")
+names(df_table_extra) <- c("Coefficients", "Estimate", "EstError", "lower95", "upper95", "Rhat")
+
 df_table_extra$Coefficients %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",

### 2022-08-05 manuscript_v2.tex
updated

@@
-\documentclass[]{elsarticle} %review=doublespace preprint=single 5p=2 column
+\documentclass[review]{elsarticle} %review=doublespace preprint=single 5p=2 column
 %%% Begin My package additions %%%%%%%%%%%%%%%%%%%
+
 \usepackage[hyphens]{url}
@@
 \usepackage{lineno} % add
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 \IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\bibliographystyle{elsarticle-harv}
+\usepackage[]{natbib}
+\bibliographystyle{plainnat}
+
 \usepackage{graphicx}
@@
             pdfborder={0 0 0}}
-\urlstyle{same}  % don't use monospace font for urls
 
@@
 
-% Pandoc citation processing
-\newlength{\cslhangindent}
-\setlength{\cslhangindent}{1.5em}
-\newlength{\csllabelwidth}
-\setlength{\csllabelwidth}{3em}
-% for Pandoc 2.8 to 2.10.1
-\newenvironment{cslreferences}%
-  {}%
-  {\par}
-% For Pandoc 2.11+
-\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
- {% don't indent paragraphs
-  \setlength{\parindent}{0pt}
-  % turn on hanging indent if param 1 is 1
-  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
-  % set entry spacing
-  \ifnum #2 > 0
-  \setlength{\parskip}{#2\baselineskip}
-  \fi
- }%
- {}
-\usepackage{calc}
-\newcommand{\CSLBlock}[1]{#1\hfill\break}
-\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
-\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
-\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
-
-% Pandoc header
+% tightlist command for lists without linebreak
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+
 \usepackage{floatrow}
@@
 \begin{document}
+
 \begin{frontmatter}
@@
 Turkish Sign Language}
+      \cortext[cor1]{Corresponding author}
 
@@
   \end{abstract}
-   \begin{keyword} Referential accessibility, signed narratives,
-pragmatic competence, delayed language acquisition, Turkish Sign
-Language\end{keyword}
+    \begin{keyword}
+    Referential accessibility, signed narratives, pragmatic competence,
+delayed language acquisition, Turkish Sign Language \sep 
+    Referential accessibility, signed narratives, pragmatic competence,
+delayed language acquisition, Turkish Sign Language
+  \end{keyword}
+
  \end{frontmatter}
@@
 organization of experiences, narrators must choose appropriate
-linguistic forms to refer to entities (Ariel, 1990; Givon, 1983; Gundel
-et al., 1993): Certain forms such as nominal expressions are commonly
-used for the first mention of a referent in a discourse, and other forms
-such as pronouns and null marking may track previously mentioned
-entities which could either be maintained across multiple clauses or
-re-introduced back into the discourse. The selection of these referring
-expressions (henceforth REs) is contingent on cognitive abilities, such
-that individuals must rely on the activation of the working memory to
-keep in mind previously referred expressions and long strings of
-information (Bamberg, 1997; Morgan, 2005). Pragmatic abilities too play
-a pivotal role in the tracking of referents. The form with the least
-amount of redundant information must be selected based on the
-accessibility levels of the referent (i.e., how active a referent is in
-the addressee's mind given a point in discourse) (Ahn, 2019; Ariel,
-1990). Because of these cognitive and pragmatic necessities, it is not
-surprising to observe that producing a coherent narrative is one of the
-latest milestones in language acquisition among children (Hickmann et
-al., 1996). As can be expected, there is a plethora of research
-examining how narrative competence is achieved in both monolingual and
-bilingual acquisition contexts among children and adults. Although
-narrative competence and referent selection have been studied
-extensively for typically developing speakers (Aksu-Koç and
-Nicolopoulou, 2015; Hickmann et al., 1996; Williams, 1988), deaf
-individuals who acquired a sign language in atypical learning contexts
-(i.e., deaf signers with linguistic deprivation) have remained mostly
-understudied with an exception of handful of studies (Becker, 2009;
-Cormier et al., 2013; Gür, 2018). Nevertheless, it is paramount that we
-investigate and describe the effects of language deprivation among deaf
-adults on narrative competence more in depth. This will help us
-understand whether the loss of systematic and frequent language input in
-the early years of life affect how deaf signers produce a narrative in
-their adulthood. Given this, the present study will focus on deaf adult
-signers with language deprivation and examine their reference tracking
-strategies in signed narratives.
+linguistic forms to refer to entities
+\citep{ariel1990, givon1983, gundel1993}: Certain forms such as nominal
+expressions are commonly used for the first mention of a referent in a
+discourse, and other forms such as pronouns and null marking may track
+previously mentioned entities which could either be maintained across
+multiple clauses or re-introduced back into the discourse. The selection
+of these referring expressions (henceforth REs) is contingent on
+cognitive abilities, such that individuals must rely on the activation
+of the working memory to keep in mind previously referred expressions
+and long strings of information \citep{morgan2005, bamberg1997}.
+Pragmatic abilities too play a pivotal role in the tracking of
+referents. The form with the least amount of redundant information must
+be selected based on the accessibility levels of the referent (i.e., how
+active a referent is in the addressee's mind given a point in discourse)
+\citep{ahn2019, ariel1990}. Because of these cognitive and pragmatic
+necessities, it is not surprising to observe that producing a coherent
+narrative is one of the latest milestones in language acquisition among
+children \citep{hickmann1996}. As can be expected, there is a plethora
+of research examining how narrative competence is achieved in both
+monolingual and bilingual acquisition contexts among children and
+adults. Although narrative competence and referent selection have been
+studied extensively for typically developing speakers
+\citep{aksu-koc2015, hickmann1996, williams1988}, deaf individuals who
+acquired a sign language in atypical learning contexts (i.e., deaf
+signers with linguistic deprivation) have remained mostly understudied
+with an exception of handful of studies
+\citep{cormier2013, gur2018, becker2009}. Nevertheless, it is paramount
+that we investigate and describe the effects of language deprivation
+among deaf adults on narrative competence more in depth. This will help
+us understand whether the loss of systematic and frequent language input
+in the early years of life affect how deaf signers produce a narrative
+in their adulthood. Given this, the present study will focus on deaf
+adult signers with language deprivation and examine their reference
+tracking strategies in signed narratives.
 
@@
 been the officially recognized language of the Deaf community in Turkey
-since 2005 (see Nuhbalaoglu, 2018 for a brief review of the historical
-development of the language). Deaf or hearing children acquiring a SL
-within a family where there is one or more deaf caregivers display the
-same developmental milestones as hearing children acquiring a spoken
-language in a hearing family (Chamberlain et al., 1999; Meier and
-Newport, 1990). However, reports indicate that over 90\% of all deaf
-individuals are born into hearing families who do not know how to sign
-(i.e., those with language deprivation) (Woll, 2013). This means that
-the auditory modality used in the household is not accessible to many
-deaf children. That is why the age of acquisition of a first functional
-language varies greatly among deaf children, with many of whom being
-deprived of their first languages typically until after age 7 when they
-start formal education in a deaf school. First language deprivation or
-delay has been found to have negative effects on certain cognitive and
-linguistic measures among late signing children. To name a few, delayed
-exposure to a first functional language is reported to negatively affect
-deaf children's morphosyntactic production and processing (Boudreault
-and Mayberry, 2006; Cheng and Mayberry, 2021; Cormier et al., 2012;
-Karadöller et al., 2017; Kayabaşı and Gökgöz, 2022; Newport, 1990) and
-narrative abilities (e.g., Becker, 2009). In our overall research, we
-are concerned with the effects of this late exposure in the ultimate
-adult forms of TİD users.
+since 2005 \citep[see][ for a brief review of the historical development
+of the language]{nuhbalaoglu2018}. Deaf or hearing children acquiring a
+SL within a family where there is one or more deaf caregivers display
+the same developmental milestones as hearing children acquiring a spoken
+language in a hearing family
+\citep{chamberlain_language_1999, meier1990}. However, reports indicate
+that over 90\% of all deaf individuals are born into hearing families
+who do not know how to sign (i.e., those with language deprivation)
+\citep{woll2013}. This means that the auditory modality used in the
+household is not accessible to many deaf children. That is why the age
+of acquisition of a first functional language varies greatly among deaf
+children, with many of whom being deprived of their first languages
+typically until after age 7 when they start formal education in a deaf
+school. First language deprivation or delay has been found to have
+negative effects on certain cognitive and linguistic measures among late
+signing children. To name a few, delayed exposure to a first functional
+language is reported to negatively affect deaf children's
+morphosyntactic production and processing
+\citep{kayabasi2020, boudreault2006, newport1990, cheng2021, karadoller2017, cormier2012}
+and narrative abilities \citep[e.g.,][]{becker2009}. In our overall
+research, we are concerned with the effects of this late exposure in the
+ultimate adult forms of TİD users.
 
@@
 accordance with the addressee's mental representations of the referents
-in the discourse is the Givenness Hierarchy (Gundel et al., 1993).
+in the discourse is the Givenness Hierarchy \citep{gundel1993}.
 According to this model, given or familiar information is expressed in
@@
 mental status of the expressed referent is already at the center of
-attention, as indicated by the present contextual information (Chafe,
-1976). The Givenness Hierarchy is akin to other previously proposed
-models of reference control within the topicality and saliency paradigm
-such as the Accessibility Hierarchy (Ariel, 1990), Topic Continuity
-(Givon, 1983), and Familiarity Scale (Prince, 1981). Gundel et al.
-(1993) argued that the distribution of forms is not random but abides by
-the Gricean Maxim of Quantity (Grice, 1975). That is to say, individuals
-must make the most informative and succinct choice possible to retain
-economy of form. When multiple expressions are available for the same
-referent as in an utterance like (\ref{ex:1}), phonetically simplest
-form is selected with the least redundant information are selected based
-on an economy principle. In the utterance, REs for ``Eddie'' are given
-in bold. All three REs (nominal, pronominal, and null marking) refer to
-the same entity; however, the quantity of marking decreases as the
-referent's accessibility increases. On this ground, for highly
-accessible referents (i.e., those that are highly active in the mind),
-linguistic forms with the least phonetic content are selected for the
-least amount of processing effort.
+attention, as indicated by the present contextual information
+\citep{chafe1976}. The Givenness Hierarchy is akin to other previously
+proposed models of reference control within the topicality and saliency
+paradigm such as the Accessibility Hierarchy \citep{ariel1990}, Topic
+Continuity \citep{givon1983}, and Familiarity Scale \citep{prince1981}.
+\citet{gundel1993} argued that the distribution of forms is not random
+but abides by the Gricean Maxim of Quantity \citep{grice1975}. That is
+to say, individuals must make the most informative and succinct choice
+possible to retain economy of form. When multiple expressions are
+available for the same referent as in an utterance like (\ref{ex:1}),
+phonetically simplest form is selected with the least redundant
+information are selected based on an economy principle. In the
+utterance, REs for ``Eddie'' are given in bold. All three REs (nominal,
+pronominal, and null marking) refer to the same entity; however, the
+quantity of marking decreases as the referent's accessibility increases.
+On this ground, for highly accessible referents (i.e., those that are
+highly active in the mind), linguistic forms with the least phonetic
+content are selected for the least amount of processing effort.
 
@@
 It is also possible to categorize REs into three according to their
-discourse status (Gullberg, 2006). A referent can be introduced or
+discourse status \citep{gullberg2006}. A referent can be introduced or
 mentioned for the first time, maintained, that is, continued across two
@@
 based on the principles of reference accessibility. For spoken
-languages, this observation is well established (Ahrenholz, 2005; Ariel,
-1990; Azar and Özyürek, 2015; Chafe, 1976; Debreslioska et al., 2013;
-Givon, 1983; Gundel et al., 1993; Hickmann et al., 1996). These research
-indicate that the introduction and re-introduction of a referent are
-often accomplished by full noun phrases which can be commonly modified
-by an adjective or demonstrative. Maintenance, on the other hand, is
-typically marked with pronominals and null markers since the continued
-referent is still highly accessible to the addressee.
+languages, this observation is well established
+\citep{ariel1990, givon1983, gundel1993, debreslioska2013, ahrenholz2005, chafe1976, hickmann1996, azar2015}.
+These research indicate that the introduction and re-introduction of a
+referent are often accomplished by full noun phrases which can be
+commonly modified by an adjective or demonstrative. Maintenance, on the
+other hand, is typically marked with pronominals and null markers since
+the continued referent is still highly accessible to the addressee.
 
@@
 languages. There is an emerging body of literature that reports the
-distribution of REs in signed narratives. Building on Gundel et al.
-(1993)'s work, Swabey (2002) demonstrated that, like other spoken
-languages such as Russian, Chinese, and Japanese, American Sign Language
-(ASL) signers use bare nouns for introduced and re-introduced discourse
-contexts whereas null marking is used for referents which are
-``in-focus'' or very accessible to the addressee. Overt REs like
-pronouns or nouns are disprefered for such referents. Swabey (2002)
-concluded that ``native'' signers in her study abided by the Gricean
-maxim of quantity and were thus economic in their reference tracking.
-Similar findings have been reported for German Sign Language (Deutsche
-Gebärdensprache - DGS) (Perniss and Özyürek, 2015) and British Sign
-Language (BSL) (Morgan, 2005). It is important to note, however, that
-these research either collapsed native and late signers into a single
-group and labeled them as ``native-like'' or explicitly compared native
-deaf signers with late deaf signers (with language deprivation) in terms
-of narrative competence.
+distribution of REs in signed narratives. Building on
+\citet{gundel1993}'s work, \citet{swabey2002} demonstrated that, like
+other spoken languages such as Russian, Chinese, and Japanese, American
+Sign Language (ASL) signers use bare nouns for introduced and
+re-introduced discourse contexts whereas null marking is used for
+referents which are ``in-focus'' or very accessible to the addressee.
+Overt REs like pronouns or nouns are disprefered for such referents.
+\citet{swabey2002} concluded that ``native'' signers in her study abided
+by the Gricean maxim of quantity and were thus economic in their
+reference tracking. Similar findings have been reported for German Sign
+Language (Deutsche Gebärdensprache - DGS) \citep{perniss2015} and
+British Sign Language (BSL) \citep{morgan2005}. It is important to note,
+however, that these research either collapsed native and late signers
+into a single group and labeled them as ``native-like'' or explicitly
+compared native deaf signers with late deaf signers (with language
+deprivation) in terms of narrative competence.
 
@@
 demonstratives where sign languages use a pointing sign, usually glossed
-as IX, which assumes similar functions (Lillo-Martin and Klima, 1990).
-In addition, the pointing sign is used to referentially anchor a
-referent to a location in signing space. For instance, when the signer
-introduces \textsc{elvan} for the first time, s/he can assign her to the
-right. This referential locus can later be used by agreement verbs and
+as IX, which assumes similar functions \citep{lillo-martin1990}. In
+addition, the pointing sign is used to referentially anchor a referent
+to a location in signing space. For instance, when the signer introduces
+\textsc{elvan} for the first time, s/he can assign her to the right.
+This referential locus can later be used by agreement verbs and
 classifier predicates (more below). For instance, in (\ref{ex:2a}) there
@@
 the verb and often cross-reference a core argument in a clause
-(Aikhenvald, 2017). Classifiers denote sets of objects based on semantic
-criteria such as size, shape, body-part, texture and sometimes function.
-Usually, the verbal classifier and the cross-referenced noun co-occur in
-the same clause when the argument is introduced into discourse. The
-classifier can track the argument on its own on subsequent occurrences.
-In sign languages, a verbal classifier morpheme is encoded by a
-\emph{handshape} that is articulated phonologically co-temporally with a
-movement, which encodes the verbal meaning (Benedicto and Brentari,
-2004; Supalla, 1990). These two infused morphemes form a complex
-predicate. Some classifier handshapes and the set of objects that they
-denote are below. Some classifier handshapes and the set of objects that
-they denote are presented below in (\ref{ex:4}).
+\citep{aikhenvald2017}. Classifiers denote sets of objects based on
+semantic criteria such as size, shape, body-part, texture and sometimes
+function. Usually, the verbal classifier and the cross-referenced noun
+co-occur in the same clause when the argument is introduced into
+discourse. The classifier can track the argument on its own on
+subsequent occurrences. In sign languages, a verbal classifier morpheme
+is encoded by a \emph{handshape} that is articulated phonologically
+co-temporally with a movement, which encodes the verbal meaning
+\citep{supalla1990, benedicto2004}. These two infused morphemes form a
+complex predicate. Some classifier handshapes and the set of objects
+that they denote are below. Some classifier handshapes and the set of
+objects that they denote are presented below in (\ref{ex:4}).
 
@@
 handling classifiers, and extension classifiers), and overt pronouns
-(pronominal index or IX) (Frederiksen and Mayberry, 2016). This interim
-section describes such constructions attested for sign languages under
-four main headings (nominals, pronouns, classifiers, and null marking).
+(pronominal index or IX) \citep{frederiksen2016}. This interim section
+describes such constructions attested for sign languages under four main
+headings (nominals, pronouns, classifiers, and null marking).
 
@@
 (\ref{ex:6c}). Fingerspelling refers to the use of the manual alphabet
-which is derived or borrowed from a surrounding written language (Kubuş,
-2008). Nouns can also be modified by classifiers as in (\ref{ex:6b})
-(Frederiksen and Mayberry, 2016) or by a prenominal or postnominal IX
-functioning as a determiner or a demonstrative as in (\ref{ex:6d})
-(Maclaughlin, 1997).
+which is derived or borrowed from a surrounding written language
+\citep{kubus2008}. Nouns can also be modified by classifiers as in
+(\ref{ex:6b}) \citep{frederiksen2016} or by a prenominal or postnominal
+IX functioning as a determiner or a demonstrative as in (\ref{ex:6d})
+\citep{maclaughlin1997}.
 
@@
 physical locus in the signing space when re-introducing a topic in
-discourse (Emmorey, 1996). The same IX function has also been identified
-for TİD (Sevinç, 2006). The example in (\ref{ex:7}) illustrates a
-spatially-anchored pronominal reference for topic re-introduction. These
-signs are thus rarely used for the first mention of an entity (Perniss
-and Özyürek, 2015). Other interpretations consider pointing signs
-without noun phrases as examples of demonstratives and not personal
-pronouns (Koulidobrova and Lillo-Martin, 2016). It is important to
+discourse \citep{emmorey1996}. The same IX function has also been
+identified for TİD \citep{sevinc2006}. The example in (\ref{ex:7})
+illustrates a spatially-anchored pronominal reference for topic
+re-introduction. These signs are thus rarely used for the first mention
+of an entity \citep{perniss2015}. Other interpretations consider
+pointing signs without noun phrases as examples of demonstratives and
+not personal pronouns \citep{koulidobrova2016}. It is important to
 mention here that in some spoken languages (e.g., German and Dutch),
-demonstratives can also refer to persons (Bosch and Umbach, 2007) and
-behave similarly to personal pronouns. However, for the purposes of the
-present study, we will refer to IX constructions as pronominal IX and
+demonstratives can also refer to persons \citep{bosch2007} and behave
+similarly to personal pronouns. However, for the purposes of the present
+study, we will refer to IX constructions as pronominal IX and
 particularly examine their non-deictic or personal pronoun realizations
@@
 Classifier predicates are highly complex morphological constructions
-which are very common in sign languages (Supalla, 1986). These are
+which are very common in sign languages \citep{supalla1986}. These are
 manually signed iconic representations of animate or inanimate objects,
@@
 including its location in the signing space, movement, path, size, and
-shape (Talmy, 2003). As in most sign languages hitherto studied,
-classifier constructions are also common and found in TİD (see Dikyuva
-et al., 2015 for a review). Following the classification of
-Engberg-Pedersen (1993), we distinguish between four types of classifier
-constructions. Whole entity classifiers (WECL) represent the entirety of
-any referent whereas body part classifiers (BPCL) only reflect a
-specific part of an animate entity (e.g., limbs, head, tongue, teeth,
-and so forth). Handling classifiers (HCL) occur with transitive verbs
-and depict how an object is handled or manipulated by agent. On this
-ground, an HCL\footnote{As discussed before, an HCL is transitive in
-  nature and depicts the handling action performed by ananimate agent.
-  This is why most research tend to consider HCLs as an extension to
-  constructed action (Cormier et al., 2013; Dudis, 2004).} depicts the
+shape \citep{talmy2003}. As in most sign languages hitherto studied,
+classifier constructions are also common and found in TİD \citep[see][
+for a review]{dikyuva2015}. Following the classification of
+\citet{engberg-pedersen1993}, we distinguish between four types of
+classifier constructions. Whole entity classifiers (WECL) represent the
+entirety of any referent whereas body part classifiers (BPCL) only
+reflect a specific part of an animate entity (e.g., limbs, head, tongue,
+teeth, and so forth). Handling classifiers (HCL) occur with transitive
+verbs and depict how an object is handled or manipulated by agent. On
+this ground, an HCL\footnote{As discussed before, an HCL is transitive
+  in nature and depicts the handling action performed by ananimate
+  agent. This is why most research tend to consider HCLs as an extension
+  to constructed action \citep{dudis2004, cormier2013}.} depicts the
 manner of the referring action while the former two constructions encode
@@
 This is often achieved with the signer taking on the role of a referent
-in the narrative and performing the actions (Metzger, 1995). CA is often
-marked by certain non-manual cues such as the facial expression and the
-breaking of the eye contact. This is to say, the narrator's head, face,
-torso, arms, and as an optional marker, hands represent those of the
-associated referent(s) (Smith and Cormier, 2014). As for the verbal
-constructions, to identify the predicate types in ASL, Padden (1986)
-suggested a triadic categorization of verbs in sign languages: plain
-verbs, agreement verbs, and spatial verbs. According to this
-classification, plain verbs (\ref{ex:9b}) do not agree with the subject
-or the object. This means that they are not inflected for number or
-person. This contrasts with agreeing verbs (\ref{ex:9c}) the starting or
-the ending point of which can be modified to specific loci for person
+in the narrative and performing the actions \citep{metzger1995}. CA is
+often marked by certain non-manual cues such as the facial expression
+and the breaking of the eye contact. This is to say, the narrator's
+head, face, torso, arms, and as an optional marker, hands represent
+those of the associated referent(s) \citep{smith2014}. As for the verbal
+constructions, to identify the predicate types in ASL,
+\citet{padden1986} suggested a triadic categorization of verbs in sign
+languages: plain verbs, agreement verbs, and spatial verbs. According to
+this classification, plain verbs (\ref{ex:9b}) do not agree with the
+subject or the object. This means that they are not inflected for number
+or person. This contrasts with agreeing verbs (\ref{ex:9c}) the starting
+or the ending point of which can be modified to specific loci for person
 and number agreement while spatial verbs (as in \textsc{ix book put} `I
@@
 verbs, they are not inflected for number or person. TİD too licenses
-null arguments. Following this classification, Sevinç (2006) argued that
-although it is common in TİD to drop arguments, there were certain
+null arguments. Following this classification, \citet{sevinc2006} argued
+that although it is common in TİD to drop arguments, there were certain
 limitations on the null marking of the arguments for agreement and plain
@@
 informative linguistic forms to indicate anaphoric or cataphoric
-references after age 10 (Hickmann et al., 1996; Rathmann et al., 2007).
-Prior to this age, both signing and speaking children fail to refer to
-entities consistently since their pragmatic abilities are yet to
-develop. The manipulation of linguistic forms for discourse functions to
-signal topic continuance or shift is difficult to master until the
-beginning of formal education. These findings are also comparable across
-different sign languages including BSL (Morgan, 2005), ASL (Loew, 1984),
-and DGS (Becker, 2009), pointing out to the fact that the mastery of
+references after age 10 \citep{hickmann1996, rathmann2007}. Prior to
+this age, both signing and speaking children fail to refer to entities
+consistently since their pragmatic abilities are yet to develop. The
+manipulation of linguistic forms for discourse functions to signal topic
+continuance or shift is difficult to master until the beginning of
+formal education. These findings are also comparable across different
+sign languages including BSL \citep{morgan2005}, ASL \citep{loew1984},
+and DGS \citep{becker2009}, pointing out to the fact that the mastery of
 reference tracking in narratives is one of the most protracted
@@
 Like in first language development, narrative skills have been
-extensively studied among second language learners (see Bel et al.,
-2015; Frederiksen and Mayberry, 2019). Comparing 11 deaf adults signers
-and 13 hearing advanced learners of Catalan Sign Language (Llengua de
-Signes Catalana - LSC), Bel et al. (2015) found that hearing signers
-resort to more overt forms (nominals and pronouns) than native signers
-to avoid potential ambiguities among referents. Nevertheless, their
-results did not reach statistical significant for the interaction of
-discourse status and acquisition group. A similar line of research came
-from Frederiksen and Mayberry (2019) who found a statistically
-significant difference between both groups, that is native signers and
-hearing L2 learners of ASL, only for classifier use especially in
-re-introduced contexts. L2 learners in the study used more nominals and
-fewer instances of null marking in number, indicative of a limited
-effect of overexplicitness among second language learners. However,
-these studies treated late and native deaf signers as a single group.
-That is, the control group consisted of both native signers who had one
-or more deaf parents and also native-like (Cormier et al., 2012) late
-deaf signers who had hearing parents and their age of acquisition
-varied. Given the preliminary evidence that topic continuity and
-salience marking is also difficult for second language learners
-regardless of modality, it has been suggested under the Interface
-Hypothesis that the syntax-discourse interface is vulnerable and
-especially challenging to acquire (Sorace, 2011) for second language
-learners. Similarly, Sorace and Serratrice (2009) discuss that this
-interface entails much more than morphosyntactic knowledge but also
-includes cognitive and pragmatic resources that help provide discourse
-coherence. They propose that the second language discourse is
-over-redundant. This means that second language learners produce more
-overt pronouns or fuller forms of language for highly accessible
-referents possibly as a compensatory strategy in an attempt to reduce
-the increasing load on cognitive processing.
+extensively studied among second language learners
+\citep[see][]{bel2015, frederiksen2019}. Comparing 11 deaf adults
+signers and 13 hearing advanced learners of Catalan Sign Language
+(Llengua de Signes Catalana - LSC), \citet{bel2015} found that hearing
+signers resort to more overt forms (nominals and pronouns) than native
+signers to avoid potential ambiguities among referents. Nevertheless,
+their results did not reach statistical significant for the interaction
+of discourse status and acquisition group. A similar line of research
+came from \citet{frederiksen2019} who found a statistically significant
+difference between both groups, that is native signers and hearing L2
+learners of ASL, only for classifier use especially in re-introduced
+contexts. L2 learners in the study used more nominals and fewer
+instances of null marking in number, indicative of a limited effect of
+overexplicitness among second language learners. However, these studies
+treated late and native deaf signers as a single group. That is, the
+control group consisted of both native signers who had one or more deaf
+parents and also native-like \citep{cormier2012} late deaf signers who
+had hearing parents and their age of acquisition varied. Given the
+preliminary evidence that topic continuity and salience marking is also
+difficult for second language learners regardless of modality, it has
+been suggested under the Interface Hypothesis that the syntax-discourse
+interface is vulnerable and especially challenging to acquire
+\citep{sorace2011} for second language learners. Similarly,
+\citet{sorace2009} discuss that this interface entails much more than
+morphosyntactic knowledge but also includes cognitive and pragmatic
+resources that help provide discourse coherence. They propose that the
+second language discourse is over-redundant. This means that second
+language learners produce more overt pronouns or fuller forms of
+language for highly accessible referents possibly as a compensatory
+strategy in an attempt to reduce the increasing load on cognitive
+processing.
 
@@
 signers (i.e., deaf signers with delayed exposure to a sign language)
-also display redundancy in their reference tracking. Becker (2009)
+also display redundancy in their reference tracking. \citet{becker2009}
 examined 6 DGS signing deaf children (3 native and 3 late)`s narrative
@@
 deprivation results in an inadequate representation of the addressee and
-less efficient use of morphosyntax. Comparably, Cormier et al. (2013)
+less efficient use of morphosyntax. Comparably, \citet{cormier2013}
 investigated the use of CA in 15 BSL among deaf native and late adult
@@
 nominals, and fewer instances of null marking when maintaining previous
-referents. Most relevantly to the present study, Gür (2018) studied the
-development of narrative skills among 10 native and 10 late deaf
+referents. Most relevantly to the present study, \citet{gur2018} studied
+the development of narrative skills among 10 native and 10 late deaf
 children and deaf adult signers of TİD. She found that native-signing
@@
 Re-introduction) and RE type (Nominal, Pronominal, Classifier, Null). We
-used ELAN Linguistic Annotation Software (Crasborn and Sloetjes, 2008)
-for data annotation. First Author who is a hearing signer of TİD
-annotated all the data and translated the narratives into Turkish
-together with a native deaf signer of TİD.
-
-Following Berman and Slobin (2013), we first identified the clause
+used ELAN Linguistic Annotation Software \citep{crasborn2008} for data
+annotation. First Author who is a hearing signer of TİD annotated all
+the data and translated the narratives into Turkish together with a
+native deaf signer of TİD.
+
+Following \citet{berman2013relating}, we first identified the clause
 boundaries in each sentence indicated by the presence of predicates and
@@
 coded the discourse status for each referent. Having adapted the
-classification of Gullberg (2006), we considered all referents mentioned
-for the first time as introduced regardless of their syntactic function
-in the sentence. Only subjects were coded for maintenance and
+classification of \citet{gullberg2006}, we considered all referents
+mentioned for the first time as introduced regardless of their syntactic
+function in the sentence. Only subjects were coded for maintenance and
 re-introduction. We coded subject referents as maintained only if they
@@
 In our analysis, we use Bayesian regression models with the help of the
-brms package (Bürkner, 2018) in R. In a Bayesian regression the prior
-and the likelihood is used to calculate the posterior. Instead of a
-single filter of significance, we get a posterior probability
+brms package \citep{burkner2018} in R. In a Bayesian regression the
+prior and the likelihood is used to calculate the posterior. Instead of
+a single filter of significance, we get a posterior probability
 distribution. We do not define our own priors and use the brms package
 defaults. We are using Bayesian regression to avoid some common pitfalls
-of other approaches (Kruschke, 2011; Wagenmakers, 2007).
+of other approaches \citep{wagenmakers2007, kruschke2011}.
 
@@
 \centering
-\includegraphics{manuscript_v2_files/figure-latex/fig:count-plot-1.pdf}
+\includegraphics{manuscript_v2_files/figure-latex/count-plot-1.pdf}
 \caption{Referent counts by Discourse Status (Introduction, Maintanence,
@@
 For more inference on the data and having pairwise comparisons, we fit a
-regression model to RE Type using Discourse Status and Group as
-predictors. We used the brms package (Bürkner, 2018) in R, with a
+regression model to RE Type using discourse status and acquisition group
+as predictors. We used the brms package \citep{burkner2018} in R, with a
 categorical distribution on the responses and sum contrasts for the
@@
 from the following analyses. For the current analysis, we did not define
-priors ourselves but used the preset uninformed priors. Figure
-2\footnote{We ran another model where we used the Nativeness predictor
-  with 3 levels 0-3:Native, 4-7:Late1, and 7-above:Late2. However, our
-  model did not converg properly. We provide the model results as a plot
-  and table in the \ref{section:d}.} shows the posterior probability
-distributions where the dot is the median estimate, the thick line is
-the 50\% credible intervals, and the thin line is the 95\% credible
-intervals. There are 3 responses we are considering in this model, and
-they are Classifier, Nominative, and Null. Classifier response is the
-level that other responses Nominative and Null are compared against.
+priors ourselves and used the preset uninformed priors.
+@ref\{fig:model-gram\}\footnote{We ran another model where we used the
+  Nativeness predictor with 3 levels 0-3:Native, 4-7:Late1, and
+  7-above:Late2. However, our model did not converge properly. We
+  provide the model results as a plot and table in the \ref{section:d}.}
+shows the posterior probability distributions where the dot is the
+median estimate, the thick line is the 50\% credible intervals, and the
+thin line is the 95\% credible intervals. There are 3 responses we are
+considering in this model. They are; Classifier, Nominative, and Null.
+Classifier response is the level that other responses Nominative and
+Null are compared against.
+
+\begin{verbatim}
+## Warning: Ignoring unknown aesthetics: height
+\end{verbatim}
 
@@
 \centering
-\includegraphics{manuscript_v2_files/figure-latex/fig:model-gram-1.pdf}
+\includegraphics{manuscript_v2_files/figure-latex/model-gram-1.pdf}
 \caption{Regression model plot for the RE Type (Nominal, Null) with the
@@
 compared to the re-introduction contexts within the greater comparison
-to classifier referents. There does not seem to be an interaction
-between the introduction status of discourse and being a native acquirer
-of TİD. However, there seems to be an interaction between maintaining a
-referent and being a native acquirer of TİD. In that, there were fewer
-nominal referents when it was a native acquirer of TİD and the discourse
-context was maintenance.
+to classifier referents. There seems to be an interaction between the
+introduction status of discourse and being a native acquirer of TİD.
+When the acquirer is a native speaker and the discourse function is
+introduction, the chances of observing a nominal referent gets even
+higher. Also, there seems to be an interaction between maintaining a
+referent and being a native acquirer of TİD. When it was a native
+acquirer of TİD and the discourse context was maintenance, the chances
+of observing a nominal referent gets even lower.
 
@@
 adult deaf signers with language deprivation in terms of their
-morphosyntactic (Boudreault and Mayberry, 2006; Cormier et al., 2012;
-Kayabaşı and Gökgöz, 2022) and narrative development (Becker, 2009;
-Cormier et al., 2013), we expected that late signers would be more
-over-redundant in their referent tracking, akin to second language
-learners (Bel et al., 2015; Frederiksen and Mayberry, 2019). That is, we
-had the expectation that late signers would use more explicit markers
-like nouns and fewer implicit markers like null marking even for highly
-accessible referents (when the discourse status is maintained or even
-re-introduced in same cases) than the control group (i.e., native
+morphosyntactic \citep{boudreault2006, cormier2012, kayabasi2020} and
+narrative development \citep{becker2009, cormier2013}, we expected that
+late signers would be more over-redundant in their referent tracking,
+akin to second language learners \citep{frederiksen2019, bel2015}. That
+is, we had the expectation that late signers would use more explicit
+markers like nouns and fewer implicit markers like null marking even for
+highly accessible referents (when the discourse status is maintained or
+even re-introduced in same cases) than the control group (i.e., native
 signers).
@@
 the present study brought back referents into the discourse more
-frequently compared to the participants in Frederiksen and Mayberry
-(2019) and Frederiksen and Mayberry (2016) which reported that
-re-introduction contexts accounted for only 7\% of the referents in ASL.
-On the other hand, maintenance percentages were lower in the present
-study compared to Frederiksen and Mayberry (2016) (40\% versus 69\%).
-Investigating the distribution of REs across discourse statuses in ASL,
-a similar study by Czubek (2017) also reported contrasting numbers (17\%
-for introduction, 64\% for maintenance, and 19\% for re-introduction).
-We suggest that the discrepancy might be due to the nature of the
-stimuli used in these research. In Frederiksen and Mayberry (2016), the
-authors used 4-6-episode-long picture story with 2 animate characters
-and 1 inanimate object. In contrast, the stimuli in the present study
-consisted of 8-episode-long video clips on average with a mean number of
-2 animate characters and 3 inanimate objects. Although the number of
-animate characters and inanimate characters in Czubek (2017)'s stimuli
+frequently compared to the participants in \citet{frederiksen2019} and
+\citet{frederiksen2016} which reported that re-introduction contexts
+accounted for only 7\% of the referents in ASL. On the other hand,
+maintenance percentages were lower in the present study compared to
+\citet{frederiksen2016} (40\% versus 69\%). Investigating the
+distribution of REs across discourse statuses in ASL, a similar study by
+\citet{czubek2017} also reported contrasting numbers (17\% for
+introduction, 64\% for maintenance, and 19\% for re-introduction). We
+suggest that the discrepancy might be due to the nature of the stimuli
+used in these research. In \citet{frederiksen2016}, the authors used
+4-6-episode-long picture story with 2 animate characters and 1 inanimate
+object. In contrast, the stimuli in the present study consisted of
+8-episode-long video clips on average with a mean number of 2 animate
+characters and 3 inanimate objects. Although the number of animate
+characters and inanimate characters in \citet{czubek2017}'s stimuli
 matched that of the present study, their narrative structure (a
@@
 of time. As a result, our signers engaged in referent re-introduction
-more than what was reported in previous studies (Czubek, 2017;
-Frederiksen and Mayberry, 2016).
+more than what was reported in previous studies
+\citep{czubek2017, frederiksen2016}.
 
@@
 finding parallels other accounts of referential accessibility suggested
-for other sign languages {[}see Frederiksen and Mayberry (2016); Czubek
-(2017); and Swabey (2002); for ASL, Morgan (2005); for BSL Bel et al.
-(2015); and Barberà and Massó (2009) for LSC. Our results also indicate
-that modified nouns (e.g., nouns that are preceded or followed by a
-classifier or IX) appeared more in introduced contexts compared to
-re-introduced contexts. This brings along the assumption that modified
-nouns are used more to track referents with lower accessibility,
-probably due to their more explicit structure.
+for other sign languages {[}see \citet{frederiksen2016};
+\citet{czubek2017}; and \citet{swabey2002}; for ASL, \citet{morgan2005};
+for BSL \citet{bel2015}; and \citet{barbera2009} for LSC. Our results
+also indicate that modified nouns (e.g., nouns that are preceded or
+followed by a classifier or IX) appeared more in introduced contexts
+compared to re-introduced contexts. This brings along the assumption
+that modified nouns are used more to track referents with lower
+accessibility, probably due to their more explicit structure.
 
@@
 were mainly used to re-introduce referents rather than maintain them.
-This finding is comparable with what has been reported for ASL (Czubek,
-2017; Frederiksen and Mayberry, 2016; Swabey, 2002) but not with LSC
-(Bel et al., 2015), BSL (Morgan, 2005) and DGS (Perniss and Özyürek,
-2015), in which the signers extensively made overt pronominal references
-(up to 30\%) to maintain entities. Comparable to our findings, It has
-been reported that pronouns do not frequently occur in the local context
-and instead follows a topic shift and thus rank higher than classifiers
-and null markers different genre (see Barberà and Massó, 2009 for LSC;
-and Nuhbalaoglu, 2018 for TİD). Considering this, the claim that the IX
-sign functions like demonstratives that are also used to track referents
-in re-introduced contexts has not been borne out (Koulidobrova and
-Lillo-Martin, 2016). Similar to these research, our findings indicate
-that the pronominal IX in TİD was used more concurrently with nominal
-constructions and pronominal IX was used very infrequently in isolation.
-The observation that overt pronouns are rare in ASL narratives comes
-from studies that include a simple picture retelling task with one
-protagonist (e.g., Frederiksen and Mayberry, 2016) as well as a more
-complex narrative design as in (Czubek, 2017). However, Czubek (2017)'s
-study also incorporated different genres and reported more instances of
-overt pronouns in personal narratives (compared to picture or video
+This finding is comparable with what has been reported for ASL
+\citep{swabey2002, frederiksen2016, czubek2017} but not with LSC
+\citep{bel2015}, BSL \citep{morgan2005} and DGS \citep{perniss2015}, in
+which the signers extensively made overt pronominal references (up to
+30\%) to maintain entities. Comparable to our findings, It has been
+reported that pronouns do not frequently occur in the local context and
+instead follows a topic shift and thus rank higher than classifiers and
+null markers different genre \citetext{\citealp[see][ for
+LSC]{barbera2009}; \citealp[and][ for TİD]{nuhbalaoglu2018}}.
+Considering this, the claim that the IX sign functions like
+demonstratives that are also used to track referents in re-introduced
+contexts has not been borne out \citep{koulidobrova2016}. Similar to
+these research, our findings indicate that the pronominal IX in TİD was
+used more concurrently with nominal constructions and pronominal IX was
+used very infrequently in isolation. The observation that overt pronouns
+are rare in ASL narratives comes from studies that include a simple
+picture retelling task with one protagonist
+\citep[e.g.,][]{frederiksen2016} as well as a more complex narrative
+design as in \citep{czubek2017}. However, \citet{czubek2017}'s study
+also incorporated different genres and reported more instances of overt
+pronouns in personal narratives (compared to picture or video
 retelling). In fact, this finding suggests that the use of pronouns
 might be genre-dependent. Moreover, another production study on referent
-tracking in ASL reported object preference for pronominals (Frederiksen
-and Mayberry, 2017). Nonetheless, we only coded subjects for maintained
-and re-introduced referents. We argue that all of these factors
-(genre-dependency as well as object preference) might contribute to the
-explanation of why overt pronouns in this data are dispreferred in favor
-of other markers for highly accessible referents such as null marking
-and classifiers.
+tracking in ASL reported object preference for pronominals
+\citep{frederiksen2017}. Nonetheless, we only coded subjects for
+maintained and re-introduced referents. We argue that all of these
+factors (genre-dependency as well as object preference) might contribute
+to the explanation of why overt pronouns in this data are dispreferred
+in favor of other markers for highly accessible referents such as null
+marking and classifiers.
 
@@
 less iconic and transparent nature of WECLs allowed signers to continue
-a previous referent without being redundant. As pointed out by Czubek
-(2017), signers prefer to rely on ExtCLs for the first mention of an
-entity because these constructions are informative and explicit by
-nature, and they provide visual and descriptive cues about an entity and
-its location in space. Nevertheless, we only observed 3 instances of
-ExtCLs for re-introduced contexts, replicating Frederiksen and Mayberry
-(2019)'s findings for ASL.
+a previous referent without being redundant. As pointed out by
+\citet{czubek2017}, signers prefer to rely on ExtCLs for the first
+mention of an entity because these constructions are informative and
+explicit by nature, and they provide visual and descriptive cues about
+an entity and its location in space. Nevertheless, we only observed 3
+instances of ExtCLs for re-introduced contexts, replicating
+\citet{frederiksen2019}'s findings for ASL.
 
@@
 of which were used for maintained contexts, respectively. This contrasts
-with some of the research (Czubek, 2017) in which agreement verbs ranked
-lowest on scale, indicating that these constructions were more implicit
-compared to plain verbs. However, there seems to be a common ground for
-the position of CA markers which have been claimed to rank higher than
-other null markers as in the present study (Barberà and Massó, 2009;
-Czubek, 2017; Frederiksen and Mayberry, 2016). In Figure
+with some of the research \citep{czubek2017} in which agreement verbs
+ranked lowest on scale, indicating that these constructions were more
+implicit compared to plain verbs. However, there seems to be a common
+ground for the position of CA markers which have been claimed to rank
+higher than other null markers as in the present study
+\citep{frederiksen2016, czubek2017, barbera2009}. In Figure
 \ref{fig:scale}, we report the distribution of REs in TİD as a function
 of referent accessibility and compare our accessibility hierarchy with
-that of (Frederiksen and Mayberry, 2016) for ASL. In the hierarchies,
-the left most REs (e.g., modified and bare nouns for TİD) are used when
-the accessibility is low. If the referent is already active in the
+that of \citep{frederiksen2016} for ASL. In the hierarchies, the left
+most REs (e.g., modified and bare nouns for TİD) are used when the
+accessibility is low. If the referent is already active in the
 addressee's mind and its accessibility is high, then the right most
@@
 both scales lack pronouns, which occurred very infrequently in the data.
-Although Frederiksen and Mayberry (2016) did not include other
-classifier types in the hierarchy, we proposed that WECLs and BPCLs must
-ranker higher than null markers but lower than ExtCLs that were mainly
-used by interlocutors to introduce a referent into the discourse
-context.
+Although \citet{frederiksen2016} did not include other classifier types
+in the hierarchy, we proposed that WECLs and BPCLs must ranker higher
+than null markers but lower than ExtCLs that were mainly used by
+interlocutors to introduce a referent into the discourse context.
 
@@
 study, a tendency for limited over-redundancy has been suggested for
-late signing children in DGS (Becker, 2009) and late signing adults in
-BSL specifically for the use of CA(Cormier et al., 2013). As for TİD, to
-our knowledge, there is only one study (Gür, 2018) that directly aimed
-at examining age of acquisition effects on adult signers' narrative
-skills. Gür (2018) observed that late signing deaf adults and children
-set the scene less frequently and used CA less accurately compared to
-their native counterparts. However, this particular investigation did
-not present any data on the distribution of REs between two acquisition
-groups. In the present study, the finding that native signers preferred
-to use null arguments even for the first mention of a referent (19 in
-total) is surprising given some of the previous literature in the
-topicality and saliency framework (Cormier et al., 2013; Swabey, 2002).
-These research reported very little or no use of null marking by native
-signers for introduced contexts.
+late signing children in DGS \citep{becker2009} and late signing adults
+in BSL specifically for the use of CA\citep{cormier2013}. As for TİD, to
+our knowledge, there is only one study \citep{gur2018} that directly
+aimed at examining age of acquisition effects on adult signers'
+narrative skills. \citet{gur2018} observed that late signing deaf adults
+and children set the scene less frequently and used CA less accurately
+compared to their native counterparts. However, this particular
+investigation did not present any data on the distribution of REs
+between two acquisition groups. In the present study, the finding that
+native signers preferred to use null arguments even for the first
+mention of a referent (19 in total) is surprising given some of the
+previous literature in the topicality and saliency framework
+\citep{cormier2013, swabey2002}. These research reported very little or
+no use of null marking by native signers for introduced contexts.
 
@@
 of the line of research that investigated deaf or hearing bimodal
-bilingual signers (Bel et al., 2015; Frederiksen and Mayberry,
-2019).\footnote{For similar results with heritage learners of sign language, see Reynolds (2018).}
+bilingual signers
+\citep{bel2015, frederiksen2019}.\footnote{For similar results with heritage learners of sign language, see Reynolds (2018).}
 The authors reported that more instances of null marking in the context
@@
 in the accessibility levels of re-introduced referents, a more
-fine-tuned protocol (as in Toole, 1996) can be used for measuring
+fine-tuned protocol \citep[as in][]{toole1996} can be used for measuring
 referent distance and the number of other possible competitors. Taken
@@
 
+\begin{verbatim}
+## Warning: Ignoring unknown aesthetics: height
+\end{verbatim}
+
 \begin{figure}
 \centering
-\includegraphics{manuscript_v2_files/figure-latex/fig:model-camera-1.pdf}
+\includegraphics{manuscript_v2_files/figure-latex/model-camera-1.pdf}
 \caption{Regression model plot for the RE Type (Nominal, Null) with the
@@
 \midrule
-muNominal\_Intercept & 1.44 & 0.17 & 1.12 & 1.81 & 1.02\\
-Null\_Intercept & -18.01 & 19.44 & -80.15 & -2.32 & 1.08\\
-Nom\_Late1vsNative & 0.25 & 0.43 & -0.52 & 1.16 & 1.02\\
-Nom\_Late2vsLate1 & -0.28 & 0.45 & -1.24 & 0.55 & 1.02\\
-Nom\_Intro & 4.24 & 0.45 & 3.50 & 5.24 & 1.06\\
-Nom\_Maintain & -5.33 & 0.30 & -5.97 & -4.79 & 1.05\\
-Nom\_L1vsN*Intro & 2.56 & 1.29 & 0.49 & 5.68 & 1.07\\
-Nom\_L2vsL1*Intro & -4.00 & 1.25 & -6.96 & -2.01 & 1.07\\
-Nom\_L1vsN*Maintain & -0.33 & 0.83 & -2.17 & 1.13 & 1.06\\
-Nom\_L2vsL1*Maintain & 0.47 & 0.83 & -1.01 & 2.29 & 1.07\\
-Null\_Late1vsNative & -6.19 & 28.69 & -75.09 & 57.19 & 1.06\\
-Null\_Late2vsLate1 & 10.31 & 26.94 & -34.28 & 88.61 & 1.06\\
-Null\_Intro & -77.04 & 77.76 & -326.00 & -14.23 & 1.08\\
-Null\_Maintain & 38.72 & 38.88 & 7.28 & 162.93 & 1.08\\
-Null\_L1vsN*Intro & -25.29 & 114.78 & -300.34 & 227.46 & 1.06\\
-Null\_L2vsL1*Intro & 40.13 & 107.78 & -137.50 & 352.79 & 1.06\\
-Null\_L1vsN*Maintain & 12.85 & 57.38 & -114.06 & 150.55 & 1.06\\
-Null\_L2vsL1*Maintain & -20.72 & 53.89 & -177.66 & 68.26 & 1.06\\
+1 & 1.23 & 0.20 & 0.83 & 1.62 & 1.00\\
+2 & -4.25 & 3.56 & -13.27 & 0.09 & 1.01\\
+3 & 0.49 & 0.54 & -0.55 & 1.56 & 1.01\\
+4 & -0.41 & 0.49 & -1.37 & 0.57 & 1.00\\
+5 & 3.76 & 0.25 & 3.29 & 4.25 & 1.00\\
+6 & -5.24 & 0.21 & -5.66 & -4.82 & 1.00\\
+7 & 1.01 & 0.63 & -0.16 & 2.30 & 1.00\\
+8 & 0.02 & 0.64 & -1.24 & 1.24 & 1.00\\
+9 & -0.33 & 0.55 & -1.43 & 0.73 & 1.00\\
+10 & -0.92 & 0.52 & -1.96 & 0.11 & 1.00\\
+11 & -13.40 & 10.69 & -40.57 & -0.41 & 1.01\\
+12 & 14.22 & 10.69 & 1.29 & 41.29 & 1.01\\
+13 & -21.98 & 14.25 & -58.10 & -4.70 & 1.01\\
+14 & 11.27 & 7.13 & 2.67 & 29.30 & 1.01\\
+15 & -54.37 & 42.74 & -162.14 & -2.68 & 1.01\\
+16 & 55.46 & 42.75 & 3.99 & 164.16 & 1.01\\
+17 & 26.68 & 21.38 & 0.88 & 80.69 & 1.01\\
+18 & -27.59 & 21.38 & -81.73 & -1.81 & 1.01\\
 \bottomrule
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-ahn2019}{}}%
-Ahn, Dorothy, 2019. The determinacy scale: A competition mechanism for
-anaphoric expressions (PhD thesis).
-
-\leavevmode\vadjust pre{\hypertarget{ref-ahrenholz2005}{}}%
-Ahrenholz, Bernt, 2005. Reference to persons and objects in the function
-of subject in learner varieties. The structure of learner varieties
-1964.
-
-\leavevmode\vadjust pre{\hypertarget{ref-aikhenvald2017}{}}%
-Aikhenvald, Alexandra, 2017. A typology of noun categorization devices.
-
-\leavevmode\vadjust pre{\hypertarget{ref-aksu-kouxe72015}{}}%
-Aksu-Koç, Ayhan, Nicolopoulou, Ageliki, 2015. Character reference in
-young children's narratives: A crosslinguistic comparison of English,
-Greek, and Turkish. Lingua, Reference and referentiality in native and
-learner grammars 155, 62--84.
-doi:\href{https://doi.org/10.1016/j.lingua.2014.04.006}{10.1016/j.lingua.2014.04.006}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ariel1990}{}}%
-Ariel, Mira, 1990. Accessing noun-phrase antecedents. Routledge.
-
-\leavevmode\vadjust pre{\hypertarget{ref-azar2015}{}}%
-Azar, Zeynep, Özyürek, Aslı, 2015. Discourse management: Reference
-tracking in speech and gesture in Turkish narratives. Dutch Journal of
-Applied Linguistics 4, 222--240.
-doi:\href{https://doi.org/10.1075/dujal.4.2.06aza}{10.1075/dujal.4.2.06aza}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bamberg1997}{}}%
-Bamberg, Michael GW, 1997. Positioning between structure and
-performance. Journal of narrative and life history 7, 335342.
-
-\leavevmode\vadjust pre{\hypertarget{ref-barberuxe02009}{}}%
-Barberà, Gemma, Massó, Guillem, 2009. Keeping track in sign language:
-Coreference annotation using centering theory, in: Talk Presented at the
-Sign Language Corpora: Linguistic Issues Workshop, University College
-London.
-
-\leavevmode\vadjust pre{\hypertarget{ref-becker2009}{}}%
-Becker, Claudia, 2009. Narrative competences of Deaf children in German
-Sign Language. Sign Language \& Linguistics 12, 113--160.
-doi:\href{https://doi.org/10.1075/sll.12.2.02bec}{10.1075/sll.12.2.02bec}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bel2015}{}}%
-Bel, Aurora, Ortells, Marta, Morgan, Gary, 2015. Reference control in
-the narratives of adult sign language learners. International Journal of
-Bilingualism 19, 608--624.
-doi:\href{https://doi.org/10.1177/1367006914527186}{10.1177/1367006914527186}
-
-\leavevmode\vadjust pre{\hypertarget{ref-benedicto2004}{}}%
-Benedicto, Elena, Brentari, Diane, 2004. Where did all the arguments
-go?: Argument-changing properties of classifiers in ASL. Natural
-Language \& Linguistic Theory 22, 743810.
-
-\leavevmode\vadjust pre{\hypertarget{ref-berman2013relating}{}}%
-Berman, Ruth A, Slobin, Dan Isaac, 2013. Relating events in narrative: A
-crosslinguistic developmental study. Psychology Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-bosch2007}{}}%
-Bosch, Peter, Umbach, Carla, 2007. Reference determination for
-demonstrative pronouns. ZAS Papers in Linguistics 48, 39--51.
-doi:\href{https://doi.org/10.21248/zaspil.48.2007.353}{10.21248/zaspil.48.2007.353}
-
-\leavevmode\vadjust pre{\hypertarget{ref-boudreault2006}{}}%
-Boudreault, Patrick, Mayberry, Rachel I., 2006. Grammatical processing
-in american sign language: Age of first-language acquisition effects in
-relation to syntactic structure. Language and Cognitive Processes 21,
-608--635.
-doi:\href{https://doi.org/10.1080/01690960500139363}{10.1080/01690960500139363}
-
-\leavevmode\vadjust pre{\hypertarget{ref-buxfcrkner2018}{}}%
-Bürkner, Paul-Christian, 2018. Advanced bayesian multilevel modeling
-with the R package brms. The R Journal 10, 395.
-doi:\href{https://doi.org/10.32614/RJ-2018-017}{10.32614/RJ-2018-017}
-
-\leavevmode\vadjust pre{\hypertarget{ref-chafe1976}{}}%
-Chafe, W., 1976.
-\href{https:///paper/Givenness\%2C-contrastiveness\%2C-definiteness\%2C-subjects\%2C-Chafe/805a12a9bee00562f6c2663722e9f7f3f5991848}{Givenness,
-contrastiveness, definiteness, subjects, topics, and point of view}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-chamberlain_language_1999}{}}%
-Chamberlain, Charlene, Morford, Jill P., Mayberry, Rachel I. (Eds.),
-1999. Language acquisition by eye. Psychology Press, Mahwah, N.J.
-
-\leavevmode\vadjust pre{\hypertarget{ref-cheng2021}{}}%
-Cheng, Qi, Mayberry, Rachel I., 2021. When event knowledge overrides
-word order in sentence comprehension: Learning a first language after
-childhood. Developmental Science.
-doi:\href{https://doi.org/10.1111/desc.13073}{10.1111/desc.13073}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cormier2012}{}}%
-Cormier, Kearsy, Schembri, Adam, Vinson, David, Orfanidou, Eleni, 2012.
-First language acquisition differs from second language acquisition in
-prelingually deaf signers: Evidence from sensitivity to grammaticality
-judgement in British Sign Language. Cognition 124, 50--65.
-doi:\href{https://doi.org/10.1016/j.cognition.2012.04.003}{10.1016/j.cognition.2012.04.003}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cormier2013}{}}%
-Cormier, Kearsy, Smith, Sandra, Zwets, Martine, 2013. Framing
-constructed action in British Sign Language narratives. Journal of
-Pragmatics 55, 119--139.
-doi:\href{https://doi.org/10.1016/j.pragma.2013.06.002}{10.1016/j.pragma.2013.06.002}
-
-\leavevmode\vadjust pre{\hypertarget{ref-crasborn2008}{}}%
-Crasborn, O., Sloetjes, H., 2008.
-\href{https://www.semanticscholar.org/paper/Enhanced-ELAN-functionality-for-sign-language-Crasborn-Sloetjes/e1623ec6b2bc3532129a3682eb263cc2e7d5389c}{Enhanced
-ELAN functionality for sign language corpora}. undefined.
-
-\leavevmode\vadjust pre{\hypertarget{ref-czubek2017}{}}%
-Czubek, Todd Alan, 2017. A comprehensive study of referring expressions
-in ASL 244.
-
-\leavevmode\vadjust pre{\hypertarget{ref-debreslioska2013}{}}%
-Debreslioska, Sandra, Özyürek, Asli, Gullberg, Marianne, Perniss,
-Pamela, 2013. Gestural viewpoint signals referent accessibility.
-Discourse Processes 50, 431--456.
-doi:\href{https://doi.org/10.1080/0163853X.2013.824286}{10.1080/0163853X.2013.824286}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dikyuva2015}{}}%
-Dikyuva, Hasan, Makaroğlu, Bahtiyar, Arık, Engin, 2015. Türk i{ş}aret
-dili dilbilgisi kitab{ı}. Aile ve Sosyal Politikalar Bakanl{ı}{ğ}{ı}
-Yay{ı}nlar{ı}: Ankara.
-
-\leavevmode\vadjust pre{\hypertarget{ref-dudis2004}{}}%
-Dudis, Paul G., 2004.
-\href{https://www.degruyter.com/view/j/cogl.2004.15.issue-2/cogl.2004.009/cogl.2004.009.xml}{Body
-partitioning and real-space blends}. Cognitive Linguistics 15.
-
-\leavevmode\vadjust pre{\hypertarget{ref-emmorey1996}{}}%
-Emmorey, Karen, 1996. The confluence of space and language in signed
-languages, in: Language, Speech, and Communication. The MIT Press,
-Cambridge, MA, US, pp. 171--209.
-
-\leavevmode\vadjust pre{\hypertarget{ref-engberg-pedersen1993}{}}%
-Engberg-Pedersen, Elisabeth, 1993. Space in danish sign language: The
-semantics and morphosyntax of the use of space in a visual language.
-
-\leavevmode\vadjust pre{\hypertarget{ref-frederiksen2016}{}}%
-Frederiksen, Anne Therese, Mayberry, Rachel I., 2016. Who's on first?
-Investigating the referential hierarchy in simple native ASL narratives.
-Lingua 180, 49--68.
-doi:\href{https://doi.org/10.1016/j.lingua.2016.03.007}{10.1016/j.lingua.2016.03.007}
-
-\leavevmode\vadjust pre{\hypertarget{ref-frederiksen2017}{}}%
-Frederiksen, Anne Therese, Mayberry, Rachel I., 2017. Pronominalizing
-subjects: Referentially empty pronouns in ASL.
-
-\leavevmode\vadjust pre{\hypertarget{ref-frederiksen2019}{}}%
-Frederiksen, Anne Therese, Mayberry, Rachel I., 2019. Reference tracking
-in early stages of different modality L2 acquisition: Limited
-over-explicitness in novice ASL signers{'} referring expressions. Second
-Language Research 35, 253--283.
-doi:\href{https://doi.org/10.1177/0267658317750220}{10.1177/0267658317750220}
-
-\leavevmode\vadjust pre{\hypertarget{ref-givon1983}{}}%
-Givon, T., 1983. \href{https://benjamins.com/catalog/tsl.3}{Topic
-continuity in discourse}. John Benjamins Publishing Company.
-
-\leavevmode\vadjust pre{\hypertarget{ref-grice1975}{}}%
-Grice, Herbert P., 1975. Logic and conversation. Brill, p. 4158.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gullberg2006}{}}%
-Gullberg, Marianne, 2006. Handling discourse: Gestures, reference
-tracking, and communication strategies in early L2: Language Learning
-Vol. 56, No. 1. Language Learning 56, 155--196.
-doi:\href{https://doi.org/10.1111/j.0023-8333.2006.00344.x}{10.1111/j.0023-8333.2006.00344.x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gundel1993}{}}%
-Gundel, Jeanette K., Hedberg, Nancy, Zacharski, Ron, 1993. Cognitive
-status and the form of referring expressions in discourse. Language 69,
-274--307. doi:\href{https://doi.org/10.2307/416535}{10.2307/416535}
-
-\leavevmode\vadjust pre{\hypertarget{ref-guxfcr2018}{}}%
-Gür, Cansu, 2018. The effects of late language acquisition on the
-development of narrative skills: Insights from Turkish Sign Language
-(PhD thesis). Erzurum.
-
-\leavevmode\vadjust pre{\hypertarget{ref-hickmann1996}{}}%
-Hickmann, Maya, Hendriks, Henriëtte, Roland, Françoise, Liang, James,
-1996. The marking of new information in children's narratives: a
-comparison of English, French, German and Mandarin Chinese*. Journal of
-Child Language 23, 591--619.
-doi:\href{https://doi.org/10.1017/S0305000900008965}{10.1017/S0305000900008965}
-
-\leavevmode\vadjust pre{\hypertarget{ref-karaduxf6ller2017}{}}%
-Karadöller, Dilay Z, Sümer, Beyza, Özyürek, Aslı, 2017. Effects of
-delayed language exposure on spatial language acquisition by signing
-children and adults 5.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kayaba2022}{}}%
-Kayabaşı, Demet, Gökgöz, Kadir, 2022. The causative-inchoative
-alternation and age-of-acquisition effects on multi-predicate
-constructions in turkish sign language. Language Learning and
-Development 0, 1--26.
-doi:\href{https://doi.org/10.1080/15475441.2022.2047690}{10.1080/15475441.2022.2047690}
-
-\leavevmode\vadjust pre{\hypertarget{ref-koulidobrova2016}{}}%
-Koulidobrova, Elena, Lillo-Martin, Diane, 2016. A {`}point{'}of inquiry:
-The case of the (non-) pronominal IX in ASL. The impact of pronominal
-form on interpretation 221, 250.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kruschke2011}{}}%
-Kruschke, John K., 2011. Bayesian assessment of null values via
-parameter estimation and model comparison. Perspectives on Psychological
-Science 6, 299312.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kubu2008}{}}%
-Kubuş, Okan, 2008. An analysis of Turkish Sign Language (T{İ}D)
-phonology and morphology (PhD thesis). Ankara.
-
-\leavevmode\vadjust pre{\hypertarget{ref-lillo-martin1990}{}}%
-Lillo-Martin, Diane, Klima, Edward S., 1990. Pointing out differences:
-ASL pronouns in syntactic theory. Theoretical issues in sign language
-research, Volume 1: Linguistics 191.
-
-\leavevmode\vadjust pre{\hypertarget{ref-loew1984}{}}%
-Loew, Ruth Carolyn, 1984.
-\href{https:///paper/Roles-and-reference-in-American-sign-language-\%3A-a-Loew/315c2c539d3bd88fbc13564775141ba4a60ac5aa}{Roles
-and reference in American sign language : a developmental perspective}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-maclaughlin1997}{}}%
-Maclaughlin, Dawn, 1997.
-\href{http://lear.unive.it/jspui/handle/11707/7016}{The structure of
-determiner phrases: Evidence from American Sign Language.} (PhD thesis).
-
-\leavevmode\vadjust pre{\hypertarget{ref-meier1990}{}}%
-Meier, Richard P., Newport, Elissa L., 1990. Out of the hands of babes:
-On a possible sign advantage in language acquisition. Language 66, 1.
-doi:\href{https://doi.org/10.2307/415277}{10.2307/415277}
-
-\leavevmode\vadjust pre{\hypertarget{ref-metzger1995}{}}%
-Metzger, Melanie, 1995. Constructed action and constructed dialogue in
-american sign language. Sociolinguistics in Deaf communities 255271.
-
-\leavevmode\vadjust pre{\hypertarget{ref-morgan2005}{}}%
-Morgan, Gary, 2005.
-\href{http://www.oxfordscholarship.com/view/10.1093/acprof:oso/9780195180947.001.0001/acprof-9780195180947}{The
-development of narrative skills in British Sign Language}, in: Schick,
-Brenda, Marschark, Marc, Spencer, Patricia Elizabeth (Eds.),. Oxford
-University Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-newport1990}{}}%
-Newport, Elissa L., 1990. Maturational constraints on language learning.
-Cognitive Science 14, 11--28.
-doi:\href{https://doi.org/10.1207/s15516709cog1401_2}{10.1207/s15516709cog1401\_2}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nuhbalaoglu2018}{}}%
-Nuhbalaoglu, Derya, 2018. Comprehension and production of referential
-expressions in German Sign Language and Turkish Sign Language: An
-empirical approach (PhD thesis). Germany.
-
-\leavevmode\vadjust pre{\hypertarget{ref-padden1986}{}}%
-Padden, Carol, 1986. Verbs and role-shifting in american sign language.
-National Association of the Deaf Silver Spring, MD, p. 57.
-
-\leavevmode\vadjust pre{\hypertarget{ref-perniss2015}{}}%
-Perniss, Pamela, Özyürek, Asli, 2015. Visible Cohesion: A comparison of
-reference tracking in sign, speech, and co-speech gesture. Topics in
-Cognitive Science 7, 36--60.
-doi:\href{https://doi.org/10.1111/tops.12122}{10.1111/tops.12122}
-
-\leavevmode\vadjust pre{\hypertarget{ref-prince1981}{}}%
-Prince, EF, 1981. Toward a taxonomy of given - new information, in:
-Cole, Peter (Ed.),. Academic Press, New York, pp. 223--255.
-
-\leavevmode\vadjust pre{\hypertarget{ref-rathmann2007}{}}%
-Rathmann, Christian, Mann, Wolfgang, Morgan, Gary, 2007. Narrative
-structure and narrative development in deaf children. Deafness \&
-Education International 9, 187--196.
-doi:\href{https://doi.org/10.1179/146431507790559932}{10.1179/146431507790559932}
-
-\leavevmode\vadjust pre{\hypertarget{ref-sevinuxe72006}{}}%
-Sevinç, Ayça Müge, 2006. Grammatical relations and word order in turkish
-sign language (t{İ}d). Unpublished master's thesis, Middle East
-Technical University, Turkey.
-
-\leavevmode\vadjust pre{\hypertarget{ref-smith2014}{}}%
-Smith, Sandra, Cormier, Kearsy, 2014. In or out?: Spatial scale and
-enactment in narratives of native and nonnative signing deaf children
-acquiring British Sign Language. Sign Language Studies 14, 275--301.
-doi:\href{https://doi.org/10.1353/sls.2014.0008}{10.1353/sls.2014.0008}
-
-\leavevmode\vadjust pre{\hypertarget{ref-sorace2011}{}}%
-Sorace, Antonella, 2011. Pinning down the concept of {``}interface{''}
-in bilingualism. Linguistic approaches to bilingualism 1, 133.
-
-\leavevmode\vadjust pre{\hypertarget{ref-sorace2009}{}}%
-Sorace, Antonella, Serratrice, Ludovica, 2009.
-\href{https://doi.org/10.1177/1367006909339810}{Internal and external
-interfaces in bilingual language development: Beyond structural
-overlap}. International Journal of Bilingualism 13, 195--210.
-
-\leavevmode\vadjust pre{\hypertarget{ref-supalla1986}{}}%
-Supalla, Ted, 1986. The classifier system in american sign language.
-Noun classes and categorization 7, 181214.
-
-\leavevmode\vadjust pre{\hypertarget{ref-supalla1990}{}}%
-Supalla, Ted, 1990. Serial verbs of motion in ASL. Theoretical issues in
-sign language research 1, 127152.
-
-\leavevmode\vadjust pre{\hypertarget{ref-swabey2002}{}}%
-Swabey, Laurie Anne, 2002. The cognitive status, form and distribution
-of referring expressions in ASL and English narratives (PhD thesis).
-Minnepolis, USA.
-
-\leavevmode\vadjust pre{\hypertarget{ref-talmy2003}{}}%
-Talmy, Leonard, 2003. The representation of spatial structure in spoken
-and signed language. Perspectives on classifier constructions in sign
-language 169195.
-
-\leavevmode\vadjust pre{\hypertarget{ref-toole1996}{}}%
-Toole, Janine, 1996. The effect of genre on referential choice. p. 263.
-doi:\href{https://doi.org/10.1075/pbns.38.16too}{10.1075/pbns.38.16too}
-
-\leavevmode\vadjust pre{\hypertarget{ref-wagenmakers2007}{}}%
-Wagenmakers, Eric-Jan, 2007. A practical solution to the pervasive
-problems of p values. Psychonomic bulletin \& review 14, 779804.
-doi:\href{https://doi.org/10.3758/bf03194105}{10.3758/bf03194105}
-
-\leavevmode\vadjust pre{\hypertarget{ref-williams1988}{}}%
-Williams, Jessica, 1988. Zero anaphora in second language acquisition: A
-comparison among three varieties of English. Studies in Second Language
-Acquisition 10, 339--370.
-doi:\href{https://doi.org/10.1017/S0272263100007488}{10.1017/S0272263100007488}
-
-\leavevmode\vadjust pre{\hypertarget{ref-woll2013}{}}%
-Woll, Bencie, 2013. Sign language and spoken language development in
-young children: Measuring vocabulary by means of the CDI, in: Meurant,
-Laurence, Sinte, Aurélie, Van Herreweghe, Mieke, Vermeerbergen, Myriam
-(Eds.),. De Gruyter, Berlin, Boston, pp. 15--34.
-doi:\href{https://doi.org/10.1515/9781614511472.15}{10.1515/9781614511472.15}
-
-\end{CSLReferences}
+\renewcommand\refname{References}
+\bibliography{./bibliography/references.bib}
 

### 2022-08-05 manuscript_v2.Rmd
model all vs camera only
The plot is added for the comparison of all vs camera only participants.

@@
 
-```{r model-camera, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
+```{r model-camera, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=3.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
 df_model_1 %<>% mutate(model = "All")
@@
   geom_point(size = 2.5, position = position_dodge(width = 0.6)) +
-  geom_errorbarh(aes(xmin = l, xmax = h), alpha = 1, height = 0, size = 1.5,position = position_dodge(width = 0.6)) +
-  geom_errorbarh(aes(xmin = ll, xmax = hh, height = 0), position = position_dodge(width = 0.6)) + vline_0(linetype = "dashed") +
+  geom_errorbarh(aes(xmin = l, xmax = h), alpha = 1, height = 0, size = 1.5, position = position_dodge(width = 0.6)) +
+  geom_errorbarh(aes(xmin = ll, xmax = hh), height=0, position = position_dodge(width = 0.6)) + vline_0(linetype = "dashed") +
   xlab("Estimate (log odds)") + ylab("Coefficients") + theme_apa()

### 2022-08-06 manuscript_v2.tex
model all vs camera only
The plot is added for the comparison of all vs camera only participants.
updated files

--- 2022-08-05 manuscript_v2.tex
@@
 exposure to a first functional language is reported to negatively affect
-deaf children's morphosyntactic production and processing (Boudreault
-and Mayberry, 2006; Cheng and Mayberry, 2021; Cormier et al., 2012;
-Karadöller et al., 2017; Kayabaşı and Gökgöz, 2022; Newport, 1990) and
-narrative abilities (e.g., Becker, 2009). In our overall research, we
-are concerned with the effects of this late exposure in the ultimate
-adult forms of TİD users.
+deaf children and adults' morphosyntactic production and processing
+(Boudreault and Mayberry, 2006; Cheng and Mayberry, 2021; Cormier et
+al., 2012; Karadöller et al., 2017; Kayabaşı et al., 2020; Newport,
+1990) and narrative abilities (e.g., Becker, 2009). In our overall
+research, we are concerned with the effects of this late exposure in the
+ultimate adult forms of TİD users.
 
@@
 referent as in an utterance like (\ref{ex:1}), phonetically simplest
-form is selected with the least redundant information are selected based
-on an economy principle. In the utterance, REs for ``Eddie'' are given
-in bold. All three REs (nominal, pronominal, and null marking) refer to
-the same entity; however, the quantity of marking decreases as the
+form is selected with the least redundant information based on an
+economy principle. In the utterance, REs for ``Eddie'' are given in
+bold. All three REs (nominal, pronominal, and null marking) refer to the
+same entity; however, the quantity of marking decreases as the
 referent's accessibility increases. On this ground, for highly
@@
 are two referents, \textsc{elvan}, who is referentially-assigned to the
-right side of the signer, represented with the ``a'' subscript and
+right side of the signer, represented with the ``A'' subscript and
 \textsc{demet}, who is referentially-assigned to the left side of the
-signer, represented with the ``b'' subscript. When the signer makes a
+signer, represented with the ``B'' subscript. When the signer makes a
 contrast in (\ref{ex:2b}), as a pronominal use, she points to the right
 side to refer to ELVAN and to the left to refer to \textsc{demet}. We
-schematically represents the referential locus assigned to the referents
+schematically represent the referential locus assigned to the referents
 in (\ref{ex:3}).
@@
 predicate. Some classifier handshapes and the set of objects that they
-denote are below. Some classifier handshapes and the set of objects that
-they denote are presented below in (\ref{ex:4}).
+denote are presented below in (\ref{ex:4}).
 
@@
 
-In (\ref{ex:5}), in the first clause, a referential locus, i.e.~locus-a,
+In (\ref{ex:5}), in the first clause, a referential locus, i.e.~locus-A,
 is assigned to the locative source argument \textsc{cage}. In the second
@@
 referential locus of the \textsc{cage} (locus-a) to the referential
-locus of the goal \textsc{tree} (locus-b).
+locus of the goal \textsc{tree} (locus-B).
 
@@
     \ex \textsc{tree} assigned to the referential locus-b. The source and goal tracked on the predicates \textsc{fly} and \textsc{land}. \\ \\
-    \textsc{bird fly[manner]-a tree-b land. CL-b}\\
+    \textsc{bird fly[manner]-a tree-b land cl-b}\\
     ‘The bird flies (from the cage assigned to the left) to a tree (assigned to the right).’
@@
 shape (Talmy, 2003). As in most sign languages hitherto studied,
-classifier constructions are also common and found in TİD (see Dikyuva
-et al., 2015 for a review). Following the classification of
+classifier constructions are also commonly found in TİD (see Dikyuva et
+al., 2015 for a review). Following the classification of
 Engberg-Pedersen (1993), we distinguish between four types of classifier
@@
 and so forth). Handling classifiers (HCL) occur with transitive verbs
-and depict how an object is handled or manipulated by agent. On this
+and depict how an object is handled or manipulated by an agent. On this
 ground, an HCL\footnote{As discussed before, an HCL is transitive in
@@
 extensively studied among second language learners (see Bel et al.,
-2015; Frederiksen and Mayberry, 2019). Comparing 11 deaf adults signers
+2015; Frederiksen and Mayberry, 2019). Comparing 11 deaf adult signers
 and 13 hearing advanced learners of Catalan Sign Language (Llengua de
@@
 to avoid potential ambiguities among referents. Nevertheless, their
-results did not reach statistical significant for the interaction of
+results did not reach statistical significance for the interaction of
 discourse status and acquisition group. A similar line of research came
@@
 
-Still, only very few research examined whether linguistically deprived
-signers (i.e., deaf signers with delayed exposure to a sign language)
-also display redundancy in their reference tracking. Becker (2009)
-examined 6 DGS signing deaf children (3 native and 3 late)`s narrative
-abilities. She reported that late signing children generated more
-explicit forms such as nominals when referring to the characters in the
-story compared to native learners, concluding that first language
+Still, only very few researchers examined whether linguistically
+deprived signers (i.e., deaf signers with delayed exposure to a sign
+language) also display redundancy in their reference tracking. Becker
+(2009) examined 6 DGS signing deaf children's (3 native and 3 late)
+narrative abilities. She reported that late signing children generated
+more explicit forms such as nominals when referring to the characters in
+the story compared to native learners, concluding that first language
 deprivation results in an inadequate representation of the addressee and
 less efficient use of morphosyntax. Comparably, Cormier et al. (2013)
-investigated the use of CA in 15 BSL among deaf native and late adult
-signers and found that late signers were redundant in topic continuity.
-In other words, the late signers produced more explicit forms like
-nominals, and fewer instances of null marking when maintaining previous
-referents. Most relevantly to the present study, Gür (2018) studied the
-development of narrative skills among 10 native and 10 late deaf
-children and deaf adult signers of TİD. She found that native-signing
-children and adults had more instances of explicit referent
-introduction, set the scene more frequently, and employed CA in an
-appropriate manner more than corresponding late-signing children and
-adults. Although this study presented preliminary findings on age of
-acquisition effects on narrative competence, it does not investigate how
-native and late signers employ different linguistic tools based on the
-accessibility of the referent. Therefore, to our knowledge, there is a
-lack of research directly examining the effects of first language
-deprivation on deaf adult signers' selection of REs in signed
-narratives.
+investigated the use of CA in BSL among 15 deaf native, early, and late
+adult signers (5 participants in each group) and found that late signers
+were redundant in topic continuity. In other words, the late signers
+produced more explicit forms like nominals, and fewer instances of null
+marking when maintaining previous referents. Most relevantly to the
+present study, Gür (2018) studied the development of narrative skills
+among 10 native and 10 late deaf children and deaf adult signers of TİD.
+She found that native-signing children and adults had more instances of
+explicit referent introduction, set the scene more frequently, and
+employed CA in an appropriate manner more than corresponding
+late-signing children and adults. Although this study presented
+preliminary findings on age of acquisition effects on narrative
+competence, it does not investigate how native and late signers employ
+different linguistic tools based on the accessibility of the referent.
+Therefore, to our knowledge, there is a lack of research directly
+examining the effects of first language deprivation on deaf adult
+signers' selection of REs in signed narratives.
 
@@
 others that have lower accessibility will be tracked with more explicit
-REs Also based on previous research, we hypothesize that late signers
-will become more over-redundant in their reference tracking compared to
+REs. Also, based on previous research, we hypothesize that late signers
+will be more over-redundant in their reference tracking compared to
 native signers. This would translate to a preference by late signers for
@@
 Although we did not set a cut-off point for nativeness, all the deaf of hearing participants indicated that they learned TİD after age 3.
-In order to capture in-group variation among late signers, we run an additional supplementary analysis}.
+In order to capture in-group variation among late signers, we run an additional supplementary analysis in which we observed very wide credible intervals due to the small group size, rendering the results uninformative.
+For the analysis, see \ref{section:d}}.
 All participants indicated that their preferred language of
@@
 4 and 7, all the other late signers were pre-lingually deaf. All late
-signers completed middle school education and the average year of
-schooling was 12.3 years (\emph{SD} = 2.8 years). All late participants
-attended at least one deaf school. While 8 participants attended two
-deaf schools (either primary and middle school, or middle and high
-school), 3 participants reported attend three deaf schools. Using the
-same scale, late signers self-rated their abilities in TİD and (spoken)
-Turkish. Both of their receptive and productive TİD abilities rated on
-average 4.9 points (\emph{SD} = 0.3 points). Late signers' mean rating
-for their reading abilities in Turkish was 3.5 points (\emph{SD} = 0.7
-points) while their writing and speaking abilities rated on average 3.4
-(\emph{SD} = 0.5 points) and 3.1 points (\emph{SD} = 0.7 points),
-respectively.
+signers completed at least middle school education and the year of
+schooling ranged between 8 and 16 years, and the average was 12.3 years
+(\emph{SD} = 2.8 years). All late participants attended at least one
+deaf school. While 8 participants attended two deaf schools (either
+primary and middle school, or middle and high school), 3 participants
+reported to attend three deaf schools. Using the same scale, late
+signers self-rated their abilities in TİD and Turkish. Both of their
+receptive and productive TİD abilities rated on average 4.9 points
+(\emph{SD} = 0.3 points). Late signers' mean rating for their reading
+abilities in Turkish was 3.5 points (\emph{SD} = 0.7 points) while their
+writing and speaking abilities rated on average 3.4 (\emph{SD} = 0.5
+points) and 3.1 points (\emph{SD} = 0.7 points), respectively.
 
@@
  & Reintroduced & A subject referent discontinued across at least one clause\\
-REAT Type & NOM: Bare Noun & A non-modified noun phrase (e.g., CAT, MOUSE)\\
+RE Type & NOM: Bare Noun & A non-modified noun phrase (e.g., CAT, MOUSE)\\
  & NOM: Modified Noun & A noun phrase modified by an index sign or an adjective\\
@@
 each clip either to the experimenter or the recording digital
-camera\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining e did not make it obligatory for participants to either narrate to the digital camera or to the experimenter for few reasons.
+camera\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining we did not make it obligatory for participants to either narrate to the digital camera or to the experimenter for few reasons.
 First, a native or native-like deaf experimenter was not always available, and a hearing late signer of TİD attended some sessions as the experimenter.
@@
 Furhermore, some signers did not feel comfortable narrating to the camera and, as a result, narrated to the deaf research assistant, if available.
-We ran an additional model excluding the 4 signers who narrated to the experimenter and compared it with the original model including these signers (see \ref{section:c} The results imply that there is not much of a difference.}.
+We ran an additional model excluding the 4 signers who narrated to the experimenter and compared it with the original model including these signers (see \ref{section:c}).
+The results imply that there is not much of a difference.}.
 
@@
 Re-introduction), its Type (Nominal, Pronominal, Classifer,
-Null)\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer Null).
+Null)\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer, and Null).
 However, we allude to the numerical distribution of the subtypes in this and following sections},
 and Participant Group (Native,
-Late)\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17).
-Treating age of acquisition as a continous variable would not be informative due to the small sample size in each interval.}.
+Late)\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17).}.
 
@@
 signers used CA more for referent re-introduction than maintenance. We
-observed a similar instances of agreement verbs for maintained and
+observed similar instances of agreement verbs for maintained and
 re-introduced contexts. In contrast, plain verbs were favored more in
@@
 For more inference on the data and having pairwise comparisons, we fit a
-regression model to RE Type using Discourse Status and Group as
-predictors. We used the brms package (Bürkner, 2018) in R, with a
+regression model to RE Type using discourse status and acquisition group
+as predictors. We used the brms package (Bürkner, 2018) in R, with a
 categorical distribution on the responses and sum contrasts for the
@@
 from the following analyses. For the current analysis, we did not define
-priors ourselves but used the preset uninformed priors. Figure
-2\footnote{We ran another model where we used the Nativeness predictor
-  with 3 levels 0-3:Native, 4-7:Late1, and 7-above:Late2. However, our
-  model did not converg properly. We provide the model results as a plot
-  and table in the \ref{section:d}.} shows the posterior probability
-distributions where the dot is the median estimate, the thick line is
-the 50\% credible intervals, and the thin line is the 95\% credible
-intervals. There are 3 responses we are considering in this model, and
-they are Classifier, Nominative, and Null. Classifier response is the
-level that other responses Nominative and Null are compared against.
+priors ourselves and used the preset uninformed priors.
+@ref\{fig:model-gram\}\footnote{We ran another model where we used the
+  Nativeness predictor with 3 levels 0-3:Native, 4-7:Late1, and
+  7-above:Late2. However, our model did not converge properly and the
+  credible intervals were very wide spread, yielding the results
+  uninformative. We provide the model results as a plot in
+  \ref{section:d}.} shows the posterior probability distributions where
+the dot is the median estimate, the thick line is the 50\% credible
+intervals, and the thin line is the 95\% credible intervals. There are 3
+responses we are considering in this model. They are; Classifier,
+Nominative, and Null. Classifier response is the level that other
+responses Nominative and Null are compared against.
+
+\begin{verbatim}
+## Warning: Ignoring unknown aesthetics: height
+\end{verbatim}
 
@@
 
-\hypertarget{general-discussion}{%
-\section{General Discussion}\label{general-discussion}}
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
 
@@
 morphosyntactic (Boudreault and Mayberry, 2006; Cormier et al., 2012;
-Kayabaşı and Gökgöz, 2022) and narrative development (Becker, 2009;
-Cormier et al., 2013), we expected that late signers would be more
+Kayabaşı et al., 2020) and narrative development (Becker, 2009; Cormier
+et al., 2013), we expected that late signers would be more
 over-redundant in their referent tracking, akin to second language
@@
 \centering
-\includegraphics{manuscript_v2_files/figure-latex/fig:model-camera-1.pdf}
+\includegraphics{manuscript_v2_files/figure-latex/model-camera-1.pdf}
 \caption{Regression model plot for the RE Type (Nominal, Null) with the
@@
 \midrule
-muNominal\_Intercept & 1.44 & 0.17 & 1.12 & 1.81 & 1.02\\
-Null\_Intercept & -18.01 & 19.44 & -80.15 & -2.32 & 1.08\\
-Nom\_Late1vsNative & 0.25 & 0.43 & -0.52 & 1.16 & 1.02\\
-Nom\_Late2vsLate1 & -0.28 & 0.45 & -1.24 & 0.55 & 1.02\\
-Nom\_Intro & 4.24 & 0.45 & 3.50 & 5.24 & 1.06\\
-Nom\_Maintain & -5.33 & 0.30 & -5.97 & -4.79 & 1.05\\
-Nom\_L1vsN*Intro & 2.56 & 1.29 & 0.49 & 5.68 & 1.07\\
-Nom\_L2vsL1*Intro & -4.00 & 1.25 & -6.96 & -2.01 & 1.07\\
-Nom\_L1vsN*Maintain & -0.33 & 0.83 & -2.17 & 1.13 & 1.06\\
-Nom\_L2vsL1*Maintain & 0.47 & 0.83 & -1.01 & 2.29 & 1.07\\
-Null\_Late1vsNative & -6.19 & 28.69 & -75.09 & 57.19 & 1.06\\
-Null\_Late2vsLate1 & 10.31 & 26.94 & -34.28 & 88.61 & 1.06\\
-Null\_Intro & -77.04 & 77.76 & -326.00 & -14.23 & 1.08\\
-Null\_Maintain & 38.72 & 38.88 & 7.28 & 162.93 & 1.08\\
-Null\_L1vsN*Intro & -25.29 & 114.78 & -300.34 & 227.46 & 1.06\\
-Null\_L2vsL1*Intro & 40.13 & 107.78 & -137.50 & 352.79 & 1.06\\
-Null\_L1vsN*Maintain & 12.85 & 57.38 & -114.06 & 150.55 & 1.06\\
-Null\_L2vsL1*Maintain & -20.72 & 53.89 & -177.66 & 68.26 & 1.06\\
+Nom\_Intercept & 1.23 & 0.20 & 0.83 & 1.62 & 1.00\\
+Null\_Intercept & -4.25 & 3.56 & -13.27 & 0.09 & 1.01\\
+Nom\_Late1vsNative & 0.49 & 0.54 & -0.55 & 1.56 & 1.01\\
+Nom\_Late2vsLate1 & -0.41 & 0.49 & -1.37 & 0.57 & 1.00\\
+Nom\_Intro & 3.76 & 0.25 & 3.29 & 4.25 & 1.00\\
+Nom\_Maintain & -5.24 & 0.21 & -5.66 & -4.82 & 1.00\\
+Nom\_L1vsN*Intro & 1.01 & 0.63 & -0.16 & 2.30 & 1.00\\
+Nom\_L2vsL1*Intro & 0.02 & 0.64 & -1.24 & 1.24 & 1.00\\
+Nom\_L1vsN*Maintain & -0.33 & 0.55 & -1.43 & 0.73 & 1.00\\
+Nom\_L2vsL1*Maintain & -0.92 & 0.52 & -1.96 & 0.11 & 1.00\\
+Null\_Late1vsNative & -13.40 & 10.69 & -40.57 & -0.41 & 1.01\\
+Null\_Late2vsLate1 & 14.22 & 10.69 & 1.29 & 41.29 & 1.01\\
+Null\_Intro & -21.98 & 14.25 & -58.10 & -4.70 & 1.01\\
+Null\_Maintain & 11.27 & 7.13 & 2.67 & 29.30 & 1.01\\
+Null\_L1vsN*Intro & -54.37 & 42.74 & -162.14 & -2.68 & 1.01\\
+Null\_L2vsL1*Intro & 55.46 & 42.75 & 3.99 & 164.16 & 1.01\\
+Null\_L1vsN*Maintain & 26.68 & 21.38 & 0.88 & 80.69 & 1.01\\
+Null\_L2vsL1*Maintain & -27.59 & 21.38 & -81.73 & -1.81 & 1.01\\
 \bottomrule
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-aksu-kouxe72015}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-aksu-koc2015}{}}%
 Aksu-Koç, Ayhan, Nicolopoulou, Ageliki, 2015. Character reference in
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-barberuxe02009}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-barbera2009}{}}%
 Barberà, Gemma, Massó, Guillem, 2009. Keeping track in sign language:
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-buxfcrkner2018}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-burkner2018}{}}%
 Bürkner, Paul-Christian, 2018. Advanced bayesian multilevel modeling
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-guxfcr2018}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-gur2018}{}}%
 Gür, Cansu, 2018. The effects of late language acquisition on the
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-karaduxf6ller2017}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-karadoller2017}{}}%
 Karadöller, Dilay Z, Sümer, Beyza, Özyürek, Aslı, 2017. Effects of
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kayaba2022}{}}%
-Kayabaşı, Demet, Gökgöz, Kadir, 2022. The causative-inchoative
-alternation and age-of-acquisition effects on multi-predicate
-constructions in turkish sign language. Language Learning and
-Development 0, 1--26.
-doi:\href{https://doi.org/10.1080/15475441.2022.2047690}{10.1080/15475441.2022.2047690}
+\leavevmode\vadjust pre{\hypertarget{ref-kayabasi2020}{}}%
+Kayabaşı, Demet, Sevgi, Hande, Özsoy, A. Sumru, 2020. Null arguments in
+turkish sign language. Morphological Complexity within and across
+Boundaries: In honour of Asli Göksel 215.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kubu2008}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-kubus2008}{}}%
 Kubuş, Okan, 2008. An analysis of Turkish Sign Language (T{İ}D)
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-sevinuxe72006}{}}%
+\leavevmode\vadjust pre{\hypertarget{ref-sevinc2006}{}}%
 Sevinç, Ayça Müge, 2006. Grammatical relations and word order in turkish

### 2022-08-06 revised_manuscript.Rmd
updated files

--- 2022-08-05 manuscript_v2.Rmd
@@
 First language deprivation or delay has been found to have negative effects on certain cognitive and linguistic measures among late signing children.
-To name a few, delayed exposure to a first functional language is reported to negatively affect deaf children's morphosyntactic production and processing [@kayabasi2020; @boudreault2006; @newport1990; @cheng2021; @karadoller2017; @cormier2012] and narrative abilities [e.g., @becker2009].
+To name a few, delayed exposure to a first functional language is reported to negatively affect deaf children and adults' morphosyntactic production and processing [@kayabasi2020; @boudreault2006; @newport1990; @cheng2021; @karadoller2017; @cormier2012] and narrative abilities [e.g., @becker2009].
 In our overall research, we are concerned with the effects of this late exposure in the ultimate adult forms of TİD users.
@@
 That is to say, individuals must make the most informative and succinct choice possible to retain economy of form.
-When multiple expressions are available for the same referent as in an utterance like (\ref{ex:1}), phonetically simplest form is selected with the least redundant information are selected based on an economy principle.
+When multiple expressions are available for the same referent as in an utterance like (\ref{ex:1}), phonetically simplest form is selected with the least redundant information based on an economy principle.
 In the utterance, REs for "Eddie" are given in bold.
@@
 This referential locus can later be used by agreement verbs and classifier predicates (more below).
-For instance, in (\ref{ex:2a}) there are two referents, \textsc{elvan}, who is referentially-assigned to the right side of the signer, represented with the "a" subscript and \textsc{demet}, who is referentially-assigned to the left side of the signer, represented with the "b" subscript.
+For instance, in (\ref{ex:2a}) there are two referents, \textsc{elvan}, who is referentially-assigned to the right side of the signer, represented with the "A" subscript and \textsc{demet}, who is referentially-assigned to the left side of the signer, represented with the "B" subscript.
 When the signer makes a contrast in (\ref{ex:2b}), as a pronominal use, she points to the right side to refer to ELVAN and to the left to refer to \textsc{demet}.
-We schematically represents the referential locus assigned to the referents in (\ref{ex:3}).
+We schematically represent the referential locus assigned to the referents in (\ref{ex:3}).
 
@@
 These two infused morphemes form a complex predicate.
-Some classifier handshapes and the set of objects that they denote are below.
 Some classifier handshapes and the set of objects that they denote are presented below in (\ref{ex:4}).
@@
 In (\ref{ex:5}), in the first clause, a referential locus, i.e.
-locus-a, is assigned to the locative source argument \textsc{cage}.
+locus-A, is assigned to the locative source argument \textsc{cage}.
 In the second clause, the core argument \textsc{bird} is cross-referenced with the classifier handshape, i.e.
 small-legged animal.
-It flies from the referential locus of the \textsc{cage} (locus-a) to the referential locus of the goal \textsc{tree} (locus-b).
+It flies from the referential locus of the \textsc{cage} (locus-a) to the referential locus of the goal \textsc{tree} (locus-B).
 
@@
 These are manually signed iconic representations of animate or inanimate objects, and they embody several properties pertaining to the nominal referent including its location in the signing space, movement, path, size, and shape [@talmy2003].
-As in most sign languages hitherto studied, classifier constructions are also common and found in TİD [see @dikyuva2015 for a review].
+As in most sign languages hitherto studied, classifier constructions are also commonly found in TİD [see @dikyuva2015 for a review].
 Following the classification of @engberg-pedersen1993, we distinguish between four types of classifier constructions.
 Whole entity classifiers (WECL) represent the entirety of any referent whereas body part classifiers (BPCL) only reflect a specific part of an animate entity (e.g., limbs, head, tongue, teeth, and so forth).
-Handling classifiers (HCL) occur with transitive verbs and depict how an object is handled or manipulated by agent.
+Handling classifiers (HCL) occur with transitive verbs and depict how an object is handled or manipulated by an agent.
 On this ground, an HCL[^1] depicts the manner of the referring action while the former two constructions encode the location of the entity as well as its motion in space.
@@
 Like in first language development, narrative skills have been extensively studied among second language learners [see @bel2015; @frederiksen2019].
-Comparing 11 deaf adults signers and 13 hearing advanced learners of Catalan Sign Language (Llengua de Signes Catalana - LSC), @bel2015 found that hearing signers resort to more overt forms (nominals and pronouns) than native signers to avoid potential ambiguities among referents.
-Nevertheless, their results did not reach statistical significant for the interaction of discourse status and acquisition group.
+Comparing 11 deaf adult signers and 13 hearing advanced learners of Catalan Sign Language (Llengua de Signes Catalana - LSC), @bel2015 found that hearing signers resort to more overt forms (nominals and pronouns) than native signers to avoid potential ambiguities among referents.
+Nevertheless, their results did not reach statistical significance for the interaction of discourse status and acquisition group.
 A similar line of research came from @frederiksen2019 who found a statistically significant difference between both groups, that is native signers and hearing L2 learners of ASL, only for classifier use especially in re-introduced contexts.
@@
 
-Still, only very few research examined whether linguistically deprived signers (i.e., deaf signers with delayed exposure to a sign language) also display redundancy in their reference tracking.
-@becker2009 examined 6 DGS signing deaf children (3 native and 3 late)'s narrative abilities.
+Still, only very few researchers examined whether linguistically deprived signers (i.e., deaf signers with delayed exposure to a sign language) also display redundancy in their reference tracking.
+@becker2009 examined 6 DGS signing deaf children's (3 native and 3 late) narrative abilities.
 She reported that late signing children generated more explicit forms such as nominals when referring to the characters in the story compared to native learners, concluding that first language deprivation results in an inadequate representation of the addressee and less efficient use of morphosyntax.
-Comparably, @cormier2013 investigated the use of CA in 15 BSL among deaf native and late adult signers and found that late signers were redundant in topic continuity.
+Comparably, @cormier2013 investigated the use of CA in BSL among 15 deaf native, early, and late adult signers (5 participants in each group) and found that late signers were redundant in topic continuity.
 In other words, the late signers produced more explicit forms like nominals, and fewer instances of null marking when maintaining previous referents.
@@
 For the re-introduction contexts, where referents which are discontinued across one or more clauses are brought back into the discourse, we expect a less clear picture.
-Some re-introduced referents will have higher accessibility and will thus be tracked with more implicit REs, while others that have lower accessibility will be tracked with more explicit REs Also based on previous research, we hypothesize that late signers will become more over-redundant in their reference tracking compared to native signers.
+Some re-introduced referents will have higher accessibility and will thus be tracked with more implicit REs, while others that have lower accessibility will be tracked with more explicit REs.
+Also, based on previous research, we hypothesize that late signers will be more over-redundant in their reference tracking compared to native signers.
 This would translate to a preference by late signers for nominals over null marking especially in maintenance and re-introduction contexts.
@@
 Although we did not set a cut-off point for nativeness, all the deaf of hearing participants indicated that they learned TİD after age 3.
-In order to capture in-group variation among late signers, we run an additional supplementary analysis}.
+In order to capture in-group variation among late signers, we run an additional supplementary analysis in which we observed very wide credible intervals due to the small group size, rendering the results uninformative.
+For the analysis, see \ref{section:d}}.
 All participants indicated that their preferred language of communication was TİD which they used for most contexts.
@@
 Except for two signers who became deaf between the ages of 4 and 7, all the other late signers were pre-lingually deaf.
-All late signers completed middle school education and the average year of schooling was 12.3 years (*SD* = 2.8 years).
+All late signers completed at least middle school education and the year of schooling ranged between 8 and 16 years, and the average was 12.3 years (*SD* = 2.8 years).
 All late participants attended at least one deaf school.
-While 8 participants attended two deaf schools (either primary and middle school, or middle and high school), 3 participants reported attend three deaf schools.
-Using the same scale, late signers self-rated their abilities in TİD and (spoken) Turkish.
+While 8 participants attended two deaf schools (either primary and middle school, or middle and high school), 3 participants reported to attend three deaf schools.
+Using the same scale, late signers self-rated their abilities in TİD and Turkish.
 Both of their receptive and productive TİD abilities rated on average 4.9 points (*SD* = 0.3 points).
@@
 All trials were recorded by a digital camera placed in front of the signers.
-The participants were instructed to watch the narrative clip as many times as they needed, and narrate the content of each clip either to the experimenter or the recording digital camera\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining e did not make it obligatory for participants to either narrate to the digital camera or to the experimenter for few reasons.
+The participants were instructed to watch the narrative clip as many times as they needed, and narrate the content of each clip either to the experimenter or the recording digital camera\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining we did not make it obligatory for participants to either narrate to the digital camera or to the experimenter for few reasons.
 First, a native or native-like deaf experimenter was not always available, and a hearing late signer of TİD attended some sessions as the experimenter.
@@
 
-# Summary of the data
-summary_df <- df_narration %>% 
-  group_by(Nativeness, ref_type) %>% 
-  summarise(Count = n()) %>% 
-  mutate(Percentage = formattable::percent(Count/sum(Count))) %>% 
-  arrange(desc(Percentage))
-
 # Descriptive plots
@@
 Each data point is a referent tracked by the participant.
-Each of the REs is classified in terms of its Discourse Status (Introduction, Maintenance, Re-introduction), its Type (Nominal, Pronominal, Classifer, Null)\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer Null).
-However, we allude to the numerical distribution of the subtypes in this and following sections}, and Participant Group (Native, Late)\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17).
-Treating age of acquisition as a continous variable would not be informative due to the small sample size in each interval.}.
+Each of the REs is classified in terms of its Discourse Status (Introduction, Maintenance, Re-introduction), its Type (Nominal, Pronominal, Classifer, Null)\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer, and Null).
+However, we allude to the numerical distribution of the subtypes in this and following sections.}, and Participant Group (Native, Late)\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17).}.
 
@@
 Additionally, signers used CA more for referent re-introduction than maintenance.
-We observed a similar instances of agreement verbs for maintained and re-introduced contexts.
+We observed similar instances of agreement verbs for maintained and re-introduced contexts.
 In contrast, plain verbs were favored more in the maintenance setting than re-introduction.
@@
 ```
-
-```{r count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), RE Type (Nominal, Pronominal, Classifier, Null), and Participant Group (Native, Late)"}
+```{r fig:count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), RE Type (Nominal, Pronominal, Classifier, Null), and Participant Group (Native, Late)"}
 df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "NULL"))
@@
 For the current analysis, we did not define priors ourselves and used the preset uninformed priors.
-\@ref{fig:model-gram}[^2] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals.
+Figure 2[^2] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals.
 There are 3 responses we are considering in this model.
@@
 [^2]: We ran another model where we used the Nativeness predictor with 3 levels 0-3:Native, 4-7:Late1, and 7-above:Late2.
-However, our model did not converge properly.
-We provide the model results as a plot and table in the \ref{section:d}.
-
-```{r model-gram, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
+However, our model did not converge properly and the credible intervals were very wide spread, yielding the results uninformative.
+We provide the model results as a plot in \ref{section:d}.
+
+```{r model-gram, echo=FALSE, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, message=FALSE, warning=FALSE, cache=FALSE, results='asis', warnings=FALSE}
 p3 <- df_model_1 %>% 
@@
 
-# General Discussion
+# Discussion
 
@@
 ```
-
 # Episodes in the stimuli
@@
   geom_errorbarh(aes(xmin = ll, xmax = hh), height=0, position = position_dodge(width = 0.6)) + vline_0(linetype = "dashed") +
-  xlab("Estimate (log odds)") + ylab("Coefficients") + theme_apa()
+  xlab("Estimate (log odds)") + ylab("Coefficients") + theme_bw() + theme_apa() +
+  scale_color_grey() +
+  scale_fill_grey() 
 p4
@@
 library(janitor)
-df <- summary(model_3) %>% .$fixed %>% as_tibble()
+df <- summary(model_3) %>% .$fixed
+df <- as.data.frame(df)
 df_table_extra <- df %>%  clean_names() %>% rownames_to_column("Coefficients") %>% dplyr::select(-7,-8)
+df_table_extra <- df %>% clean_names() %>% rownames_to_column("Coefficients") %>% dplyr::select(-7,-8)
 
@@
 
-df_table_extra$Coefficients %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
+df_table_extra$Coefficients %<>% dplyr::recode(`muNominal_Intercept` = "Nom_Intercept",
                                         `muNull_Intercept` = "Null_Intercept",

### 2022-08-06 revised_manuscript.tex
updated files

--- 2022-08-06 manuscript_v2.tex
@@
 Null)\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer, and Null).
-However, we allude to the numerical distribution of the subtypes in this and following sections},
+However, we allude to the numerical distribution of the subtypes in this and following sections.},
 and Participant Group (Native,
@@
 \centering
-\includegraphics{manuscript_v2_files/figure-latex/count-plot-1.pdf}
+\includegraphics{revised_manuscript_files/figure-latex/fig:count-plot-1.pdf}
 \caption{Referent counts by Discourse Status (Introduction, Maintanence,
@@
 from the following analyses. For the current analysis, we did not define
-priors ourselves and used the preset uninformed priors.
-@ref\{fig:model-gram\}\footnote{We ran another model where we used the
-  Nativeness predictor with 3 levels 0-3:Native, 4-7:Late1, and
-  7-above:Late2. However, our model did not converge properly and the
-  credible intervals were very wide spread, yielding the results
-  uninformative. We provide the model results as a plot in
-  \ref{section:d}.} shows the posterior probability distributions where
-the dot is the median estimate, the thick line is the 50\% credible
-intervals, and the thin line is the 95\% credible intervals. There are 3
-responses we are considering in this model. They are; Classifier,
-Nominative, and Null. Classifier response is the level that other
-responses Nominative and Null are compared against.
-
-\begin{verbatim}
-## Warning: Ignoring unknown aesthetics: height
-\end{verbatim}
+priors ourselves and used the preset uninformed priors. Figure
+2\footnote{We ran another model where we used the Nativeness predictor
+  with 3 levels 0-3:Native, 4-7:Late1, and 7-above:Late2. However, our
+  model did not converge properly and the credible intervals were very
+  wide spread, yielding the results uninformative. We provide the model
+  results as a plot in \ref{section:d}.} shows the posterior probability
+distributions where the dot is the median estimate, the thick line is
+the 50\% credible intervals, and the thin line is the 95\% credible
+intervals. There are 3 responses we are considering in this model. They
+are; Classifier, Nominative, and Null. Classifier response is the level
+that other responses Nominative and Null are compared against.
 
@@
 \centering
-\includegraphics{manuscript_v2_files/figure-latex/model-gram-1.pdf}
+\includegraphics{revised_manuscript_files/figure-latex/model-gram-1.pdf}
 \caption{Regression model plot for the RE Type (Nominal, Null) with the
@@
 \centering
-\includegraphics{manuscript_v2_files/figure-latex/model-camera-1.pdf}
+\includegraphics{revised_manuscript_files/figure-latex/model-camera-1.pdf}
 \caption{Regression model plot for the RE Type (Nominal, Null) with the

### 2022-08-06 revised_manuscript.Rmd
updated files

@@
 There are 3 responses we are considering in this model.
-They are; Classifier, Nominative, and Null.
+They are Classifier, Nominative, and Null.
 Classifier response is the level that other responses Nominative and Null are compared against.
@@
 [^3]: As in other sign languages, in TİD, the lexical meaning of signs has an effect on the morphological grouping.
-For instance, usually (di)transitive agreement verbs denote some sort of physical or metaphorical transfer of an object or abstract notion (Meir 1999).
+For instance, usually (di)transitive agreement verbs denote some sort of physical or metaphorical transfer of an object or abstract notion [@meir1999].
 As for plain verbs, both phonology of a sign and its semantics have an effect on morphology.
@@
 These generalizations noted we should also be cautious in labelling a verb as categorially one type or the other.
-For instance, some intransitive plain verbs like \textsc{fall} or \textsc{jump} in TİD may display spatial person agreement with the single argument in the clause (Gökgöz and Sevgi, 2020).
-See also for a similar discussion in Brazilian Sign Language.}.
+For instance, some intransitive plain verbs like \textsc{fall} or \textsc{jump} in TİD may display spatial person agreement with the single argument in the clause [@gokgoz2020].
+See @lourenço2018 also for a similar discussion in Brazilian Sign Language.}.
 
@@
 
-```{r model-camera, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=3.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
+```{r model-camera, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction), Participant Group (Native, Late) and Narration Style (all responses vs.
+only camera)", fig.height=3.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
 df_model_1 %<>% mutate(model = "All")
@@
 p4
-```
+
+```
+
+*Note:* "All" indicates model results including all participants (both camera and experimenter) whereas "camera" indicates the results from only those who narrated to the camera.
+That is, the 4 signers who narrated to the experimenter were from analysis in the second dataset.
 

### 2022-08-06 revised_manuscript.tex
updated files

@@
 intervals. There are 3 responses we are considering in this model. They
-are; Classifier, Nominative, and Null. Classifier response is the level
+are Classifier, Nominative, and Null. Classifier response is the level
 that other responses Nominative and Null are compared against.
@@
   verbs denote some sort of physical or metaphorical transfer of an
-  object or abstract notion (Meir 1999). As for plain verbs, both
+  object or abstract notion (Meir, 1999). As for plain verbs, both
   phonology of a sign and its semantics have an effect on morphology.
@@
   or \textsc{jump} in TİD may display spatial person agreement with the
-  single argument in the clause (Gökgöz and Sevgi, 2020). See also for a
-  similar discussion in Brazilian Sign Language.\}.}. Agreement verbs
-and CA were the following implicit markers of reference, 0.77 and 0.72
-of which were used for maintained contexts, respectively. This contrasts
-with some of the research (Czubek, 2017) in which agreement verbs ranked
-lowest on scale, indicating that these constructions were more implicit
-compared to plain verbs. However, there seems to be a common ground for
-the position of CA markers which have been claimed to rank higher than
-other null markers as in the present study (Barberà and Massó, 2009;
-Czubek, 2017; Frederiksen and Mayberry, 2016). In Figure
-\ref{fig:scale}, we report the distribution of REs in TİD as a function
-of referent accessibility and compare our accessibility hierarchy with
-that of (Frederiksen and Mayberry, 2016) for ASL. In the hierarchies,
-the left most REs (e.g., modified and bare nouns for TİD) are used when
-the accessibility is low. If the referent is already active in the
-addressee's mind and its accessibility is high, then the right most
-markers (e.g., agreement and plain verbs for TİD) are selected.
+  single argument in the clause (Gökgöz and Sevgi, 2020). See Lourenço
+  and Wilbur (2018) also for a similar discussion in Brazilian Sign
+  Language.\}.}. Agreement verbs and CA were the following implicit
+markers of reference, 0.77 and 0.72 of which were used for maintained
+contexts, respectively. This contrasts with some of the research
+(Czubek, 2017) in which agreement verbs ranked lowest on scale,
+indicating that these constructions were more implicit compared to plain
+verbs. However, there seems to be a common ground for the position of CA
+markers which have been claimed to rank higher than other null markers
+as in the present study (Barberà and Massó, 2009; Czubek, 2017;
+Frederiksen and Mayberry, 2016). In Figure \ref{fig:scale}, we report
+the distribution of REs in TİD as a function of referent accessibility
+and compare our accessibility hierarchy with that of (Frederiksen and
+Mayberry, 2016) for ASL. In the hierarchies, the left most REs (e.g.,
+modified and bare nouns for TİD) are used when the accessibility is low.
+If the referent is already active in the addressee's mind and its
+accessibility is high, then the right most markers (e.g., agreement and
+plain verbs for TİD) are selected.
 
@@
 \caption{Regression model plot for the RE Type (Nominal, Null) with the
-predictors Discourse Status (Introduction, Maintenance, Re-introduction)
-and Participant Group (Native, Late)}
+predictors Discourse Status (Introduction, Maintenance,
+Re-introduction), Participant Group (Native, Late) and Narration Style
+(all responses vs.~only camera)}
 \end{figure}
+
+\emph{Note:} ``All'' indicates model results including all participants
+(both camera and experimenter) whereas ``camera'' indicates the results
+from only those who narrated to the camera. That is, the 4 signers who
+narrated to the experimenter were from analysis in the second dataset.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-gokgoz2020}{}}%
+Gökgöz, Kadir, Sevgi, Hande, 2020. Aspects of clause structure and
+morphology in turkish sign language. Morphological Complexity within and
+across Boundaries: In Honour of Asli Göksel 215, 315.
+
 \leavevmode\vadjust pre{\hypertarget{ref-grice1975}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-lourenuxe7o2018}{}}%
+Lourenço, Guilherme, Wilbur, Ronnie B., 2018. Are plain verbs really
+plain?: Co-localization as the agreement marker in sign languages.
+FEAST. Formal and Experimental Advances in Sign language Theory 2, 6881.
+
 \leavevmode\vadjust pre{\hypertarget{ref-maclaughlin1997}{}}%
@@
 doi:\href{https://doi.org/10.2307/415277}{10.2307/415277}
+
+\leavevmode\vadjust pre{\hypertarget{ref-meir1999}{}}%
+Meir, Irit, 1999. A perfect marker in israeli sign language. Sign
+Language \& Linguistics 2, 4362.
 

### 2022-08-06 revised_manuscript.Rmd
updated files

@@
 
-In our first set of hypotheses regarding referent tracking strategies, we had the general expectation that signers would use more explicit forms (mainly nominals) for referent introduction and less implicit forms such as null marking, pronominal use, and classifiers.
+In our first set of hypotheses regarding referent tracking strategies, we had the general expectation that signers would use more explicit forms (mainly nominals) for referent introduction and less implicit forms such as null marking, pronominal use, and classifiers for maintenance.
 In re-introduced contexts, we predicted that signers would use both explicit and implicit markers depending on the accessibility of the tracked referent.
@@
 For age of acquisition effects, based on what is proposed for child and adult deaf signers with language deprivation in terms of their morphosyntactic [@boudreault2006; @cormier2012; @kayabasi2020] and narrative development [@becker2009; @cormier2013], we expected that late signers would be more over-redundant in their referent tracking, akin to second language learners [@frederiksen2019; @bel2015].
-That is, we had the expectation that late signers would use more explicit markers like nouns and fewer implicit markers like null marking even for highly accessible referents (when the discourse status is maintained or even re-introduced in same cases) than the control group (i.e., native signers).
+That is, we had the expectation that late signers would use more explicit markers like nouns and fewer implicit markers like null marking even for highly accessible referents (when the discourse status is maintained or even re-introduced in same cases) than native signers.
 
@@
 Investigating the distribution of REs across discourse statuses in ASL, a similar study by @czubek2017 also reported contrasting numbers (17% for introduction, 64% for maintenance, and 19% for re-introduction).
-We suggest that the discrepancy might be due to the nature of the stimuli used in these research.
-In @frederiksen2016, the authors used 4-6-episode-long picture story with 2 animate characters and 1 inanimate object.
+We suggest that the discrepancy might be due to the nature of the stimuli.
+In @frederiksen2016, the authors used a 4-6-episode-long picture story with 2 animate characters and 1 inanimate object.
 In contrast, the stimuli in the present study consisted of 8-episode-long video clips on average with a mean number of 2 animate characters and 3 inanimate objects.
@@
 
-As for the distribution of REs across different discourses statuses, we observed that signers overwhelmingly preferred to use nominals over other REs for the first mention of a referent and used null markers such as CA, and agreement and plain verb for referents with high accessibility (i.e., when the discourse status was maintained).
-This finding parallels other accounts of referential accessibility suggested for other sign languages [see @frederiksen2016; @czubek2017; and @swabey2002; for ASL, @morgan2005; for BSL @bel2015; and @barbera2009 for LSC.
+As for the distribution of REs across different discourses statuses, we observed that signers overwhelmingly preferred to use nominals over other REs for the first mention of a referent and used null markers such as CA, agreement and plain verb for referents with high accessibility (i.e., when the discourse status was maintained).
+This finding parallels other accounts of referential accessibility suggested for other sign languages (see @frederiksen2016; @czubek2017; and @swabey2002; for ASL, @morgan2005; for BSL @bel2015; and @barbera2009 for LSC).
 Our results also indicate that modified nouns (e.g., nouns that are preceded or followed by a classifier or IX) appeared more in introduced contexts compared to re-introduced contexts.
@@
 This finding is comparable with what has been reported for ASL [@swabey2002; @frederiksen2016; @czubek2017] but not with LSC [@bel2015], BSL [@morgan2005] and DGS [@perniss2015], in which the signers extensively made overt pronominal references (up to 30%) to maintain entities.
-Comparable to our findings, It has been reported that pronouns do not frequently occur in the local context and instead follows a topic shift and thus rank higher than classifiers and null markers different genre [see @barbera2009 for LSC; and @nuhbalaoglu2018 for TİD].
+Comparable to our findings, it has been reported that pronouns do not frequently occur in the local context and instead follows a topic shift and thus rank higher than classifiers and null markers [see @barbera2009 for LSC; and @nuhbalaoglu2018 for TİD].
 Considering this, the claim that the IX sign functions like demonstratives that are also used to track referents in re-introduced contexts has not been borne out [@koulidobrova2016].
@@
 The observation that overt pronouns are rare in ASL narratives comes from studies that include a simple picture retelling task with one protagonist [e.g., @frederiksen2016] as well as a more complex narrative design as in [@czubek2017].
-However, @czubek2017's study also incorporated different genres and reported more instances of overt pronouns in personal narratives (compared to picture or video retelling).
+However, @czubek2017 also incorporated different genres and reported more instances of overt pronouns in personal narratives (compared to picture or video retelling).
 In fact, this finding suggests that the use of pronouns might be genre-dependent.
@@
 A deeper look into the allocation of classifier constructions revealed that WECLs were the most common type of classifiers for the maintenance of a referent followed by BPCLs and ExtCLs.
-We observed a similar alignment for re-introduced contexts.
+We observed a similar distibution for re-introduced contexts.
 In contrast, a great number of the overall ExtCLs were used to introduce a new referent into the discourse.
-That is to say, whereas WECLs were the preferred classifier types to signal highly accessible referents, ExtCLs constructions were mainly used for referents which were not easily accessible.
+That is to say, whereas WECLs were the preferred classifier types to signal highly accessible referents, ExtCL constructions were mainly used for referents which were not easily accessible.
 We suggest that the less iconic and transparent nature of WECLs allowed signers to continue a previous referent without being redundant.
 As pointed out by @czubek2017, signers prefer to rely on ExtCLs for the first mention of an entity because these constructions are informative and explicit by nature, and they provide visual and descriptive cues about an entity and its location in space.
-Nevertheless, we only observed `r sum(df_narration$GTR1_plot == "SASS" & df_narration$Discourse == "Re-introduction")` instances of ExtCLs for re-introduced contexts, replicating @frederiksen2019's findings for ASL.
+We only observed `r sum(df_narration$GTR1_plot == "SASS" & df_narration$Discourse == "Re-introduction")` instances of ExtCLs for re-introduced contexts, replicating @frederiksen2019's findings for ASL.
 
@@
 In Figure \ref{fig:scale}, we report the distribution of REs in TİD as a function of referent accessibility and compare our accessibility hierarchy with that of [@frederiksen2016] for ASL.
-In the hierarchies, the left most REs (e.g., modified and bare nouns for TİD) are used when the accessibility is low.
-If the referent is already active in the addressee's mind and its accessibility is high, then the right most markers (e.g., agreement and plain verbs for TİD) are selected.
+In the hierarchies, the leftmost REs (e.g., modified and bare nouns for TİD) are used when the accessibility is low.
+If the referent is already active in the addressee's mind and its accessibility is high, then the rightmost markers (e.g., agreement and plain verbs for TİD) are selected.
 
@@
 \begin{figure}
-\caption{Comparing the present study and Frederikse and Mayberry (2016) in terms of proposed accessibility hierarchies}
+\caption{Comparing the present study and Frederiksen and Mayberry (2016) in terms of proposed accessibility hierarchies}
 \label{fig:scale}
@@
 The results in this study only somewhat support our second set of hypotheses that we set for the effects of age of acquisition on participants' RE selection.
-That is, while being a native acquirer of TİD increased the odds of using null marking in reference tracking across different discourse statuses, we failed to find such a difference for overall nominal production.
+That is, while being a native acquirer of TİD increased the odds of using null marking in reference tracking across different discourse statuses, we failed to find a difference for overall nominal production.
 Taking a closer look at the interaction terms, we observed that native signers, compared to late signers, were less likely to use nominal constructions when the discourse status was maintained.
@@
 Given the paucity of research directed at the late signing group, we now turn to the referent tracking strategies observed for second language learners of a sign language.
-Our study replicated some of the findings of the line of research that investigated deaf or hearing bimodal bilingual signers [@bel2015; @frederiksen2019].\footnote{For similar results with heritage learners of sign language, see Reynolds (2018).} The authors reported that more instances of null marking in the context of referent introduction was attested for native signers than deaf or hearing bimodal bilingual signers.
+Our study replicated some of the findings of the line of research that investigated second language learners of a sign language [@bel2015; @frederiksen2019].
+The authors reported that more instances of null marking in the context of referent introduction was attested for native signers than second language learners.
 Although using null marking to introduce a referent into the discourse would normally result in referential ambiguity, it is important to note that our stimuli consisted of 10 interconnected narratives with 2 or more recurring characters.
@@
 Then, our results imply a difference in pragmatic sensitivity to referential accessibility between native and late signers, which is more robustly reflected in introduced contexts.
+To test this implication, one could use the same number and length of stimuli from different cartoons with different characters and see if the tendency to introduce a referent without an overt nominal for the first time is dependent on the task or not.
 
@@
 
-The present study aimed to investigate the reference tracking strategies of native and late deaf adult signers in TİD narratives.
+The present study aimed to investigate the reference tracking strategies in TİD narratives and whether native and late deaf TİD signers differ from each other in these strategies.
 In line with our expectations, we found that the distribution of REs (nominal, pronominal, classifier, and null marking) followed from previous accounts of saliency and referential accessibility.
-For introduced contexts, signers produced more nominals and ExtCLs, and to maintain a previous referent, hey mainly used null arguments (CA, and plain and agreement verbs) and certain classifiers (e.g., BPCL and WECL) to track highly accessible referents.
+For introduced contexts, signers produced more nominals and ExtCLs, and to maintain a previous referent, they mainly used null arguments (in clauses with CA, agreement and plain verbs) and certain classifiers (e.g., BPCL and WECL) to track highly accessible referents.
 We observed only very few instances of pronouns (mostly produced by late signers), unlike our expectations.
@@
 Native signers were less likely to use a nominal for maintained contexts and more likely to use null marking for introduced contexts than late signers.
-We conclude that native signers might only partially adhere to pragmatic norms and leverage accessibility better than late signers since they use less form when the referent is more active in the mind.
+Following @swabey2002's intituition, we conclude that native signers must adhere to pragmatic norms and leverage accessibility better than late signers only to a certain extent (given our finding of limited overexplicitness for late signers) since they use less form when the referent is more active in the mind.
 

### 2022-08-06 revised_manuscript.tex
updated files

@@
 forms (mainly nominals) for referent introduction and less implicit
-forms such as null marking, pronominal use, and classifiers. In
-re-introduced contexts, we predicted that signers would use both
-explicit and implicit markers depending on the accessibility of the
-tracked referent.
+forms such as null marking, pronominal use, and classifiers for
+maintenance. In re-introduced contexts, we predicted that signers would
+use both explicit and implicit markers depending on the accessibility of
+the tracked referent.
 
@@
 accessible referents (when the discourse status is maintained or even
-re-introduced in same cases) than the control group (i.e., native
-signers).
+re-introduced in same cases) than native signers.
 
@@
 We suggest that the discrepancy might be due to the nature of the
-stimuli used in these research. In Frederiksen and Mayberry (2016), the
-authors used 4-6-episode-long picture story with 2 animate characters
-and 1 inanimate object. In contrast, the stimuli in the present study
-consisted of 8-episode-long video clips on average with a mean number of
-2 animate characters and 3 inanimate objects. Although the number of
-animate characters and inanimate characters in Czubek (2017)'s stimuli
-matched that of the present study, their narrative structure (a
-6-episode-long picture story) was still shorter compared to our
-narrative stimuli. We believe fewer possible referents and/or a simpler
-narrative structure in these studies may have contributed to the lower
-number of re-introduced contexts compared to the present study. Due to
-the story length and composition in this study, the participants had to
-track more recurring animate or inanimate referents for a longer period
-of time. As a result, our signers engaged in referent re-introduction
-more than what was reported in previous studies (Czubek, 2017;
-Frederiksen and Mayberry, 2016).
+stimuli. In Frederiksen and Mayberry (2016), the authors used a
+4-6-episode-long picture story with 2 animate characters and 1 inanimate
+object. In contrast, the stimuli in the present study consisted of
+8-episode-long video clips on average with a mean number of 2 animate
+characters and 3 inanimate objects. Although the number of animate
+characters and inanimate characters in Czubek (2017)'s stimuli matched
+that of the present study, their narrative structure (a 6-episode-long
+picture story) was still shorter compared to our narrative stimuli. We
+believe fewer possible referents and/or a simpler narrative structure in
+these studies may have contributed to the lower number of re-introduced
+contexts compared to the present study. Due to the story length and
+composition in this study, the participants had to track more recurring
+animate or inanimate referents for a longer period of time. As a result,
+our signers engaged in referent re-introduction more than what was
+reported in previous studies (Czubek, 2017; Frederiksen and Mayberry,
+2016).
 
@@
 other REs for the first mention of a referent and used null markers such
-as CA, and agreement and plain verb for referents with high
-accessibility (i.e., when the discourse status was maintained). This
-finding parallels other accounts of referential accessibility suggested
-for other sign languages {[}see Frederiksen and Mayberry (2016); Czubek
-(2017); and Swabey (2002); for ASL, Morgan (2005); for BSL Bel et al.
-(2015); and Barberà and Massó (2009) for LSC. Our results also indicate
-that modified nouns (e.g., nouns that are preceded or followed by a
+as CA, agreement and plain verb for referents with high accessibility
+(i.e., when the discourse status was maintained). This finding parallels
+other accounts of referential accessibility suggested for other sign
+languages (see Frederiksen and Mayberry (2016); Czubek (2017); and
+Swabey (2002); for ASL, Morgan (2005); for BSL Bel et al. (2015); and
+Barberà and Massó (2009) for LSC). Our results also indicate that
+modified nouns (e.g., nouns that are preceded or followed by a
 classifier or IX) appeared more in introduced contexts compared to
@@
 2015), in which the signers extensively made overt pronominal references
-(up to 30\%) to maintain entities. Comparable to our findings, It has
+(up to 30\%) to maintain entities. Comparable to our findings, it has
 been reported that pronouns do not frequently occur in the local context
 and instead follows a topic shift and thus rank higher than classifiers
-and null markers different genre (see Barberà and Massó, 2009 for LSC;
-and Nuhbalaoglu, 2018 for TİD). Considering this, the claim that the IX
-sign functions like demonstratives that are also used to track referents
-in re-introduced contexts has not been borne out (Koulidobrova and
+and null markers (see Barberà and Massó, 2009 for LSC; and Nuhbalaoglu,
+2018 for TİD). Considering this, the claim that the IX sign functions
+like demonstratives that are also used to track referents in
+re-introduced contexts has not been borne out (Koulidobrova and
 Lillo-Martin, 2016). Similar to these research, our findings indicate
@@
 protagonist (e.g., Frederiksen and Mayberry, 2016) as well as a more
-complex narrative design as in (Czubek, 2017). However, Czubek (2017)'s
-study also incorporated different genres and reported more instances of
-overt pronouns in personal narratives (compared to picture or video
+complex narrative design as in (Czubek, 2017). However, Czubek (2017)
+also incorporated different genres and reported more instances of overt
+pronouns in personal narratives (compared to picture or video
 retelling). In fact, this finding suggests that the use of pronouns
@@
 of a referent followed by BPCLs and ExtCLs. We observed a similar
-alignment for re-introduced contexts. In contrast, a great number of the
-overall ExtCLs were used to introduce a new referent into the discourse.
-That is to say, whereas WECLs were the preferred classifier types to
-signal highly accessible referents, ExtCLs constructions were mainly
-used for referents which were not easily accessible. We suggest that the
-less iconic and transparent nature of WECLs allowed signers to continue
-a previous referent without being redundant. As pointed out by Czubek
-(2017), signers prefer to rely on ExtCLs for the first mention of an
-entity because these constructions are informative and explicit by
+distibution for re-introduced contexts. In contrast, a great number of
+the overall ExtCLs were used to introduce a new referent into the
+discourse. That is to say, whereas WECLs were the preferred classifier
+types to signal highly accessible referents, ExtCL constructions were
+mainly used for referents which were not easily accessible. We suggest
+that the less iconic and transparent nature of WECLs allowed signers to
+continue a previous referent without being redundant. As pointed out by
+Czubek (2017), signers prefer to rely on ExtCLs for the first mention of
+an entity because these constructions are informative and explicit by
 nature, and they provide visual and descriptive cues about an entity and
-its location in space. Nevertheless, we only observed 3 instances of
-ExtCLs for re-introduced contexts, replicating Frederiksen and Mayberry
-(2019)'s findings for ASL.
+its location in space. We only observed 3 instances of ExtCLs for
+re-introduced contexts, replicating Frederiksen and Mayberry (2019)'s
+findings for ASL.
 
@@
 and compare our accessibility hierarchy with that of (Frederiksen and
-Mayberry, 2016) for ASL. In the hierarchies, the left most REs (e.g.,
+Mayberry, 2016) for ASL. In the hierarchies, the leftmost REs (e.g.,
 modified and bare nouns for TİD) are used when the accessibility is low.
 If the referent is already active in the addressee's mind and its
-accessibility is high, then the right most markers (e.g., agreement and
+accessibility is high, then the rightmost markers (e.g., agreement and
 plain verbs for TİD) are selected.
@@
 TİD increased the odds of using null marking in reference tracking
-across different discourse statuses, we failed to find such a difference
-for overall nominal production. Taking a closer look at the interaction
+across different discourse statuses, we failed to find a difference for
+overall nominal production. Taking a closer look at the interaction
 terms, we observed that native signers, compared to late signers, were
@@
 learners of a sign language. Our study replicated some of the findings
-of the line of research that investigated deaf or hearing bimodal
-bilingual signers (Bel et al., 2015; Frederiksen and Mayberry,
-2019).\footnote{For similar results with heritage learners of sign language, see Reynolds (2018).}
-The authors reported that more instances of null marking in the context
-of referent introduction was attested for native signers than deaf or
-hearing bimodal bilingual signers. Although using null marking to
-introduce a referent into the discourse would normally result in
-referential ambiguity, it is important to note that our stimuli
-consisted of 10 interconnected narratives with 2 or more recurring
-characters. It may be possible that native signers, having already
-introduced the referent with an explicit nominal marker in a different
-discourse context, found it felicitous to refer to the same entity with
-more implicit markers in the following narratives even for introduced
-contexts. Then, our results imply a difference in pragmatic sensitivity
-to referential accessibility between native and late signers, which is
-more robustly reflected in introduced contexts.
+of the line of research that investigated second language learners of a
+sign language (Bel et al., 2015; Frederiksen and Mayberry, 2019). The
+authors reported that more instances of null marking in the context of
+referent introduction was attested for native signers than second
+language learners. Although using null marking to introduce a referent
+into the discourse would normally result in referential ambiguity, it is
+important to note that our stimuli consisted of 10 interconnected
+narratives with 2 or more recurring characters. It may be possible that
+native signers, having already introduced the referent with an explicit
+nominal marker in a different discourse context, found it felicitous to
+refer to the same entity with more implicit markers in the following
+narratives even for introduced contexts. Then, our results imply a
+difference in pragmatic sensitivity to referential accessibility between
+native and late signers, which is more robustly reflected in introduced
+contexts. To test this implication, one could use the same number and
+length of stimuli from different cartoons with different characters and
+see if the tendency to introduce a referent without an overt nominal for
+the first time is dependent on the task or not.
 
@@
 The present study aimed to investigate the reference tracking strategies
-of native and late deaf adult signers in TİD narratives. In line with
-our expectations, we found that the distribution of REs (nominal,
-pronominal, classifier, and null marking) followed from previous
-accounts of saliency and referential accessibility. For introduced
-contexts, signers produced more nominals and ExtCLs, and to maintain a
-previous referent, hey mainly used null arguments (CA, and plain and
-agreement verbs) and certain classifiers (e.g., BPCL and WECL) to track
-highly accessible referents. We observed only very few instances of
-pronouns (mostly produced by late signers), unlike our expectations. The
-ones we observed were mostly reserved for re-introduced contexts along
-with nominals. Based on our findings across the two groups, we proposed
-a hierarchy of referent accessibility for TİD. As for the age of
+in TİD narratives and whether native and late deaf TİD signers differ
+from each other in these strategies. In line with our expectations, we
+found that the distribution of REs (nominal, pronominal, classifier, and
+null marking) followed from previous accounts of saliency and
+referential accessibility. For introduced contexts, signers produced
+more nominals and ExtCLs, and to maintain a previous referent, they
+mainly used null arguments (in clauses with CA, agreement and plain
+verbs) and certain classifiers (e.g., BPCL and WECL) to track highly
+accessible referents. We observed only very few instances of pronouns
+(mostly produced by late signers), unlike our expectations. The ones we
+observed were mostly reserved for re-introduced contexts along with
+nominals. Based on our findings across the two groups, we proposed a
+hierarchy of referent accessibility for TİD. As for the age of
 acquisition effects, we observed limited over-redundancy for signers
@@
 and more likely to use null marking for introduced contexts than late
-signers. We conclude that native signers might only partially adhere to
-pragmatic norms and leverage accessibility better than late signers
-since they use less form when the referent is more active in the mind.
+signers. Following Swabey (2002)'s intituition, we conclude that native
+signers must adhere to pragmatic norms and leverage accessibility better
+than late signers only to a certain extent (given our finding of limited
+overexplicitness for late signers) since they use less form when the
+referent is more active in the mind.
 

### 2022-08-06 revised_manuscript.Rmd
small change

@@
 This does not mean that late signers used more null markers in maintenance contexts.
-It means that the use of null markers did not linearly increase as the addition of using a null marker instead of a classifier as a native plus using null markers instead of a classifier for maintaining a referent versus re-introducing it.
+It means that the use of null markers did not linearly increase as the sum of using a null marker as a native plus using null markers for maintaining a referent versus re-introducing it.
 

### 2022-08-06 revised_manuscript.Rmd
updated files

--- 2022-08-06 revised_manuscript.Rmd
@@
 times: true
-numbersections: true
+numbersections: trueand w
 classoption: "review"
@@
 That is why the age of acquisition of a first functional language varies greatly among deaf children, with many of whom being deprived of their first languages typically until after age 7 when they start formal education in a deaf school.
+Despite this, many of the early intervention programs for late signing children prioritize spoken language development and discourage the use of a sign language [@kemaloglu2012].
+@ministryofnationaleducation2020 identifies 72 deaf schools in Turkey for the K-12 level.
+Even when deaf children with hearing parents start their formal education in one of these schools, they do not receive proper instruction in TİD since the teachers themselves are mainly hearing and only have a very rudimentary knowledge of a sign language [@ilkbasaran2015].
 First language deprivation or delay has been found to have negative effects on certain cognitive and linguistic measures among late signing children.
@@
 However, this particular observation is not peculiar to spoken languages.
-There is an emerging body of literature that reports the distribution of REs in signed narratives.
+There is an emerging body of literature that reports the distribution of REs in signed narratives [@czubek2017; @swabey2002; @frederiksen2016; @perniss2015; @morgan2005; @hodge2019; @ahlgren1994].
 Building on @gundel1993's work, @swabey2002 demonstrated that, like other spoken languages such as Russian, Chinese, and Japanese, American Sign Language (ASL) signers use bare nouns for introduced and re-introduced discourse contexts whereas null marking is used for referents which are "in-focus" or very accessible to the addressee.
@@
 
-### Signed vs. Spoken REs
-
-Commonly in both spoken and sign languages nouns and noun phrases are used as REs.
-Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions [@lillo-martin1990].
-In addition, the pointing sign is used to referentially anchor a referent to a location in signing space.
-For instance, when the signer introduces \textsc{elvan} for the first time, s/he can assign her to the right.
-This referential locus can later be used by agreement verbs and classifier predicates (more below).
-For instance, in (\ref{ex:2a}) there are two referents, \textsc{elvan}, who is referentially-assigned to the right side of the signer, represented with the "A" subscript and \textsc{demet}, who is referentially-assigned to the left side of the signer, represented with the "B" subscript.
-When the signer makes a contrast in (\ref{ex:2b}), as a pronominal use, she points to the right side to refer to ELVAN and to the left to refer to \textsc{demet}.
-We schematically represent the referential locus assigned to the referents in (\ref{ex:3}).
-
-```{=tex}
-\begin{exe}
-    \ex 
-\begin{xlist}
-    \ex \label{ex:2a}
-    \textsc{lab place elvan ix\textsubscript{a} ten years work. demet ix\textsubscript{b} last year start.} \\
-    ‘In the lab, Elvan has been working for 10 years. Demet started last year.’
-    \\
-    \ex \label{ex:2b}
-    \textsc{ix\textsubscript{a} e-l-a-n know. ix\textsubscript{b} learn.}\\
-    ‘She (Elvan) knows ELAN (the annotation program). She (Demet) is learning.’
-    \end{xlist}
-    \end{exe}
-
-    \begin{exe}
-    \ex \label{ex:3} Signing space, signer, example referential loci for the referents \textsc{elvan} (a) and \textsc{demet} (b) \\
-\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
-
-\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
-%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300
-
-%Curve Lines [id:da7006577339157659] 
-\draw    (100.25,77.81) .. controls (136.19,41.85) and (183.62,42.17) .. (222.5,78.13) ;
-%Straight Lines [id:da8428827309782937] 
-\draw    (112.9,128.8) -- (210.21,128.8) ;
-%Curve Lines [id:da39290455666055124] 
-\draw    (222.19,100.13) .. controls (221.89,103.9) and (217.55,123.77) .. (210.21,128.8) ;
-%Curve Lines [id:da9909903731270634] 
-\draw    (101.17,100.45) .. controls (100.86,104.22) and (108.2,124.02) .. (112.9,128.8) ;
-
-% Text Node
-\draw (88.75,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-b }};
-% Text Node
-\draw (91.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Demet)}};
-% Text Node
-\draw (208.25,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-a }};
-% Text Node
-\draw (212.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Elvan)}};
-% Text Node
-\draw (144,115.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Signer}};
-\end{tikzpicture}
-    \end{exe}
-```
-Grammatical gender as in Romance languages or noun classes as in Bantu languages are some reference tracking tools that are familiar from spoken languages.
-A somehow less known reference-tracking tool in spoken languages is *verbal classifiers*, which morphologically attach to the verb and often cross-reference a core argument in a clause [@aikhenvald2017].
-Classifiers denote sets of objects based on semantic criteria such as size, shape, body-part, texture and sometimes function.
-Usually, the verbal classifier and the cross-referenced noun co-occur in the same clause when the argument is introduced into discourse.
-The classifier can track the argument on its own on subsequent occurrences.
-In sign languages, a verbal classifier morpheme is encoded by a *handshape* that is articulated phonologically co-temporally with a movement, which encodes the verbal meaning [@supalla1990; @benedicto2004].
-These two infused morphemes form a complex predicate.
-Some classifier handshapes and the set of objects that they denote are presented below in (\ref{ex:4}).
-
-```{=tex}
-\begin{exe}
-\ex \label{ex:4}
-\includegraphics[width=25pt]{pictures/flathand.png}: flat entities, surfaces, vehicles.\\
-\includegraphics[width=25pt]{pictures/flat.png}: human agent holding a flat entity.\\
-\includegraphics[width=25pt]{pictures/small_legged.png}: small animal legs.\\
-\includegraphics[width=25pt]{pictures/index.png}: thin entities (like a pencil); upright beings.\\
-\includegraphics[width=25pt]{pictures/hold.png}: human agent holding a thin entity (like a pencil or a branch of flower). 
-\end{exe}
-```
-In (\ref{ex:5}), in the first clause, a referential locus, i.e.
-locus-A, is assigned to the locative source argument \textsc{cage}.
-In the second clause, the core argument \textsc{bird} is cross-referenced with the classifier handshape, i.e.
-small-legged animal.
-It flies from the referential locus of the \textsc{cage} (locus-a) to the referential locus of the goal \textsc{tree} (locus-B).
-
-```{=tex}
-\begin{exe}
-    \ex \label{ex:5}
-\begin{xlist}
-    \ex \textsc{cage} assigned to the referential locus-a. \\ \\
-    \textsc{garden bird cage-a open} \\
-    ‘In the garden, there is a cage and it is open.’
-    \\
-
-    \ex \textsc{tree} assigned to the referential locus-b. The source and goal tracked on the predicates \textsc{fly} and \textsc{land}. \\ \\
-    \textsc{bird fly[manner]-a tree-b land cl-b}\\
-    ‘The bird flies (from the cage assigned to the left) to a tree (assigned to the right).’
-    \end{xlist}
-    \end{exe}
-    \hfill (Adapted from Gökgöz and Sevgi 2020:318)
-```
 ## Linguistic Tools for Tracking Referents in Signed Narratives
@@
 
-\clearpage
+### Nominals
+
+In TİD and many other sign languages, it is possible to use bare noun phrases as in (\ref{ex:6a}) as well as fingerspelled nouns as in (\ref{ex:6c}).
+Fingerspelling refers to the use of the manual alphabet which is derived or borrowed from a surrounding written language [@kubus2008].
+Nouns can also be modified by classifiers as in (\ref{ex:6b}) [@frederiksen2016] or by a prenominal or postnominal IX functioning as a determiner or a demonstrative as in (\ref{ex:6d}) [@maclaughlin1997].
+
+### Pronouns
+
+The IX sign (extended index finger) has various functions in sign languages, one of which is the pronominal use which refers to established nominals by pointing to a previously assigned abstract or physical locus in the signing space when re-introducing a topic in discourse [@emmorey1996].
+The same IX function has also been identified for TİD [@sevinc2006].
+The example in (\ref{ex:7}) illustrates a spatially-anchored pronominal reference for topic re-introduction.
+These signs are thus rarely used for the first mention of an entity [@perniss2015].
+Other interpretations consider pointing signs without noun phrases as examples of demonstratives and not personal pronouns [@koulidobrova2016].
+However, for the purposes of the present study, we will refer to IX constructions as pronominal IX and particularly examine their non-deictic or personal pronoun realizations as a possible discourse strategy to mainly mark topic shift among native and late signers of TİD.
+
+### Classifiers
+
+Classifier predicates are highly complex morphological constructions which are very common in sign languages [@supalla1986].
+These are manually signed iconic representations of animate or inanimate objects, and they embody several properties pertaining to the nominal referent including its location in the signing space, movement, path, size, and shape [@talmy2003].
+As in most sign languages hitherto studied, classifier constructions are also commonly found in TİD [see @dikyuva2015 for a review].
+Following the classification of @engberg-pedersen1993, we distinguish between four types of classifier constructions.
+Whole entity classifiers (WECL) represent the entirety of any referent whereas body part classifiers (BPCL) only reflect a specific part of an animate entity (e.g., limbs, head, tongue, teeth, and so forth).
+Handling classifiers (HCL[^1]) occur with transitive verbs and depict how an object is handled or manipulated by an agent.
+Extension classifiers (ExtCL), on the other hand, denote the size and shape of an entity by following a trajectory to trace its shape.
+We exemplify these constructions in (\ref{ex:8}).
+
+[^1]: As discussed before, an HCL is transitive in nature and depicts the handling action performed by ananimate agent.
+This is why most research tend to consider HCLs as an extension to constructed action [@dudis2004; @cormier2013].
 
@@
 \z \z
+\clearpage
 \begin{multicols}{4}
@@
 ```
-### Nominals
-
-In TİD and many other sign languages, it is possible to use bare noun phrases as in (\ref{ex:6a}) as well as fingerspelled nouns as in (\ref{ex:6c}).
-Fingerspelling refers to the use of the manual alphabet which is derived or borrowed from a surrounding written language [@kubus2008].
-Nouns can also be modified by classifiers as in (\ref{ex:6b}) [@frederiksen2016] or by a prenominal or postnominal IX functioning as a determiner or a demonstrative as in (\ref{ex:6d}) [@maclaughlin1997].
-
-### Pronouns
-
-The IX sign (extended index finger) has various functions in sign languages, one of which is the pronominal use which refers to established nominals by pointing to a previously assigned abstract or physical locus in the signing space when re-introducing a topic in discourse [@emmorey1996].
-The same IX function has also been identified for TİD [@sevinc2006].
-The example in (\ref{ex:7}) illustrates a spatially-anchored pronominal reference for topic re-introduction.
-These signs are thus rarely used for the first mention of an entity [@perniss2015].
-Other interpretations consider pointing signs without noun phrases as examples of demonstratives and not personal pronouns [@koulidobrova2016].
-It is important to mention here that in some spoken languages (e.g., German and Dutch), demonstratives can also refer to persons [@bosch2007] and behave similarly to personal pronouns.
-However, for the purposes of the present study, we will refer to IX constructions as pronominal IX and particularly examine their non-deictic or personal pronoun realizations as a possible discourse strategy to mainly mark topic shift among native and late signers of TİD.
-
-### Classifiers
-
-Classifier predicates are highly complex morphological constructions which are very common in sign languages [@supalla1986].
-These are manually signed iconic representations of animate or inanimate objects, and they embody several properties pertaining to the nominal referent including its location in the signing space, movement, path, size, and shape [@talmy2003].
-As in most sign languages hitherto studied, classifier constructions are also commonly found in TİD [see @dikyuva2015 for a review].
-Following the classification of @engberg-pedersen1993, we distinguish between four types of classifier constructions.
-Whole entity classifiers (WECL) represent the entirety of any referent whereas body part classifiers (BPCL) only reflect a specific part of an animate entity (e.g., limbs, head, tongue, teeth, and so forth).
-Handling classifiers (HCL) occur with transitive verbs and depict how an object is handled or manipulated by an agent.
-On this ground, an HCL[^1] depicts the manner of the referring action while the former two constructions encode the location of the entity as well as its motion in space.
-Extension classifiers (ExtCL), on the other hand, denote the size and shape of an entity by following a trajectory to trace its shape.
-We exemplify these constructions in (\ref{ex:8}).
-
-[^1]: As discussed before, an HCL is transitive in nature and depicts the handling action performed by ananimate agent.
-This is why most research tend to consider HCLs as an extension to constructed action [@dudis2004; @cormier2013].
-
 ### Null marking
@@
 
+## Signed vs. Spoken REs
+
+Commonly in both spoken and sign languages nouns and noun phrases are used as REs.
+Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions [@lillo-martin1990].
+It is important to mention here that in some spoken languages (e.g., German and Dutch), demonstratives can also refer to persons [@bosch2007] and behave similarly to personal pronouns.
+In addition, the pointing sign is used to referentially anchor a referent to a location in signing space.
+For instance, when the signer introduces \textsc{elvan} for the first time, s/he can assign her to the right.
+This referential locus can later be used by agreement verbs and classifier predicates (more below).
+For instance, in (\ref{ex:2a}) there are two referents, \textsc{elvan}, who is referentially-assigned to the right side of the signer, represented with the "A" subscript and \textsc{demet}, who is referentially-assigned to the left side of the signer, represented with the "B" subscript.
+When the signer makes a contrast in (\ref{ex:2b}), as a pronominal use, she points to the right side to refer to ELVAN and to the left to refer to \textsc{demet}.
+We schematically represent the referential locus assigned to the referents in (\ref{ex:3}).
+
+```{=tex}
+\begin{exe}
+    \ex 
+\begin{xlist}
+    \ex \label{ex:2a}
+    \textsc{lab place elvan ix\textsubscript{a} ten years work. demet ix\textsubscript{b} last year start.} \\
+    ‘In the lab, Elvan has been working for 10 years. Demet started last year.’
+    \\
+    \ex \label{ex:2b}
+    \textsc{ix\textsubscript{a} e-l-a-n know. ix\textsubscript{b} learn.}\\
+    ‘She (Elvan) knows ELAN (the annotation program). She (Demet) is learning.’
+    \end{xlist}
+    \end{exe}
+
+    \begin{exe}
+    \ex \label{ex:3} Signing space, signer, example referential loci for the referents \textsc{elvan} (a) and \textsc{demet} (b) \\
+\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
+
+\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
+%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300
+
+%Curve Lines [id:da7006577339157659] 
+\draw    (100.25,77.81) .. controls (136.19,41.85) and (183.62,42.17) .. (222.5,78.13) ;
+%Straight Lines [id:da8428827309782937] 
+\draw    (112.9,128.8) -- (210.21,128.8) ;
+%Curve Lines [id:da39290455666055124] 
+\draw    (222.19,100.13) .. controls (221.89,103.9) and (217.55,123.77) .. (210.21,128.8) ;
+%Curve Lines [id:da9909903731270634] 
+\draw    (101.17,100.45) .. controls (100.86,104.22) and (108.2,124.02) .. (112.9,128.8) ;
+
+% Text Node
+\draw (88.75,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-b }};
+% Text Node
+\draw (91.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Demet)}};
+% Text Node
+\draw (208.25,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-a }};
+% Text Node
+\draw (212.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Elvan)}};
+% Text Node
+\draw (144,115.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Signer}};
+\end{tikzpicture}
+    \end{exe}
+```
+Grammatical gender as in Romance languages or noun classes as in Bantu languages are some reference tracking tools that are familiar from spoken languages.
+A somehow less known reference-tracking tool in spoken languages is *verbal classifiers*, which morphologically attach to the verb and often cross-reference a core argument in a clause [@aikhenvald2017].
+As outlined in the previous section, classifiers denote sets of objects based on semantic criteria such as size, shape, body-part, texture and sometimes function.
+Usually, the verbal classifier and the cross-referenced noun co-occur in the same clause when the argument is introduced into discourse.
+The classifier can track the argument on its own on subsequent occurrences.
+In sign languages, a verbal classifier morpheme is encoded by a *handshape* that is articulated phonologically co-temporally with a movement, which encodes the verbal meaning [@supalla1990; @benedicto2004].
+These two infused morphemes form a complex predicate.
+Some classifier handshapes and the set of objects that they denote are presented below in (\ref{ex:4}).
+
+```{=tex}
+\begin{exe}
+\ex \label{ex:4}
+\includegraphics[width=25pt]{pictures/flathand.png}: flat entities, surfaces, vehicles.\\
+\includegraphics[width=25pt]{pictures/flat.png}: human agent holding a flat entity.\\
+\includegraphics[width=25pt]{pictures/small_legged.png}: small animal legs.\\
+\includegraphics[width=25pt]{pictures/index.png}: thin entities (like a pencil); upright beings.\\
+\includegraphics[width=25pt]{pictures/hold.png}: human agent holding a thin entity (like a pencil or a branch of flower). 
+\end{exe}
+```
+In (\ref{ex:5}), in the first clause, a referential locus, i.e.
+locus-A, is assigned to the locative source argument \textsc{cage}.
+In the second clause, the core argument \textsc{bird} is cross-referenced with the classifier handshape, i.e.
+small-legged animal.
+It flies from the referential locus of the \textsc{cage} (locus-a) to the referential locus of the goal \textsc{tree} (locus-B).
+
+```{=tex}
+\begin{exe}
+    \ex \label{ex:5}
+\begin{xlist}
+    \ex \textsc{cage} assigned to the referential locus-a. \\ \\
+    \textsc{garden bird cage-a open} \\
+    ‘In the garden, there is a cage and it is open.’
+    \\
+
+    \ex \textsc{tree} assigned to the referential locus-b. The source and goal tracked on the predicates \textsc{fly} and \textsc{land}. \\ \\
+    \textsc{bird fly[manner]-a tree-b land cl-b}\\
+    ‘The bird flies (from the cage assigned to the left) to a tree (assigned to the right).’
+    \end{xlist}
+    \end{exe}
+    \hfill (Adapted from Gökgöz and Sevgi 2020:318)
+```
 ## Learning How to Track Referents in Sign Languages

### 2022-08-06 revised_manuscript.tex
updated files

@@
 deprived of their first languages typically until after age 7 when they
-start formal education in a deaf school. First language deprivation or
-delay has been found to have negative effects on certain cognitive and
-linguistic measures among late signing children. To name a few, delayed
-exposure to a first functional language is reported to negatively affect
-deaf children and adults' morphosyntactic production and processing
-(Boudreault and Mayberry, 2006; Cheng and Mayberry, 2021; Cormier et
-al., 2012; Karadöller et al., 2017; Kayabaşı et al., 2020; Newport,
-1990) and narrative abilities (e.g., Becker, 2009). In our overall
-research, we are concerned with the effects of this late exposure in the
-ultimate adult forms of TİD users.
+start formal education in a deaf school. Despite this, many of the early
+intervention programs for late signing children prioritize spoken
+language development and discourage the use of a sign language
+(Kemaloğlu and Kemaloğlu, 2012). Ministry of National Education (2020)
+identifies 72 deaf schools in Turkey for the K-12 level. Even when deaf
+children with hearing parents start their formal education in one of
+these schools, they do not receive proper instruction in TİD since the
+teachers themselves are mainly hearing and only have a very rudimentary
+knowledge of a sign language (Ilkbasaran, 2015). First language
+deprivation or delay has been found to have negative effects on certain
+cognitive and linguistic measures among late signing children. To name a
+few, delayed exposure to a first functional language is reported to
+negatively affect deaf children and adults' morphosyntactic production
+and processing (Boudreault and Mayberry, 2006; Cheng and Mayberry, 2021;
+Cormier et al., 2012; Karadöller et al., 2017; Kayabaşı et al., 2020;
+Newport, 1990) and narrative abilities (e.g., Becker, 2009). In our
+overall research, we are concerned with the effects of this late
+exposure in the ultimate adult forms of TİD users.
 
@@
 languages. There is an emerging body of literature that reports the
-distribution of REs in signed narratives. Building on Gundel et al.
-(1993)'s work, Swabey (2002) demonstrated that, like other spoken
-languages such as Russian, Chinese, and Japanese, American Sign Language
-(ASL) signers use bare nouns for introduced and re-introduced discourse
-contexts whereas null marking is used for referents which are
+distribution of REs in signed narratives (Ahlgren and Bergman, 1994;
+Czubek, 2017; Frederiksen and Mayberry, 2016; Hodge et al., 2019;
+Morgan, 2005; Perniss and Özyürek, 2015; Swabey, 2002). Building on
+Gundel et al. (1993)'s work, Swabey (2002) demonstrated that, like other
+spoken languages such as Russian, Chinese, and Japanese, American Sign
+Language (ASL) signers use bare nouns for introduced and re-introduced
+discourse contexts whereas null marking is used for referents which are
 ``in-focus'' or very accessible to the addressee. Overt REs like
@@
 
-\hypertarget{signed-vs.-spoken-res}{%
-\subsubsection{Signed vs.~Spoken REs}\label{signed-vs.-spoken-res}}
-
-Commonly in both spoken and sign languages nouns and noun phrases are
-used as REs. Spoken languages use pronouns, determines and
-demonstratives where sign languages use a pointing sign, usually glossed
-as IX, which assumes similar functions (Lillo-Martin and Klima, 1990).
-In addition, the pointing sign is used to referentially anchor a
-referent to a location in signing space. For instance, when the signer
-introduces \textsc{elvan} for the first time, s/he can assign her to the
-right. This referential locus can later be used by agreement verbs and
-classifier predicates (more below). For instance, in (\ref{ex:2a}) there
-are two referents, \textsc{elvan}, who is referentially-assigned to the
-right side of the signer, represented with the ``A'' subscript and
-\textsc{demet}, who is referentially-assigned to the left side of the
-signer, represented with the ``B'' subscript. When the signer makes a
-contrast in (\ref{ex:2b}), as a pronominal use, she points to the right
-side to refer to ELVAN and to the left to refer to \textsc{demet}. We
-schematically represent the referential locus assigned to the referents
-in (\ref{ex:3}).
-
-\begin{exe}
-    \ex 
-\begin{xlist}
-    \ex \label{ex:2a}
-    \textsc{lab place elvan ix\textsubscript{a} ten years work. demet ix\textsubscript{b} last year start.} \\
-    ‘In the lab, Elvan has been working for 10 years. Demet started last year.’
-    \\
-    \ex \label{ex:2b}
-    \textsc{ix\textsubscript{a} e-l-a-n know. ix\textsubscript{b} learn.}\\
-    ‘She (Elvan) knows ELAN (the annotation program). She (Demet) is learning.’
-    \end{xlist}
-    \end{exe}
-
-    \begin{exe}
-    \ex \label{ex:3} Signing space, signer, example referential loci for the referents \textsc{elvan} (a) and \textsc{demet} (b) \\
-\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
-
-\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
-%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300
-
-%Curve Lines [id:da7006577339157659] 
-\draw    (100.25,77.81) .. controls (136.19,41.85) and (183.62,42.17) .. (222.5,78.13) ;
-%Straight Lines [id:da8428827309782937] 
-\draw    (112.9,128.8) -- (210.21,128.8) ;
-%Curve Lines [id:da39290455666055124] 
-\draw    (222.19,100.13) .. controls (221.89,103.9) and (217.55,123.77) .. (210.21,128.8) ;
-%Curve Lines [id:da9909903731270634] 
-\draw    (101.17,100.45) .. controls (100.86,104.22) and (108.2,124.02) .. (112.9,128.8) ;
-
-% Text Node
-\draw (88.75,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-b }};
-% Text Node
-\draw (91.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Demet)}};
-% Text Node
-\draw (208.25,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-a }};
-% Text Node
-\draw (212.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Elvan)}};
-% Text Node
-\draw (144,115.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Signer}};
-\end{tikzpicture}
-    \end{exe}
-
-Grammatical gender as in Romance languages or noun classes as in Bantu
-languages are some reference tracking tools that are familiar from
-spoken languages. A somehow less known reference-tracking tool in spoken
-languages is \emph{verbal classifiers}, which morphologically attach to
-the verb and often cross-reference a core argument in a clause
-(Aikhenvald, 2017). Classifiers denote sets of objects based on semantic
-criteria such as size, shape, body-part, texture and sometimes function.
-Usually, the verbal classifier and the cross-referenced noun co-occur in
-the same clause when the argument is introduced into discourse. The
-classifier can track the argument on its own on subsequent occurrences.
-In sign languages, a verbal classifier morpheme is encoded by a
-\emph{handshape} that is articulated phonologically co-temporally with a
-movement, which encodes the verbal meaning (Benedicto and Brentari,
-2004; Supalla, 1990). These two infused morphemes form a complex
-predicate. Some classifier handshapes and the set of objects that they
-denote are presented below in (\ref{ex:4}).
-
-\begin{exe}
-\ex \label{ex:4}
-\includegraphics[width=25pt]{pictures/flathand.png}: flat entities, surfaces, vehicles.\\
-\includegraphics[width=25pt]{pictures/flat.png}: human agent holding a flat entity.\\
-\includegraphics[width=25pt]{pictures/small_legged.png}: small animal legs.\\
-\includegraphics[width=25pt]{pictures/index.png}: thin entities (like a pencil); upright beings.\\
-\includegraphics[width=25pt]{pictures/hold.png}: human agent holding a thin entity (like a pencil or a branch of flower). 
-\end{exe}
-
-In (\ref{ex:5}), in the first clause, a referential locus, i.e.~locus-A,
-is assigned to the locative source argument \textsc{cage}. In the second
-clause, the core argument \textsc{bird} is cross-referenced with the
-classifier handshape, i.e.~small-legged animal. It flies from the
-referential locus of the \textsc{cage} (locus-a) to the referential
-locus of the goal \textsc{tree} (locus-B).
-
-\begin{exe}
-    \ex \label{ex:5}
-\begin{xlist}
-    \ex \textsc{cage} assigned to the referential locus-a. \\ \\
-    \textsc{garden bird cage-a open} \\
-    ‘In the garden, there is a cage and it is open.’
-    \\
-
-    \ex \textsc{tree} assigned to the referential locus-b. The source and goal tracked on the predicates \textsc{fly} and \textsc{land}. \\ \\
-    \textsc{bird fly[manner]-a tree-b land cl-b}\\
-    ‘The bird flies (from the cage assigned to the left) to a tree (assigned to the right).’
-    \end{xlist}
-    \end{exe}
-    \hfill (Adapted from Gökgöz and Sevgi 2020:318)
-
 \hypertarget{linguistic-tools-for-tracking-referents-in-signed-narratives}{%
@@
 
-\clearpage
+\hypertarget{nominals}{%
+\subsubsection{Nominals}\label{nominals}}
+
+In TİD and many other sign languages, it is possible to use bare noun
+phrases as in (\ref{ex:6a}) as well as fingerspelled nouns as in
+(\ref{ex:6c}). Fingerspelling refers to the use of the manual alphabet
+which is derived or borrowed from a surrounding written language (Kubuş,
+2008). Nouns can also be modified by classifiers as in (\ref{ex:6b})
+(Frederiksen and Mayberry, 2016) or by a prenominal or postnominal IX
+functioning as a determiner or a demonstrative as in (\ref{ex:6d})
+(Maclaughlin, 1997).
+
+\hypertarget{pronouns}{%
+\subsubsection{Pronouns}\label{pronouns}}
+
+The IX sign (extended index finger) has various functions in sign
+languages, one of which is the pronominal use which refers to
+established nominals by pointing to a previously assigned abstract or
+physical locus in the signing space when re-introducing a topic in
+discourse (Emmorey, 1996). The same IX function has also been identified
+for TİD (Sevinç, 2006). The example in (\ref{ex:7}) illustrates a
+spatially-anchored pronominal reference for topic re-introduction. These
+signs are thus rarely used for the first mention of an entity (Perniss
+and Özyürek, 2015). Other interpretations consider pointing signs
+without noun phrases as examples of demonstratives and not personal
+pronouns (Koulidobrova and Lillo-Martin, 2016). However, for the
+purposes of the present study, we will refer to IX constructions as
+pronominal IX and particularly examine their non-deictic or personal
+pronoun realizations as a possible discourse strategy to mainly mark
+topic shift among native and late signers of TİD.
+
+\hypertarget{classifiers}{%
+\subsubsection{Classifiers}\label{classifiers}}
+
+Classifier predicates are highly complex morphological constructions
+which are very common in sign languages (Supalla, 1986). These are
+manually signed iconic representations of animate or inanimate objects,
+and they embody several properties pertaining to the nominal referent
+including its location in the signing space, movement, path, size, and
+shape (Talmy, 2003). As in most sign languages hitherto studied,
+classifier constructions are also commonly found in TİD (see Dikyuva et
+al., 2015 for a review). Following the classification of
+Engberg-Pedersen (1993), we distinguish between four types of classifier
+constructions. Whole entity classifiers (WECL) represent the entirety of
+any referent whereas body part classifiers (BPCL) only reflect a
+specific part of an animate entity (e.g., limbs, head, tongue, teeth,
+and so forth). Handling classifiers (HCL\footnote{As discussed before,
+  an HCL is transitive in nature and depicts the handling action
+  performed by ananimate agent. This is why most research tend to
+  consider HCLs as an extension to constructed action (Cormier et al.,
+  2013; Dudis, 2004).}) occur with transitive verbs and depict how an
+object is handled or manipulated by an agent. Extension classifiers
+(ExtCL), on the other hand, denote the size and shape of an entity by
+following a trajectory to trace its shape. We exemplify these
+constructions in (\ref{ex:8}).
 
@@
 \footnotetext{We use small capitals to indicate sign glosses.}
-
-\hypertarget{nominals}{%
-\subsubsection{Nominals}\label{nominals}}
-
-In TİD and many other sign languages, it is possible to use bare noun
-phrases as in (\ref{ex:6a}) as well as fingerspelled nouns as in
-(\ref{ex:6c}). Fingerspelling refers to the use of the manual alphabet
-which is derived or borrowed from a surrounding written language (Kubuş,
-2008). Nouns can also be modified by classifiers as in (\ref{ex:6b})
-(Frederiksen and Mayberry, 2016) or by a prenominal or postnominal IX
-functioning as a determiner or a demonstrative as in (\ref{ex:6d})
-(Maclaughlin, 1997).
-
-\hypertarget{pronouns}{%
-\subsubsection{Pronouns}\label{pronouns}}
-
-The IX sign (extended index finger) has various functions in sign
-languages, one of which is the pronominal use which refers to
-established nominals by pointing to a previously assigned abstract or
-physical locus in the signing space when re-introducing a topic in
-discourse (Emmorey, 1996). The same IX function has also been identified
-for TİD (Sevinç, 2006). The example in (\ref{ex:7}) illustrates a
-spatially-anchored pronominal reference for topic re-introduction. These
-signs are thus rarely used for the first mention of an entity (Perniss
-and Özyürek, 2015). Other interpretations consider pointing signs
-without noun phrases as examples of demonstratives and not personal
-pronouns (Koulidobrova and Lillo-Martin, 2016). It is important to
-mention here that in some spoken languages (e.g., German and Dutch),
-demonstratives can also refer to persons (Bosch and Umbach, 2007) and
-behave similarly to personal pronouns. However, for the purposes of the
-present study, we will refer to IX constructions as pronominal IX and
-particularly examine their non-deictic or personal pronoun realizations
-as a possible discourse strategy to mainly mark topic shift among native
-and late signers of TİD.
-
-\hypertarget{classifiers}{%
-\subsubsection{Classifiers}\label{classifiers}}
-
-Classifier predicates are highly complex morphological constructions
-which are very common in sign languages (Supalla, 1986). These are
-manually signed iconic representations of animate or inanimate objects,
-and they embody several properties pertaining to the nominal referent
-including its location in the signing space, movement, path, size, and
-shape (Talmy, 2003). As in most sign languages hitherto studied,
-classifier constructions are also commonly found in TİD (see Dikyuva et
-al., 2015 for a review). Following the classification of
-Engberg-Pedersen (1993), we distinguish between four types of classifier
-constructions. Whole entity classifiers (WECL) represent the entirety of
-any referent whereas body part classifiers (BPCL) only reflect a
-specific part of an animate entity (e.g., limbs, head, tongue, teeth,
-and so forth). Handling classifiers (HCL) occur with transitive verbs
-and depict how an object is handled or manipulated by an agent. On this
-ground, an HCL\footnote{As discussed before, an HCL is transitive in
-  nature and depicts the handling action performed by ananimate agent.
-  This is why most research tend to consider HCLs as an extension to
-  constructed action (Cormier et al., 2013; Dudis, 2004).} depicts the
-manner of the referring action while the former two constructions encode
-the location of the entity as well as its motion in space. Extension
-classifiers (ExtCL), on the other hand, denote the size and shape of an
-entity by following a trajectory to trace its shape. We exemplify these
-constructions in (\ref{ex:8}).
 
@@
 verbs.
+
+\hypertarget{signed-vs.-spoken-res}{%
+\subsection{Signed vs.~Spoken REs}\label{signed-vs.-spoken-res}}
+
+Commonly in both spoken and sign languages nouns and noun phrases are
+used as REs. Spoken languages use pronouns, determines and
+demonstratives where sign languages use a pointing sign, usually glossed
+as IX, which assumes similar functions (Lillo-Martin and Klima, 1990).
+It is important to mention here that in some spoken languages (e.g.,
+German and Dutch), demonstratives can also refer to persons (Bosch and
+Umbach, 2007) and behave similarly to personal pronouns. In addition,
+the pointing sign is used to referentially anchor a referent to a
+location in signing space. For instance, when the signer introduces
+\textsc{elvan} for the first time, s/he can assign her to the right.
+This referential locus can later be used by agreement verbs and
+classifier predicates (more below). For instance, in (\ref{ex:2a}) there
+are two referents, \textsc{elvan}, who is referentially-assigned to the
+right side of the signer, represented with the ``A'' subscript and
+\textsc{demet}, who is referentially-assigned to the left side of the
+signer, represented with the ``B'' subscript. When the signer makes a
+contrast in (\ref{ex:2b}), as a pronominal use, she points to the right
+side to refer to ELVAN and to the left to refer to \textsc{demet}. We
+schematically represent the referential locus assigned to the referents
+in (\ref{ex:3}).
+
+\begin{exe}
+    \ex 
+\begin{xlist}
+    \ex \label{ex:2a}
+    \textsc{lab place elvan ix\textsubscript{a} ten years work. demet ix\textsubscript{b} last year start.} \\
+    ‘In the lab, Elvan has been working for 10 years. Demet started last year.’
+    \\
+    \ex \label{ex:2b}
+    \textsc{ix\textsubscript{a} e-l-a-n know. ix\textsubscript{b} learn.}\\
+    ‘She (Elvan) knows ELAN (the annotation program). She (Demet) is learning.’
+    \end{xlist}
+    \end{exe}
+
+    \begin{exe}
+    \ex \label{ex:3} Signing space, signer, example referential loci for the referents \textsc{elvan} (a) and \textsc{demet} (b) \\
+\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
+
+\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
+%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300
+
+%Curve Lines [id:da7006577339157659] 
+\draw    (100.25,77.81) .. controls (136.19,41.85) and (183.62,42.17) .. (222.5,78.13) ;
+%Straight Lines [id:da8428827309782937] 
+\draw    (112.9,128.8) -- (210.21,128.8) ;
+%Curve Lines [id:da39290455666055124] 
+\draw    (222.19,100.13) .. controls (221.89,103.9) and (217.55,123.77) .. (210.21,128.8) ;
+%Curve Lines [id:da9909903731270634] 
+\draw    (101.17,100.45) .. controls (100.86,104.22) and (108.2,124.02) .. (112.9,128.8) ;
+
+% Text Node
+\draw (88.75,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-b }};
+% Text Node
+\draw (91.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Demet)}};
+% Text Node
+\draw (208.25,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-a }};
+% Text Node
+\draw (212.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Elvan)}};
+% Text Node
+\draw (144,115.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Signer}};
+\end{tikzpicture}
+    \end{exe}
+
+Grammatical gender as in Romance languages or noun classes as in Bantu
+languages are some reference tracking tools that are familiar from
+spoken languages. A somehow less known reference-tracking tool in spoken
+languages is \emph{verbal classifiers}, which morphologically attach to
+the verb and often cross-reference a core argument in a clause
+(Aikhenvald, 2017). As outlined in the previous section, classifiers
+denote sets of objects based on semantic criteria such as size, shape,
+body-part, texture and sometimes function. Usually, the verbal
+classifier and the cross-referenced noun co-occur in the same clause
+when the argument is introduced into discourse. The classifier can track
+the argument on its own on subsequent occurrences. In sign languages, a
+verbal classifier morpheme is encoded by a \emph{handshape} that is
+articulated phonologically co-temporally with a movement, which encodes
+the verbal meaning (Benedicto and Brentari, 2004; Supalla, 1990). These
+two infused morphemes form a complex predicate. Some classifier
+handshapes and the set of objects that they denote are presented below
+in (\ref{ex:4}).
+
+\begin{exe}
+\ex \label{ex:4}
+\includegraphics[width=25pt]{pictures/flathand.png}: flat entities, surfaces, vehicles.\\
+\includegraphics[width=25pt]{pictures/flat.png}: human agent holding a flat entity.\\
+\includegraphics[width=25pt]{pictures/small_legged.png}: small animal legs.\\
+\includegraphics[width=25pt]{pictures/index.png}: thin entities (like a pencil); upright beings.\\
+\includegraphics[width=25pt]{pictures/hold.png}: human agent holding a thin entity (like a pencil or a branch of flower). 
+\end{exe}
+
+In (\ref{ex:5}), in the first clause, a referential locus, i.e.~locus-A,
+is assigned to the locative source argument \textsc{cage}. In the second
+clause, the core argument \textsc{bird} is cross-referenced with the
+classifier handshape, i.e.~small-legged animal. It flies from the
+referential locus of the \textsc{cage} (locus-a) to the referential
+locus of the goal \textsc{tree} (locus-B).
+
+\begin{exe}
+    \ex \label{ex:5}
+\begin{xlist}
+    \ex \textsc{cage} assigned to the referential locus-a. \\ \\
+    \textsc{garden bird cage-a open} \\
+    ‘In the garden, there is a cage and it is open.’
+    \\
+
+    \ex \textsc{tree} assigned to the referential locus-b. The source and goal tracked on the predicates \textsc{fly} and \textsc{land}. \\ \\
+    \textsc{bird fly[manner]-a tree-b land cl-b}\\
+    ‘The bird flies (from the cage assigned to the left) to a tree (assigned to the right).’
+    \end{xlist}
+    \end{exe}
+    \hfill (Adapted from Gökgöz and Sevgi 2020:318)
 
@@
 \begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-ahlgren1994}{}}%
+Ahlgren, Inger, Bergman, Brita, 1994. Reference in narratives. I.
+Ahlgren, B. Bergman \& M.
+
 \leavevmode\vadjust pre{\hypertarget{ref-ahn2019}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-hodge2019}{}}%
+Hodge, Gabrielle, Ferrara, Lindsay N., Anible, Benjamin D., 2019. The
+semiotic diversity of doing reference in a deaf signed language. Journal
+of Pragmatics 143, 3353.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ilkbasaran2015}{}}%
+Ilkbasaran, Deniz, 2015.
+\href{https://www.proquest.com/docview/1656169535/abstract/16EA71911A644FC3PQ/1}{Literacies,
+mobilities and agencies of deaf youth in turkey: Constraints and
+opportunities in the 21st century} (PhD thesis). Ann Arbor, United
+States.
+
 \leavevmode\vadjust pre{\hypertarget{ref-karadoller2017}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-kemaloglu2012}{}}%
+Kemaloğlu, Yusuf Kemal, Kemaloğlu, Pınar Yaprak, 2012. The history of
+sign language and deaf education in Turkey. Kulak burun bogaz ihtisas
+dergisi: KBB = Journal of ear, nose, and throat 22, 65--76.
+doi:\href{https://doi.org/10.5606/kbbihtisas.2012.013}{10.5606/kbbihtisas.2012.013}
+
 \leavevmode\vadjust pre{\hypertarget{ref-koulidobrova2016}{}}%
@@
 american sign language. Sociolinguistics in Deaf communities 255271.
+
+\leavevmode\vadjust pre{\hypertarget{ref-ministryofnationaleducation2020}{}}%
+Ministry of National Education, 2020.
+\href{http://sgb.meb.gov.tr/meb_iys_dosyalar/2020_09/04144812_meb_istatistikleri_orgun_\%20egitim_2019_2020.pdf}{Milli
+e{ğ}itim istatistikleri, örgün e{ğ}itim 2019/{'}20 {[}national education
+statistics, formal education 2019/{'}20{]}}.
 

### 2022-08-06 revised_manuscript.Rmd
updated files

@@
 Other interpretations consider pointing signs without noun phrases as examples of demonstratives and not personal pronouns [@koulidobrova2016].
+It is also important to mention here that in some spoken languages (e.g., German and Dutch), demonstratives can also refer to persons [@bosch2007] and behave similarly to personal pronouns.
 However, for the purposes of the present study, we will refer to IX constructions as pronominal IX and particularly examine their non-deictic or personal pronoun realizations as a possible discourse strategy to mainly mark topic shift among native and late signers of TİD.
@@
 Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions [@lillo-martin1990].
-It is important to mention here that in some spoken languages (e.g., German and Dutch), demonstratives can also refer to persons [@bosch2007] and behave similarly to personal pronouns.
 In addition, the pointing sign is used to referentially anchor a referent to a location in signing space.
@@
 
-We collected `r nrow(df_narration)` data points as a result of the experiment.
-Each data point is a referent tracked by the participant.
-Each of the REs is classified in terms of its Discourse Status (Introduction, Maintenance, Re-introduction), its Type (Nominal, Pronominal, Classifer, Null)\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer, and Null).
-However, we allude to the numerical distribution of the subtypes in this and following sections.}, and Participant Group (Native, Late)\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17).}.
+We collected `r nrow(df_narration)` data points as a result of the experiment.\footnote{For the purposes of this double-blind review, we share all our data and code in an anonymous online repository that can be accessed here: \url{https://anonymous.4open.science/r/reference_tracking_revision-A542/}} Each data point is a referent tracked by the participant.
+Each of the REs is classified in terms of its Discourse Status (Introduction, Maintenance, Re-introduction), its Type (Nominal, Pronominal, Classifer, Null),\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer, and Null).
+However, we allude to the numerical distribution of the subtypes in this and following sections.} and Participant Group (Native, Late).\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17).}
 
@@
 ```
-Table \ref{tab:my-table2} shows that for all discourse statuses the most numerous nominal type was \textsc{bare nom} which was followed by nouns modified by index or classifiers.
-In number, signers used \textsc{mod nom} more in introduction and re-introduction compared to maintenance settings.
+Table \ref{tab:my-table2} shows that for all discourse statuses the most numerous nominal type was bare nominal which was followed by nouns modified by index or classifiers.
+In number, signers used modified nouns more in introduction and re-introduction compared to maintenance settings.
 Signers produced a total number of `r sum(df_narration$ref_type == "Pronominal")` pronouns (Table \ref{tab:my-table3}).

### 2022-08-06 revised_manuscript.tex
updated files

@@
 without noun phrases as examples of demonstratives and not personal
-pronouns (Koulidobrova and Lillo-Martin, 2016). However, for the
-purposes of the present study, we will refer to IX constructions as
-pronominal IX and particularly examine their non-deictic or personal
-pronoun realizations as a possible discourse strategy to mainly mark
-topic shift among native and late signers of TİD.
+pronouns (Koulidobrova and Lillo-Martin, 2016). It is also important to
+mention here that in some spoken languages (e.g., German and Dutch),
+demonstratives can also refer to persons (Bosch and Umbach, 2007) and
+behave similarly to personal pronouns. However, for the purposes of the
+present study, we will refer to IX constructions as pronominal IX and
+particularly examine their non-deictic or personal pronoun realizations
+as a possible discourse strategy to mainly mark topic shift among native
+and late signers of TİD.
 
@@
 as IX, which assumes similar functions (Lillo-Martin and Klima, 1990).
-It is important to mention here that in some spoken languages (e.g.,
-German and Dutch), demonstratives can also refer to persons (Bosch and
-Umbach, 2007) and behave similarly to personal pronouns. In addition,
-the pointing sign is used to referentially anchor a referent to a
-location in signing space. For instance, when the signer introduces
-\textsc{elvan} for the first time, s/he can assign her to the right.
-This referential locus can later be used by agreement verbs and
+In addition, the pointing sign is used to referentially anchor a
+referent to a location in signing space. For instance, when the signer
+introduces \textsc{elvan} for the first time, s/he can assign her to the
+right. This referential locus can later be used by agreement verbs and
 classifier predicates (more below). For instance, in (\ref{ex:2a}) there
@@
 
-We collected 4142 data points as a result of the experiment. Each data
-point is a referent tracked by the participant. Each of the REs is
-classified in terms of its Discourse Status (Introduction, Maintenance,
-Re-introduction), its Type (Nominal, Pronominal, Classifer,
-Null)\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer, and Null).
-However, we allude to the numerical distribution of the subtypes in this and following sections.},
+We collected 4142 data points as a result of the
+experiment.\footnote{For the purposes of this double-blind review, we share all our data and code in an anonymous online repository that can be accessed here: \url{https://anonymous.4open.science/r/reference_tracking_revision-A542/}}
+Each data point is a referent tracked by the participant. Each of the
+REs is classified in terms of its Discourse Status (Introduction,
+Maintenance, Re-introduction), its Type (Nominal, Pronominal, Classifer,
+Null),\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer, and Null).
+However, we allude to the numerical distribution of the subtypes in this and following sections.}
 and Participant Group (Native,
-Late)\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17).}.
+Late).\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17).}
 
@@
 Table \ref{tab:my-table2} shows that for all discourse statuses the most
-numerous nominal type was \textsc{bare nom} which was followed by nouns
-modified by index or classifiers. In number, signers used
-\textsc{mod nom} more in introduction and re-introduction compared to
-maintenance settings. Signers produced a total number of 40 pronouns
-(Table \ref{tab:my-table3}). A great majority of the pronominal
-constructions were used to re-introduce a referent into the discourse.
-We observed only a very small amount of pronoun use for referent
-introduction and maintenance.
+numerous nominal type was bare nominal which was followed by nouns
+modified by index or classifiers. In number, signers used modified nouns
+more in introduction and re-introduction compared to maintenance
+settings. Signers produced a total number of 40 pronouns (Table
+\ref{tab:my-table3}). A great majority of the pronominal constructions
+were used to re-introduce a referent into the discourse. We observed
+only a very small amount of pronoun use for referent introduction and
+maintenance.
 
@@
 maintenance contexts. It means that the use of null markers did not
-linearly increase as the addition of using a null marker instead of a
-classifier as a native plus using null markers instead of a classifier
-for maintaining a referent versus re-introducing it.
+linearly increase as the sum of using a null marker as a native plus
+using null markers for maintaining a referent versus re-introducing it.
 

### 2022-08-06 revised_manuscript.Rmd
update

@@
 In order to capture in-group variation among late signers, we run an additional supplementary analysis in which we observed very wide credible intervals due to the small group size, rendering the results uninformative.
-For the analysis, see \ref{section:d}}.
+For the analysis, see \ref{section:d}.}.
 All participants indicated that their preferred language of communication was TİD which they used for most contexts.
@@
 In addition to TİD, the home language for some native signers (n = 4) in early childhood included the use of home sign\footnote{Home signs are gesture-based communication systems that emerge between the deaf children and their primary caregivers.
-Although home sign communication may have internal consistency, they are not fully-flegged linguistic systems like natural languages}.
+Although home sign communication may have internal consistency, they are not fully-flegged linguistic systems like natural languages.}.
 The age of the native participants ranged between 18 and 35 (*M* = 26.4 years, *SD* = 5.0 years).
@@
 For instance, some intransitive plain verbs like \textsc{fall} or \textsc{jump} in TİD may display spatial person agreement with the single argument in the clause [@gokgoz2020].
-See @lourenço2018 also for a similar discussion in Brazilian Sign Language.}.
+See @lourenço2018 also for a similar discussion in Brazilian Sign Language.
 

### 2022-08-06 revised_manuscript.tex
update

@@
 In order to capture in-group variation among late signers, we run an additional supplementary analysis in which we observed very wide credible intervals due to the small group size, rendering the results uninformative.
-For the analysis, see \ref{section:d}}.
+For the analysis, see \ref{section:d}.}.
 All participants indicated that their preferred language of
@@
 sign\footnote{Home signs are gesture-based communication systems that emerge between the deaf children and their primary caregivers.
-Although home sign communication may have internal consistency, they are not fully-flegged linguistic systems like natural languages}.
+Although home sign communication may have internal consistency, they are not fully-flegged linguistic systems like natural languages.}.
 The age of the native participants ranged between 18 and 35 (\emph{M} =
@@
   and Wilbur (2018) also for a similar discussion in Brazilian Sign
-  Language.\}.}. Agreement verbs and CA were the following implicit
-markers of reference, 0.77 and 0.72 of which were used for maintained
-contexts, respectively. This contrasts with some of the research
-(Czubek, 2017) in which agreement verbs ranked lowest on scale,
-indicating that these constructions were more implicit compared to plain
-verbs. However, there seems to be a common ground for the position of CA
-markers which have been claimed to rank higher than other null markers
-as in the present study (Barberà and Massó, 2009; Czubek, 2017;
-Frederiksen and Mayberry, 2016). In Figure \ref{fig:scale}, we report
-the distribution of REs in TİD as a function of referent accessibility
-and compare our accessibility hierarchy with that of (Frederiksen and
-Mayberry, 2016) for ASL. In the hierarchies, the leftmost REs (e.g.,
-modified and bare nouns for TİD) are used when the accessibility is low.
-If the referent is already active in the addressee's mind and its
-accessibility is high, then the rightmost markers (e.g., agreement and
-plain verbs for TİD) are selected.
+  Language.}. Agreement verbs and CA were the following implicit markers
+of reference, 0.77 and 0.72 of which were used for maintained contexts,
+respectively. This contrasts with some of the research (Czubek, 2017) in
+which agreement verbs ranked lowest on scale, indicating that these
+constructions were more implicit compared to plain verbs. However, there
+seems to be a common ground for the position of CA markers which have
+been claimed to rank higher than other null markers as in the present
+study (Barberà and Massó, 2009; Czubek, 2017; Frederiksen and Mayberry,
+2016). In Figure \ref{fig:scale}, we report the distribution of REs in
+TİD as a function of referent accessibility and compare our
+accessibility hierarchy with that of (Frederiksen and Mayberry, 2016)
+for ASL. In the hierarchies, the leftmost REs (e.g., modified and bare
+nouns for TİD) are used when the accessibility is low. If the referent
+is already active in the addressee's mind and its accessibility is high,
+then the rightmost markers (e.g., agreement and plain verbs for TİD) are
+selected.
 

### 2023-01-28 revised_manuscript.Rmd
update
Removed participant 13 and updated labels for the plots.

@@
 library(bayestestR)
+
+# theme function for ggplot2
 theme_apa <- function(base_size = 10, base_family = "serif", box = FALSE) {
@@
 
+# remove subject 13, native but home language is home sign instead of TiD
+df_narration %<>% subset(Subject != 13)
+
 # Coding nominal, Zero, Classifier, and Pronominal categories
@@
                                         `b_muNull_Intercept` = "Null_Intercept",
-                                        `b_muNominal_Nativeness1` = "Nom_Native",
-                                        `b_muNull_Nativeness1` = "Null_Native",
+                                        `b_muNominal_Nativeness1` = "Nom_DoD",
+                                        `b_muNull_Nativeness1` = "Null_DoD",
                                         `b_muNominal_Discourse1` = "Nom_Intro",
                                         `b_muNominal_Discourse2` = "Nom_Maintain",
-                                        'b_muNominal_Nativeness1:Discourse1' = 'Nom_Native*Intro',
-                                        'b_muNominal_Nativeness1:Discourse2' = 'Nom_Native*Maintain',
-                                        'b_muNull_Nativeness1:Discourse1' = 'Null_Native*Intro',
-                                        'b_muNull_Nativeness1:Discourse2' = 'Null_Native*Maintain',
+                                        'b_muNominal_Nativeness1:Discourse1' = 'Nom_DoD*Intro',
+                                        'b_muNominal_Nativeness1:Discourse2' = 'Nom_DoD*Maintain',
+                                        'b_muNull_Nativeness1:Discourse1' = 'Null_DoD*Intro',
+                                        'b_muNull_Nativeness1:Discourse2' = 'Null_DoD*Maintain',
                                         `b_muNull_Discourse1` = "Null_Intro",
@@
                                         `b_muNominal_Group2M1` = "Nom_Late1-Late2",
-                                        `b_muNominal_Group3M2` = "Nom_Native-Late1",
+                                        `b_muNominal_Group3M2` = "Nom_Early-Late1",
                                         `b_muNominal_Discourse1` = "Nom_Intro",
@@
                                         `b_muNominal_Group2M1:Discourse1` = 'Nom_L1-L2*Intro',
-                                        `b_muNominal_Group3M2:Discourse1` = 'Nom_N-L1*Intro',
+                                        `b_muNominal_Group3M2:Discourse1` = 'Nom_E-L1*Intro',
                                         `b_muNominal_Group2M1:Discourse2` = 'Nom_L1-L2*Maintain',
-                                        `b_muNominal_Group3M2:Discourse2` = 'Nom_N-L1*Maintain',
+                                        `b_muNominal_Group3M2:Discourse2` = 'Nom_E-L1*Maintain',
                                         `b_muNull_Group2M1` = "Null_Late1-Late2",
-                                        `b_muNull_Group3M2` = "Null_Native-Late1",
+                                        `b_muNull_Group3M2` = "Null_Early-Late1",
                                         `b_muNull_Discourse1` = "Null_Intro",
@@
                                         `b_muNull_Group2M1:Discourse1` = 'Null_L1-L2*Intro',
-                                        `b_muNull_Group3M2:Discourse1` = 'Null_N-L1*Intro',
+                                        `b_muNull_Group3M2:Discourse1` = 'Null_E-L1*Intro',
                                         `b_muNull_Group2M1:Discourse2` = 'Null_L1-L2*Maintain',
-                                        `b_muNull_Group3M2:Discourse2` = 'Null_N-L1*Maintain')
+                                        `b_muNull_Group3M2:Discourse2` = 'Null_E-L1*Maintain')
 saveRDS(df_model_3, "./rds/df_model_3.rds")
+
 #df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "NULL"))

### 2023-01-31 revised_manuscript.Rmd
updated

@@
   chunk_output_type: console
-title: Reference Tracking Strategies of Native and Late Signers in Turkish Sign Language
+title: Reference Tracking Strategies of Deaf Adult Signers in Turkish Sign Language
 abstract: |
-  We investigated the reference tracking strategies among native and late adult deaf signers of Turkish Sign Language.
-Consistent with the theories of saliency and referential accessibility, regardless of their acquisiton groups, signers mainly used nominals and extension classifiers for introducing referents.
-To maintain a referent in the previous clause, signers chose null marking (e.g., constructed action, agreement and plain verbs).
+  We investigated the reference tracking strategies among deaf adults with signing deaf parents (DoD) and adult deaf signers with non-signing hearing parents or caregivers (DoH) who had late exposure to Turkish Sign Language (TİD).
+Consistent with the theories of saliency and referential accessibility, regardless of their acquisiton groups or parental hearing status, signers mainly used nominals and extension classifiers for introducing referents.
+To maintain a referent from the previous clause, the signers used zero anaphora (e.g., constructed action, agreement and plain verbs).
 For topic shifts or re-introduced contexts, nominals and pronouns were chiefly favored although we observed very little pronominal use.
-As for age of acquisition effects, we only report limited over-redundancy for late signers who used null marking less compared to native signers especially for introduced and maintained contexts.
-Given our limited support for redundancy by late signers, we conclude that native signers might utilize the spatial affordances of the visual-spatial modality better than late signers only to a certain extent.
+As for the effect of native acquisition from deaf caregivers, we only report limited over-redundancy for DoH signers who used zero anaphora less compared to DoD especially for introduced and maintained contexts.
+We conclude that DoH signers are still able to achieve native-like competency in terms of reference tracking in simple narratives but DoD signers utilize the spatial affordances of the visual-spatial modality better than DoH signers only to a certain extent.
 keywords: "Referential accessibility, signed narratives, pragmatic competence, delayed language acquisition, Turkish Sign Language"
@@
    - \usepackage{multicol}
-   - \setlength{\columnsep}{-3cm}
+   - \setlength{\columnsep}{-2cm}
    - \usepackage{setspace}
@@
 
-Narratives are extensive discourses that construct an important part of human communication.
-In order to create a systematic and coherent organization of experiences, narrators must choose appropriate linguistic forms to refer to entities [@ariel1990; @givon1983; @gundel1993]: Certain forms such as nominal expressions are commonly used for the first mention of a referent in a discourse, and other forms such as pronouns and null marking may track previously mentioned entities which could either be maintained across multiple clauses or re-introduced back into the discourse.
+Narratives are extensive discourses that form an important part of human communication.
+In order to create a systematic and coherent organization of experiences, narrators must choose appropriate linguistic forms to refer to entities [@ariel1990; @givon1983; @gundel1993].
+Certain forms such as nominal expressions are commonly used for the first mention of a referent in a discourse, and other forms such as pronouns and zero anaphora may track previously mentioned entities which could either be maintained across multiple clauses or re-introduced back into the discourse.
 The selection of these referring expressions (henceforth REs) is contingent on cognitive abilities, such that individuals must rely on the activation of the working memory to keep in mind previously referred expressions and long strings of information [@morgan2005; @bamberg1997].
@@
 As can be expected, there is a plethora of research examining how narrative competence is achieved in both monolingual and bilingual acquisition contexts among children and adults.
-Although narrative competence and referent selection have been studied extensively for typically developing speakers [@aksu-koc2015; @hickmann1996; @williams1988], deaf individuals who acquired a sign language in atypical learning contexts (i.e., deaf signers with linguistic deprivation) have remained mostly understudied with an exception of handful of studies [@cormier2013; @gur2018; @becker2009].
-Nevertheless, it is paramount that we investigate and describe the effects of language deprivation among deaf adults on narrative competence more in depth.
-This will help us understand whether the loss of systematic and frequent language input in the early years of life affect how deaf signers produce a narrative in their adulthood.
-Given this, the present study will focus on deaf adult signers with language deprivation and examine their reference tracking strategies in signed narratives.
-
-Sign Languages are natural languages of the Deaf communities all around the world.
+Although narrative competence and referent selection have been studied extensively for typically developing speakers [@aksu-koc2015; @hickmann1996; @williams1988], deaf individuals who acquired a sign language in atypical learning contexts (i.e., deaf signers with delayed exposure to a sign language) have remained mostly understudied with an exception of a handful of studies [@cormier2013; @gur2018; @becker2009].
+Nevertheless, it is paramount that we investigate and describe the effects of late exposure to a sign language among deaf adults growing up in a signing vs.
+non-signing environment on narrative competence more in depth.
+This will help us understand whether the lack of systematic and frequent language exposure in the early years of life affect how deaf signers produce a narrative in their adulthood.
+Given this, the present study will focus on two groups of signers and examine their reference tracking strategies in signed narratives: (i) deaf of deaf (DoD) adult signers (i.e., signers who have deaf caregivers and natively acquired a sign language from birth) and (ii) deaf of hearing (DoH) adult signers (i.e., signers who have hearing non-signing caregivers and state to have started acquiring a sign language later in life).
+
+Sign languages are the natural languages of the Deaf communities all around the world.
 They are independent languages, which do not derive from majority spoken language(s) but emerge naturally when deaf people form a community.
@@
 Deaf or hearing children acquiring a SL within a family where there is one or more deaf caregivers display the same developmental milestones as hearing children acquiring a spoken language in a hearing family [@chamberlain_language_1999; @meier1990].
-However, reports indicate that over 90% of all deaf individuals are born into hearing families who do not know how to sign (i.e., those with language deprivation) [@woll2013].
+However, reports indicate that over 90% of all deaf individuals are born into hearing families (DoH) who do not know how to sign [@woll2013].
 This means that the auditory modality used in the household is not accessible to many deaf children.
-That is why the age of acquisition of a first functional language varies greatly among deaf children, with many of whom being deprived of their first languages typically until after age 7 when they start formal education in a deaf school.
+That is why the age of acquisition of a first functional language varies greatly among deaf children, with many of whom being deprived of their first languages typically until after age 7 when they start formal education in a deaf school, especially in the context of Turkey [@sari_analysis_2005].
 Despite this, many of the early intervention programs for late signing children prioritize spoken language development and discourage the use of a sign language [@kemaloglu2012].
-@ministryofnationaleducation2020 identifies 72 deaf schools in Turkey for the K-12 level.
-Even when deaf children with hearing parents start their formal education in one of these schools, they do not receive proper instruction in TİD since the teachers themselves are mainly hearing and only have a very rudimentary knowledge of a sign language [@ilkbasaran2015].
-First language deprivation or delay has been found to have negative effects on certain cognitive and linguistic measures among late signing children.
-To name a few, delayed exposure to a first functional language is reported to negatively affect deaf children and adults' morphosyntactic production and processing [@kayabasi2020; @boudreault2006; @newport1990; @cheng2021; @karadoller2017; @cormier2012] and narrative abilities [e.g., @becker2009].
+The @ministryofnationaleducation2020 identifies 72 deaf schools in Turkey for the K-12 level.
+Even when deaf children with hearing parents start their formal education in one of these schools, they do not receive proper instruction in TİD since the teachers themselves are mainly hearing and only have a very rudimentary knowledge of a sign language [@ilkbasaran2015; @sari_analysis_2005].
+In some cases, it has been argued that late exposure to a sign language among certain DoH signers has negative effects on some cognitive and linguistic measures.
+To name a few, delayed exposure to a first functional language is reported to negatively affect deaf children and adults' morphosyntactic production and processing [@kayaba2022; @boudreault2006; @newport1990; @cheng2021; @karadoller2017; @cormier2012] and narrative abilities [e.g., @becker2009].
 In our overall research, we are concerned with the effects of this late exposure in the ultimate adult forms of TİD users.
@@
 Among the theoretical frameworks that aim to explain that different linguistic forms are selected in accordance with the addressee's mental representations of the referents in the discourse is the Givenness Hierarchy [@gundel1993].
-According to this model, given or familiar information is expressed in more attenuated forms such as pronominal use and null marking since the mental status of the expressed referent is already at the center of attention, as indicated by the present contextual information [@chafe1976].
+According to this model, given or familiar information is expressed in more attenuated forms such as pronouns and zero anaphora (i.e., there is no explicit mention of the referent) since the mental status of the expressed referent is already at the center of attention, as indicated by the present contextual information [@chafe1976].
 The Givenness Hierarchy is akin to other previously proposed models of reference control within the topicality and saliency paradigm such as the Accessibility Hierarchy [@ariel1990], Topic Continuity [@givon1983], and Familiarity Scale [@prince1981].
@@
 In the utterance, REs for "Eddie" are given in bold.
-All three REs (nominal, pronominal, and null marking) refer to the same entity; however, the quantity of marking decreases as the referent's accessibility increases.
+All three REs (nominal, pronominal, and zero anaphora) refer to the same entity; however, the quantity of marking decreases as the referent's accessibility increases.
 On this ground, for highly accessible referents (i.e., those that are highly active in the mind), linguistic forms with the least phonetic content are selected for the least amount of processing effort.
@@
 These research indicate that the introduction and re-introduction of a referent are often accomplished by full noun phrases which can be commonly modified by an adjective or demonstrative.
-Maintenance, on the other hand, is typically marked with pronominals and null markers since the continued referent is still highly accessible to the addressee.
+Maintenance, on the other hand, is typically marked with pronominals and zero anaphora since the continued referent is still highly accessible to the addressee.
 
@@
 There is an emerging body of literature that reports the distribution of REs in signed narratives [@czubek2017; @swabey2002; @frederiksen2016; @perniss2015; @morgan2005; @hodge2019; @ahlgren1994].
-Building on @gundel1993's work, @swabey2002 demonstrated that, like other spoken languages such as Russian, Chinese, and Japanese, American Sign Language (ASL) signers use bare nouns for introduced and re-introduced discourse contexts whereas null marking is used for referents which are "in-focus" or very accessible to the addressee.
+Building on the work by @gundel1993, @swabey2002 demonstrated that, like other languages such as Russian, Chinese, and Japanese, American Sign Language (ASL) signers use bare nouns for introduced and re-introduced discourse contexts whereas zero anaphora is used for referents which are "in-focus" or very accessible to the addressee.
 Overt REs like pronouns or nouns are disprefered for such referents.
-@swabey2002 concluded that "native" signers in her study abided by the Gricean maxim of quantity and were thus economic in their reference tracking.
+@swabey2002 concluded that "native" signers in her study abided by the Gricean maxim of quantity and thus used more economic forms of reference tracking.
 Similar findings have been reported for German Sign Language (Deutsche Gebärdensprache - DGS) [@perniss2015] and British Sign Language (BSL) [@morgan2005].
-It is important to note, however, that these research either collapsed native and late signers into a single group and labeled them as "native-like" or explicitly compared native deaf signers with late deaf signers (with language deprivation) in terms of narrative competence.
+It is important to note, however, that this research either collapsed DoD and DoH signers into a single group and labeled them as "native-like" or explicitly compared native deaf signers with late deaf signers (with delayed exposure to a sign language) in terms of narrative competence.
 
@@
 
-The linguistic inventory for discourse referents can include nominal constructions (bare nouns, fingerspelled nouns, and modified nouns), null markers (plain verb, agreement verb, and constructed action), classifier predicates (whole entity classifiers, body part classifiers, handling classifiers, and extension classifiers), and overt pronouns (pronominal index or IX) [@frederiksen2016].
-This interim section describes such constructions attested for sign languages under four main headings (nominals, pronouns, classifiers, and null marking).
+The linguistic inventory for discourse referents can include nominal constructions (bare nouns, fingerspelled nouns, and modified nouns), null or zero anaphors\footnote{Zero anaphora does not refer to the lack of a referent.
+It only alludes to the non-presence of a more explicit anaphoric form for the referent such as a nominal, pronominal, or classifiers.} (predicated by a plain verb, an agreement verb, or constructed action), classifier predicates (whole entity classifiers, body part classifiers, handling classifiers, and extension classifiers), and overt pronouns (pronominal index or IX) [@frederiksen2016].
+This interim section describes such forms attested for sign languages under four main headings (nominals, pronouns, classifiers, and zero anaphora).
 
@@
 
-In TİD and many other sign languages, it is possible to use bare noun phrases as in (\ref{ex:6a}) as well as fingerspelled nouns as in (\ref{ex:6c}).
-Fingerspelling refers to the use of the manual alphabet which is derived or borrowed from a surrounding written language [@kubus2008].
-Nouns can also be modified by classifiers as in (\ref{ex:6b}) [@frederiksen2016] or by a prenominal or postnominal IX functioning as a determiner or a demonstrative as in (\ref{ex:6d}) [@maclaughlin1997].
-
-### Pronouns
-
-The IX sign (extended index finger) has various functions in sign languages, one of which is the pronominal use which refers to established nominals by pointing to a previously assigned abstract or physical locus in the signing space when re-introducing a topic in discourse [@emmorey1996].
-The same IX function has also been identified for TİD [@sevinc2006].
-The example in (\ref{ex:7}) illustrates a spatially-anchored pronominal reference for topic re-introduction.
-These signs are thus rarely used for the first mention of an entity [@perniss2015].
-Other interpretations consider pointing signs without noun phrases as examples of demonstratives and not personal pronouns [@koulidobrova2016].
-It is also important to mention here that in some spoken languages (e.g., German and Dutch), demonstratives can also refer to persons [@bosch2007] and behave similarly to personal pronouns.
-However, for the purposes of the present study, we will refer to IX constructions as pronominal IX and particularly examine their non-deictic or personal pronoun realizations as a possible discourse strategy to mainly mark topic shift among native and late signers of TİD.
-
-### Classifiers
-
-Classifier predicates are highly complex morphological constructions which are very common in sign languages [@supalla1986].
-These are manually signed iconic representations of animate or inanimate objects, and they embody several properties pertaining to the nominal referent including its location in the signing space, movement, path, size, and shape [@talmy2003].
-As in most sign languages hitherto studied, classifier constructions are also commonly found in TİD [see @dikyuva2015 for a review].
-Following the classification of @engberg-pedersen1993, we distinguish between four types of classifier constructions.
-Whole entity classifiers (WECL) represent the entirety of any referent whereas body part classifiers (BPCL) only reflect a specific part of an animate entity (e.g., limbs, head, tongue, teeth, and so forth).
-Handling classifiers (HCL[^1]) occur with transitive verbs and depict how an object is handled or manipulated by an agent.
-Extension classifiers (ExtCL), on the other hand, denote the size and shape of an entity by following a trajectory to trace its shape.
-We exemplify these constructions in (\ref{ex:8}).
-
-[^1]: As discussed before, an HCL is transitive in nature and depicts the handling action performed by ananimate agent.
-This is why most research tend to consider HCLs as an extension to constructed action [@dudis2004; @cormier2013].
+In TİD and many other sign languages, it is possible to use bare nouns as in \ref{ex:2a} as well as fingerspelled nouns as in \ref{ex:2b}.
+Fingerspelling refers to the use of the manual aphabet to spell all or part of the letters of a word borrowed from a surrounding written language [@kubus2008].
+Nouns can also be modified by a prenominal or postnominal extended index finger pointing sign, glossed as IX, functioning as a determiner or a demonstrative as in \ref{ex:2c} [@maclaughlin1997] or by a modifying classifier within the same NP as in \ref{ex:2d} [@frederiksen2016].\footnote{Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions (Lillo-Martin and Klima, 1990).
+In addition, the pointing sign is used to referentially anchor a referent to a location in signing space.}
 
 ```{=tex}
-\begin{multicols}{2}
+\newpage
 \begin{exe}
@@
 \begin{xlist}
-\ex \label{ex:6a} Full noun \\\glll
-\includegraphics[width=50pt]{pictures/1a_1.png}\\ \textsc{mouse\footnotemark} \\ \\ 
-\ex \label{ex:6b} Noun modified by classifier \\\glll
-\includegraphics[width=50pt]{pictures/1c_2.png} \includegraphics[width=50pt]{pictures/1c_1.png} \\ \textsc{soap} \textsc{soap.cl} \\ \\ 
-\ex \label{ex:6c} Fingerspelled noun\\\glll 
-\includegraphics[width=50pt]{pictures/1b_1.png} \includegraphics[width=50pt]{pictures/1b_2.png} \includegraphics[width=50pt]{pictures/1b_3.png}\\ \textsc{t-} \textsc{o-} \textsc{m} \\ \\
-\ex \label{ex:6d} Noun modified by index (\textsc{IX})\\\glll 
-\includegraphics[width=50pt]{pictures/1d_1.png} \includegraphics[width=50pt]{pictures/1d_2.png}\\ \textsc{ix} \textsc{cat} \\ \\
+\ex \label{ex:2a} Full noun \\\glll
+{} \includegraphics[width=35pt]{pictures/p1.png} {} \\ \textbf{Gloss:} \textsc{mouse\footnotemark} \textsc{sick} \\ 
+\textbf{Status:} Intro. \\
+\glt \textbf{Context:} Narration 6, Event 1. Here, the signer introduces the mouse, Jerry, at the beginning of the narration with a bare noun.
+\glt \textbf{Translation:} There is a sick mouse.
+\vspace{0.35cm}
+
+\ex \label{ex:2b} Fingerspelled noun \\\glll
+{} \includegraphics[width=40pt]{pictures/p2.1.png} \includegraphics[width=40pt]{pictures/p2.2.png} \includegraphics[width=40pt]{pictures/p2.3.png} {} \\ \textbf{Gloss:} \textsc{t-} \textsc{-o-} \textsc{-m} \textsc{come.back} \\
+\textbf{Status:} Re-intro. \\
+\glt \textbf{Context:} Narration 7, Event 2.
+The signer introduces Tom two sentences before.
+She introduces the black cat in the preceding clause and then here re-introduces Tom by using fingerspelling.
+\glt \textbf{Translation:} Tom comes back.
+\vspace{0.35cm}
+
+\ex \label{ex:2c} Noun modified by index (IX) \\\glll
+{} \includegraphics[width=45pt]{pictures/p3.1.png} \includegraphics[width=40pt]{pictures/p3.2.png} {} {} {} \\ \textbf{Gloss:} \textsc{ix} \textsc{cat} \textsc{pipe.extend} \textsc{go.through} \textsc{exit} \\
+\textbf{Status:} Re-intro. \\
+\glt \textbf{Context:} Narration 1, Event 6. The signer introduces the cat in the first clause of the first event and re-introduces it several times including this last event of the narrative.
+\glt \textbf{Translation:} The cat goes through and exits the thin pipe.
+\vspace{0.35cm}
+
+\ex \label{ex:2d} Noun modified by classifier (CL) \\\glll
+{} \includegraphics[width=35pt]{pictures/p4.1.png} \includegraphics[width=35pt]{pictures/p4.2.png} \\ \textbf{Gloss:} \textsc{soap} \textsc{soap.cl} \\ 
+\textbf{Status:} Intro. \\
+\glt \textbf{Context:} Narration 5, Event 3. Here the signer introduces the soap to the discourse. The signer produces the lexical sign, \textsc{soap}, plus a classifier, \textsc{soap.cl}.
+\glt \textbf{Translation:} The soap, which is like this.
 \end{xlist}
 \end{exe}
-\end{multicols}
+
+\footnotetext{We use small capitals to indicate sign glosses.}
+```
+### Referential Use of Space and Pronouns
+
+In addition to being a modifier of a noun, the IX sign can also be pronominal by referring to established nominals through pointing to a previously assigned abstract or physical locus in the signing space when re-introducing a topic in discourse [@emmorey1996].
+This function has also been identified for TİD [@sevinc2006].
+The example in \ref{ex:3} illustrates this spatially-anchored pronominal use.
+
+```{=tex}
 \ea 
-\ea \label{ex:7} Pronominal \\\glll
-\includegraphics[width=50pt]{pictures/2_1.png} \includegraphics[width=50pt]{pictures/2_2.png} \\ \textsc{ix} \textsc{sick} \\ \\
+\ea \label{ex:3} Pronominal \\\glll
+{} \includegraphics[width=35pt]{pictures/p5.png} {} {} {} {} \\ \textbf{Gloss:} \textsc{ix} \textsc{look} \textsc{cat.sad(constructed action)} \textsc{sad} \textsc{sick} \\ 
+\textbf{Status:} Re-intro. \\
+\glt \textbf{Context:} Narration 9, Event 1: The signer introduces the mouse then signs an action of the cat. Here, he re-introduces the mouse with a pointing sign that functions as a pronoun.
+\glt \textbf{Translation:} It (the mouse) looks sad and sick.
 \z \z
-\clearpage
-\begin{multicols}{4}
-\begin{exe}
-\ex \label{ex:8}
-\begin{xlist}
-\ex Whole entity \\\glll 
-\includegraphics[width=50pt]{pictures/3e.png} \\ \textsc{fish.fall}\\ \\
-\ex Body part \\\glll
-\includegraphics[width=50pt]{pictures/3b.png} \\ \textsc{walk} \\ \\ 
-\ex Handling \\\glll 
-\includegraphics[width=50pt]{pictures/3c.png} \\ \textsc{hold} \\ \\
-\ex Extension \\\glll 
-\includegraphics[width=50pt]{pictures/3d_1.png} \includegraphics[width=50pt]{pictures/3a.png}\\ \textsc{plank} {\textsc{dh: mouse}} \\ {} {\textsc{ndh: tunnel}}\\ \\
-\end{xlist}
-\end{exe}
-\end{multicols}
-\begin{multicols}{3}
+```
+These pronominal signs are rarely used for the first mention of an entity [@perniss2015].
+Other researchers consider pointing signs without noun phrases as examples of demonstratives and not personal pronouns [@koulidobrova2016].
+It is also important to mention here that in some spoken languages (e.g., German and Dutch), demonstratives can also refer to persons [@bosch2007] and behave similarly to personal pronouns.
+However, for the purposes of the present study, we will refer to IX signs as pronominal IX and particularly examine their non-deictic or personal pronoun realizations as a possible discourse strategy to mainly mark topic shift (re-introduction) among DoD and DoH signers of TİD.
+
+### Classifiers
+
+In sign languages, a verbal classifier morpheme is morpho-phonologically encoded by a handshape.
+This handshape is articulated co-temporally with a movement that encodes the verbal meaning [@supalla1990; @benedicto2004].
+Semantically, the classifiers denote sets of objects based on criteria such as animacy, size, shape, body-part, texture and sometimes function [@talmy2003].
+The classifier articulated simultaneously with a movement is called a classifier predicate and the sentence in which such a predicate is used is usually called a classifier construction in sign language research.
+As in most sign languages hitherto studied, classifier constructions are also commonly found in TİD [see @dikyuva2015 for a review].
+Following the classification of [@engberg-pedersen1993], in TİD, we distinguish between four types of classifier handshapes.
+Whole entity classifiers (WECL) represent the entirety of a referent (e.g.
+an entire flat horizontal entity, or an upright entity) (e.g., \ref{ex:4a}) whereas body part classifiers (BPCL) represent a specific part of an animate entity (e.g., limbs, head, mouth, tongue, or teeth, etc.) (e.g.,\ref{ex:4b}).
+Handling classifiers (HCL) occur with transitive verbal roots and represent both the size, shape and surface of an object and how that object is handled or manipulated by an agent (e.g., \ref{ex:4c}).
+Extension classifiers (ExtCL), on the other hand, represent the size and shape of an entity by following a trajectory to trace its shape (e.g., \ref{ex:4d}).
+We exemplify these handshapes with the discourse status of the referent they track and a description of the previous discourse when available.
+
+```{=tex}
 \begin{exe}
@@
 \begin{xlist}
-\ex \label{ex:9a} Constructed action \\\glll 
-\includegraphics[width=50pt]{pictures/4c.png} \\ \textsc{struggle} \\ \\
-
-\ex \label{ex:9b} Plain verb  \\\glll
-\includegraphics[width=50pt]{pictures/4a_2.png}  \\ \textsc{enter}\\ \\
-
-\ex \label{ex:9c} Agreement verb  \\\glll 
-\includegraphics[width=50pt]{pictures/4b_1.png} 
-\includegraphics[width=50pt]{pictures/4b_2.png} \\ {\textsc{$_{1}$hit$_{3}$}} {} \\ \\ 
+\ex \label{ex:4a} Whole entity classifier  (WECL) \\\glll
+{} \includegraphics[width=40pt]{pictures/p6.png} \\ \textbf{Gloss:} \textsc{two.mice.go.together} \\ 
+\textbf{Status:} Maintain. \\
+\glt \textbf{Context:} Narration 4, Event 2: The signer introduces Jerry and the robot mouse in the preceding clause. She maintains reference to them by using a whole entity classifier on each hand. 
+\glt \textbf{Translation:} (Jerry and the robot mouse from the previous context) go together.
+\vspace{0.35cm}
+
+\ex \label{ex:4b} Body part classifier (BPCL) \\\glll
+{} \includegraphics[width=40pt]{pictures/p7.png} \\ \textbf{Gloss:} \textsc{cat.move.thin.legs} \\ 
+\textbf{Status:} Maintain. \\
+\glt \textbf{Context:} Narration 1, Event 6: The signer re-introduces the cat three clauses before and maintains reference to him in the two clauses before and here with a body part classifier that refers to the legs of the cat.
+\glt \textbf{Translation:} (The cat) walks (i.e. moving his thin legs forward).
+\vspace{0.35cm}
+
+\ex \label{ex:4c} Handling classifier (HCL) \\\glll
+{} {} {} {} {} \includegraphics[width=40pt]{pictures/p8.png} \\ \textbf{Gloss:} {} \textsc{rope see-3a} \textsc{soon} \textsc{go} \textsc{cat.pull.rope} \\ 
+\textbf{Status:} Maintain. \\
+\glt \textbf{Context:} Narration 5, Event 2: The signer introduces Tom three clauses before.
+She maintains reference to him in the preceding two clauses and here she still maintains reference to him by using a handling classifier on both hands.
+\glt \textbf{Translation:} When (he/Tom) sees the rope, (he) soon goes to pull it.
+\vspace{0.35cm}
+
+\ex \label{ex:4d} Extension classifier (ExtCL) \\\glll
+{} \includegraphics[width=80pt]{pictures/p9.png} \\ \textbf{Gloss:} \textsc{cat.extend} \\ 
+\textbf{Status:} Maintain. \\
+\glt \textbf{Context:} Narration 1, Event 6: The signer re-introduces Tom two clauses before and she maintains reference to him with constructed action in the previous clause and here with an extension classifier.
+\glt \textbf{Translation:} (The cat) extends thin.
 \end{xlist}
 \end{exe}
-\end{multicols}
-\footnotetext{We use small capitals to indicate sign glosses.}
-```
-### Null marking
-
-Parallel to spoken languages, many sign languages have been reported to allow null subjects and/or objects and make use of null arguments when referring to highly accessible referents which are still active in the mental representation of the addressee.
+```
+### Zero Anaphora
+
+Parallel to spoken languages, many sign languages have been reported to allow null subjects and/or objects and make use of null arguments when referring to highly accessible referents that are still active in the mental representation of the addressee.
 In signed discourse, this can be achieved by employing certain verbal constructions or constructed action (CA).
-The latter construction exemplified in (\ref{ex:9a}) refers to cases in which the signers engage with the space in front of themselves as if they are interacting with the objects and characters in the story.
+The latter construction exemplified in \ref{ex:5a} refers to cases in which the signers engage with the space in front of themselves as if they are characters in the story interacting with objects and other characters.
 This is often achieved with the signer taking on the role of a referent in the narrative and performing the actions [@metzger1995].
@@
 As for the verbal constructions, to identify the predicate types in ASL, @padden1986 suggested a triadic categorization of verbs in sign languages: plain verbs, agreement verbs, and spatial verbs.
-According to this classification, plain verbs (\ref{ex:9b}) do not agree with the subject or the object.
+According to this classification, plain verbs (as in \ref{ex:5b}) do not agree with the subject or the object.
 This means that they are not inflected for number or person.
-This contrasts with agreeing verbs (\ref{ex:9c}) the starting or the ending point of which can be modified to specific loci for person and number agreement while spatial verbs (as in \textsc{ix book put} 'I placed the book there') agree with locative arguments but like plain verbs, they are not inflected for number or person.
+This contrasts with agreeing verbs (as in \ref{ex:5c}) the starting or the ending point of which can be modified to specific loci for person and number agreement while spatial verbs (as in \textsc{ix book put} 'I placed the book there') agree with locative arguments but like plain verbs, they are not inflected for number or person.
 TİD too licenses null arguments.
@@
 
-## Signed vs. Spoken REs
-
-Commonly in both spoken and sign languages nouns and noun phrases are used as REs.
-Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions [@lillo-martin1990].
-In addition, the pointing sign is used to referentially anchor a referent to a location in signing space.
-For instance, when the signer introduces \textsc{elvan} for the first time, s/he can assign her to the right.
-This referential locus can later be used by agreement verbs and classifier predicates (more below).
-For instance, in (\ref{ex:2a}) there are two referents, \textsc{elvan}, who is referentially-assigned to the right side of the signer, represented with the "A" subscript and \textsc{demet}, who is referentially-assigned to the left side of the signer, represented with the "B" subscript.
-When the signer makes a contrast in (\ref{ex:2b}), as a pronominal use, she points to the right side to refer to ELVAN and to the left to refer to \textsc{demet}.
-We schematically represent the referential locus assigned to the referents in (\ref{ex:3}).
+In terms of referential accessibility, zero anaphors predicated by such constructions in narratives are assumed to rank last in the hierarchy, preceded by nominal classifiers and then pronouns and other nominal forms.
+The reason that we distinguish zero anaphora from classifer handshapes or predicates (which can also be considered implicit markers of reference) is because classifiers typically denote certain topographic or semantic properties (e.g., size and shape) of the referent noun due to their iconic nature.
+They are not as explicit as pronouns or nouns since they do not target a specific individual (e.g., \textsc{eddie}) in a set but rather they refer to a set of individuals that are semantically similar to each other [@morgan2005].
+For example, the extended index finger can be used to refer to a human entity.
+Finally, zero anaphors predicated by constructed action, agreement and plain verbs are least explicit forms of referent tracking since the implicit nature of such constructions do not easily restrict the referent set based on cues such as size, shape or semantic category.
+In other words, zero anaphora requires preceding context or information the most compared to other forms.
 
@@
 \begin{xlist}
-    \ex \label{ex:2a}
-    \textsc{lab place elvan ix\textsubscript{a} ten years work. demet ix\textsubscript{b} last year start.} \\
-    ‘In the lab, Elvan has been working for 10 years. Demet started last year.’
-    \\
-    \ex \label{ex:2b}
-    \textsc{ix\textsubscript{a} e-l-a-n know. ix\textsubscript{b} learn.}\\
-    ‘She (Elvan) knows ELAN (the annotation program). She (Demet) is learning.’
+\ex \label{ex:5a} Constructed action \\\glll
+{} \includegraphics[width=40pt]{pictures/p10.png} \\ \textbf{Gloss:} \textsc{mouse.stay.like.this} \\ 
+\textbf{Status:} Maintained. \\
+\glt \textbf{Context:} Narration 6, Event 9: The signer re-introduces Jerry in the previous clause and maintains reference to him by using this constructed action.
+\glt \textbf{Translation:} (The mouse) stays like this.
+\vspace{0.35cm}
+
+\ex \label{ex:5b} Plain verb \\\glll
+{} \includegraphics[width=80pt]{pictures/p11.png} \\ \textbf{Gloss:} \textsc{run.away} \\ 
+\textbf{Status:} Re-introduced. \\
+\glt \textbf{Context:} Narration 5, Event 4: In this sentence, the signer re-introduces Jerry with the plain verb \textsc{run.away}.
+\glt \textbf{Translation:} (Jerry) runs away.
+\vspace{0.35cm}
+
+\ex \label{ex:5c} Agreement verb \\\glll
+{} \includegraphics[width=80pt]{pictures/p12.png} \\ \textbf{Gloss:} \textsc{3a(mouse)-hit-3b(cat)} \\ 
+\textbf{Status:} Maintained. \\
+\glt \textbf{Context:} Narration 8, Event 2: The signer introduces Jerry in the preceding clause and maintains reference to him by using the agreement verb \textsc{hit}.
+\glt \textbf{Translation:} (Jerry) hits (Tom).
+\vspace{0.35cm}
+
 \end{xlist}
 \end{exe}
-
-    \begin{exe}
-    \ex \label{ex:3} Signing space, signer, example referential loci for the referents \textsc{elvan} (a) and \textsc{demet} (b) \\
-\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
-
-\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
-%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300
-
-%Curve Lines [id:da7006577339157659] 
-\draw    (100.25,77.81) .. controls (136.19,41.85) and (183.62,42.17) .. (222.5,78.13) ;
-%Straight Lines [id:da8428827309782937] 
-\draw    (112.9,128.8) -- (210.21,128.8) ;
-%Curve Lines [id:da39290455666055124] 
-\draw    (222.19,100.13) .. controls (221.89,103.9) and (217.55,123.77) .. (210.21,128.8) ;
-%Curve Lines [id:da9909903731270634] 
-\draw    (101.17,100.45) .. controls (100.86,104.22) and (108.2,124.02) .. (112.9,128.8) ;
-
-% Text Node
-\draw (88.75,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-b }};
-% Text Node
-\draw (91.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Demet)}};
-% Text Node
-\draw (208.25,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-a }};
-% Text Node
-\draw (212.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Elvan)}};
-% Text Node
-\draw (144,115.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Signer}};
-\end{tikzpicture}
-    \end{exe}
-```
-Grammatical gender as in Romance languages or noun classes as in Bantu languages are some reference tracking tools that are familiar from spoken languages.
-A somehow less known reference-tracking tool in spoken languages is *verbal classifiers*, which morphologically attach to the verb and often cross-reference a core argument in a clause [@aikhenvald2017].
-As outlined in the previous section, classifiers denote sets of objects based on semantic criteria such as size, shape, body-part, texture and sometimes function.
-Usually, the verbal classifier and the cross-referenced noun co-occur in the same clause when the argument is introduced into discourse.
-The classifier can track the argument on its own on subsequent occurrences.
-In sign languages, a verbal classifier morpheme is encoded by a *handshape* that is articulated phonologically co-temporally with a movement, which encodes the verbal meaning [@supalla1990; @benedicto2004].
-These two infused morphemes form a complex predicate.
-Some classifier handshapes and the set of objects that they denote are presented below in (\ref{ex:4}).
-
-```{=tex}
-\begin{exe}
-\ex \label{ex:4}
-\includegraphics[width=25pt]{pictures/flathand.png}: flat entities, surfaces, vehicles.\\
-\includegraphics[width=25pt]{pictures/flat.png}: human agent holding a flat entity.\\
-\includegraphics[width=25pt]{pictures/small_legged.png}: small animal legs.\\
-\includegraphics[width=25pt]{pictures/index.png}: thin entities (like a pencil); upright beings.\\
-\includegraphics[width=25pt]{pictures/hold.png}: human agent holding a thin entity (like a pencil or a branch of flower). 
-\end{exe}
-```
-In (\ref{ex:5}), in the first clause, a referential locus, i.e.
-locus-A, is assigned to the locative source argument \textsc{cage}.
-In the second clause, the core argument \textsc{bird} is cross-referenced with the classifier handshape, i.e.
-small-legged animal.
-It flies from the referential locus of the \textsc{cage} (locus-a) to the referential locus of the goal \textsc{tree} (locus-B).
-
-```{=tex}
-\begin{exe}
-    \ex \label{ex:5}
-\begin{xlist}
-    \ex \textsc{cage} assigned to the referential locus-a. \\ \\
-    \textsc{garden bird cage-a open} \\
-    ‘In the garden, there is a cage and it is open.’
-    \\
-
-    \ex \textsc{tree} assigned to the referential locus-b. The source and goal tracked on the predicates \textsc{fly} and \textsc{land}. \\ \\
-    \textsc{bird fly[manner]-a tree-b land cl-b}\\
-    ‘The bird flies (from the cage assigned to the left) to a tree (assigned to the right).’
-    \end{xlist}
-    \end{exe}
-    \hfill (Adapted from Gökgöz and Sevgi 2020:318)
 ```
@@
 Like in first language development, narrative skills have been extensively studied among second language learners [see @bel2015; @frederiksen2019].
-Comparing 11 deaf adult signers and 13 hearing advanced learners of Catalan Sign Language (Llengua de Signes Catalana - LSC), @bel2015 found that hearing signers resort to more overt forms (nominals and pronouns) than native signers to avoid potential ambiguities among referents.
+Comparing 11 deaf adult signers and 13 hearing advanced learners of Catalan Sign Language (Llengua de Signes Catalana - LSC), @bel2015 found that hearing signers resort to more overt forms (nominals and pronouns) than native deaf signers to avoid potential ambiguities among referents.
 Nevertheless, their results did not reach statistical significance for the interaction of discourse status and acquisition group.
-A similar line of research came from @frederiksen2019 who found a statistically significant difference between both groups, that is native signers and hearing L2 learners of ASL, only for classifier use especially in re-introduced contexts.
-L2 learners in the study used more nominals and fewer instances of null marking in number, indicative of a limited effect of overexplicitness among second language learners.
-However, these studies treated late and native deaf signers as a single group.
-That is, the control group consisted of both native signers who had one or more deaf parents and also native-like [@cormier2012] late deaf signers who had hearing parents and their age of acquisition varied.
-Given the preliminary evidence that topic continuity and salience marking is also difficult for second language learners regardless of modality, it has been suggested under the Interface Hypothesis that the syntax-discourse interface is vulnerable and especially challenging to acquire [@sorace2011] for second language learners.
-Similarly, @sorace2009 discuss that this interface entails much more than morphosyntactic knowledge but also includes cognitive and pragmatic resources that help provide discourse coherence.
-They propose that the second language discourse is over-redundant.
-This means that second language learners produce more overt pronouns or fuller forms of language for highly accessible referents possibly as a compensatory strategy in an attempt to reduce the increasing load on cognitive processing.
-
-Still, only very few researchers examined whether linguistically deprived signers (i.e., deaf signers with delayed exposure to a sign language) also display redundancy in their reference tracking.
+However, the control group here consisted of both deaf signers who had one or more deaf parents and also deaf signers who had hearing parents and their age of acquisition varied.
+Out of 11 deaf L1 signers in Bel et al.'s [-@bel2015] control sample, 5 had deaf parents and acquired LSC from birth, and the other 6 had hearing parents but their exposure to LSC started before age 5.
+A similar line of research came from @frederiksen2019 who found a statistically significant difference between both groups, that is DoD signers and hearing L2 learners of ASL, only for classifier use especially in re-introduced contexts.
+L2 learners in the study used more nominals and fewer instances of zero anaphora in number, indicative of a limited effect of overexplicitness among second language learners.
+Although Frederiksen and Mayberry's control group included some signers with hearing parents, they had all learned to sign from birth since they received sign language input from older deaf siblings.
+
+Still, only very few researchers explicitly compared DoD (i.e., deaf signers with native exposure to a sign language) and DoH (i.e., deaf signers with delayed exposure to a sign language) also display redundancy in their reference tracking.
 @becker2009 examined 6 DGS signing deaf children's (3 native and 3 late) narrative abilities.
 She reported that late signing children generated more explicit forms such as nominals when referring to the characters in the story compared to native learners, concluding that first language deprivation results in an inadequate representation of the addressee and less efficient use of morphosyntax.
-Comparably, @cormier2013 investigated the use of CA in BSL among 15 deaf native, early, and late adult signers (5 participants in each group) and found that late signers were redundant in topic continuity.
-In other words, the late signers produced more explicit forms like nominals, and fewer instances of null marking when maintaining previous referents.
+Comparably, @cormier2013 investigated the use of CA in BSL among 15 deaf native, early, and late adult signers (5 participants in each group) and found that DoH signers were redundant in topic continuity.
+In other words, the DoH signers produced more explicit forms like nominals, and fewer instances of zero anaphora when maintaining previous referents.
 Most relevantly to the present study, @gur2018 studied the development of narrative skills among 10 native and 10 late deaf children and deaf adult signers of TİD.
 She found that native-signing children and adults had more instances of explicit referent introduction, set the scene more frequently, and employed CA in an appropriate manner more than corresponding late-signing children and adults.
-Although this study presented preliminary findings on age of acquisition effects on narrative competence, it does not investigate how native and late signers employ different linguistic tools based on the accessibility of the referent.
-Therefore, to our knowledge, there is a lack of research directly examining the effects of first language deprivation on deaf adult signers' selection of REs in signed narratives.
-
-Considering this, it is not very clear whether the negative effects on narrative abilities observed among late signing deaf children hold when they become adults.
-Understanding the impact of language deprivation or delayed first language learning on narrative competence is crucial for three reasons.
-First, studying this population will shed light on the possible effect of lack of early language experience on future cognitive and pragmatic abilities since the narrative competence is dependent on such faculties.
-Second, examining reference tracking strategies in particular will also inform about the linguistic inventory that native and late deaf signers use to refer to entities.
+Although this study presented preliminary findings on age of acquisition effects on narrative competence, it does not investigate how DoD and DoH signers employ different linguistic tools based on the accessibility of the referent.
+Therefore, to our knowledge, there is a lack of research directly examining the effects of first language delay especially among deaf adult signers with non-signing hearing parents on the selection of REs in signed narratives.
+
+Considering this, it is also not very clear whether the age of acquisition effects on narrative abilities, observed among late signing deaf children, hold when they become adults.
+Understanding the impact of delayed first language learning on narrative competence is crucial for three reasons.
+First, studying signers from this population will shed light on the possible effect of lack of early language experience on future cognitive and pragmatic abilities since the narrative competence is dependent on such faculties.
+Second, examining reference tracking strategies in particular will also inform about the linguistic inventory that DoD and DoH signers use to refer to entities.
 Finally, any possible difference between the two acquisition groups in terms of narrative competence will have further practical implications for the importance of early access to a first functional sign language to retain future linguistic and cognitive abilities.
@@
 
-In an attempt at addressing this gap in the literature, the present study aims to investigate the reference tracking strategies of native and late adult signers in TİD narratives with a story-telling paradigm.
-We explore the forms available for referent tracking in TİD and examine the nature of RE selection as a possible indicator of pragmatic competence between our acquisition groups.
-
-Overall, considering what has been suggested for the distribution of REs in signed narratives, we expect more explicit forms like nominals for referent introduction and and less explicit forms such as null marking, pronominal and/or classifier use for referent maintainance.
+In an attempt at addressing this gap in the literature, the present study aims to investigate the reference tracking strategies of DoD and DoH adult signers in TİD narratives with a story-telling paradigm.
+We explore the forms available for referent tracking in TİD and examine the nature of RE selection as an indicator of possible pragmatic differences between the acquisition groups.
+
+Overall, considering what has been suggested for the distribution of REs in signed narratives, we expect more explicit forms like nominals for referent introduction and and less explicit forms such as zero anaphora, pronominal and/or classifier use for referent maintainance.
 For the re-introduction contexts, where referents which are discontinued across one or more clauses are brought back into the discourse, we expect a less clear picture.
 Some re-introduced referents will have higher accessibility and will thus be tracked with more implicit REs, while others that have lower accessibility will be tracked with more explicit REs.
-Also, based on previous research, we hypothesize that late signers will be more over-redundant in their reference tracking compared to native signers.
-This would translate to a preference by late signers for nominals over null marking especially in maintenance and re-introduction contexts.
-In contrast, we would expect that native signers use more null marking to track referents.
+Also, based on previous research, we hypothesize that DoH signers will be more over-redundant in their reference tracking compared to DoD signers.
+This would translate to a preference by DoH signers for nominals over zero anaphora especially in maintenance and re-introduction contexts.
+In contrast, we would expect that DoD signers use more zero anaphora to track referents.
 
@@
 
-We tested 29 deaf adult signers, 15 with exposure to TİD from birth (i.e., native signers), 14 with exposure between ages 3 and 17 (i.e., late signers) (*M*\textsubscript{Age of acquisition} = 7.7 years, *SD* = 3.3 years).
-The criterion to be classified as a late signer was to have hearing and non-signing caregivers.
-All the signers satisfying this criterion were regarded as late signers \footnote{Note that we are essentially comparing deaf of deaf and deaf of hearing adult signers.
-Although we did not set a cut-off point for nativeness, all the deaf of hearing participants indicated that they learned TİD after age 3.
-In order to capture in-group variation among late signers, we run an additional supplementary analysis in which we observed very wide credible intervals due to the small group size, rendering the results uninformative.
-For the analysis, see \ref{section:d}.}.
-All participants indicated that their preferred language of communication was TİD which they used for most contexts.
+We tested 29 deaf adult signers, 15 DoD signers with exposure to TİD from birth from deaf caregivers (e.g., parents or siblings), 14 DoH signers who had hearing caregivers and were exposed to TİD between ages 3 and 17.
+(*M*\textsubscript{Age of acquisition} = 7.7 years, *SD* = 3.3 years).
+As a result, the criterion to be classified as a DoH signer was to have hearing and non-signing caregivers.
+All the signers satisfying this criterion were regarded as such.\footnote{Although we did not set a cut-off point for nativeness, all the deaf of hearing participants indicated that they learned TİD after age 3.
+In order to capture in-group variation among DoH signers, we run an additional supplementary analysis in which we observed very wide credible intervals due to the small group size, rendering the results uninformative.
+For the analysis, see \ref{section:d}.} All participants indicated that their preferred language of communication was TİD which they used for most contexts.
 Detailed information about the linguistic and educational background of the participants can be found in \ref{section:a}.
-
-All participating native signers (7 males, 8 females) were born into deaf families and started acquiring TİD from at least one of their caregivers after birth.
-The parents of 14 participating native signers were deaf.
+1 DoD signer was excluded from the data analysis because they self-reported to have signing deaf parents but also only used home sign (a more crude form of communication) at home.
+As a result, we present data from only 14 DoD signers.
+
+All participating DoD signers (7 males, 7 females) were born into deaf signing families and started acquiring TİD from at least one of their caregivers after birth.
+The parents of 14 participating DoD signers were deaf.
 One native signer had a deaf father and a hearing mother.
-In addition to TİD, the home language for some native signers (n = 4) in early childhood included the use of home sign\footnote{Home signs are gesture-based communication systems that emerge between the deaf children and their primary caregivers.
+In addition to TİD, the home language for some DoD signers (n = 4) in early childhood included the use of home sign.\footnote{Home signs are gesture-based communication systems that emerge between the deaf children and their primary caregivers.
 Although home sign communication may have internal consistency, they are not fully-flegged linguistic systems like natural languages.}.
-The age of the native participants ranged between 18 and 35 (*M* = 26.4 years, *SD* = 5.0 years).
-This also corresponded to the total average years of TİD use for native signers.
-All native signers were pre-lingually deaf, that is they were either born deaf or they had become deaf before the age of 3, with the exception of one signer who became deaf between the ages of 4 and 7.
-Native signers reported to complete at least high school education and their level of education (LOE) was between 12 and 16 years (*M* = 12.7 years, *SD* = 1.5 years).
+The age of the native participants ranged between 18 and 35 (*M* = 26.4 years, *SD* = 5.0 years) This also corresponded to the total average years of TİD use for DoD signers.
+All DoD signers were pre-lingually deaf, that is they were either born deaf or they had become deaf before the age of 3, with the exception of one signer who became deaf between the ages of 4 and 7.
+DoD signers reported to complete at least high school education and their level of education (LOE) was between 12 and 16 years (*M* = 12.7 years, *SD* = 1.5 years).
 In addition, all native participants attended at least two deaf schools, one being primary and the other middle school.
@@
 
-Comparably, late learners (8 males, 6 females) were all born into hearing families and started learning TİD later in life after age 3, usually with the start of formal education.
-The age of sign language acquisition for late signers ranged between years 3-17 (*M* = 8.7 years, *SD* = 3.5 years).
-The parents of all participating late signers were hearing, and the reported home languages used in childhood mainly consisted of Turkish and home sign.
-The average age of late signers was 34.6 years (*SD* = 8 years) and ranged between 24 and 50.
-The derived mean number of years for TİD use following its acquisition was comparable to native signers with 25.9 years (*SD* = 7.9 years).
-Except for two signers who became deaf between the ages of 4 and 7, all the other late signers were pre-lingually deaf.
-All late signers completed at least middle school education and the year of schooling ranged between 8 and 16 years, and the average was 12.3 years (*SD* = 2.8 years).
-All late participants attended at least one deaf school.
+Comparably, DoH signers (8 males, 6 females) were all born into hearing non-signing families and started learning TİD later in life after age 3, usually with the start of formal education.
+The age of sign language acquisition for DoH signers ranged between years 3-17 (*M* = 8.7 years, *SD* = 3.5 years).
+The caregivers of all participating DoH signers were hearing, and the reported home languages used in childhood mainly consisted of Turkish and home sign.
+The average age of DoH signers was 34.6 years (*SD* = 8 years) and ranged between 24 and 50.
+The derived mean number of years for TİD use following its acquisition was comparable to DoD signers with 25.9 years (*SD* = 7.9 years).
+Except for two signers who became deaf between the ages of 4 and 7, all the other DoH signers were pre-lingually deaf.
+All DoH signers completed at least middle school education and the year of schooling ranged between 8 and 16 years, and the average was 12.3 years (*SD* = 2.8 years).
+All of the DoH group attended at least one deaf school.
 While 8 participants attended two deaf schools (either primary and middle school, or middle and high school), 3 participants reported to attend three deaf schools.
-Using the same scale, late signers self-rated their abilities in TİD and Turkish.
+Using the same scale, the signers self-rated their abilities in TİD and Turkish.
 Both of their receptive and productive TİD abilities rated on average 4.9 points (*SD* = 0.3 points).
-Late signers' mean rating for their reading abilities in Turkish was 3.5 points (*SD* = 0.7 points) while their writing and speaking abilities rated on average 3.4 (*SD* = 0.5 points) and 3.1 points (*SD* = 0.7 points), respectively.
-
-Native and late participants' sex, average years of schooling, TİD use following its acquisition, and their self-reported receptive and productive TİD ratings, and their reading, writing, and speaking ratings for Turkish did not statistically differ; Bayesian *p*'s \> 0.1.
+DoH signers' mean rating for their reading abilities in Turkish was 3.5 points (*SD* = 0.7 points) while their writing and speaking abilities rated on average 3.4 (*SD* = 0.5 points) and 3.1 points (*SD* = 0.7 points), respectively.
+
+Both groups' sex, average years of schooling, TİD use following its acquisition, and their self-reported receptive and productive TİD ratings, and their reading, writing, and speaking ratings for Turkish did not statistically differ; Bayesian *p*'s \> 0.1.
 However, participants' ages statistically differed between the two acquisition groups; *t*(27) = 3.4, Bayesian *p* \< 0.01.
-Overall, the native group consisted of younger participants.
-This was to ensure that the average years of TİD use among native signers approximated those of late signers.
+Overall, the DoD group consisted of younger participants.
+This was to ensure that the average years of TİD use among DoD signers approximated those of DoH signers.
 
@@
 
-We used a story-telling paradigm to examine how native and late signers of TİD track reference in signed discourse.
+We used a story-telling paradigm to examine how DoD and DoH signers of TİD track reference in signed discourse.
 Participants were presented with 10 short video clips extracted from the cartoon *Tom and Jerry* and asked to retell them.
@@
 We instructed the participants to narrate as they would to a deaf friend.
+We gave this instruction to make sure that signers naturally sign, and simply do not try to simplify or adjust their language to the level of the addressee.
 All trials were recorded by a digital camera placed in front of the signers.
-The participants were instructed to watch the narrative clip as many times as they needed, and narrate the content of each clip either to the experimenter or the recording digital camera\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining we did not make it obligatory for participants to either narrate to the digital camera or to the experimenter for few reasons.
+The participants were instructed to watch the narrative clip as many times as they needed, and narrate the content of each clip either to the experimenter or the recording digital camera.\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining narrated to the camera.
+We did not make it obligatory for participants to  narrate to the digital camera or to the experimenter for few reasons.
 First, a native or native-like deaf experimenter was not always available, and a hearing late signer of TİD attended some sessions as the experimenter.
@@
 We ran an additional model excluding the 4 signers who narrated to the experimenter and compared it with the original model including these signers (see \ref{section:c}).
-The results imply that there is not much of a difference.}.
+The results imply that there is not much of a difference.}
 
@@
               "Fingerspelled noun", "CL noun", "Noun CL") ~ "Nominal",
-  GTR1 %in% c("Constructed action", "Constructed Action", "Plain verb", "Person agreement verb") ~ "Null",
+  GTR1 %in% c("Constructed action", "Constructed Action", "Plain verb", "Person agreement verb") ~ "Zero",
   GTR1 %in%  c("Whole entity CL", "Body part CL",
@@
   GTR1 %in% c("Point noun", "Noun Point","CL noun", "Noun CL") ~ "M NOM",
-  GTR1 %in% c("Point noun", "Noun Point","CL noun", "Noun CL") ~ "FS NOM",
+  GTR1 %in% c("Fingerspelled noun") ~ "FS NOM",
   GTR1 %in% c("Constructed action", "Constructed Action") ~ "CA",
@@
 df_narration$GTR1 %<>% as.factor() 
-df_narration$Nativeness %<>% dplyr::recode(`0` = "Late", `1` = "Native") %>% as.factor() %>% reorder.factor(new.order = c("Native","Late"))
+df_narration$Nativeness %<>% dplyr::recode(`0` = "DoH", `1` = "DoD") %>% as.factor() %>% reorder.factor(new.order = c("DoD","DoH"))
 df_narration$Discourse %<>% as.factor()
@@
 # Descriptive plots
-df_narration$ref_type2 <- df_narration$ref_type %>% recode('Classifier' = 'CL', 'Null' = 'NULL',
+df_narration$ref_type2 <- df_narration$ref_type %>% recode('Classifier' = 'CL', 'Zero' = 'ZERO',
                                                            'Pronominal' = 'PRO', 'Nominal' = 'NOM')
@@
 # Contrast coding
-df_model1$Nativeness %<>% reorder.factor(new.order = c("Native","Late"))
+df_model1$Nativeness %<>% reorder.factor(new.order = c("DoD","DoH"))
 contrasts(df_model1$Nativeness) <- contr.sum(2)/2
@@
 df_model_1$parameter %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
-                                        `b_muNull_Intercept` = "Null_Intercept",
+                                        `b_muNull_Intercept` = "Zero_Intercept",
                                         `b_muNominal_Nativeness1` = "Nom_DoD",
-                                        `b_muNull_Nativeness1` = "Null_DoD",
+                                        `b_muNull_Nativeness1` = "Zero_DoD",
                                         `b_muNominal_Discourse1` = "Nom_Intro",
@@
                                         'b_muNominal_Nativeness1:Discourse2' = 'Nom_DoD*Maintain',
-                                        'b_muNull_Nativeness1:Discourse1' = 'Null_DoD*Intro',
-                                        'b_muNull_Nativeness1:Discourse2' = 'Null_DoD*Maintain',
-                                        `b_muNull_Discourse1` = "Null_Intro",
-                                        `b_muNull_Discourse2` = "Null_Maintain")
+                                        'b_muNull_Nativeness1:Discourse1' = 'Zero_DoD*Intro',
+                                        'b_muNull_Nativeness1:Discourse2' = 'Zero_DoD*Maintain',
+                                        `b_muNull_Discourse1` = "Zero_Intro",
+                                        `b_muNull_Discourse2` = "Zero_Maintain")
 saveRDS(df_model_1, "./rds/df_model_1.rds")
@@
 df_model_3$parameter %<>% dplyr::recode(`b_muNominal_Intercept` = "Nom_Intercept",
-                                        `b_muNull_Intercept` = "Null_Intercept",
+                                        `b_muNull_Intercept` = "Zero_Intercept",
                                         `b_muNominal_Group2M1` = "Nom_Late1-Late2",
@@
                                         `b_muNominal_Group3M2:Discourse2` = 'Nom_E-L1*Maintain',
-                                        `b_muNull_Group2M1` = "Null_Late1-Late2",
-                                        `b_muNull_Group3M2` = "Null_Early-Late1",
-                                        `b_muNull_Discourse1` = "Null_Intro",
-                                        `b_muNull_Discourse2` = "Null_Maintain",
-                                        `b_muNull_Group2M1:Discourse1` = 'Null_L1-L2*Intro',
-                                        `b_muNull_Group3M2:Discourse1` = 'Null_E-L1*Intro',
-                                        `b_muNull_Group2M1:Discourse2` = 'Null_L1-L2*Maintain',
-                                        `b_muNull_Group3M2:Discourse2` = 'Null_E-L1*Maintain')
+                                        `b_muNull_Group2M1` = "Zero_Late1-Late2",
+                                        `b_muNull_Group3M2` = "Zero_Early-Late1",
+                                        `b_muNull_Discourse1` = "Zero_Intro",
+                                        `b_muNull_Discourse2` = "Zero_Maintain",
+                                        `b_muNull_Group2M1:Discourse1` = 'Zero_L1-L2*Intro',
+                                        `b_muNull_Group3M2:Discourse1` = 'Zero_E-L1*Intro',
+                                        `b_muNull_Group2M1:Discourse2` = 'Zero_L1-L2*Maintain',
+                                        `b_muNull_Group3M2:Discourse2` = 'Zero_E-L1*Maintain')
 saveRDS(df_model_3, "./rds/df_model_3.rds")
@@
 
-We coded the discourse status (Introduction, Maintenance, Re-introduction) and RE type (Nominal, Pronominal, Classifier, Null).
+We coded the discourse status (Introduction, Maintenance, Re-introduction) and RE type (Nominal, Pronominal, Classifier, Zero Anaphora).
 We used ELAN Linguistic Annotation Software [@crasborn2008] for data annotation.
@@
 To this end, the referents which had been discontinued across one or more sentences were all regarded re-introduced into the discourse.
-On a third tier, we had the RE type for a single reference with the main categories nominal, pronominal, null marking, and classifier.
+On a third tier, we had the RE type for a single reference with the main categories nominal, pronominal, zero anaphora, and classifier.
 Under nominal, we identified bare nouns, modified nouns, and fingerspelled nouns.
-For classifiers, we distinguished among three types of constructions: WECL, BPCL, and ExtCL.
-Finally, we had CA, agreement, plain tags for null constructions.
+For classifiers, we distinguished among three types of handshapes: WECL, BPCL, and ExtCL.
+Finally, we had CA, agreement, plain tags for zero anaphora.
 We were unable to observe any instance of HCLs occurring in isolation (i.e., not part of CA).
+We should also highlight here that we coded referent status as nominal if the clause included an explicit nominal referent.
+For example, if constructed action, an index sign, or a classifier preceded or followed its explicitly signed nominal referent, the clause was regarded as having nominal reference (i.e., modified noun).
 The annotation tiers are presented in Table \ref{tab:tab-2}.
@@
 The researcher is usually interested in the remaining coefficients which give information about predictor level comparisons.
-
-In a multinomial distribution, the responses are in a categorical relationship where the levels can be A,B,C, and more.
+Each coefficient is a comparison.
+That is why for a predictor with two levels A and B, the plots or the results only have "A" or "B" as labels.
+
+In a multinomial distribution, the responses are in a categorical relationship where the levels can be A, B, C, and more.
 These levels are not ordered in time, they are not cumulative, and they are not additive.
-If we were to make an analogy, they are like countries on Earth which can have differences and similarities, but they don't have to.
+If we were to make an analogy, they are like countries on Earth which can have differences and similarities, but they do not have to, and they are treated as distinct entities.
 In a multinomial distribution, one of the responses becomes the level of comparison and other levels are compared against it.
@@
 If a considerable amount (\>90%) of the posterior probability distribution is towards a sign (-/+) we interpret it as a categorial effect, if a good amount (\>50%) of the distribution is towards a sign (-/+) we interpret it as a tendency, and we see the remaining distributions as indifference.
-In addition to these, we look at how the distribution skewed in the extremes or how \`wide' it is from the median estimate to inform us mainly about uncertainty and variation.
+In addition to these, we look at how the distribution is skewed in the extremes or how \`wide' it is from the median estimate to inform us mainly about uncertainty and variation.
 
@@
 We collected `r nrow(df_narration)` data points as a result of the experiment.\footnote{For the purposes of this double-blind review, we share all our data and code in an anonymous online repository that can be accessed here: \url{https://anonymous.4open.science/r/reference_tracking_revision-A542/}} Each data point is a referent tracked by the participant.
-Each of the REs is classified in terms of its Discourse Status (Introduction, Maintenance, Re-introduction), its Type (Nominal, Pronominal, Classifer, Null),\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer, and Null).
-However, we allude to the numerical distribution of the subtypes in this and following sections.} and Participant Group (Native, Late).\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17).}
+Each of the REs is classified in terms of its Discourse Status (Introduction, Maintenance, Re-introduction), its Type (Nominal, Pronominal, Classifer, Zero Anaphora),\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer, and Zero Anaphora).
+However, we allude to the numerical distribution of the subtypes in this and following sections.} and Participant Group (DoD, DoH).\footnote{Similarly, we consider DoH signers a single group in our analyses and disregard the differing age of acquisition intervals (0-3, 4-7, 8-12, 13-17).}
 
@@
 For the introduction of a referent, the majority of the REs were nominal constructions (Table \ref{tab:my-table}).
-However, we observed a sporadic use of classifiers and null marking respectively.
-When the discourse status was maintained, signers extensively used null markers to refer to the entities in the immediately previous clause.
-Classifier constructions followed null markers in frequency in maintained contexts.
+However, we observed a sporadic use of classifiers and zero anaphora respectively.
+When the discourse status was maintained, signers extensively used zero anaphora to refer to the entities in the immediately previous clause.
+Classifier constructions followed zero anaphora in frequency in maintained contexts.
 We observed only a few instances of nominal use for referent maintenance.
-In re-introduction contexts, signers used a greater proportion of nominal markings which were followed by null, classifier, and pronominal use.
+In re-introduction contexts, signers used a greater proportion of nominal markings which were followed by zero anaphora, classifier, and pronominal use.
 
@@
                      & Introduced    & Maintained     & Re-introduced \\ \hline
-\textit{Nominal:}    & 92.78\% (989) & 5.63\% (93)    & 63.65\% (907) \\
-\textit{Pronominal:} & 0.09\% (1)    & 0.24\% (4)     & 2.46\% (35)   \\
-\textit{Classifier:} & 5.35\% (57)   & 18.29\% (302)  & 8\% (114)     \\
-\textit{Null:}       & 1.78\% (19)   & 75.83\% (1252) & 25.89\% (369) \\ \hline
+\textit{Nominal:}    & 92.7\% (956) & 5.9\% (93)    & 63.7\% (884) \\
+\textit{Pronominal:} & 0.1\% (1)    & 0.3\% (4)     & 2.5\% (35)   \\
+\textit{Classifier:} & 5.3\% (55)   & 18.3\% (291)  & 7.8\% (108)     \\
+\textit{Null:}       & 1.8\% (19)   & 75.6\% (1201) & 26\% (360) \\ \hline
 \end{tabular}%
@@
                         & Bare Noun  & Modified Noun & FS Noun \\ \hline
-\textit{Introduced:}    & 48.93\% (842) & 53.56\% (139)     & 100\% (8) \\
-\textit{Maintained:}    & 4.82\% (83)  & 3.85\% (10)      & 0\% (0) \\
-\textit{Re-introduced:} & 46.25\% (796) & 42.69\% (111)     & 0\% (0) \\ \hline
+\textit{Introduced:}    & 48.7\% (813) & 53\% (134)     & 75\% (9) \\
+\textit{Maintained:}    & 5\% (83)  & 4\% (10)      & 0\% (0) \\
+\textit{Re-introduced:} & 46.3\% (772) & 43\% (109)     & 25\% (3) \\ \hline
 \end{tabular}
@@
 
-Among the null markers, CA was most frequently employed, and it was followed by plain verbs and agreement verbs (Table \ref{tab:my-table5}).
-Overall, signers used very few null markings for the introduction of a referent and the majority of the existing ones were CA.
+Among the markers of zero anaphora, CA was most frequently employed, and it was followed by plain verbs and agreement verbs (Table \ref{tab:my-table5}).
+Overall, signers used very few markers of zero anaphora for the introduction of a referent and the majority of the existing ones were CA.
 Additionally, signers used CA more for referent re-introduction than maintenance.
@@
                         & WECL          & BPCL        & ExtCL         \\ \hline
-\textit{Introduced:}    & 2.14\% (6)    & 2.4\% (3)   & 71.64\% (48) \\
-\textit{Maintained:}    & 67.26\% (189) & 77.6\% (97) & 23.88\% (16) \\
-\textit{Re-introduced:} & 30.6\% (86)   & 20\% (25)   & 4.48\% (3)   \\ \hline
+\textit{Introduced:}    & 2.2\% (6)    & 2.5\% (3)   & 70.8\% (46) \\
+\textit{Maintained:}    & 67.5\% (181) & 77.7\% (94) & 24.6\% (16) \\
+\textit{Re-introduced:} & 30.2\% (81)   & 19.8\% (24)   & 4.6\% (3)   \\ \hline
 \end{tabular}
@@
 ```
-Across the discourse statuses, there is also a difference between participant groups in the use of null marking for reference tracking.
-Native signers of TİD used it more (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "Native") / nrow(df_narration), digits=2)`) compared to the late signers (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "Late") / nrow(df_narration), digits=2)`).
-Comparably, late signers used nominals more frequently (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "Late") / nrow(df_narration), digits=2)`) than native signers (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "Native") / nrow(df_narration), digits=2)`).
+Across the discourse statuses, there is also a difference between participant groups in the use of zero anaphora for reference tracking.
+DoD signers of TİD used it more frequently (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "DoD") / nrow(df_narration), digits=2)`) compared to the DoH signers (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "DoH") / nrow(df_narration), digits=2)`).
+Comparably, DoH signers used nominals more frequently (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "DoH") / nrow(df_narration), digits=2)`) than DoD signers (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "DoD") / nrow(df_narration), digits=2)`).
 Figure 1 shows the numerical distribution of the responses by RE type and Group.
@@
 \small
-\caption{Distribution of null markers by discourse status}
+\caption{Distribution of zero anaphora by discourse status}
 \label{tab:my-table5}
@@
                         & Constructed Action & Plain Verb    & Agreement Verb \\ \hline
-\textit{Introduced:}    & 1.58\% (17)        & 0.28\% (1)    & 0.5\% (1)      \\
-\textit{Maintained:}    & 71.64\% (773)      & 89.97\% (323) & 77.23\% (156)  \\
-\textit{Re-introduced:} & 26.78\% (289)      & 9.75\% (35)   & 22.28\% (45)   \\ \hline
+\textit{Introduced:}    & 1.6\% (17)        & 0.3\% (1)    & 0.5\% (1)      \\
+\textit{Maintained:}    & 71.5\% (749)      & 89.3\% (302) & 76.9\% (150)  \\
+\textit{Re-introduced:} & 26.8\% (281)      & 10.4\% (35)   & 22.6\% (44)   \\ \hline
 \end{tabular}
@@
 ```
-```{r fig:count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), RE Type (Nominal, Pronominal, Classifier, Null), and Participant Group (Native, Late)"}
-df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "NULL"))
+```{r fig:count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), RE Type (Nominal, Pronominal, Classifier, Zero Anaphora), and Participant Group (DoD, DoH)"}
+df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "ZERO"))
 
@@
              position = position_dodge(width = 0.85), size = 2) +
-  facet_grid(~Discourse) + ylab("Count") + xlab("REATs") +
+  facet_grid(~Discourse) + ylab("Count") + xlab("REs") + 
+  labs(fill = "Group", color = "Group") +
   scale_color_grey() +
@@
 For more inference on the data and having pairwise comparisons, we fit a regression model to RE Type using discourse status and acquisition group as predictors.
-We used the brms package [@burkner2018] in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late).
+We used the brms package [@burkner2018] in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (DoD, DoH).
 As the overall number of produced pronouns was very little (40 data points), we excluded pronominal use from the following analyses.
 For the current analysis, we did not define priors ourselves and used the preset uninformed priors.
-Figure 2[^2] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals.
+Figure 2[^1] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals.
 There are 3 responses we are considering in this model.
-They are Classifier, Nominative, and Null.
-Classifier response is the level that other responses Nominative and Null are compared against.
-
-[^2]: We ran another model where we used the Nativeness predictor with 3 levels 0-3:Native, 4-7:Late1, and 7-above:Late2.
+They are Classifier, Nominal, and Zero Anaphora.
+Classifier response is the level that other responses Nominal and Zero Anaphora are compared against.
+
+[^1]: We ran another model where we used the self-reported acquisition time as a predictor with 3 levels 0-3:Early, 4-7:Late1, and 7-above:Late2.
 However, our model did not converge properly and the credible intervals were very wide spread, yielding the results uninformative.
@@
 
-```{r model-gram, echo=FALSE, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=2.5, fig.pos="hbt!", fig.width=4, message=FALSE, warning=FALSE, cache=FALSE, results='asis', warnings=FALSE}
+```{r model-gram, echo=FALSE, fig.cap="Regression model plot for the RE Type (Nominal, Zero, Classifier) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (DoD, DoH)", fig.height=2.5, fig.pos="hbt!", fig.width=4, message=FALSE, warning=FALSE, cache=FALSE, results='asis', warnings=FALSE}
 p3 <- df_model_1 %>% 
@@
 It means that participants produced more nominals than classifiers in general regardless of participant group and discourse status.
-The second coefficient shows that null markings were also used more than classifiers by the participants in general, but not as much as nominals.
-
-In the comparison of nominals and classifiers, being a native acquirer of TİD did not have an effect.
+The second coefficient shows that zero anaphora was also used more than classifiers by the participants in general, but not as much as nominals.
+
+In the comparison of nominals and classifiers, acquiring TİD from deaf signing caregivers did not have an effect.
 Nominals were used more when the discourse status was introduction compared to re-introduction.
@@
 In contrast to the introduction status, nominals were used less in maintenance contexts compared to the re-introduction contexts within the greater comparison to classifier referents.
-There seems to be an interaction between the introduction status of discourse and being a native acquirer of TİD.
-When the acquirer is a native speaker and the discourse function is introduction, the chances of observing a nominal referent gets even higher.
-Also, there seems to be an interaction between maintaining a referent and being a native acquirer of TİD.
-When it was a native acquirer of TİD and the discourse context was maintenance, the chances of observing a nominal referent gets even lower.
-
-Compared to classifiers, being a native acquirer of TİD increased the odds of using null marking in general.
-Null constructions were used less in introduction contexts than in re-introduction.
+There seems to be an interaction between the introduction status of discourse and being a DoD signer of TİD.
+When the acquirer is a DoD signer and the discourse function is introduction, the chances of observing a nominal referent gets higher.\footnote{Remember that this is an interaction term and the grand comparison is nominal vs classifier use.} Also, there seems to be an interaction between maintaining a referent and being a DoD signer of TİD.
+When it was a DoD signer of TİD and the discourse context was maintenance, the chances of observing a nominal referent gets lower.\footnote{This coefficient too is an interaction term and needs to be considered withing the comparison of nominal versus classifier use.}
+
+Compared to classifiers, being a DoD signer of TİD increased the odds of using zero anaphora in general.
+Zero anaphora constructions were used less in introduction contexts than in re-introduction.
 This contrasts with the nominals (4th coefficient) as they were used more in referent introduction versus re-introduction.
-In contrast to introduction contexts, null markers were used more in referent maintenance compared to re-introduction, yet again within the greater comparison to the classifiers.
-There seems to be an interaction between introducing a referent and being a native acquirer of TİD.
-In that, the odds of using null marking increases if the participant is a native acquirer of TİD and the referent is used in an introduction context.
+In contrast to introduction contexts, zero anaphora was used more in referent maintenance compared to re-introduction, yet again within the greater comparison to the classifiers.
+There seems to be an interaction between introducing a referent and being a DoD signer of TİD.
+In that, the odds of using zero anaphora increases if the participant is a DoD signer of TİD and the referent is used in an introduction context.
 There is another interaction with regards to referent maintenance, contrasting the first interaction.
-In that, the odds of using null markers decrease if the participant is a native acquirer of TİD and the referent is maintained.
-This does not mean that late signers used more null markers in maintenance contexts.
-It means that the use of null markers did not linearly increase as the sum of using a null marker as a native plus using null markers for maintaining a referent versus re-introducing it.
+In that, the odds of using zero anaphora decrease if the participant is a DoD signer of TİD and the referent is maintained.
+This does not mean that DoH signers used more zero anaphora in maintenance contexts.
+It means that the use of zero anaphora did not linearly increase as the sum of using a zero anaphora marker as a DoD signer plus using zero anaphora for maintaining a referent versus re-introducing it.
 
@@
 
-In our study, we aimed to (i) describe how signers track referents in signed discourse in TİD and (ii) examine whether language deprivation results in more over-redundant referent selection.
-To this end, we investigated the distribution of referring expressions (REs) that the native and late signers select when employing referents in different discourse statuses (introduced, maintained, and re-introduced).
-
-In our first set of hypotheses regarding referent tracking strategies, we had the general expectation that signers would use more explicit forms (mainly nominals) for referent introduction and less implicit forms such as null marking, pronominal use, and classifiers for maintenance.
+In our study, we aimed to (i) describe how signers track referents in signed discourse in TİD and (ii) examine whether late acquisition of TİD results in more over-redundant referent selection.
+To this end, we investigated the distribution of referring expressions (REs) that the DoD and DoH signers select when employing referents in different discourse statuses (introduced, maintained, and re-introduced).
+
+In our first set of hypotheses regarding referent tracking strategies, we had the general expectation that signers would use more explicit forms (mainly nominals) for referent introduction and less implicit forms such as zero anaphora, pronominals, and classifiers for maintenance.
 In re-introduced contexts, we predicted that signers would use both explicit and implicit markers depending on the accessibility of the tracked referent.
 
-For age of acquisition effects, based on what is proposed for child and adult deaf signers with language deprivation in terms of their morphosyntactic [@boudreault2006; @cormier2012; @kayabasi2020] and narrative development [@becker2009; @cormier2013], we expected that late signers would be more over-redundant in their referent tracking, akin to second language learners [@frederiksen2019; @bel2015].
-That is, we had the expectation that late signers would use more explicit markers like nouns and fewer implicit markers like null marking even for highly accessible referents (when the discourse status is maintained or even re-introduced in same cases) than native signers.
+For age of acquisition effects, based on what is proposed for child and adult deaf signers with delayed language exposure in terms of their morphosyntactic [@boudreault2006; @cormier2012; @kayaba2022] and narrative development [@becker2009; @cormier2013], we expected that DoH signers would be more over-redundant in their referent tracking, akin to second language learners [@frederiksen2019; @bel2015].
+That is, we had the expectation that DoH signers would use more explicit markers like nouns and fewer implicit markers like zero anaphora even for highly accessible referents (when the discourse status is maintained or even re-introduced in same cases) than DoD signers.
 
@@
 
-The results of the present study partly support our hypotheses.
-In line with our expectations, regardless of their acquisition groups (native or late) (i) signers mainly used nominals to introduce a referent into the discourse, (ii) used null marking and classifiers but not pronouns for maintained contexts, and (iii) employed both nominals and less explicit markers (classifiers, null markers and pronouns) to re-introduce a referent.
+The results of the present study only partly support our hypotheses.
+In line with our expectations, regardless of the participant groups (DoD or DoH) (i) signers mainly used nominals to introduce a referent into the discourse, (ii) used zero anaphora and classifiers but not pronouns for maintained contexts, and (iii) employed both nominals and less explicit markers (classifiers, zero anaphora and pronouns) to re-introduce a referent.
 
@@
 In contrast, the stimuli in the present study consisted of 8-episode-long video clips on average with a mean number of 2 animate characters and 3 inanimate objects.
-Although the number of animate characters and inanimate characters in @czubek2017's stimuli matched that of the present study, their narrative structure (a 6-episode-long picture story) was still shorter compared to our narrative stimuli.
+Although the number of animate characters and inanimate characters in Czubek's [-@czubek2017] stimuli matched that of the present study, their narrative structure (a 6-episode-long picture story) was still shorter compared to our narrative stimuli.
 We believe fewer possible referents and/or a simpler narrative structure in these studies may have contributed to the lower number of re-introduced contexts compared to the present study.
 Due to the story length and composition in this study, the participants had to track more recurring animate or inanimate referents for a longer period of time.
-As a result, our signers engaged in referent re-introduction more than what was reported in previous studies [@czubek2017; @frederiksen2016].
-
-As for the distribution of REs across different discourses statuses, we observed that signers overwhelmingly preferred to use nominals over other REs for the first mention of a referent and used null markers such as CA, agreement and plain verb for referents with high accessibility (i.e., when the discourse status was maintained).
+As a result, our participants might have engaged in referent re-introduction more than what was reported in previous studies [@czubek2017; @frederiksen2016].
+
+As for the distribution of REs across different discourses statuses, we observed that signers overwhelmingly preferred to use nominals over other REs for the first mention of a referent and used zero anaphora such as CA, agreement and plain verb for referents with high accessibility (i.e., when the discourse status was maintained).
 This finding parallels other accounts of referential accessibility suggested for other sign languages (see @frederiksen2016; @czubek2017; and @swabey2002; for ASL, @morgan2005; for BSL @bel2015; and @barbera2009 for LSC).
 Our results also indicate that modified nouns (e.g., nouns that are preceded or followed by a classifier or IX) appeared more in introduced contexts compared to re-introduced contexts.
-This brings along the assumption that modified nouns are used more to track referents with lower accessibility, probably due to their more explicit structure.
+This indicates that modified nouns are used more to track referents with lower accessibility, probably due to their more explicit structure.
 
@@
 This finding is comparable with what has been reported for ASL [@swabey2002; @frederiksen2016; @czubek2017] but not with LSC [@bel2015], BSL [@morgan2005] and DGS [@perniss2015], in which the signers extensively made overt pronominal references (up to 30%) to maintain entities.
-Comparable to our findings, it has been reported that pronouns do not frequently occur in the local context and instead follows a topic shift and thus rank higher than classifiers and null markers [see @barbera2009 for LSC; and @nuhbalaoglu2018 for TİD].
+Comparable to our findings, it has been reported that pronouns do not frequently occur in the local context and instead follows a topic shift and thus rank higher than classifiers and zero anaphora [see @barbera2009 for LSC; and @nuhbalaoglu2018 for TİD].
 Considering this, the claim that the IX sign functions like demonstratives that are also used to track referents in re-introduced contexts has not been borne out [@koulidobrova2016].
@@
 Moreover, another production study on referent tracking in ASL reported object preference for pronominals [@frederiksen2017].
-Nonetheless, we only coded subjects for maintained and re-introduced referents.
-We argue that all of these factors (genre-dependency as well as object preference) might contribute to the explanation of why overt pronouns in this data are dispreferred in favor of other markers for highly accessible referents such as null marking and classifiers.
-
-A deeper look into the allocation of classifier constructions revealed that WECLs were the most common type of classifiers for the maintenance of a referent followed by BPCLs and ExtCLs.
+However, we only coded subjects for maintained and re-introduced referents.
+We argue that all of these factors (genre-dependency as well as object preference) might contribute to the explanation of why overt pronouns in our data are dispreferred in favor of other markers for highly accessible referents such as zero anaphora and classifiers.
+
+A deeper look into the allocation of classifier handshapes revealed that WECLs were the most common type of classifiers for the maintenance of a referent followed by BPCLs and ExtCLs.
 We observed a similar distibution for re-introduced contexts.
@@
 That is to say, whereas WECLs were the preferred classifier types to signal highly accessible referents, ExtCL constructions were mainly used for referents which were not easily accessible.
-We suggest that the less iconic and transparent nature of WECLs allowed signers to continue a previous referent without being redundant.
+We suggest that the less iconic and transparent nature of WECLs allowed signers to continue a previous referent without having to use an explict form.
 As pointed out by @czubek2017, signers prefer to rely on ExtCLs for the first mention of an entity because these constructions are informative and explicit by nature, and they provide visual and descriptive cues about an entity and its location in space.
-We only observed `r sum(df_narration$GTR1_plot == "SASS" & df_narration$Discourse == "Re-introduction")` instances of ExtCLs for re-introduced contexts, replicating @frederiksen2019's findings for ASL.
-
-As for the distribution of the null markers, approximately 90% of the plain verbs were used to maintain a previous referent, making these constructions least implicit REs used for TİD[^3].
+We only observed `r sum(df_narration$GTR1_plot == "SASS" & df_narration$Discourse == "Re-introduction")` instances of ExtCLs for re-introduced contexts, replicating Frederiksen and Mayberry's [-@frederiksen2019] findings for ASL.
+
+As for the distribution of the zero anaphora, approximately 90% of the plain verbs were used to maintain a previous referent and this makes these constructions least implicit REs used for TİD[^2].
 Agreement verbs and CA were the following implicit markers of reference, `r sum(df_narration$GTR1 == "Person agreement verb" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1 == "Person agreement verb")` and `r sum(df_narration$GTR1_plot == "CA" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1_plot == "CA")` of which were used for maintained contexts, respectively.
-This contrasts with some of the research [@czubek2017] in which agreement verbs ranked lowest on scale, indicating that these constructions were more implicit compared to plain verbs.
-However, there seems to be a common ground for the position of CA markers which have been claimed to rank higher than other null markers as in the present study [@frederiksen2016; @czubek2017; @barbera2009].
+This contrasts with some of the research [@czubek2017] in which agreement verbs ranked lowest on the scale, indicating that these constructions were more implicit compared to plain verbs.
+However, there seems to be a common ground for the position of CA markers which have been claimed to rank higher than other markers of zero anaphora as in the present study [@frederiksen2016; @czubek2017; @barbera2009].
 In Figure \ref{fig:scale}, we report the distribution of REs in TİD as a function of referent accessibility and compare our accessibility hierarchy with that of [@frederiksen2016] for ASL.
@@
 
-[^3]: As in other sign languages, in TİD, the lexical meaning of signs has an effect on the morphological grouping.
+[^2]: As in other sign languages, in TİD, the lexical meaning of signs has an effect on the morphological grouping.
 For instance, usually (di)transitive agreement verbs denote some sort of physical or metaphorical transfer of an object or abstract notion [@meir1999].
@@
 ```
-Both hierarchies illustrated in Figure \ref{fig:scale} consider nominals and ExtCLs as markers of lowly accessibly referents.
-Likewise, in both hierarchies, as markers of highly accessible referents, different types of null markers are further distinguished.
+Both hierarchies illustrated in Figure \ref{fig:scale} treat nominals and ExtCLs as markers of lowly accessibly referents.
+Likewise, in both hierarchies, as markers of highly accessible referents, different types of zero anaphora markers are further distinguished.
 Plain verbs rank the lowest on the scales, preceded by agreement verbs and CA.
 On a similar ground, both scales lack pronouns, which occurred very infrequently in the data.
-Although @frederiksen2016 did not include other classifier types in the hierarchy, we proposed that WECLs and BPCLs must ranker higher than null markers but lower than ExtCLs that were mainly used by interlocutors to introduce a referent into the discourse context.
-
-## Age of acquisition effects on referent tracking
-
-The results in this study only somewhat support our second set of hypotheses that we set for the effects of age of acquisition on participants' RE selection.
-That is, while being a native acquirer of TİD increased the odds of using null marking in reference tracking across different discourse statuses, we failed to find a difference for overall nominal production.
-Taking a closer look at the interaction terms, we observed that native signers, compared to late signers, were less likely to use nominal constructions when the discourse status was maintained.
-Native acquisition further increased the odds of null markers when the discourse status was introduced.
+Although @frederiksen2016 did not include other classifier types in the hierarchy, we propose that WECLs and BPCLs must rank higher than zero anaphora but lower than ExtCLs that were mainly used by interlocutors to introduce a referent into the discourse context.
+
+## Effect of native acquisition on referent tracking
+
+The results in this study only somewhat support our second set of hypotheses that we set for the effects of having signing deaf parents on participants' RE selection.
+That is, while being a DoD signer of TİD increased the odds of using zero anaphora in reference tracking across different discourse statuses, we failed to find a difference for overall nominal production.
+Taking a closer look at the interaction terms, we observed that DoD signers, compared to DoH signers, were less likely to use nominal constructions when the discourse status was maintained.
+Native acquisition from deaf caregivers further increased the odds of zero anaphora when the discourse status was introduced.
 However, we found the opposite alignment for the interaction of maintenance and native acquisition.
-That is, the coefficient for this interaction term slightly decreased the combined main effects of having a maintained context and being a native signer.
-This means that these main effects were greater than the interaction or added effects, pointing out to the implication that the type of discourse status (introduced, maintained, re-introduced) accounts for the distribution of REs in TİD narratives better than age of acquisition (native, late) which, in comparison, had a less salient effect.
-
-While late signers were only partially over-redundant in maintained contexts, they were more so in introduced contexts.
-In line with our study, a tendency for limited over-redundancy has been suggested for late signing children in DGS [@becker2009] and late signing adults in BSL specifically for the use of CA[@cormier2013].
+That is, the coefficient for this interaction term slightly decreased the combined main effects of having a maintained context and being a DoD signer.
+This means that these main effects were greater than the interaction or added effects, pointing out to the implication that the type of discourse status (introduced, maintained, re-introduced) accounts for the distribution of REs in TİD narratives better than native acquisition from deaf caregivers (DoD, DoH) which, in comparison, had a less salient effect.
+
+While DoH signers were only partially over-redundant in maintained contexts, they were more so in introduced contexts.
+In line with our study, a tendency for limited over-redundancy has been suggested for late signing children in DGS [@becker2009] and late signing adults in BSL specifically for the use of CA [@cormier2013].
 As for TİD, to our knowledge, there is only one study [@gur2018] that directly aimed at examining age of acquisition effects on adult signers' narrative skills.
@@
 However, this particular investigation did not present any data on the distribution of REs between two acquisition groups.
-In the present study, the finding that native signers preferred to use null arguments even for the first mention of a referent (`r sum(df_narration$ref_type == "Null" & df_narration$Discourse == "Introduction")` in total) is surprising given some of the previous literature in the topicality and saliency framework [@cormier2013; @swabey2002].
-These research reported very little or no use of null marking by native signers for introduced contexts.
-
-Given the paucity of research directed at the late signing group, we now turn to the referent tracking strategies observed for second language learners of a sign language.
+In the present study, the finding that DoD signers preferred to use null arguments even for the first mention of a referent (`r sum(df_narration$ref_type == "Zero" & df_narration$Discourse == "Introduction")` in total) is surprising given some of the previous literature in the topicality and saliency framework [@cormier2013; @swabey2002].
+These research reported very little or no use of zero anaphora by DoD signers for introduced contexts.
+
+Given the paucity of research directed at the DoH or late signing group, we now turn to the referent tracking strategies observed for second language learners of a sign language.
 Our study replicated some of the findings of the line of research that investigated second language learners of a sign language [@bel2015; @frederiksen2019].
-The authors reported that more instances of null marking in the context of referent introduction was attested for native signers than second language learners.
-Although using null marking to introduce a referent into the discourse would normally result in referential ambiguity, it is important to note that our stimuli consisted of 10 interconnected narratives with 2 or more recurring characters.
-It may be possible that native signers, having already introduced the referent with an explicit nominal marker in a different discourse context, found it felicitous to refer to the same entity with more implicit markers in the following narratives even for introduced contexts.
-Then, our results imply a difference in pragmatic sensitivity to referential accessibility between native and late signers, which is more robustly reflected in introduced contexts.
+The authors reported that more instances of zero anaphora in the context of referent introduction was attested for DoD signers than second language learners.
+Although using zero anaphora to introduce a referent into the discourse would normally result in referential ambiguity, it is important to note that our stimuli consisted of 10 interconnected narratives with 2 or more recurring characters.
+It may be possible that DoD signers, having already introduced the referent with an explicit nominal marker in a different discourse context, found it felicitous to refer to the same entity with more implicit markers in the following narratives even for introduced contexts.
+Then, our results imply a difference in pragmatic sensitivity to referential accessibility between DoD and DoH signers, which is more robustly reflected in introduced contexts.
 To test this implication, one could use the same number and length of stimuli from different cartoons with different characters and see if the tendency to introduce a referent without an overt nominal for the first time is dependent on the task or not.
 
-Since we only observed a small difference between the two acquisition groups for the use of certain REs, the aforementioned findings allude to the implication that the effect on language deprivation on pragmatic factors should also be considered in different contexts.
+In other contexts, we found only very small differences between the two participant groups.
+This means that adult signers with late exposure to sign language in our study were still able to achieve native-like competency in terms of reference tracking in simple narratives.
+Nevertheless, DoD and DoH signers still exhibit some differences especially in terms of pragmatic sensitivity to economy of form despite using TİD as their preferred communication language for similar numbers of years.
+Since we only observed a small difference between the two acquisition groups for the use of certain REs, the aforementioned findings allude to the implication that the effect on native acquisition on pragmatic factors should also be considered in different contexts.
 For one, both groups of signers can be tested with different genres of narratives and with a more complex story design.
 In order to capture possible variation in the accessibility levels of re-introduced referents, a more fine-tuned protocol [as in @toole1996] can be used for measuring referent distance and the number of other possible competitors.
-Taken altogether, our results accentuates the necessity of engaging in signed narrative production early in life as well as the importance of early intervention for late signing deaf children.
+Furthermore, our focus in this study was to look at referent tracking strategies between DoD and DoH signers.
+However, for future research, it might also be useful to look at pre and post-lingual acquisiton of TİD.
+We only had 3 post-lingual signers of TİD.\footnote{When we looked at the random effects by participants, they did not differ from their acquisition group (DoD or DoH).}
+
+\clearpage
 
@@
 The present study aimed to investigate the reference tracking strategies in TİD narratives and whether native and late deaf TİD signers differ from each other in these strategies.
-In line with our expectations, we found that the distribution of REs (nominal, pronominal, classifier, and null marking) followed from previous accounts of saliency and referential accessibility.
+In line with our expectations, we found that the distribution of REs (nominal, pronominal, classifier, and zero anaphora) followed from previous accounts of saliency and referential accessibility.
 For introduced contexts, signers produced more nominals and ExtCLs, and to maintain a previous referent, they mainly used null arguments (in clauses with CA, agreement and plain verbs) and certain classifiers (e.g., BPCL and WECL) to track highly accessible referents.
-We observed only very few instances of pronouns (mostly produced by late signers), unlike our expectations.
+We observed only very few instances of pronouns (mostly produced by DoH signers), unlike our expectations.
 The ones we observed were mostly reserved for re-introduced contexts along with nominals.
 Based on our findings across the two groups, we proposed a hierarchy of referent accessibility for TİD.
-As for the age of acquisition effects, we observed limited over-redundancy for signers with delayed language acquisition (i.e., late signers) who employed fewer implicit forms of reference tracking compared to native signers.
-Native signers were less likely to use a nominal for maintained contexts and more likely to use null marking for introduced contexts than late signers.
-Following @swabey2002's intituition, we conclude that native signers must adhere to pragmatic norms and leverage accessibility better than late signers only to a certain extent (given our finding of limited overexplicitness for late signers) since they use less form when the referent is more active in the mind.
+As for the effects of native acquisition, we observed limited over-redundancy for signers with delayed language acquisition (i.e., DoH signers) who employed fewer implicit forms of reference tracking compared to DoD signers.
+DoD signers were less likely to use a nominal for maintained contexts and more likely to use zero anaphora for introduced contexts than DoH signers.
+
+We conclude that DoH signers have overall native-like competency in simple narratives, however, following @swabey2002's intuition, our findings also suggest that DoD signers adhere to pragmatic norms and leverage accessibility better than DoH signers to a certain extent (given our finding of limited overexplicitness for DoH signers) since they use less form when the referent is more active in the mind.
+To avoid this possible discrepancy between the groups, our results accentuates the necessity of engaging in signed narrative production early in life as well as the importance of early intervention for DoH or late signing deaf children.
+This is especially a crucial message for policymakers, educators, and health professionals who work with deaf and hard-of-hearing children.
+Providing early and effective access to sign language from very early on can help bridge the gap and ensure equal opportunities for language development and success in life for DoH signers.
 
@@
           caption ="Participant demographic information based on self-report",
-          note = "AOA = age of acquisition, LOE = level of education, PS = primary school, MS = middle school, HS = high school, TR = (Spoken) Turkish.",
+          note = "AOA = age of acquisition, LOE = level of education, PS = primary school, MS = middle school, HS = high school, TR = (Spoken) Turkish.
+*We excluded Participant 13 from our data analysis because the signer indicated to have deaf parents, thus grouped as DoD, but also reported only home sign as the home language and not TİD.
+Since our classification of DoD included the criterion of having deaf and TİD-signing parents, we decided to remove this participant's data.
+Also, Participant 10 reported to use TİD at home despite having hearing caregivers which may go against our classification criterion for being DoH that is to have non-signing hearing caregivers.
+However, Participant 10 informed us that the parents used very rudimentary sign language and that they often did not understand what he was signing.",
           escape = FALSE,

### 2023-01-31 revised_manuscript.tex
updated

@@
             pdfauthor={},
-            pdftitle={Reference Tracking Strategies of Native and Late Signers in Turkish Sign Language},
+            pdftitle={Reference Tracking Strategies of Deaf Adult Signers in Turkish Sign Language},
             colorlinks=false,
@@
 \usepackage{multicol}
-\setlength{\columnsep}{-3cm}
+\setlength{\columnsep}{-2cm}
 \usepackage{setspace}
@@
 
-  \title{Reference Tracking Strategies of Native and Late Signers in
-Turkish Sign Language}
+  \title{Reference Tracking Strategies of Deaf Adult Signers in Turkish
+Sign Language}
 
   \begin{abstract}
-  We investigated the reference tracking strategies among native and
-  late adult deaf signers of Turkish Sign Language. Consistent with the
-  theories of saliency and referential accessibility, regardless of
-  their acquisiton groups, signers mainly used nominals and extension
-  classifiers for introducing referents. To maintain a referent in the
-  previous clause, signers chose null marking (e.g., constructed action,
-  agreement and plain verbs). For topic shifts or re-introduced
-  contexts, nominals and pronouns were chiefly favored although we
-  observed very little pronominal use. As for age of acquisition
-  effects, we only report limited over-redundancy for late signers who
-  used null marking less compared to native signers especially for
-  introduced and maintained contexts. Given our limited support for
-  redundancy by late signers, we conclude that native signers might
-  utilize the spatial affordances of the visual-spatial modality better
-  than late signers only to a certain extent.
+  We investigated the reference tracking strategies among deaf adults
+  with signing deaf parents (DoD) and adult deaf signers with
+  non-signing hearing parents or caregivers (DoH) who had late exposure
+  to Turkish Sign Language (TİD). Consistent with the theories of
+  saliency and referential accessibility, regardless of their acquisiton
+  groups or parental hearing status, signers mainly used nominals and
+  extension classifiers for introducing referents. To maintain a
+  referent from the previous clause, the signers used zero anaphora
+  (e.g., constructed action, agreement and plain verbs). For topic
+  shifts or re-introduced contexts, nominals and pronouns were chiefly
+  favored although we observed very little pronominal use. As for the
+  effect of native acquisition from deaf caregivers, we only report
+  limited over-redundancy for DoH signers who used zero anaphora less
+  compared to DoD especially for introduced and maintained contexts. We
+  conclude that DoH signers are still able to achieve native-like
+  competency in terms of reference tracking in simple narratives but DoD
+  signers utilize the spatial affordances of the visual-spatial modality
+  better than DoH signers only to a certain extent.
   \end{abstract}
@@
 
-Narratives are extensive discourses that construct an important part of
-human communication. In order to create a systematic and coherent
-organization of experiences, narrators must choose appropriate
-linguistic forms to refer to entities (Ariel, 1990; Givon, 1983; Gundel
-et al., 1993): Certain forms such as nominal expressions are commonly
-used for the first mention of a referent in a discourse, and other forms
-such as pronouns and null marking may track previously mentioned
-entities which could either be maintained across multiple clauses or
-re-introduced back into the discourse. The selection of these referring
-expressions (henceforth REs) is contingent on cognitive abilities, such
-that individuals must rely on the activation of the working memory to
-keep in mind previously referred expressions and long strings of
-information (Bamberg, 1997; Morgan, 2005). Pragmatic abilities too play
-a pivotal role in the tracking of referents. The form with the least
-amount of redundant information must be selected based on the
-accessibility levels of the referent (i.e., how active a referent is in
-the addressee's mind given a point in discourse) (Ahn, 2019; Ariel,
-1990). Because of these cognitive and pragmatic necessities, it is not
-surprising to observe that producing a coherent narrative is one of the
-latest milestones in language acquisition among children (Hickmann et
-al., 1996). As can be expected, there is a plethora of research
-examining how narrative competence is achieved in both monolingual and
-bilingual acquisition contexts among children and adults. Although
-narrative competence and referent selection have been studied
-extensively for typically developing speakers (Aksu-Koç and
-Nicolopoulou, 2015; Hickmann et al., 1996; Williams, 1988), deaf
-individuals who acquired a sign language in atypical learning contexts
-(i.e., deaf signers with linguistic deprivation) have remained mostly
-understudied with an exception of handful of studies (Becker, 2009;
-Cormier et al., 2013; Gür, 2018). Nevertheless, it is paramount that we
-investigate and describe the effects of language deprivation among deaf
-adults on narrative competence more in depth. This will help us
-understand whether the loss of systematic and frequent language input in
-the early years of life affect how deaf signers produce a narrative in
-their adulthood. Given this, the present study will focus on deaf adult
-signers with language deprivation and examine their reference tracking
-strategies in signed narratives.
-
-Sign Languages are natural languages of the Deaf communities all around
-the world. They are independent languages, which do not derive from
-majority spoken language(s) but emerge naturally when deaf people form a
-community. Turkish Sign Language (Türk İşaret Dili -- TİD) is and has
-been the officially recognized language of the Deaf community in Turkey
-since 2005 (see Nuhbalaoglu, 2018 for a brief review of the historical
-development of the language). Deaf or hearing children acquiring a SL
-within a family where there is one or more deaf caregivers display the
-same developmental milestones as hearing children acquiring a spoken
-language in a hearing family (Chamberlain et al., 1999; Meier and
-Newport, 1990). However, reports indicate that over 90\% of all deaf
-individuals are born into hearing families who do not know how to sign
-(i.e., those with language deprivation) (Woll, 2013). This means that
-the auditory modality used in the household is not accessible to many
-deaf children. That is why the age of acquisition of a first functional
-language varies greatly among deaf children, with many of whom being
-deprived of their first languages typically until after age 7 when they
-start formal education in a deaf school. Despite this, many of the early
-intervention programs for late signing children prioritize spoken
-language development and discourage the use of a sign language
-(Kemaloğlu and Kemaloğlu, 2012). Ministry of National Education (2020)
-identifies 72 deaf schools in Turkey for the K-12 level. Even when deaf
-children with hearing parents start their formal education in one of
-these schools, they do not receive proper instruction in TİD since the
-teachers themselves are mainly hearing and only have a very rudimentary
-knowledge of a sign language (Ilkbasaran, 2015). First language
-deprivation or delay has been found to have negative effects on certain
-cognitive and linguistic measures among late signing children. To name a
-few, delayed exposure to a first functional language is reported to
-negatively affect deaf children and adults' morphosyntactic production
-and processing (Boudreault and Mayberry, 2006; Cheng and Mayberry, 2021;
-Cormier et al., 2012; Karadöller et al., 2017; Kayabaşı et al., 2020;
-Newport, 1990) and narrative abilities (e.g., Becker, 2009). In our
-overall research, we are concerned with the effects of this late
-exposure in the ultimate adult forms of TİD users.
+Narratives are extensive discourses that form an important part of human
+communication. In order to create a systematic and coherent organization
+of experiences, narrators must choose appropriate linguistic forms to
+refer to entities (Ariel, 1990; Givon, 1983; Gundel et al., 1993).
+Certain forms such as nominal expressions are commonly used for the
+first mention of a referent in a discourse, and other forms such as
+pronouns and zero anaphora may track previously mentioned entities which
+could either be maintained across multiple clauses or re-introduced back
+into the discourse. The selection of these referring expressions
+(henceforth REs) is contingent on cognitive abilities, such that
+individuals must rely on the activation of the working memory to keep in
+mind previously referred expressions and long strings of information
+(Bamberg, 1997; Morgan, 2005). Pragmatic abilities too play a pivotal
+role in the tracking of referents. The form with the least amount of
+redundant information must be selected based on the accessibility levels
+of the referent (i.e., how active a referent is in the addressee's mind
+given a point in discourse) (Ahn, 2019; Ariel, 1990). Because of these
+cognitive and pragmatic necessities, it is not surprising to observe
+that producing a coherent narrative is one of the latest milestones in
+language acquisition among children (Hickmann et al., 1996). As can be
+expected, there is a plethora of research examining how narrative
+competence is achieved in both monolingual and bilingual acquisition
+contexts among children and adults. Although narrative competence and
+referent selection have been studied extensively for typically
+developing speakers (Aksu-Koç and Nicolopoulou, 2015; Hickmann et al.,
+1996; Williams, 1988), deaf individuals who acquired a sign language in
+atypical learning contexts (i.e., deaf signers with delayed exposure to
+a sign language) have remained mostly understudied with an exception of
+a handful of studies (Becker, 2009; Cormier et al., 2013; Gür, 2018).
+Nevertheless, it is paramount that we investigate and describe the
+effects of late exposure to a sign language among deaf adults growing up
+in a signing vs.~non-signing environment on narrative competence more in
+depth. This will help us understand whether the lack of systematic and
+frequent language exposure in the early years of life affect how deaf
+signers produce a narrative in their adulthood. Given this, the present
+study will focus on two groups of signers and examine their reference
+tracking strategies in signed narratives: (i) deaf of deaf (DoD) adult
+signers (i.e., signers who have deaf caregivers and natively acquired a
+sign language from birth) and (ii) deaf of hearing (DoH) adult signers
+(i.e., signers who have hearing non-signing caregivers and state to have
+started acquiring a sign language later in life).
+
+Sign languages are the natural languages of the Deaf communities all
+around the world. They are independent languages, which do not derive
+from majority spoken language(s) but emerge naturally when deaf people
+form a community. Turkish Sign Language (Türk İşaret Dili -- TİD) is and
+has been the officially recognized language of the Deaf community in
+Turkey since 2005 (see Nuhbalaoglu, 2018 for a brief review of the
+historical development of the language). Deaf or hearing children
+acquiring a SL within a family where there is one or more deaf
+caregivers display the same developmental milestones as hearing children
+acquiring a spoken language in a hearing family (Chamberlain et al.,
+1999; Meier and Newport, 1990). However, reports indicate that over 90\%
+of all deaf individuals are born into hearing families (DoH) who do not
+know how to sign (Woll, 2013). This means that the auditory modality
+used in the household is not accessible to many deaf children. That is
+why the age of acquisition of a first functional language varies greatly
+among deaf children, with many of whom being deprived of their first
+languages typically until after age 7 when they start formal education
+in a deaf school, especially in the context of Turkey (Sari, 2005).
+Despite this, many of the early intervention programs for late signing
+children prioritize spoken language development and discourage the use
+of a sign language (Kemaloğlu and Kemaloğlu, 2012). The Ministry of
+National Education (2020) identifies 72 deaf schools in Turkey for the
+K-12 level. Even when deaf children with hearing parents start their
+formal education in one of these schools, they do not receive proper
+instruction in TİD since the teachers themselves are mainly hearing and
+only have a very rudimentary knowledge of a sign language (Ilkbasaran,
+2015; Sari, 2005). In some cases, it has been argued that late exposure
+to a sign language among certain DoH signers has negative effects on
+some cognitive and linguistic measures. To name a few, delayed exposure
+to a first functional language is reported to negatively affect deaf
+children and adults' morphosyntactic production and processing
+(Boudreault and Mayberry, 2006; Cheng and Mayberry, 2021; Cormier et
+al., 2012; Karadöller et al., 2017; Kayabaşı and Gökgöz, 2022; Newport,
+1990) and narrative abilities (e.g., Becker, 2009). In our overall
+research, we are concerned with the effects of this late exposure in the
+ultimate adult forms of TİD users.
 
@@
 According to this model, given or familiar information is expressed in
-more attenuated forms such as pronominal use and null marking since the
-mental status of the expressed referent is already at the center of
-attention, as indicated by the present contextual information (Chafe,
-1976). The Givenness Hierarchy is akin to other previously proposed
-models of reference control within the topicality and saliency paradigm
-such as the Accessibility Hierarchy (Ariel, 1990), Topic Continuity
-(Givon, 1983), and Familiarity Scale (Prince, 1981). Gundel et al.
-(1993) argued that the distribution of forms is not random but abides by
-the Gricean Maxim of Quantity (Grice, 1975). That is to say, individuals
-must make the most informative and succinct choice possible to retain
-economy of form. When multiple expressions are available for the same
-referent as in an utterance like (\ref{ex:1}), phonetically simplest
-form is selected with the least redundant information based on an
-economy principle. In the utterance, REs for ``Eddie'' are given in
-bold. All three REs (nominal, pronominal, and null marking) refer to the
-same entity; however, the quantity of marking decreases as the
-referent's accessibility increases. On this ground, for highly
-accessible referents (i.e., those that are highly active in the mind),
-linguistic forms with the least phonetic content are selected for the
-least amount of processing effort.
+more attenuated forms such as pronouns and zero anaphora (i.e., there is
+no explicit mention of the referent) since the mental status of the
+expressed referent is already at the center of attention, as indicated
+by the present contextual information (Chafe, 1976). The Givenness
+Hierarchy is akin to other previously proposed models of reference
+control within the topicality and saliency paradigm such as the
+Accessibility Hierarchy (Ariel, 1990), Topic Continuity (Givon, 1983),
+and Familiarity Scale (Prince, 1981). Gundel et al. (1993) argued that
+the distribution of forms is not random but abides by the Gricean Maxim
+of Quantity (Grice, 1975). That is to say, individuals must make the
+most informative and succinct choice possible to retain economy of form.
+When multiple expressions are available for the same referent as in an
+utterance like (\ref{ex:1}), phonetically simplest form is selected with
+the least redundant information based on an economy principle. In the
+utterance, REs for ``Eddie'' are given in bold. All three REs (nominal,
+pronominal, and zero anaphora) refer to the same entity; however, the
+quantity of marking decreases as the referent's accessibility increases.
+On this ground, for highly accessible referents (i.e., those that are
+highly active in the mind), linguistic forms with the least phonetic
+content are selected for the least amount of processing effort.
 
@@
 by an adjective or demonstrative. Maintenance, on the other hand, is
-typically marked with pronominals and null markers since the continued
+typically marked with pronominals and zero anaphora since the continued
 referent is still highly accessible to the addressee.
@@
 Czubek, 2017; Frederiksen and Mayberry, 2016; Hodge et al., 2019;
-Morgan, 2005; Perniss and Özyürek, 2015; Swabey, 2002). Building on
-Gundel et al. (1993)'s work, Swabey (2002) demonstrated that, like other
-spoken languages such as Russian, Chinese, and Japanese, American Sign
+Morgan, 2005; Perniss and Özyürek, 2015; Swabey, 2002). Building on the
+work by Gundel et al. (1993), Swabey (2002) demonstrated that, like
+other languages such as Russian, Chinese, and Japanese, American Sign
 Language (ASL) signers use bare nouns for introduced and re-introduced
-discourse contexts whereas null marking is used for referents which are
+discourse contexts whereas zero anaphora is used for referents which are
 ``in-focus'' or very accessible to the addressee. Overt REs like
@@
 concluded that ``native'' signers in her study abided by the Gricean
-maxim of quantity and were thus economic in their reference tracking.
-Similar findings have been reported for German Sign Language (Deutsche
-Gebärdensprache - DGS) (Perniss and Özyürek, 2015) and British Sign
-Language (BSL) (Morgan, 2005). It is important to note, however, that
-these research either collapsed native and late signers into a single
+maxim of quantity and thus used more economic forms of reference
+tracking. Similar findings have been reported for German Sign Language
+(Deutsche Gebärdensprache - DGS) (Perniss and Özyürek, 2015) and British
+Sign Language (BSL) (Morgan, 2005). It is important to note, however,
+that this research either collapsed DoD and DoH signers into a single
 group and labeled them as ``native-like'' or explicitly compared native
-deaf signers with late deaf signers (with language deprivation) in terms
-of narrative competence.
+deaf signers with late deaf signers (with delayed exposure to a sign
+language) in terms of narrative competence.
 
@@
 constructions (bare nouns, fingerspelled nouns, and modified nouns),
-null markers (plain verb, agreement verb, and constructed action),
+null or zero
+anaphors\footnote{Zero anaphora does not refer to the lack of a referent.
+It only alludes to the non-presence of a more explicit anaphoric form for the referent such as a nominal, pronominal, or classifiers.}
+(predicated by a plain verb, an agreement verb, or constructed action),
 classifier predicates (whole entity classifiers, body part classifiers,
@@
 (pronominal index or IX) (Frederiksen and Mayberry, 2016). This interim
-section describes such constructions attested for sign languages under
-four main headings (nominals, pronouns, classifiers, and null marking).
+section describes such forms attested for sign languages under four main
+headings (nominals, pronouns, classifiers, and zero anaphora).
 
@@
 
-In TİD and many other sign languages, it is possible to use bare noun
-phrases as in (\ref{ex:6a}) as well as fingerspelled nouns as in
-(\ref{ex:6c}). Fingerspelling refers to the use of the manual alphabet
-which is derived or borrowed from a surrounding written language (Kubuş,
-2008). Nouns can also be modified by classifiers as in (\ref{ex:6b})
-(Frederiksen and Mayberry, 2016) or by a prenominal or postnominal IX
-functioning as a determiner or a demonstrative as in (\ref{ex:6d})
-(Maclaughlin, 1997).
-
-\hypertarget{pronouns}{%
-\subsubsection{Pronouns}\label{pronouns}}
-
-The IX sign (extended index finger) has various functions in sign
-languages, one of which is the pronominal use which refers to
-established nominals by pointing to a previously assigned abstract or
-physical locus in the signing space when re-introducing a topic in
-discourse (Emmorey, 1996). The same IX function has also been identified
-for TİD (Sevinç, 2006). The example in (\ref{ex:7}) illustrates a
-spatially-anchored pronominal reference for topic re-introduction. These
-signs are thus rarely used for the first mention of an entity (Perniss
-and Özyürek, 2015). Other interpretations consider pointing signs
-without noun phrases as examples of demonstratives and not personal
-pronouns (Koulidobrova and Lillo-Martin, 2016). It is also important to
-mention here that in some spoken languages (e.g., German and Dutch),
-demonstratives can also refer to persons (Bosch and Umbach, 2007) and
-behave similarly to personal pronouns. However, for the purposes of the
-present study, we will refer to IX constructions as pronominal IX and
-particularly examine their non-deictic or personal pronoun realizations
-as a possible discourse strategy to mainly mark topic shift among native
-and late signers of TİD.
-
-\hypertarget{classifiers}{%
-\subsubsection{Classifiers}\label{classifiers}}
-
-Classifier predicates are highly complex morphological constructions
-which are very common in sign languages (Supalla, 1986). These are
-manually signed iconic representations of animate or inanimate objects,
-and they embody several properties pertaining to the nominal referent
-including its location in the signing space, movement, path, size, and
-shape (Talmy, 2003). As in most sign languages hitherto studied,
-classifier constructions are also commonly found in TİD (see Dikyuva et
-al., 2015 for a review). Following the classification of
-Engberg-Pedersen (1993), we distinguish between four types of classifier
-constructions. Whole entity classifiers (WECL) represent the entirety of
-any referent whereas body part classifiers (BPCL) only reflect a
-specific part of an animate entity (e.g., limbs, head, tongue, teeth,
-and so forth). Handling classifiers (HCL\footnote{As discussed before,
-  an HCL is transitive in nature and depicts the handling action
-  performed by ananimate agent. This is why most research tend to
-  consider HCLs as an extension to constructed action (Cormier et al.,
-  2013; Dudis, 2004).}) occur with transitive verbs and depict how an
-object is handled or manipulated by an agent. Extension classifiers
-(ExtCL), on the other hand, denote the size and shape of an entity by
-following a trajectory to trace its shape. We exemplify these
-constructions in (\ref{ex:8}).
-
-\begin{multicols}{2}
+In TİD and many other sign languages, it is possible to use bare nouns
+as in \ref{ex:2a} as well as fingerspelled nouns as in \ref{ex:2b}.
+Fingerspelling refers to the use of the manual aphabet to spell all or
+part of the letters of a word borrowed from a surrounding written
+language (Kubuş, 2008). Nouns can also be modified by a prenominal or
+postnominal extended index finger pointing sign, glossed as IX,
+functioning as a determiner or a demonstrative as in \ref{ex:2c}
+(Maclaughlin, 1997) or by a modifying classifier within the same NP as
+in \ref{ex:2d} (Frederiksen and Mayberry,
+2016).\footnote{Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions (Lillo-Martin and Klima, 1990).
+In addition, the pointing sign is used to referentially anchor a referent to a location in signing space.}
+
+\newpage
 \begin{exe}
@@
 \begin{xlist}
-\ex \label{ex:6a} Full noun \\\glll
-\includegraphics[width=50pt]{pictures/1a_1.png}\\ \textsc{mouse\footnotemark} \\ \\ 
-\ex \label{ex:6b} Noun modified by classifier \\\glll
-\includegraphics[width=50pt]{pictures/1c_2.png} \includegraphics[width=50pt]{pictures/1c_1.png} \\ \textsc{soap} \textsc{soap.cl} \\ \\ 
-\ex \label{ex:6c} Fingerspelled noun\\\glll 
-\includegraphics[width=50pt]{pictures/1b_1.png} \includegraphics[width=50pt]{pictures/1b_2.png} \includegraphics[width=50pt]{pictures/1b_3.png}\\ \textsc{t-} \textsc{o-} \textsc{m} \\ \\
-\ex \label{ex:6d} Noun modified by index (\textsc{IX})\\\glll 
-\includegraphics[width=50pt]{pictures/1d_1.png} \includegraphics[width=50pt]{pictures/1d_2.png}\\ \textsc{ix} \textsc{cat} \\ \\
+\ex \label{ex:2a} Full noun \\\glll
+{} \includegraphics[width=35pt]{pictures/p1.png} {} \\ \textbf{Gloss:} \textsc{mouse\footnotemark} \textsc{sick} \\ 
+\textbf{Status:} Intro. \\
+\glt \textbf{Context:} Narration 6, Event 1. Here, the signer introduces the mouse, Jerry, at the beginning of the narration with a bare noun.
+\glt \textbf{Translation:} There is a sick mouse.
+\vspace{0.35cm}
+
+\ex \label{ex:2b} Fingerspelled noun \\\glll
+{} \includegraphics[width=40pt]{pictures/p2.1.png} \includegraphics[width=40pt]{pictures/p2.2.png} \includegraphics[width=40pt]{pictures/p2.3.png} {} \\ \textbf{Gloss:} \textsc{t-} \textsc{-o-} \textsc{-m} \textsc{come.back} \\
+\textbf{Status:} Re-intro. \\
+\glt \textbf{Context:} Narration 7, Event 2.
+The signer introduces Tom two sentences before.
+She introduces the black cat in the preceding clause and then here re-introduces Tom by using fingerspelling.
+\glt \textbf{Translation:} Tom comes back.
+\vspace{0.35cm}
+
+\ex \label{ex:2c} Noun modified by index (IX) \\\glll
+{} \includegraphics[width=45pt]{pictures/p3.1.png} \includegraphics[width=40pt]{pictures/p3.2.png} {} {} {} \\ \textbf{Gloss:} \textsc{ix} \textsc{cat} \textsc{pipe.extend} \textsc{go.through} \textsc{exit} \\
+\textbf{Status:} Re-intro. \\
+\glt \textbf{Context:} Narration 1, Event 6. The signer introduces the cat in the first clause of the first event and re-introduces it several times including this last event of the narrative.
+\glt \textbf{Translation:} The cat goes through and exits the thin pipe.
+\vspace{0.35cm}
+
+\ex \label{ex:2d} Noun modified by classifier (CL) \\\glll
+{} \includegraphics[width=35pt]{pictures/p4.1.png} \includegraphics[width=35pt]{pictures/p4.2.png} \\ \textbf{Gloss:} \textsc{soap} \textsc{soap.cl} \\ 
+\textbf{Status:} Intro. \\
+\glt \textbf{Context:} Narration 5, Event 3. Here the signer introduces the soap to the discourse. The signer produces the lexical sign, \textsc{soap}, plus a classifier, \textsc{soap.cl}.
+\glt \textbf{Translation:} The soap, which is like this.
 \end{xlist}
 \end{exe}
-\end{multicols}
+
+\footnotetext{We use small capitals to indicate sign glosses.}
+
+\hypertarget{referential-use-of-space-and-pronouns}{%
+\subsubsection{Referential Use of Space and
+Pronouns}\label{referential-use-of-space-and-pronouns}}
+
+In addition to being a modifier of a noun, the IX sign can also be
+pronominal by referring to established nominals through pointing to a
+previously assigned abstract or physical locus in the signing space when
+re-introducing a topic in discourse (Emmorey, 1996). This function has
+also been identified for TİD (Sevinç, 2006). The example in \ref{ex:3}
+illustrates this spatially-anchored pronominal use.
+
 \ea 
-\ea \label{ex:7} Pronominal \\\glll
-\includegraphics[width=50pt]{pictures/2_1.png} \includegraphics[width=50pt]{pictures/2_2.png} \\ \textsc{ix} \textsc{sick} \\ \\
+\ea \label{ex:3} Pronominal \\\glll
+{} \includegraphics[width=35pt]{pictures/p5.png} {} {} {} {} \\ \textbf{Gloss:} \textsc{ix} \textsc{look} \textsc{cat.sad(constructed action)} \textsc{sad} \textsc{sick} \\ 
+\textbf{Status:} Re-intro. \\
+\glt \textbf{Context:} Narration 9, Event 1: The signer introduces the mouse then signs an action of the cat. Here, he re-introduces the mouse with a pointing sign that functions as a pronoun.
+\glt \textbf{Translation:} It (the mouse) looks sad and sick.
 \z \z
-\clearpage
-\begin{multicols}{4}
-\begin{exe}
-\ex \label{ex:8}
-\begin{xlist}
-\ex Whole entity \\\glll 
-\includegraphics[width=50pt]{pictures/3e.png} \\ \textsc{fish.fall}\\ \\
-\ex Body part \\\glll
-\includegraphics[width=50pt]{pictures/3b.png} \\ \textsc{walk} \\ \\ 
-\ex Handling \\\glll 
-\includegraphics[width=50pt]{pictures/3c.png} \\ \textsc{hold} \\ \\
-\ex Extension \\\glll 
-\includegraphics[width=50pt]{pictures/3d_1.png} \includegraphics[width=50pt]{pictures/3a.png}\\ \textsc{plank} {\textsc{dh: mouse}} \\ {} {\textsc{ndh: tunnel}}\\ \\
-\end{xlist}
-\end{exe}
-\end{multicols}
-\begin{multicols}{3}
+
+These pronominal signs are rarely used for the first mention of an
+entity (Perniss and Özyürek, 2015). Other researchers consider pointing
+signs without noun phrases as examples of demonstratives and not
+personal pronouns (Koulidobrova and Lillo-Martin, 2016). It is also
+important to mention here that in some spoken languages (e.g., German
+and Dutch), demonstratives can also refer to persons (Bosch and Umbach,
+2007) and behave similarly to personal pronouns. However, for the
+purposes of the present study, we will refer to IX signs as pronominal
+IX and particularly examine their non-deictic or personal pronoun
+realizations as a possible discourse strategy to mainly mark topic shift
+(re-introduction) among DoD and DoH signers of TİD.
+
+\hypertarget{classifiers}{%
+\subsubsection{Classifiers}\label{classifiers}}
+
+In sign languages, a verbal classifier morpheme is morpho-phonologically
+encoded by a handshape. This handshape is articulated co-temporally with
+a movement that encodes the verbal meaning (Benedicto and Brentari,
+2004; Supalla, 1990). Semantically, the classifiers denote sets of
+objects based on criteria such as animacy, size, shape, body-part,
+texture and sometimes function (Talmy, 2003). The classifier articulated
+simultaneously with a movement is called a classifier predicate and the
+sentence in which such a predicate is used is usually called a
+classifier construction in sign language research. As in most sign
+languages hitherto studied, classifier constructions are also commonly
+found in TİD (see Dikyuva et al., 2015 for a review). Following the
+classification of (Engberg-Pedersen, 1993), in TİD, we distinguish
+between four types of classifier handshapes. Whole entity classifiers
+(WECL) represent the entirety of a referent (e.g.~an entire flat
+horizontal entity, or an upright entity) (e.g., \ref{ex:4a}) whereas
+body part classifiers (BPCL) represent a specific part of an animate
+entity (e.g., limbs, head, mouth, tongue, or teeth, etc.)
+(e.g.,\ref{ex:4b}). Handling classifiers (HCL) occur with transitive
+verbal roots and represent both the size, shape and surface of an object
+and how that object is handled or manipulated by an agent (e.g.,
+\ref{ex:4c}). Extension classifiers (ExtCL), on the other hand,
+represent the size and shape of an entity by following a trajectory to
+trace its shape (e.g., \ref{ex:4d}). We exemplify these handshapes with
+the discourse status of the referent they track and a description of the
+previous discourse when available.
+
 \begin{exe}
@@
 \begin{xlist}
-\ex \label{ex:9a} Constructed action \\\glll 
-\includegraphics[width=50pt]{pictures/4c.png} \\ \textsc{struggle} \\ \\
-
-\ex \label{ex:9b} Plain verb  \\\glll
-\includegraphics[width=50pt]{pictures/4a_2.png}  \\ \textsc{enter}\\ \\
-
-\ex \label{ex:9c} Agreement verb  \\\glll 
-\includegraphics[width=50pt]{pictures/4b_1.png} 
-\includegraphics[width=50pt]{pictures/4b_2.png} \\ {\textsc{$_{1}$hit$_{3}$}} {} \\ \\ 
+\ex \label{ex:4a} Whole entity classifier  (WECL) \\\glll
+{} \includegraphics[width=40pt]{pictures/p6.png} \\ \textbf{Gloss:} \textsc{two.mice.go.together} \\ 
+\textbf{Status:} Maintain. \\
+\glt \textbf{Context:} Narration 4, Event 2: The signer introduces Jerry and the robot mouse in the preceding clause. She maintains reference to them by using a whole entity classifier on each hand. 
+\glt \textbf{Translation:} (Jerry and the robot mouse from the previous context) go together.
+\vspace{0.35cm}
+
+\ex \label{ex:4b} Body part classifier (BPCL) \\\glll
+{} \includegraphics[width=40pt]{pictures/p7.png} \\ \textbf{Gloss:} \textsc{cat.move.thin.legs} \\ 
+\textbf{Status:} Maintain. \\
+\glt \textbf{Context:} Narration 1, Event 6: The signer re-introduces the cat three clauses before and maintains reference to him in the two clauses before and here with a body part classifier that refers to the legs of the cat.
+\glt \textbf{Translation:} (The cat) walks (i.e. moving his thin legs forward).
+\vspace{0.35cm}
+
+\ex \label{ex:4c} Handling classifier (HCL) \\\glll
+{} {} {} {} {} \includegraphics[width=40pt]{pictures/p8.png} \\ \textbf{Gloss:} {} \textsc{rope see-3a} \textsc{soon} \textsc{go} \textsc{cat.pull.rope} \\ 
+\textbf{Status:} Maintain. \\
+\glt \textbf{Context:} Narration 5, Event 2: The signer introduces Tom three clauses before.
+She maintains reference to him in the preceding two clauses and here she still maintains reference to him by using a handling classifier on both hands.
+\glt \textbf{Translation:} When (he/Tom) sees the rope, (he) soon goes to pull it.
+\vspace{0.35cm}
+
+\ex \label{ex:4d} Extension classifier (ExtCL) \\\glll
+{} \includegraphics[width=80pt]{pictures/p9.png} \\ \textbf{Gloss:} \textsc{cat.extend} \\ 
+\textbf{Status:} Maintain. \\
+\glt \textbf{Context:} Narration 1, Event 6: The signer re-introduces Tom two clauses before and she maintains reference to him with constructed action in the previous clause and here with an extension classifier.
+\glt \textbf{Translation:} (The cat) extends thin.
 \end{xlist}
 \end{exe}
-\end{multicols}
-\footnotetext{We use small capitals to indicate sign glosses.}
-
-\hypertarget{null-marking}{%
-\subsubsection{Null marking}\label{null-marking}}
+
+\hypertarget{zero-anaphora}{%
+\subsubsection{Zero Anaphora}\label{zero-anaphora}}
 
@@
 allow null subjects and/or objects and make use of null arguments when
-referring to highly accessible referents which are still active in the
+referring to highly accessible referents that are still active in the
 mental representation of the addressee. In signed discourse, this can be
 achieved by employing certain verbal constructions or constructed action
-(CA). The latter construction exemplified in (\ref{ex:9a}) refers to
-cases in which the signers engage with the space in front of themselves
-as if they are interacting with the objects and characters in the story.
-This is often achieved with the signer taking on the role of a referent
-in the narrative and performing the actions (Metzger, 1995). CA is often
-marked by certain non-manual cues such as the facial expression and the
-breaking of the eye contact. This is to say, the narrator's head, face,
-torso, arms, and as an optional marker, hands represent those of the
-associated referent(s) (Smith and Cormier, 2014). As for the verbal
-constructions, to identify the predicate types in ASL, Padden (1986)
-suggested a triadic categorization of verbs in sign languages: plain
-verbs, agreement verbs, and spatial verbs. According to this
-classification, plain verbs (\ref{ex:9b}) do not agree with the subject
-or the object. This means that they are not inflected for number or
-person. This contrasts with agreeing verbs (\ref{ex:9c}) the starting or
-the ending point of which can be modified to specific loci for person
-and number agreement while spatial verbs (as in \textsc{ix book put} `I
-placed the book there') agree with locative arguments but like plain
-verbs, they are not inflected for number or person. TİD too licenses
-null arguments. Following this classification, Sevinç (2006) argued that
-although it is common in TİD to drop arguments, there were certain
-limitations on the null marking of the arguments for agreement and plain
-verbs.
-
-\hypertarget{signed-vs.-spoken-res}{%
-\subsection{Signed vs.~Spoken REs}\label{signed-vs.-spoken-res}}
-
-Commonly in both spoken and sign languages nouns and noun phrases are
-used as REs. Spoken languages use pronouns, determines and
-demonstratives where sign languages use a pointing sign, usually glossed
-as IX, which assumes similar functions (Lillo-Martin and Klima, 1990).
-In addition, the pointing sign is used to referentially anchor a
-referent to a location in signing space. For instance, when the signer
-introduces \textsc{elvan} for the first time, s/he can assign her to the
-right. This referential locus can later be used by agreement verbs and
-classifier predicates (more below). For instance, in (\ref{ex:2a}) there
-are two referents, \textsc{elvan}, who is referentially-assigned to the
-right side of the signer, represented with the ``A'' subscript and
-\textsc{demet}, who is referentially-assigned to the left side of the
-signer, represented with the ``B'' subscript. When the signer makes a
-contrast in (\ref{ex:2b}), as a pronominal use, she points to the right
-side to refer to ELVAN and to the left to refer to \textsc{demet}. We
-schematically represent the referential locus assigned to the referents
-in (\ref{ex:3}).
+(CA). The latter construction exemplified in \ref{ex:5a} refers to cases
+in which the signers engage with the space in front of themselves as if
+they are characters in the story interacting with objects and other
+characters. This is often achieved with the signer taking on the role of
+a referent in the narrative and performing the actions (Metzger, 1995).
+CA is often marked by certain non-manual cues such as the facial
+expression and the breaking of the eye contact. This is to say, the
+narrator's head, face, torso, arms, and as an optional marker, hands
+represent those of the associated referent(s) (Smith and Cormier, 2014).
+As for the verbal constructions, to identify the predicate types in ASL,
+Padden (1986) suggested a triadic categorization of verbs in sign
+languages: plain verbs, agreement verbs, and spatial verbs. According to
+this classification, plain verbs (as in \ref{ex:5b}) do not agree with
+the subject or the object. This means that they are not inflected for
+number or person. This contrasts with agreeing verbs (as in \ref{ex:5c})
+the starting or the ending point of which can be modified to specific
+loci for person and number agreement while spatial verbs (as in
+\textsc{ix book put} `I placed the book there') agree with locative
+arguments but like plain verbs, they are not inflected for number or
+person. TİD too licenses null arguments. Following this classification,
+Sevinç (2006) argued that although it is common in TİD to drop
+arguments, there were certain limitations on the null marking of the
+arguments for agreement and plain verbs.
+
+In terms of referential accessibility, zero anaphors predicated by such
+constructions in narratives are assumed to rank last in the hierarchy,
+preceded by nominal classifiers and then pronouns and other nominal
+forms. The reason that we distinguish zero anaphora from classifer
+handshapes or predicates (which can also be considered implicit markers
+of reference) is because classifiers typically denote certain
+topographic or semantic properties (e.g., size and shape) of the
+referent noun due to their iconic nature. They are not as explicit as
+pronouns or nouns since they do not target a specific individual (e.g.,
+\textsc{eddie}) in a set but rather they refer to a set of individuals
+that are semantically similar to each other (Morgan, 2005). For example,
+the extended index finger can be used to refer to a human entity.
+Finally, zero anaphors predicated by constructed action, agreement and
+plain verbs are least explicit forms of referent tracking since the
+implicit nature of such constructions do not easily restrict the
+referent set based on cues such as size, shape or semantic category. In
+other words, zero anaphora requires preceding context or information the
+most compared to other forms.
 
@@
 \begin{xlist}
-    \ex \label{ex:2a}
-    \textsc{lab place elvan ix\textsubscript{a} ten years work. demet ix\textsubscript{b} last year start.} \\
-    ‘In the lab, Elvan has been working for 10 years. Demet started last year.’
-    \\
-    \ex \label{ex:2b}
-    \textsc{ix\textsubscript{a} e-l-a-n know. ix\textsubscript{b} learn.}\\
-    ‘She (Elvan) knows ELAN (the annotation program). She (Demet) is learning.’
+\ex \label{ex:5a} Constructed action \\\glll
+{} \includegraphics[width=40pt]{pictures/p10.png} \\ \textbf{Gloss:} \textsc{mouse.stay.like.this} \\ 
+\textbf{Status:} Maintained. \\
+\glt \textbf{Context:} Narration 6, Event 9: The signer re-introduces Jerry in the previous clause and maintains reference to him by using this constructed action.
+\glt \textbf{Translation:} (The mouse) stays like this.
+\vspace{0.35cm}
+
+\ex \label{ex:5b} Plain verb \\\glll
+{} \includegraphics[width=80pt]{pictures/p11.png} \\ \textbf{Gloss:} \textsc{run.away} \\ 
+\textbf{Status:} Re-introduced. \\
+\glt \textbf{Context:} Narration 5, Event 4: In this sentence, the signer re-introduces Jerry with the plain verb \textsc{run.away}.
+\glt \textbf{Translation:} (Jerry) runs away.
+\vspace{0.35cm}
+
+\ex \label{ex:5c} Agreement verb \\\glll
+{} \includegraphics[width=80pt]{pictures/p12.png} \\ \textbf{Gloss:} \textsc{3a(mouse)-hit-3b(cat)} \\ 
+\textbf{Status:} Maintained. \\
+\glt \textbf{Context:} Narration 8, Event 2: The signer introduces Jerry in the preceding clause and maintains reference to him by using the agreement verb \textsc{hit}.
+\glt \textbf{Translation:} (Jerry) hits (Tom).
+\vspace{0.35cm}
+
 \end{xlist}
 \end{exe}
-
-    \begin{exe}
-    \ex \label{ex:3} Signing space, signer, example referential loci for the referents \textsc{elvan} (a) and \textsc{demet} (b) \\
-\tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
-
-\begin{tikzpicture}[x=0.75pt,y=0.75pt,yscale=-1,xscale=1]
-%uncomment if require: \path (0,300); %set diagram left start at 0, and has height of 300
-
-%Curve Lines [id:da7006577339157659] 
-\draw    (100.25,77.81) .. controls (136.19,41.85) and (183.62,42.17) .. (222.5,78.13) ;
-%Straight Lines [id:da8428827309782937] 
-\draw    (112.9,128.8) -- (210.21,128.8) ;
-%Curve Lines [id:da39290455666055124] 
-\draw    (222.19,100.13) .. controls (221.89,103.9) and (217.55,123.77) .. (210.21,128.8) ;
-%Curve Lines [id:da9909903731270634] 
-\draw    (101.17,100.45) .. controls (100.86,104.22) and (108.2,124.02) .. (112.9,128.8) ;
-
-% Text Node
-\draw (88.75,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-b }};
-% Text Node
-\draw (91.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Demet)}};
-% Text Node
-\draw (208.25,80.75) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Locus-a }};
-% Text Node
-\draw (212.75,90.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\tiny (Elvan)}};
-% Text Node
-\draw (144,115.25) node [anchor=north west][inner sep=0.75pt]   [align=left] {{\footnotesize Signer}};
-\end{tikzpicture}
-    \end{exe}
-
-Grammatical gender as in Romance languages or noun classes as in Bantu
-languages are some reference tracking tools that are familiar from
-spoken languages. A somehow less known reference-tracking tool in spoken
-languages is \emph{verbal classifiers}, which morphologically attach to
-the verb and often cross-reference a core argument in a clause
-(Aikhenvald, 2017). As outlined in the previous section, classifiers
-denote sets of objects based on semantic criteria such as size, shape,
-body-part, texture and sometimes function. Usually, the verbal
-classifier and the cross-referenced noun co-occur in the same clause
-when the argument is introduced into discourse. The classifier can track
-the argument on its own on subsequent occurrences. In sign languages, a
-verbal classifier morpheme is encoded by a \emph{handshape} that is
-articulated phonologically co-temporally with a movement, which encodes
-the verbal meaning (Benedicto and Brentari, 2004; Supalla, 1990). These
-two infused morphemes form a complex predicate. Some classifier
-handshapes and the set of objects that they denote are presented below
-in (\ref{ex:4}).
-
-\begin{exe}
-\ex \label{ex:4}
-\includegraphics[width=25pt]{pictures/flathand.png}: flat entities, surfaces, vehicles.\\
-\includegraphics[width=25pt]{pictures/flat.png}: human agent holding a flat entity.\\
-\includegraphics[width=25pt]{pictures/small_legged.png}: small animal legs.\\
-\includegraphics[width=25pt]{pictures/index.png}: thin entities (like a pencil); upright beings.\\
-\includegraphics[width=25pt]{pictures/hold.png}: human agent holding a thin entity (like a pencil or a branch of flower). 
-\end{exe}
-
-In (\ref{ex:5}), in the first clause, a referential locus, i.e.~locus-A,
-is assigned to the locative source argument \textsc{cage}. In the second
-clause, the core argument \textsc{bird} is cross-referenced with the
-classifier handshape, i.e.~small-legged animal. It flies from the
-referential locus of the \textsc{cage} (locus-a) to the referential
-locus of the goal \textsc{tree} (locus-B).
-
-\begin{exe}
-    \ex \label{ex:5}
-\begin{xlist}
-    \ex \textsc{cage} assigned to the referential locus-a. \\ \\
-    \textsc{garden bird cage-a open} \\
-    ‘In the garden, there is a cage and it is open.’
-    \\
-
-    \ex \textsc{tree} assigned to the referential locus-b. The source and goal tracked on the predicates \textsc{fly} and \textsc{land}. \\ \\
-    \textsc{bird fly[manner]-a tree-b land cl-b}\\
-    ‘The bird flies (from the cage assigned to the left) to a tree (assigned to the right).’
-    \end{xlist}
-    \end{exe}
-    \hfill (Adapted from Gökgöz and Sevgi 2020:318)
 
@@
 Signes Catalana - LSC), Bel et al. (2015) found that hearing signers
-resort to more overt forms (nominals and pronouns) than native signers
-to avoid potential ambiguities among referents. Nevertheless, their
-results did not reach statistical significance for the interaction of
-discourse status and acquisition group. A similar line of research came
-from Frederiksen and Mayberry (2019) who found a statistically
-significant difference between both groups, that is native signers and
-hearing L2 learners of ASL, only for classifier use especially in
-re-introduced contexts. L2 learners in the study used more nominals and
-fewer instances of null marking in number, indicative of a limited
-effect of overexplicitness among second language learners. However,
-these studies treated late and native deaf signers as a single group.
-That is, the control group consisted of both native signers who had one
-or more deaf parents and also native-like (Cormier et al., 2012) late
-deaf signers who had hearing parents and their age of acquisition
-varied. Given the preliminary evidence that topic continuity and
-salience marking is also difficult for second language learners
-regardless of modality, it has been suggested under the Interface
-Hypothesis that the syntax-discourse interface is vulnerable and
-especially challenging to acquire (Sorace, 2011) for second language
-learners. Similarly, Sorace and Serratrice (2009) discuss that this
-interface entails much more than morphosyntactic knowledge but also
-includes cognitive and pragmatic resources that help provide discourse
-coherence. They propose that the second language discourse is
-over-redundant. This means that second language learners produce more
-overt pronouns or fuller forms of language for highly accessible
-referents possibly as a compensatory strategy in an attempt to reduce
-the increasing load on cognitive processing.
-
-Still, only very few researchers examined whether linguistically
-deprived signers (i.e., deaf signers with delayed exposure to a sign
-language) also display redundancy in their reference tracking. Becker
-(2009) examined 6 DGS signing deaf children's (3 native and 3 late)
-narrative abilities. She reported that late signing children generated
-more explicit forms such as nominals when referring to the characters in
-the story compared to native learners, concluding that first language
-deprivation results in an inadequate representation of the addressee and
-less efficient use of morphosyntax. Comparably, Cormier et al. (2013)
-investigated the use of CA in BSL among 15 deaf native, early, and late
-adult signers (5 participants in each group) and found that late signers
-were redundant in topic continuity. In other words, the late signers
-produced more explicit forms like nominals, and fewer instances of null
-marking when maintaining previous referents. Most relevantly to the
-present study, Gür (2018) studied the development of narrative skills
-among 10 native and 10 late deaf children and deaf adult signers of TİD.
-She found that native-signing children and adults had more instances of
-explicit referent introduction, set the scene more frequently, and
-employed CA in an appropriate manner more than corresponding
-late-signing children and adults. Although this study presented
-preliminary findings on age of acquisition effects on narrative
-competence, it does not investigate how native and late signers employ
-different linguistic tools based on the accessibility of the referent.
-Therefore, to our knowledge, there is a lack of research directly
-examining the effects of first language deprivation on deaf adult
-signers' selection of REs in signed narratives.
-
-Considering this, it is not very clear whether the negative effects on
-narrative abilities observed among late signing deaf children hold when
-they become adults. Understanding the impact of language deprivation or
+resort to more overt forms (nominals and pronouns) than native deaf
+signers to avoid potential ambiguities among referents. Nevertheless,
+their results did not reach statistical significance for the interaction
+of discourse status and acquisition group. However, the control group
+here consisted of both deaf signers who had one or more deaf parents and
+also deaf signers who had hearing parents and their age of acquisition
+varied. Out of 11 deaf L1 signers in Bel et al.'s (2015) control sample,
+5 had deaf parents and acquired LSC from birth, and the other 6 had
+hearing parents but their exposure to LSC started before age 5. A
+similar line of research came from Frederiksen and Mayberry (2019) who
+found a statistically significant difference between both groups, that
+is DoD signers and hearing L2 learners of ASL, only for classifier use
+especially in re-introduced contexts. L2 learners in the study used more
+nominals and fewer instances of zero anaphora in number, indicative of a
+limited effect of overexplicitness among second language learners.
+Although Frederiksen and Mayberry's control group included some signers
+with hearing parents, they had all learned to sign from birth since they
+received sign language input from older deaf siblings.
+
+Still, only very few researchers explicitly compared DoD (i.e., deaf
+signers with native exposure to a sign language) and DoH (i.e., deaf
+signers with delayed exposure to a sign language) also display
+redundancy in their reference tracking. Becker (2009) examined 6 DGS
+signing deaf children's (3 native and 3 late) narrative abilities. She
+reported that late signing children generated more explicit forms such
+as nominals when referring to the characters in the story compared to
+native learners, concluding that first language deprivation results in
+an inadequate representation of the addressee and less efficient use of
+morphosyntax. Comparably, Cormier et al. (2013) investigated the use of
+CA in BSL among 15 deaf native, early, and late adult signers (5
+participants in each group) and found that DoH signers were redundant in
+topic continuity. In other words, the DoH signers produced more explicit
+forms like nominals, and fewer instances of zero anaphora when
+maintaining previous referents. Most relevantly to the present study,
+Gür (2018) studied the development of narrative skills among 10 native
+and 10 late deaf children and deaf adult signers of TİD. She found that
+native-signing children and adults had more instances of explicit
+referent introduction, set the scene more frequently, and employed CA in
+an appropriate manner more than corresponding late-signing children and
+adults. Although this study presented preliminary findings on age of
+acquisition effects on narrative competence, it does not investigate how
+DoD and DoH signers employ different linguistic tools based on the
+accessibility of the referent. Therefore, to our knowledge, there is a
+lack of research directly examining the effects of first language delay
+especially among deaf adult signers with non-signing hearing parents on
+the selection of REs in signed narratives.
+
+Considering this, it is also not very clear whether the age of
+acquisition effects on narrative abilities, observed among late signing
+deaf children, hold when they become adults. Understanding the impact of
 delayed first language learning on narrative competence is crucial for
-three reasons. First, studying this population will shed light on the
-possible effect of lack of early language experience on future cognitive
-and pragmatic abilities since the narrative competence is dependent on
-such faculties. Second, examining reference tracking strategies in
-particular will also inform about the linguistic inventory that native
-and late deaf signers use to refer to entities. Finally, any possible
+three reasons. First, studying signers from this population will shed
+light on the possible effect of lack of early language experience on
+future cognitive and pragmatic abilities since the narrative competence
+is dependent on such faculties. Second, examining reference tracking
+strategies in particular will also inform about the linguistic inventory
+that DoD and DoH signers use to refer to entities. Finally, any possible
 difference between the two acquisition groups in terms of narrative
@@
 In an attempt at addressing this gap in the literature, the present
-study aims to investigate the reference tracking strategies of native
-and late adult signers in TİD narratives with a story-telling paradigm.
-We explore the forms available for referent tracking in TİD and examine
-the nature of RE selection as a possible indicator of pragmatic
-competence between our acquisition groups.
+study aims to investigate the reference tracking strategies of DoD and
+DoH adult signers in TİD narratives with a story-telling paradigm. We
+explore the forms available for referent tracking in TİD and examine the
+nature of RE selection as an indicator of possible pragmatic differences
+between the acquisition groups.
 
@@
 in signed narratives, we expect more explicit forms like nominals for
-referent introduction and and less explicit forms such as null marking,
+referent introduction and and less explicit forms such as zero anaphora,
 pronominal and/or classifier use for referent maintainance. For the
@@
 others that have lower accessibility will be tracked with more explicit
-REs. Also, based on previous research, we hypothesize that late signers
-will be more over-redundant in their reference tracking compared to
-native signers. This would translate to a preference by late signers for
-nominals over null marking especially in maintenance and re-introduction
-contexts. In contrast, we would expect that native signers use more null
-marking to track referents.
+REs. Also, based on previous research, we hypothesize that DoH signers
+will be more over-redundant in their reference tracking compared to DoD
+signers. This would translate to a preference by DoH signers for
+nominals over zero anaphora especially in maintenance and
+re-introduction contexts. In contrast, we would expect that DoD signers
+use more zero anaphora to track referents.
 
@@
 
-We tested 29 deaf adult signers, 15 with exposure to TİD from birth
-(i.e., native signers), 14 with exposure between ages 3 and 17 (i.e.,
-late signers) (\emph{M}\textsubscript{Age of acquisition} = 7.7 years,
-\emph{SD} = 3.3 years). The criterion to be classified as a late signer
-was to have hearing and non-signing caregivers. All the signers
-satisfying this criterion were regarded as late signers
-\footnote{Note that we are essentially comparing deaf of deaf and deaf of hearing adult signers.
-Although we did not set a cut-off point for nativeness, all the deaf of hearing participants indicated that they learned TİD after age 3.
-In order to capture in-group variation among late signers, we run an additional supplementary analysis in which we observed very wide credible intervals due to the small group size, rendering the results uninformative.
-For the analysis, see \ref{section:d}.}.
+We tested 29 deaf adult signers, 15 DoD signers with exposure to TİD
+from birth from deaf caregivers (e.g., parents or siblings), 14 DoH
+signers who had hearing caregivers and were exposed to TİD between ages
+3 and 17. (\emph{M}\textsubscript{Age of acquisition} = 7.7 years,
+\emph{SD} = 3.3 years). As a result, the criterion to be classified as a
+DoH signer was to have hearing and non-signing caregivers. All the
+signers satisfying this criterion were regarded as
+such.\footnote{Although we did not set a cut-off point for nativeness, all the deaf of hearing participants indicated that they learned TİD after age 3.
+In order to capture in-group variation among DoH signers, we run an additional supplementary analysis in which we observed very wide credible intervals due to the small group size, rendering the results uninformative.
+For the analysis, see \ref{section:d}.}
 All participants indicated that their preferred language of
@@
 information about the linguistic and educational background of the
-participants can be found in \ref{section:a}.
-
-All participating native signers (7 males, 8 females) were born into
-deaf families and started acquiring TİD from at least one of their
-caregivers after birth. The parents of 14 participating native signers
-were deaf. One native signer had a deaf father and a hearing mother. In
-addition to TİD, the home language for some native signers (n = 4) in
-early childhood included the use of home
-sign\footnote{Home signs are gesture-based communication systems that emerge between the deaf children and their primary caregivers.
+participants can be found in \ref{section:a}. 1 DoD signer was excluded
+from the data analysis because they self-reported to have signing deaf
+parents but also only used home sign (a more crude form of
+communication) at home. As a result, we present data from only 14 DoD
+signers.
+
+All participating DoD signers (7 males, 7 females) were born into deaf
+signing families and started acquiring TİD from at least one of their
+caregivers after birth. The parents of 14 participating DoD signers were
+deaf. One native signer had a deaf father and a hearing mother. In
+addition to TİD, the home language for some DoD signers (n = 4) in early
+childhood included the use of home
+sign.\footnote{Home signs are gesture-based communication systems that emerge between the deaf children and their primary caregivers.
 Although home sign communication may have internal consistency, they are not fully-flegged linguistic systems like natural languages.}.
 The age of the native participants ranged between 18 and 35 (\emph{M} =
-26.4 years, \emph{SD} = 5.0 years). This also corresponded to the total
-average years of TİD use for native signers. All native signers were
+26.4 years, \emph{SD} = 5.0 years) This also corresponded to the total
+average years of TİD use for DoD signers. All DoD signers were
 pre-lingually deaf, that is they were either born deaf or they had
 become deaf before the age of 3, with the exception of one signer who
-became deaf between the ages of 4 and 7. Native signers reported to
+became deaf between the ages of 4 and 7. DoD signers reported to
 complete at least high school education and their level of education
@@
 
-Comparably, late learners (8 males, 6 females) were all born into
-hearing families and started learning TİD later in life after age 3,
+Comparably, DoH signers (8 males, 6 females) were all born into hearing
+non-signing families and started learning TİD later in life after age 3,
 usually with the start of formal education. The age of sign language
-acquisition for late signers ranged between years 3-17 (\emph{M} = 8.7
-years, \emph{SD} = 3.5 years). The parents of all participating late
+acquisition for DoH signers ranged between years 3-17 (\emph{M} = 8.7
+years, \emph{SD} = 3.5 years). The caregivers of all participating DoH
 signers were hearing, and the reported home languages used in childhood
-mainly consisted of Turkish and home sign. The average age of late
+mainly consisted of Turkish and home sign. The average age of DoH
 signers was 34.6 years (\emph{SD} = 8 years) and ranged between 24 and
 50. The derived mean number of years for TİD use following its
-acquisition was comparable to native signers with 25.9 years (\emph{SD}
-= 7.9 years). Except for two signers who became deaf between the ages of
-4 and 7, all the other late signers were pre-lingually deaf. All late
+acquisition was comparable to DoD signers with 25.9 years (\emph{SD} =
+7.9 years). Except for two signers who became deaf between the ages of 4
+and 7, all the other DoH signers were pre-lingually deaf. All DoH
 signers completed at least middle school education and the year of
 schooling ranged between 8 and 16 years, and the average was 12.3 years
-(\emph{SD} = 2.8 years). All late participants attended at least one
-deaf school. While 8 participants attended two deaf schools (either
-primary and middle school, or middle and high school), 3 participants
-reported to attend three deaf schools. Using the same scale, late
-signers self-rated their abilities in TİD and Turkish. Both of their
-receptive and productive TİD abilities rated on average 4.9 points
-(\emph{SD} = 0.3 points). Late signers' mean rating for their reading
-abilities in Turkish was 3.5 points (\emph{SD} = 0.7 points) while their
-writing and speaking abilities rated on average 3.4 (\emph{SD} = 0.5
-points) and 3.1 points (\emph{SD} = 0.7 points), respectively.
-
-Native and late participants' sex, average years of schooling, TİD use
-following its acquisition, and their self-reported receptive and
-productive TİD ratings, and their reading, writing, and speaking ratings
-for Turkish did not statistically differ; Bayesian \emph{p}'s
-\textgreater{} 0.1. However, participants' ages statistically differed
-between the two acquisition groups; \emph{t}(27) = 3.4, Bayesian
-\emph{p} \textless{} 0.01. Overall, the native group consisted of
-younger participants. This was to ensure that the average years of TİD
-use among native signers approximated those of late signers.
+(\emph{SD} = 2.8 years). All of the DoH group attended at least one deaf
+school. While 8 participants attended two deaf schools (either primary
+and middle school, or middle and high school), 3 participants reported
+to attend three deaf schools. Using the same scale, the signers
+self-rated their abilities in TİD and Turkish. Both of their receptive
+and productive TİD abilities rated on average 4.9 points (\emph{SD} =
+0.3 points). DoH signers' mean rating for their reading abilities in
+Turkish was 3.5 points (\emph{SD} = 0.7 points) while their writing and
+speaking abilities rated on average 3.4 (\emph{SD} = 0.5 points) and 3.1
+points (\emph{SD} = 0.7 points), respectively.
+
+Both groups' sex, average years of schooling, TİD use following its
+acquisition, and their self-reported receptive and productive TİD
+ratings, and their reading, writing, and speaking ratings for Turkish
+did not statistically differ; Bayesian \emph{p}'s \textgreater{} 0.1.
+However, participants' ages statistically differed between the two
+acquisition groups; \emph{t}(27) = 3.4, Bayesian \emph{p} \textless{}
+0.01. Overall, the DoD group consisted of younger participants. This was
+to ensure that the average years of TİD use among DoD signers
+approximated those of DoH signers.
 
@@
 
-We used a story-telling paradigm to examine how native and late signers
-of TİD track reference in signed discourse. Participants were presented
+We used a story-telling paradigm to examine how DoD and DoH signers of
+TİD track reference in signed discourse. Participants were presented
 with 10 short video clips extracted from the cartoon \emph{Tom and
@@
  & PRO: Pronominal IX & An index sign used in isolation to refer to an entity in discourse\\
- & CL: Whole Entity Classifier & A manual construction that refers to the entirety of an entity\\
- & CL: Body Part Classifier & A manual construction that refers to a part of an entity\\
- & CL: Handling Classifier & A manual construction that refers to the handling of an item\\
- & CL: Extension Classifier & A manual construction that denotes the size and shape of an entity\\
+ & CL: Whole Entity Classifier & A handshape that refers to the entirety of an entity\\
+ & CL: Body Part Classifier & A handshape that refers to a part of an entity\\
+ & CL: Handling Classifier & A handshape that refers to the handling of an item\\
+ & CL: Extension Classifier & A handshape that denotes the size and shape of an entity\\
  & NULL: Constructed Action & A multifunctional construction where the signers take on the role of a referent and perform their actions\\
- & NULL: Plain Verb & A non-agreeing verb (e.g., LOVE)\\
- & NULL: Agreement Verb & A verbal construction the beginning and ending of which agree with person and number\\
+ & NULL: Plain Verb & A non-agreeing predicate (e.g., LOVE)\\
+ & NULL: Agreement Verb & A predicate the beginning and ending of which agree with the person and number\\
 \bottomrule
@@
 learner of TİD. We instructed the participants to narrate as they would
-to a deaf friend. All trials were recorded by a digital camera placed in
-front of the signers. The participants were instructed to watch the
-narrative clip as many times as they needed, and narrate the content of
-each clip either to the experimenter or the recording digital
-camera\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining we did not make it obligatory for participants to either narrate to the digital camera or to the experimenter for few reasons.
+to a deaf friend. We gave this instruction to make sure that signers
+naturally sign, and simply do not try to simplify or adjust their
+language to the level of the addressee. All trials were recorded by a
+digital camera placed in front of the signers. The participants were
+instructed to watch the narrative clip as many times as they needed, and
+narrate the content of each clip either to the experimenter or the
+recording digital
+camera.\footnote{Only 4 participants (2 native, 2 late) narrated to the experimenter whereas the remaining narrated to the camera.
+We did not make it obligatory for participants to  narrate to the digital camera or to the experimenter for few reasons.
 First, a native or native-like deaf experimenter was not always available, and a hearing late signer of TİD attended some sessions as the experimenter.
@@
 We coded the discourse status (Introduction, Maintenance,
-Re-introduction) and RE type (Nominal, Pronominal, Classifier, Null). We
-used ELAN Linguistic Annotation Software (Crasborn and Sloetjes, 2008)
-for data annotation. First Author who is a hearing signer of TİD
-annotated all the data and translated the narratives into Turkish
-together with a native deaf signer of TİD.
+Re-introduction) and RE type (Nominal, Pronominal, Classifier, Zero
+Anaphora). We used ELAN Linguistic Annotation Software (Crasborn and
+Sloetjes, 2008) for data annotation. First Author who is a hearing
+signer of TİD annotated all the data and translated the narratives into
+Turkish together with a native deaf signer of TİD.
 
@@
 RE type for a single reference with the main categories nominal,
-pronominal, null marking, and classifier. Under nominal, we identified
+pronominal, zero anaphora, and classifier. Under nominal, we identified
 bare nouns, modified nouns, and fingerspelled nouns. For classifiers, we
-distinguished among three types of constructions: WECL, BPCL, and ExtCL.
-Finally, we had CA, agreement, plain tags for null constructions. We
-were unable to observe any instance of HCLs occurring in isolation
-(i.e., not part of CA). The annotation tiers are presented in Table
-\ref{tab:tab-2}.
+distinguished among three types of handshapes: WECL, BPCL, and ExtCL.
+Finally, we had CA, agreement, plain tags for zero anaphora. We were
+unable to observe any instance of HCLs occurring in isolation (i.e., not
+part of CA). We should also highlight here that we coded referent status
+as nominal if the clause included an explicit nominal referent. For
+example, if constructed action, an index sign, or a classifier preceded
+or followed its explicitly signed nominal referent, the clause was
+regarded as having nominal reference (i.e., modified noun). The
+annotation tiers are presented in Table \ref{tab:tab-2}.
 
@@
 interested in the remaining coefficients which give information about
-predictor level comparisons.
+predictor level comparisons. Each coefficient is a comparison. That is
+why for a predictor with two levels A and B, the plots or the results
+only have ``A'' or ``B'' as labels.
 
 In a multinomial distribution, the responses are in a categorical
-relationship where the levels can be A,B,C, and more. These levels are
+relationship where the levels can be A, B, C, and more. These levels are
 not ordered in time, they are not cumulative, and they are not additive.
 If we were to make an analogy, they are like countries on Earth which
-can have differences and similarities, but they don't have to. In a
-multinomial distribution, one of the responses becomes the level of
-comparison and other levels are compared against it. In a way, there are
-multiple binomial distributions. However, the intercepts in a
-multinomial distribution end up having more of an interest for the
-researcher. This is because the intercept is a comparison of two levels
-of the response to the exclusion of the remaining response levels. There
-is another consideration that should go into interpreting the results of
-a multinomial distribution. Different from the intercepts, the remaining
-coefficients are predictor level comparisons within the grand comparison
-of the response levels. To indicate which predictor level comparison
-belongs to which response comparison, we name our coefficients with
-prefixes followed by an underscore.
+can have differences and similarities, but they do not have to, and they
+are treated as distinct entities. In a multinomial distribution, one of
+the responses becomes the level of comparison and other levels are
+compared against it. In a way, there are multiple binomial
+distributions. However, the intercepts in a multinomial distribution end
+up having more of an interest for the researcher. This is because the
+intercept is a comparison of two levels of the response to the exclusion
+of the remaining response levels. There is another consideration that
+should go into interpreting the results of a multinomial distribution.
+Different from the intercepts, the remaining coefficients are predictor
+level comparisons within the grand comparison of the response levels. To
+indicate which predictor level comparison belongs to which response
+comparison, we name our coefficients with prefixes followed by an
+underscore.
 
@@
 remaining distributions as indifference. In addition to these, we look
-at how the distribution skewed in the extremes or how `wide' it is from
-the median estimate to inform us mainly about uncertainty and variation.
+at how the distribution is skewed in the extremes or how `wide' it is
+from the median estimate to inform us mainly about uncertainty and
+variation.
 
@@
 
-We collected 4142 data points as a result of the
+We collected 4007 data points as a result of the
 experiment.\footnote{For the purposes of this double-blind review, we share all our data and code in an anonymous online repository that can be accessed here: \url{https://anonymous.4open.science/r/reference_tracking_revision-A542/}}
@@
 Maintenance, Re-introduction), its Type (Nominal, Pronominal, Classifer,
-Null),\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer, and Null).
+Zero
+Anaphora),\footnote{In our statistical analyses, we will not consider the subtypes of REs but rather group them by the four proposed main types (Nominal, Pronominal, Classifer, and Zero Anaphora).
 However, we allude to the numerical distribution of the subtypes in this and following sections.}
-and Participant Group (Native,
-Late).\footnote{Similarly, we treat late signers as a single group in our analyses and disregard the differing age of acquisition intervals among the late signers (0-3, 4-7, 8-12, 13-17).}
-
-Across our acquisition groups, we observed 1066 (0.26) instances of
-referent introduction, 1651 (0.4) instances of referent maintenance, and
-1425 (0.34) instances of referent re-introduction. For the introduction
+and Participant Group (DoD,
+DoH).\footnote{Similarly, we consider DoH signers a single group in our analyses and disregard the differing age of acquisition intervals (0-3, 4-7, 8-12, 13-17).}
+
+Across our acquisition groups, we observed 1031 (0.26) instances of
+referent introduction, 1589 (0.4) instances of referent maintenance, and
+1387 (0.35) instances of referent re-introduction. For the introduction
 of a referent, the majority of the REs were nominal constructions (Table
 \ref{tab:my-table}). However, we observed a sporadic use of classifiers
-and null marking respectively. When the discourse status was maintained,
-signers extensively used null markers to refer to the entities in the
-immediately previous clause. Classifier constructions followed null
-markers in frequency in maintained contexts. We observed only a few
-instances of nominal use for referent maintenance. In re-introduction
-contexts, signers used a greater proportion of nominal markings which
-were followed by null, classifier, and pronominal use.
+and zero anaphora respectively. When the discourse status was
+maintained, signers extensively used zero anaphora to refer to the
+entities in the immediately previous clause. Classifier constructions
+followed zero anaphora in frequency in maintained contexts. We observed
+only a few instances of nominal use for referent maintenance. In
+re-introduction contexts, signers used a greater proportion of nominal
+markings which were followed by zero anaphora, classifier, and
+pronominal use.
 
@@
 
-Among the null markers, CA was most frequently employed, and it was
-followed by plain verbs and agreement verbs (Table \ref{tab:my-table5}).
-Overall, signers used very few null markings for the introduction of a
-referent and the majority of the existing ones were CA. Additionally,
-signers used CA more for referent re-introduction than maintenance. We
-observed similar instances of agreement verbs for maintained and
-re-introduced contexts. In contrast, plain verbs were favored more in
-the maintenance setting than re-introduction.
+Among the markers of zero anaphora, CA was most frequently employed, and
+it was followed by plain verbs and agreement verbs (Table
+\ref{tab:my-table5}). Overall, signers used very few markers of zero
+anaphora for the introduction of a referent and the majority of the
+existing ones were CA. Additionally, signers used CA more for referent
+re-introduction than maintenance. We observed similar instances of
+agreement verbs for maintained and re-introduced contexts. In contrast,
+plain verbs were favored more in the maintenance setting than
+re-introduction.
 
@@
 Across the discourse statuses, there is also a difference between
-participant groups in the use of null marking for reference tracking.
-Native signers of TİD used it more (0.23) compared to the late signers
-(0.16). Comparably, late signers used nominals more frequently (0.25)
-than native signers (0.23). Figure 1 shows the numerical distribution of
-the responses by RE type and Group.
+participant groups in the use of zero anaphora for reference tracking.
+DoD signers of TİD used it more frequently (0.22) compared to the DoH
+signers (0.17). Comparably, DoH signers used nominals more frequently
+(0.26) than DoD signers (0.22). Figure 1 shows the numerical
+distribution of the responses by RE type and Group.
 
@@
 \caption{Referent counts by Discourse Status (Introduction, Maintanence,
-Re-introduction), RE Type (Nominal, Pronominal, Classifier, Null), and
-Participant Group (Native, Late)}
+Re-introduction), RE Type (Nominal, Pronominal, Classifier, Zero
+Anaphora), and Participant Group (DoD, DoH)}
 \end{figure}
@@
 predictors Discourse Status (Introduction, Maintenance, Re-introduction)
-and Participant Group (Native, Late). As the overall number of produced
+and Participant Group (DoD, DoH). As the overall number of produced
 pronouns was very little (40 data points), we excluded pronominal use
@@
 priors ourselves and used the preset uninformed priors. Figure
-2\footnote{We ran another model where we used the Nativeness predictor
-  with 3 levels 0-3:Native, 4-7:Late1, and 7-above:Late2. However, our
-  model did not converge properly and the credible intervals were very
-  wide spread, yielding the results uninformative. We provide the model
-  results as a plot in \ref{section:d}.} shows the posterior probability
-distributions where the dot is the median estimate, the thick line is
-the 50\% credible intervals, and the thin line is the 95\% credible
-intervals. There are 3 responses we are considering in this model. They
-are Classifier, Nominative, and Null. Classifier response is the level
-that other responses Nominative and Null are compared against.
+2\footnote{We ran another model where we used the self-reported
+  acquisition time as a predictor with 3 levels 0-3:Early, 4-7:Late1,
+  and 7-above:Late2. However, our model did not converge properly and
+  the credible intervals were very wide spread, yielding the results
+  uninformative. We provide the model results as a plot in
+  \ref{section:d}.} shows the posterior probability distributions where
+the dot is the median estimate, the thick line is the 50\% credible
+intervals, and the thin line is the 95\% credible intervals. There are 3
+responses we are considering in this model. They are Classifier,
+Nominal, and Zero Anaphora. Classifier response is the level that other
+responses Nominal and Zero Anaphora are compared against.
 
@@
 \includegraphics{revised_manuscript_files/figure-latex/model-gram-1.pdf}
-\caption{Regression model plot for the RE Type (Nominal, Null) with the
-predictors Discourse Status (Introduction, Maintenance, Re-introduction)
-and Participant Group (Native, Late)}
+\caption{Regression model plot for the RE Type (Nominal, Zero,
+Classifier) with the predictors Discourse Status (Introduction,
+Maintenance, Re-introduction) and Participant Group (DoD, DoH)}
 \end{figure}
@@
 participant group and discourse status. The second coefficient shows
-that null markings were also used more than classifiers by the
+that zero anaphora was also used more than classifiers by the
 participants in general, but not as much as nominals.
 
-In the comparison of nominals and classifiers, being a native acquirer
-of TİD did not have an effect. Nominals were used more when the
-discourse status was introduction compared to re-introduction. Again,
-the whole comparison is made against a classifier. In contrast to the
-introduction status, nominals were used less in maintenance contexts
+In the comparison of nominals and classifiers, acquiring TİD from deaf
+signing caregivers did not have an effect. Nominals were used more when
+the discourse status was introduction compared to re-introduction.
+Again, the whole comparison is made against a classifier. In contrast to
+the introduction status, nominals were used less in maintenance contexts
 compared to the re-introduction contexts within the greater comparison
 to classifier referents. There seems to be an interaction between the
-introduction status of discourse and being a native acquirer of TİD.
-When the acquirer is a native speaker and the discourse function is
-introduction, the chances of observing a nominal referent gets even
-higher. Also, there seems to be an interaction between maintaining a
-referent and being a native acquirer of TİD. When it was a native
-acquirer of TİD and the discourse context was maintenance, the chances
-of observing a nominal referent gets even lower.
-
-Compared to classifiers, being a native acquirer of TİD increased the
-odds of using null marking in general. Null constructions were used less
-in introduction contexts than in re-introduction. This contrasts with
-the nominals (4th coefficient) as they were used more in referent
+introduction status of discourse and being a DoD signer of TİD. When the
+acquirer is a DoD signer and the discourse function is introduction, the
+chances of observing a nominal referent gets
+higher.\footnote{Remember that this is an interaction term and the grand comparison is nominal vs classifier use.}
+Also, there seems to be an interaction between maintaining a referent
+and being a DoD signer of TİD. When it was a DoD signer of TİD and the
+discourse context was maintenance, the chances of observing a nominal
+referent gets
+lower.\footnote{This coefficient too is an interaction term and needs to be considered withing the comparison of nominal versus classifier use.}
+
+Compared to classifiers, being a DoD signer of TİD increased the odds of
+using zero anaphora in general. Zero anaphora constructions were used
+less in introduction contexts than in re-introduction. This contrasts
+with the nominals (4th coefficient) as they were used more in referent
 introduction versus re-introduction. In contrast to introduction
-contexts, null markers were used more in referent maintenance compared
+contexts, zero anaphora was used more in referent maintenance compared
 to re-introduction, yet again within the greater comparison to the
 classifiers. There seems to be an interaction between introducing a
-referent and being a native acquirer of TİD. In that, the odds of using
-null marking increases if the participant is a native acquirer of TİD
-and the referent is used in an introduction context. There is another
+referent and being a DoD signer of TİD. In that, the odds of using zero
+anaphora increases if the participant is a DoD signer of TİD and the
+referent is used in an introduction context. There is another
 interaction with regards to referent maintenance, contrasting the first
-interaction. In that, the odds of using null markers decrease if the
-participant is a native acquirer of TİD and the referent is maintained.
-This does not mean that late signers used more null markers in
-maintenance contexts. It means that the use of null markers did not
-linearly increase as the sum of using a null marker as a native plus
-using null markers for maintaining a referent versus re-introducing it.
+interaction. In that, the odds of using zero anaphora decrease if the
+participant is a DoD signer of TİD and the referent is maintained. This
+does not mean that DoH signers used more zero anaphora in maintenance
+contexts. It means that the use of zero anaphora did not linearly
+increase as the sum of using a zero anaphora marker as a DoD signer plus
+using zero anaphora for maintaining a referent versus re-introducing it.
 
@@
 In our study, we aimed to (i) describe how signers track referents in
-signed discourse in TİD and (ii) examine whether language deprivation
+signed discourse in TİD and (ii) examine whether late acquisition of TİD
 results in more over-redundant referent selection. To this end, we
 investigated the distribution of referring expressions (REs) that the
-native and late signers select when employing referents in different
+DoD and DoH signers select when employing referents in different
 discourse statuses (introduced, maintained, and re-introduced).
@@
 forms (mainly nominals) for referent introduction and less implicit
-forms such as null marking, pronominal use, and classifiers for
+forms such as zero anaphora, pronominals, and classifiers for
 maintenance. In re-introduced contexts, we predicted that signers would
@@
 For age of acquisition effects, based on what is proposed for child and
-adult deaf signers with language deprivation in terms of their
+adult deaf signers with delayed language exposure in terms of their
 morphosyntactic (Boudreault and Mayberry, 2006; Cormier et al., 2012;
-Kayabaşı et al., 2020) and narrative development (Becker, 2009; Cormier
-et al., 2013), we expected that late signers would be more
+Kayabaşı and Gökgöz, 2022) and narrative development (Becker, 2009;
+Cormier et al., 2013), we expected that DoH signers would be more
 over-redundant in their referent tracking, akin to second language
 learners (Bel et al., 2015; Frederiksen and Mayberry, 2019). That is, we
-had the expectation that late signers would use more explicit markers
-like nouns and fewer implicit markers like null marking even for highly
+had the expectation that DoH signers would use more explicit markers
+like nouns and fewer implicit markers like zero anaphora even for highly
 accessible referents (when the discourse status is maintained or even
-re-introduced in same cases) than native signers.
+re-introduced in same cases) than DoD signers.
 
@@
 
-The results of the present study partly support our hypotheses. In line
-with our expectations, regardless of their acquisition groups (native or
-late) (i) signers mainly used nominals to introduce a referent into the
-discourse, (ii) used null marking and classifiers but not pronouns for
+The results of the present study only partly support our hypotheses. In
+line with our expectations, regardless of the participant groups (DoD or
+DoH) (i) signers mainly used nominals to introduce a referent into the
+discourse, (ii) used zero anaphora and classifiers but not pronouns for
 maintained contexts, and (iii) employed both nominals and less explicit
-markers (classifiers, null markers and pronouns) to re-introduce a
+markers (classifiers, zero anaphora and pronouns) to re-introduce a
 referent.
@@
 characters and 3 inanimate objects. Although the number of animate
-characters and inanimate characters in Czubek (2017)'s stimuli matched
+characters and inanimate characters in Czubek's (2017) stimuli matched
 that of the present study, their narrative structure (a 6-episode-long
@@
 animate or inanimate referents for a longer period of time. As a result,
-our signers engaged in referent re-introduction more than what was
-reported in previous studies (Czubek, 2017; Frederiksen and Mayberry,
-2016).
+our participants might have engaged in referent re-introduction more
+than what was reported in previous studies (Czubek, 2017; Frederiksen
+and Mayberry, 2016).
 
@@
 observed that signers overwhelmingly preferred to use nominals over
-other REs for the first mention of a referent and used null markers such
-as CA, agreement and plain verb for referents with high accessibility
-(i.e., when the discourse status was maintained). This finding parallels
-other accounts of referential accessibility suggested for other sign
-languages (see Frederiksen and Mayberry (2016); Czubek (2017); and
-Swabey (2002); for ASL, Morgan (2005); for BSL Bel et al. (2015); and
-Barberà and Massó (2009) for LSC). Our results also indicate that
-modified nouns (e.g., nouns that are preceded or followed by a
+other REs for the first mention of a referent and used zero anaphora
+such as CA, agreement and plain verb for referents with high
+accessibility (i.e., when the discourse status was maintained). This
+finding parallels other accounts of referential accessibility suggested
+for other sign languages (see Frederiksen and Mayberry (2016); Czubek
+(2017); and Swabey (2002); for ASL, Morgan (2005); for BSL Bel et al.
+(2015); and Barberà and Massó (2009) for LSC). Our results also indicate
+that modified nouns (e.g., nouns that are preceded or followed by a
 classifier or IX) appeared more in introduced contexts compared to
-re-introduced contexts. This brings along the assumption that modified
-nouns are used more to track referents with lower accessibility,
-probably due to their more explicit structure.
+re-introduced contexts. This indicates that modified nouns are used more
+to track referents with lower accessibility, probably due to their more
+explicit structure.
 
@@
 and instead follows a topic shift and thus rank higher than classifiers
-and null markers (see Barberà and Massó, 2009 for LSC; and Nuhbalaoglu,
+and zero anaphora (see Barberà and Massó, 2009 for LSC; and Nuhbalaoglu,
 2018 for TİD). Considering this, the claim that the IX sign functions
@@
 tracking in ASL reported object preference for pronominals (Frederiksen
-and Mayberry, 2017). Nonetheless, we only coded subjects for maintained
-and re-introduced referents. We argue that all of these factors
+and Mayberry, 2017). However, we only coded subjects for maintained and
+re-introduced referents. We argue that all of these factors
 (genre-dependency as well as object preference) might contribute to the
-explanation of why overt pronouns in this data are dispreferred in favor
-of other markers for highly accessible referents such as null marking
+explanation of why overt pronouns in our data are dispreferred in favor
+of other markers for highly accessible referents such as zero anaphora
 and classifiers.
 
-A deeper look into the allocation of classifier constructions revealed
-that WECLs were the most common type of classifiers for the maintenance
-of a referent followed by BPCLs and ExtCLs. We observed a similar
-distibution for re-introduced contexts. In contrast, a great number of
-the overall ExtCLs were used to introduce a new referent into the
-discourse. That is to say, whereas WECLs were the preferred classifier
-types to signal highly accessible referents, ExtCL constructions were
-mainly used for referents which were not easily accessible. We suggest
-that the less iconic and transparent nature of WECLs allowed signers to
-continue a previous referent without being redundant. As pointed out by
-Czubek (2017), signers prefer to rely on ExtCLs for the first mention of
-an entity because these constructions are informative and explicit by
+A deeper look into the allocation of classifier handshapes revealed that
+WECLs were the most common type of classifiers for the maintenance of a
+referent followed by BPCLs and ExtCLs. We observed a similar distibution
+for re-introduced contexts. In contrast, a great number of the overall
+ExtCLs were used to introduce a new referent into the discourse. That is
+to say, whereas WECLs were the preferred classifier types to signal
+highly accessible referents, ExtCL constructions were mainly used for
+referents which were not easily accessible. We suggest that the less
+iconic and transparent nature of WECLs allowed signers to continue a
+previous referent without having to use an explict form. As pointed out
+by Czubek (2017), signers prefer to rely on ExtCLs for the first mention
+of an entity because these constructions are informative and explicit by
 nature, and they provide visual and descriptive cues about an entity and
 its location in space. We only observed 3 instances of ExtCLs for
-re-introduced contexts, replicating Frederiksen and Mayberry (2019)'s
+re-introduced contexts, replicating Frederiksen and Mayberry's (2019)
 findings for ASL.
 
-As for the distribution of the null markers, approximately 90\% of the
-plain verbs were used to maintain a previous referent, making these
-constructions least implicit REs used for TİD\footnote{As in other sign
-  languages, in TİD, the lexical meaning of signs has an effect on the
-  morphological grouping. For instance, usually (di)transitive agreement
-  verbs denote some sort of physical or metaphorical transfer of an
-  object or abstract notion (Meir, 1999). As for plain verbs, both
+As for the distribution of the zero anaphora, approximately 90\% of the
+plain verbs were used to maintain a previous referent and this makes
+these constructions least implicit REs used for TİD\footnote{As in other
+  sign languages, in TİD, the lexical meaning of signs has an effect on
+  the morphological grouping. For instance, usually (di)transitive
+  agreement verbs denote some sort of physical or metaphorical transfer
+  of an object or abstract notion (Meir, 1999). As for plain verbs, both
   phonology of a sign and its semantics have an effect on morphology.
@@
 respectively. This contrasts with some of the research (Czubek, 2017) in
-which agreement verbs ranked lowest on scale, indicating that these
+which agreement verbs ranked lowest on the scale, indicating that these
 constructions were more implicit compared to plain verbs. However, there
 seems to be a common ground for the position of CA markers which have
-been claimed to rank higher than other null markers as in the present
-study (Barberà and Massó, 2009; Czubek, 2017; Frederiksen and Mayberry,
-2016). In Figure \ref{fig:scale}, we report the distribution of REs in
-TİD as a function of referent accessibility and compare our
-accessibility hierarchy with that of (Frederiksen and Mayberry, 2016)
-for ASL. In the hierarchies, the leftmost REs (e.g., modified and bare
-nouns for TİD) are used when the accessibility is low. If the referent
-is already active in the addressee's mind and its accessibility is high,
-then the rightmost markers (e.g., agreement and plain verbs for TİD) are
-selected.
+been claimed to rank higher than other markers of zero anaphora as in
+the present study (Barberà and Massó, 2009; Czubek, 2017; Frederiksen
+and Mayberry, 2016). In Figure \ref{fig:scale}, we report the
+distribution of REs in TİD as a function of referent accessibility and
+compare our accessibility hierarchy with that of (Frederiksen and
+Mayberry, 2016) for ASL. In the hierarchies, the leftmost REs (e.g.,
+modified and bare nouns for TİD) are used when the accessibility is low.
+If the referent is already active in the addressee's mind and its
+accessibility is high, then the rightmost markers (e.g., agreement and
+plain verbs for TİD) are selected.
 
@@
 
-Both hierarchies illustrated in Figure \ref{fig:scale} consider nominals
+Both hierarchies illustrated in Figure \ref{fig:scale} treat nominals
 and ExtCLs as markers of lowly accessibly referents. Likewise, in both
 hierarchies, as markers of highly accessible referents, different types
-of null markers are further distinguished. Plain verbs rank the lowest
-on the scales, preceded by agreement verbs and CA. On a similar ground,
-both scales lack pronouns, which occurred very infrequently in the data.
-Although Frederiksen and Mayberry (2016) did not include other
-classifier types in the hierarchy, we proposed that WECLs and BPCLs must
-ranker higher than null markers but lower than ExtCLs that were mainly
+of zero anaphora markers are further distinguished. Plain verbs rank the
+lowest on the scales, preceded by agreement verbs and CA. On a similar
+ground, both scales lack pronouns, which occurred very infrequently in
+the data. Although Frederiksen and Mayberry (2016) did not include other
+classifier types in the hierarchy, we propose that WECLs and BPCLs must
+rank higher than zero anaphora but lower than ExtCLs that were mainly
 used by interlocutors to introduce a referent into the discourse
@@
 
-\hypertarget{age-of-acquisition-effects-on-referent-tracking}{%
-\subsection{Age of acquisition effects on referent
-tracking}\label{age-of-acquisition-effects-on-referent-tracking}}
+\hypertarget{effect-of-native-acquisition-on-referent-tracking}{%
+\subsection{Effect of native acquisition on referent
+tracking}\label{effect-of-native-acquisition-on-referent-tracking}}
 
 The results in this study only somewhat support our second set of
-hypotheses that we set for the effects of age of acquisition on
-participants' RE selection. That is, while being a native acquirer of
-TİD increased the odds of using null marking in reference tracking
-across different discourse statuses, we failed to find a difference for
-overall nominal production. Taking a closer look at the interaction
-terms, we observed that native signers, compared to late signers, were
-less likely to use nominal constructions when the discourse status was
-maintained. Native acquisition further increased the odds of null
-markers when the discourse status was introduced. However, we found the
-opposite alignment for the interaction of maintenance and native
-acquisition. That is, the coefficient for this interaction term slightly
-decreased the combined main effects of having a maintained context and
-being a native signer. This means that these main effects were greater
-than the interaction or added effects, pointing out to the implication
-that the type of discourse status (introduced, maintained,
+hypotheses that we set for the effects of having signing deaf parents on
+participants' RE selection. That is, while being a DoD signer of TİD
+increased the odds of using zero anaphora in reference tracking across
+different discourse statuses, we failed to find a difference for overall
+nominal production. Taking a closer look at the interaction terms, we
+observed that DoD signers, compared to DoH signers, were less likely to
+use nominal constructions when the discourse status was maintained.
+Native acquisition from deaf caregivers further increased the odds of
+zero anaphora when the discourse status was introduced. However, we
+found the opposite alignment for the interaction of maintenance and
+native acquisition. That is, the coefficient for this interaction term
+slightly decreased the combined main effects of having a maintained
+context and being a DoD signer. This means that these main effects were
+greater than the interaction or added effects, pointing out to the
+implication that the type of discourse status (introduced, maintained,
 re-introduced) accounts for the distribution of REs in TİD narratives
-better than age of acquisition (native, late) which, in comparison, had
-a less salient effect.
-
-While late signers were only partially over-redundant in maintained
+better than native acquisition from deaf caregivers (DoD, DoH) which, in
+comparison, had a less salient effect.
+
+While DoH signers were only partially over-redundant in maintained
 contexts, they were more so in introduced contexts. In line with our
@@
 late signing children in DGS (Becker, 2009) and late signing adults in
-BSL specifically for the use of CA(Cormier et al., 2013). As for TİD, to
-our knowledge, there is only one study (Gür, 2018) that directly aimed
-at examining age of acquisition effects on adult signers' narrative
-skills. Gür (2018) observed that late signing deaf adults and children
-set the scene less frequently and used CA less accurately compared to
-their native counterparts. However, this particular investigation did
-not present any data on the distribution of REs between two acquisition
-groups. In the present study, the finding that native signers preferred
-to use null arguments even for the first mention of a referent (19 in
-total) is surprising given some of the previous literature in the
-topicality and saliency framework (Cormier et al., 2013; Swabey, 2002).
-These research reported very little or no use of null marking by native
-signers for introduced contexts.
-
-Given the paucity of research directed at the late signing group, we now
-turn to the referent tracking strategies observed for second language
-learners of a sign language. Our study replicated some of the findings
-of the line of research that investigated second language learners of a
-sign language (Bel et al., 2015; Frederiksen and Mayberry, 2019). The
-authors reported that more instances of null marking in the context of
-referent introduction was attested for native signers than second
-language learners. Although using null marking to introduce a referent
-into the discourse would normally result in referential ambiguity, it is
-important to note that our stimuli consisted of 10 interconnected
-narratives with 2 or more recurring characters. It may be possible that
-native signers, having already introduced the referent with an explicit
-nominal marker in a different discourse context, found it felicitous to
-refer to the same entity with more implicit markers in the following
-narratives even for introduced contexts. Then, our results imply a
-difference in pragmatic sensitivity to referential accessibility between
-native and late signers, which is more robustly reflected in introduced
-contexts. To test this implication, one could use the same number and
-length of stimuli from different cartoons with different characters and
-see if the tendency to introduce a referent without an overt nominal for
-the first time is dependent on the task or not.
-
-Since we only observed a small difference between the two acquisition
-groups for the use of certain REs, the aforementioned findings allude to
-the implication that the effect on language deprivation on pragmatic
-factors should also be considered in different contexts. For one, both
-groups of signers can be tested with different genres of narratives and
-with a more complex story design. In order to capture possible variation
-in the accessibility levels of re-introduced referents, a more
-fine-tuned protocol (as in Toole, 1996) can be used for measuring
-referent distance and the number of other possible competitors. Taken
-altogether, our results accentuates the necessity of engaging in signed
-narrative production early in life as well as the importance of early
-intervention for late signing deaf children.
+BSL specifically for the use of CA (Cormier et al., 2013). As for TİD,
+to our knowledge, there is only one study (Gür, 2018) that directly
+aimed at examining age of acquisition effects on adult signers'
+narrative skills. Gür (2018) observed that late signing deaf adults and
+children set the scene less frequently and used CA less accurately
+compared to their native counterparts. However, this particular
+investigation did not present any data on the distribution of REs
+between two acquisition groups. In the present study, the finding that
+DoD signers preferred to use null arguments even for the first mention
+of a referent (19 in total) is surprising given some of the previous
+literature in the topicality and saliency framework (Cormier et al.,
+2013; Swabey, 2002). These research reported very little or no use of
+zero anaphora by DoD signers for introduced contexts.
+
+Given the paucity of research directed at the DoH or late signing group,
+we now turn to the referent tracking strategies observed for second
+language learners of a sign language. Our study replicated some of the
+findings of the line of research that investigated second language
+learners of a sign language (Bel et al., 2015; Frederiksen and Mayberry,
+2019). The authors reported that more instances of zero anaphora in the
+context of referent introduction was attested for DoD signers than
+second language learners. Although using zero anaphora to introduce a
+referent into the discourse would normally result in referential
+ambiguity, it is important to note that our stimuli consisted of 10
+interconnected narratives with 2 or more recurring characters. It may be
+possible that DoD signers, having already introduced the referent with
+an explicit nominal marker in a different discourse context, found it
+felicitous to refer to the same entity with more implicit markers in the
+following narratives even for introduced contexts. Then, our results
+imply a difference in pragmatic sensitivity to referential accessibility
+between DoD and DoH signers, which is more robustly reflected in
+introduced contexts. To test this implication, one could use the same
+number and length of stimuli from different cartoons with different
+characters and see if the tendency to introduce a referent without an
+overt nominal for the first time is dependent on the task or not.
+
+In other contexts, we found only very small differences between the two
+participant groups. This means that adult signers with late exposure to
+sign language in our study were still able to achieve native-like
+competency in terms of reference tracking in simple narratives.
+Nevertheless, DoD and DoH signers still exhibit some differences
+especially in terms of pragmatic sensitivity to economy of form despite
+using TİD as their preferred communication language for similar numbers
+of years. Since we only observed a small difference between the two
+acquisition groups for the use of certain REs, the aforementioned
+findings allude to the implication that the effect on native acquisition
+on pragmatic factors should also be considered in different contexts.
+For one, both groups of signers can be tested with different genres of
+narratives and with a more complex story design. In order to capture
+possible variation in the accessibility levels of re-introduced
+referents, a more fine-tuned protocol (as in Toole, 1996) can be used
+for measuring referent distance and the number of other possible
+competitors. Furthermore, our focus in this study was to look at
+referent tracking strategies between DoD and DoH signers. However, for
+future research, it might also be useful to look at pre and post-lingual
+acquisiton of TİD. We only had 3 post-lingual signers of
+TİD.\footnote{When we looked at the random effects by participants, they did not differ from their acquisition group (DoD or DoH).}
+
+\clearpage
 
@@
 found that the distribution of REs (nominal, pronominal, classifier, and
-null marking) followed from previous accounts of saliency and
+zero anaphora) followed from previous accounts of saliency and
 referential accessibility. For introduced contexts, signers produced
@@
 accessible referents. We observed only very few instances of pronouns
-(mostly produced by late signers), unlike our expectations. The ones we
+(mostly produced by DoH signers), unlike our expectations. The ones we
 observed were mostly reserved for re-introduced contexts along with
 nominals. Based on our findings across the two groups, we proposed a
-hierarchy of referent accessibility for TİD. As for the age of
-acquisition effects, we observed limited over-redundancy for signers
-with delayed language acquisition (i.e., late signers) who employed
-fewer implicit forms of reference tracking compared to native signers.
-Native signers were less likely to use a nominal for maintained contexts
-and more likely to use null marking for introduced contexts than late
-signers. Following Swabey (2002)'s intituition, we conclude that native
-signers must adhere to pragmatic norms and leverage accessibility better
-than late signers only to a certain extent (given our finding of limited
-overexplicitness for late signers) since they use less form when the
-referent is more active in the mind.
+hierarchy of referent accessibility for TİD. As for the effects of
+native acquisition, we observed limited over-redundancy for signers with
+delayed language acquisition (i.e., DoH signers) who employed fewer
+implicit forms of reference tracking compared to DoD signers. DoD
+signers were less likely to use a nominal for maintained contexts and
+more likely to use zero anaphora for introduced contexts than DoH
+signers.
+
+We conclude that DoH signers have overall native-like competency in
+simple narratives, however, following Swabey (2002)'s intuition, our
+findings also suggest that DoD signers adhere to pragmatic norms and
+leverage accessibility better than DoH signers to a certain extent
+(given our finding of limited overexplicitness for DoH signers) since
+they use less form when the referent is more active in the mind. To
+avoid this possible discrepancy between the groups, our results
+accentuates the necessity of engaging in signed narrative production
+early in life as well as the importance of early intervention for DoH or
+late signing deaf children. This is especially a crucial message for
+policymakers, educators, and health professionals who work with deaf and
+hard-of-hearing children. Providing early and effective access to sign
+language from very early on can help bridge the gap and ensure equal
+opportunities for language development and success in life for DoH
+signers.
 
@@
 \midrule
-2 & Native & M & 0-3 & 33 & 12 & PS, MS, HS & TİD, home sign\\
-3 & Native & F & 0-3 & 27 & 12 & PS, MS, HS & TİD, home sign\\
-13 & Native & F & 0-3 & 27 & 12 & PS, MS, HS & Home sign\\
-14 & Native & M & 0-3 & 29 & 12 & PS, MS, HS & TİD, home sign\\
-18 & Native & M & 0-3 & 28 & 12 & PS, MS & TİD\\
-20 & Native & M & 0-3 & 32 & 12 & PS, MS & TİD\\
-22 & Native & M & 0-3 & 35 & 12 & PS, MS, HS & TİD\\
-24 & Native & F & 0-3 & 23 & 12 & PS, MS, HS & TİD\\
-26 & Native & F & 0-3 & 30 & 12 & PS, MS, HS & TİD\\
-27 & Native & F & 0-3 & 25 & 12 & PS, MS & TİD\\
-29 & Native & F & 0-3 & 24 & 14 & PS, MS, HS & TİD\\
-30 & Native & M & 0-3 & 18 & 12 & PS, MS & TİD\\
-33 & Native & F & 0-3 & 21 & 16 & PS, MS, HS & TİD\\
-34 & Native & M & 0-3 & 19 & 16 & PS, MS, HS & TİD\\
-38 & Native & F & 0-3 & 25 & 12 & PS, MS & TİD\\
-10 & Late & F & 4-7 & 35 & 14 & PS, MS & TİD, home sign, TR\\
-23 & Late & M & 4-7 & 36 & 12 & MS, HS & TR, home sign\\
-4 & Late & M & 4-7 & 33 & 12 & PS, HS & TR\\
-16 & Late & F & 4-7 & 33 & 12 & PS & TR, home sign\\
-21 & Late & F & 4-7 & 31 & 14 & PS, MS & Home sign\\
-25 & Late & M & 4-7 & 35 & 12 & PS, MS, HS & Home sign\\
-35 & Late & F & 8-12 & 49 & 8 & PS, MS & Home sign, TR\\
-36 & Late & M & 8-12 & 43 & 16 & PS, MS & TR\\
-12 & Late & M & 8-12 & 28 & 16 & MS & TR\\
-15 & Late & F & 8-12 & 24 & 12 & PS, MS, HS & TR, TİD\\
-19 & Late & F & 8-12 & 24 & 8 & PS, MS & Home sign\\
-28 & Late & M & 8-12 & 33 & 12 & PS, MS & Home sign\\
-31 & Late & M & 13-17 & 50 & 8 & MS & TR, home sign\\
-32 & Late & M & 13-17 & 31 & 16 & PS, MS, HS & TR\\
+2 & DoD & M & 0-3 & 33 & 12 & PS, MS, HS & TİD, home sign\\
+3 & DoD & F & 0-3 & 27 & 12 & PS, MS, HS & TİD, home sign\\
+*13 & DoD & F & 0-3 & 27 & 12 & PS, MS, HS & Home sign\\
+14 & DoD & M & 0-3 & 29 & 12 & PS, MS, HS & TİD, home sign\\
+18 & DoD & M & 0-3 & 28 & 12 & PS, MS & TİD\\
+20 & DoD & M & 0-3 & 32 & 12 & PS, MS & TİD\\
+22 & DoD & M & 0-3 & 35 & 12 & PS, MS, HS & TİD\\
+24 & DoD & F & 0-3 & 23 & 12 & PS, MS, HS & TİD\\
+26 & DoD & F & 0-3 & 30 & 12 & PS, MS, HS & TİD\\
+27 & DoD & F & 0-3 & 25 & 12 & PS, MS & TİD\\
+29 & DoD & F & 0-3 & 24 & 14 & PS, MS, HS & TİD\\
+30 & DoD & M & 0-3 & 18 & 12 & PS, MS & TİD\\
+33 & DoD & F & 0-3 & 21 & 16 & PS, MS, HS & TİD\\
+34 & DoD & M & 0-3 & 19 & 16 & PS, MS, HS & TİD\\
+38 & DoD & F & 0-3 & 25 & 12 & PS, MS & TİD\\
+*10 & DoH & F & 4-7 & 35 & 14 & PS, MS & TİD, home sign, TR\\
+23 & DoH & M & 4-7 & 36 & 12 & MS, HS & TR, home sign\\
+4 & DoH & M & 4-7 & 33 & 12 & PS, HS & TR\\
+16 & DoH & F & 4-7 & 33 & 12 & PS & TR, home sign\\
+21 & DoH & F & 4-7 & 31 & 14 & PS, MS & Home sign\\
+25 & DoH & M & 4-7 & 35 & 12 & PS, MS, HS & Home sign\\
+35 & DoH & F & 8-12 & 49 & 8 & PS, MS & Home sign, TR\\
+36 & DoH & M & 8-12 & 43 & 16 & PS, MS & TR\\
+12 & DoH & M & 8-12 & 28 & 16 & MS & TR\\
+15 & DoH & F & 8-12 & 24 & 12 & PS, MS, HS & TR\\
+19 & DoH & F & 8-12 & 24 & 8 & PS, MS & Home sign\\
+28 & DoH & M & 8-12 & 33 & 12 & PS, MS & Home sign\\
+31 & DoH & M & 13-17 & 50 & 8 & MS & TR, home sign\\
+32 & DoH & M & 13-17 & 31 & 16 & PS, MS, HS & TR\\
 \bottomrule
@@
 \begin{tablenotes}[para]
-\normalsize{\textit{Note.} AOA = age of acquisition, LOE = level of education, PS = primary school, MS = middle school, HS = high school, TR = (Spoken) Turkish.}
+\normalsize{\textit{Note.} AOA = age of acquisition, LOE = level of education, PS = primary school, MS = middle school, HS = high school, TR = (Spoken) Turkish.
+*We excluded Participant 13 from our data analysis because the signer indicated to have deaf parents, thus grouped as DoD, but also reported only home sign as the home language and not TİD.
+Since our classification of DoD included the criterion of having deaf and TİD-signing parents, we decided to remove this participant's data.
+Also, Participant 10 reported to use TİD at home despite having hearing caregivers which may go against our classification criterion for being DoH that is to have non-signing hearing caregivers.
+However, Participant 10 informed us that the parents used very rudimentary sign language and that they often did not understand what he was signing.}
 \end{tablenotes}
@@
 \midrule
-Nom\_Intercept & 1.23 & 0.20 & 0.83 & 1.62 & 1.00\\
-Null\_Intercept & -4.25 & 3.56 & -13.27 & 0.09 & 1.01\\
-Nom\_Late1vsNative & 0.49 & 0.54 & -0.55 & 1.56 & 1.01\\
-Nom\_Late2vsLate1 & -0.41 & 0.49 & -1.37 & 0.57 & 1.00\\
-Nom\_Intro & 3.76 & 0.25 & 3.29 & 4.25 & 1.00\\
-Nom\_Maintain & -5.24 & 0.21 & -5.66 & -4.82 & 1.00\\
-Nom\_L1vsN*Intro & 1.01 & 0.63 & -0.16 & 2.30 & 1.00\\
-Nom\_L2vsL1*Intro & 0.02 & 0.64 & -1.24 & 1.24 & 1.00\\
-Nom\_L1vsN*Maintain & -0.33 & 0.55 & -1.43 & 0.73 & 1.00\\
-Nom\_L2vsL1*Maintain & -0.92 & 0.52 & -1.96 & 0.11 & 1.00\\
-Null\_Late1vsNative & -13.40 & 10.69 & -40.57 & -0.41 & 1.01\\
-Null\_Late2vsLate1 & 14.22 & 10.69 & 1.29 & 41.29 & 1.01\\
-Null\_Intro & -21.98 & 14.25 & -58.10 & -4.70 & 1.01\\
-Null\_Maintain & 11.27 & 7.13 & 2.67 & 29.30 & 1.01\\
-Null\_L1vsN*Intro & -54.37 & 42.74 & -162.14 & -2.68 & 1.01\\
-Null\_L2vsL1*Intro & 55.46 & 42.75 & 3.99 & 164.16 & 1.01\\
-Null\_L1vsN*Maintain & 26.68 & 21.38 & 0.88 & 80.69 & 1.01\\
-Null\_L2vsL1*Maintain & -27.59 & 21.38 & -81.73 & -1.81 & 1.01\\
+Nom\_Intercept & 1.24 & 0.21 & 0.83 & 1.68 & 1.00\\
+Null\_Intercept & -4.77 & 4.19 & -16.09 & 0.09 & 1.01\\
+Nom\_Late1vsNative & 0.43 & 0.57 & -0.71 & 1.56 & 1.00\\
+Nom\_Late2vsLate1 & -0.34 & 0.51 & -1.32 & 0.68 & 1.01\\
+Nom\_Intro & 3.76 & 0.26 & 3.27 & 4.27 & 1.00\\
+Nom\_Maintain & -5.24 & 0.22 & -5.67 & -4.80 & 1.00\\
+Nom\_L1vsN*Intro & 1.05 & 0.67 & -0.23 & 2.41 & 1.00\\
+Nom\_L2vsL1*Intro & -0.07 & 0.67 & -1.41 & 1.22 & 1.00\\
+Nom\_L1vsN*Maintain & -0.35 & 0.57 & -1.49 & 0.75 & 1.00\\
+Nom\_L2vsL1*Maintain & -0.87 & 0.54 & -1.89 & 0.20 & 1.00\\
+Null\_Late1vsNative & -15.00 & 12.57 & -48.88 & -0.48 & 1.01\\
+Null\_Late2vsLate1 & 15.88 & 12.57 & 1.30 & 49.79 & 1.01\\
+Null\_Intro & -24.10 & 16.74 & -69.41 & -4.68 & 1.01\\
+Null\_Maintain & 12.30 & 8.37 & 2.60 & 35.02 & 1.01\\
+Null\_L1vsN*Intro & -60.77 & 50.28 & -195.99 & -2.68 & 1.01\\
+Null\_L2vsL1*Intro & 61.91 & 50.26 & 3.62 & 197.45 & 1.01\\
+Null\_L1vsN*Maintain & 29.89 & 25.14 & 0.81 & 97.46 & 1.01\\
+Null\_L2vsL1*Maintain & -30.89 & 25.13 & -98.46 & -1.73 & 1.01\\
 \bottomrule
@@
 1964.
-
-\leavevmode\vadjust pre{\hypertarget{ref-aikhenvald2017}{}}%
-Aikhenvald, Alexandra, 2017. A typology of noun categorization devices.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-dudis2004}{}}%
-Dudis, Paul G., 2004.
-\href{https://www.degruyter.com/view/j/cogl.2004.15.issue-2/cogl.2004.009/cogl.2004.009.xml}{Body
-partitioning and real-space blends}. Cognitive Linguistics 15.
-
 \leavevmode\vadjust pre{\hypertarget{ref-emmorey1996}{}}%
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-kayabasi2020}{}}%
-Kayabaşı, Demet, Sevgi, Hande, Özsoy, A. Sumru, 2020. Null arguments in
-turkish sign language. Morphological Complexity within and across
-Boundaries: In honour of Asli Göksel 215.
+\leavevmode\vadjust pre{\hypertarget{ref-kayaba2022}{}}%
+Kayabaşı, Demet, Gökgöz, Kadir, 2022. The causative-inchoative
+alternation and age-of-acquisition effects on multi-predicate
+constructions in turkish sign language. Language Learning and
+Development 0, 1--26.
+doi:\href{https://doi.org/10.1080/15475441.2022.2047690}{10.1080/15475441.2022.2047690}
 
@@
 phonology and morphology (PhD thesis). Ankara.
-
-\leavevmode\vadjust pre{\hypertarget{ref-lillo-martin1990}{}}%
-Lillo-Martin, Diane, Klima, Edward S., 1990. Pointing out differences:
-ASL pronouns in syntactic theory. Theoretical issues in sign language
-research, Volume 1: Linguistics 191.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-sari_analysis_2005}{}}%
+Sari, Hakan, 2005. An analysis of the relationship between identity
+patterns of turkish deaf adolescents and the communication modes used in
+special residential schools for the hearing impaired and deaf. Deafness
+\& Education International 7, 206--222.
+doi:\href{https://doi.org/10.1179/146431505790560329}{10.1179/146431505790560329}
+
 \leavevmode\vadjust pre{\hypertarget{ref-sevinc2006}{}}%
@@
 doi:\href{https://doi.org/10.1353/sls.2014.0008}{10.1353/sls.2014.0008}
-
-\leavevmode\vadjust pre{\hypertarget{ref-sorace2011}{}}%
-Sorace, Antonella, 2011. Pinning down the concept of {``}interface{''}
-in bilingualism. Linguistic approaches to bilingualism 1, 133.
-
-\leavevmode\vadjust pre{\hypertarget{ref-sorace2009}{}}%
-Sorace, Antonella, Serratrice, Ludovica, 2009.
-\href{https://doi.org/10.1177/1367006909339810}{Internal and external
-interfaces in bilingual language development: Beyond structural
-overlap}. International Journal of Bilingualism 13, 195--210.
-
-\leavevmode\vadjust pre{\hypertarget{ref-supalla1986}{}}%
-Supalla, Ted, 1986. The classifier system in american sign language.
-Noun classes and categorization 7, 181214.
 

### 2023-01-31 revised_manuscript.Rmd
update

@@
 However, this particular observation is not peculiar to spoken languages.
-There is an emerging body of literature that reports the distribution of REs in signed narratives [@czubek2017; @swabey2002; @frederiksen2016; @perniss2015; @morgan2005; @hodge2019; @ahlgren1994].
+There is an emerging body of literature that reports the distribution of REs in signed narratives [@czubek2017; @ferrara2022; @swabey2002; @frederiksen2016; @perniss2015; @morgan2005; @hodge2019; @ahlgren1994].
 Building on the work by @gundel1993, @swabey2002 demonstrated that, like other languages such as Russian, Chinese, and Japanese, American Sign Language (ASL) signers use bare nouns for introduced and re-introduced discourse contexts whereas zero anaphora is used for referents which are "in-focus" or very accessible to the addressee.
@@
 It is important to note, however, that this research either collapsed DoD and DoH signers into a single group and labeled them as "native-like" or explicitly compared native deaf signers with late deaf signers (with delayed exposure to a sign language) in terms of narrative competence.
+Most recently, @ferrara2022 examined corpora including video-recorded narratives from five different sign languages including Irish Sign Language, Finnish Sign Language, Norwegian Sign Language and Swedish Sign Language and investigated how these languages established and maintained referents.
+Like other observations, this research showed that, across each of these languages, bare nouns were used to introduce new referents and zero anaphora or more reduced forms were preferred for referent maintenance.
 
@@
 {} \includegraphics[width=40pt]{pictures/p10.png} \\ \textbf{Gloss:} \textsc{mouse.stay.like.this} \\ 
-\textbf{Status:} Maintained. \\
+\textbf{Status:} Maintain. \\
 \glt \textbf{Context:} Narration 6, Event 9: The signer re-introduces Jerry in the previous clause and maintains reference to him by using this constructed action.
@@
 {} \includegraphics[width=80pt]{pictures/p11.png} \\ \textbf{Gloss:} \textsc{run.away} \\ 
-\textbf{Status:} Re-introduced. \\
+\textbf{Status:} Re-introduce. \\
 \glt \textbf{Context:} Narration 5, Event 4: In this sentence, the signer re-introduces Jerry with the plain verb \textsc{run.away}.
@@
 {} \includegraphics[width=80pt]{pictures/p12.png} \\ \textbf{Gloss:} \textsc{3a(mouse)-hit-3b(cat)} \\ 
-\textbf{Status:} Maintained. \\
+\textbf{Status:} Maintain. \\
 \glt \textbf{Context:} Narration 8, Event 2: The signer introduces Jerry in the preceding clause and maintains reference to him by using the agreement verb \textsc{hit}.
@@
 
-For age of acquisition effects, based on what is proposed for child and adult deaf signers with delayed language exposure in terms of their morphosyntactic [@boudreault2006; @cormier2012; @kayaba2022] and narrative development [@becker2009; @cormier2013], we expected that DoH signers would be more over-redundant in their referent tracking, akin to second language learners [@frederiksen2019; @bel2015].
+For age of acquisition effects, based on what is proposed for child and adult deaf signers with delayed language exposure in terms of their morphosyntactic [@boudreault2006; @cormier2012; @kayaba2022] and narrative development [@becker2009; @cormier2013], we expected that DoH signers would be more over-redundant in their referent tracking, perhaps akin to second language learners [@frederiksen2019; @bel2015].
 That is, we had the expectation that DoH signers would use more explicit markers like nouns and fewer implicit markers like zero anaphora even for highly accessible referents (when the discourse status is maintained or even re-introduced in same cases) than DoD signers.
@@
 As for the distribution of REs across different discourses statuses, we observed that signers overwhelmingly preferred to use nominals over other REs for the first mention of a referent and used zero anaphora such as CA, agreement and plain verb for referents with high accessibility (i.e., when the discourse status was maintained).
-This finding parallels other accounts of referential accessibility suggested for other sign languages (see @frederiksen2016; @czubek2017; and @swabey2002; for ASL, @morgan2005; for BSL @bel2015; and @barbera2009 for LSC).
+This finding parallels other accounts of referential accessibility suggested for other sign languages (see @frederiksen2016; @czubek2017; and @swabey2002; for ASL, @morgan2005; for BSL @bel2015; @barbera2009 for LSC; @ferrara2022 for other sign languages).
 Our results also indicate that modified nouns (e.g., nouns that are preceded or followed by a classifier or IX) appeared more in introduced contexts compared to re-introduced contexts.

### 2023-01-31 revised_manuscript.tex
update

@@
 distribution of REs in signed narratives (Ahlgren and Bergman, 1994;
-Czubek, 2017; Frederiksen and Mayberry, 2016; Hodge et al., 2019;
-Morgan, 2005; Perniss and Özyürek, 2015; Swabey, 2002). Building on the
-work by Gundel et al. (1993), Swabey (2002) demonstrated that, like
-other languages such as Russian, Chinese, and Japanese, American Sign
-Language (ASL) signers use bare nouns for introduced and re-introduced
-discourse contexts whereas zero anaphora is used for referents which are
-``in-focus'' or very accessible to the addressee. Overt REs like
-pronouns or nouns are disprefered for such referents. Swabey (2002)
-concluded that ``native'' signers in her study abided by the Gricean
-maxim of quantity and thus used more economic forms of reference
-tracking. Similar findings have been reported for German Sign Language
-(Deutsche Gebärdensprache - DGS) (Perniss and Özyürek, 2015) and British
-Sign Language (BSL) (Morgan, 2005). It is important to note, however,
-that this research either collapsed DoD and DoH signers into a single
-group and labeled them as ``native-like'' or explicitly compared native
-deaf signers with late deaf signers (with delayed exposure to a sign
-language) in terms of narrative competence.
+Czubek, 2017; Ferrara et al., 2022; Frederiksen and Mayberry, 2016;
+Hodge et al., 2019; Morgan, 2005; Perniss and Özyürek, 2015; Swabey,
+2002). Building on the work by Gundel et al. (1993), Swabey (2002)
+demonstrated that, like other languages such as Russian, Chinese, and
+Japanese, American Sign Language (ASL) signers use bare nouns for
+introduced and re-introduced discourse contexts whereas zero anaphora is
+used for referents which are ``in-focus'' or very accessible to the
+addressee. Overt REs like pronouns or nouns are disprefered for such
+referents. Swabey (2002) concluded that ``native'' signers in her study
+abided by the Gricean maxim of quantity and thus used more economic
+forms of reference tracking. Similar findings have been reported for
+German Sign Language (Deutsche Gebärdensprache - DGS) (Perniss and
+Özyürek, 2015) and British Sign Language (BSL) (Morgan, 2005). It is
+important to note, however, that this research either collapsed DoD and
+DoH signers into a single group and labeled them as ``native-like'' or
+explicitly compared native deaf signers with late deaf signers (with
+delayed exposure to a sign language) in terms of narrative competence.
+Most recently, Ferrara et al. (2022) examined corpora including
+video-recorded narratives from five different sign languages including
+Irish Sign Language, Finnish Sign Language, Norwegian Sign Language and
+Swedish Sign Language and investigated how these languages established
+and maintained referents. Like other observations, this research showed
+that, across each of these languages, bare nouns were used to introduce
+new referents and zero anaphora or more reduced forms were preferred for
+referent maintenance.
 
@@
 Cormier et al., 2013), we expected that DoH signers would be more
-over-redundant in their referent tracking, akin to second language
-learners (Bel et al., 2015; Frederiksen and Mayberry, 2019). That is, we
-had the expectation that DoH signers would use more explicit markers
-like nouns and fewer implicit markers like zero anaphora even for highly
-accessible referents (when the discourse status is maintained or even
-re-introduced in same cases) than DoD signers.
+over-redundant in their referent tracking, perhaps akin to second
+language learners (Bel et al., 2015; Frederiksen and Mayberry, 2019).
+That is, we had the expectation that DoH signers would use more explicit
+markers like nouns and fewer implicit markers like zero anaphora even
+for highly accessible referents (when the discourse status is maintained
+or even re-introduced in same cases) than DoD signers.
 
@@
 (2017); and Swabey (2002); for ASL, Morgan (2005); for BSL Bel et al.
-(2015); and Barberà and Massó (2009) for LSC). Our results also indicate
-that modified nouns (e.g., nouns that are preceded or followed by a
-classifier or IX) appeared more in introduced contexts compared to
-re-introduced contexts. This indicates that modified nouns are used more
-to track referents with lower accessibility, probably due to their more
-explicit structure.
+(2015); Barberà and Massó (2009) for LSC; Ferrara et al. (2022) for
+other sign languages). Our results also indicate that modified nouns
+(e.g., nouns that are preceded or followed by a classifier or IX)
+appeared more in introduced contexts compared to re-introduced contexts.
+This indicates that modified nouns are used more to track referents with
+lower accessibility, probably due to their more explicit structure.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-ferrara2022}{}}%
+Ferrara, Lindsay, Anible, Benjamin, Hodge, Gabrielle, Jantunen, Tommi,
+Leeson, Lorraine, Mesch, Johanna, Nilsson, Anna-Lena, 2022. A
+cross-linguistic comparison of reference across five signed languages.
+Linguistic Typology.
+
 \leavevmode\vadjust pre{\hypertarget{ref-frederiksen2016}{}}%

### 2023-01-31 revised_manuscript.Rmd
update

@@
           caption ="Annotation Tiers with their Tags and Descriptions",
-          note = "NOM = nominal, PRO = pronominal, CL = classifier, NULL = null anaphora.",
+          note = "NOM = nominal, PRO = pronominal, CL = classifier, ZERO = zero anaphora.",
           escape = FALSE,
@@
 
-\clearpage
-
 # Conclusion

### 2023-01-31 revised_manuscript.tex
update

@@
 RE Type & NOM: Bare Noun & A non-modified noun phrase (e.g., CAT, MOUSE)\\
- & NOM: Modified Noun & A noun phrase modified by an index sign or an adjective\\
+ & NOM: Modified Noun & A noun phrase modified by an index sign, classifier etc.\\
  & NOM: Fingerspelled Noun & Borrowed nouns from Turkish spelled by the manual alphabet (e.g., K-E-D-İ ‘cat’)\\
@@
  & CL: Extension Classifier & A handshape that denotes the size and shape of an entity\\
- & NULL: Constructed Action & A multifunctional construction where the signers take on the role of a referent and perform their actions\\
- & NULL: Plain Verb & A non-agreeing predicate (e.g., LOVE)\\
- & NULL: Agreement Verb & A predicate the beginning and ending of which agree with the person and number\\
+ & ZERO: Constructed Action & A multifunctional construction where the signers take on the role of a referent and perform their actions\\
+ & ZERO: Plain Verb & A non-agreeing predicate (e.g., LOVE)\\
+ & ZERO: Agreement Verb & A predicate the beginning and ending of which agree with the person and number\\
 \bottomrule
@@
 \begin{tablenotes}[para]
-\normalsize{\textit{Note.} NOM = nominal, PRO = pronominal, CL = classifier, NULL = null anaphora.}
+\normalsize{\textit{Note.} NOM = nominal, PRO = pronominal, CL = classifier, ZERO = zero anaphora.}
 \end{tablenotes}
@@
 TİD.\footnote{When we looked at the random effects by participants, they did not differ from their acquisition group (DoD or DoH).}
-
-\clearpage
 

### 2023-05-01 revised_manuscript.Rmd
updated

@@
 However, this particular observation is not peculiar to spoken languages.
-There is an emerging body of literature that reports the distribution of REs in signed narratives [@czubek2017; @ferrara2022; @swabey2002; @frederiksen2016; @perniss2015; @morgan2005; @hodge2019; @ahlgren1994].
+There is an emerging body of literature that reports the distribution of REs in signed narratives [@czubek2017; @ferrara2022; @swabey2002; @frederiksen2016; @perniss2015; @morgan2005; @hodge2019; @ahlgren1994; @garcia_reference_2014; @garcia2018impersonal].
 Building on the work by @gundel1993, @swabey2002 demonstrated that, like other languages such as Russian, Chinese, and Japanese, American Sign Language (ASL) signers use bare nouns for introduced and re-introduced discourse contexts whereas zero anaphora is used for referents which are "in-focus" or very accessible to the addressee.
@@
 In TİD and many other sign languages, it is possible to use bare nouns as in \ref{ex:2a} as well as fingerspelled nouns as in \ref{ex:2b}.
-Fingerspelling refers to the use of the manual aphabet to spell all or part of the letters of a word borrowed from a surrounding written language [@kubus2008].
+Fingerspelling refers to the use of the manual alphabet to spell all or part of the letters of a word borrowed from a surrounding written language [@kubus2008].
 Nouns can also be modified by a prenominal or postnominal extended index finger pointing sign, glossed as IX, functioning as a determiner or a demonstrative as in \ref{ex:2c} [@maclaughlin1997] or by a modifying classifier within the same NP as in \ref{ex:2d} [@frederiksen2016].\footnote{Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions (Lillo-Martin and Klima, 1990).
@@
 In order to capture in-group variation among DoH signers, we run an additional supplementary analysis in which we observed very wide credible intervals due to the small group size, rendering the results uninformative.
-For the analysis, see \ref{section:d}.} All participants indicated that their preferred language of communication was TİD which they used for most contexts.
-Detailed information about the linguistic and educational background of the participants can be found in \ref{section:a}.
+For the analysis, see \ref{section:d}.} Detailed information about the linguistic and educational background of the participants can be found in \ref{section:a}.
+
+All participants indicated that their preferred language of communication was TİD which they used for most contexts.
 1 DoD signer was excluded from the data analysis because they self-reported to have signing deaf parents but also only used home sign (a more crude form of communication) at home.
@@
 For classifiers, we distinguished among three types of handshapes: WECL, BPCL, and ExtCL.
-Finally, we had CA, agreement, plain tags for zero anaphora.
+Finally, we had CA[^1], agreement, plain tags for zero anaphora.
 We were unable to observe any instance of HCLs occurring in isolation (i.e., not part of CA).
@@
 The annotation tiers are presented in Table \ref{tab:tab-2}.
+
+[^1]: Note that CA can also be further subdivided into overt, reduced, and subtle CA, following [@cormier_rethinking_2015].
+In overt CA, the signer is "fully in character" and does not use any other lexical or classifier constructions.
+In reduced CA, the signer can involve a small amount of narration or lexical material while tracking a referent using CA.
+However, subtle CA refers to cases where the referent is primarily tracked by a non-CA construction (e.g., nominal, pronominal, verbal or classifier constructions) but the signer subtly uses a few CA articulators at the same time (e.g., enacting the eye-gaze of the character).
+We only coded overt and reduced types as reference tracking by CA, and disregarded the subtle type.
+We'd also like to acknowledge that mixed CA, or simultaneous tracking of two or multiple references through CA and other tools can be possible [@dudis2004; @liddell2003].
+In such cases, as we explain here, we only coded only the subject of a clause for maintenance and re-introduction.
+If the referent was in subject position and tracked by using CA, we considered this an instance of CA, as well.
 
@@
 ```
-```{r fig:count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), RE Type (Nominal, Pronominal, Classifier, Zero Anaphora), and Participant Group (DoD, DoH)"}
+```{r fig-count-plot, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE, fig.width=4.75, fig.height=3.5, fig.pos="hbt!",fig.cap="Referent counts by Discourse Status (Introduction, Maintanence, Re-introduction), RE Type (Nominal, Pronominal, Classifier, Zero Anaphora), and Participant Group (DoD, DoH)"}
 df_narration$ref_type2 <- factor(df_narration$ref_type2,levels = c("NOM", "PRO", "CL", "ZERO"))
@@
 For the current analysis, we did not define priors ourselves and used the preset uninformed priors.
-Figure 2[^1] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals.
+Figure 2[^2] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals.
 There are 3 responses we are considering in this model.
@@
 
-[^1]: We ran another model where we used the self-reported acquisition time as a predictor with 3 levels 0-3:Early, 4-7:Late1, and 7-above:Late2.
+[^2]: We ran another model where we used the self-reported acquisition time as a predictor with 3 levels 0-3:Early, 4-7:Late1, and 7-above:Late2.
 However, our model did not converge properly and the credible intervals were very wide spread, yielding the results uninformative.
@@
 
-As for the distribution of the zero anaphora, approximately 90% of the plain verbs were used to maintain a previous referent and this makes these constructions least implicit REs used for TİD[^2].
+As for the distribution of the zero anaphora, approximately 90% of the plain verbs were used to maintain a previous referent and this makes these constructions least implicit REs used for TİD[^3].
 Agreement verbs and CA were the following implicit markers of reference, `r sum(df_narration$GTR1 == "Person agreement verb" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1 == "Person agreement verb")` and `r sum(df_narration$GTR1_plot == "CA" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1_plot == "CA")` of which were used for maintained contexts, respectively.
@@
 However, there seems to be a common ground for the position of CA markers which have been claimed to rank higher than other markers of zero anaphora as in the present study [@frederiksen2016; @czubek2017; @barbera2009].
-In Figure \ref{fig:scale}, we report the distribution of REs in TİD as a function of referent accessibility and compare our accessibility hierarchy with that of [@frederiksen2016] for ASL.
+In Figure \ref{fig-scale}, we report the distribution of REs in TİD as a function of referent accessibility and compare our accessibility hierarchy with that of [@frederiksen2016] for ASL.
 In the hierarchies, the leftmost REs (e.g., modified and bare nouns for TİD) are used when the accessibility is low.
@@
 
-[^2]: As in other sign languages, in TİD, the lexical meaning of signs has an effect on the morphological grouping.
+[^3]: As in other sign languages, in TİD, the lexical meaning of signs has an effect on the morphological grouping.
 For instance, usually (di)transitive agreement verbs denote some sort of physical or metaphorical transfer of an object or abstract notion [@meir1999].
@@
 \caption{Comparing the present study and Frederiksen and Mayberry (2016) in terms of proposed accessibility hierarchies}
-\label{fig:scale}
+\label{fig-scale}
 \tikzset{every picture/.style={line width=0.75pt}} %set default line width to 0.75pt
@@
 ```
-Both hierarchies illustrated in Figure \ref{fig:scale} treat nominals and ExtCLs as markers of lowly accessibly referents.
+Both hierarchies illustrated in Figure \ref{fig-scale} treat nominals and ExtCLs as markers of lowly accessibly referents.
 Likewise, in both hierarchies, as markers of highly accessible referents, different types of zero anaphora markers are further distinguished.
@@
 However, this particular investigation did not present any data on the distribution of REs between two acquisition groups.
-In the present study, the finding that DoD signers preferred to use null arguments even for the first mention of a referent (`r sum(df_narration$ref_type == "Zero" & df_narration$Discourse == "Introduction")` in total) is surprising given some of the previous literature in the topicality and saliency framework [@cormier2013; @swabey2002].
+In the present study, the finding that DoD signers preferred to use null arguments even for the first mention of a referent (`r sum(df_narration$ref_type == "Zero" & df_narration$Discourse == "Introduction")` in total) is surprising given some of the previous literature in the topicality and saliency framework [@cormier2013; @swabey2002; @ferrara2022].
 These research reported very little or no use of zero anaphora by DoD signers for introduced contexts.

### 2023-05-01 revised_manuscript.tex
updated

@@
 Czubek, 2017; Ferrara et al., 2022; Frederiksen and Mayberry, 2016;
-Hodge et al., 2019; Morgan, 2005; Perniss and Özyürek, 2015; Swabey,
-2002). Building on the work by Gundel et al. (1993), Swabey (2002)
-demonstrated that, like other languages such as Russian, Chinese, and
-Japanese, American Sign Language (ASL) signers use bare nouns for
-introduced and re-introduced discourse contexts whereas zero anaphora is
-used for referents which are ``in-focus'' or very accessible to the
-addressee. Overt REs like pronouns or nouns are disprefered for such
-referents. Swabey (2002) concluded that ``native'' signers in her study
-abided by the Gricean maxim of quantity and thus used more economic
-forms of reference tracking. Similar findings have been reported for
-German Sign Language (Deutsche Gebärdensprache - DGS) (Perniss and
-Özyürek, 2015) and British Sign Language (BSL) (Morgan, 2005). It is
-important to note, however, that this research either collapsed DoD and
-DoH signers into a single group and labeled them as ``native-like'' or
-explicitly compared native deaf signers with late deaf signers (with
-delayed exposure to a sign language) in terms of narrative competence.
-Most recently, Ferrara et al. (2022) examined corpora including
-video-recorded narratives from five different sign languages including
-Irish Sign Language, Finnish Sign Language, Norwegian Sign Language and
-Swedish Sign Language and investigated how these languages established
-and maintained referents. Like other observations, this research showed
-that, across each of these languages, bare nouns were used to introduce
-new referents and zero anaphora or more reduced forms were preferred for
-referent maintenance.
+Garcia et al., 2018; Garcia and Sallandre, 2014; Hodge et al., 2019;
+Morgan, 2005; Perniss and Özyürek, 2015; Swabey, 2002). Building on the
+work by Gundel et al. (1993), Swabey (2002) demonstrated that, like
+other languages such as Russian, Chinese, and Japanese, American Sign
+Language (ASL) signers use bare nouns for introduced and re-introduced
+discourse contexts whereas zero anaphora is used for referents which are
+``in-focus'' or very accessible to the addressee. Overt REs like
+pronouns or nouns are disprefered for such referents. Swabey (2002)
+concluded that ``native'' signers in her study abided by the Gricean
+maxim of quantity and thus used more economic forms of reference
+tracking. Similar findings have been reported for German Sign Language
+(Deutsche Gebärdensprache - DGS) (Perniss and Özyürek, 2015) and British
+Sign Language (BSL) (Morgan, 2005). It is important to note, however,
+that this research either collapsed DoD and DoH signers into a single
+group and labeled them as ``native-like'' or explicitly compared native
+deaf signers with late deaf signers (with delayed exposure to a sign
+language) in terms of narrative competence. Most recently, Ferrara et
+al. (2022) examined corpora including video-recorded narratives from
+five different sign languages including Irish Sign Language, Finnish
+Sign Language, Norwegian Sign Language and Swedish Sign Language and
+investigated how these languages established and maintained referents.
+Like other observations, this research showed that, across each of these
+languages, bare nouns were used to introduce new referents and zero
+anaphora or more reduced forms were preferred for referent maintenance.
 
@@
 as in \ref{ex:2a} as well as fingerspelled nouns as in \ref{ex:2b}.
-Fingerspelling refers to the use of the manual aphabet to spell all or
+Fingerspelling refers to the use of the manual alphabet to spell all or
 part of the letters of a word borrowed from a surrounding written
@@
 For the analysis, see \ref{section:d}.}
+Detailed information about the linguistic and educational background of
+the participants can be found in \ref{section:a}.
+
 All participants indicated that their preferred language of
-communication was TİD which they used for most contexts. Detailed
-information about the linguistic and educational background of the
-participants can be found in \ref{section:a}. 1 DoD signer was excluded
-from the data analysis because they self-reported to have signing deaf
-parents but also only used home sign (a more crude form of
+communication was TİD which they used for most contexts. 1 DoD signer
+was excluded from the data analysis because they self-reported to have
+signing deaf parents but also only used home sign (a more crude form of
 communication) at home. As a result, we present data from only 14 DoD
@@
 distinguished among three types of handshapes: WECL, BPCL, and ExtCL.
-Finally, we had CA, agreement, plain tags for zero anaphora. We were
-unable to observe any instance of HCLs occurring in isolation (i.e., not
-part of CA). We should also highlight here that we coded referent status
-as nominal if the clause included an explicit nominal referent. For
-example, if constructed action, an index sign, or a classifier preceded
-or followed its explicitly signed nominal referent, the clause was
-regarded as having nominal reference (i.e., modified noun). The
-annotation tiers are presented in Table \ref{tab:tab-2}.
+Finally, we had CA\footnote{Note that CA can also be further subdivided
+  into overt, reduced, and subtle CA, following (Cormier et al., 2015).
+  In overt CA, the signer is ``fully in character'' and does not use any
+  other lexical or classifier constructions. In reduced CA, the signer
+  can involve a small amount of narration or lexical material while
+  tracking a referent using CA. However, subtle CA refers to cases where
+  the referent is primarily tracked by a non-CA construction (e.g.,
+  nominal, pronominal, verbal or classifier constructions) but the
+  signer subtly uses a few CA articulators at the same time (e.g.,
+  enacting the eye-gaze of the character). We only coded overt and
+  reduced types as reference tracking by CA, and disregarded the subtle
+  type. We'd also like to acknowledge that mixed CA, or simultaneous
+  tracking of two or multiple references through CA and other tools can
+  be possible (Dudis, 2004; Liddell, 2003). In such cases, as we explain
+  here, we only coded only the subject of a clause for maintenance and
+  re-introduction. If the referent was in subject position and tracked
+  by using CA, we considered this an instance of CA, as well.},
+agreement, plain tags for zero anaphora. We were unable to observe any
+instance of HCLs occurring in isolation (i.e., not part of CA). We
+should also highlight here that we coded referent status as nominal if
+the clause included an explicit nominal referent. For example, if
+constructed action, an index sign, or a classifier preceded or followed
+its explicitly signed nominal referent, the clause was regarded as
+having nominal reference (i.e., modified noun). The annotation tiers are
+presented in Table \ref{tab:tab-2}.
 
@@
 \centering
-\includegraphics{revised_manuscript_files/figure-latex/fig:count-plot-1.pdf}
+\includegraphics{revised_manuscript_files/figure-latex/fig-count-plot-1.pdf}
 \caption{Referent counts by Discourse Status (Introduction, Maintanence,
@@
 the present study (Barberà and Massó, 2009; Czubek, 2017; Frederiksen
-and Mayberry, 2016). In Figure \ref{fig:scale}, we report the
+and Mayberry, 2016). In Figure \ref{fig-scale}, we report the
 distribution of REs in TİD as a function of referent accessibility and
@@
 
-Both hierarchies illustrated in Figure \ref{fig:scale} treat nominals
+Both hierarchies illustrated in Figure \ref{fig-scale} treat nominals
 and ExtCLs as markers of lowly accessibly referents. Likewise, in both
@@
 literature in the topicality and saliency framework (Cormier et al.,
-2013; Swabey, 2002). These research reported very little or no use of
-zero anaphora by DoD signers for introduced contexts.
+2013; Ferrara et al., 2022; Swabey, 2002). These research reported very
+little or no use of zero anaphora by DoD signers for introduced
+contexts.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-cormier_rethinking_2015}{}}%
+Cormier, Kearsy, Smith, Sandra, Sevcikova-Sehyr, Zed, 2015. Rethinking
+constructed action. {SL}\&L 18, 167--204.
+doi:\href{https://doi.org/10.1075/sll.18.2.01cor}{10.1075/sll.18.2.01cor}
+
 \leavevmode\vadjust pre{\hypertarget{ref-cormier2013}{}}%
@@
 Yay{ı}nlar{ı}: Ankara.
+
+\leavevmode\vadjust pre{\hypertarget{ref-dudis2004}{}}%
+Dudis, Paul G., 2004.
+\href{https://www.degruyter.com/view/j/cogl.2004.15.issue-2/cogl.2004.009/cogl.2004.009.xml}{Body
+partitioning and real-space blends}. Cognitive Linguistics 15.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-garcia_reference_2014}{}}%
+Garcia, Brigitte, Sallandre, Marie-Anne, 2014. Reference resolution in
+french sign language ({LSF}), in: Cabredo Hofherr, Patricia,
+Zribi-Hertz, Anne (Eds.), Crosslinguistic Studies on Noun Phrase
+Structure and Reference. Brill, Leiden, pp. 316--364.
+doi:\href{https://doi.org/10.1163/9789004261440_012}{10.1163/9789004261440\_012}
+
+\leavevmode\vadjust pre{\hypertarget{ref-garcia2018impersonal}{}}%
+Garcia, Brigitte, Sallandre, Marie-Anne, L'Huillier, Marie-Thérèse,
+2018. Impersonal human reference in french sign language (LSF). Sign
+Language \& Linguistics 21, 307--333.
+
 \leavevmode\vadjust pre{\hypertarget{ref-givon1983}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-liddell2003}{}}%
+Liddell, Scott K., 2003. Grammar, gesture, and meaning in American Sign
+Language. Cambridge University Press.
+
 \leavevmode\vadjust pre{\hypertarget{ref-loew1984}{}}%

### 2023-05-02 revised_manuscript.Rmd
model results added
Added the model results to the appendix.

@@
    - \usepackage{graphicx}
-   - \usepackage{./sty/langsci-gb4e}
+   - \usepackage{langsci-gb4e}
    - \usepackage{caption}
@@
 
-[^2]: We ran another model where we used the self-reported acquisition time as a predictor with 3 levels 0-3:Early, 4-7:Late1, and 7-above:Late2.
+[^2]: The regression model results are given in \ref{section:modelResult}.
+We ran another model where we used the self-reported acquisition time as a predictor with 3 levels 0-3:Early, 4-7:Late1, and 7-above:Late2.
 However, our model did not converge properly and the credible intervals were very wide spread, yielding the results uninformative.
@@
 
+# Regression model results for RE Type
+
+\label{section:modelResult}
+
+```{r echo=FALSE, message=FALSE, cache=FALSE, results='asis', warning=FALSE}
+colnames(df_model_1)[1:4] <- c("Coefficients","outerWidth","innerWidth", "pointEstimate")
+df_model_1 %<>% mutate(across(.cols = "Coefficients", ~str_replace( ., "_","\\\\_")))
+apa_table(df_model_1,
+          format = "latex",
+          booktabs = T,
+          escape = FALSE,
+          landscape = FALSE,
+          font_size = "small",
+          placement = "hbt!",
+          caption ="Regression model results for the RE Type (Nominal, Zero, Classifier) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (DoD, DoH)"
+)
+```
+
+\clearpage
+
 # Additional model with 3 levels for Nativeness
@@
                                         `muNull_Group3M2:Discourse2` = 'Null_L2vsL1*Maintain')
-df_table_extra %<>% mutate( across(.cols = "Coefficients",~str_replace( ., "_","\\\\_")))
+df_table_extra %<>% mutate(across(.cols = "Coefficients",~str_replace( ., "_","\\\\_")))
 

### 2023-05-02 revised_manuscript.tex
model results added
Added the model results to the appendix.

@@
 \usepackage{graphicx}
-\usepackage{./sty/langsci-gb4e}
+\usepackage{langsci-gb4e}
 \usepackage{caption}
@@
 Sign Language}
+      \cortext[cor1]{Corresponding author}
 
@@
 of experiences, narrators must choose appropriate linguistic forms to
-refer to entities (Ariel, 1990; Givon, 1983; Gundel et al., 1993).
-Certain forms such as nominal expressions are commonly used for the
-first mention of a referent in a discourse, and other forms such as
-pronouns and zero anaphora may track previously mentioned entities which
-could either be maintained across multiple clauses or re-introduced back
-into the discourse. The selection of these referring expressions
-(henceforth REs) is contingent on cognitive abilities, such that
-individuals must rely on the activation of the working memory to keep in
-mind previously referred expressions and long strings of information
-(Bamberg, 1997; Morgan, 2005). Pragmatic abilities too play a pivotal
+refer to entities \citep{ariel1990, givon1983, gundel1993}. Certain
+forms such as nominal expressions are commonly used for the first
+mention of a referent in a discourse, and other forms such as pronouns
+and zero anaphora may track previously mentioned entities which could
+either be maintained across multiple clauses or re-introduced back into
+the discourse. The selection of these referring expressions (henceforth
+REs) is contingent on cognitive abilities, such that individuals must
+rely on the activation of the working memory to keep in mind previously
+referred expressions and long strings of information
+\citep{morgan2005, bamberg1997}. Pragmatic abilities too play a pivotal
 role in the tracking of referents. The form with the least amount of
@@
 of the referent (i.e., how active a referent is in the addressee's mind
-given a point in discourse) (Ahn, 2019; Ariel, 1990). Because of these
+given a point in discourse) \citep{ahn2019, ariel1990}. Because of these
 cognitive and pragmatic necessities, it is not surprising to observe
 that producing a coherent narrative is one of the latest milestones in
-language acquisition among children (Hickmann et al., 1996). As can be
+language acquisition among children \citep{hickmann1996}. As can be
 expected, there is a plethora of research examining how narrative
@@
 referent selection have been studied extensively for typically
-developing speakers (Aksu-Koç and Nicolopoulou, 2015; Hickmann et al.,
-1996; Williams, 1988), deaf individuals who acquired a sign language in
-atypical learning contexts (i.e., deaf signers with delayed exposure to
-a sign language) have remained mostly understudied with an exception of
-a handful of studies (Becker, 2009; Cormier et al., 2013; Gür, 2018).
-Nevertheless, it is paramount that we investigate and describe the
-effects of late exposure to a sign language among deaf adults growing up
-in a signing vs.~non-signing environment on narrative competence more in
-depth. This will help us understand whether the lack of systematic and
-frequent language exposure in the early years of life affect how deaf
-signers produce a narrative in their adulthood. Given this, the present
-study will focus on two groups of signers and examine their reference
-tracking strategies in signed narratives: (i) deaf of deaf (DoD) adult
-signers (i.e., signers who have deaf caregivers and natively acquired a
-sign language from birth) and (ii) deaf of hearing (DoH) adult signers
-(i.e., signers who have hearing non-signing caregivers and state to have
+developing speakers \citep{aksu-koc2015, hickmann1996, williams1988},
+deaf individuals who acquired a sign language in atypical learning
+contexts (i.e., deaf signers with delayed exposure to a sign language)
+have remained mostly understudied with an exception of a handful of
+studies \citep{cormier2013, gur2018, becker2009}. Nevertheless, it is
+paramount that we investigate and describe the effects of late exposure
+to a sign language among deaf adults growing up in a signing
+vs.~non-signing environment on narrative competence more in depth. This
+will help us understand whether the lack of systematic and frequent
+language exposure in the early years of life affect how deaf signers
+produce a narrative in their adulthood. Given this, the present study
+will focus on two groups of signers and examine their reference tracking
+strategies in signed narratives: (i) deaf of deaf (DoD) adult signers
+(i.e., signers who have deaf caregivers and natively acquired a sign
+language from birth) and (ii) deaf of hearing (DoH) adult signers (i.e.,
+signers who have hearing non-signing caregivers and state to have
 started acquiring a sign language later in life).
@@
 has been the officially recognized language of the Deaf community in
-Turkey since 2005 (see Nuhbalaoglu, 2018 for a brief review of the
-historical development of the language). Deaf or hearing children
+Turkey since 2005 \citep[see][ for a brief review of the historical
+development of the language]{nuhbalaoglu2018}. Deaf or hearing children
 acquiring a SL within a family where there is one or more deaf
 caregivers display the same developmental milestones as hearing children
-acquiring a spoken language in a hearing family (Chamberlain et al.,
-1999; Meier and Newport, 1990). However, reports indicate that over 90\%
-of all deaf individuals are born into hearing families (DoH) who do not
-know how to sign (Woll, 2013). This means that the auditory modality
-used in the household is not accessible to many deaf children. That is
-why the age of acquisition of a first functional language varies greatly
-among deaf children, with many of whom being deprived of their first
-languages typically until after age 7 when they start formal education
-in a deaf school, especially in the context of Turkey (Sari, 2005).
-Despite this, many of the early intervention programs for late signing
-children prioritize spoken language development and discourage the use
-of a sign language (Kemaloğlu and Kemaloğlu, 2012). The Ministry of
-National Education (2020) identifies 72 deaf schools in Turkey for the
-K-12 level. Even when deaf children with hearing parents start their
-formal education in one of these schools, they do not receive proper
-instruction in TİD since the teachers themselves are mainly hearing and
-only have a very rudimentary knowledge of a sign language (Ilkbasaran,
-2015; Sari, 2005). In some cases, it has been argued that late exposure
-to a sign language among certain DoH signers has negative effects on
-some cognitive and linguistic measures. To name a few, delayed exposure
-to a first functional language is reported to negatively affect deaf
-children and adults' morphosyntactic production and processing
-(Boudreault and Mayberry, 2006; Cheng and Mayberry, 2021; Cormier et
-al., 2012; Karadöller et al., 2017; Kayabaşı and Gökgöz, 2022; Newport,
-1990) and narrative abilities (e.g., Becker, 2009). In our overall
+acquiring a spoken language in a hearing family
+\citep{chamberlain_language_1999, meier1990}. However, reports indicate
+that over 90\% of all deaf individuals are born into hearing families
+(DoH) who do not know how to sign \citep{woll2013}. This means that the
+auditory modality used in the household is not accessible to many deaf
+children. That is why the age of acquisition of a first functional
+language varies greatly among deaf children, with many of whom being
+deprived of their first languages typically until after age 7 when they
+start formal education in a deaf school, especially in the context of
+Turkey \citep{sari_analysis_2005}. Despite this, many of the early
+intervention programs for late signing children prioritize spoken
+language development and discourage the use of a sign language
+\citep{kemaloglu2012}. The \citet{ministryofnationaleducation2020}
+identifies 72 deaf schools in Turkey for the K-12 level. Even when deaf
+children with hearing parents start their formal education in one of
+these schools, they do not receive proper instruction in TİD since the
+teachers themselves are mainly hearing and only have a very rudimentary
+knowledge of a sign language \citep{ilkbasaran2015, sari_analysis_2005}.
+In some cases, it has been argued that late exposure to a sign language
+among certain DoH signers has negative effects on some cognitive and
+linguistic measures. To name a few, delayed exposure to a first
+functional language is reported to negatively affect deaf children and
+adults' morphosyntactic production and processing
+\citep{kayaba2022, boudreault2006, newport1990, cheng2021, karadoller2017, cormier2012}
+and narrative abilities \citep[e.g.,][]{becker2009}. In our overall
 research, we are concerned with the effects of this late exposure in the
@@
 expressed referent is already at the center of attention, as indicated
-by the present contextual information (Chafe, 1976). The Givenness
+by the present contextual information \citep{chafe1976}. The Givenness
 Hierarchy is akin to other previously proposed models of reference
 control within the topicality and saliency paradigm such as the
-Accessibility Hierarchy (Ariel, 1990), Topic Continuity (Givon, 1983),
-and Familiarity Scale (Prince, 1981). Gundel et al. (1993) argued that
-the distribution of forms is not random but abides by the Gricean Maxim
-of Quantity (Grice, 1975). That is to say, individuals must make the
-most informative and succinct choice possible to retain economy of form.
-When multiple expressions are available for the same referent as in an
-utterance like (\ref{ex:1}), phonetically simplest form is selected with
-the least redundant information based on an economy principle. In the
-utterance, REs for ``Eddie'' are given in bold. All three REs (nominal,
-pronominal, and zero anaphora) refer to the same entity; however, the
-quantity of marking decreases as the referent's accessibility increases.
-On this ground, for highly accessible referents (i.e., those that are
-highly active in the mind), linguistic forms with the least phonetic
-content are selected for the least amount of processing effort.
+Accessibility Hierarchy \citep{ariel1990}, Topic Continuity
+\citep{givon1983}, and Familiarity Scale \citep{prince1981}.
+\citet{gundel1993} argued that the distribution of forms is not random
+but abides by the Gricean Maxim of Quantity \citep{grice1975}. That is
+to say, individuals must make the most informative and succinct choice
+possible to retain economy of form. When multiple expressions are
+available for the same referent as in an utterance like (\ref{ex:1}),
+phonetically simplest form is selected with the least redundant
+information based on an economy principle. In the utterance, REs for
+``Eddie'' are given in bold. All three REs (nominal, pronominal, and
+zero anaphora) refer to the same entity; however, the quantity of
+marking decreases as the referent's accessibility increases. On this
+ground, for highly accessible referents (i.e., those that are highly
+active in the mind), linguistic forms with the least phonetic content
+are selected for the least amount of processing effort.
 
@@
 based on the principles of reference accessibility. For spoken
-languages, this observation is well established (Ahrenholz, 2005; Ariel,
-1990; Azar and Özyürek, 2015; Chafe, 1976; Debreslioska et al., 2013;
-Givon, 1983; Gundel et al., 1993; Hickmann et al., 1996). These research
-indicate that the introduction and re-introduction of a referent are
-often accomplished by full noun phrases which can be commonly modified
-by an adjective or demonstrative. Maintenance, on the other hand, is
-typically marked with pronominals and zero anaphora since the continued
-referent is still highly accessible to the addressee.
+languages, this observation is well established
+\citep{ariel1990, givon1983, gundel1993, debreslioska2013, ahrenholz2005, chafe1976, hickmann1996, azar2015}.
+These research indicate that the introduction and re-introduction of a
+referent are often accomplished by full noun phrases which can be
+commonly modified by an adjective or demonstrative. Maintenance, on the
+other hand, is typically marked with pronominals and zero anaphora since
+the continued referent is still highly accessible to the addressee.
 
@@
 languages. There is an emerging body of literature that reports the
-distribution of REs in signed narratives (Ahlgren and Bergman, 1994;
-Czubek, 2017; Ferrara et al., 2022; Frederiksen and Mayberry, 2016;
-Garcia et al., 2018; Garcia and Sallandre, 2014; Hodge et al., 2019;
-Morgan, 2005; Perniss and Özyürek, 2015; Swabey, 2002). Building on the
-work by Gundel et al. (1993), Swabey (2002) demonstrated that, like
-other languages such as Russian, Chinese, and Japanese, American Sign
-Language (ASL) signers use bare nouns for introduced and re-introduced
-discourse contexts whereas zero anaphora is used for referents which are
-``in-focus'' or very accessible to the addressee. Overt REs like
-pronouns or nouns are disprefered for such referents. Swabey (2002)
-concluded that ``native'' signers in her study abided by the Gricean
-maxim of quantity and thus used more economic forms of reference
-tracking. Similar findings have been reported for German Sign Language
-(Deutsche Gebärdensprache - DGS) (Perniss and Özyürek, 2015) and British
-Sign Language (BSL) (Morgan, 2005). It is important to note, however,
-that this research either collapsed DoD and DoH signers into a single
-group and labeled them as ``native-like'' or explicitly compared native
-deaf signers with late deaf signers (with delayed exposure to a sign
-language) in terms of narrative competence. Most recently, Ferrara et
-al. (2022) examined corpora including video-recorded narratives from
-five different sign languages including Irish Sign Language, Finnish
-Sign Language, Norwegian Sign Language and Swedish Sign Language and
-investigated how these languages established and maintained referents.
-Like other observations, this research showed that, across each of these
-languages, bare nouns were used to introduce new referents and zero
-anaphora or more reduced forms were preferred for referent maintenance.
+distribution of REs in signed narratives
+\citep{czubek2017, ferrara2022, swabey2002, frederiksen2016, perniss2015, morgan2005, hodge2019, ahlgren1994, garcia_reference_2014, garcia2018impersonal}.
+Building on the work by \citet{gundel1993}, \citet{swabey2002}
+demonstrated that, like other languages such as Russian, Chinese, and
+Japanese, American Sign Language (ASL) signers use bare nouns for
+introduced and re-introduced discourse contexts whereas zero anaphora is
+used for referents which are ``in-focus'' or very accessible to the
+addressee. Overt REs like pronouns or nouns are disprefered for such
+referents. \citet{swabey2002} concluded that ``native'' signers in her
+study abided by the Gricean maxim of quantity and thus used more
+economic forms of reference tracking. Similar findings have been
+reported for German Sign Language (Deutsche Gebärdensprache - DGS)
+\citep{perniss2015} and British Sign Language (BSL) \citep{morgan2005}.
+It is important to note, however, that this research either collapsed
+DoD and DoH signers into a single group and labeled them as
+``native-like'' or explicitly compared native deaf signers with late
+deaf signers (with delayed exposure to a sign language) in terms of
+narrative competence. Most recently, \citet{ferrara2022} examined
+corpora including video-recorded narratives from five different sign
+languages including Irish Sign Language, Finnish Sign Language,
+Norwegian Sign Language and Swedish Sign Language and investigated how
+these languages established and maintained referents. Like other
+observations, this research showed that, across each of these languages,
+bare nouns were used to introduce new referents and zero anaphora or
+more reduced forms were preferred for referent maintenance.
 
@@
 handling classifiers, and extension classifiers), and overt pronouns
-(pronominal index or IX) (Frederiksen and Mayberry, 2016). This interim
-section describes such forms attested for sign languages under four main
+(pronominal index or IX) \citep{frederiksen2016}. This interim section
+describes such forms attested for sign languages under four main
 headings (nominals, pronouns, classifiers, and zero anaphora).
@@
 part of the letters of a word borrowed from a surrounding written
-language (Kubuş, 2008). Nouns can also be modified by a prenominal or
-postnominal extended index finger pointing sign, glossed as IX,
+language \citep{kubus2008}. Nouns can also be modified by a prenominal
+or postnominal extended index finger pointing sign, glossed as IX,
 functioning as a determiner or a demonstrative as in \ref{ex:2c}
-(Maclaughlin, 1997) or by a modifying classifier within the same NP as
-in \ref{ex:2d} (Frederiksen and Mayberry,
-2016).\footnote{Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions (Lillo-Martin and Klima, 1990).
+\citep{maclaughlin1997} or by a modifying classifier within the same NP
+as in \ref{ex:2d}
+\citep{frederiksen2016}.\footnote{Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions (Lillo-Martin and Klima, 1990).
 In addition, the pointing sign is used to referentially anchor a referent to a location in signing space.}
@@
 previously assigned abstract or physical locus in the signing space when
-re-introducing a topic in discourse (Emmorey, 1996). This function has
-also been identified for TİD (Sevinç, 2006). The example in \ref{ex:3}
-illustrates this spatially-anchored pronominal use.
+re-introducing a topic in discourse \citep{emmorey1996}. This function
+has also been identified for TİD \citep{sevinc2006}. The example in
+\ref{ex:3} illustrates this spatially-anchored pronominal use.
 
@@
 These pronominal signs are rarely used for the first mention of an
-entity (Perniss and Özyürek, 2015). Other researchers consider pointing
-signs without noun phrases as examples of demonstratives and not
-personal pronouns (Koulidobrova and Lillo-Martin, 2016). It is also
-important to mention here that in some spoken languages (e.g., German
-and Dutch), demonstratives can also refer to persons (Bosch and Umbach,
-2007) and behave similarly to personal pronouns. However, for the
-purposes of the present study, we will refer to IX signs as pronominal
-IX and particularly examine their non-deictic or personal pronoun
-realizations as a possible discourse strategy to mainly mark topic shift
-(re-introduction) among DoD and DoH signers of TİD.
+entity \citep{perniss2015}. Other researchers consider pointing signs
+without noun phrases as examples of demonstratives and not personal
+pronouns \citep{koulidobrova2016}. It is also important to mention here
+that in some spoken languages (e.g., German and Dutch), demonstratives
+can also refer to persons \citep{bosch2007} and behave similarly to
+personal pronouns. However, for the purposes of the present study, we
+will refer to IX signs as pronominal IX and particularly examine their
+non-deictic or personal pronoun realizations as a possible discourse
+strategy to mainly mark topic shift (re-introduction) among DoD and DoH
+signers of TİD.
 
@@
 encoded by a handshape. This handshape is articulated co-temporally with
-a movement that encodes the verbal meaning (Benedicto and Brentari,
-2004; Supalla, 1990). Semantically, the classifiers denote sets of
-objects based on criteria such as animacy, size, shape, body-part,
-texture and sometimes function (Talmy, 2003). The classifier articulated
-simultaneously with a movement is called a classifier predicate and the
-sentence in which such a predicate is used is usually called a
-classifier construction in sign language research. As in most sign
-languages hitherto studied, classifier constructions are also commonly
-found in TİD (see Dikyuva et al., 2015 for a review). Following the
-classification of (Engberg-Pedersen, 1993), in TİD, we distinguish
-between four types of classifier handshapes. Whole entity classifiers
-(WECL) represent the entirety of a referent (e.g.~an entire flat
-horizontal entity, or an upright entity) (e.g., \ref{ex:4a}) whereas
-body part classifiers (BPCL) represent a specific part of an animate
-entity (e.g., limbs, head, mouth, tongue, or teeth, etc.)
+a movement that encodes the verbal meaning
+\citep{supalla1990, benedicto2004}. Semantically, the classifiers denote
+sets of objects based on criteria such as animacy, size, shape,
+body-part, texture and sometimes function \citep{talmy2003}. The
+classifier articulated simultaneously with a movement is called a
+classifier predicate and the sentence in which such a predicate is used
+is usually called a classifier construction in sign language research.
+As in most sign languages hitherto studied, classifier constructions are
+also commonly found in TİD \citep[see][ for a review]{dikyuva2015}.
+Following the classification of \citep{engberg-pedersen1993}, in TİD, we
+distinguish between four types of classifier handshapes. Whole entity
+classifiers (WECL) represent the entirety of a referent (e.g.~an entire
+flat horizontal entity, or an upright entity) (e.g., \ref{ex:4a})
+whereas body part classifiers (BPCL) represent a specific part of an
+animate entity (e.g., limbs, head, mouth, tongue, or teeth, etc.)
 (e.g.,\ref{ex:4b}). Handling classifiers (HCL) occur with transitive
@@
 characters. This is often achieved with the signer taking on the role of
-a referent in the narrative and performing the actions (Metzger, 1995).
-CA is often marked by certain non-manual cues such as the facial
-expression and the breaking of the eye contact. This is to say, the
-narrator's head, face, torso, arms, and as an optional marker, hands
-represent those of the associated referent(s) (Smith and Cormier, 2014).
+a referent in the narrative and performing the actions
+\citep{metzger1995}. CA is often marked by certain non-manual cues such
+as the facial expression and the breaking of the eye contact. This is to
+say, the narrator's head, face, torso, arms, and as an optional marker,
+hands represent those of the associated referent(s) \citep{smith2014}.
 As for the verbal constructions, to identify the predicate types in ASL,
-Padden (1986) suggested a triadic categorization of verbs in sign
+\citet{padden1986} suggested a triadic categorization of verbs in sign
 languages: plain verbs, agreement verbs, and spatial verbs. According to
@@
 person. TİD too licenses null arguments. Following this classification,
-Sevinç (2006) argued that although it is common in TİD to drop
+\citet{sevinc2006} argued that although it is common in TİD to drop
 arguments, there were certain limitations on the null marking of the
@@
 \textsc{eddie}) in a set but rather they refer to a set of individuals
-that are semantically similar to each other (Morgan, 2005). For example,
-the extended index finger can be used to refer to a human entity.
-Finally, zero anaphors predicated by constructed action, agreement and
-plain verbs are least explicit forms of referent tracking since the
-implicit nature of such constructions do not easily restrict the
-referent set based on cues such as size, shape or semantic category. In
-other words, zero anaphora requires preceding context or information the
-most compared to other forms.
+that are semantically similar to each other \citep{morgan2005}. For
+example, the extended index finger can be used to refer to a human
+entity. Finally, zero anaphors predicated by constructed action,
+agreement and plain verbs are least explicit forms of referent tracking
+since the implicit nature of such constructions do not easily restrict
+the referent set based on cues such as size, shape or semantic category.
+In other words, zero anaphora requires preceding context or information
+the most compared to other forms.
 
@@
 Like in first language development, narrative skills have been
-extensively studied among second language learners (see Bel et al.,
-2015; Frederiksen and Mayberry, 2019). Comparing 11 deaf adult signers
+extensively studied among second language learners
+\citep[see][]{bel2015, frederiksen2019}. Comparing 11 deaf adult signers
 and 13 hearing advanced learners of Catalan Sign Language (Llengua de
-Signes Catalana - LSC), Bel et al. (2015) found that hearing signers
+Signes Catalana - LSC), \citet{bel2015} found that hearing signers
 resort to more overt forms (nominals and pronouns) than native deaf
@@
 also deaf signers who had hearing parents and their age of acquisition
-varied. Out of 11 deaf L1 signers in Bel et al.'s (2015) control sample,
-5 had deaf parents and acquired LSC from birth, and the other 6 had
-hearing parents but their exposure to LSC started before age 5. A
-similar line of research came from Frederiksen and Mayberry (2019) who
+varied. Out of 11 deaf L1 signers in Bel et al.'s \citeyearpar{bel2015}
+control sample, 5 had deaf parents and acquired LSC from birth, and the
+other 6 had hearing parents but their exposure to LSC started before age
+5. A similar line of research came from \citet{frederiksen2019} who
 found a statistically significant difference between both groups, that
@@
 signers with delayed exposure to a sign language) also display
-redundancy in their reference tracking. Becker (2009) examined 6 DGS
-signing deaf children's (3 native and 3 late) narrative abilities. She
-reported that late signing children generated more explicit forms such
-as nominals when referring to the characters in the story compared to
-native learners, concluding that first language deprivation results in
-an inadequate representation of the addressee and less efficient use of
-morphosyntax. Comparably, Cormier et al. (2013) investigated the use of
+redundancy in their reference tracking. \citet{becker2009} examined 6
+DGS signing deaf children's (3 native and 3 late) narrative abilities.
+She reported that late signing children generated more explicit forms
+such as nominals when referring to the characters in the story compared
+to native learners, concluding that first language deprivation results
+in an inadequate representation of the addressee and less efficient use
+of morphosyntax. Comparably, \citet{cormier2013} investigated the use of
 CA in BSL among 15 deaf native, early, and late adult signers (5
@@
 maintaining previous referents. Most relevantly to the present study,
-Gür (2018) studied the development of narrative skills among 10 native
-and 10 late deaf children and deaf adult signers of TİD. She found that
-native-signing children and adults had more instances of explicit
-referent introduction, set the scene more frequently, and employed CA in
-an appropriate manner more than corresponding late-signing children and
-adults. Although this study presented preliminary findings on age of
-acquisition effects on narrative competence, it does not investigate how
-DoD and DoH signers employ different linguistic tools based on the
-accessibility of the referent. Therefore, to our knowledge, there is a
-lack of research directly examining the effects of first language delay
-especially among deaf adult signers with non-signing hearing parents on
-the selection of REs in signed narratives.
+\citet{gur2018} studied the development of narrative skills among 10
+native and 10 late deaf children and deaf adult signers of TİD. She
+found that native-signing children and adults had more instances of
+explicit referent introduction, set the scene more frequently, and
+employed CA in an appropriate manner more than corresponding
+late-signing children and adults. Although this study presented
+preliminary findings on age of acquisition effects on narrative
+competence, it does not investigate how DoD and DoH signers employ
+different linguistic tools based on the accessibility of the referent.
+Therefore, to our knowledge, there is a lack of research directly
+examining the effects of first language delay especially among deaf
+adult signers with non-signing hearing parents on the selection of REs
+in signed narratives.
 
@@
 Re-introduction) and RE type (Nominal, Pronominal, Classifier, Zero
-Anaphora). We used ELAN Linguistic Annotation Software (Crasborn and
-Sloetjes, 2008) for data annotation. First Author who is a hearing
+Anaphora). We used ELAN Linguistic Annotation Software
+\citep{crasborn2008} for data annotation. First Author who is a hearing
 signer of TİD annotated all the data and translated the narratives into
@@
 
-Following Berman and Slobin (2013), we first identified the clause
+Following \citet{berman2013relating}, we first identified the clause
 boundaries in each sentence indicated by the presence of predicates and
@@
 Finally, we had CA\footnote{Note that CA can also be further subdivided
-  into overt, reduced, and subtle CA, following (Cormier et al., 2015).
-  In overt CA, the signer is ``fully in character'' and does not use any
-  other lexical or classifier constructions. In reduced CA, the signer
-  can involve a small amount of narration or lexical material while
-  tracking a referent using CA. However, subtle CA refers to cases where
-  the referent is primarily tracked by a non-CA construction (e.g.,
-  nominal, pronominal, verbal or classifier constructions) but the
-  signer subtly uses a few CA articulators at the same time (e.g.,
-  enacting the eye-gaze of the character). We only coded overt and
-  reduced types as reference tracking by CA, and disregarded the subtle
-  type. We'd also like to acknowledge that mixed CA, or simultaneous
-  tracking of two or multiple references through CA and other tools can
-  be possible (Dudis, 2004; Liddell, 2003). In such cases, as we explain
-  here, we only coded only the subject of a clause for maintenance and
+  into overt, reduced, and subtle CA, following
+  \citep{cormier_rethinking_2015}. In overt CA, the signer is ``fully in
+  character'' and does not use any other lexical or classifier
+  constructions. In reduced CA, the signer can involve a small amount of
+  narration or lexical material while tracking a referent using CA.
+  However, subtle CA refers to cases where the referent is primarily
+  tracked by a non-CA construction (e.g., nominal, pronominal, verbal or
+  classifier constructions) but the signer subtly uses a few CA
+  articulators at the same time (e.g., enacting the eye-gaze of the
+  character). We only coded overt and reduced types as reference
+  tracking by CA, and disregarded the subtle type. We'd also like to
+  acknowledge that mixed CA, or simultaneous tracking of two or multiple
+  references through CA and other tools can be possible
+  \citep{dudis2004, liddell2003}. In such cases, as we explain here, we
+  only coded only the subject of a clause for maintenance and
   re-introduction. If the referent was in subject position and tracked
@@
 
+\begin{verbatim}
+## Warning: The dot-dot notation (`..count..`) was deprecated in ggplot2 3.4.0.
+## i Please use `after_stat(count)` instead.
+## This warning is displayed once every 8 hours.
+## Call `lifecycle::last_lifecycle_warnings()` to see where this warning was
+## generated.
+\end{verbatim}
+
 \begin{figure}
@@
 regression model to RE Type using discourse status and acquisition group
-as predictors. We used the brms package (Bürkner, 2018) in R, with a
+as predictors. We used the brms package \citep{burkner2018} in R, with a
 categorical distribution on the responses and sum contrasts for the
@@
 priors ourselves and used the preset uninformed priors. Figure
-2\footnote{We ran another model where we used the self-reported
-  acquisition time as a predictor with 3 levels 0-3:Early, 4-7:Late1,
-  and 7-above:Late2. However, our model did not converge properly and
-  the credible intervals were very wide spread, yielding the results
-  uninformative. We provide the model results as a plot in
+2\footnote{The regression model results are given in
+  \ref{section:modelResult}. We ran another model where we used the
+  self-reported acquisition time as a predictor with 3 levels 0-3:Early,
+  4-7:Late1, and 7-above:Late2. However, our model did not converge
+  properly and the credible intervals were very wide spread, yielding
+  the results uninformative. We provide the model results as a plot in
   \ref{section:d}.} shows the posterior probability distributions where
@@
 adult deaf signers with delayed language exposure in terms of their
-morphosyntactic (Boudreault and Mayberry, 2006; Cormier et al., 2012;
-Kayabaşı and Gökgöz, 2022) and narrative development (Becker, 2009;
-Cormier et al., 2013), we expected that DoH signers would be more
-over-redundant in their referent tracking, perhaps akin to second
-language learners (Bel et al., 2015; Frederiksen and Mayberry, 2019).
-That is, we had the expectation that DoH signers would use more explicit
-markers like nouns and fewer implicit markers like zero anaphora even
-for highly accessible referents (when the discourse status is maintained
-or even re-introduced in same cases) than DoD signers.
+morphosyntactic \citep{boudreault2006, cormier2012, kayaba2022} and
+narrative development \citep{becker2009, cormier2013}, we expected that
+DoH signers would be more over-redundant in their referent tracking,
+perhaps akin to second language learners
+\citep{frederiksen2019, bel2015}. That is, we had the expectation that
+DoH signers would use more explicit markers like nouns and fewer
+implicit markers like zero anaphora even for highly accessible referents
+(when the discourse status is maintained or even re-introduced in same
+cases) than DoD signers.
 
@@
 the present study brought back referents into the discourse more
-frequently compared to the participants in Frederiksen and Mayberry
-(2019) and Frederiksen and Mayberry (2016) which reported that
-re-introduction contexts accounted for only 7\% of the referents in ASL.
-On the other hand, maintenance percentages were lower in the present
-study compared to Frederiksen and Mayberry (2016) (40\% versus 69\%).
-Investigating the distribution of REs across discourse statuses in ASL,
-a similar study by Czubek (2017) also reported contrasting numbers (17\%
-for introduction, 64\% for maintenance, and 19\% for re-introduction).
-We suggest that the discrepancy might be due to the nature of the
-stimuli. In Frederiksen and Mayberry (2016), the authors used a
-4-6-episode-long picture story with 2 animate characters and 1 inanimate
-object. In contrast, the stimuli in the present study consisted of
-8-episode-long video clips on average with a mean number of 2 animate
-characters and 3 inanimate objects. Although the number of animate
-characters and inanimate characters in Czubek's (2017) stimuli matched
-that of the present study, their narrative structure (a 6-episode-long
-picture story) was still shorter compared to our narrative stimuli. We
-believe fewer possible referents and/or a simpler narrative structure in
-these studies may have contributed to the lower number of re-introduced
+frequently compared to the participants in \citet{frederiksen2019} and
+\citet{frederiksen2016} which reported that re-introduction contexts
+accounted for only 7\% of the referents in ASL. On the other hand,
+maintenance percentages were lower in the present study compared to
+\citet{frederiksen2016} (40\% versus 69\%). Investigating the
+distribution of REs across discourse statuses in ASL, a similar study by
+\citet{czubek2017} also reported contrasting numbers (17\% for
+introduction, 64\% for maintenance, and 19\% for re-introduction). We
+suggest that the discrepancy might be due to the nature of the stimuli.
+In \citet{frederiksen2016}, the authors used a 4-6-episode-long picture
+story with 2 animate characters and 1 inanimate object. In contrast, the
+stimuli in the present study consisted of 8-episode-long video clips on
+average with a mean number of 2 animate characters and 3 inanimate
+objects. Although the number of animate characters and inanimate
+characters in Czubek's \citeyearpar{czubek2017} stimuli matched that of
+the present study, their narrative structure (a 6-episode-long picture
+story) was still shorter compared to our narrative stimuli. We believe
+fewer possible referents and/or a simpler narrative structure in these
+studies may have contributed to the lower number of re-introduced
 contexts compared to the present study. Due to the story length and
@@
 our participants might have engaged in referent re-introduction more
-than what was reported in previous studies (Czubek, 2017; Frederiksen
-and Mayberry, 2016).
+than what was reported in previous studies
+\citep{czubek2017, frederiksen2016}.
 
@@
 finding parallels other accounts of referential accessibility suggested
-for other sign languages (see Frederiksen and Mayberry (2016); Czubek
-(2017); and Swabey (2002); for ASL, Morgan (2005); for BSL Bel et al.
-(2015); Barberà and Massó (2009) for LSC; Ferrara et al. (2022) for
-other sign languages). Our results also indicate that modified nouns
-(e.g., nouns that are preceded or followed by a classifier or IX)
-appeared more in introduced contexts compared to re-introduced contexts.
-This indicates that modified nouns are used more to track referents with
-lower accessibility, probably due to their more explicit structure.
+for other sign languages (see \citet{frederiksen2016};
+\citet{czubek2017}; and \citet{swabey2002}; for ASL, \citet{morgan2005};
+for BSL \citet{bel2015}; \citet{barbera2009} for LSC;
+\citet{ferrara2022} for other sign languages). Our results also indicate
+that modified nouns (e.g., nouns that are preceded or followed by a
+classifier or IX) appeared more in introduced contexts compared to
+re-introduced contexts. This indicates that modified nouns are used more
+to track referents with lower accessibility, probably due to their more
+explicit structure.
 
@@
 were mainly used to re-introduce referents rather than maintain them.
-This finding is comparable with what has been reported for ASL (Czubek,
-2017; Frederiksen and Mayberry, 2016; Swabey, 2002) but not with LSC
-(Bel et al., 2015), BSL (Morgan, 2005) and DGS (Perniss and Özyürek,
-2015), in which the signers extensively made overt pronominal references
-(up to 30\%) to maintain entities. Comparable to our findings, it has
-been reported that pronouns do not frequently occur in the local context
-and instead follows a topic shift and thus rank higher than classifiers
-and zero anaphora (see Barberà and Massó, 2009 for LSC; and Nuhbalaoglu,
-2018 for TİD). Considering this, the claim that the IX sign functions
-like demonstratives that are also used to track referents in
-re-introduced contexts has not been borne out (Koulidobrova and
-Lillo-Martin, 2016). Similar to these research, our findings indicate
-that the pronominal IX in TİD was used more concurrently with nominal
-constructions and pronominal IX was used very infrequently in isolation.
-The observation that overt pronouns are rare in ASL narratives comes
-from studies that include a simple picture retelling task with one
-protagonist (e.g., Frederiksen and Mayberry, 2016) as well as a more
-complex narrative design as in (Czubek, 2017). However, Czubek (2017)
-also incorporated different genres and reported more instances of overt
+This finding is comparable with what has been reported for ASL
+\citep{swabey2002, frederiksen2016, czubek2017} but not with LSC
+\citep{bel2015}, BSL \citep{morgan2005} and DGS \citep{perniss2015}, in
+which the signers extensively made overt pronominal references (up to
+30\%) to maintain entities. Comparable to our findings, it has been
+reported that pronouns do not frequently occur in the local context and
+instead follows a topic shift and thus rank higher than classifiers and
+zero anaphora \citetext{\citealp[see][ for
+LSC]{barbera2009}; \citealp[and][ for TİD]{nuhbalaoglu2018}}.
+Considering this, the claim that the IX sign functions like
+demonstratives that are also used to track referents in re-introduced
+contexts has not been borne out \citep{koulidobrova2016}. Similar to
+these research, our findings indicate that the pronominal IX in TİD was
+used more concurrently with nominal constructions and pronominal IX was
+used very infrequently in isolation. The observation that overt pronouns
+are rare in ASL narratives comes from studies that include a simple
+picture retelling task with one protagonist
+\citep[e.g.,][]{frederiksen2016} as well as a more complex narrative
+design as in \citep{czubek2017}. However, \citet{czubek2017} also
+incorporated different genres and reported more instances of overt
 pronouns in personal narratives (compared to picture or video
@@
 might be genre-dependent. Moreover, another production study on referent
-tracking in ASL reported object preference for pronominals (Frederiksen
-and Mayberry, 2017). However, we only coded subjects for maintained and
-re-introduced referents. We argue that all of these factors
+tracking in ASL reported object preference for pronominals
+\citep{frederiksen2017}. However, we only coded subjects for maintained
+and re-introduced referents. We argue that all of these factors
 (genre-dependency as well as object preference) might contribute to the
@@
 previous referent without having to use an explict form. As pointed out
-by Czubek (2017), signers prefer to rely on ExtCLs for the first mention
-of an entity because these constructions are informative and explicit by
-nature, and they provide visual and descriptive cues about an entity and
-its location in space. We only observed 3 instances of ExtCLs for
-re-introduced contexts, replicating Frederiksen and Mayberry's (2019)
-findings for ASL.
+by \citet{czubek2017}, signers prefer to rely on ExtCLs for the first
+mention of an entity because these constructions are informative and
+explicit by nature, and they provide visual and descriptive cues about
+an entity and its location in space. We only observed 3 instances of
+ExtCLs for re-introduced contexts, replicating Frederiksen and
+Mayberry's \citeyearpar{frederiksen2019} findings for ASL.
 
@@
   agreement verbs denote some sort of physical or metaphorical transfer
-  of an object or abstract notion (Meir, 1999). As for plain verbs, both
-  phonology of a sign and its semantics have an effect on morphology.
-  For instance, \textsc{eat} in TİD is signed nearby the mouth and since
-  it needs to remain in that phonological location, it is not suitable
-  for agreement and thus it is regarded as a plain verb. Likewise the
-  experiencer verb \textsc{love} is signed on the torso and as such it
-  does not take part in agreement. These generalizations noted we should
-  also be cautious in labelling a verb as categorially one type or the
-  other. For instance, some intransitive plain verbs like \textsc{fall}
-  or \textsc{jump} in TİD may display spatial person agreement with the
-  single argument in the clause (Gökgöz and Sevgi, 2020). See Lourenço
-  and Wilbur (2018) also for a similar discussion in Brazilian Sign
-  Language.}. Agreement verbs and CA were the following implicit markers
-of reference, 0.77 and 0.72 of which were used for maintained contexts,
-respectively. This contrasts with some of the research (Czubek, 2017) in
-which agreement verbs ranked lowest on the scale, indicating that these
-constructions were more implicit compared to plain verbs. However, there
-seems to be a common ground for the position of CA markers which have
-been claimed to rank higher than other markers of zero anaphora as in
-the present study (Barberà and Massó, 2009; Czubek, 2017; Frederiksen
-and Mayberry, 2016). In Figure \ref{fig-scale}, we report the
-distribution of REs in TİD as a function of referent accessibility and
-compare our accessibility hierarchy with that of (Frederiksen and
-Mayberry, 2016) for ASL. In the hierarchies, the leftmost REs (e.g.,
-modified and bare nouns for TİD) are used when the accessibility is low.
-If the referent is already active in the addressee's mind and its
-accessibility is high, then the rightmost markers (e.g., agreement and
-plain verbs for TİD) are selected.
+  of an object or abstract notion \citep{meir1999}. As for plain verbs,
+  both phonology of a sign and its semantics have an effect on
+  morphology. For instance, \textsc{eat} in TİD is signed nearby the
+  mouth and since it needs to remain in that phonological location, it
+  is not suitable for agreement and thus it is regarded as a plain verb.
+  Likewise the experiencer verb \textsc{love} is signed on the torso and
+  as such it does not take part in agreement. These generalizations
+  noted we should also be cautious in labelling a verb as categorially
+  one type or the other. For instance, some intransitive plain verbs
+  like \textsc{fall} or \textsc{jump} in TİD may display spatial person
+  agreement with the single argument in the clause \citep{gokgoz2020}.
+  See \citet{lourenço2018} also for a similar discussion in Brazilian
+  Sign Language.}. Agreement verbs and CA were the following implicit
+markers of reference, 0.77 and 0.72 of which were used for maintained
+contexts, respectively. This contrasts with some of the research
+\citep{czubek2017} in which agreement verbs ranked lowest on the scale,
+indicating that these constructions were more implicit compared to plain
+verbs. However, there seems to be a common ground for the position of CA
+markers which have been claimed to rank higher than other markers of
+zero anaphora as in the present study
+\citep{frederiksen2016, czubek2017, barbera2009}. In Figure
+\ref{fig-scale}, we report the distribution of REs in TİD as a function
+of referent accessibility and compare our accessibility hierarchy with
+that of \citep{frederiksen2016} for ASL. In the hierarchies, the
+leftmost REs (e.g., modified and bare nouns for TİD) are used when the
+accessibility is low. If the referent is already active in the
+addressee's mind and its accessibility is high, then the rightmost
+markers (e.g., agreement and plain verbs for TİD) are selected.
 
@@
 ground, both scales lack pronouns, which occurred very infrequently in
-the data. Although Frederiksen and Mayberry (2016) did not include other
+the data. Although \citet{frederiksen2016} did not include other
 classifier types in the hierarchy, we propose that WECLs and BPCLs must
@@
 study, a tendency for limited over-redundancy has been suggested for
-late signing children in DGS (Becker, 2009) and late signing adults in
-BSL specifically for the use of CA (Cormier et al., 2013). As for TİD,
-to our knowledge, there is only one study (Gür, 2018) that directly
+late signing children in DGS \citep{becker2009} and late signing adults
+in BSL specifically for the use of CA \citep{cormier2013}. As for TİD,
+to our knowledge, there is only one study \citep{gur2018} that directly
 aimed at examining age of acquisition effects on adult signers'
-narrative skills. Gür (2018) observed that late signing deaf adults and
-children set the scene less frequently and used CA less accurately
+narrative skills. \citet{gur2018} observed that late signing deaf adults
+and children set the scene less frequently and used CA less accurately
 compared to their native counterparts. However, this particular
@@
 of a referent (19 in total) is surprising given some of the previous
-literature in the topicality and saliency framework (Cormier et al.,
-2013; Ferrara et al., 2022; Swabey, 2002). These research reported very
-little or no use of zero anaphora by DoD signers for introduced
+literature in the topicality and saliency framework
+\citep{cormier2013, swabey2002, ferrara2022}. These research reported
+very little or no use of zero anaphora by DoD signers for introduced
 contexts.
@@
 findings of the line of research that investigated second language
-learners of a sign language (Bel et al., 2015; Frederiksen and Mayberry,
-2019). The authors reported that more instances of zero anaphora in the
-context of referent introduction was attested for DoD signers than
-second language learners. Although using zero anaphora to introduce a
-referent into the discourse would normally result in referential
-ambiguity, it is important to note that our stimuli consisted of 10
-interconnected narratives with 2 or more recurring characters. It may be
-possible that DoD signers, having already introduced the referent with
-an explicit nominal marker in a different discourse context, found it
-felicitous to refer to the same entity with more implicit markers in the
-following narratives even for introduced contexts. Then, our results
-imply a difference in pragmatic sensitivity to referential accessibility
-between DoD and DoH signers, which is more robustly reflected in
-introduced contexts. To test this implication, one could use the same
-number and length of stimuli from different cartoons with different
-characters and see if the tendency to introduce a referent without an
-overt nominal for the first time is dependent on the task or not.
+learners of a sign language \citep{bel2015, frederiksen2019}. The
+authors reported that more instances of zero anaphora in the context of
+referent introduction was attested for DoD signers than second language
+learners. Although using zero anaphora to introduce a referent into the
+discourse would normally result in referential ambiguity, it is
+important to note that our stimuli consisted of 10 interconnected
+narratives with 2 or more recurring characters. It may be possible that
+DoD signers, having already introduced the referent with an explicit
+nominal marker in a different discourse context, found it felicitous to
+refer to the same entity with more implicit markers in the following
+narratives even for introduced contexts. Then, our results imply a
+difference in pragmatic sensitivity to referential accessibility between
+DoD and DoH signers, which is more robustly reflected in introduced
+contexts. To test this implication, one could use the same number and
+length of stimuli from different cartoons with different characters and
+see if the tendency to introduce a referent without an overt nominal for
+the first time is dependent on the task or not.
 
@@
 possible variation in the accessibility levels of re-introduced
-referents, a more fine-tuned protocol (as in Toole, 1996) can be used
-for measuring referent distance and the number of other possible
+referents, a more fine-tuned protocol \citep[as in][]{toole1996} can be
+used for measuring referent distance and the number of other possible
 competitors. Furthermore, our focus in this study was to look at
@@
 We conclude that DoH signers have overall native-like competency in
-simple narratives, however, following Swabey (2002)'s intuition, our
-findings also suggest that DoD signers adhere to pragmatic norms and
+simple narratives, however, following \citet{swabey2002}'s intuition,
+our findings also suggest that DoD signers adhere to pragmatic norms and
 leverage accessibility better than DoH signers to a certain extent
@@
 
+\hypertarget{regression-model-results-for-re-type}{%
+\section{Regression model results for RE
+Type}\label{regression-model-results-for-re-type}}
+
+\label{section:modelResult}
+
+\begin{table}[hbt!]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:unnamed-chunk-4}Regression model results for the RE Type (Nominal, Zero, Classifier) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (DoD, DoH)}
+
+\small{
+
+\begin{tabular}{llllllllll}
+\toprule
+Coefficients & \multicolumn{1}{c}{outerWidth} & \multicolumn{1}{c}{innerWidth} & \multicolumn{1}{c}{pointEstimate} & \multicolumn{1}{c}{ll} & \multicolumn{1}{c}{l} & \multicolumn{1}{c}{m} & \multicolumn{1}{c}{h} & \multicolumn{1}{c}{hh} & \multicolumn{1}{c}{model}\\
+\midrule
+Nom\_Intercept & 0.90 & 0.50 & median & 0.86 & 1.06 & 1.18 & 1.31 & 1.50 & All\\
+Zero\_Intercept & 0.90 & 0.50 & median & 0.11 & 0.26 & 0.37 & 0.46 & 0.59 & All\\
+Nom\_DoD & 0.90 & 0.50 & median & -0.67 & -0.31 & -0.06 & 0.18 & 0.55 & All\\
+Nom\_Intro & 0.90 & 0.50 & median & 3.34 & 3.56 & 3.72 & 3.88 & 4.13 & All\\
+Nom\_Maintain & 0.90 & 0.50 & median & -5.70 & -5.49 & -5.34 & -5.20 & -4.98 & All\\
+Nom\_DoD*Intro & 0.90 & 0.50 & median & -0.09 & 0.34 & 0.65 & 0.98 & 1.44 & All\\
+Nom\_DoD*Maintain & 0.90 & 0.50 & median & -1.86 & -1.45 & -1.16 & -0.88 & -0.46 & All\\
+Zero\_DoD & 0.90 & 0.50 & median & 0.45 & 0.72 & 0.89 & 1.10 & 1.40 & All\\
+Zero\_Intro & 0.90 & 0.50 & median & -4.77 & -4.22 & -3.87 & -3.55 & -3.13 & All\\
+Zero\_Maintain & 0.90 & 0.50 & median & 1.75 & 1.99 & 2.17 & 2.37 & 2.66 & All\\
+Zero\_DoD*Intro & 0.90 & 0.50 & median & 0.10 & 0.93 & 1.58 & 2.26 & 3.32 & All\\
+Zero\_DoD*Maintain & 0.90 & 0.50 & median & -1.95 & -1.38 & -0.99 & -0.64 & -0.17 & All\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\clearpage
+
 \hypertarget{additional-model-with-3-levels-for-nativeness}{%
@@
 
-\caption{\label{tab:unnamed-chunk-4}Model results for Nativeness with 3 levels}
+\caption{\label{tab:unnamed-chunk-5}Model results for Nativeness with 3 levels}
 
@@
 
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-ahlgren1994}{}}%
-Ahlgren, Inger, Bergman, Brita, 1994. Reference in narratives. I.
-Ahlgren, B. Bergman \& M.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ahn2019}{}}%
-Ahn, Dorothy, 2019. The determinacy scale: A competition mechanism for
-anaphoric expressions (PhD thesis).
-
-\leavevmode\vadjust pre{\hypertarget{ref-ahrenholz2005}{}}%
-Ahrenholz, Bernt, 2005. Reference to persons and objects in the function
-of subject in learner varieties. The structure of learner varieties
-1964.
-
-\leavevmode\vadjust pre{\hypertarget{ref-aksu-koc2015}{}}%
-Aksu-Koç, Ayhan, Nicolopoulou, Ageliki, 2015. Character reference in
-young children's narratives: A crosslinguistic comparison of English,
-Greek, and Turkish. Lingua, Reference and referentiality in native and
-learner grammars 155, 62--84.
-doi:\href{https://doi.org/10.1016/j.lingua.2014.04.006}{10.1016/j.lingua.2014.04.006}
-
-\leavevmode\vadjust pre{\hypertarget{ref-ariel1990}{}}%
-Ariel, Mira, 1990. Accessing noun-phrase antecedents. Routledge.
-
-\leavevmode\vadjust pre{\hypertarget{ref-azar2015}{}}%
-Azar, Zeynep, Özyürek, Aslı, 2015. Discourse management: Reference
-tracking in speech and gesture in Turkish narratives. Dutch Journal of
-Applied Linguistics 4, 222--240.
-doi:\href{https://doi.org/10.1075/dujal.4.2.06aza}{10.1075/dujal.4.2.06aza}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bamberg1997}{}}%
-Bamberg, Michael GW, 1997. Positioning between structure and
-performance. Journal of narrative and life history 7, 335342.
-
-\leavevmode\vadjust pre{\hypertarget{ref-barbera2009}{}}%
-Barberà, Gemma, Massó, Guillem, 2009. Keeping track in sign language:
-Coreference annotation using centering theory, in: Talk Presented at the
-Sign Language Corpora: Linguistic Issues Workshop, University College
-London.
-
-\leavevmode\vadjust pre{\hypertarget{ref-becker2009}{}}%
-Becker, Claudia, 2009. Narrative competences of Deaf children in German
-Sign Language. Sign Language \& Linguistics 12, 113--160.
-doi:\href{https://doi.org/10.1075/sll.12.2.02bec}{10.1075/sll.12.2.02bec}
-
-\leavevmode\vadjust pre{\hypertarget{ref-bel2015}{}}%
-Bel, Aurora, Ortells, Marta, Morgan, Gary, 2015. Reference control in
-the narratives of adult sign language learners. International Journal of
-Bilingualism 19, 608--624.
-doi:\href{https://doi.org/10.1177/1367006914527186}{10.1177/1367006914527186}
-
-\leavevmode\vadjust pre{\hypertarget{ref-benedicto2004}{}}%
-Benedicto, Elena, Brentari, Diane, 2004. Where did all the arguments
-go?: Argument-changing properties of classifiers in ASL. Natural
-Language \& Linguistic Theory 22, 743810.
-
-\leavevmode\vadjust pre{\hypertarget{ref-berman2013relating}{}}%
-Berman, Ruth A, Slobin, Dan Isaac, 2013. Relating events in narrative: A
-crosslinguistic developmental study. Psychology Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-bosch2007}{}}%
-Bosch, Peter, Umbach, Carla, 2007. Reference determination for
-demonstrative pronouns. ZAS Papers in Linguistics 48, 39--51.
-doi:\href{https://doi.org/10.21248/zaspil.48.2007.353}{10.21248/zaspil.48.2007.353}
-
-\leavevmode\vadjust pre{\hypertarget{ref-boudreault2006}{}}%
-Boudreault, Patrick, Mayberry, Rachel I., 2006. Grammatical processing
-in american sign language: Age of first-language acquisition effects in
-relation to syntactic structure. Language and Cognitive Processes 21,
-608--635.
-doi:\href{https://doi.org/10.1080/01690960500139363}{10.1080/01690960500139363}
-
-\leavevmode\vadjust pre{\hypertarget{ref-burkner2018}{}}%
-Bürkner, Paul-Christian, 2018. Advanced bayesian multilevel modeling
-with the R package brms. The R Journal 10, 395.
-doi:\href{https://doi.org/10.32614/RJ-2018-017}{10.32614/RJ-2018-017}
-
-\leavevmode\vadjust pre{\hypertarget{ref-chafe1976}{}}%
-Chafe, W., 1976.
-\href{https:///paper/Givenness\%2C-contrastiveness\%2C-definiteness\%2C-subjects\%2C-Chafe/805a12a9bee00562f6c2663722e9f7f3f5991848}{Givenness,
-contrastiveness, definiteness, subjects, topics, and point of view}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-chamberlain_language_1999}{}}%
-Chamberlain, Charlene, Morford, Jill P., Mayberry, Rachel I. (Eds.),
-1999. Language acquisition by eye. Psychology Press, Mahwah, N.J.
-
-\leavevmode\vadjust pre{\hypertarget{ref-cheng2021}{}}%
-Cheng, Qi, Mayberry, Rachel I., 2021. When event knowledge overrides
-word order in sentence comprehension: Learning a first language after
-childhood. Developmental Science.
-doi:\href{https://doi.org/10.1111/desc.13073}{10.1111/desc.13073}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cormier2012}{}}%
-Cormier, Kearsy, Schembri, Adam, Vinson, David, Orfanidou, Eleni, 2012.
-First language acquisition differs from second language acquisition in
-prelingually deaf signers: Evidence from sensitivity to grammaticality
-judgement in British Sign Language. Cognition 124, 50--65.
-doi:\href{https://doi.org/10.1016/j.cognition.2012.04.003}{10.1016/j.cognition.2012.04.003}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cormier_rethinking_2015}{}}%
-Cormier, Kearsy, Smith, Sandra, Sevcikova-Sehyr, Zed, 2015. Rethinking
-constructed action. {SL}\&L 18, 167--204.
-doi:\href{https://doi.org/10.1075/sll.18.2.01cor}{10.1075/sll.18.2.01cor}
-
-\leavevmode\vadjust pre{\hypertarget{ref-cormier2013}{}}%
-Cormier, Kearsy, Smith, Sandra, Zwets, Martine, 2013. Framing
-constructed action in British Sign Language narratives. Journal of
-Pragmatics 55, 119--139.
-doi:\href{https://doi.org/10.1016/j.pragma.2013.06.002}{10.1016/j.pragma.2013.06.002}
-
-\leavevmode\vadjust pre{\hypertarget{ref-crasborn2008}{}}%
-Crasborn, O., Sloetjes, H., 2008.
-\href{https://www.semanticscholar.org/paper/Enhanced-ELAN-functionality-for-sign-language-Crasborn-Sloetjes/e1623ec6b2bc3532129a3682eb263cc2e7d5389c}{Enhanced
-ELAN functionality for sign language corpora}. undefined.
-
-\leavevmode\vadjust pre{\hypertarget{ref-czubek2017}{}}%
-Czubek, Todd Alan, 2017. A comprehensive study of referring expressions
-in ASL 244.
-
-\leavevmode\vadjust pre{\hypertarget{ref-debreslioska2013}{}}%
-Debreslioska, Sandra, Özyürek, Asli, Gullberg, Marianne, Perniss,
-Pamela, 2013. Gestural viewpoint signals referent accessibility.
-Discourse Processes 50, 431--456.
-doi:\href{https://doi.org/10.1080/0163853X.2013.824286}{10.1080/0163853X.2013.824286}
-
-\leavevmode\vadjust pre{\hypertarget{ref-dikyuva2015}{}}%
-Dikyuva, Hasan, Makaroğlu, Bahtiyar, Arık, Engin, 2015. Türk i{ş}aret
-dili dilbilgisi kitab{ı}. Aile ve Sosyal Politikalar Bakanl{ı}{ğ}{ı}
-Yay{ı}nlar{ı}: Ankara.
-
-\leavevmode\vadjust pre{\hypertarget{ref-dudis2004}{}}%
-Dudis, Paul G., 2004.
-\href{https://www.degruyter.com/view/j/cogl.2004.15.issue-2/cogl.2004.009/cogl.2004.009.xml}{Body
-partitioning and real-space blends}. Cognitive Linguistics 15.
-
-\leavevmode\vadjust pre{\hypertarget{ref-emmorey1996}{}}%
-Emmorey, Karen, 1996. The confluence of space and language in signed
-languages, in: Language, Speech, and Communication. The MIT Press,
-Cambridge, MA, US, pp. 171--209.
-
-\leavevmode\vadjust pre{\hypertarget{ref-engberg-pedersen1993}{}}%
-Engberg-Pedersen, Elisabeth, 1993. Space in danish sign language: The
-semantics and morphosyntax of the use of space in a visual language.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ferrara2022}{}}%
-Ferrara, Lindsay, Anible, Benjamin, Hodge, Gabrielle, Jantunen, Tommi,
-Leeson, Lorraine, Mesch, Johanna, Nilsson, Anna-Lena, 2022. A
-cross-linguistic comparison of reference across five signed languages.
-Linguistic Typology.
-
-\leavevmode\vadjust pre{\hypertarget{ref-frederiksen2016}{}}%
-Frederiksen, Anne Therese, Mayberry, Rachel I., 2016. Who's on first?
-Investigating the referential hierarchy in simple native ASL narratives.
-Lingua 180, 49--68.
-doi:\href{https://doi.org/10.1016/j.lingua.2016.03.007}{10.1016/j.lingua.2016.03.007}
-
-\leavevmode\vadjust pre{\hypertarget{ref-frederiksen2017}{}}%
-Frederiksen, Anne Therese, Mayberry, Rachel I., 2017. Pronominalizing
-subjects: Referentially empty pronouns in ASL.
-
-\leavevmode\vadjust pre{\hypertarget{ref-frederiksen2019}{}}%
-Frederiksen, Anne Therese, Mayberry, Rachel I., 2019. Reference tracking
-in early stages of different modality L2 acquisition: Limited
-over-explicitness in novice ASL signers{'} referring expressions. Second
-Language Research 35, 253--283.
-doi:\href{https://doi.org/10.1177/0267658317750220}{10.1177/0267658317750220}
-
-\leavevmode\vadjust pre{\hypertarget{ref-garcia_reference_2014}{}}%
-Garcia, Brigitte, Sallandre, Marie-Anne, 2014. Reference resolution in
-french sign language ({LSF}), in: Cabredo Hofherr, Patricia,
-Zribi-Hertz, Anne (Eds.), Crosslinguistic Studies on Noun Phrase
-Structure and Reference. Brill, Leiden, pp. 316--364.
-doi:\href{https://doi.org/10.1163/9789004261440_012}{10.1163/9789004261440\_012}
-
-\leavevmode\vadjust pre{\hypertarget{ref-garcia2018impersonal}{}}%
-Garcia, Brigitte, Sallandre, Marie-Anne, L'Huillier, Marie-Thérèse,
-2018. Impersonal human reference in french sign language (LSF). Sign
-Language \& Linguistics 21, 307--333.
-
-\leavevmode\vadjust pre{\hypertarget{ref-givon1983}{}}%
-Givon, T., 1983. \href{https://benjamins.com/catalog/tsl.3}{Topic
-continuity in discourse}. John Benjamins Publishing Company.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gokgoz2020}{}}%
-Gökgöz, Kadir, Sevgi, Hande, 2020. Aspects of clause structure and
-morphology in turkish sign language. Morphological Complexity within and
-across Boundaries: In Honour of Asli Göksel 215, 315.
-
-\leavevmode\vadjust pre{\hypertarget{ref-grice1975}{}}%
-Grice, Herbert P., 1975. Logic and conversation. Brill, p. 4158.
-
-\leavevmode\vadjust pre{\hypertarget{ref-gullberg2006}{}}%
-Gullberg, Marianne, 2006. Handling discourse: Gestures, reference
-tracking, and communication strategies in early L2: Language Learning
-Vol. 56, No. 1. Language Learning 56, 155--196.
-doi:\href{https://doi.org/10.1111/j.0023-8333.2006.00344.x}{10.1111/j.0023-8333.2006.00344.x}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gundel1993}{}}%
-Gundel, Jeanette K., Hedberg, Nancy, Zacharski, Ron, 1993. Cognitive
-status and the form of referring expressions in discourse. Language 69,
-274--307. doi:\href{https://doi.org/10.2307/416535}{10.2307/416535}
-
-\leavevmode\vadjust pre{\hypertarget{ref-gur2018}{}}%
-Gür, Cansu, 2018. The effects of late language acquisition on the
-development of narrative skills: Insights from Turkish Sign Language
-(PhD thesis). Erzurum.
-
-\leavevmode\vadjust pre{\hypertarget{ref-hickmann1996}{}}%
-Hickmann, Maya, Hendriks, Henriëtte, Roland, Françoise, Liang, James,
-1996. The marking of new information in children's narratives: a
-comparison of English, French, German and Mandarin Chinese*. Journal of
-Child Language 23, 591--619.
-doi:\href{https://doi.org/10.1017/S0305000900008965}{10.1017/S0305000900008965}
-
-\leavevmode\vadjust pre{\hypertarget{ref-hodge2019}{}}%
-Hodge, Gabrielle, Ferrara, Lindsay N., Anible, Benjamin D., 2019. The
-semiotic diversity of doing reference in a deaf signed language. Journal
-of Pragmatics 143, 3353.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ilkbasaran2015}{}}%
-Ilkbasaran, Deniz, 2015.
-\href{https://www.proquest.com/docview/1656169535/abstract/16EA71911A644FC3PQ/1}{Literacies,
-mobilities and agencies of deaf youth in turkey: Constraints and
-opportunities in the 21st century} (PhD thesis). Ann Arbor, United
-States.
-
-\leavevmode\vadjust pre{\hypertarget{ref-karadoller2017}{}}%
-Karadöller, Dilay Z, Sümer, Beyza, Özyürek, Aslı, 2017. Effects of
-delayed language exposure on spatial language acquisition by signing
-children and adults 5.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kayaba2022}{}}%
-Kayabaşı, Demet, Gökgöz, Kadir, 2022. The causative-inchoative
-alternation and age-of-acquisition effects on multi-predicate
-constructions in turkish sign language. Language Learning and
-Development 0, 1--26.
-doi:\href{https://doi.org/10.1080/15475441.2022.2047690}{10.1080/15475441.2022.2047690}
-
-\leavevmode\vadjust pre{\hypertarget{ref-kemaloglu2012}{}}%
-Kemaloğlu, Yusuf Kemal, Kemaloğlu, Pınar Yaprak, 2012. The history of
-sign language and deaf education in Turkey. Kulak burun bogaz ihtisas
-dergisi: KBB = Journal of ear, nose, and throat 22, 65--76.
-doi:\href{https://doi.org/10.5606/kbbihtisas.2012.013}{10.5606/kbbihtisas.2012.013}
-
-\leavevmode\vadjust pre{\hypertarget{ref-koulidobrova2016}{}}%
-Koulidobrova, Elena, Lillo-Martin, Diane, 2016. A {`}point{'}of inquiry:
-The case of the (non-) pronominal IX in ASL. The impact of pronominal
-form on interpretation 221, 250.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kruschke2011}{}}%
-Kruschke, John K., 2011. Bayesian assessment of null values via
-parameter estimation and model comparison. Perspectives on Psychological
-Science 6, 299312.
-
-\leavevmode\vadjust pre{\hypertarget{ref-kubus2008}{}}%
-Kubuş, Okan, 2008. An analysis of Turkish Sign Language (T{İ}D)
-phonology and morphology (PhD thesis). Ankara.
-
-\leavevmode\vadjust pre{\hypertarget{ref-liddell2003}{}}%
-Liddell, Scott K., 2003. Grammar, gesture, and meaning in American Sign
-Language. Cambridge University Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-loew1984}{}}%
-Loew, Ruth Carolyn, 1984.
-\href{https:///paper/Roles-and-reference-in-American-sign-language-\%3A-a-Loew/315c2c539d3bd88fbc13564775141ba4a60ac5aa}{Roles
-and reference in American sign language : a developmental perspective}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-lourenuxe7o2018}{}}%
-Lourenço, Guilherme, Wilbur, Ronnie B., 2018. Are plain verbs really
-plain?: Co-localization as the agreement marker in sign languages.
-FEAST. Formal and Experimental Advances in Sign language Theory 2, 6881.
-
-\leavevmode\vadjust pre{\hypertarget{ref-maclaughlin1997}{}}%
-Maclaughlin, Dawn, 1997.
-\href{http://lear.unive.it/jspui/handle/11707/7016}{The structure of
-determiner phrases: Evidence from American Sign Language.} (PhD thesis).
-
-\leavevmode\vadjust pre{\hypertarget{ref-meier1990}{}}%
-Meier, Richard P., Newport, Elissa L., 1990. Out of the hands of babes:
-On a possible sign advantage in language acquisition. Language 66, 1.
-doi:\href{https://doi.org/10.2307/415277}{10.2307/415277}
-
-\leavevmode\vadjust pre{\hypertarget{ref-meir1999}{}}%
-Meir, Irit, 1999. A perfect marker in israeli sign language. Sign
-Language \& Linguistics 2, 4362.
-
-\leavevmode\vadjust pre{\hypertarget{ref-metzger1995}{}}%
-Metzger, Melanie, 1995. Constructed action and constructed dialogue in
-american sign language. Sociolinguistics in Deaf communities 255271.
-
-\leavevmode\vadjust pre{\hypertarget{ref-ministryofnationaleducation2020}{}}%
-Ministry of National Education, 2020.
-\href{http://sgb.meb.gov.tr/meb_iys_dosyalar/2020_09/04144812_meb_istatistikleri_orgun_\%20egitim_2019_2020.pdf}{Milli
-e{ğ}itim istatistikleri, örgün e{ğ}itim 2019/{'}20 {[}national education
-statistics, formal education 2019/{'}20{]}}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-morgan2005}{}}%
-Morgan, Gary, 2005.
-\href{http://www.oxfordscholarship.com/view/10.1093/acprof:oso/9780195180947.001.0001/acprof-9780195180947}{The
-development of narrative skills in British Sign Language}, in: Schick,
-Brenda, Marschark, Marc, Spencer, Patricia Elizabeth (Eds.),. Oxford
-University Press.
-
-\leavevmode\vadjust pre{\hypertarget{ref-newport1990}{}}%
-Newport, Elissa L., 1990. Maturational constraints on language learning.
-Cognitive Science 14, 11--28.
-doi:\href{https://doi.org/10.1207/s15516709cog1401_2}{10.1207/s15516709cog1401\_2}
-
-\leavevmode\vadjust pre{\hypertarget{ref-nuhbalaoglu2018}{}}%
-Nuhbalaoglu, Derya, 2018. Comprehension and production of referential
-expressions in German Sign Language and Turkish Sign Language: An
-empirical approach (PhD thesis). Germany.
-
-\leavevmode\vadjust pre{\hypertarget{ref-padden1986}{}}%
-Padden, Carol, 1986. Verbs and role-shifting in american sign language.
-National Association of the Deaf Silver Spring, MD, p. 57.
-
-\leavevmode\vadjust pre{\hypertarget{ref-perniss2015}{}}%
-Perniss, Pamela, Özyürek, Asli, 2015. Visible Cohesion: A comparison of
-reference tracking in sign, speech, and co-speech gesture. Topics in
-Cognitive Science 7, 36--60.
-doi:\href{https://doi.org/10.1111/tops.12122}{10.1111/tops.12122}
-
-\leavevmode\vadjust pre{\hypertarget{ref-prince1981}{}}%
-Prince, EF, 1981. Toward a taxonomy of given - new information, in:
-Cole, Peter (Ed.),. Academic Press, New York, pp. 223--255.
-
-\leavevmode\vadjust pre{\hypertarget{ref-rathmann2007}{}}%
-Rathmann, Christian, Mann, Wolfgang, Morgan, Gary, 2007. Narrative
-structure and narrative development in deaf children. Deafness \&
-Education International 9, 187--196.
-doi:\href{https://doi.org/10.1179/146431507790559932}{10.1179/146431507790559932}
-
-\leavevmode\vadjust pre{\hypertarget{ref-sari_analysis_2005}{}}%
-Sari, Hakan, 2005. An analysis of the relationship between identity
-patterns of turkish deaf adolescents and the communication modes used in
-special residential schools for the hearing impaired and deaf. Deafness
-\& Education International 7, 206--222.
-doi:\href{https://doi.org/10.1179/146431505790560329}{10.1179/146431505790560329}
-
-\leavevmode\vadjust pre{\hypertarget{ref-sevinc2006}{}}%
-Sevinç, Ayça Müge, 2006. Grammatical relations and word order in turkish
-sign language (t{İ}d). Unpublished master's thesis, Middle East
-Technical University, Turkey.
-
-\leavevmode\vadjust pre{\hypertarget{ref-smith2014}{}}%
-Smith, Sandra, Cormier, Kearsy, 2014. In or out?: Spatial scale and
-enactment in narratives of native and nonnative signing deaf children
-acquiring British Sign Language. Sign Language Studies 14, 275--301.
-doi:\href{https://doi.org/10.1353/sls.2014.0008}{10.1353/sls.2014.0008}
-
-\leavevmode\vadjust pre{\hypertarget{ref-supalla1990}{}}%
-Supalla, Ted, 1990. Serial verbs of motion in ASL. Theoretical issues in
-sign language research 1, 127152.
-
-\leavevmode\vadjust pre{\hypertarget{ref-swabey2002}{}}%
-Swabey, Laurie Anne, 2002. The cognitive status, form and distribution
-of referring expressions in ASL and English narratives (PhD thesis).
-Minnepolis, USA.
-
-\leavevmode\vadjust pre{\hypertarget{ref-talmy2003}{}}%
-Talmy, Leonard, 2003. The representation of spatial structure in spoken
-and signed language. Perspectives on classifier constructions in sign
-language 169195.
-
-\leavevmode\vadjust pre{\hypertarget{ref-toole1996}{}}%
-Toole, Janine, 1996. The effect of genre on referential choice. p. 263.
-doi:\href{https://doi.org/10.1075/pbns.38.16too}{10.1075/pbns.38.16too}
-
-\leavevmode\vadjust pre{\hypertarget{ref-wagenmakers2007}{}}%
-Wagenmakers, Eric-Jan, 2007. A practical solution to the pervasive
-problems of p values. Psychonomic bulletin \& review 14, 779804.
-doi:\href{https://doi.org/10.3758/bf03194105}{10.3758/bf03194105}
-
-\leavevmode\vadjust pre{\hypertarget{ref-williams1988}{}}%
-Williams, Jessica, 1988. Zero anaphora in second language acquisition: A
-comparison among three varieties of English. Studies in Second Language
-Acquisition 10, 339--370.
-doi:\href{https://doi.org/10.1017/S0272263100007488}{10.1017/S0272263100007488}
-
-\leavevmode\vadjust pre{\hypertarget{ref-woll2013}{}}%
-Woll, Bencie, 2013. Sign language and spoken language development in
-young children: Measuring vocabulary by means of the CDI, in: Meurant,
-Laurence, Sinte, Aurélie, Van Herreweghe, Mieke, Vermeerbergen, Myriam
-(Eds.),. De Gruyter, Berlin, Boston, pp. 15--34.
-doi:\href{https://doi.org/10.1515/9781614511472.15}{10.1515/9781614511472.15}
-
-\end{CSLReferences}
+\renewcommand\refname{References}
+\bibliography{./bibliography/references.bib}
 

### 2023-05-04 revised_manuscript.Rmd
update

@@
 colnames(df_model_1)[1:4] <- c("Coefficients","outerWidth","innerWidth", "pointEstimate")
-df_model_1 %<>% mutate(across(.cols = "Coefficients", ~str_replace( ., "_","\\\\_")))
+df_model_1 %<>% dplyr::select(-model) %>% mutate(across(.cols = "Coefficients", ~str_replace( ., "_","\\\\_")))
 apa_table(df_model_1,

### 2023-05-04 revised_manuscript.tex
update

--- 2023-05-01 revised_manuscript.tex
@@
 children prioritize spoken language development and discourage the use
-of a sign language (Kemaloğlu and Kemaloğlu, 2012). The Ministry of
-National Education (2020) identifies 72 deaf schools in Turkey for the
-K-12 level. Even when deaf children with hearing parents start their
-formal education in one of these schools, they do not receive proper
+of a sign language (Kemaloğlu and Kemaloğlu, 2012). The National
+Education (2020) identifies 72 deaf schools in Turkey for the K-12
+level. Even when deaf children with hearing parents start their formal
+education in one of these schools, they do not receive proper
 instruction in TİD since the teachers themselves are mainly hearing and
@@
 
+\hypertarget{regression-model-results-for-re-type}{%
+\section{Regression model results for RE
+Type}\label{regression-model-results-for-re-type}}
+
+\label{section:modelResult}
+
+\begin{table}[hbt!]
+
+\begin{center}
+\begin{threeparttable}
+
+\caption{\label{tab:unnamed-chunk-4}Regression model results for the RE Type (Nominal, Zero, Classifier) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (DoD, DoH)}
+
+\small{
+
+\begin{tabular}{lllllllll}
+\toprule
+Coefficients & \multicolumn{1}{c}{outerWidth} & \multicolumn{1}{c}{innerWidth} & \multicolumn{1}{c}{pointEstimate} & \multicolumn{1}{c}{ll} & \multicolumn{1}{c}{l} & \multicolumn{1}{c}{m} & \multicolumn{1}{c}{h} & \multicolumn{1}{c}{hh}\\
+\midrule
+Nom\_Intercept & 0.90 & 0.50 & median & 0.86 & 1.06 & 1.18 & 1.31 & 1.50\\
+Zero\_Intercept & 0.90 & 0.50 & median & 0.11 & 0.26 & 0.37 & 0.46 & 0.59\\
+Nom\_DoD & 0.90 & 0.50 & median & -0.67 & -0.31 & -0.06 & 0.18 & 0.55\\
+Nom\_Intro & 0.90 & 0.50 & median & 3.34 & 3.56 & 3.72 & 3.88 & 4.13\\
+Nom\_Maintain & 0.90 & 0.50 & median & -5.70 & -5.49 & -5.34 & -5.20 & -4.98\\
+Nom\_DoD*Intro & 0.90 & 0.50 & median & -0.09 & 0.34 & 0.65 & 0.98 & 1.44\\
+Nom\_DoD*Maintain & 0.90 & 0.50 & median & -1.86 & -1.45 & -1.16 & -0.88 & -0.46\\
+Zero\_DoD & 0.90 & 0.50 & median & 0.45 & 0.72 & 0.89 & 1.10 & 1.40\\
+Zero\_Intro & 0.90 & 0.50 & median & -4.77 & -4.22 & -3.87 & -3.55 & -3.13\\
+Zero\_Maintain & 0.90 & 0.50 & median & 1.75 & 1.99 & 2.17 & 2.37 & 2.66\\
+Zero\_DoD*Intro & 0.90 & 0.50 & median & 0.10 & 0.93 & 1.58 & 2.26 & 3.32\\
+Zero\_DoD*Maintain & 0.90 & 0.50 & median & -1.95 & -1.38 & -0.99 & -0.64 & -0.17\\
+\bottomrule
+\end{tabular}
+
+}
+
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\clearpage
+
 \hypertarget{additional-model-with-3-levels-for-nativeness}{%
@@
 \leavevmode\vadjust pre{\hypertarget{ref-chamberlain_language_1999}{}}%
-Chamberlain, Charlene, Morford, Jill P., Mayberry, Rachel I. (Eds.),
-1999. Language acquisition by eye. Psychology Press, Mahwah, N.J.
+Chamberlain, Charlene, Morford, Jill P, Mayberry, Rachel I, 1999.
+Language acquisition by eye. Psychology Press, Mahwah, N.J.
 
@@
 Czubek, Todd Alan, 2017. A comprehensive study of referring expressions
-in ASL 244.
+in ASL (PhD thesis).
 
@@
 semantics and morphosyntax of the use of space in a visual language.
+Gallaudet University Press.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-ministryofnationaleducation2020}{}}%
-Ministry of National Education, 2020.
-\href{http://sgb.meb.gov.tr/meb_iys_dosyalar/2020_09/04144812_meb_istatistikleri_orgun_\%20egitim_2019_2020.pdf}{Milli
-e{ğ}itim istatistikleri, örgün e{ğ}itim 2019/{'}20 {[}national education
-statistics, formal education 2019/{'}20{]}}.
-
 \leavevmode\vadjust pre{\hypertarget{ref-morgan2005}{}}%
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-ministryofnationaleducation2020}{}}%
+National Education, Ministry of, 2020.
+\href{http://sgb.meb.gov.tr/meb_iys_dosyalar/2020_09/04144812_meb_istatistikleri_orgun_\%20egitim_2019_2020.pdf}{Milli
+e{ğ}itim istatistikleri, örgün e{ğ}itim 2019/{'}20 {[}national education
+statistics, formal education 2019/{'}20{]}}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-newport1990}{}}%

### 2023-05-04 revised_manuscript.Rmd
updated

@@
 Fingerspelling refers to the use of the manual alphabet to spell all or part of the letters of a word borrowed from a surrounding written language [@kubus2008].
-Nouns can also be modified by a prenominal or postnominal extended index finger pointing sign, glossed as IX, functioning as a determiner or a demonstrative as in \ref{ex:2c} [@maclaughlin1997] or by a modifying classifier within the same NP as in \ref{ex:2d} [@frederiksen2016].\footnote{Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions (Lillo-Martin and Klima, 1990).
-In addition, the pointing sign is used to referentially anchor a referent to a location in signing space.}
+Nouns can also be modified by a prenominal or postnominal extended index finger pointing sign, glossed as IX, functioning as a determiner or a demonstrative as in \ref{ex:2c} [@maclaughlin1997] or by a modifying classifier within the same NP as in \ref{ex:2d} [@frederiksen2016].[^1]
+
+[^1]: Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions [@lillo-martin1990].
+In addition, the pointing sign is used to referentially anchor a referent to a location in signing space.
 
 ```{=tex}
-\newpage
 \begin{exe}
@@
 For classifiers, we distinguished among three types of handshapes: WECL, BPCL, and ExtCL.
-Finally, we had CA[^1], agreement, plain tags for zero anaphora.
+Finally, we had CA[^2], agreement, plain tags for zero anaphora.
 We were unable to observe any instance of HCLs occurring in isolation (i.e., not part of CA).
@@
 
-[^1]: Note that CA can also be further subdivided into overt, reduced, and subtle CA, following [@cormier_rethinking_2015].
+[^2]: Note that CA can also be further subdivided into overt, reduced, and subtle CA, following [@cormier_rethinking_2015].
 In overt CA, the signer is "fully in character" and does not use any other lexical or classifier constructions.
@@
 In re-introduction contexts, signers used a greater proportion of nominal markings which were followed by zero anaphora, classifier, and pronominal use.
+
+Table \ref{tab:my-table2} shows that for all discourse statuses the most numerous nominal type was bare nominal which was followed by nouns modified by index or classifiers.
+In number, signers used modified nouns more in introduction and re-introduction compared to maintenance settings.
+Signers produced a total number of `r sum(df_narration$ref_type == "Pronominal")` pronouns (Table \ref{tab:my-table3}).
+A great majority of the pronominal constructions were used to re-introduce a referent into the discourse.
+We observed only a very small amount of pronoun use for referent introduction and maintenance.
 
@@
 \end{table}
-```
-Table \ref{tab:my-table2} shows that for all discourse statuses the most numerous nominal type was bare nominal which was followed by nouns modified by index or classifiers.
-In number, signers used modified nouns more in introduction and re-introduction compared to maintenance settings.
-Signers produced a total number of `r sum(df_narration$ref_type == "Pronominal")` pronouns (Table \ref{tab:my-table3}).
-A great majority of the pronominal constructions were used to re-introduce a referent into the discourse.
-We observed only a very small amount of pronoun use for referent introduction and maintenance.
-
-```{=tex}
+
 \begin{table}
@@
 ```
-As for the classifier constructions (Table \ref{tab:my-table4}), WECL was the most common type, followed by BPCL and ExtCL.
-For the first mention of a referent, signers mostly preferred to use ExtCL.
-However, for referent maintenance and re-introduction WECL was used more compared to BPCL and ExtCL.
-
-Among the markers of zero anaphora, CA was most frequently employed, and it was followed by plain verbs and agreement verbs (Table \ref{tab:my-table5}).
-Overall, signers used very few markers of zero anaphora for the introduction of a referent and the majority of the existing ones were CA.
-Additionally, signers used CA more for referent re-introduction than maintenance.
-We observed similar instances of agreement verbs for maintained and re-introduced contexts.
-In contrast, plain verbs were favored more in the maintenance setting than re-introduction.
-
 ```{=tex}
@@
 ```
-Across the discourse statuses, there is also a difference between participant groups in the use of zero anaphora for reference tracking.
-DoD signers of TİD used it more frequently (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "DoD") / nrow(df_narration), digits=2)`) compared to the DoH signers (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "DoH") / nrow(df_narration), digits=2)`).
-Comparably, DoH signers used nominals more frequently (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "DoH") / nrow(df_narration), digits=2)`) than DoD signers (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "DoD") / nrow(df_narration), digits=2)`).
-Figure 1 shows the numerical distribution of the responses by RE type and Group.
-
 ```{=tex}
@@
 
+Across the discourse statuses, there is also a difference between participant groups in the use of zero anaphora for reference tracking.
+DoD signers of TİD used it more frequently (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "DoD") / nrow(df_narration), digits=2)`) compared to the DoH signers (`r round(sum(df_narration$GTR1_CAT == "Zero Anaphor" & df_narration$Nativeness == "DoH") / nrow(df_narration), digits=2)`).
+Comparably, DoH signers used nominals more frequently (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "DoH") / nrow(df_narration), digits=2)`) than DoD signers (`r round(sum(df_narration$GTR1_CAT == "Nominal" & df_narration$Nativeness == "DoD") / nrow(df_narration), digits=2)`).
+Figure 1 shows the numerical distribution of the responses by RE type and Group.
+
 For more inference on the data and having pairwise comparisons, we fit a regression model to RE Type using discourse status and acquisition group as predictors.
@@
 For the current analysis, we did not define priors ourselves and used the preset uninformed priors.
-Figure 2[^2] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals.
+Figure 2[^3] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals.
 There are 3 responses we are considering in this model.
@@
 
-[^2]: The regression model results are given in \ref{section:modelResult}.
+[^3]: The regression model results are given in \ref{section:modelResult}.
 We ran another model where we used the self-reported acquisition time as a predictor with 3 levels 0-3:Early, 4-7:Late1, and 7-above:Late2.
@@
 
-As for the distribution of the zero anaphora, approximately 90% of the plain verbs were used to maintain a previous referent and this makes these constructions least implicit REs used for TİD[^3].
+As for the distribution of the zero anaphora, approximately 90% of the plain verbs were used to maintain a previous referent and this makes these constructions least implicit REs used for TİD[^4].
 Agreement verbs and CA were the following implicit markers of reference, `r sum(df_narration$GTR1 == "Person agreement verb" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1 == "Person agreement verb")` and `r sum(df_narration$GTR1_plot == "CA" & df_narration$Discourse == "Maintenance") / sum(df_narration$GTR1_plot == "CA")` of which were used for maintained contexts, respectively.
@@
 
-[^3]: As in other sign languages, in TİD, the lexical meaning of signs has an effect on the morphological grouping.
+[^4]: As in other sign languages, in TİD, the lexical meaning of signs has an effect on the morphological grouping.
 For instance, usually (di)transitive agreement verbs denote some sort of physical or metaphorical transfer of an object or abstract notion [@meir1999].

### 2023-05-04 revised_manuscript.tex
updated

@@
 (Maclaughlin, 1997) or by a modifying classifier within the same NP as
-in \ref{ex:2d} (Frederiksen and Mayberry,
-2016).\footnote{Spoken languages use pronouns, determines and demonstratives where sign languages use a pointing sign, usually glossed as IX, which assumes similar functions (Lillo-Martin and Klima, 1990).
-In addition, the pointing sign is used to referentially anchor a referent to a location in signing space.}
-
-\newpage
+in \ref{ex:2d} (Frederiksen and Mayberry, 2016).\footnote{Spoken
+  languages use pronouns, determines and demonstratives where sign
+  languages use a pointing sign, usually glossed as IX, which assumes
+  similar functions (Lillo-Martin and Klima, 1990). In addition, the
+  pointing sign is used to referentially anchor a referent to a location
+  in signing space.}
+
 \begin{exe}
@@
 
+Table \ref{tab:my-table2} shows that for all discourse statuses the most
+numerous nominal type was bare nominal which was followed by nouns
+modified by index or classifiers. In number, signers used modified nouns
+more in introduction and re-introduction compared to maintenance
+settings. Signers produced a total number of 40 pronouns (Table
+\ref{tab:my-table3}). A great majority of the pronominal constructions
+were used to re-introduce a referent into the discourse. We observed
+only a very small amount of pronoun use for referent introduction and
+maintenance.
+
 \begin{table}
@@
 
-Table \ref{tab:my-table2} shows that for all discourse statuses the most
-numerous nominal type was bare nominal which was followed by nouns
-modified by index or classifiers. In number, signers used modified nouns
-more in introduction and re-introduction compared to maintenance
-settings. Signers produced a total number of 40 pronouns (Table
-\ref{tab:my-table3}). A great majority of the pronominal constructions
-were used to re-introduce a referent into the discourse. We observed
-only a very small amount of pronoun use for referent introduction and
-maintenance.
-
 \begin{table}
@@
 
-As for the classifier constructions (Table \ref{tab:my-table4}), WECL
-was the most common type, followed by BPCL and ExtCL. For the first
-mention of a referent, signers mostly preferred to use ExtCL. However,
-for referent maintenance and re-introduction WECL was used more compared
-to BPCL and ExtCL.
-
-Among the markers of zero anaphora, CA was most frequently employed, and
-it was followed by plain verbs and agreement verbs (Table
-\ref{tab:my-table5}). Overall, signers used very few markers of zero
-anaphora for the introduction of a referent and the majority of the
-existing ones were CA. Additionally, signers used CA more for referent
-re-introduction than maintenance. We observed similar instances of
-agreement verbs for maintained and re-introduced contexts. In contrast,
-plain verbs were favored more in the maintenance setting than
-re-introduction.
-
 \begin{table}
@@
 
-Across the discourse statuses, there is also a difference between
-participant groups in the use of zero anaphora for reference tracking.
-DoD signers of TİD used it more frequently (0.22) compared to the DoH
-signers (0.17). Comparably, DoH signers used nominals more frequently
-(0.26) than DoD signers (0.22). Figure 1 shows the numerical
-distribution of the responses by RE type and Group.
-
 \begin{table}
@@
 \end{figure}
+
+Across the discourse statuses, there is also a difference between
+participant groups in the use of zero anaphora for reference tracking.
+DoD signers of TİD used it more frequently (0.22) compared to the DoH
+signers (0.17). Comparably, DoH signers used nominals more frequently
+(0.26) than DoD signers (0.22). Figure 1 shows the numerical
+distribution of the responses by RE type and Group.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-lillo-martin1990}{}}%
+Lillo-Martin, Diane, Klima, Edward S., 1990. Pointing out differences:
+ASL pronouns in syntactic theory. Theoretical issues in sign language
+research, Volume 1: Linguistics 191.
+
 \leavevmode\vadjust pre{\hypertarget{ref-loew1984}{}}%

### 2023-05-04 revised_manuscript.Rmd
updated

@@
 
-In our first set of hypotheses regarding referent tracking strategies, we had the general expectation that signers would use more explicit forms (mainly nominals) for referent introduction and less implicit forms such as zero anaphora, pronominals, and classifiers for maintenance.
+Regarding referent tracking strategies, we had the general expectation that signers would use more explicit forms (mainly nominals) for referent introduction and less implicit forms such as zero anaphora, pronominals, and classifiers for maintenance.
 In re-introduced contexts, we predicted that signers would use both explicit and implicit markers depending on the accessibility of the tracked referent.
 
-For age of acquisition effects, based on what is proposed for child and adult deaf signers with delayed language exposure in terms of their morphosyntactic [@boudreault2006; @cormier2012; @kayaba2022] and narrative development [@becker2009; @cormier2013], we expected that DoH signers would be more over-redundant in their referent tracking, perhaps akin to second language learners [@frederiksen2019; @bel2015].
+Especially, for age of acquisition effects, based on what is proposed for child and adult deaf signers with delayed language exposure in terms of their morphosyntactic [@boudreault2006; @cormier2012; @kayaba2022] and narrative development [@becker2009; @cormier2013], we expected that DoH signers would be more over-redundant in their referent tracking, perhaps akin to second language learners [@frederiksen2019; @bel2015].
 That is, we had the expectation that DoH signers would use more explicit markers like nouns and fewer implicit markers like zero anaphora even for highly accessible referents (when the discourse status is maintained or even re-introduced in same cases) than DoD signers.
@@
 
-The results of the present study only partly support our hypotheses.
+Mainly, the results of the present study support our initial set of hypotheses.
 In line with our expectations, regardless of the participant groups (DoD or DoH) (i) signers mainly used nominals to introduce a referent into the discourse, (ii) used zero anaphora and classifiers but not pronouns for maintained contexts, and (iii) employed both nominals and less explicit markers (classifiers, zero anaphora and pronouns) to re-introduce a referent.
 
-However, the proportion of the observed discourse statuses (26% for introduction, 40% for maintenance, 34% for re-introduction across acquisition groups) in this study differed from that of some other studies.
+By contrast, the proportion of the observed discourse statuses (26% for introduction, 40% for maintenance, 34% for re-introduction across acquisition groups) in this study differed from that of some other studies.
 For a comparison, while the percentage of introduced referents were similar in both research (26% versus 24%), the participants in the present study brought back referents into the discourse more frequently compared to the participants in @frederiksen2019 and @frederiksen2016 which reported that re-introduction contexts accounted for only 7% of the referents in ASL.
@@
 
-Interestingly and to our surprise, the signers in this study used pronominal IX very sporadically with only 40 observed instances that were mainly used to re-introduce referents rather than maintain them.
+Yet, surprisingly, the signers in this study used pronominal IX very sporadically with only 40 observed instances that were mainly used to re-introduce referents rather than maintain them.
 This finding is comparable with what has been reported for ASL [@swabey2002; @frederiksen2016; @czubek2017] but not with LSC [@bel2015], BSL [@morgan2005] and DGS [@perniss2015], in which the signers extensively made overt pronominal references (up to 30%) to maintain entities.
@@
 
-A deeper look into the allocation of classifier handshapes revealed that WECLs were the most common type of classifiers for the maintenance of a referent followed by BPCLs and ExtCLs.
+Regarding the allocation of classifier handshapes, we found that WECLs were the most common type of classifiers for the maintenance of a referent followed by BPCLs and ExtCLs.
 We observed a similar distibution for re-introduced contexts.
@@
 ```
-Both hierarchies illustrated in Figure \ref{fig-scale} treat nominals and ExtCLs as markers of lowly accessibly referents.
-Likewise, in both hierarchies, as markers of highly accessible referents, different types of zero anaphora markers are further distinguished.
+Keep in mind that both hierarchies illustrated in Figure \ref{fig-scale} treat nominals and ExtCLs as markers of lowly accessibly referents.
+In both hierarchies, as markers of highly accessible referents, different types of zero anaphora markers are further distinguished.
 Plain verbs rank the lowest on the scales, preceded by agreement verbs and CA.
@@
 
-The results in this study only somewhat support our second set of hypotheses that we set for the effects of having signing deaf parents on participants' RE selection.
+Likewise, our results only partly support the second set of hypotheses that we set for the effects of having signing deaf parents on participants' RE selection.
 That is, while being a DoD signer of TİD increased the odds of using zero anaphora in reference tracking across different discourse statuses, we failed to find a difference for overall nominal production.
@@
 
-While DoH signers were only partially over-redundant in maintained contexts, they were more so in introduced contexts.
+It can be observed that while DoH signers were only partially over-redundant in maintained contexts, they were more so in introduced contexts.
 In line with our study, a tendency for limited over-redundancy has been suggested for late signing children in DGS [@becker2009] and late signing adults in BSL specifically for the use of CA [@cormier2013].

### 2023-05-04 revised_manuscript.tex
updated

@@
 
-In our first set of hypotheses regarding referent tracking strategies,
-we had the general expectation that signers would use more explicit
-forms (mainly nominals) for referent introduction and less implicit
-forms such as zero anaphora, pronominals, and classifiers for
-maintenance. In re-introduced contexts, we predicted that signers would
-use both explicit and implicit markers depending on the accessibility of
-the tracked referent.
-
-For age of acquisition effects, based on what is proposed for child and
-adult deaf signers with delayed language exposure in terms of their
-morphosyntactic (Boudreault and Mayberry, 2006; Cormier et al., 2012;
-Kayabaşı and Gökgöz, 2022) and narrative development (Becker, 2009;
-Cormier et al., 2013), we expected that DoH signers would be more
+Regarding referent tracking strategies, we had the general expectation
+that signers would use more explicit forms (mainly nominals) for
+referent introduction and less implicit forms such as zero anaphora,
+pronominals, and classifiers for maintenance. In re-introduced contexts,
+we predicted that signers would use both explicit and implicit markers
+depending on the accessibility of the tracked referent.
+
+Especially, for age of acquisition effects, based on what is proposed
+for child and adult deaf signers with delayed language exposure in terms
+of their morphosyntactic (Boudreault and Mayberry, 2006; Cormier et al.,
+2012; Kayabaşı and Gökgöz, 2022) and narrative development (Becker,
+2009; Cormier et al., 2013), we expected that DoH signers would be more
 over-redundant in their referent tracking, perhaps akin to second
@@
 
-The results of the present study only partly support our hypotheses. In
-line with our expectations, regardless of the participant groups (DoD or
-DoH) (i) signers mainly used nominals to introduce a referent into the
-discourse, (ii) used zero anaphora and classifiers but not pronouns for
-maintained contexts, and (iii) employed both nominals and less explicit
-markers (classifiers, zero anaphora and pronouns) to re-introduce a
-referent.
-
-However, the proportion of the observed discourse statuses (26\% for
+Mainly, the results of the present study support our initial set of
+hypotheses. In line with our expectations, regardless of the participant
+groups (DoD or DoH) (i) signers mainly used nominals to introduce a
+referent into the discourse, (ii) used zero anaphora and classifiers but
+not pronouns for maintained contexts, and (iii) employed both nominals
+and less explicit markers (classifiers, zero anaphora and pronouns) to
+re-introduce a referent.
+
+By contrast, the proportion of the observed discourse statuses (26\% for
 introduction, 40\% for maintenance, 34\% for re-introduction across
@@
 
-Interestingly and to our surprise, the signers in this study used
-pronominal IX very sporadically with only 40 observed instances that
-were mainly used to re-introduce referents rather than maintain them.
-This finding is comparable with what has been reported for ASL (Czubek,
-2017; Frederiksen and Mayberry, 2016; Swabey, 2002) but not with LSC
-(Bel et al., 2015), BSL (Morgan, 2005) and DGS (Perniss and Özyürek,
-2015), in which the signers extensively made overt pronominal references
-(up to 30\%) to maintain entities. Comparable to our findings, it has
-been reported that pronouns do not frequently occur in the local context
-and instead follows a topic shift and thus rank higher than classifiers
-and zero anaphora (see Barberà and Massó, 2009 for LSC; and Nuhbalaoglu,
+Yet, surprisingly, the signers in this study used pronominal IX very
+sporadically with only 40 observed instances that were mainly used to
+re-introduce referents rather than maintain them. This finding is
+comparable with what has been reported for ASL (Czubek, 2017;
+Frederiksen and Mayberry, 2016; Swabey, 2002) but not with LSC (Bel et
+al., 2015), BSL (Morgan, 2005) and DGS (Perniss and Özyürek, 2015), in
+which the signers extensively made overt pronominal references (up to
+30\%) to maintain entities. Comparable to our findings, it has been
+reported that pronouns do not frequently occur in the local context and
+instead follows a topic shift and thus rank higher than classifiers and
+zero anaphora (see Barberà and Massó, 2009 for LSC; and Nuhbalaoglu,
 2018 for TİD). Considering this, the claim that the IX sign functions
@@
 
-A deeper look into the allocation of classifier handshapes revealed that
-WECLs were the most common type of classifiers for the maintenance of a
+Regarding the allocation of classifier handshapes, we found that WECLs
+were the most common type of classifiers for the maintenance of a
 referent followed by BPCLs and ExtCLs. We observed a similar distibution
@@
 
-Both hierarchies illustrated in Figure \ref{fig-scale} treat nominals
-and ExtCLs as markers of lowly accessibly referents. Likewise, in both
-hierarchies, as markers of highly accessible referents, different types
-of zero anaphora markers are further distinguished. Plain verbs rank the
-lowest on the scales, preceded by agreement verbs and CA. On a similar
-ground, both scales lack pronouns, which occurred very infrequently in
-the data. Although Frederiksen and Mayberry (2016) did not include other
-classifier types in the hierarchy, we propose that WECLs and BPCLs must
-rank higher than zero anaphora but lower than ExtCLs that were mainly
-used by interlocutors to introduce a referent into the discourse
-context.
+Keep in mind that both hierarchies illustrated in Figure \ref{fig-scale}
+treat nominals and ExtCLs as markers of lowly accessibly referents. In
+both hierarchies, as markers of highly accessible referents, different
+types of zero anaphora markers are further distinguished. Plain verbs
+rank the lowest on the scales, preceded by agreement verbs and CA. On a
+similar ground, both scales lack pronouns, which occurred very
+infrequently in the data. Although Frederiksen and Mayberry (2016) did
+not include other classifier types in the hierarchy, we propose that
+WECLs and BPCLs must rank higher than zero anaphora but lower than
+ExtCLs that were mainly used by interlocutors to introduce a referent
+into the discourse context.
 
@@
 
-The results in this study only somewhat support our second set of
-hypotheses that we set for the effects of having signing deaf parents on
+Likewise, our results only partly support the second set of hypotheses
+that we set for the effects of having signing deaf parents on
 participants' RE selection. That is, while being a DoD signer of TİD
@@
 
-While DoH signers were only partially over-redundant in maintained
-contexts, they were more so in introduced contexts. In line with our
-study, a tendency for limited over-redundancy has been suggested for
-late signing children in DGS (Becker, 2009) and late signing adults in
-BSL specifically for the use of CA (Cormier et al., 2013). As for TİD,
-to our knowledge, there is only one study (Gür, 2018) that directly
-aimed at examining age of acquisition effects on adult signers'
-narrative skills. Gür (2018) observed that late signing deaf adults and
-children set the scene less frequently and used CA less accurately
-compared to their native counterparts. However, this particular
-investigation did not present any data on the distribution of REs
-between two acquisition groups. In the present study, the finding that
-DoD signers preferred to use null arguments even for the first mention
-of a referent (19 in total) is surprising given some of the previous
-literature in the topicality and saliency framework (Cormier et al.,
-2013; Ferrara et al., 2022; Swabey, 2002). These research reported very
-little or no use of zero anaphora by DoD signers for introduced
+It can be observed that while DoH signers were only partially
+over-redundant in maintained contexts, they were more so in introduced
+contexts. In line with our study, a tendency for limited over-redundancy
+has been suggested for late signing children in DGS (Becker, 2009) and
+late signing adults in BSL specifically for the use of CA (Cormier et
+al., 2013). As for TİD, to our knowledge, there is only one study (Gür,
+2018) that directly aimed at examining age of acquisition effects on
+adult signers' narrative skills. Gür (2018) observed that late signing
+deaf adults and children set the scene less frequently and used CA less
+accurately compared to their native counterparts. However, this
+particular investigation did not present any data on the distribution of
+REs between two acquisition groups. In the present study, the finding
+that DoD signers preferred to use null arguments even for the first
+mention of a referent (19 in total) is surprising given some of the
+previous literature in the topicality and saliency framework (Cormier et
+al., 2013; Ferrara et al., 2022; Swabey, 2002). These research reported
+very little or no use of zero anaphora by DoD signers for introduced
 contexts.
